### 2014-12-03 external/notes/cropland-bias-notes.tex
Added latex document outlining paper approach, with figures. Created new Rmd that describes analysis drawing random sample of kml ids from SA grid over the subsistence farming areas.  Saved this as sqlite to give to Stephanie for mapping. This is divided into two datasets--one has full sample of 200 sites, the other a random draw of 100 from that, which will be the first ones we map.  Started qgis project in which to start mapping this.

\documentclass[11 pt]{article}
\usepackage[left=0.95in,top=0.9in,bottom=0.9in,right=0.95in,nohead,nofoot]{geometry}
\usepackage[hidelinks]{hyperref}  % get rid of boxes around links 
\usepackage[usenames]{color}
\hypersetup{colorlinks=true,citecolor=black,urlcolor=blue}
\usepackage[left,running]{lineno}
% \linenumbers
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\makeatletter
\renewcommand{\@makefnmark}{\hbox{\textsuperscript{\normalsize{\@thefnmark}}}}
\makeatother
\usepackage{graphicx, fullpage}
% \usepackage[parfill]{parskip}
\usepackage{float}
\usepackage{color}

\begin{document}
\pagestyle{empty}

{\centering\section*{How Biases Scale in Landcover Data, With Implications for Area-based Estimates}}

\section*{Overview}
This paper explores the biases inherent in current landcover products' estimates of cropland cover, and the implications that these biases have for research making area-related inferences using these data (e.g. estimates of carbon stocks; crop production and yield; land use change studies). This analysis was made possible by the existence of a high accuracy, hand-digitized cropland data layer available for South Africa, which provides an objective and spatially extensive standard of ``truth'' against which landcover products can be assessed.

\section*{Framing}
This paper has relevance for any field that makes use of landcover data, so I want to try target a high visibility journal. Nature would be nice, but that is a big ask, The problem is that an editor might look at this very quickly and say that is too specialized or local, so it has to have a pretty good sales pitch if we are to avoid having it go into a trade journal (e.g. \emph{Remote Sensing of Environment}). PloS One might be a good fit, given that it accepts papers across multiple disciplines provided they are technically sound. Thoughts (including ideas for a better title)?

General approach/rationale:

\begin{enumerate}
  \item Satellite-derived landcover data are a key ingredient for understanding many aspects of environmental, social, and economic change. Examples:
    \begin{itemize}
      \item Food production estimates (e.g. yield gaps) - anything by RamanFoley, Rachel's work, etc.
      \item Carbon stocks (Tier 1 vegetated carbon estimates) - Ruesch and Gibbs
      \item Habitat loss data.
      \item Attempts to estimate land-atmosphere interactions, hydrological models, etc. 
    \end{itemize}
  \item It is understood that there are errors inherent in these data, particularly global-scale products, and that there is great discrepancy between these. This is particularly true of developing regions that will experience the largest amount of land surface change this century, such as Africa.  
    \begin{itemize} 
      \item Cite to Fritz's papers (from geowiki). 
      \item Emphasize what they did so that we can differentiate--they look primarily at disagreements between products, and how much it affects country level estimates. Also focus on older generation datasets. 
    \end{itemize}
  \item What is less known is how much bias these landcover errors might introduce into research findings that depend on areas estimates drawn from landcover data. Such an assessment depends on having accurate, spatially extensive landcover data to assess large areas, which are typically not available. 
  \item In this study, we use a unique, high accuracy agricultural landcover dataset to assess the biases in the most current landcover datasets, looking at bias as a function of spatial scale and landcover fraction.  
  \item Our goal is to provide insight--for users of landcover products across a range of disciplines--into: 
    \begin{itemize}
      \item The scales at which area-based inferences drawn from global landcover products are appropriate
      \item How variations within landcover properties may influence bias
      \item Which datasets are most appropriate for making inferences
    \end{itemize}
\end{enumerate}

\section*{Key Results}
\begin{enumerate}
  \item Bias as a function of scale
    \begin{enumerate}
    \item At 1 km resolution all landcover products are still fairly biased.  Actual bias is lowest in the geowiki dataset (+5\% under-estimate, on average, Figure 2), followed by the higher resolution South Africa landcover (SALC) dataset (8\% over-estimate, Figure 2), but the absolute bias is much higher in geowiki (23\%, Figure 3) than the SA dataset 11\%). The other products (MODIS and GlobCover) are grossly biased at the 1 km resolution (21-34\% under-estimates; 31-35\% mean bias).
      \begin{itemize}
         \item This discrepancy between actual and absolute biases for the two datasets (geowiki \& SALC) reflect the spatial variability of biases--geowiki underestimates cropland fraction in the west and overestimate in the north and east, whereas SALC's biases are all of over-estimation. 
      \end{itemize}
    \item Bias drops to acceptable levels quickly for geowiki--at 5X5 km, mean bias is just 1\% (overestimated). The absolute bias for this dataset is 10\% or lower from 10X10 km resolution and coarser.  
    \item The SA dataset's bias is fairly consistent but low across all levels of aggregation, amounting to no more than an 8\% overestimate of cropland with absolute bias of similar magnitude. 
    \item MODIS and GlobCover biases (mostly of underestimation) do not dissipate until the higher levels of aggregation. MODIS's actual bias (under-estimation) falls below 10\% at 20 km resolution, but the absolute bias remains above 10\% until more than a 100-fold aggregation is done ($>$100 km resolution).  For GlobCover, it needs to be more than 200-fold.  
    \end{enumerate}
  \item Bias as a function of cropland cover
    \begin{enumerate}
      \item Biases are highest when pixels contain intermediate levels of cropland cover.  This is particularly evident in the distributions of absolute biases (Figure 3) at the finer spatial scales ($\leq$20 km). The highest biases are associated with cropland cover of 50-75\%.
       \begin{itemize} 
         \item Classification algorithms are thus more error-prone where landcover is mixed/heterogenous.
        \end{itemize}
      \item The exception to this lies in the GlobCover dataset, where bias primarily increases as a function of cropland cover. The reason for this is that GlobCover's cropland classes do not provide for 100\% cropland cover, so aggregation tends to exacerbate underestimates. 
      \begin{itemize} 
        \item An example illustrats this: take 4 1 km pixels, 2 of which are 100\% cropland, 2 of which are other cover types. Imagine a landcover product classifies 3 of these as cropland (2 correct, 1 an error of commission), using a cropland class that is defined as 50\% cropland. Aggregating the actual fraction by a factor of 4 will result in a new 4 km pixel having 50\% cropland, whereas aggregating the landcover product's pixels will give just 38\% cropland, even when factoring in the incorrect classification.)
        \end{itemize}
    \end{enumerate}
  \item Bias as a function of method
      \begin{enumerate}
      \item Higher resolution and ensemble-based approaches have less bias.
      \item geowiki represents a fusion of multiple coarse resolution data sources that has undergone extensive validation using a crowdsourcing approach
      \item the SALC dataset is based on 30 m landsat data, but incorporates a range of ancillary data and expert judgement
      \item MODIS and GlobCover data are effectively single source/single algorithm. 
    \end{enumerate}
\end{enumerate}

%\vspace{0.5 in}
\begin{figure}[H]
\includegraphics[width = \linewidth]{../figures/bias_map.pdf}
\caption{The spatial distribution of biases in cropland cover estimates from four landcover products, calculated by subtracting each of their cropland percentage estimates from that of GeoTerraImage's cropland data layer. The biases were calculated for several different levels of, ranging from the original 1 km resolution up to 300 km (only 1, 20, 50, and 100 km results are shown here). Blue areas are those where the landcover products underestimated croplands, and red areas are where they overestimated. The GlobCover and MODIS results are drawn from biases calculated from the intermediate cropland fraction estimates per pixel.}
\end{figure}
%\vspace{-0.3in}

\begin{figure}[H]
\begin{center}
\includegraphics[width = \linewidth]{../figures/bias_act.pdf}
\caption{The statistical distribution of biases in cropland percentage estimates, in relation to different levels of cropland cover measured by the GeoTerraImage data (grouped into classes spanning 5\%). The top row shows biases at 1 km resolution for each of the four assessed landcover datasets, including the 2.5$^{th}$, 25$^{th}$, 50$^{th}$, 75$^{th}$, and 95$^{th}$ percentile values (grey box plots), as well as the means (colored lines). Rows 2-5 illustrate how mean biases for each landcover product (rows) changes in relation to the level of aggregation (columns). Mean biases greater than 40\% are truncated in order to improve plot clarity (the maximum truncated value is provided at the top right of each plot in grey). The mean bias across all cropland classes is provided at the bottom of each panel. Each 5\% cropland classes is represented by its upper bound on the X-axis (e.g. 5 for the 0-5\% class).}
\end{center}
\end{figure}

\begin{figure}[H]
\includegraphics[width = \linewidth]{../figures/bias_abs.pdf}
\caption{The statistical distribution of \emph{absolute} biases in cropland percentage estimates, in relation to different levels of cropland cover measured by the GeoTerraImage data (grouped into classes spanning 5\%). The top row shows \emph{absolute} biases at 1 km resolution for each of the four assessed landcover datasets, including the 2.5$^{th}$, 25$^{th}$, 50$^{th}$, 75$^{th}$, and 95$^{th}$ percentile values (grey box plots), as well as the means (colored lines). Rows 2-5 illustrate how mean \emph{absolute} biases for each landcover product (rows) changes in relation to the level of aggregation (columns). Mean biases greater than 40\% are truncated in order to improve plot clarity (the maximum truncated value is provided at the top right of each plot in grey). The mean \emph{absolute} bias across all cropland classes is provided at the bottom of each panel. Each 5\% cropland classes is represented by its upper bound on the X-axis (e.g. 5 for the 0-5\% class).}
\end{figure}

\section*{Implications}
\begin{enumerate}
  \item At finer resolutions, only high resolution or newer, fusion-based landcover products are appropriate for making area-based inferences
    \begin{itemize}
      \item At 1-10 km resolution, high resolution products are least biased. Such products are typically not available globally, but new technologies (processing, high resolution imagery) are making it increasingly feasible to generate them (cite Hansen's forest product)
      \item The geowiki dataset has sufficiently low biases at resolutions of 20 km and coarser. It also demonstrates a methodological aspect that is evident in other fields (climate modeling, crop modeling, species distribution modeling)--ensemble based approaches return more accurate results than single method approaches.  
    \end{itemize}
  \item If you have to rely on single source datasets, e.g. for mapping landcover change, than substantial aggregation is necessary before area-based inferences can be made with confidence. Country-scales or higher should be appropriate. 
  \begin{itemize} 
    \item Illustrative examples: 
      \begin{itemize} 
        \item Examples from actual papers - Convergence of yield or carbon estimates at regional scale from our (Tim) savannas paper, emphasizing discrepancy between cropland fractions; discuss that Tier 1 carbon estimates from Gibbs et al use the GLC2000 dataset, the predecessor of GlobCover.  
         \item Maybe another example from the literature - habitat conversion, or crop yield estimates, or thinking about how a land surface model might be impacted (Justin?)
      \end{itemize}
    \item But care should be taken when aggregating--the characteristics of cropland in a country will influence this bias. 
      \begin{itemize}
        \item For example, if a high percentage of a region's cropland is of a mixed/mosaic type (e.g. fields interleaved with natural vegetation, such that $\sim$50\% is cropland), then aggregated biases may be higher.  
        \item This is the sort of situation that is more likely to occur in the smallholder dominated systems, which are dominant in many other African countries (where a high proportion of trees in fields may exacerbate classification errors). It is less likely in larger-scale commercial agriculture of the sort that dominates South Africa. 
        \item We could not test this aspect here, as we had to explicitly exclude smallholder/communal farmlands from our analysis because they were not mapped with the same precision, so the degree to which bias in this sort of cropland differs from commercial croplands requires further investigation \textbf{this sets up one aspect of the second paper, Stephanie}.  
      \end{itemize}
  \end{itemize}

\end{enumerate}

\section*{Methods}
\subsection*{Figures, tables, stats needed}
\begin{itemize}
  \item Distribution of accuracy assessment plots
  \item Close-up of polygons (maybe)
  \item Country level bias means/bar plot of means
\end{itemize}

\end{document}

### 2014-12-29 external/notes/cropland-bias-notes.tex
Updated notes to reflect change from 20 to 25 km resolution, and land surface model bias example.

@@
       \item Attempts to estimate land-atmosphere interactions, hydrological models, etc.
+[Here we will run land surface model with GTI cropland fraction providing the cropland surface, and we will then alter this using the bias rasters to see how this propagates into ET and runoff estimation errors, etc]
     \end{itemize}
@@
 \caption{The spatial distribution of biases in cropland cover estimates from four landcover products, calculated by subtracting each of their cropland percentage estimates from that of GeoTerraImage's cropland data layer.
-The biases were calculated for several different levels of, ranging from the original 1 km resolution up to 300 km (only 1, 20, 50, and 100 km results are shown here).
+The biases were calculated for several different levels of, ranging from the original 1 km resolution up to 300 km (only 1, 25, 50, and 100 km results are shown here).
 Blue areas are those where the landcover products underestimated croplands, and red areas are where they overestimated.

### 2015-06-02 Rmd/compare-landcover.Rmd
Moved entire folder to Dropbox/publications. Renamed vignettes folder to Rmd; committed some changes to Rd files that were lingering.

---
title: "Calculating cropland biases"
author: "Lyndon Estes"
date: "01 October 2014"
output:
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes
---

## Overview
This section of the analysis calculates the differences between the GTI cropland percentages and those from the various landcover products. It draws from results generated by the [cropland pre-processing](https://github.com/PrincetonUniversity/SAcropland/blob/master/vignettes/sa-cropland-analysis.Rmd). (Note to self here: I want to have a transportable link between chapters here, rather than a direct link to the other RMD document in 
github).

## Analyses
### Load in data 
1. Set up paths
```{r, eval = FALSE}
library(lmisc)
library(raster)
library(SAcropland)
p_root <- proj_root("SAcropland")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
```

2. Load cropland grids 
  + Derived from global landcover products (note: it turns out GLC-Share and geowiki are the same thing)

```{r, eval = FALSE}
salc <- raster(full_path(p_data, "sa_ag_masked.tif")) * 100
globsa_list_sum <- lapply(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
modsa_list_sum <- lapply(full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
glcsa <- raster(full_path(p_data, "glcsa_masked.tif"))
geow <- raster(full_path(p_data, "geowikisa_masked.tif"))
lclist <- c(salc, globsa_list_sum, modsa_list_sum, glcsa, geow)  # into list
names(lclist) <- c("sa30", "globmin", "globmu", "globmax", "modmin", "modmu", "modmax", "glcsh", "geow")
```

  + SA landcover data
```{r, eval = FALSE}
load(full_path(p_data, "MZshapes.Rdata"))  
paths <- full_path(p_data, dir(p_data, paste("cover*.*sum_mask*", sep = "")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
```
[[back to top]][Overview]

### Calculate differences at different levels of aggregation
This steps draws on functions I wrote to calculate the actual and absolute differences between a given "master" grid and lists of other grids. This comparison is done at multiple aggregations, ranging from 1-600 km resolution cropland averages. The lists have two levels of nesting, here the upper provides the resolution, the inner provides the different landcover datasets.

```{r, eval = FALSE}
# Aggregate rasters
#fact <- c(5, 10, 25, 50, 100, 200, 300)
fact <- 25
lc_agg <- aggregate_rast_list(fact, lclist)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters

# Calculate differences
dlist_act <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)),
                            gti_agg, lc_agg, "a - b", silent = FALSE)  # actual
dlist_abs <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)), 
                            gti_agg, lc_agg, "abs(a - b)", silent = FALSE)  # absolute
dlist_gti <- rast_list_math(list("g2007", names(gti_agg), "g2011"), 
                            gti_agg, gti_agg, "a - b", silent = FALSE)

save(dlist_act, file = full_path(p_data, "d_grid_act.rda"))
save(dlist_abs, file = full_path(p_data, "d_grid_abs.rda"))

# Write out the gti cover raster and the actual bias rasters for running impact of bias examples in land
# surface model (25 km resolution)
r <- gti_agg$f25$g2011
projection(r) <- projection(salc)
writeRaster(r, filename = full_path(p_data, "gti_2011_25km.tif"))
b <- brick(stack(lapply(c("sa30", "globmu", "modmu", "geow"), function(x) {
  dlist_act[[x]]$f25$g2011
})))
projection(b) <- projection(salc)
writeRaster(b, filename = full_path(p_data, "lc_bias_25km.tif"))

# r <- raster(full_path(p_data, "gti_2011_25km.tif"))
# lcb <- brick(full_path(p_data, "lc_bias_25km.tif"))
# plot(r - lcb, axes = FALSE, box = FALSE)
# plot(lcb)
# plot(round(lc_agg$f25$geow - (r - lcb$lc_bias_25km.4), 7))

```
[[back to top]][Overview]
### Bias as function of cropcover
+ First create NA mask at 1 km resolution, then weighting raster based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc) 
```{r, eval = FALSE}
# NA mask and weights
namask <- !is.na(gti[[1]])  
area_wgts <- aggregate_rast_list(fact, list(namask), fun = sum)
namask[namask == 0] <- NA
writeRaster(namask, filename = full_path(p_data, "namask.tif"), overwrite = TRUE)

# cropland cover bins
binv <- seq(0, 100, 5)
gti_bins <- lapply(gti_agg, function(x) lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE)))
```

+ Calculate biases, pooling MODIS and GlobCover across all version
Step 1. Reshape lists, so that all MODIS and GlobCover variants are pooled by scale
```{r, eval = FALSE}
# Set up indexing parameters
fact_all <- c(1, fact)# full aggregation levels
lev_vec <- paste("f", fact_all, sep = "") 
nm_rt <- c("sa30", "glob", "mod", "glcsh", "geow")

# Reshape the list of actual landcover differences
dlists <- lapply(list(dlist_act, dlist_abs), function(i) {
  resh <- lapply(nm_rt, function(x) {  # select sensor
    l2 <- lapply(lev_vec, function(y) {  # isolate by scale
      l3 <- lapply(i[grep(x, names(i))], function(z) z[[y]])  # sensor by scale
      named_out(l3, names(i)[grep(x, names(i))])
      })
    named_out(l2, lev_vec)
  })
  named_out(resh, nm_rt)
})
names(dlists) <- c("act", "abs")
rm(dlist_act, dlist_abs)  # remove original to make space

```

+ Then calculate biases across datasets
```{r, eval = FALSE}
dang <- Sys.time()
dstats_all <- lapply(dlists, function(i) {
  print("processing 1 of 2 main bias lists")
  dstats <- lapply(names(i), function(x) {
    print(paste("..", x))
    l1 <- lapply(lev_vec, function(y) {
      print(paste("....", y))
      d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(j) {
        print(paste("......bin", j))
        d2 <- do.call(rbind, lapply(1:length(i[[x]][[y]]), function(z) {
          print(paste("........variant", z))
          l3 <- i[[x]][[y]][[z]]
          d3 <- do.call(rbind, lapply(1:length(l3), function(k) {
            rs <- gti_bins[[y]][[k]] == j    
            rs[rs == 0] <- NA
            o <- rs * l3[[k]]
            w <- area_wgts[[y]][[1]] * rs
            oo <- cbind.data.frame("lc" = rep(names(i)[z], length(w)), 
                                   getValues(o), getValues(w))
          }))
          d3f <- d3[!is.na(d3[, 2]), ]
        }))
        if(nrow(d2) > (length(i[[x]][[y]]) * 10)) {
        qs <- c(nrow(d2), box_stats(d2[, 2], weighted = TRUE, 
                                    weight.opts = list("weights" = d2[, 3])))
        } else if((nrow(d2) > 0) & (nrow(d2) < length(i[[x]][[y]]) * 20)) {
          qs <- c(rep(NA, 6), Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], 
                                              na.rm = FALSE))
        } else {
          qs <- rep(NA, 7)
        }
      }))
      cbind(binv[-1], d1)
      })
    named_out(l1, lev_vec)
  })
  named_out(dstats, names(i))
})
dut <- Sys.time() - dang  # 33 minutes
save(dstats_all, file = full_path(p_data, "bias_stats.rda"))
```
[[back to top]][Overview]

+ Finally, differences between gti2007 and 2011
```{r, eval = FALSE}
dstats_gti <- lapply(lev_vec, function(x) {
    #x <- lev_vec[[2]]
    i <-  dlist_gti$g2007[[x]][[1]]
    print(paste("..", x))
    d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(y) {
      #y <- 1
      print(paste("....", y))
      rs <- gti_bins[[x]][[2]] == y    
      rs[rs == 0] <- NA
      o <- rs * i
      w <- area_wgts[[x]][[1]] * rs
      oo <- cbind.data.frame(rep(y, length(w)), getValues(o), getValues(w))
      oof <- oo[!is.na(oo[, 2]), ]
      if(nrow(oof) > 10) {
        qs <- c(nrow(oof), box_stats(oof[, 2], weighted = TRUE, 
                                     weight.opts = list("weights" = oof[, 3])))
      } else if((nrow(oof) > 0) & (nrow(oof) < 10)) {
        qs <- c(rep(NA, 6), Hmisc::wtd.mean(oof[, 2], weights = oof[, 3], na.rm = FALSE))
      } else {
        qs <- rep(NA, 7)
      }
    }))
    cbind(binv[-1], d1)
})
save(dstats_gti, file = full_path(p_data, "bias_stats_gti.rda"))

```
Checks: compared number of non-NA cells coming out of each GTI version at each level of aggregation (fine); compared cell counts between previous committed version and this one (difference exists, but due to fact that gtimu is no longer considered; checked that previous versions of difference rasters and reshapes of them were referencing the correct GTI-landcover bias rasters (i.e. to make sure rasters weren't incorrectly indexed: all fine)); box_stats with weighted quantile option compared to previous commit where wtd.quantile was used separately (fine). Compared results from original roughed out version of approach for calculating bias statistics with results from new outputs list (using geowiki at 20 km aggregation as the comparison)

## Next steps
Plot bias results.  
[[back to top]][Overview]

### 2015-06-02 Rmd/draw-subsistence-map-sample.Rmd
Moved entire folder to Dropbox/publications. Renamed vignettes folder to Rmd; committed some changes to Rd files that were lingering.

---
title: "Select subsistence mapping sample"
author: "Lyndon Estes"
date: "24 November 2014"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
---

##Selecting areas for mapping in small-scale/subsistence cropland
GTI did not map communal farmlands with the same precision as commercial and other croplands in SA, so we need to come up with a sub-sample of these areas for further mapping and analysis. _Note: This will probably go into another project eventually._

```{r, eval = FALSE}
library(raster)
library(lmisc)
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/external/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
```

Load in GTI2011 dataset, the SA grid (which provides the grid IDs for the whole country), and then multiple the two to assign sample IDs to just the subsistence/communal farming areas
```{r, eval = FALSE}
cover2011 <- brick(full_path(p_data, "cover2011.tif"))

# load in SA grid
sagrid <- brick(full_path(p_data, "sagrid.tif"))

subs_crops <- cover2011[[2]] > 0  # select those areas
subs_crops[subs_crops == 0] <- NA
projection(subs_crops) <- projection(sagrid)

# create sample ids for the subsistence areas
subs_sids <- subs_crops * sagrid
#subsgrid <- subs_sids
```

Pull a random sample from those, and convert to polygons
```{r, eval = FALSE}
set.seed(234)
samp <- sampleRandom(subs_sids, 200, na.rm = TRUE, asRaster = TRUE)
subsgrid[!subsgrid %in% samp] <- NA
subs_polys <- rasterToPolygons(samp, dissolve = TRUE)
colnames(subs_polys@data) <- "ID"
subs_polys$name <- paste("SA", subs_polys$ID, sep = "")
writeOGR(subs_polys, dsn = full_path(p_data, "SA_subsistence_sample.sqlite"), "SA_subs", driver = "SQLite", 
         dataset_options = c("SPATIALITE = yes"))
#writeOGR(obj = subs_polys, dsn = p_data, "SA_subsistence_sample", driver = "ESRI Shapefile")  # identical

```

Then create a template for mapping fields in QGIS (with openlayers plugins providing the satellite imagery, as with mappingafrica.princeton.edu), and also do a randomized reduction of the initial draw of 200 (down to 100). This preserves a larger draw if we find we need more than 100 mapped samples. 
```{r, eval = FALSE}
map_polys <- subs_polys[1, ]
map_polys$fid <- 1

set.seed(234)
subs100 <- sample(subs_polys$ID, size = 100, replace = FALSE)
write <- write(subs100, file = full_path(p_data, "fields100.txt"))

writeOGR(subs_polys[subs_polys$ID %in% subs100, ], dsn = full_path(p_data, "SA_subsistence_sample100.sqlite"),
         "SA_subs100", driver = "SQLite", dataset_options = c("SPATIALITE = yes"))

```

### 2015-06-02 Rmd/fetch-land-cover.Rmd
Moved entire folder to Dropbox/publications. Renamed vignettes folder to Rmd; committed some changes to Rd files that were lingering.

---
title: "Fetching and pre-processing landcover data"
author: "Lyndon Estes"
date: "04/11/2014"
output: html_document
---

### Introduction
Fetching and pre-processing of global landcover datasets to be analyzed relative to the SA dataset: the MODIS 2011 landcover product for South Africa was downloaded, the GLC-SHARE dataset, the 2009 GlobCover product, South Africa's 2009 landcover product, and geo-wiki's cropland fusion map. These were converted to 1 km$^2$ percentage cropland estimates after making several versions of MODIS and GlobCover products to account for mixed cropland classes.  A further processing was applied to the two GTI derived cover images.

### Analyses
#### Set-up
```{r, eval = FALSE}
library(SAcropland)
library(gdalUtils)
library(RCurl)
library(RPostgreSQL)
library(rgdal)
library(rgeos)
library(raster)
library(lmisc)
```

Paths
```{r, eval = FALSE}
#setwd("/u/lestes/analyses/SAcropland/")
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/external/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
```

Background data, projections, database connections, and SA grid
```{r, eval = FALSE}
# Get SA shape
if(!file.exists(full_path(p_data, "MZshapes.Rdata"))) {
  system("wget http://dl.dropboxusercontent.com/u/31890709/MZshapes.Rdata")
}
load(full_path(p_data, "MZshapes.Rdata"))
sa_buf <- gBuffer(sa.shp, width = 3000)

# Connection
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "SouthAfricaSandbox", user = "postgis", password = "P0stG1S")

# projections, etc
prjsrid <- 97490  # Albers
prj_sql <- paste("select proj4text from spatial_ref_sys where srid=", prjsrid, sep = "")
prjstr <- dbGetQuery(con, prj_sql)$proj4text
gcsid <- 4326  # GCS
gcs_sql <- paste("select proj4text from spatial_ref_sys where srid=", gcsid, sep = "")
gcsstr <- dbGetQuery(con, gcs_sql)$proj4text

# Convert SA to GCS for cropping extent
sa_buf_gcs <- spTransform(sa_buf, CRSobj = CRS(gcsstr))

# Create 1 km grid from sa1kilo to rectify grids against that 
if(!file.exists(full_path(p_data, "sagrid.tif"))) {  # Wrap in if statement in case grid already made
  sql <- paste("SELECT gid, id, ST_AsText(ST_Centroid(geom)) as center FROM sa1kilo")
  geom.tab <- dbGetQuery(con, sql)
  coord.mat <- do.call(rbind, lapply(1:nrow(geom.tab), function(y) {
    strip <- gsub("POINT|\\(|\\)", "", geom.tab[y, 3])
    coord.mat <- do.call(rbind, strsplit(strsplit(strip, ",")[[1]], " "))
  }))
  class(coord.mat) <- "numeric"
  point.tab <- cbind(geom.tab[, 1:2], coord.mat)
  colnames(point.tab)[3:4] <- c("x", "y") 
  pointsXYZ <- point.tab[, c(3:4, 2)]
  sa_r <- rasterFromXYZ(pointsXYZ)
  projection(sa_r) <- sa.buf@proj4string
  sa_r <- writeRaster(sa_r, filename = full_path(p_data, "sagrid.tif"), overwrite = TRUE)
} else {
  sa_r <- raster(full_path(p_data, "sagrid.tif"))
}
```

#### Bring in landcover datasets
Let's start with the improved/fusion datasets, GLC share and geo-wiki's
```{r, eval = FALSE}
if(!file.exists(full_path(p_data, "glcsa_masked.tif"))) {  # see if file exists, to avoid redoing step
  path <- "/u/lestes/spatial_data/glc_share/"
  url <- "http://www.fao.org/geonetwork/srv/en/resources.get?id=47948&fname=GlcShare_v10_02.zip&acces=private"
  download.file(url, method = "auto", destfile = "glc_share.zip")
  unzip("glc_share.zip", exdir = "glc_share")
  glc <- raster(full_path(path, "glc_shv10_02.Tif"))
  projection(glc) <- gcsstr
  glcSA <- crop(glc, y=sa_buf_gcs, file = full_path(path, "glcSA.tif"), overwrite = TRUE)  # crop to SA
  gdalwarp(srcfile = glcSA@file@name, dstfile = full_path(path, "glcSA_alb.tif"), t_srs = prjstr, 
           tr = c(1000, 1000), r = "bilinear")
  glcSA_alb <- raster(full_path(path, "glcSA_alb.tif"))

  # Warp to SA grid
  glcsa <- resample(glcSA_alb, sa_r, method = "bilinear", filename = full_path(path, "glcSA_alb_rect.tif"))
  glcsa_m <- mask(glcsa, sa_r, file = full_path(p_data, "glcsa_masked.tif"), overwrite = TRUE)
} else {
  glcsa_m <- raster(full_path(p_data, "glcsa_masked.tif"))
}

if(!file.exists(full_path(p_data, "geowikisa_masked.tif"))) {  # see if file exists, to avoid redoing step 
   # first downloaded manually from password protected geo-wiki site
  path <- "/u/lestes/spatial_data/geowiki/"
  unzip("cropland_hybrid_14052014v8.zip", exdir= "geowiki") 
  geowiki <- raster(full_path(path, "Hybrid_14052014V8.img"))
  projection(geowiki) <- gcsstr
  geowikiSA <- crop(geowiki, y=sa_buf_gcs, file = full_path(path, "geowikiSA.tif"))  # crop to extent of SA
  gdalwarp(srcfile = geowikiSA@file@name, dstfile = full_path(path, "geowikiSA_alb.tif"), t_srs = prjstr, 
           tr = c(1000, 1000), r = "bilinear")
  geowikisa_alb <- raster(full_path(path, "geowikiSA_alb.tif"))
  geowikisa <- resample(geowikisa_alb, sa_r, method = "bilinear", 
                        filename = full_path(path, "geowikisa_alb_rect.tif"))    # Warp to SA grid
  geowikisa_m <- mask(glcsa, sa_r, file = full_path(p_data, "geowikisa_masked.tif"), overwrite = TRUE)
} else {
  geowikisa_m <- raster(full_path(p_data, "geowikisa_masked.tif"))
}
```

And the standard products, Globcover 2009 and MODIS
```{r, eval = FALSE}
path <- "/u/lestes/spatial_data/globcover_2009/"
if(!file.exists(full_path(path, "globSA_alb.tif"))) {
  url <- "http://due.esrin.esa.int/globcover/LandCover2009/Globcover2009_V2.3_Global_.zip"
  download.file(url, method="auto", destfile = "globcover_2009.zip")
  unzip("globcover_2009.zip", exdir= "globcover_2009")
  glob <- raster(full_path(path, "GLOBCOVER_L4_200901_200912_V2.3.tif"))
  globSA <- crop(glob, y=sa_buf_gcs)  # crop to extent of SA
  gdalwarp(srcfile = globSA@file@name, dstfile = full_path(path, "globSA_alb.tif"), t_srs = prjstr, 
           tr = c(300, 300))
  globSA_alb <- raster(full_path(path, "globSA_alb.tif"))
} else {
  globSA_alb <- raster(full_path(path, "globSA_alb.tif"))
}

# MODIS landcover 
moddir <- "/u/lestes/spatial_data/MCD12Q1/"
if(!file.exists(full_path(moddir, "modis_type_1_alb.tif"))) {
  tiles <- expand.grid("h" = c(19, 20), "v" = c(11, 12))
  tiles <- apply(tiles, 1, function(x) paste("h", x[1], "v", x[2], sep = ""))
  url <- "http://e4ftl01.cr.usgs.gov/MOTA/MCD12Q1.051/2011.01.01/"
  tile.names <- getURL(url, verbose=TRUE, dirlistonly = TRUE) 
  tile.names1 <- strsplit(tile.names, "alt")[[1]]
  tile.names2 <- unique(substr(tile.names1, 20, 64))
  modnames <- tile.names2[sapply(tiles, function(x) grep(x, tile.names2))]
  lapply(modnames, function(x) {
    url <- paste("http://e4ftl01.cr.usgs.gov/MOTA/MCD12Q1.051/2011.01.01/", x, sep = "")
    download.file(url, method="auto", destfile = x)
  })
  dir.create("MCD12Q1")
  file.copy(from = dir(pattern = "hdf"), to = paste(getwd(), "MCD12Q1", dir(pattern = "hdf"), sep = "/"))
  file.remove(dir(pattern = "hdf"))

  modis <- dir(moddir)

  # Translate to geotiff
  lapply(1:5, function(x) {
    batch_gdal_translate(infiles = modis, outdir = moddir, 
                         outsuffix = paste("_LC_", x, ".tif", sep = ""), sd_index = x)  
  })
  sdnames <- lapply(1:5, function(x) dir(moddir, pattern  = paste("LC_", x, ".tif", sep = "")))

  # warp to albers, mosaic, and clip to SA extent
  lapply(1:5, function(x) {
    print(paste("processing type", x))
    gdalwarp(srcfile = sdnames[[x]], dstfile = paste("modis_type_", x, "_alb.tif", sep = ""), 
             t_srs = prjstr, tr = res(r), te = bbox(sa.buf)[1:4])
  })  
  mod_1 <- raster("modis_type1_alb.tif")
} else {
  mod_1 <- raster(full_path(moddir, "modis_type_1_alb.tif"))
}

```

And then South Africa's landcover database (30 m)
```{r, eval = FALSE}
path <- "/u/lestes/spatial_data/sa_lc_2009/"
setwd("/u/lestes/spatial_data/")
lcnm <- c("sa_ag.tif", "sa_af.tif")
if(length(dir(p_data, pattern = "sa_.*._masked.tif")) != 2) {
  url <- "http://planet.uwc.ac.za/BGISdownloads/landcover_2009.zip"
  download.file(url, method="auto", destfile = "sa_landcover_2009.zip")
  unzip("landcover_2009.zip", exdir = "sa_lc_2009")

  setwd(path)
  gdal_translate(src_dataset = "landcover", dst_dataset = "sa_lc_2009.tif", of = "GTiff", ot = "BYTE")
  #system(paste("rm -r", full_path(path, "landcover")))
  #file.remove("../landcover_2009.zip")

  # process masks--extract extract landcover class 2, cultivation, and class 6, plantations, resample, mask 
  # with gdal tools
  av <- c(2, 6)
  a <- "sa_lc_2009.tif"
  dang <- Sys.time() 
  salc_l <- lapply(1:2, function(j) {
    print(paste("extracting cover for", lcnm[j]))
    gdal_calc(cstr = paste("A==", av[j], sep = ""), x = list("A" = a), type = "Byte", filename = lcnm[j])
    nm <- full_path(p_data, paste(gsub("\\.tif", "", lcnm[j]), "_masked.tif", sep = ""))
    ext <- sapply(c("xmin", "ymin", "xmax", "ymax"), function(x) slot(extent(sa_r), x))
    print(paste("warping and masking", gsub(p_data, "", nm)))
    gdalwarp(srcfile = lcnm[j], t_srs = projection(sa_r), dstfile = nm, r = "average", ot = "Float32", 
             te = ext, srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), of = "GTiff", verbose = TRUE,
             overwrite = TRUE)
    file.remove(lcnm[j])
    r <- raster(nm)
  })
  Sys.time() - dang

} else {
  lcmnm <- paste(gsub("\\.tif", "", lcnm), "_masked.tif", sep = "")
  salc_l <- lapply(lcmnm, function(x) raster(full_path(p_data, x)))
} 
setwd(p_data)
```

Create mask to remove sugarcane growing regions
```{r, eval = FALSE}
if(!file.exists(full_path(p_data, "kzn_cane_masked.tif"))) {
  path <- "/u/lestes/spatial_data/kzn_landcover/"
  setwd(kznpath)
  gdal_translate(src_dataset = "kznlc11v1w31", dst_dataset= "kznlandcover.tif", of = "GTiff")
  system(paste("rm -r", full_path(path, "kznlc11v1w31")))
  kznlc <- raster(full_path(npath, "kznlandcover.tif"))

  # system call to gdal_calc.py to extract landcover class 2, cultivation
  kznlconm <- "kzn_cane.tif"  # file.remove(kznlconm)
  kzninnm <- strsplit(kznlc@file@name, "/")[[1]][length(strsplit(kznlc@file@name, "/")[[1]])]
  # kzninnm <- strsplit(kznlc.ss@file@name, "/")[[1]][length(strsplit(kznlc.ss@file@name, "/")[[1]])]
  pcalc <- paste("gdal_calc.py -A ", kzninnm, " --outfile=", kznlconm, 
                 " --overwrite --calc='(A==9)|(A==10)'", sep = "")
  system.time(system(pcalc))  # 49 seconds
  kzn_cane <- raster(full_path(path, kznlconm))
  # cext <- extent(c(20000, 40000, -3220000, -3200000))  # crops to small extent for examination
  # kznlc.ss <- crop(kznlc, cext, filename = full_path(kznpath, "kznlandcover_crop.tif"), overwrite = TRUE)
  # plot(kznlc.ss)
  # scane1 <- (kznlc.ss == 9); scane2 <- (kznlc.ss == 10)
  # plot(scane1, add = TRUE, col = c("transparent", "red"))
  # plot(scane2, add = TRUE, col = c("transparent", "purple"))
  # chk <- crop(kzn_cane, cext)
  # plot(chk); plot(scane1, add = TRUE, col = c("transparent", "red"), legend = FALSE) 
  # plot(scane2, add = TRUE, col = c("transparent", "red"), legend = FALSE)
  # file.remove("kznlandcover_crop.tif")
  # rm(chk, kznlc.ss, scane1, scane2)
  file.remove("kznlandcover.tif")

  # Warp and resample to SA grid
  setwd(p_data)
  gdalwarp(srcfile = kzn_cane@file@name, t_srs = projection(sa_r), dstfile = "kzn_cane_1km.tif", 
           r = "average", ot = "Float32", srcnodata = 255, tr = c(1000, 1000), of = "GTiff", 
           verbose = TRUE, overwrite = TRUE)
  kzn_1km <- raster(full_path(p_data, "kzn_cane_1km.tif"))
  kzn_1km_r <- resample(kzn_1km, sa_r, method = "bilinear")
  kzn_1km_m <- mask(kzn_1km_r, sa_r, filename = full_path(p_data, "kzn_cane_masked.tif"), 
                    overwrite = TRUE)
} else {
  kzn_1km_m <- raster(full_path(p_data, "kzn_cane_masked.tif"))
}

```

Create low, medium, high cover fraction variants from MODIS classes 12 and 14 (from the IGBP-DIS scheme). Class 14 is a cropland mosaic class, with minimum of 10% and a max of 60%, so create variants of 10%, 35% (mean) and 60%. 
```{r, eval = FALSE}
if(any(!file.exists(full_path(p_data, dir(p_data, pattern = "mod_1*.*1kmrect_sum.tif"))))) {
  # MODIS, using main classification scheme (IGBP-DIS)
  mod_1_crop <- stack(mod_1 == 12, mod_1 == 14)
  mod_1_cropb <- brick(mod_1_crop)
  names(mod_1_crop) <- c("cropland", "mosaic")

  # assign percentages
  minset <- c(100, 10)  # assume 10% is minimum crop cover for MODIS IGBP mosaic class
  meanset <- c(100, 35)  # assume 35% is mean crop cover
  maxset <- c(100, 60)  # assume 60% is mean crop cover  
  modsa_min <- assignLCPct(mod_1_cropb, minset, fname = full_path(p_data, "mod_1_min.tif"))
  modsa_mu <- assignLCPct(mod_1_cropb, meanset, fname = full_path(p_data, "mod_1_mu.tif"))
  modsa_max <- assignLCPct(mod_1_cropb, maxset, fname = full_path(p_data, "mod_1_max.tif"))

  # aggregate, resample, mask to SA grid
  modsa_list <- lapply(list(modsa_min, modsa_mu, modsa_max), function(x) {
    nm <- gsub("//.tif", "", x@file@name)
    print(paste("Processing", nm))
    print("...aggregating")
    agg <- aggregate(x, fact = 2, na.rm = TRUE)  
    print("...resampling")
    crop1km <- resample(agg, sa_r, method = "bilinear")
    print("...masking")
    crop1km <- mask(crop1km, mask = sa_r, filename = paste(nm, "_1kmrect.tif", sep = ""), overwrite = TRUE)
    return(crop1km)
  })

  # And them create single raster with summed proportions
  modsa_list_sum <- lapply(modsa_list, function(x) {
    nm <- paste(gsub("\\.tif", "", x@file@name), "_sum.tif", sep = "")
    paste(nm)
    calc(x, sum, filename = nm, overwrite = TRUE)
  })
} else {
  modsa_list_sum <- lapply(full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
}  
setwd(p_data)
```

And the same for GlobCover. We want Globcover class 11 (irrigated cropland), class 14 (rainfed), 20 (50-70% cropland), and 30 (20-50% cropland). For the latter two classes we create variants assumung the minimum, mean, and max of earh class
```{r, eval = FALSE}
if(any(!file.exists(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif"))))) {
  l <- lapply(c(11, 14, 20, 30), function(x) r <- globSA_alb == x)  # pull out classes
  s <- stack(l)
  globSA_cropb <- brick(full_path(p_data, "globSA_crop.grd"))
  rm(s, l)

  # assign percentages
  minset <- c(100, 100, 50, 20)
  maxset <- c(100, 100, 70, 50)
  meanset <- c(100, 100, 60, 35)
  globsa_min <- assignLCPct(globSA_cropb, minset, fname = full_path(p_data, "globSA_300_min.tif"))
  globsa_mu <- assignLCPct(globSA_cropb, meanset, fname = full_path(p_data, "globSA_300_mu.tif"))
  globsa_max <- assignLCPct(globSA_cropb, maxset, fname = full_path(p_data, "globSA_300_max.tif"))
  glob300list <- lapply(dir(pattern = "globSA_300+.+tif")[c(2:3, 1)], raster)

  # Aggregate them and resample and mask to SA grid
  globsa_list <- lapply(list(globsa_min, globsa_mu, globsa_max), function(x) {
    nm <- gsub("_300|.tif", "", x@file@name)
    print(paste("Processing", nm))
    print("...aggregating")
    agg <- aggregate(x, fact = 3, na.rm = TRUE)  # equivalent to summing, dividing by 900
    print("...resampling")
    crop1km <- resample(agg, sa_r, method = "bilinear")
    print("...masking")
    crop1km <- mask(crop1km, mask = sa_r, filename = paste(nm, "_1kmrect.tif", sep = ""), 
                    overwrite = TRUE)
    return(crop1km)
  })

  # And them create single raster with summed proportions
  globsa_list_sum <- lapply(globsa_list, function(x) {
    nm <- paste(gsub("\\.tif", "", x@file@name), "_sum.tif", sep = "")
    paste(nm)
    calc(x, sum, filename = nm)
  })
} else {
  globsa_list_sum <- lapply(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
} 
```

Bring in GTI cropland fractions to set up masks
```{r, eval = FALSE}
# Bring in SA landcover sets and mask
cover2007 <- brick(full_path(p_data, "cover2007.tif"))
cover2011 <- brick(full_path(p_data, "cover2011.tif"))

# sum these also with and without horticulture
sumfun <- function(x) sum(x, na.rm = TRUE)
gti <- lapply(list(cover2007, cover2011), function(y) {
  r <- calc(y[[-6]], sumfun)
  r[r > 100] <- 100
  writeRaster(r, file = gsub("\\.tif", "sum.tif", y@file@name), overwrite = TRUE)
})

# Leave commented out in case want to bring back horticulture class
# gti_hort <- lapply(list(cover2007, cover2011), function(y) {
#   r <- calc(y, sumfun)
#   r[r > 100] <- 100
#   writeRaster(r, file = gsub("\\.tif", "sum_h.tif", y@file@name), overwrite = TRUE)
# })

# Remove communal farmlands
sust_mask <- !is.na(cover2007[[2]]) | !is.na(cover2011[[2]])  # remove communal areas from both dates
sust_mask[sust_mask > 0] <- NA

# Set up SA mask including sugarcane and forestry -- add in Mpumalanga sugar cane when it becomes available
m1 <- kzn_1km_m
m1[is.na(m1)] <- 0
m1[m1 > 0] <- NA
m2 <- salc_l[[2]]
m2[is.na(m2)] <- 0
m2[m2 > 0] <- NA
sa_masks <- sa_r > 0
sa_masks <- m1 + m2 + sa_masks
rm(m1, m2)
full_mask <- writeRaster(sa_masks + sust_mask, file = full_path(p_data, "mask.tif"), overwrite = TRUE)

# Apply masks
lapply(gti, function(x) {
  out_name <- paste(gsub("\\.tif", "", x@file@name), "_mask.tif", sep = "")
  mask(x, full_mask, file = out_name, overwrite = TRUE)
})

```

### 2015-06-02 Rmd/landcover-bias-plots.Rmd
Moved entire folder to Dropbox/publications. Renamed vignettes folder to Rmd; committed some changes to Rd files that were lingering.

---
title: "Plotting cropland biases"
author: "Lyndon Estes"
date: "06 October 2014"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
---

## Plotting bias figures
### Biases in space
Load in necessary datasets

```{r, eval = FALSE}
library(raster)
library(lmisc)
library(R.utils)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/external/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")

# Load in datasets
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
load(full_path(p_data, "d_grid_abs.rda"))  # actual diffence grids
load(full_path(p_data, "bias_stats.rda"))  # bias statistics
load(full_path(p_data, "bias_stats_gti.rda"))  # bias statistics
```

Plot maps of biases at different resolutions, for publication output (and display here). To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and then masked before plotting.   
```{r, eval = FALSE}
namask <- raster(full_path(p_data, "namask.tif"))
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data, excluding Modis and GlobCov extremes
lev <- c("f1", "f25", "f50", "f100")  # just show four levels

# disggregate selected rasters at selected levels
disagg <- lapply(snms, function(x) {
  l1 <- lapply(lev, function(y) {
    if(y == "f1") {
      r <- dlist_act[[x]][[y]]$g2011
    } else {
      r <- disaggregate(dlist_act[[x]][[y]]$g2011, fact = as.numeric(gsub("f", "", y)))
      r <- raster::mask(crop(r, namask), namask)
    }
  })
  named_out(l1, lev)
})
names(disagg) <- snms
```

Plot of selected rasters and levels 
```{r, eval = FALSE}
# Plotting colors
cols <- colorRampPalette(c("red", "grey80", "blue4"))
brks <- c(-100.1, -75, -50, -25, -10, 0, 10, 25, 50, 75, 100.1)
n_cols <- length(brks) - 1
legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "bias_map.pdf"), height = 6, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols(n_cols), 
         breaks = brks, legend = FALSE)
  if(j == 1) mtext(mcap[i], side = 2, line = 1, cex = cx)
  if(i == 1) mtext(lev2[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext, 
            legend.vals = round(brks), 
            longdims = c(0.2, 0.8), shortdims = c(0.06, 0.01), 
            colvec = cols(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0.5, -0.5)), 
            cex.val = cx, textcol = lcol, bordercol = lcol)
dev.off()
```
The map below was created by the code above, showing biases at 1 km, 25 km, 50 km, 
and 100 km from South Africa's landcover product, the mean estimates from 
GlobCover and MODIS, and Geowiki/GLC-SHARE.

<p align="center"><img src="../external/figures/bias_map.pdf" width="400px"></p>

[[back to top]][Plotting bias figures]

### Bias in relation to field density

```{r, eval = FALSE}
btype <- c("bias_act.pdf", "bias_abs.pdf")
# plot matrix
nc <- 4
m <- matrix(rep(c(rep(1, 8), rep(2, 8), rep(3, 8), rep(4, 8)), 2), nrow = 6, ncol = 32, byrow = TRUE) 
m2 <- unlist(lapply(seq(5, by = 1, length = 8 * 4), function(x) rep(x, nc)))
m3 <- do.call(rbind, lapply(seq(1, length(m2), by = 8 * nc), function(x) {
  do.call(rbind, rep(list(m2[x:(x + 8 * nc - 1)]), nc))
})) 
cols <- c("red", "orange3", "green4", "blue", "yellow")
bcol <- c("grey50", "grey50", "grey50", "grey")
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
levnms <- paste(gsub("f", "", lev), "km")
fcol <- "grey"
lev <- names(dstats_all$act$sa30)
lvec <- seq(5, 100, 5)
pp <- list("x" = c(-2, 21), "y" = c(0, 0), 
           "x21" = list(c(0, 9), c(9, 21)),
           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), seq(10, 100, 10)), 
           "y1lwd" = list(2, 1),  
           "y2lwd" = 0.5, #c(1, 0.5, 1, 0.5, 0.5, 0.5, 0.5, 1, 0.5, 1),
           "y2lc" = list(c("black", rep("grey80", 4), "black", rep("grey80", 4), 
                           rep("grey80", 4), "black", rep("grey80", 4), "black"),
                         rep("grey80", 10)),
           "y3lwd" = rep(2, length(lev)), 
           "yl" = list(c(-100, 100), c(0, 100)), "yl2" = list(c(-40, 40), c(0, 60)), 
           "xl" = c(1, 20),
           "xaxl" = unlist(lapply(seq(5, 100, 10), function(x) c(x, " "))),
           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), " "),
           "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
           "yax2" = list(unlist(lapply(seq(-40, 40, 20), function(x) c(" ", x)))[-1], 
                         unlist(lapply(seq(0, 60, 20), function(x) c(" ", x)))[-1]),
           "yaxp" = list(-100, 0), 
           "ycol" = c("grey70"))  # plot parameters

for(bt in 1:2) {
  pdf(full_path(p_fig, btype[bt]), width = 7, height = 7)
  lmat <- rbind(m, m3)
  l <- layout(lmat)
  par(mar = c(3, 0, 0, 0), oma = c(3, 4, 2, 2))
  snms <- names(dstats_all[[bt]])[names(dstats_all[[bt]]) != "glcsh"]
  dset <- dstats_all[[bt]][snms]
  for(i in 1:length(snms)) {
    #bt <- 1
    d <- dset[[snms[i]]]$f1
    n <- nrow(d)
    plot(pp$x, pp$y, ylim = pp$yl[[bt]], xlim = pp$xl, type = "l", 
         lwd = pp$y1lwd[[bt]], axes = FALSE, 
         xlab = "", ylab = "")
    if(i == 1) axis(2, at = pp$yax1[[bt]], labels = pp$yax1[[bt]], las = 2, 
                    mgp = c(1, 0.6, 0))
    axis(1, at = 1:20, labels = pp$xaxl, las = 2, mgp = c(1, 0.4, 0), tcl = -0.3, 
         pos = pp$yaxp[[bt]], cex = 0.8)
    for(z in 1:length(pp$y2[[bt]])) lines(pp$x, rep(pp$y2[[bt]][z], 2), 
                                          lwd = pp$y2lwd, col = pp$y2lc[[bt]][z])
    for(k in 1:n) {
      a <- seq(grep("2.5", colnames(d)), by = 1, length.out = 6)
      if(!is.na(d[k, 2])) boxplot_v(k, d[k, a], n, 150, 10, bcol, fcol, pcex = 0.5, 
                                    lwd = c(1, 0.5, 2))
    }
    lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = 3)
    if(i == 1) mtext(ifelse(bt == 1, "% bias", "% bias (absolute)"), side = 2, 
                     line = 2.2, cex = 0.8)
    mtext(lcnms[i], side = 3, line = 0, col = cols[i])
  }
  mtext("% cropland", side = 1, outer = TRUE, line = -36.7, cex = 0.8)
  par(mar = c(0.5, 0, 0.5, 0.3))
  for(i in 1:length(snms)) {
    highcounter <- rep(0, length(lev))
    for(j in 1:length(lev)) {
      plot(pp$x, pp$y, ylim = pp$yl2[[bt]], xlim = pp$xl, type = "l", lwd = 1, 
           axes = FALSE, xlab = "", ylab = "", xaxs = "i")
      for(z in 1:length(pp$y2[[bt]])) {
        lines(pp$x, rep(pp$y2[[bt]][z], 2), lwd = pp$y2lwd, col = pp$y2lc[[bt]][z])
      }
      d <- dset[[i]][[lev[j]]]
      n <- nrow(d)
      lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = pp$y3lwd[j])
      muv <- round(mean(d[, 8], na.rm = TRUE))
      if(j == length(lev)) {
        mtext(expression(mu), side = 1, cex = 0.8, col = "grey10", 
              line = ifelse(bt == 1, -1.5, -1.8), adj = 1.1)
      }
      text(10, pp$yl2[[bt]][1] + 10, labels = muv, cex = 1.3, col = "grey10")
      toohigh <- which(d[, "mu"] > pp$yl2[[bt]][2])
      if(length(toohigh) > 0) {
        highcounter[j] <- 1
        xi <- toohigh[which.max(d[toohigh, "mu"])]
        text(xi, y = pp$yl2[[bt]][2], labels = round(d[xi, "mu"]), col = "grey50", 
             xpd = NA)
      }
      if(j == length(lev)) {
        if(sum(highcounter) > 0) mtext("max", side = 3, cex = 0.7, col = "grey50", 
                                       line = -0.7, adj = 1.1)
      }
      if(j == 1) {
        axis(2, at = seq(pp$yl2[[bt]][1], pp$yl2[[bt]][2], 10), 
             labels = pp$yax2[[bt]], las = 2, mgp = c(1, 0.6, 0))
      }
      if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 0.8)
      axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
           tcl = -0.7, lwd.ticks = 1.2, 
           pos = pp$yl2[[bt]][1])
      axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
           tcl = -0.3, pos = pp$yl2[[bt]][1])
      if(i == length(snms)) {
        axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
             tcl = -0.3, pos = pp$yl2[[bt]][1])
      }
    }
  }
  mtext(ifelse(bt == 1, "% bias", "% bias (absolute)"), side = 2, line = 2.2, 
        cex = 0.8, outer = TRUE, adj = 0.35)
  mtext("% cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
  dev.off()
}

#   for(j in 1:6) { #j <- 1
#     plot(pp$x, pp$y, ylim = pp$yl, xlim = pp$xl, type = "l", lwd = 2, axes = FALSE, xlab = "", ylab = "")
#     for(z in 1:length(pp$y2)) {
#       for(q in pp$x21) lines(q, rep(pp$y2[z], 2), lwd = pp$y2lwd[z], col = pp$ycol)
#     }
#     if(i == 1) axis(2, at = seq(-100, 100, 50), labels = seq(-100, 100, 50), las = 2)
#     #if(i != 1) axis(2, at = seq(-100, 100, 50), labels = rep("", 5))
#     if(j == 6) axis(1, at = 1:20, labels = lvec, las = 2, mgp = c(1, 0.7, 0), tcl = -0.3)
#     df <- d[[j]]
#     for(k in 2:n) {
#       a <- seq(grep("2.5", colnames(df)), by = 1, length.out = 6)
#       if(!is.na(df[k, 2])) boxplot_v(k, df[k, a], n, 150, 10, bcol, fcol, pcex = 0.5, lwd = c(1, 0.5, 2))
#       if(!is.na(df[k, 7])) lines(1:19, df[-20, "mu"], pch = 20, col = cols[i], lwd = 1.5)
#     }
#     print(mean(df[, 8], na.rm = TRUE))
#   }
# }
# dev.off()

```

### 2015-06-02 Rmd/overview.Rmd
Moved entire folder to Dropbox/publications. Renamed vignettes folder to Rmd; committed some changes to Rd files that were lingering.

---
title: "Analysis of cropland biases"
author: "Lyndon Estes"
date: "15 November 2014"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes
---

# Introduction

## Sidebar
I am just learning how to use Rmarkdown with Rmd files to document analyses, primarily analyses that are more interactive in nature and notrepeated many times. Code that will be re-used multiple times will go into a library.

One things I am noticing is that the knit process for compiling the notebook does not find functions within code chunks unless the libraries are called within the code chunks themselves. It seems like each code chunk is its own environment.  The solution to this is to do something like `raster::raster()` when you want a function, or run this where chunks are set to {r, eval = FALSE}, and just run the code chunk as if you were in a normal R script working interactively.  I have already set up the library for SAcropland to call the packages its needs so they are loaded up in the global environment.

### 2015-06-02 Rmd/build-reports.R

@@ -1,0 +1,24 @@
+# build-reports.R
+# Builds output Rmd files
+
+library(lmisc)
+require(knitr)
+require(markdown)
+
+p_root <- proj_root("SAcropland")
+p_vignettes <- full_path(p_root, "SAcropland/vignettes")
+
+# Create .md, .html, and .pdf files
+knit("vignettes/overview.Rmd", output = "vignettes/overview.md")
+markdownToHTML("vignettes/overview.md", "vignettes/overview.html", 
+               options=c("use_xhml"))
+
+knit("vignettes/compare-landcover.Rmd", 
+     output = "vignettes/compare-landcover.md")
+markdownToHTML("vignettes/compare-landcover.md", 
+               "vignettes/compare-landcover.html", options=c("use_xhml"))
+
+knit("vignettes/landcover-bias-plots.Rmd", 
+     output = "vignettes/landcover-bias-plots.md")
+markdownToHTML("vignettes/landcover-bias-plots.md", 
+               "vignettes/landcover-bias-plots.html", options=c("use_xhml"))

### 2015-06-02 vignettes/build-reports.R

@@ -1,20 +1,0 @@
-# build-reports.R
-# Builds output Rmd files
-
-library(lmisc)
-require(knitr)
-require(markdown)
-
-p_root <- proj_root("SAcropland")
-p_vignettes <- full_path(p_root, "SAcropland/vignettes")
-
-# Create .md, .html, and .pdf files
-knit("vignettes/overview.Rmd", output = "vignettes/overview.md")
-markdownToHTML("vignettes/overview.md", "vignettes/overview.html", options=c("use_xhml"))
-
-knit("vignettes/compare-landcover.Rmd", output = "vignettes/compare-landcover.md")
-markdownToHTML("vignettes/compare-landcover.md", "vignettes/compare-landcover.html", options=c("use_xhml"))
-
-knit("vignettes/landcover-bias-plots.Rmd", output = "vignettes/landcover-bias-plots.md")
-markdownToHTML("vignettes/landcover-bias-plots.md", "vignettes/landcover-bias-plots.html", 
-               options=c("use_xhml"))

### 2015-06-04 Rmd/compare-landcover.Rmd
1. New version of figures, with boxplot version of 1 km pushed to separate figure for supplemental. 2. Pushed paper related material (figures, etc) to new paper folder. 2. Tested out polynomial regressions in biases in compare-landcover.Rmd

@@
 globsa_list_sum <- lapply(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
-modsa_list_sum <- lapply(full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
+fnames <- full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)])
+modsa_list_sum <- lapply(fnames, raster)
 glcsa <- raster(full_path(p_data, "glcsa_masked.tif"))
@@
 lclist <- c(salc, globsa_list_sum, modsa_list_sum, glcsa, geow)  # into list
-names(lclist) <- c("sa30", "globmin", "globmu", "globmax", "modmin", "modmu", "modmax", "glcsh", "geow")
+names(lclist) <- c("sa30", "globmin", "globmu", "globmax", "modmin", "modmu",
+                   "modmax", "glcsh", "geow")
 ```
@@
 #fact <- c(5, 10, 25, 50, 100, 200, 300)
-fact <- 25
+#fact <- 25
 lc_agg <- aggregate_rast_list(fact, lclist)   # landcover rasters 
@@
 [[back to top]][Overview]
+
 ### Bias as function of cropcover
@@
 namask[namask == 0] <- NA
-writeRaster(namask, filename = full_path(p_data, "namask.tif"), overwrite = TRUE)
+writeRaster(namask, filename = full_path(p_data, "namask.tif"), 
+            overwrite = TRUE)
 
@@
 binv <- seq(0, 100, 5)
-gti_bins <- lapply(gti_agg, function(x) lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE)))
+gti_bins <- lapply(gti_agg, function(x) {
+  lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE))
+})
 ```
@@
 
++ Collected absolute differences with weights for regression plots. Test first on 25 km
+```{r, eval=FALSE}
+gti_2011_25 <- raster(full_path(p_data, "gti_2011_25km.tif"))
+snms2 <- c("sa30", "modmu", "globmu", "geow")
+km25 <- stack(lapply(dlist_abs[snms2], function(x) x$f25$g2011))
+s <- stack(list(area_wgts$f25[[1]], gti_2011_25, km25))
+# s <- s[[-10]]
+library(agroEcoTradeoffs)
+DT <- as.data.table.raster(s)
+DT <- DT[!is.na(namask), ]
+setnames(DT, "gti_2011_25km", "A")
+DTL <- do.call(rbind, lapply(3:ncol(DT), function(x) {
+  DTS <- DT[, c(1, 2, x), with = FALSE]
+  setnames(DTS, colnames(DT)[x], new = "D")
+}))
+head(DTL)
+DTL[, plot(A, D, pch = 20, cex = 0.1)]
+dlm <- lm(D ~ poly(A, degree = 2), weights = namask, data = DTL)
+lines(predict.lm(dlm, newdata = data.frame("A" = 1:100)))
+
+DT[, plot(A, geow, pch = 20, cex = 0.1)]
+dlm <- lm(geow ~ poly(A, degree = 2), data = DT)
+lines(predict.lm(dlm, newdata = data.frame("A" = 1:100)))
+
+```
+
 + Finally, differences between gti2007 and 2011
@@
       } else if((nrow(oof) > 0) & (nrow(oof) < 10)) {
-        qs <- c(rep(NA, 6), Hmisc::wtd.mean(oof[, 2], weights = oof[, 3], na.rm = FALSE))
+        qs <- c(rep(NA, 6), 
+                Hmisc::wtd.mean(oof[, 2], weights = oof[, 3], na.rm = FALSE))
       } else {

### 2015-06-04 Rmd/landcover-bias-plots.Rmd
1. New version of figures, with boxplot version of 1 km pushed to separate figure for supplemental. 2. Pushed paper related material (figures, etc) to new paper folder. 2. Tested out polynomial regressions in biases in compare-landcover.Rmd

@@
 p_root <- proj_root("SAcropland")
-p_fig <- full_path(p_root, "SAcropland/external/figures/")
+p_fig <- full_path(p_root, "SAcropland/paper/figures/")
 p_data <- full_path(p_root, "SAcropland/external/ext_data/")
@@
 namask <- raster(full_path(p_data, "namask.tif"))
-snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data, excluding Modis and GlobCov extremes
+snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
 lev <- c("f1", "f25", "f50", "f100")  # just show four levels
@@
     } else {
-      r <- disaggregate(dlist_act[[x]][[y]]$g2011, fact = as.numeric(gsub("f", "", y)))
+      r <- disaggregate(dlist_act[[x]][[y]]$g2011, 
+                        fact = as.numeric(gsub("f", "", y)))
       r <- raster::mask(crop(r, namask), namask)
@@
 ```
-The map below was created by the code above, showing biases at 1 km, 25 km, 50 km, 
-and 100 km from South Africa's landcover product, the mean estimates from 
+The map below was created by the code above, showing biases at 1 km, 25 km, 50 km, and 100 km from South Africa's landcover product, the mean estimates from 
 GlobCover and MODIS, and Geowiki/GLC-SHARE.
 
-<p align="center"><img src="../external/figures/bias_map.pdf" width="400px"></p>
+<p align="center"><img src="../paper/figures/bias_map.pdf" width="400px"></p>
 
@@
 ### Bias in relation to field density
+1. Supplementary figures - bias plots at 1 km with boxplots.
 
 ```{r, eval = FALSE}
-btype <- c("bias_act.pdf", "bias_abs.pdf")
+#btype <- c("bias_act.pdf", "bias_abs.pdf")
 # plot matrix
 nc <- 4
-m <- matrix(rep(c(rep(1, 8), rep(2, 8), rep(3, 8), rep(4, 8)), 2), nrow = 6, ncol = 32, byrow = TRUE) 
-m2 <- unlist(lapply(seq(5, by = 1, length = 8 * 4), function(x) rep(x, nc)))
-m3 <- do.call(rbind, lapply(seq(1, length(m2), by = 8 * nc), function(x) {
-  do.call(rbind, rep(list(m2[x:(x + 8 * nc - 1)]), nc))
-})) 
+m <- matrix(rep(c(rep(1, 8), rep(2, 8), rep(3, 8), rep(4, 8)), 2), nrow = 6, 
+            ncol = 32, byrow = TRUE) 
 cols <- c("red", "orange3", "green4", "blue", "yellow")
@@
 lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
-levnms <- paste(gsub("f", "", lev), "km")
 fcol <- "grey"
@@
            "x21" = list(c(0, 9), c(9, 21)),
-           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), seq(10, 100, 10)), 
+           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), 
+                       seq(10, 100, 10)), 
            "y1lwd" = list(2, 1),  
@@
            "y2lc" = list(c("black", rep("grey80", 4), "black", rep("grey80", 4), 
-                           rep("grey80", 4), "black", rep("grey80", 4), "black"),
+                           rep("grey80", 4), "black", rep("grey80", 4),"black"),
                          rep("grey80", 10)),
-           "y3lwd" = rep(2, length(lev)), 
-           "yl" = list(c(-100, 100), c(0, 100)), "yl2" = list(c(-40, 40), c(0, 60)), 
-           "xl" = c(1, 20),
+           "yl" = list(c(-100, 100), c(0, 100)), 
+           #"yl2" = list(c(-40, 40), c(0, 60)), "xl" = c(1, 20),
+           "yl2" = c(-20, 60), "xl" = c(1, 20),
            "xaxl" = unlist(lapply(seq(5, 100, 10), function(x) c(x, " "))),
-           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), " "),
+           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), 
+                            " "),
            "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
-           "yax2" = list(unlist(lapply(seq(-40, 40, 20), function(x) c(" ", x)))[-1], 
-                         unlist(lapply(seq(0, 60, 20), function(x) c(" ", x)))[-1]),
-           "yaxp" = list(-100, 0), 
+           #"yax2" = list(
+            # unlist(lapply(seq(-40, 40, 20), function(x) c(" ", x)))[-1], 
+            # unlist(lapply(seq(0, 60, 20), function(x) c(" ", x)))[-1]),
+           #"yaxp" = list(-100, 0), 
            "ycol" = c("grey70"))  # plot parameters
 
+pdf(full_path(p_fig, "biases_1km.pdf"), width = 7, height = 5)
+lmat <- rbind(m, m + max(m))
+l <- layout(lmat)
+par(mar = c(3, 0, 0, 0), oma = c(1, 4, 2, 2))
 for(bt in 1:2) {
-  pdf(full_path(p_fig, btype[bt]), width = 7, height = 7)
-  lmat <- rbind(m, m3)
-  l <- layout(lmat)
-  par(mar = c(3, 0, 0, 0), oma = c(3, 4, 2, 2))
   snms <- names(dstats_all[[bt]])[names(dstats_all[[bt]]) != "glcsh"]
   dset <- dstats_all[[bt]][snms]
-  for(i in 1:length(snms)) {
-    #bt <- 1
+  for(i in 1:length(snms)) { i
     d <- dset[[snms[i]]]$f1
@@
                     mgp = c(1, 0.6, 0))
-    axis(1, at = 1:20, labels = pp$xaxl, las = 2, mgp = c(1, 0.4, 0), tcl = -0.3, 
-         pos = pp$yaxp[[bt]], cex = 0.8)
-    for(z in 1:length(pp$y2[[bt]])) lines(pp$x, rep(pp$y2[[bt]][z], 2), 
-                                          lwd = pp$y2lwd, col = pp$y2lc[[bt]][z])
+    axis(1, at = 1:20, labels = pp$xaxl, las = 2, mgp = c(1, 0.4, 0), 
+         tcl = -0.3, pos = pp$yaxp[[bt]], cex = 0.8)
+    for(z in 1:length(pp$y2[[bt]])) {
+      lines(pp$x, rep(pp$y2[[bt]][z], 2), lwd =pp$y2lwd, col = pp$y2lc[[bt]][z])
+    }
     for(k in 1:n) {
       a <- seq(grep("2.5", colnames(d)), by = 1, length.out = 6)
-      if(!is.na(d[k, 2])) boxplot_v(k, d[k, a], n, 150, 10, bcol, fcol, pcex = 0.5, 
+      if(!is.na(d[k, 2])) {
+        boxplot_v(k, d[k, a], n, 150, 10, bcol, fcol, pcex = 0.5, 
                   lwd = c(1, 0.5, 2))
+      }
     }
@@
                      line = 2.2, cex = 0.8)
-    mtext(lcnms[i], side = 3, line = 0, col = cols[i])
+    if(bt == 1) mtext(lcnms[i], side = 3, line = 0, col = cols[i])
   }
-  mtext("% cropland", side = 1, outer = TRUE, line = -36.7, cex = 0.8)
-  par(mar = c(0.5, 0, 0.5, 0.3))
+}
+mtext("% cropland", side = 1, outer = TRUE, line = -1, cex = 0.8)
+dev.off()
+```
+
+2. Bias with scale
+```{r, eval=FALSE}
+lev <- lev[!lev %in% c("f5", "f300")]
+levnms <- paste(gsub("f", "", lev), "km")
+m2 <- unlist(lapply(seq(1, by = 1, length = length(lev) * 4), function(x) {
+  rep(x, nc)
+}))
+m3 <- do.call(rbind, lapply(seq(1, length(m2), length(lev) * nc), function(x) {
+  do.call(rbind, rep(list(m2[x:(x + length(lev) * nc - 1)]), nc))
+}))
+
+yl <- seq(pp$yl2[1], pp$yl2[2], 10)
+ylcol <- c("black", "grey", "black", rep("grey", 5), "black")
+lt <- c(5, 1)
+ylwd <- c(1.5, 2)
+
+pdf(full_path(p_fig, "biases_1-200km.pdf"), width = 7.5, height = 7)
+l <- layout(m3)
+# i <- 1; j <- 1
+par(mar = c(0.5, 0, 1, 0.3), oma = c(5, 4, 2, 2))
 for(i in 1:length(snms)) {
@@
   for(j in 1:length(lev)) {
-      plot(pp$x, pp$y, ylim = pp$yl2[[bt]], xlim = pp$xl, type = "l", lwd = 1, 
+    plot(pp$x, pp$y, ylim = pp$yl2, xlim = pp$xl, type = "l", lwd = 1, 
          axes = FALSE, xlab = "", ylab = "", xaxs = "i")
-      for(z in 1:length(pp$y2[[bt]])) {
-        lines(pp$x, rep(pp$y2[[bt]][z], 2), lwd = pp$y2lwd, col = pp$y2lc[[bt]][z])
+    for(z in 1:length(yl)) {
+      lines(pp$x, rep(yl[z], 2), lwd = pp$y2lwd, col = ylcol[z])
     }
-      d <- dset[[i]][[lev[j]]]
+    muv <- rep(NA, 2)
+    for(bt in 1:2) {
+      d <- dstats_all[[bt]][[i]][[j]]
       n <- nrow(d)
-      lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = pp$y3lwd[j])
-      muv <- round(mean(d[, 8], na.rm = TRUE))
-      if(j == length(lev)) {
-        mtext(expression(mu), side = 1, cex = 0.8, col = "grey10", 
-              line = ifelse(bt == 1, -1.5, -1.8), adj = 1.1)
-      }
-      text(10, pp$yl2[[bt]][1] + 10, labels = muv, cex = 1.3, col = "grey10")
-      toohigh <- which(d[, "mu"] > pp$yl2[[bt]][2])
+      lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = ylwd[bt], 
+            lty = lt[bt])
+      muv[bt] <- round(mean(d[, 8], na.rm = TRUE))
+      toohigh <- which(d[, "mu"] > pp$yl2[2])
       if(length(toohigh) > 0) {
@@
         xi <- toohigh[which.max(d[toohigh, "mu"])]
-        text(xi, y = pp$yl2[[bt]][2], labels = round(d[xi, "mu"]), col = "grey50", 
-             xpd = NA)
+        text(xi, y = pp$yl2[2] + 4, labels = round(d[xi, "mu"]), cex = 1.1,
+             col = "grey50", xpd = NA)
       }
-      if(j == length(lev)) {
-        if(sum(highcounter) > 0) mtext("max", side = 3, cex = 0.7, col = "grey50", 
-                                       line = -0.7, adj = 1.1)
     }
-      if(j == 1) {
-        axis(2, at = seq(pp$yl2[[bt]][1], pp$yl2[[bt]][2], 10), 
-             labels = pp$yax2[[bt]], las = 2, mgp = c(1, 0.6, 0))
-      }
-      if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 0.8)
+    text(5, pp$yl2[2] - 14, labels = paste(muv, collapse = " / "), cex = 1.1, 
+         col = "grey10")
+    if(j == 1) axis(2, at = yl, labels = yl, las = 2, mgp = c(1, 0.6, 0))
+    if(j == 1) mtext(lcnms[i], side = 2, line = 2.8, cex = 1, col = cols[i])
+    if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 1.2)
     axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
-           tcl = -0.7, lwd.ticks = 1.2, 
-           pos = pp$yl2[[bt]][1])
+         tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[1])
     axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
-           tcl = -0.3, pos = pp$yl2[[bt]][1])
+         tcl = -0.3, pos = pp$yl2[1])
     if(i == length(snms)) {
       axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
-             tcl = -0.3, pos = pp$yl2[[bt]][1])
+           tcl = -0.3, pos = pp$yl2[1])
     }
@@
 }
-  mtext(ifelse(bt == 1, "% bias", "% bias (absolute)"), side = 2, line = 2.2, 
-        cex = 0.8, outer = TRUE, adj = 0.35)
+mtext("% bias", side = 2, line = 2.1, cex = 0.8, outer = TRUE, adj = 0.5)
 mtext("% cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
+par(xpd = NA)
+x <- grconvertX(0.75, from = "ndc", to = "user")
+y <- grconvertY(0.07, from = "ndc", to = "user")
+legend(x = x, y = y, legend = c("actual bias", "absolute bias"), lty = lt, 
+       lwd = ylwd, bty = "n", cex = 1.2)
 dev.off()
-}
+
+# DF <- do.call(rbind.data.frame, lapply(snms[-2], function(x) {
+#   dstats_all[[2]][[x]]$f25[, c(1, 8)]
+# }))
+# dlm <- lm(mu ~ poly(V1, degree = 2), data = DF)
+# plot(DF[, 1], DF[, 2])
+# lines(predict.lm(dlm, newdata = data.frame("V1" = 1:100)))
+
+# for(bt in 1:2) {
+#   pdf(full_path(p_fig, btype[bt]), width = 7, height = 7)
+#   lmat <- rbind(m, m3)
+#   l <- layout(lmat)
+#   par(mar = c(3, 0, 0, 0), oma = c(3, 4, 2, 2))
+#   snms <- names(dstats_all[[bt]])[names(dstats_all[[bt]]) != "glcsh"]
+#   dset <- dstats_all[[bt]][snms]
+#   for(i in 1:length(snms)) { i
+#     #bt <- 1
+#     d <- dset[[snms[i]]]$f1
+#     n <- nrow(d)
+#     plot(pp$x, pp$y, ylim = pp$yl[[bt]], xlim = pp$xl, type = "l", 
+#          lwd = pp$y1lwd[[bt]], axes = FALSE, 
+#          xlab = "", ylab = "")
+#     if(i == 1) axis(2, at = pp$yax1[[bt]], labels = pp$yax1[[bt]], las = 2, 
+#                     mgp = c(1, 0.6, 0))
+#     axis(1, at = 1:20, labels = pp$xaxl, las = 2, mgp = c(1, 0.4, 0), 
+#          tcl = -0.3, pos = pp$yaxp[[bt]], cex = 0.8)
+#     for(z in 1:length(pp$y2[[bt]])) {
+#       lines(pp$x, rep(pp$y2[[bt]][z], 2), lwd =pp$y2lwd, col = pp$y2lc[[bt]][z])
+#     }
+#     for(k in 1:n) {
+#       a <- seq(grep("2.5", colnames(d)), by = 1, length.out = 6)
+#       if(!is.na(d[k, 2])) {
+#         boxplot_v(k, d[k, a], n, 150, 10, bcol, fcol, pcex = 0.5, 
+#                   lwd = c(1, 0.5, 2))
+#       }
+#     }
+#     lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = 3)
+#     if(i == 1) mtext(ifelse(bt == 1, "% bias", "% bias (absolute)"), side = 2, 
+#                      line = 2.2, cex = 0.8)
+#     mtext(lcnms[i], side = 3, line = 0, col = cols[i])
+#   }
+#   mtext("% cropland", side = 1, outer = TRUE, line = -36.7, cex = 0.8)
+#   par(mar = c(0.5, 0, 0.5, 0.3))
+#   for(i in 1:length(snms)) {
+#     highcounter <- rep(0, length(lev))
+#     for(j in 1:length(lev)) {
+#       plot(pp$x, pp$y, ylim = pp$yl2[[bt]], xlim = pp$xl, type = "l", lwd = 1, 
+#            axes = FALSE, xlab = "", ylab = "", xaxs = "i")
+#       for(z in 1:length(pp$y2[[bt]])) {
+#         lines(pp$x, rep(pp$y2[[bt]][z], 2), lwd = pp$y2lwd, 
+#               col = pp$y2lc[[bt]][z])
+#       }
+#       d <- dset[[i]][[lev[j]]]
+#       n <- nrow(d)
+#       lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = pp$y3lwd[j])
+#       muv <- round(mean(d[, 8], na.rm = TRUE))
+#       if(j == length(lev)) {
+#         mtext(expression(mu), side = 1, cex = 0.8, col = "grey10", 
+#               line = ifelse(bt == 1, -1.5, -1.8), adj = 1.1)
+#       }
+#       text(10, pp$yl2[[bt]][1] + 10, labels = muv, cex = 1.3, col = "grey10")
+#       toohigh <- which(d[, "mu"] > pp$yl2[[bt]][2])
+#       if(length(toohigh) > 0) {
+#         highcounter[j] <- 1
+#         xi <- toohigh[which.max(d[toohigh, "mu"])]
+#         text(xi, y = pp$yl2[[bt]][2], labels = round(d[xi, "mu"]), 
+#              col = "grey50", xpd = NA)
+#       }
+#       if(j == length(lev)) {
+#         if(sum(highcounter) > 0) {
+#           mtext("max", side = 3, cex = 0.7, col = "grey50", line = -0.7, 
+#                 adj = 1.1)
+#         }
+#       }
+#       if(j == 1) {
+#         axis(2, at = seq(pp$yl2[[bt]][1], pp$yl2[[bt]][2], 10), 
+#              labels = pp$yax2[[bt]], las = 2, mgp = c(1, 0.6, 0))
+#       }
+#       if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 0.8)
+#       axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
+#            tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[[bt]][1])
+#       axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
+#            tcl = -0.3, pos = pp$yl2[[bt]][1])
+#       if(i == length(snms)) {
+#         axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
+#              tcl = -0.3, pos = pp$yl2[[bt]][1])
+#       }
+#     }
+#   }
+#   mtext(ifelse(bt == 1, "% bias", "% bias (absolute)"), side = 2, line = 2.2, 
+#         cex = 0.8, outer = TRUE, adj = 0.35)
+#   mtext("% cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
+#   dev.off()
+# }
 

### 2015-06-24 Rmd/compare-landcover.Rmd
Many changes made, some revisions to existing Rmds, but main work done was to create magisterial-district-bias.Rmd and associated figures.

@@
 salc <- raster(full_path(p_data, "sa_ag_masked.tif")) * 100
-globsa_list_sum <- lapply(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
+ds <- full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)])
+globsa_list_sum <- lapply(ds, raster)
 fnames <- full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)])

### 2015-06-24 Rmd/landcover-bias-plots.Rmd
Many changes made, some revisions to existing Rmds, but main work done was to create magisterial-district-bias.Rmd and associated figures.

@@
 
-# DF <- do.call(rbind.data.frame, lapply(snms[-2], function(x) {
-#   dstats_all[[2]][[x]]$f25[, c(1, 8)]
-# }))
-# dlm <- lm(mu ~ poly(V1, degree = 2), data = DF)
-# plot(DF[, 1], DF[, 2])
-# lines(predict.lm(dlm, newdata = data.frame("V1" = 1:100)))
-
-# for(bt in 1:2) {
-#   pdf(full_path(p_fig, btype[bt]), width = 7, height = 7)
-#   lmat <- rbind(m, m3)
-#   l <- layout(lmat)
-#   par(mar = c(3, 0, 0, 0), oma = c(3, 4, 2, 2))
-#   snms <- names(dstats_all[[bt]])[names(dstats_all[[bt]]) != "glcsh"]
-#   dset <- dstats_all[[bt]][snms]
-#   for(i in 1:length(snms)) { i
-#     #bt <- 1
-#     d <- dset[[snms[i]]]$f1
-#     n <- nrow(d)
-#     plot(pp$x, pp$y, ylim = pp$yl[[bt]], xlim = pp$xl, type = "l", 
-#          lwd = pp$y1lwd[[bt]], axes = FALSE, 
-#          xlab = "", ylab = "")
-#     if(i == 1) axis(2, at = pp$yax1[[bt]], labels = pp$yax1[[bt]], las = 2, 
-#                     mgp = c(1, 0.6, 0))
-#     axis(1, at = 1:20, labels = pp$xaxl, las = 2, mgp = c(1, 0.4, 0), 
-#          tcl = -0.3, pos = pp$yaxp[[bt]], cex = 0.8)
-#     for(z in 1:length(pp$y2[[bt]])) {
-#       lines(pp$x, rep(pp$y2[[bt]][z], 2), lwd =pp$y2lwd, col = pp$y2lc[[bt]][z])
-#     }
-#     for(k in 1:n) {
-#       a <- seq(grep("2.5", colnames(d)), by = 1, length.out = 6)
-#       if(!is.na(d[k, 2])) {
-#         boxplot_v(k, d[k, a], n, 150, 10, bcol, fcol, pcex = 0.5, 
-#                   lwd = c(1, 0.5, 2))
-#       }
-#     }
-#     lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = 3)
-#     if(i == 1) mtext(ifelse(bt == 1, "% bias", "% bias (absolute)"), side = 2, 
-#                      line = 2.2, cex = 0.8)
-#     mtext(lcnms[i], side = 3, line = 0, col = cols[i])
-#   }
-#   mtext("% cropland", side = 1, outer = TRUE, line = -36.7, cex = 0.8)
-#   par(mar = c(0.5, 0, 0.5, 0.3))
-#   for(i in 1:length(snms)) {
-#     highcounter <- rep(0, length(lev))
-#     for(j in 1:length(lev)) {
-#       plot(pp$x, pp$y, ylim = pp$yl2[[bt]], xlim = pp$xl, type = "l", lwd = 1, 
-#            axes = FALSE, xlab = "", ylab = "", xaxs = "i")
-#       for(z in 1:length(pp$y2[[bt]])) {
-#         lines(pp$x, rep(pp$y2[[bt]][z], 2), lwd = pp$y2lwd, 
-#               col = pp$y2lc[[bt]][z])
-#       }
-#       d <- dset[[i]][[lev[j]]]
-#       n <- nrow(d)
-#       lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = pp$y3lwd[j])
-#       muv <- round(mean(d[, 8], na.rm = TRUE))
-#       if(j == length(lev)) {
-#         mtext(expression(mu), side = 1, cex = 0.8, col = "grey10", 
-#               line = ifelse(bt == 1, -1.5, -1.8), adj = 1.1)
-#       }
-#       text(10, pp$yl2[[bt]][1] + 10, labels = muv, cex = 1.3, col = "grey10")
-#       toohigh <- which(d[, "mu"] > pp$yl2[[bt]][2])
-#       if(length(toohigh) > 0) {
-#         highcounter[j] <- 1
-#         xi <- toohigh[which.max(d[toohigh, "mu"])]
-#         text(xi, y = pp$yl2[[bt]][2], labels = round(d[xi, "mu"]), 
-#              col = "grey50", xpd = NA)
-#       }
-#       if(j == length(lev)) {
-#         if(sum(highcounter) > 0) {
-#           mtext("max", side = 3, cex = 0.7, col = "grey50", line = -0.7, 
-#                 adj = 1.1)
-#         }
-#       }
-#       if(j == 1) {
-#         axis(2, at = seq(pp$yl2[[bt]][1], pp$yl2[[bt]][2], 10), 
-#              labels = pp$yax2[[bt]], las = 2, mgp = c(1, 0.6, 0))
-#       }
-#       if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 0.8)
-#       axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
-#            tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[[bt]][1])
-#       axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
-#            tcl = -0.3, pos = pp$yl2[[bt]][1])
-#       if(i == length(snms)) {
-#         axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
-#              tcl = -0.3, pos = pp$yl2[[bt]][1])
-#       }
-#     }
-#   }
-#   mtext(ifelse(bt == 1, "% bias", "% bias (absolute)"), side = 2, line = 2.2, 
-#         cex = 0.8, outer = TRUE, adj = 0.35)
-#   mtext("% cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
-#   dev.off()
-# }
-
-#   for(j in 1:6) { #j <- 1
-#     plot(pp$x, pp$y, ylim = pp$yl, xlim = pp$xl, type = "l", lwd = 2, axes = FALSE, xlab = "", ylab = "")
-#     for(z in 1:length(pp$y2)) {
-#       for(q in pp$x21) lines(q, rep(pp$y2[z], 2), lwd = pp$y2lwd[z], col = pp$ycol)
-#     }
-#     if(i == 1) axis(2, at = seq(-100, 100, 50), labels = seq(-100, 100, 50), las = 2)
-#     #if(i != 1) axis(2, at = seq(-100, 100, 50), labels = rep("", 5))
-#     if(j == 6) axis(1, at = 1:20, labels = lvec, las = 2, mgp = c(1, 0.7, 0), tcl = -0.3)
-#     df <- d[[j]]
-#     for(k in 2:n) {
-#       a <- seq(grep("2.5", colnames(df)), by = 1, length.out = 6)
-#       if(!is.na(df[k, 2])) boxplot_v(k, df[k, a], n, 150, 10, bcol, fcol, pcex = 0.5, lwd = c(1, 0.5, 2))
-#       if(!is.na(df[k, 7])) lines(1:19, df[-20, "mu"], pch = 20, col = cols[i], lwd = 1.5)
-#     }
-#     print(mean(df[, 8], na.rm = TRUE))
-#   }
-# }
-# dev.off()
-
+DF <- do.call(rbind.data.frame, lapply(snms[4], function(x) {
+  dstats_all[[2]][[x]]$f1[, c(1, 8)]
+}))
+dlm <- lm(mu ~ poly(V1, degree = 2), data = DF)
+plot(DF[, 1], DF[, 2])
+lines(predict.lm(dlm, newdata = data.frame("V1" = 1:110)))
 ```

### 2015-06-24 Rmd/magisterial-district-bias.Rmd
Many changes made, some revisions to existing Rmds, but main work done was to create magisterial-district-bias.Rmd and associated figures.

---
title: "District-level cropcover and bias"
author: "Lyndon Estes"
date: "June 4, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
---

## Bias in relation to typical crop cover

### Data
```{r, eval = FALSE}
library(raster)
library(lmisc)
library(nlme)
library(mgcv)
library(ape)
library(fitdistrplus)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")

# Load in datasets
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data2, "ZAF_adm2.Rdata"))  # magisterial districts
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
load(full_path(p_data, "d_grid_abs.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask*")))
gti <- raster(paths)  # gti 2011
gti@crs <- sa.shp@proj4string

# Transform magisterial districts, remove Prince Edward Islands
md <- spTransform(gadm, CRSobj = sa.shp@proj4string)
md <- md[md$ID_2 != 313, ]
round(mean(rgeos::gArea(md, byid = TRUE) / 1000000))
```

### Analyses

1. Extract the cropland percentages for magisterial district, as well as the absolute biases. 
```{r, eval=FALSE}
# GTI data
# mean cropland cover
gti_md <- extract(x = gti, y = md, progress = "text")
save(gti_md, file = "external/ext_data/gti_md.rda")
# load("external/ext_data/gti_md.rda")
gti_mu <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
}))

# mean no zeros
gti_mu0 <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x[x > 0], na.rm = TRUE), 1))
}))

# how many contribute when non-cropland removed
nonzero <- sapply(gti_md, function(x) length(x[!is.na(x) & (x > 0)]))

# median
gti_median <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(median(x, na.rm = TRUE), 1))
}))

gti_median0 <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), 
    round(median(x[x > 0], na.rm = TRUE), 1))
}))

par(mar = c(0, 0, 0, 0))
plot(md[323, ])
plot(gti, add = TRUE)
plot(md[323, ], add = TRUE)

# Absolute biases at 1 km
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_abs1km <- stack(lapply(dlist_abs[snms], function(x) x$f1$g2011))
dlist_abs1km@crs <- sa.shp@proj4string
dlist_abs1km_md <- extract(x = dlist_abs1km, y = md, progress = "text")
save(dlist_abs1km_md, file = "external/ext_data/dlist_abs1km_md.rda")
# load("external/ext_data/dlist_abs1km_md.rda")

# check counts, primarily of NA values for each dataset
dlist_counts <- t(sapply(dlist_abs1km_md, function(x) {
  apply(x, 2, function(y) {
    c(length(y), length(which(is.na(y))))#, round(mean(y, na.rm = TRUE), 1))
  })
}))

chkfun <- function(x, tol = 0.001) {
  return(apply(x, 1, function(x) abs(max(x) - min(x)) < tol))
}
a <- seq(1, 7, 2)
b <- seq(2, 8, 2)
dlist_counts[which(!chkfun(dlist_counts[, a])), a]
dlist_counts[which(!chkfun(dlist_counts[, b])), b]

# bind stats
# MD
# some have more NAs than others, so select which LC set has max NAs
nas <- apply(dlist_counts[, seq(2, 8, 2)], 1, max)
all(gti_mu[, 1] == dlist_counts[, 1])

# then calculate mean per LC set per district
# mean
dlist_mu <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
  round(colMeans(x, na.rm = TRUE), 1)
})))

# mean no zero
dlist_mu0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
  tab <- round(x)
  round(apply(tab, 2, function(x) mean(x[x > 0], na.rm = TRUE)), 1)
})))

# median
dlist_median <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
  round(matrixStats::colMedians(x, na.rm = TRUE), 1)
})))
colnames(dlist_median) <- c("nas", snms)

dlist_median0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
  tab <- round(x)
  round(apply(tab, 2, function(x) median(x[x > 0], na.rm = TRUE)), 1)
})))

# bind function
bind_stats <- function(x, y, coords = xy, 
                       snms = c("sa30", "globmu", "modmu", "geow")) {
  dat <- cbind.data.frame(x, y)
  dat$nd <- apply(dat[, c(2, 4)], 1, max)  # max of GTI and LC NAs
  colnames(dat)[c(1, 3)] <- c("N", "gti")
  dat <- dat[, -2]
  dat <- cbind(coords, dat)
  dat$dmu4 <- round(rowMeans(dat[, snms]), 1)  # mean across all 4
  dat$dmu3 <- round(rowMeans(dat[, snms[snms != "globmu"]]), 1) # mu - globcov
  dat
}

# centroids for districts
xy <- as.data.frame(rgeos::gCentroid(md, byid = TRUE))  # get centroid coords

# bind to gti for comparisons 
mu <- bind_stats(x = gti_mu, y = dlist_mu)
mu$w <- mu$N - mu$nd  # weight - N contributing pixels
mu0 <- bind_stats(x = gti_mu0, y = dlist_mu0)
mu0$w <- nonzero  # weight - N contributing pixels (non-zeros)
med <- bind_stats(x = gti_median, y = dlist_median)
med$w <- med$N - med$nd  # weight - N contributing pixels
med0 <- bind_stats(x = gti_median0, y = dlist_median0)
med0$w <- nonzero  # weight - N contributing pixels (non-zeros)
```

2. Run regressions and check for spatial autocorrelation
```{r, eval=FALSE}
# check the polygons and rows line up for coordinates
# i <- 300
# a <- extract(x = gti, y = md[i, ], progress = "text")
# length(a[[1]])
# length(which(is.na(a[[1]])))
# mean(a[[1]], na.rm = TRUE)
# mu[i, ]

proc_df <- function(x, nm) {
  nms <- which(colnames(x) %in% c("x", "y", "gti", nm, "w"))
  dat <- x[, nms]
  colnames(dat)[3:4] <- c("pv", "dv")
  return(dat)
}

nms <- colnames(mu)[match(c(snms, "dmu4", "dmu3"), colnames(mu))]
lm_mu <- lapply(nms, function(x) {
  dat <- proc_df(x = mu, nm = x)
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})

lm_mu0 <- lapply(nms, function(x) {
  dat <- proc_df(x = mu0, nm = x)
  dat <- dat[!is.na(dat$pv), ]
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})

lm_med <- lapply(nms, function(x) {
  dat <- proc_df(x = med, nm = x)
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})

lm_med0 <- lapply(nms, function(x) {
  dat <- proc_df(x = med0, nm = x)
  dat <- dat[!is.na(dat$pv), ]
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})
```

Plot the fitted LM models
```{r, eval=FALSE}
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
cols <- c("red", "orange3", "green4", "blue", "black", "grey50")
lw <- c(rep(1, 4), 3, 3)
lt <- c(rep(2, 4), 1, 1)
xl <- c(0, 100)
yl <- c(0, round(max(mu[, snms] / 10 )) * 10)

# mean cropland in districts
pdf(full_path(p_fig, "biases_magdist.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- mu[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_mu)) {
  dlm <- lm_mu[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()

# mean non-zero cropland
yl <- c(0, round(max(mu0[, snms] / 10, na.rm = TRUE)) * 10)
pdf(full_path(p_fig, "biases_magdist_no0.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- mu0[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_mu0)) {
  dlm <- lm_mu0[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()

# median cropland
yl <- c(0, round(max(med[, snms] / 10, na.rm = TRUE)) * 10)
pdf(full_path(p_fig, "biases_magdist_med.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- med[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_med)) {
  dlm <- lm_med[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()

# median of non-zero cropland
yl <- c(0, round(max(med0[, snms] / 10, na.rm = TRUE)) * 10)
pdf(full_path(p_fig, "biases_magdist_med0.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- med0[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_med0)) {
  dlm <- lm_med0[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()
```

Check for spatial autocorrelation in residuals
```{r, eval=FALSE}
dists <- as.matrix(dist(mu[, c("x", "y")]))
dists_inv <- 1 / dists
diag(dists_inv) <- 0
for(i in 1:length(snms)) {
  print(Moran.I(residuals(lm_mu[[i]]$model), dists_inv)$p.value)  
}

dists <- as.matrix(dist(lm_mu0[[1]]$dat[, c("x", "y")]))
dists_inv <- 1 / dists
diag(dists_inv) <- 0
for(i in 1:length(snms)) {
  print(Moran.I(residuals(lm_mu0[[i]]$model), dists_inv)$p.value)  
} 
```

Present and significant. To deal with it, first try to use:

3. A generalized least squares (GLS) with spatial autocovariance terms.

```{r, eval=FALSE}
# mean of all data
i <- "geow"  # geowiki
dat <- proc_df(x = mu, nm = i)
dgls <- gls(dv ~ poly(pv, 2), weights = ~1 / w, data = dat)  # orthogonal polys
dgls_exp <- update(dgls, correlation = corExp(form = ~x + y, nugget = TRUE))
plot(Variogram(dgls, form = ~x+y, resType = "n"))
plot(Variogram(dgls_exp, form = ~x+y, resType = "n"))  # funny shape remains

# cropland only mean
dat <- proc_df(x = mu0, nm = i)
dat <- dat[!is.na(dat$pv), ]
dgls <- gls(dv ~ poly(pv, 2), weights = ~1 / w, data = dat)  # orthogonal polys
dgls_exp <- update(dgls, correlation = corExp(form = ~x + y, nugget = TRUE))
plot(Variogram(dgls, form = ~x+y, resType = "n"))
plot(Variogram(dgls_exp, form = ~x+y, resType = "n"))  # funny shape remains
```

That seemed inadequate for removing autocorrelation in residuals. The GLS did not seem to properly deal with residual autocorrelation, probably because the spatial structure is complex, given the pattern of rainfall and cropland in it.  A better approach is to include coordinates directly in the model, leveraging the 2-D smoothing capabilities of the GAM model in the mgcv library.

3. Use a GAM instead

First check fit of model family, and whether assumptions of normally distributed errors are met. 
```{r, eval=FALSE}
# stats.stackexchange.com/questions/99425/distribution-for-percentage-data
# quartz()
dat <- mu0[, "sa30"]
dat <- dat[!is.na(dat)]
descdist(dat, discrete = FALSE)  # check potential distributions

# Check what distributions work best on data (non-cropland removed)
dists <- c("norm", "lnorm", "exp", "cauchy", "gamma", "logis", "beta","weibull")
fit_distrs <- lapply(snms, function(x) {
  dat <- mu0[, x]
  dat <- dat[!is.na(dat)] / 100
  fits <- lapply(dists, function(x) fitdist(dat, x))
  names(fits) <- dists
  fits
})
names(fit_distrs) <- snms
lapply(fit_distrs, function(x) sort(sapply(x, function(y) y$aic)))
plot(fit_distrs[[1]]$lnorm)  # lognorm shows up as best fit

# Matters more on residuals from model fit, however. 
# dataset
# dat <- proc_df(x = mu0, nm = snms[4])  # interactively change response  
dat <- proc_df(x = med0, nm = snms[1])  # and input data
dat <- dat[!is.na(dat$pv), ]

# GAM assuming Gaussian
dgam <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
            method = "ML")
summary(dgam)
AIC(dgam)
gam.check(dgam)
summary(fitdist(residuals(dgam), "norm"))
plot(fitdist(residuals(dgam), "norm"))

# GAM Gaussian with log link
dgam2 <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
            family = gaussian(link = "log"), method = "ML")
summary(dgam2)
AIC(dgam2)
gam.check(dgam2)
summary(fitdist(residuals(dgam2), "norm"))
plot(fitdist(residuals(dgam2), "norm"))

# GAM w/Gamma
dgam3 <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
            family = Gamma(link = "log"), method = "ML")
summary(dgam3)
AIC(dgam3)  # bizarre large AIC
gam.check(dgam3)  # crazy deviance residuals
summary(fitdist(residuals(dgam3), "norm"))
plot(fitdist(residuals(dgam3), "norm"))

# GAM w/log
dgam4 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
             method = "ML")
summary(dgam4)
AIC(dgam4)
gam.check(dgam4)
summary(fitdist(residuals(dgam4), "norm"))
plot(fitdist(residuals(dgam4), "norm"))
# termplot(dgam4, terms = "poly(pv, 2)", ask = FALSE)
# pred <- predict(dgam4, newdata = newdat, type = "response", se.fit = TRUE)

# above, with smooth fit instead of quadratic
dgam4b <- gam(log(dv) ~ s(pv, k = 3) + s(x, y), weights = w, data = dat, 
             method = "ML")
summary(dgam4b)
AIC(dgam4b)
gam.check(dgam4)
summary(fitdist(residuals(dgam4b), "norm"))
plot(fitdist(residuals(dgam4b), "norm"))

plot.gam(dgam4b, residuals = TRUE, select = 1, se = TRUE, all.terms = TRUE, 
         pages = 0)

```

Testing each of the main landcover sets for model shape, the best fit in terms of AIC and assumption of normality of residuals is from a log-normal model.

### Plot final results

1. Helper functions
```{r, eval = FALSE}

predfunc <- function(model, newdat) {
  pred <- predict(model, newdata = newdat, type = "response", se.fit = TRUE)
  pdat <- transform(newdat, fit = pred$fit)
  pdat <- transform(pdat, up = fit + (1.96 * pred$se.fit),
                    lo = fit - (1.96 * pred$se.fit))
  return(pdat)
}

termfunc <- function(model, newdat) {
  pred <- predict(model, newdata = newdat, type = "terms", se.fit = TRUE)
  pdat <- transform(newdat, fit = pred$fit[, 1] + coef(model)[1])
  pdat <- transform(pdat, up = fit + (1.96 * pred$se.fit[, 1]),
                    lo = fit - (1.96 * pred$se.fit[, 1]))
  return(pdat)
}

polyfunc <- function(pdat) {
  coord <- cbind.data.frame("x" = c(rev(pdat$pv), pdat$pv), 
                            "y" = c(rev(pdat$lo), pdat$up))
}

makeTransparent<-function(someColor, alpha = 100) {
  newColor <- col2rgb(someColor)
  apply(newColor, 2, function(x) {
    rgb(red = x[1], green = x[2], blue = x[3],alpha = alpha, maxColorValue=255)
  })
}
```

2. GAM fits
```{r, eval = FALSE}
gam_mods <- lapply(nms, function(i) {
  dat1 <- proc_df(x = mu0, nm = i)
  dat1 <- dat1[!is.na(dat1$pv), ]
  dat2 <- proc_df(x = med0, nm = i)
  dat2 <- dat2[!is.na(dat2$pv), ]
  dgam1 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat1, 
               method = "ML")
  dgam2 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat2, 
               method = "ML")
  newdat <- with(dat1, 
                 data.frame(pv = 0:100, "x" = mean(dat1$x), "y" = mean(dat1$y)))
  pdat1 <- predfunc(dgam1, newdat)
  pdat2 <- predfunc(dgam2, newdat)
  return(list("mumod" = dgam1, "medmod" = dgam2, "pdat1" = pdat1, 
              "pdat2" = pdat2,  "data1" = dat1, "data2" = dat2))  
})
```

3. GAM plots
```{r, eval = FALSE}
plot(xl, yl2, pch = "", ylab = "% bias)", xlab = "% cropland", xaxs = "i",
     yaxs = "i", yaxt = "n")

round(log(seq(1, 90, 10)), 2)
exp(seq(0, 4.5, 0.5))

ptcols <- sapply(cols, function(x) makeTransparent(x, 80))  # CI colors
lw <- c(rep(2, 4), 3, 3)
yl1 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat1$fit))), 1))
yl2 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat2$fit))), 1))
pdf(full_path(p_fig, "biases_md_lnorm_gam_mu0.pdf"), width = 7, height = 7)
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
plot(xl, yl2, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i",
     mgp = c(1.5, 0.5, 0), yaxs = "i", yaxt = "n")
axis(2, at = 0:4, round(exp(c(-10, 1:4))))
for(i in 1:length(gam_mods)) {
  m <- gam_mods[[i]]
  p <- polyfunc(m$pdat1)
  polygon(p$x, p$y, col = ptcols[i], border = NA)
  lines(fit ~ pv, data = m$pdat1, col = cols[i], lwd = 1.5, lty = 1)
}
legend("topleft", c(mcap, "All 4", "No GlobCover"), lty = 1, lwd = 1.5,
       col = cols, bty = "n", xpd = NA)#, title = "Fit")
dev.off()

pdf(full_path(p_fig, "biases_md_lnorm_gam_med0.pdf"), width = 7, height = 7)
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2))
plot(xl, yl2, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i",
     mgp = c(1.5, 0.5, 0), yaxs = "i", yaxt = "n")
axis(2, at = 0:4, round(exp(c(-10, 1:4))))
for(i in 1:length(gam_mods)) {
  m <- gam_mods[[i]]
  p <- polyfunc(m$pdat2)
  polygon(p$x, p$y, col = ptcols[i], border = NA)
  lines(fit ~ pv, data = m$pdat2, col = cols[i], lwd = 1.5, lty = 1)
}
legend("topleft", c(mcap, "All 4", "No GlobCover"), lty = 1, lwd = 1.5,
       col = cols, bty = "n", xpd = NA)#, title = "Fit")
dev.off()

# Check p-values for all model terms
lapply(gam_mods, function(x) {
  m1 <- summary(x$mumod)
  m2 <- summary(x$medmod)
  rbind("mu" = round(c(m1$p.table[, 4], m1$s.pv), 3), 
        "med" = round(c(m2$p.table[, 4], m2$s.pv), 3))
})  # all terms significant at p < 0.001 except 2nd order term on GlobCov med
```

Final plot of magisterial districts, for supplementals
```{r, eval=FALSE}
provs <- unique(md$ID_1)
greys <- paste0("grey", seq(15, 95, 10))
pdf(full_path(p_fig, "md_map.pdf"), width = 7, height = 7, bg = "transparent")
par(mar = rep(0, 4))
plot(sa.shp, lty = 0)
for(i in 1:length(provs)) {
  plot(md[md$ID_1 == provs[i], ], col = greys[i], #border = "transparent", 
       add = TRUE, lwd = 1)
}
dev.off()
```

### 2015-06-26 Rmd/fetch-land-cover.Rmd
Created markdown file to select magisterial districts for agent-based model analysis.  Edited formats in older fetch-landcover.Rmd file.

@@
 drv <- dbDriver("PostgreSQL")
-con <- dbConnect(drv, dbname = "SouthAfricaSandbox", user = "postgis", password = "P0stG1S")
+con <- dbConnect(drv, dbname = "SouthAfricaSandbox", user = "postgis", 
+                 password = "P0stG1S")
 
@@
 prjsrid <- 97490  # Albers
-prj_sql <- paste("select proj4text from spatial_ref_sys where srid=", prjsrid, sep = "")
+prj_sql <- paste0("select proj4text from spatial_ref_sys where srid=", prjsrid)
 prjstr <- dbGetQuery(con, prj_sql)$proj4text
 gcsid <- 4326  # GCS
-gcs_sql <- paste("select proj4text from spatial_ref_sys where srid=", gcsid, sep = "")
+gcs_sql <- paste0("select proj4text from spatial_ref_sys where srid=", gcsid)
 gcsstr <- dbGetQuery(con, gcs_sql)$proj4text
@@
 # Create 1 km grid from sa1kilo to rectify grids against that 
-if(!file.exists(full_path(p_data, "sagrid.tif"))) {  # Wrap in if statement in case grid already made
-  sql <- paste("SELECT gid, id, ST_AsText(ST_Centroid(geom)) as center FROM sa1kilo")
+if(!file.exists(full_path(p_data, "sagrid.tif"))) {  # already made? 
+  sql <- paste("SELECT gid, id, ST_AsText(ST_Centroid(geom)) as center", 
+               "FROM sa1kilo")
   geom.tab <- dbGetQuery(con, sql)
@@
   projection(sa_r) <- sa.buf@proj4string
-  sa_r <- writeRaster(sa_r, filename = full_path(p_data, "sagrid.tif"), overwrite = TRUE)
+  sa_r <- writeRaster(sa_r, filename = full_path(p_data, "sagrid.tif"), 
+                      overwrite = TRUE)
 } else {
@@
 ```{r, eval = FALSE}
-if(!file.exists(full_path(p_data, "glcsa_masked.tif"))) {  # see if file exists, to avoid redoing step
+if(!file.exists(full_path(p_data, "glcsa_masked.tif"))) {  # see if file exists,  
+  # to avoid redoing step
   path <- "/u/lestes/spatial_data/glc_share/"
@@
   projection(glc) <- gcsstr
-  glcSA <- crop(glc, y=sa_buf_gcs, file = full_path(path, "glcSA.tif"), overwrite = TRUE)  # crop to SA
-  gdalwarp(srcfile = glcSA@file@name, dstfile = full_path(path, "glcSA_alb.tif"), t_srs = prjstr, 
+  glcSA <- crop(glc, y=sa_buf_gcs, file = full_path(path, "glcSA.tif"), 
+                overwrite = TRUE)  # crop to SA
+  gdalwarp(srcfile = glcSA@file@name, 
+           dstfile = full_path(path, "glcSA_alb.tif"), t_srs = prjstr, 
            tr = c(1000, 1000), r = "bilinear")
@@
   # Warp to SA grid
-  glcsa <- resample(glcSA_alb, sa_r, method = "bilinear", filename = full_path(path, "glcSA_alb_rect.tif"))
-  glcsa_m <- mask(glcsa, sa_r, file = full_path(p_data, "glcsa_masked.tif"), overwrite = TRUE)
+  glcsa <- resample(glcSA_alb, sa_r, method = "bilinear", 
+                    filename = full_path(path, "glcSA_alb_rect.tif"))
+  glcsa_m <- mask(glcsa, sa_r, file = full_path(p_data, "glcsa_masked.tif"),
+                  overwrite = TRUE)
 } else {
@@
   projection(geowiki) <- gcsstr
-  geowikiSA <- crop(geowiki, y=sa_buf_gcs, file = full_path(path, "geowikiSA.tif"))  # crop to extent of SA
-  gdalwarp(srcfile = geowikiSA@file@name, dstfile = full_path(path, "geowikiSA_alb.tif"), t_srs = prjstr, 
+  geowikiSA <- crop(geowiki, y=sa_buf_gcs, 
+                    file = full_path(path, "geowikiSA.tif"))  # crop to SA
+  gdalwarp(srcfile = geowikiSA@file@name, 
+           dstfile = full_path(path, "geowikiSA_alb.tif"), t_srs = prjstr, 
            tr = c(1000, 1000), r = "bilinear")
   geowikisa_alb <- raster(full_path(path, "geowikiSA_alb.tif"))
-  geowikisa <- resample(geowikisa_alb, sa_r, method = "bilinear", 
-                        filename = full_path(path, "geowikisa_alb_rect.tif"))    # Warp to SA grid
-  geowikisa_m <- mask(glcsa, sa_r, file = full_path(p_data, "geowikisa_masked.tif"), overwrite = TRUE)
+  geowikisa <- resample(geowikisa_alb, sa_r, method = "bilinear", # Warp to SA
+                        filename = full_path(path, "geowikisa_alb_rect.tif"))    
+  geowikisa_m <- mask(glcsa, sa_r, 
+                      file = full_path(p_data, "geowikisa_masked.tif"), 
+                      overwrite = TRUE)
 } else {
@@
   globSA <- crop(glob, y=sa_buf_gcs)  # crop to extent of SA
-  gdalwarp(srcfile = globSA@file@name, dstfile = full_path(path, "globSA_alb.tif"), t_srs = prjstr, 
+  gdalwarp(srcfile = globSA@file@name, 
+           dstfile = full_path(path, "globSA_alb.tif"), t_srs = prjstr, 
            tr = c(300, 300))
@@
   lapply(modnames, function(x) {
-    url <- paste("http://e4ftl01.cr.usgs.gov/MOTA/MCD12Q1.051/2011.01.01/", x, sep = "")
+    url <- paste0("http://e4ftl01.cr.usgs.gov/MOTA/MCD12Q1.051/2011.01.01/", x)
     download.file(url, method="auto", destfile = x)
@@
   dir.create("MCD12Q1")
-  file.copy(from = dir(pattern = "hdf"), to = paste(getwd(), "MCD12Q1", dir(pattern = "hdf"), sep = "/"))
+  file.copy(from = dir(pattern = "hdf"), 
+            to = paste(getwd(), "MCD12Q1", dir(pattern = "hdf"), sep = "/"))
   file.remove(dir(pattern = "hdf"))
@@
     batch_gdal_translate(infiles = modis, outdir = moddir, 
-                         outsuffix = paste("_LC_", x, ".tif", sep = ""), sd_index = x)  
+                         outsuffix = paste("_LC_", x, ".tif", sep = ""), 
+                         sd_index = x)  
   })
-  sdnames <- lapply(1:5, function(x) dir(moddir, pattern  = paste("LC_", x, ".tif", sep = "")))
+  sdnames <- lapply(1:5, function(x) dir(moddir, 
+                                         pattern = paste0("LC_", x, ".tif")))
 
@@
     print(paste("processing type", x))
-    gdalwarp(srcfile = sdnames[[x]], dstfile = paste("modis_type_", x, "_alb.tif", sep = ""), 
+    gdalwarp(srcfile = sdnames[[x]], 
+             dstfile = paste0("modis_type_", x, "_alb.tif"), 
              t_srs = prjstr, tr = res(r), te = bbox(sa.buf)[1:4])
@@
   setwd(path)
-  gdal_translate(src_dataset = "landcover", dst_dataset = "sa_lc_2009.tif", of = "GTiff", ot = "BYTE")
+  gdal_translate(src_dataset = "landcover", dst_dataset = "sa_lc_2009.tif", 
+                 of = "GTiff", ot = "BYTE")
   #system(paste("rm -r", full_path(path, "landcover")))
@@
 
-  # process masks--extract extract landcover class 2, cultivation, and class 6, plantations, resample, mask 
-  # with gdal tools
+  # process masks--extract extract landcover class 2, cultivation, and class 6,
+  # plantations, resample, mask with gdal tools
   av <- c(2, 6)
@@
     print(paste("extracting cover for", lcnm[j]))
-    gdal_calc(cstr = paste("A==", av[j], sep = ""), x = list("A" = a), type = "Byte", filename = lcnm[j])
-    nm <- full_path(p_data, paste(gsub("\\.tif", "", lcnm[j]), "_masked.tif", sep = ""))
-    ext <- sapply(c("xmin", "ymin", "xmax", "ymax"), function(x) slot(extent(sa_r), x))
+    gdal_calc(cstr = paste("A==", av[j], sep = ""), x = list("A" = a), 
+              type = "Byte", filename = lcnm[j])
+    nm <- full_path(p_data, 
+                    paste(gsub("\\.tif", "", lcnm[j]), "_masked.tif", sep = ""))
+    ext <- sapply(c("xmin", "ymin", "xmax", "ymax"), function(x) {
+      slot(extent(sa_r), x)
+    })
     print(paste("warping and masking", gsub(p_data, "", nm)))
-    gdalwarp(srcfile = lcnm[j], t_srs = projection(sa_r), dstfile = nm, r = "average", ot = "Float32", 
-             te = ext, srcnodata = 255, dstnodata = 255, tr = c(1000, 1000), of = "GTiff", verbose = TRUE,
+    gdalwarp(srcfile = lcnm[j], t_srs = projection(sa_r), dstfile = nm, 
+             r = "average", ot = "Float32", te = ext, srcnodata = 255, 
+             dstnodata = 255, tr = c(1000, 1000), of = "GTiff", verbose = TRUE,
              overwrite = TRUE)
@@
   setwd(kznpath)
-  gdal_translate(src_dataset = "kznlc11v1w31", dst_dataset= "kznlandcover.tif", of = "GTiff")
+  gdal_translate(src_dataset = "kznlc11v1w31", dst_dataset= "kznlandcover.tif",
+                 of = "GTiff")
   system(paste("rm -r", full_path(path, "kznlc11v1w31")))
@@
   kznlconm <- "kzn_cane.tif"  # file.remove(kznlconm)
-  kzninnm <- strsplit(kznlc@file@name, "/")[[1]][length(strsplit(kznlc@file@name, "/")[[1]])]
+  ss <- strsplit(kznlc@file@name, "/")
+  kzninnm <- ss[[1]][length(ss[[1]])]
   # kzninnm <- strsplit(kznlc.ss@file@name, "/")[[1]][length(strsplit(kznlc.ss@file@name, "/")[[1]])]
@@
   setwd(p_data)
-  gdalwarp(srcfile = kzn_cane@file@name, t_srs = projection(sa_r), dstfile = "kzn_cane_1km.tif", 
-           r = "average", ot = "Float32", srcnodata = 255, tr = c(1000, 1000), of = "GTiff", 
+  gdalwarp(srcfile = kzn_cane@file@name, t_srs = projection(sa_r), 
+           dstfile = "kzn_cane_1km.tif", r = "average", ot = "Float32", 
+           srcnodata = 255, tr = c(1000, 1000), of = "GTiff", 
            verbose = TRUE, overwrite = TRUE)
@@
   kzn_1km_r <- resample(kzn_1km, sa_r, method = "bilinear")
-  kzn_1km_m <- mask(kzn_1km_r, sa_r, filename = full_path(p_data, "kzn_cane_masked.tif"), 
+  kzn_1km_m <- mask(kzn_1km_r, sa_r, 
+                    filename = full_path(p_data, "kzn_cane_masked.tif"), 
                     overwrite = TRUE)
@@
 ```{r, eval = FALSE}
-if(any(!file.exists(full_path(p_data, dir(p_data, pattern = "mod_1*.*1kmrect_sum.tif"))))) {
+if(any(!file.exists(full_path(p_data, 
+                              dir(p_data, 
+                                  pattern = "mod_1*.*1kmrect_sum.tif"))))) {
   # MODIS, using main classification scheme (IGBP-DIS)
@@
   # assign percentages
-  minset <- c(100, 10)  # assume 10% is minimum crop cover for MODIS IGBP mosaic class
+  minset <- c(100, 10)  # assume 10% is min crop cover for MODIS mosaic class
   meanset <- c(100, 35)  # assume 35% is mean crop cover
   maxset <- c(100, 60)  # assume 60% is mean crop cover  
-  modsa_min <- assignLCPct(mod_1_cropb, minset, fname = full_path(p_data, "mod_1_min.tif"))
-  modsa_mu <- assignLCPct(mod_1_cropb, meanset, fname = full_path(p_data, "mod_1_mu.tif"))
-  modsa_max <- assignLCPct(mod_1_cropb, maxset, fname = full_path(p_data, "mod_1_max.tif"))
+  modsa_min <- assignLCPct(mod_1_cropb, minset, 
+                           fname = full_path(p_data, "mod_1_min.tif"))
+  modsa_mu <- assignLCPct(mod_1_cropb, meanset, 
+                          fname = full_path(p_data, "mod_1_mu.tif"))
+  modsa_max <- assignLCPct(mod_1_cropb, maxset, 
+                           fname = full_path(p_data, "mod_1_max.tif"))
 
@@
     print("...masking")
-    crop1km <- mask(crop1km, mask = sa_r, filename = paste(nm, "_1kmrect.tif", sep = ""), overwrite = TRUE)
+    crop1km <- mask(crop1km, mask = sa_r, filename = paste0(nm, "_1kmrect.tif"),
+                    overwrite = TRUE)
     return(crop1km)

### 2015-06-26 Rmd/magisterial-district-ABM-select.Rmd
Created markdown file to select magisterial districts for agent-based model analysis.  Edited formats in older fetch-landcover.Rmd file.

---
title: "Districts for ABM"
author: "Lyndon Estes"
date: "June 26, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
---

## Overview

Selection of magisterial districts and associated landcover sets for use in ABM example.

## Data
```{r, eval=FALSE}
library(raster)
library(rgdal)
library(lmisc)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")

# landcover data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*07sum_mask*")))
gti <- raster(paths)  # gti 2007 rather than 2011, to match census data

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2007))
lc_list <- gti - dlist_1km

# magisterial district data cropland statistics (2011, but only to look at for 
# selection 
load("external/ext_data/gti_md.rda")
mu <- data.frame(t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
})))
colnames(mu) <- c("N", "na", "f")

# district shapes, filtered by N missing data, minimum fraction, and area
md <- readOGR("external/ext_data/mag_dists.sqlite", layer = "mag_dists")
mdfilt <- mu[mu$na < 20 & mu$f > 5 & mu$N < 1500, ]  
md@data[mdsel, 7:8]
mdsel <- as.numeric(row.names(mdfilt))

# Examine, including in QGIS
plot(md)
plot(md[mdsel[mdsel < 314], ], col = "red", add = TRUE)  # filter out Cape ones
writeOGR(md[mdsel[mdsel < 314], ], dsn = "external/ext_data/md_abms.sqlite",
         layer = "md_abms", driver = "SQLite")

```

### Selection

Looking at the data in QGIS, including some rainfall maps from earlier papers and the inter-tubes, it looks like the following districts are a good match:

   + Clocolan (269)
   + Ficksburg (274)
   + Fouriesburg (275)
   + Marquard (288)

```{r, eval=FALSE}
abm_md <- md[mdsel[mdsel %in% c(269, 274:275, 288)], ]
writeOGR(abm_md, dsn = "external/ext_data/md4.sqlite", layer = "md4", 
         driver = "SQLite")

# Let's compare the fractions between the GTI 2007 and GTI 2011 datasets for 
# interest's sake
gti_abm <- extract(gti, abm_md)
abm_mu <- data.frame(t(sapply(gti_abm, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
})))
colnames(abm_mu) <- c("N", "na", "f")
abm_mu$f - mu[mdsel[mdsel %in% c(269, 274:275, 288)], "f"]
 # -1.0  0.0 -0.4 -0.2  # tinf differences!
```

### Extraction
```{r, eval=FALSE}
gti4 <- crop(gti, extent(abm_md))
lc4 <- crop(lc_list, extent(abm_md))
cb <- brick(stack(gti4, lc4), filename = "external/ext_data/cropland-md4.tif")
```

### Visuals
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(raster)
library(rgdal)
library(lmisc)

setwd(full_path(proj_root("SAcropland"), "SAcropland"))
md <- readOGR("external/ext_data/mag_dists.sqlite", layer = "mag_dists")
abm_md <- readOGR("external/ext_data/md4.sqlite", layer = "md4")
cb <- brick("external/ext_data/cropland-md4.tif")
```

```{r}
par(mar = c(0, 0, 0, 0))
plot(md, col = "grey", border = "transparent")
plot(abm_md, col = "red", border = "grey30", add = TRUE)
```

The four selected districts.

```{r}
par(mar = c(0, 0, 0, 0))
plot(cb[[1]], axes = FALSE, box = FALSE)
plot(abm_md, add = TRUE)
```

2007 cropland fraction in the four districts.

```{r}
par(mar = c(0, 0, 0, 0), mfrow = c(2, 2))
for(i in 2:5) {
  plot(cb[[i]], axes = FALSE, box = FALSE, zlim = c(0, 101))
  plot(abm_md, add = TRUE)
}
```

Cropland fractions estimated by South Africa's landcover product, GLobCover, MODIS, and Geo-Wiki for the same 4 districts. Note the substantial variability.

### 2015-06-30 Rmd/carbon-bias.Rmd
Rmd for carbon-bias estimation added

---
title: "Carbon bias"
author: "Lyndon Estes"
date: "June 24, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
---

## Bias in carbon stock estimates

Based on the Ruesch & Gibbs (2008) approach for estimating carbon density.

## Data

Paths
```{r, eval=FALSE}
library(raster)
library(rgdal)
library(lmisc)
library(data.table)
library(SAcropland)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")
p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")
```

Starting with the carbon look-up tables used for Africa by Ruesch & Gibbs (2008), inputs downloaded from [here](http://cdiac.ornl.gov/epubs/ndp/global_carbon/carbon_documentation.html#methods).
```{r, eval=FALSE}
cfiles <- dir(p_carb, pattern = "m6", full.names = TRUE)
nms <- gsub("*.*m6|\\.txt", "", cfiles)
cval_list <- lapply(1:length(cfiles), function(x) {
  tab <- read.table(cfiles[x])
  tab <- tab[ c(1, 3)]
  colnames(tab) <- c("CL", nms[x])
  tab
})

# merge into single carbon table
mergefun <- function(x, y) merge(x, y, by = "CL", all.x = TRUE, all.y = TRUE)
ctab <- Reduce(mergefun, cval_list)

# Read in lookup table key
key <- readLines(dir(p_carb, pattern = "key", full.names = TRUE))[25:44]
key2 <- readLines(dir(p_carb, pattern = "key", full.names = TRUE))[47:57]
mtch <- sapply(gregexpr("[0-9]", key2), max)
lcs <- sapply(1:length(key2), function(x) substr(key2[x], 1, mtch[x]))
lcs <- gsub(" & ", ",", gsub("-", ":", lcs))

# Reshape and reduce according to carbon classes
ctab2 <- t(data.frame(sapply(lcs, function(x) {
  ind <- eval(parse(text = paste0("c(", x, ")")))
  as.numeric(as.vector(
    round(colMeans(ctab[ctab$CL %in% ind, -1], na.rm = TRUE))))
})))
rownames(ctab2) <- 1:nrow(ctab2)
colnames(ctab2) <- nms
ctab2 <- ctab2 * 0.01  # convert to tons/ha from 1000 kg/ha
ctab2 <- cbind(transform(lcs), ctab2)
colnames(ctab2)[1] <- "class"
```

Further compress classes that have the same carbon value.
    + 1:3, 6:8 => 1 (broadleaf and mixed forests)
    + 4:5 => Drop
    + 9, 10; 17 => 2 (Secondary forests, forest/cropland mosaic)
    + 11, 12, 15 => 3 (shrublands)
    + 20:23 => drop (water, snow, artificial surfaces)
    + 19 => drop (bare areas)

```{r, eval=FALSE}
ctabf <- round(rbind(colMeans(ctab2[c(1, 3), 2:ncol(ctab2)], na.rm = TRUE), 
                     colMeans(ctab2[10:11, 2:ncol(ctab2)], na.rm = TRUE), 
                     ctab2[4:6, 2:ncol(ctab2)]))

# Read in ecofloristic regions and calculate their areas for Africa
ecoflora <- readOGR("external/ext_data/africa_ecofloristic_zones.sqlite", 
                    layer = "africa_ecofloristic_zones")
ecoflora@proj4string <- CRS("+proj=longlat +datum=WGS84 +no_defs")
afalb <- paste0("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 ", 
                "+y_0=0 +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")
ecoflora_alb <- spTransform(ecoflora, CRS = CRS(afalb))
ecoflora_num <- cbind.data.frame(unique(ecoflora_alb@data),  
                                 "ID" = c("06", "08", "09", "15", "NA", "18",
                                          "20", "16", "17", "WA", "19", "07"),
                                 stringsAsFactors = FALSE)
ecoflora_alb$ID <- rep("NA", nrow(ecoflora_alb))
head(ecoflora_alb)
ecoflora_alb$ID <- ecoflora_num[match(ecoflora_alb$gez_term, 
                                      ecoflora_num$gez_term), "ID"]
# ecoflora_alb@data <- sp::merge(ecoflora_alb@data, ecoflora_num, 
#                                by = "gez_term", sort = FALSE, keep = TRUE)
ecoflora_alb@data[sample(1:nrow(ecoflora_alb@data), 2), ]
ecoflora_alb$area <- round(rgeos::gArea(ecoflora_alb, byid = TRUE) / 1000000, 1)
ecoareas <- sapply(ecoflora_num$ID, function(x) {
  sum(ecoflora_alb@data[ecoflora_alb$ID == x, "area"])
})
ecoareas <- ecoareas[!names(ecoareas) %in% c("WA", "NA")]
ecowgts <- ecoareas / sum(ecoareas)
ecowgts <- ecowgts[sort(names(ecowgts))]

# par(mar = rep(0, 4))
# plot(ecoflora_alb)
# plot(ecoflora_alb[ecoflora_alb$ID == "17", ], col = "red", add = TRUE)

i <- 1:ncol(ctabf)
ctabf$mu <- round(sapply(1:nrow(ctabf), function(x) {
  sum(ctabf[x, ] * ecowgts, na.rm = TRUE)
}), 1)
ctabf <- cbind(ctabf, t(apply(ctabf[, i], 1, function(x) range(x, na.rm=TRUE))))
colnames(ctabf)[(ncol(ctabf) - 1):ncol(ctabf)] <- c("min", "max")
rownames(ctabf) <- 1:nrow(ctabf)
LC <- c("forest", "second", "shrub", "grass", "sparse")
ctabo <- data.frame(t(ctabf[, c("mu", "min", "max")]))  # reclass table
colnames(ctabo) <- LC

```

### Analysis
1. Prepare raster data
```{r, eval=FALSE}
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100) 
plot(lc_list[[1]])
plot(gti)
# tst <- raster("external/ext_data/geowikisa_masked.tif")  # check
# plot(round(tst / 100 - lc_list[[4]], 4)) # okay

fact <- c(5, 25, 50, 100)
lc_agg <- aggregate_rast_list(fact, lc_list)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, list("gti" = gti))  # GTI rasters
```

2. Create cropland carbon estimates
```{r, eval=FALSE}

# Apply carbon estimates
cc <- ctab2[ctab2$class == 16, 2]
carbon <- function(fcrop, cropC, noncropC) {
  carb <- fcrop * cropC + noncropC * (1 - fcrop)
  return(round(carb, 2))
}

cc <- ctab2[ctab2$class == 16, 2]
gti_c <- lapply(gti_agg, function(x) {
  r <- x$gti
  s <- stack(lapply(1:ncol(ctabo), function(y) {
    carbon(r, cc, ctabo[1, y])
  }))
  names(s) <- colnames(ctabo)
  s
})

lc_c <- lapply(lc_agg, function(x) {
  lc <- lapply(x, function(j) {
    s <- stack(lapply(1:ncol(ctabo), function(y) {
      carbon(j, cc, ctabo[1, y])
    }))
    names(s) <- colnames(ctabo)
    s
  })
  names(lc) <- names(x)
  lc
})

# checks 
tst <- cbind(sample(1:5, 10, replace = TRUE), 
             sample(1:4, 10, replace = TRUE),
             sample(1:5, 10, replace = TRUE))
sapply(1:nrow(tst), function(i) {
  cellStats(lc_c[[tst[i, 1]]][[tst[i, 2]]][[tst[i, 3]]] - 
              carbon(lc_agg[[tst[i, 1]]][[tst[i, 2]]], cc, ctabo[1, tst[i, 3]]),
              sum)
})
# all zeroes

```

3. Difference the carbon datasets
```{r, eval=FALSE}
# percent difference
pct_diff <- function(x, y) (x - y) / x * 100
c_pct_diff <- lapply(1:length(gti_c), function(x) {
  dif <- lapply(1:length(lc_c[[x]]), function(y) {
    s <- stack(lapply(1:nlayers(lc_c[[x]][[y]]), function(z) {
      p <- pct_diff(gti_c[[x]][[z]], lc_c[[x]][[y]][[z]])
    }))
    names(s) <- colnames(ctabo)
    s
  })
  names(dif) <- names(lc_c[[x]])
  dif
})
names(c_pct_diff) <- names(lc_c)

# checks 
tst <- cbind(sample(1:5, 10, replace = TRUE), 
             sample(1:4, 10, replace = TRUE),
             sample(1:5, 10, replace = TRUE))
sapply(1:nrow(tst), function(i) {
  x <- gti_c[[tst[i, 1]]][[tst[i, 3]]]
  y <- lc_c[[tst[i, 1]]][[tst[i, 2]]][[tst[i, 3]]]
  z <- c_pct_diff[[tst[i, 1]]][[tst[i, 2]]][[tst[i, 3]]]
  cellStats(z - ((x - y) / x * 100), sum)
})
# zeroes

# Then for a map plot figure, calculate the mean pixel-wise percent difference
c_pct_diff_mu <- lapply(c_pct_diff, function(x) {
  lapply(x, function(y) calc(y, mean))
})

# # disggregate selected rasters at selected levels for plotting
namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
lev <- names(c_pct_diff_mu) 
disagg <- lapply(snms, function(x) {
  l1 <- lapply(lev, function(y) {
    if(y == "f1") {
      r <- c_pct_diff_mu[[y]][[x]]
    } else {
      r <- raster::disaggregate(c_pct_diff_mu[[y]][[x]], 
                                fact = as.numeric(gsub("f", "", y)))
      r <- raster::mask(crop(r, namask), namask)
    }
  })
  named_out(l1, lev)
})
names(disagg) <- snms
c_pct_diff_mu$f1

stats <- lapply(disagg, function(x) {
  sapply(x, function(y) {
    c(cellStats(y, mean), quantile(y))
  })
})

plot(c_pct_diff_mu$f1$geow - c_pct_diff_mu$f1$sa30)

# Plotting colors
cols <- colorRampPalette(c("red", "grey80", "blue4"))
brks <- c(-100.1, -75, -50, -25, -10, 0, 10, 25, 50, 75, 100.1)
n_cols <- length(brks) - 1
legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "bias_map.pdf"), height = 6, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols(n_cols), 
         breaks = brks, legend = FALSE)
  if(j == 1) mtext(mcap[i], side = 2, line = 1, cex = cx)
  if(i == 1) mtext(lev2[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext, 
            legend.vals = round(brks), 
            longdims = c(0.2, 0.8), shortdims = c(0.06, 0.01), 
            colvec = cols(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0.5, -0.5)), 
            cex.val = cx, textcol = lcol, bordercol = lcol)
dev.off()

```

2. Convert them to data.tables and remove NA data

```{r, eval=FALSE}
# GTI
# gti0 <- 1 - gti / 100  # non cropland fractions for GTI data
# gti_c <- gti / 100 * ctabo$crops[1]  # GTI cropland carbon
gti_dt <- agroEcoTradeoffs::as.data.table.raster(gti / 100, xy = TRUE)
gti_dt[, id:= 1:.N]
setkeyv(gti_dt, c("x", "y"))
setnames(gti_dt, "cover2011sum_mask", "fA")

lc_dt <- lapply(1:nlayers(lc_list), function(x) {
  DT <- agroEcoTradeoffs::as.data.table.raster(lc_list[[x]] / 100, xy = TRUE)
  DT[, id := 1:.N]
  setkeyv(DT, c("x", "y"))
  setnames(DT, names(lc_list[[x]]), "fA")
  #DT <- DT[!is.na(fA), ]
})

# find NAs common to all datasets
na1 <- gti_dt[is.na(fA)]
na2 <- rbind(na1, do.call(rbind, lapply(lc_dt, function(x) x[is.na(fA)])))
setkey(na2)
na2 <- na2[!duplicated(na2)]
nrow(na2[unique(id)]) == nrow(na2)

# Remove all common NA values, to tidy up size
gti_dt <- gti_dt[!na2]
gti_dt[, id := NULL]
lc_dt <- lapply(lc_dt, function(x) {
  DT <- x[!na2]
  DT[, id := NULL]
})

# plot(agroEcoTradeoffs::dt_to_raster(gti_dt, gti@crs)[[1]])
```

```{r}
for(i in 1:(ncol(ctabo))) {
  nm <- colnames(ctabo)[i]
  for(j in 1) { #:3) {
    nmv <- paste0(nm, "_", rownames(ctabo)[j])
    gti_dt[, c(nmv) := carbon(fA, noncropC = ctabo[j, nm]), with = FALSE] 
  }
}

# To GTI
for(i in 1:(ncol(ctabo))) {
  nm <- colnames(ctabo)[i]
  for(j in 1) { #:3) {
    nmv <- paste0(nm, "_", rownames(ctabo)[j])
    gti_dt[, c(nmv) := carbon(fA, noncropC = ctabo[j, nm]), with = FALSE] 
  }
}

# To landcover sets
for(k in 1:length(lc_dt)) {
  for(i in 1:(ncol(ctabo))) {
    nm <- colnames(ctabo)[i]
    for(j in 1) {#:3) {
      nmv <- paste0(nm, "_", rownames(ctabo)[j])
      lc_dt[[k]][, c(nmv) := carbon(fA, noncropC = ctabo[j, nm]), with = FALSE] 
    }
  }
}

```

```{r, eval=FALSE}
# Actual difference
cdiff_dt <- lapply(1:length(lc_dt), function(x) {
  ind <- colnames(gti_dt)[4:ncol(gti_dt)]
  DT <- copy(gti_dt)  # so gti_dt is not modified
  DT2 <- copy(lc_dt[[x]])
  DT[, fA := NULL]
  DT[, `:=`(ind, Map('-', DT[, ind, with = FALSE], 
                     DT2[, ind, with = FALSE])), with = FALSE]
})

cdiff_dts <- lapply(cdiff_dt, function(x) {
  ind <- colnames(x)[3:ncol(x)]
  DT <- copy(x)  # so gti_dt is not modified
  DTo <- DT[, 1:2, with = FALSE]
  DTo[, mx := rowMaxs(as.matrix(DT[, ind, with = FALSE]))]
  DTo[, mu := rowMeans(DT[, ind, with = FALSE])]
  DTo[, mn := rowMins(as.matrix(DT[, ind, with = FALSE]))]
})

perc_diff <- function(x, y) round((x - y) / x * 100, 1)

# Difference as a percent
cdiffpct_dt <- lapply(1:length(lc_dt), function(x) {
  ind <- colnames(gti_dt)[4:ncol(gti_dt)]
  DT <- copy(gti_dt)  # so gti_dt is not modified
  DT2 <- copy(lc_dt[[x]])
  DT[, fA := NULL]
  DT[, `:=` (ind, Map(perc_diff, x = DT[, ind, with = FALSE], 
                      y = DT2[, ind, with = FALSE])), with = FALSE]
})

```

### 2015-07-06 Rmd/carbon-bias.Rmd
Initial versions of carbon bias estimation complete. Added utility functions used in various analyses to library. Minor edit to magisterial-district-bias.Rmd

@@
     toc: yes 
+    number_sections: true
 ---
@@
 ## Bias in carbon stock estimates
-
 Based on the Ruesch & Gibbs (2008) approach for estimating carbon density.
@@
 
-Paths
+### Paths
 ```{r, eval=FALSE}
@@
 library(SAcropland)
+library(RColorBrewer)
 
@@
 ```
+
+### Carbon values
 
@@
 ```
+<a href="#top">Back to top</a>
 
-Further compress classes that have the same carbon value.
+### Further compress classes that have the same carbon value.
+
   + 1:3, 6:8 => 1 (broadleaf and mixed forests)
@@
 ```
+<a href="#top">Back to top</a>
 
-### Analysis
-1. Prepare raster data
+## Analysis
+### Prepare raster data
 ```{r, eval=FALSE}
@@
 
-fact <- c(5, 25, 50, 100)
+# aggregate rasters
+fact <- c(5, 10, 25, 50, 100)
 lc_agg <- aggregate_rast_list(fact, lc_list)   # landcover rasters 
 gti_agg <- aggregate_rast_list(fact, list("gti" = gti))  # GTI rasters
+
+# need NA mask for weighting of aggregated value
+namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
+namask2 <- !is.na(namask)  # set NAs to zero
+area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
+
+# cropland cover bins, for looking at error as a function of cover
+binv <- seq(0, 1, 0.05)
+gti_bins <- lapply(gti_agg, function(x) {
+  cut(x$gti, breaks = binv, include.lowest = TRUE)
+})
+
 ```
+<a href="#top">Back to top</a>
 
-2. Create cropland carbon estimates
+### Create cropland carbon estimates
 ```{r, eval=FALSE}
-
 # Apply carbon estimates
@@
 
-# checks 
+# checks--right rasters being referenced? 
 tst <- cbind(sample(1:5, 10, replace = TRUE), 
@@
               sum)
-})
-# all zeroes
+}) # all zeroes
 
 ```
+<a href="#top">Back to top</a>
 
-3. Difference the carbon datasets
+### Difference the carbon datasets
 ```{r, eval=FALSE}
@@
 
-# checks 
+# checks - right rasters being referenced 
 tst <- cbind(sample(1:5, 10, replace = TRUE), 
@@
   cellStats(z - ((x - y) / x * 100), sum)
-})
-# zeroes
+})  # zeroes
 
@@
 names(disagg) <- snms
-c_pct_diff_mu$f1
 
@@
   sapply(x, function(y) {
-    c(cellStats(y, mean), quantile(y))
+    c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
   })
 })
+```
+<a href="#top">Back to top</a>
 
-plot(c_pct_diff_mu$f1$geow - c_pct_diff_mu$f1$sa30)
+## Outputs
+### Carbon bias maps
+```{r, eval=FALSE}
+load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
 
 # Plotting colors
-cols <- colorRampPalette(c("red", "grey80", "blue4"))
-brks <- c(-100.1, -75, -50, -25, -10, 0, 10, 25, 50, 75, 100.1)
-n_cols <- length(brks) - 1
+lims <- c(ceiling(min(sapply(stats, function(x) x[3, ]))), 
+          floor(max(sapply(stats, function(x) x[21, ]))))
+rng <- range(sapply(stats, range))
+brks <- c(rng[1], lims[1], -45, -20, -10, -5, -1, 1, 5, 10, lims[2], rng[2])
+#n_cols <- length(brks) - 1
+colsall <- brewer.pal(n = 11, "Spectral")
+cols <- c(colsall[1:6], "grey80", colsall[c(7, 9, 10, 11)])
+#colorRampPalette(c("red", "tan", "grey80"))
+#cols2 <- colorRampPalette(c("grey80", "green4", "blue"))
+#cols <- c(cols1(8), c(cols2(5)))
+
+# brks <- c(-100.1, -75, -50, -25, -10, 0, 10, 25, 50, 75, 100.1)
+# n_cols <- length(brks) - 1
 legtext <- "% Difference"
@@
 mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
+lev <- names(disagg[[1]])[-c(2:3)]
 lev2 <- c("1 km", "25 km", "50 km", "100 km")
-pdf(full_path(p_fig, "bias_map.pdf"), height = 6, width = 7)
+pdf(full_path(p_fig, "carbon_bias_map.pdf"), height = 6, width = 7)
 par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
@@
     plot(sa.shp, lty = 0)
-    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols(n_cols), 
+    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols, 
          breaks = brks, legend = FALSE)
@@
             longdims = c(0.2, 0.8), shortdims = c(0.06, 0.01), 
-            colvec = cols(length(brks) - 1), 
+            colvec = cols, #(length(brks) - 1), 
             srt = c(270, 0), horiz = TRUE, textside = "bottom", 
-            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0.5, -0.5)), 
+            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0, -0.5)), 
             cex.val = cx, textcol = lcol, bordercol = lcol)
+dev.off()
+```
+<a href="#top">Back to top</a>
+
+### Carbon bias statistics
+#### Calculate weighted bias statistics
+```{r, eval=FALSE}
+# Helper functions to pass into data.table
+# not used - switch on if quantiles needs
+# bfn <- function(x, y) {
+#   box_stats(x, weighted = TRUE, weight.opts = list("weights" = y))
+# }
+# bfna <- function(x, y) {
+#   box_stats(abs(x), weighted = TRUE, weight.opts = list("weights" = y))
+# }
+wm <- function(x, w) stats::weighted.mean(x, w)
+wma <- function(x, w) stats::weighted.mean(abs(x), w)
+
+# check constancy of < 1/2% cropland being excluded
+areas <- sapply(lc_agg, function(x) res(x[[1]])[1]^2 / 10000)
+plot(areas, (areas * 0.005))  # scales
+(areas * 0.005)  # 1/2 to 5000 ha
+
+lev_vec <- names(c_pct_diff)
+# created mask for non-cropland areas, unioning GTI and each LC, filtering out
+# areas of no-cropland (<1/2% total cover)
+lc_union <- lapply(lev_vec, function(x) {
+  lcb <- lapply(snms, function(y) {
+    gti_gt0 <- Which(round(gti_agg[[x]]$gti * 100) > 0)
+    lc_gt0 <- Which(round(lc_agg[[x]][[y]] * 100) > 0)
+    all_gt0 <- gti_gt0 + lc_gt0
+    all_gt0[all_gt0 > 0] <- 1
+    all_gt0
+  })
+  named_out(lcb, snms)
+})
+names(lc_union) <- lev_vec
+# plot(lc_union$f50$globmu)
+# plot(gti_agg$f50$gti)
+
+# selection variables
+nms1 <- c(colnames(ctabo), "wgt")
+nms2 <- colnames(ctabo)
+# bvals <- names(box_stats(sample(1:100, 200, replace = TRUE)))
+bvals <- "mu"
+
+dang <- Sys.time()
+  #print(paste("..", x))
+cb_stats <- lapply(lev_vec, function(x) {  # level
+  #x <- lev_vec[4]
+  print(paste(".", x))
+  #levr <- names(c_pct_diff[[x]])
+  l1 <- lapply(names(c_pct_diff[[x]]), function(y) {
+    #y <- levr[1]
+    print(paste("...", y))
+    lc <- c_pct_diff[[x]][[y]]
+    lcmask <- lc_union[[x]][[y]]
+
+    # stack raster bins, cropland bins, and landcover set
+    s <- stack(list("bin" = gti_bins[[x]], "wgt" = area_wgts[[x]][[1]], 
+                    "mask" = lc_union[[x]][[y]], c_pct_diff[[x]][[y]]))
+    DT <- na.omit(agroEcoTradeoff::as.data.table.raster(s))
+    setkey(DT, "bin")
+    # indexing columns
+    #front1 <- data.table("bin" = rep("all", 6), "N" = rep(nrow(DT), 6))
+    #front2 <- data.table("bin" = rep("abs", 6), "N" = rep(nrow(DT), 6))
+    # stats, overall means
+    #m <- cbind(bvals, DT[mask == 1, lapply(.SD, bfn, wgt), .SDcols = nms1])
+    #ma <- cbind(bvals, DT[mask == 1, lapply(.SD, bfna, wgt), .SDcols = nms1])
+    # stats per bin
+    #b <- cbind(bvals, 
+    #           DT[mask == 1, lapply(.SD, bfn, wgt), by = bin, .SDcols = nms1])
+    #ba <- cbind(bvals, 
+    #            DT[mask == 1, lapply(.SD, bfna, wgt), by = bin, .SDcols = nms1])    
+    # output
+    #odt <- rbind(cbind(front1, mu), cbind(front2, muabs), 
+    #             binl[bin][, bin := paste0("a", bin)], 
+    #             binl[binabs][, bin := paste0("b", bin)])[, wgt := NULL]
+
+    fr1 <- data.table("bvals" = c("m", "m0", "ma", "ma0"))
+    fr2 <- data.table("bin" = "all", "N" = nrow(DT))
+    binl <- DT[, .N, by = bin]  # n obs per bin
+    a <- round(rbind(DT[mask == 1, lapply(.SD, wm, wgt), .SDcols = nms1], 
+                     DT[, lapply(.SD, wm, wgt), .SDcols = nms1], 
+                     DT[mask == 1, lapply(.SD, wma, wgt), .SDcols = nms1],
+                     DT[, lapply(.SD, wma, wgt), .SDcols = nms1]), 2)
+    dtl <- list(DT[mask == 1, lapply(.SD, wm, wgt), by = bin, .SDcols = nms1],
+                DT[, lapply(.SD, wm, wgt), by = bin, .SDcols = nms1], 
+                DT[mask == 1, lapply(.SD, wma, wgt), by = bin, .SDcols = nms1], 
+                DT[, lapply(.SD, wma, wgt), by = bin, .SDcols = nms1])
+    b <- rbindlist(lapply(1:4, function(x) cbind(fr1[x], round(dtl[[x]], 2))))
+    setkey(b, "bin")
+    odt <- rbind(cbind(fr1, rbind(cbind(fr2, a))), binl[b])
+  })
+  named_out(l1, names(c_pct_diff[[x]]))
+})      
+dut <- Sys.time() - dang  # 3.65 minutes (vs 33 in earlier incarnation), 
+# 24 sec for means only
+names(cb_stats) <- lev_vec
+#save(cb_stats, file = "external/ext_data/carbon_bias_tables.rda")
+save(cb_stats, file = "external/ext_data/carbon_bias_tables_mus.rda")
+
+extract_stat <- function(namevec, stats, i1, i2) {
+  estats <- do.call(rbindlist(lapply(namevec, function(i) {
+    data.table(sapply(stats, function(x) {
+      unlist(sapply(x, function(y) {
+        v <- y[bin == i1 & bvals == i2, i, with = FALSE]
+        if(nrow(v) == 0) v <- rbindlist(list(v, as.list(NA)))
+        v
+      }))
+    }))
+  }))
+  out <- cbind(do.call(rbind, strsplit(rownames(estats), "\\.")), 
+               data.table(round(estats, 2)))
+  setnames(out, old = names(mu), new = )
+  return(out)
+}
+N <- function(namevec, stats, i1, i2) {
+  estats <- data.table(do.call(rbind, lapply(namevec, function(i) {
+    o <- sapply(stats, function(x) {
+      sapply(x, function(y) y[bin == i1 & bvals == i2, N][1])
+    })
+  })))
+  return(estats)
+}
+```
+<a href="#top">Back to top</a>
+
+#### Plot means against scale, supplemental version 
+```{r, eval=FALSE}
+a <- extract_stat(LC, cb_stats, "all", "m")  # mean abs crop only
+b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+setcolorder(a, c(2:1, 3:ncol(a)))
+mu <- rbind(a, b)
+
+a <- extract_stat(LC, cb_stats, "all", "ma")  # mean abs crop only
+b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+setcolorder(b, c(2:1, 3:ncol(b)))
+mua <- rbind(a, b)
+
+a <- extract_stat(LC, cb_stats, "all", "m0")  # mean abs crop only
+b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+setcolorder(b, c(2:1, 3:ncol(b)))
+mu0 <- rbind(a, b)
+
+a <- extract_stat(LC, cb_stats, "all", "ma0")  # mean abs crop only
+b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+setcolorder(b, c(2:1, 3:ncol(b)))
+mu0a <- rbind(a, b)
+
+#yr <- range(all_l[, 3:8, with = FALSE], all_u[, 3:8, with = FALSE])
+# yl <- range(all_mus[, 3:8, with = FALSE])
+
+cols <- c("red", "orange3", "green4", "blue")
+lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
+pnms <- c("Forest", "Secondary", "Shrubland", "Grassland", "Sparse")
+lw <- 1.5
+
+yl <- range(mu[, 3:8, with = FALSE], mua[, 3:8, with = FALSE])
+yl <- round(yl / 10) * 10
+yax <- seq(yl[1], yl[2], 10)
+xax <- seq(1, 6, 5 / (length(yax) - 1))
+
+# Version 1 of mean carbon bias against scale - probably for supplementals, if 
+# retained
+pdf(full_path(p_fig, "carbon_veg_scale.pdf"), height = 7, width = 5)
+par(mfrow = c(3, 2), mar = c(1, 0, 1, 1.25), mgp = c(1, 0.5, 0), tcl = -0.3,  
+    oma = c(6, 4, 1, 0))
+for(f in 1:length(LC)) {
+  plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+       ylab = "", xlab = "", 
+       panel.first = grid(NA, length(yax), lwd = 1, lty = 1))
+  #grid(NA, NULL, lwd = 1, lty = 1)
+  mtext(side = 3, text = pnms[f], line = 0)
+  if(f == 5) axis(1, at = 1:6, labels = c(1, fact))
+  if(f %in% c(1, 3, 5)) axis(2, at = yax, labels = yax, las = 2)
+  if(f %in% c(2, 4)) axis(2, at = yax, labels = rep("", length(yax)), las = 2)
+  for(i in 1:4) {
+    lines(1:6, mu[V1 == snms[i] & V2 == LC[f], lev_vec, with = FALSE], 
+          col = cols[i], lty = 2, lwd = lw)
+  }
+  for(i in 1:4) {
+    lines(1:6, mua[V1 == snms[i] & V2 == LC[f], lev_vec, with = FALSE], 
+          col = cols[i], lty = 1, lwd = lw)
+  }
+}
+plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "", 
+     panel.first = grid(NA, length(yax), lwd = 1, lty = 1))
+for(i in 1:4) {
+ lines(1:6, all_mu[V1 == snms[i], lev_vec, with = FALSE], col = cols[i], 
+       lty = 2, lwd = lw)
+}
+for(i in 1:4) {
+  lines(1:6, all_mua[V1 == snms[i], lev_vec, with = FALSE], col = cols[i],
+        lty = 1, lwd = lw)
+}
+mtext(side = 3, text = "Vegetation mean", line = 0)
+axis(1, at = 1:6, labels = c(1, fact))
+axis(2, at = yax, labels = rep("", length(yax)), las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 1.5)
+mtext(side = 2, text = "% bias", outer = TRUE, line = 2)
+par(xpd = NA)
+x <- grconvertX(0.7, from = "ndc", to = "user")
+y <- grconvertY(0.075, from = "ndc", to = "user")
+legend(x = x, y = y, legend = c("actual bias", "absolute bias"), lty = c(2, 1), 
+       lwd = 2, bty = "n", cex = 1.2)
+x <- grconvertX(0.1, from = "ndc", to = "user")
+y <- grconvertY(0.11, from = "ndc", to = "user")
+legend(x = x, y = y, legend = mcap, lty = 1, col = cols, lwd = 2, bty = "n",
+       cex = 1.2)
+dev.off()
+
+# Statistics from non-croplands areas included in means, for supplementals
+nmrepl <- cbind(snms, lcnms)
+mu0[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mu0a[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mu0[, V1 := nmrepl[match(V1, nmrepl), 2]]
+mu0a[, V1 := nmrepl[match(V1, nmrepl), 2]]
+mu0s <- rbind(cbind(stat = "actual", mu0), cbind(stat = "absolute", mu0a))
+mu0s_xtab <- xtable::xtable(mu0s)
+print(mu0s_xtab, type = "latex", file = "paper/figures/C-bias-mean-all-SA.tex")
+
+# Statistics from non-croplands areas included in means, for supplementals
+nmrepl <- cbind(snms, lcnms)
+mu[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mua[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mu[, V1 := nmrepl[match(V1, nmrepl), 2]]
+mua[, V1 := nmrepl[match(V1, nmrepl), 2]]
+mus <- rbind(cbind(stat = "actual", mu), cbind(stat = "absolute", mua))
+mus_xtab <- xtable::xtable(mus)
+print(mu0s_xtab, type = "latex", 
+      file = "paper/figures/C-bias-mean-cropland-SA.tex")
+
+```
+<a href="#top">Back to top</a>
+
+#### Main text version of mean carbon bias against scale
+```{r, eval=FALSE}
+alph <- c(225, 75)
+#Ccols <- brewer.pal(n = 7, name = "Greys")
+LC2 <- c(LC[-grep("second|grass", LC)], "all")
+# xax <- seq(0, 7, 6 / (length(yax) - 2)); length(xax)
+x <- c(0, 3, 6, 9, 12, 15)
+w <- 3 / 8
+xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
+#x + xo[j] + o[k]
+xa <- sapply(x, function(x) x + xo)
+mutype <- c("mua", "mu")
+
+o <- c(0, w)
+pchs <- c("*", "+", "o", "-")
+cx <- c(1, 0.7, 0.6, 1)
+g1 <- "grey90"
+# pchShow()
+# ctabo
+xl <- c(-0.5, 18)
+yl <- c(-60, 70)
+shd <- c(4.5, 10.5, 16.5)
+
+pdf(full_path(p_fig, "carbon_veg_scale2.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")#,
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+grid(NA, 13, lwd = 1, lty = 1)
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey50")
+for(i in 1:length(lev_vec)) {
+  for(j in 1:length(snms)) {
+    for(k in 1:length(mutype)) {
+      dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
+      pcol <- makeTransparent(cols[j], alpha = alph[k])
+      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
+      for(z in 1:length(LC2)) {
+        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
+               pch = pchs[z], type = "pt")
+      }
+    }
+  }
+}
+axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "% bias", outer = TRUE, line = 1)
+legend(x = 12.4, y = -40, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = -40, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = -42, labels = "Absolute", srt = 45, adj = c(0, 0), cex = 0.8)
+text(x = 12.2, y = -42, labels = "Actual", srt = 45, adj = c(0, 0), cex = 0.8)
+legend(x = 12.1, y = 65, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
+       pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)
 dev.off()
@@
 ```
+<a href="#top">Back to top</a>
 
-2. Convert them to data.tables and remove NA data
+```{r, eval=FALSE, echo=FALSE}
+# Plot means against cropland density (to possibly be run - for now, 
+# too much information)
+mub <- do.call(rbind, lapply(1:20, function(x) {
+  cbind("bin" = x, extract_stat(LC, cb_stats, x, "m"))
+}))
+muba <- do.call(rbind, lapply(1:20, function(x) {
+  cbind("bin" = x, extract_stat(LC, cb_stats, x, "ma"))
+}))
+n <- do.call(rbind, lapply(1:20, function(x) {
+  cbind("bin" = x, N(LC, cb_stats, x, "m"))
+}))
+#n[, mean(f50), by = bin]
 
-```{r, eval=FALSE}
-# GTI
-# gti0 <- 1 - gti / 100  # non cropland fractions for GTI data
-# gti_c <- gti / 100 * ctabo$crops[1]  # GTI cropland carbon
-gti_dt <- agroEcoTradeoffs::as.data.table.raster(gti / 100, xy = TRUE)
-gti_dt[, id:= 1:.N]
-setkeyv(gti_dt, c("x", "y"))
-setnames(gti_dt, "cover2011sum_mask", "fA")
+pp <- list("x" = c(-2, 21), "y" = c(0, 0), 
+           "x21" = list(c(0, 9), c(9, 21)),
+           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), 
+                       seq(10, 100, 10)), 
+           "y1lwd" = list(2, 1),  
+           "y2lwd" = 0.5, 
+           "y2lc" = list(c("black", rep("grey80", 4), "black", rep("grey80", 4), 
+                           rep("grey80", 4), "black", rep("grey80", 4),"black"),
+                         rep("grey80", 10)),
+           "yl" = list(c(-100, 100), c(0, 100)), 
+           "yl2" = c(-100, 100), "xl" = c(1, 20),
+           "xaxl" = unlist(lapply(seq(5, 100, 10), function(x) c(x, " "))),
+           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), 
+                            " "),
+           "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
+           "ycol" = c("grey70"))  # plot parameters
 
-lc_dt <- lapply(1:nlayers(lc_list), function(x) {
-  DT <- agroEcoTradeoffs::as.data.table.raster(lc_list[[x]] / 100, xy = TRUE)
-  DT[, id := 1:.N]
-  setkeyv(DT, c("x", "y"))
-  setnames(DT, names(lc_list[[x]]), "fA")
-  #DT <- DT[!is.na(fA), ]
-})
+yax <- seq(-100, 100, 10)
+xax <- seq(0, 22, 20 / (length(yax) - 2))
 
-# find NAs common to all datasets
-na1 <- gti_dt[is.na(fA)]
-na2 <- rbind(na1, do.call(rbind, lapply(lc_dt, function(x) x[is.na(fA)])))
-setkey(na2)
-na2 <- na2[!duplicated(na2)]
-nrow(na2[unique(id)]) == nrow(na2)
+pdf(full_path(p_fig, "carbon_biases_1-100km_v1.pdf"), width = 7.5, height = 7)
+par(mfrow = c(6, 4), mar = c(0.5, 0, 0.5, 0.5), mgp = c(1, 0.5, 0), tcl = -0.3,  
+    oma = c(6, 4, 1, 0))
+for(f in 1:length(LC)) {
+  for(j in c(1, 4:6)) {#j <- 1; f <- 1
+      plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+       ylab = "", xlab = "", 
+       panel.first = grid(NA, length(yax) - 1, lwd = 1, lty = 1))
+      lines(pp$x, pp$y, lwd = 1)
+    for(i in 1:4) {
+      mub[V1 == snms[i] & V2 == LC[f], 
+          lines(get(lev_vec[j]), col = cols[i], lty = 2, lwd = lw)]
+      muba[V1 == snms[i] & V2 == LC[f], 
+           lines(get(lev_vec[j]), col = cols[i], lty = 1, lwd = lw)]
+    }
+  }
+}
+dev.off()
 
-# Remove all common NA values, to tidy up size
-gti_dt <- gti_dt[!na2]
-gti_dt[, id := NULL]
-lc_dt <- lapply(lc_dt, function(x) {
-  DT <- x[!na2]
-  DT[, id := NULL]
-})
+ys <- c(-100, 100)
+yax <- seq(ys[1], ys[2], 10)
+xax <- seq(0, 22, 20 / (length(yax) - 2))
 
-# plot(agroEcoTradeoffs::dt_to_raster(gti_dt, gti@crs)[[1]])
-```
-
-```{r}
-for(i in 1:(ncol(ctabo))) {
-  nm <- colnames(ctabo)[i]
-  for(j in 1) { #:3) {
-    nmv <- paste0(nm, "_", rownames(ctabo)[j])
-    gti_dt[, c(nmv) := carbon(fA, noncropC = ctabo[j, nm]), with = FALSE] 
+#yl <- list(c(-100, 100), c(0, 100))
+yl <- seq(yax[1], yax[length(yax)], 10)
+yll <- as.character(yl)
+yll[seq(2, length(yll) - 1, 2)] <- ""
+Ccols <- brewer.pal(n = 6, name = "Dark2") 
+xaxl = ifelse(seq(5, 100, 5) %in% c(5, 25, 50, 75, 100), seq(5, 100, 5), " ")
+x <- mub[V1 == snms[f] & V2 == LC[j], get(lev_vec[i])]
+y <- 100
+toohigh <- function(x, y, col = "black", yadj = 0, cex = 1.1) {
+  h <- which(abs(x) > abs(y))
+  if(length(h) > 0) {
+    xi <- h[which.max(h)]
+    text(xi, y = y + yadj, labels = round(x[xi]), cex = cex, col = col, 
+         xpd = NA)
   }
@@
 
-# To GTI
-for(i in 1:(ncol(ctabo))) {
-  nm <- colnames(ctabo)[i]
-  for(j in 1) { #:3) {
-    nmv <- paste0(nm, "_", rownames(ctabo)[j])
-    gti_dt[, c(nmv) := carbon(fA, noncropC = ctabo[j, nm]), with = FALSE] 
+pdf(full_path(p_fig, "carbon_biases_1-100km_v2.pdf"), width = 7.5, height = 7)
+par(mfrow = c(4, 4), mar = c(0.5, 0, 0.5, 0.5), mgp = c(1, 0.5, 0), tcl = -0.3,  
+    oma = c(6, 4, 1, 0))
+for(f in 1:length(snms)) {
+  for(i in c(1, 4:6)) {#i <- 1; j <- 1; f <- 2
+      plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+           ylab = "", xlab = "", 
+           panel.first = grid(NA, (length(yax) - 1) / 2, lwd = 1, lty = 1))
+      lines(pp$x, pp$y, lwd = 1)
+    for(j in 1:length(LC)) {
+      e <- parse(text = "V1 == snms[f] & V2 == LC[j]")
+      mub[eval(e), 
+          lines(get(lev_vec[i]), col = Ccols[j], lty = 2, lwd = lw)]
+
+      mub[V1 == snms[f] & V2 == LC[j], 
+          lines(get(lev_vec[i]), col = Ccols[j], lty = 2, lwd = lw)]
+      mub[V1 == snms[f] & V2 == LC[j], toohigh(get(lev_vec[i]), ys[1])]
+      muba[V1 == snms[f] & V2 == LC[j], 
+           lines(get(lev_vec[i]), col = Ccols[j], lty = 1, lwd = lw)]
+      axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
+           tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[1])
+      axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
+           tcl = -0.3, pos = pp$yl2[1])
+      if(f == length(snms)) {
+        axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
+             tcl = -0.3, pos = pp$yl2[1])
       }
-}
-
-# To landcover sets
-for(k in 1:length(lc_dt)) {
-  for(i in 1:(ncol(ctabo))) {
-    nm <- colnames(ctabo)[i]
-    for(j in 1) {#:3) {
-      nmv <- paste0(nm, "_", rownames(ctabo)[j])
-      lc_dt[[k]][, c(nmv) := carbon(fA, noncropC = ctabo[j, nm]), with = FALSE] 
+      if(i == 1) axis(2, at = yl, labels = yll, las = 2, mgp = c(1, 0.6, 0))
+      if(i == 1 & j == 1) {
+        mtext(lcnms[f], side = 2, line = 2.8, cex = 1, col = cols[f])
       }
@@
   }
+}
+dev.off()
+
+i1 <- "sa30"; i2 <- "forest"; s1 <- "f1"
+mub[V1 == i1 & V2 == i2, ]
+mub[V1 == i1 & V2 == i2, plot(bin, f1, type = "l")]
+mubin[V1 == i1 & V2 == i2, plot(bin, f1, type = "l")]
+```
+
+```{r, eval=FALSE, echo=FALSE}
+# Checks on data values (embedded code)
+# Carbon values and carbon biases - random selection of cells from c_tab_diff
+# compared to recalculated values for those points
+i <- c_pct_diff
+for(chk in list(c("modmu", "f25", "forest"), c("geow", "f10", "shrub"), 
+                c("sa30", "f50", "sparse"))) {
+  x <- chk[1] #names(i[[1]])[3]
+  y <- chk[2] #names(i)[4]
+  tp <- chk[3]
+  r <- i[[y]][[x]][[tp]]
+  rid <- which(values(Which(r != 0)))
+  sid <- sample(rid, 20)
+  v1 <- r[sid]  # value to check
+  sav <- gti_agg[[y]]$gti[sid]
+  lcv <- lc_agg[[y]][[x]][sid]
+  c1 <- carbon(sav, cropC = 5, noncropC = ctabo[1, tp])
+  c2 <- carbon(lcv, cropC = 5, noncropC = ctabo[1, tp])
+  #round((sav * 5) + (ctabo[1, tp] * (1 - sav)), 2) == c1
+  #round((lcv * 5) + (ctabo[1, tp] * (1 - lcv)), 2) == c2
+  print(all(v1 == (c1 - c2) / c1 * 100))
+}  # all true
+
+# Carbon bias by bins 
+# DT-derived values
+mu0a <- do.call(rbind, lapply(1:20, function(x) {
+  cbind("bin" = x, extract_stat(LC, cb_stats, x, "ma0"))
+}))
+cb_stats$f1$sa30
+cb_stats$f10$sa30[get("bin") == 19, ]
+
+# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
+# version is finding correct results
+for(chk in list(c("modmu", "f25", "forest"), c("geow", "f10", "shrub"), 
+                c("sa30", "f50", "sparse"))) {
+  x <- chk[1] #names(i[[1]])[3]
+  y <- chk[2] #names(i)[4]
+  tp <- chk[3]
+  print(paste(".", x, "..", y, "..", tp))
+  d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(j) {
+    #print(paste("......bin", j))
+    d2 <- do.call(rbind, lapply(tp, function(z) {#:length(i[[y]][[x]]),..
+      #print(paste("........variant", z))
+      l3 <- abs(i[[y]][[x]][[z]])  # i[[y]][[x]][[z]]
+      #print(paste("........vegetation:", names(l3)))
+      d3 <- do.call(rbind, lapply(1, function(k) { #:length(l3)...
+        rs <- gti_bins[[y]][[1]] == j #[[k]] == j    
+        rs[rs == 0] <- NA
+        o <- rs * l3 #[[k]]
+        w <- area_wgts[[y]][[1]] * rs
+        oo <- cbind.data.frame("lc" = rep(paste0(x, y, z), length(w)), 
+                               getValues(o), getValues(w))
+      }))
+      d3f <- d3[!is.na(d3[, 2]), ]
+    }))
+    if(nrow(d2) > 0) { #(length(i[[x]][[y]]) * 10)) {
+      qs <- c(nrow(d2), box_stats(d2[, 2], weighted = TRUE, 
+                                  weight.opts = list("weights" = d2[, 3])))
+    } else if((nrow(d2) > 0) & (nrow(d2) < length(i[[x]][[y]]) * 20)) {
+      qs <- c(rep(NA, 6), Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], 
+                                          na.rm = FALSE))
+    } else {
+      qs <- rep(NA, 7)
+    }
+  }))
+  #print(round(d1[, "mu"], 2) == mubin[V1 == x & V2 == tp, y, with = FALSE], 
+  #          na.rm = TRUE)
+  #print(all(round(d1[, "mu"], 2) == mubin[V1 == x & V2 == tp, y, with = FALSE], 
+  #          na.rm = TRUE))
+  print(all(round(d1[, "mu"], 2) == 
+              mu0a[V1 == x & V2 == tp, y, with = FALSE], na.rm = TRUE))
+}
+# all true
+
+# another check of DT-derived stats against a raster-calculated variant
+tst <- sapply(names(c_pct_diff)[1], function(x) {
+  lapply(names(c_pct_diff[[1]]), function(y) {
+    cellStats(c_pct_diff[[x]][[y]]$forest, mean)
+  })
+})
+mean(round(unlist(tst)), 2) == mean(round(mu0[V2 == "forest", f1]), 2)
+
+gti_gt0 <- Which(gti_agg$f1$gti > 0)
+lc_gt0 <- calc(stack(lapply(lc_agg$f1, function(x) Which(x > 0))), sum)
+all_gt0 <- gti_gt0 + lc_gt0
+all_gt0 <- all_gt0 / all_gt0
+
+filts <- sapply(names(c_pct_diff), function(x) {
+  gti_gt0 <- Which(gti_agg[[x]]$gti > 0)
+  lc_gt0 <- calc(stack(lapply(lc_agg[[x]], function(j) Which(j > 0))), sum)
+  all_gt0 <- gti_gt0 + lc_gt0
+  all_gt0 <- all_gt0 / all_gt0
+})
+
+tst <- sapply(names(c_pct_diff), function(x) {
+  lapply(names(c_pct_diff[[x]]), function(y) {
+    cellStats(filts[[x]] * c_pct_diff[[x]][[y]]$forest, mean)
+  })
+})
+
+tstabs <- sapply(names(c_pct_diff), function(x) {
+  lapply(names(c_pct_diff[[x]]), function(y) {
+    cellStats(abs(filts[[x]] * c_pct_diff[[x]][[y]]$forest), mean)
+  })
+})
+
+plot(c(1, 6), range(tst, tstabs), pch = "", xlab = )
+for(i in 1:nrow(tst)) lines(1:6, tst[i, ], lty = 2)
+for(i in 1:nrow(tstabs)) lines(1:6, tstabs[i, ], col = "red")
+
+cellStats(all_gt0 * c_pct_diff$f1$sa30$sparse, mean)
+cellStats(all_gt0 * c_pct_diff$f1$sa30$forest, median)
+cellStats(c_pct_diff$f1$sa30$forest, mean)
+cellStats(all_gt0 * c_pct_diff$f1$sa30$forest, mean)
+
+gti_
 
 ```
-
-```{r, eval=FALSE}
-# Actual difference
-cdiff_dt <- lapply(1:length(lc_dt), function(x) {
-  ind <- colnames(gti_dt)[4:ncol(gti_dt)]
-  DT <- copy(gti_dt)  # so gti_dt is not modified
-  DT2 <- copy(lc_dt[[x]])
-  DT[, fA := NULL]
-  DT[, `:=`(ind, Map('-', DT[, ind, with = FALSE], 
-                     DT2[, ind, with = FALSE])), with = FALSE]
-})
-
-cdiff_dts <- lapply(cdiff_dt, function(x) {
-  ind <- colnames(x)[3:ncol(x)]
-  DT <- copy(x)  # so gti_dt is not modified
-  DTo <- DT[, 1:2, with = FALSE]
-  DTo[, mx := rowMaxs(as.matrix(DT[, ind, with = FALSE]))]
-  DTo[, mu := rowMeans(DT[, ind, with = FALSE])]
-  DTo[, mn := rowMins(as.matrix(DT[, ind, with = FALSE]))]
-})
-
-perc_diff <- function(x, y) round((x - y) / x * 100, 1)
-
-# Difference as a percent
-cdiffpct_dt <- lapply(1:length(lc_dt), function(x) {
-  ind <- colnames(gti_dt)[4:ncol(gti_dt)]
-  DT <- copy(gti_dt)  # so gti_dt is not modified
-  DT2 <- copy(lc_dt[[x]])
-  DT[, fA := NULL]
-  DT[, `:=` (ind, Map(perc_diff, x = DT[, ind, with = FALSE], 
-                      y = DT2[, ind, with = FALSE])), with = FALSE]
-})
-
-```

### 2015-07-06 R/utilities.R

@@ -1,0 +1,81 @@
+#' GAM Prediction function
+#' @param model Model object
+#' @param newdat data.frame havng dummy values for predictors
+#' @return Vector of predicted values with confidence intervals, for curve 
+#' plotting
+#' @keywords internal
+#' @export
+predfunc <- function(model, newdat) {
+  pred <- predict(model, newdata = newdat, type = "response", se.fit = TRUE)
+  pdat <- transform(newdat, fit = pred$fit)
+  pdat <- transform(pdat, up = fit + (1.96 * pred$se.fit),
+                    lo = fit - (1.96 * pred$se.fit))
+  return(pdat)
+}
+
+#' Predict values from single GAM model term
+#' @param model Model object
+#' @param newdat data.frame havng dummy values for predictors
+#' @return Vector of predicted values with confidence intervals, for curve 
+#' plotting
+#' @keywords internal
+#' @export 
+termfunc <- function(model, newdat) {
+  pred <- predict(model, newdata = newdat, type = "terms", se.fit = TRUE)
+  pdat <- transform(newdat, fit = pred$fit[, 1] + coef(model)[1])
+  pdat <- transform(pdat, up = fit + (1.96 * pred$se.fit[, 1]),
+                    lo = fit - (1.96 * pred$se.fit[, 1]))
+  return(pdat)
+}
+
+#' Create arbitrary polygon on plot
+#' @param x coordinate on which to center polygon on x-axis
+#' @param y upper and lower coordinates for polygon on y-axis
+#' @param w polygon width in x dimension
+#' @param col fill color
+#' @param bcol border color
+#' @param lwd border width
+#' @return plotted polygon on previously called plot
+#' @keywords internal
+#' @export
+polyfunc2 <- function(x, y, w = 0.5, col, bcol, lwd = 1) {
+  x <- c((x - 0.5 * w), (x + 0.5 * w))
+  polygon(cbind.data.frame("x" = c(x, rev(x)), "y" = y[c(1, 1, 2, 2)]), 
+          col = col, lwd = lwd, border = bcol)
+}
+
+#' Make any color transparent
+#' @param someColor Color name, e.g. "red"
+#' @param alpha Transparency value, from 0-255
+#' @return Transparent color value
+#' @note Function found at 
+#' stackoverflow.com/questions/8047668/transparent-equivalent-of-given-color
+#' @keywords internal
+#' @export
[…]

### 2015-07-15 Rmd/carbon-bias.Rmd
much of the way through yield-bias.Rmd analysis. Added new function disaggregate_rast_list. Various tweaks and nudges.

@@
                     "mask" = lc_union[[x]][[y]], c_pct_diff[[x]][[y]]))
-    DT <- na.omit(agroEcoTradeoff::as.data.table.raster(s))
+    DT <- na.omit(as.data.table.raster(s))
     setkey(DT, "bin")
@@
 
-# Statistics from non-croplands areas included in means, for supplementals
-nmrepl <- cbind(snms, lcnms)
-mu0[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mu0a[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mu0[, V1 := nmrepl[match(V1, nmrepl), 2]]
-mu0a[, V1 := nmrepl[match(V1, nmrepl), 2]]
-mu0s <- rbind(cbind(stat = "actual", mu0), cbind(stat = "absolute", mu0a))
-mu0s_xtab <- xtable::xtable(mu0s)
-print(mu0s_xtab, type = "latex", file = "paper/figures/C-bias-mean-all-SA.tex")
-
-# Statistics from non-croplands areas included in means, for supplementals
-nmrepl <- cbind(snms, lcnms)
-mu[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mua[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mu[, V1 := nmrepl[match(V1, nmrepl), 2]]
-mua[, V1 := nmrepl[match(V1, nmrepl), 2]]
-mus <- rbind(cbind(stat = "actual", mu), cbind(stat = "absolute", mua))
-mus_xtab <- xtable::xtable(mus)
-print(mu0s_xtab, type = "latex", 
-      file = "paper/figures/C-bias-mean-cropland-SA.tex")
-
 ```
@@
 ```{r, eval=FALSE}
-alph <- c(225, 75)
+alph <- c(225, 40)
 #Ccols <- brewer.pal(n = 7, name = "Greys")
@@
 <a href="#top">Back to top</a>
+
+#### Supplementary mean bias tables
+```{r, eval=FALSE}
+# Statistics from non-croplands areas included in means, for supplementals
+nmrepl <- cbind(snms, lcnms)
+mu0[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mu0a[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mu0[, V1 := nmrepl[match(V1, nmrepl), 2]]
+mu0a[, V1 := nmrepl[match(V1, nmrepl), 2]]
+mu0s <- rbind(cbind(stat = "actual", mu0), cbind(stat = "absolute", mu0a))
+mu0s_xtab <- xtable::xtable(mu0s)
+print(mu0s_xtab, type = "latex", file = "paper/figures/C-bias-mean-all-SA.tex")
+
+# Statistics from non-croplands areas included in means, for supplementals
+nmrepl <- cbind(snms, lcnms)
+mu[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mua[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mu[, V1 := nmrepl[match(V1, nmrepl), 2]]
+mua[, V1 := nmrepl[match(V1, nmrepl), 2]]
+mus <- rbind(cbind(stat = "actual", mu), cbind(stat = "absolute", mua))
+mus_xtab <- xtable::xtable(mus)
+print(mu0s_xtab, type = "latex", 
+      file = "paper/figures/C-bias-mean-cropland-SA.tex")
+```
 
@@
 
-gti_
-
 ```

### 2015-07-15 Rmd/yield-bias.Rmd
much of the way through yield-bias.Rmd analysis. Added new function disaggregate_rast_list. Various tweaks and nudges.

---
title: "Yield bias"
author: "Lyndon Estes"
date: "July 6, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Methods to examine

Ramankutty et al's (2008) procedure for calculating cropland area.

1. Divide provincial cropland area estimates by total area in each province. 
2. Linear model calculates cropland proportions from model of all satellite landcover classes, creating a 5 minute gridded estimate of cropland in each pixel.
3. These grids are aggregated to administrative level (provinces in our case), and then compared with the provincial statistics. A correction factor is then applied, and _pycnophylactically_ interpolated, which is then used to adjust the gridded estimates.

$$ fcropland_{x,y} = ucf_{x,y} * cf_{x,y}$$

Where ucf is the correction factor and cf is the step 1 modeled cropland fraction estimate for a given pixel (x, y).

Monfreda et al (2008) then use $fcropland_{x,y}$ in a three step process to calculate the area in administrative units for each crop type, in the following steps:

1. They divide $fcropland_{x,y}$ by the area of the administrative unit to calculate the fraction of each unit taken up by a particular crop.

2. They then disaggregate this fraction back onto the gridded data as follows:

$$ fcrop_{x,y} = fcropland_{x,y}\frac{croparea_{pu}}{cropland_{pu}}$$

They assign yields for each crop--obtained from the finest administrative unit available, uniformly across pixels having a non-zero proportion of that crop within the administrative unit.

They rescaled both area and yield within sub-administrative units to match country totals derived from FAOStat from 1997-2003.

## Key Questions

1. How does these processes create bias if they are not aggregated by the administrative units to which they are scaled? i.e. If you aggregate by grid cells rather than by administrative units, how does this affect estimates of total production (and therefore things such as calories to be gained by closing yield gaps)

2. Do any studies use it in this particular way?

    + Foley et al (2011) estimate amount of calories that could be gained by closing yield gaps.  
    + Find one or two others - maybe Roy with yield trends, or Mueller et al 2012.

# Data

1. Use SA provinces with GTI to create the amount of agricultural land estimates per province, akin to provincial statistics used by Ramankutty et al (2008). Note: they mentioned SA has 11 administrative units they used, but the publications they cited ([here](http://www.statssa.gov.za/publications/DiscussLandAccounts/DiscussLandAccounts.pdf)) has just 9 provinces, so maybe they mistakenly counted Lesotho and Swaziland?

```{r, eval=FALSE}
library(SAcropland)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
#p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")

# SA provinces
prov <- readOGR("external/ext_data/provinces.sqlite", layer = "provinces")
prov <- prov[prov$id_1 != 9, ]  # remove Princeton Edward Islands
prov[prov$id_1 == 10, "id_1"] <- 9 # reset WC to id 10
load("external/ext_data/MZshapes.Rdata")

# cropland data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100)

# plot(lc_list[[1]])
# plot(gti)
# tst <- raster("external/ext_data/geowikisa_masked.tif")  # check
# plot(round(tst / 100 - lc_list[[4]], 4)) # okay

# Create namask and area_wgts, removing all NA areas across datasets
sumna <- function(x) sum(x, na.rm = FALSE)
namask <- calc(stack(gti, stack(lc_list)), sumna)
namask[namask > 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(is.na(gti), sum); cellStats(namask2, sum)
fact <- c(5, 10, 25, 50, 100)
# area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
# area_wgts <- lapply(area_wgts, function(x) x[[1]])  # unlist on inner loop
# plot(area_wgts$f25)

```

# Analyses
## Cropland fractions

### Get provincial $cf$ factors from 1 km GTI dataset
```{r, eval=FALSE}
# rasterize provinces and stack with gti raster, convert to data.table
provr <- rasterize(prov, y = gti, field = "id_1", 
                   filename = "external/ext_data/provinces.tif")
# provr <- raster("external/ext_data/provinces.tif")
provr <- mask(provr, namask)
gtis <- stack(list("prov" = provr, "f" = gti))

# calculate fractions using raster::extract with provinces, to compare speed
prov_est <- extract(gti, prov, progress = "text")
prov_estNA <- lapply(prov_est, function(x) x[!is.na(x)])
cftest <- cbind("carea" = sapply(prov_estNA, sum), 
                "parea" = sapply(prov_estNA, length), 
                "cf" = sapply(prov_estNA, sum) / sapply(prov_estNA, length))

# data.table calculations
gti_dt <- na.omit(as.data.table.raster(gtis, xy = TRUE))
setkey(gti_dt, prov)
cf <- gti_dt[, list("carea" = sum(f), "parea" = length(f), 
                    "cf" = sum(f) / length(f)), by = prov]
# cftest - cf[, 2:4, with = FALSE]  # close, but diff caused by few NAs in provr

# data.table approach is much more efficient

```
<a href="#top">Back to top</a>

### Calculate cropland fractions from 1 km landcover sets

Derived at 1 km resolution rather 10 km (as in Ramankutty et al, 2008), using data.tables to calculate provincial-level correction factors for each
```{r, eval=FALSE}
lcs <- stack(list("prov" = provr, stack(lc_list)))
lcs_dt <- na.omit(as.data.table.raster(lcs, xy = TRUE))
setkey(lcs_dt, prov)
fc <- function(x) sum(x) / length(x)  # fraction functions
lc_cf <- lapply(list(sum, length, fc), function(x) {
  lcs_dt[, lapply(.SD, x), by = prov, .SDcols = snms]
})
names(lc_cf) <- colnames(cf)[-1]
# lcs_dt[, .N, by = prov] == gti_dt[, .N, by = prov]

# correction factors
pcf <- cbind("prov" = prov$id_1, 
             sapply(snms, function(x) cf$cf / lc_cf$cf[, get(x)]))
pcf <- data.table(pcf)

# tst1 <- disaggregate(tst, fact = 10)
# namask <- raster("external/ext_data/namask.tif")
# tst1m <- mask(crop(tst1, namask), mask = namask)
# plot(tst1m - gti$cover2011sum_mask)

```
<a href="#top">Back to top</a>

### Re-scale landcover cropland fraction using $cf$ factors
```{r, eval=FALSE}
# multiply each lc fraction value by correction factor for the particular 
# province
lcs_dta <- copy(lcs_dt)
for(j in snms) {
  for(g in pcf$prov) {
    lcs_dta[prov == g, j := (get(j) * pcf[prov == g, get(j)]), with = FALSE] 
  }
}

# check to see if DT syntax correct
for(i in 1:9) {
  for(j in snms) {
    print(all(lcs_dta[prov == i, get(j)][1:10] == 
                (lcs_dt[prov == i, get(j)] * pcf[i, get(j)])[1:10]))
  }
}

# check to see if the factors resulted in same total area per province
fchk <- lcs_dta[, lapply(.SD, sum), by = prov, .SDcols = snms]
par(mfrow = c(2, 2), mar = rep(1, 4))
for(i in snms) plot(cf[, carea], fchk[, get(i)])  # yup

# how many pixels have fraction > 1 
for(i in 1:9) {
  for(j in snms) {
    print(paste(j, ":", i, ":", 
                  lcs_dta[prov == i, length(which(get(j) > 1))] / 
                            lcs_dta[prov == i, .N]))
  }
}  # several in globcover, modis, and geow

# Adjust these to equal 1 
lcs_dta2 <- copy(lcs_dta)

# function to adjust fractions to fall within 1 while maintaining statistics
force1 <- function(x) {
  a <- x
  while(any(a > 1)) {
    reall <- sum(a[a > 1] - 1)  # total of parts of x > 1, to reallocate
    ind <- which(a < 1)  # parts of x less than 1
    a[a > 1] <- 1
    af <- (reall + sum(a[a < 1])) / sum(a[a < 1])
    a[a < 1] <- a[a < 1] * af
  }
  return(a)
}
# check if works on modis set
hist(lcs_dta[prov == 8, modmu])
hist(force1(lcs_dta[prov == 8, modmu]))
round(sum(force1(lcs_dta[prov == 8, modmu])), 2) == 
  round(sum(lcs_dta[prov == 8, modmu]), 2)

for(j in snms) {
  for(g in pcf$prov) {
    lcs_dta2[prov == g, j := force1(get(j)), with = FALSE] 
  }
}

# Check to see if all fractions <= 1 and that provincial sums are equal
for(i in 1:9) {
  for(j in snms) {
    print(paste(j, ":", i, ":", 
                  lcs_dta2[prov == i, length(which(get(j) > 1))] / 
                            lcs_dta[prov == i, .N], ":", 
                round(lcs_dta2[prov == i, sum(get(j))], 4) == 
                  round(lcs_dta[prov == i, sum(get(j))], 4)))
  }
}  # check

# convert back to rasters
lc_adj <- dt_to_raster(lcs_dta2, CRSobj = sa.shp@proj4string)
lc_adj <- dropLayer(lc_adj, i = 1)
```

### Aggregate adjusted rasters, compare cropland fractions
```{r, eval=FALSE}
tl <- lapply(1:4, function(x) lc_adj[[x]])
names(tl) <- snms
lc_agg <- aggregate_rast_list(fact, tl) # landcover rasters 
rm(tl)
gtic <- crop(gti, lc_adj$sa30)
gti_agg <- aggregate_rast_list(fact, list("gti" = gtic))  # GTI rasters

# compare extents to make sure no offsets - look at in QGIS
# writeRaster(gti_agg$f1$gti, filename = "external/ext_data/test/gti_crop.tif")
# writeRaster(lc_agg$f1$sa30, filename = "external/ext_data/test/sa30_crop.tif")
# looks fine

# tst <- gti_agg$f5$gti - lc_agg$f5$sa30
# plot(tst * 100)

#pct_diff <- function(x, y) (x - y) / x * 100
cf_pct_diff <- lapply(1:length(gti_agg), function(x) {
  dif <- lapply(1:length(lc_agg[[x]]), function(y) {
    d <- gti_agg[[x]][[1]] - lc_agg[[x]][[y]]
  })
  named_out(dif, snms)
})
names(cf_pct_diff) <- names(lc_agg)

cf_pct_diff <- lapply(snms, function(x) {
  dif <- lapply(names(lc_agg), function(y) {
    d <- gti_agg[[y]][[1]] - lc_agg[[y]][[x]]
  })
  named_out(dif, names(lc_agg))
})
names(cf_pct_diff) <- snms

# disaggregate for plotting
namaskc <- crop(namask, gtic)
lev <- names(cf_pct_diff[[1]])
disagg <- disaggregate_rast_list(snms, lev, cf_pct_diff, crop(namask, gtic))

stats <- lapply(disagg, function(x) {
  sapply(x, function(y) {
    c(cellStats(y * 100, mean), quantile(y * 100, seq(0, 1, 0.05)))
  })
})
```
<a href="#top">Back to top</a>

Plot for supplementary data
```{r, eval=FALSE}
lims <- c(ceiling(min(sapply(stats, function(x) x[3, ]))), 
          floor(max(sapply(stats, function(x) x[21, ]))))
rng <- range(sapply(stats, range))
brks <- c(rng[1], -50, -30, -20, -10, -5, -1, 1, 5, 10, 20, 30, 50, rng[2])
cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)

legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev <- names(disagg[[1]])[-c(2:3)]
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "cropland_adj_bias_map.pdf"), height = 6, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]] * 100, add = TRUE, col = cols, 
         breaks = brks, legend = FALSE)
  if(j == 1) mtext(mcap[i], side = 2, line = 1, cex = cx)
  if(i == 1) mtext(lev2[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext, 
            legend.vals = round(brks), 
            longdims = c(0.2, 0.8), shortdims = c(0.06, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0, -0.5)), 
            cex.val = cx, textcol = lcol, bordercol = lcol)
dev.off()
```
<a href="#top">Back to top</a>

## Yield disaggregation

Following Monfreda et al's (2008) methods. They appear to have used the 2002 district census, so we will use the more recent 2007 census.

### Process magisterial districts and census data
```{r, eval=FALSE}
# Load in yield dataset and magisterial district polygons
load(full_path(p_data2, "ZAF_adm2.Rdata"))  # magisterial districts
yld <- fread(full_path(p_data, "maize_wheat_2007.csv"))

# Transform magisterial districts, remove Prince Edward Islands
md <- spTransform(gadm, CRSobj = sa.shp@proj4string)
md <- md[md$ID_2 != 313, ]
# round(mean(rgeos::gArea(md, byid = TRUE) / 1000000))
md@data <- md@data[, c("PID", "ID_2", "NAME_1", "NAME_2")]  # trim down columns
# fix a few names that occur more than once to make unique for merging properly
md@data[md$ID_2 == 43, "NAME_2"] <- "MiddelburgEC"
md@data[md$ID_2 == 143, "NAME_2"] <- "RichmondKZ"
md@data[md$ID_2 == 86, "NAME_2"] <- "HeidelbergG"

# check names in two datasets
# match(md@data$NAME_2, yld$district)
# match(yld$district, md@data$NAME_2)
# yld$district[which(!yld$district %in% md@data$NAME_2)]
# sort(md@data$NAME_2[which(!md@data$NAME_2 %in% yld$district)])

# some offline fixing of names ensues
# write.csv(md@data, file = "external/ext_data/mdrect.csv")
# writeOGR(md, dsn = "external/ext_data/mdcheck.sqlite", layer = "mdcheck", 
#          driver = "SQLite")

# reread fixed data, merge a few districts' data where there are synonyms or 1
# district subsumed by another--sum yields, because these districts were 
# probably broken into smaller pieces
# fix same names for merging properly
yld[district == "Middelburg" & province == "Eastern Cape", 
    district := "MiddelburgEC"]
yld[district == "Richmond" & province == "KwaZulu-Natal", 
    district := "RichmondKZ"]
yld[district == "Heidelberg" & province == "Gauteng", 
    district := "HeidelbergG"]

# sumna <- function(x) sum(x, na.rm = TRUE)
mrg <- rbindlist(lapply(unique(yld$merge)[-1], function(x) {
  yld[merge == x, lapply(.SD, sumna), .SDcols = grep("mz|wh", names(yld))]
}))
nms <- sapply(unique(yld$merge)[-1], function(x) {
  nm <- yld[merge == x, district]
})
mnms <- unname(sapply(nms, function(x) x[x %in% md$NAME_2][1]))

keep <- names(yld)[-c(2:5)]
yld2 <- rbind(yld[!district %in% unname(unlist(nms)), keep, with = FALSE], 
              cbind("district" = mnms, mrg))  # adjusted yield dataset

# merge with magisterial district data
mdyld <- sp::merge(md@data, yld2, by.x = "NAME_2", by.y = "district", 
                   all.x = TRUE)  # merge
mdyld <- mdyld[match(mdyld$NAME_2, md@data$NAME_2), ]  # reorder rows correctly
nrow(mdyld) == nrow(md@data)  # check

# calculate yields and total planted ha
mdyld$mzha <- rowSums(mdyld[, grep("mz_ha", colnames(mdyld))])
mdyld$mzyld <- round(rowSums(mdyld[, grep("mz_pr", colnames(mdyld))]) /
                       mdyld$mzha, 1)
mdyld$whha <- rowSums(mdyld[, grep("wh_ha", colnames(mdyld))])
mdyld$whyld <- round(rowSums(mdyld[, grep("wh_pr", colnames(mdyld))]) / 
                       mdyld$whha, 1)

# join to md data
m <- mdyld[match(md$ID_2, mdyld$ID_2), 
           c("ID_2", "NAME_2", "mzha", "mzyld", "whha", "whyld")]
colnames(m)[1:2] <- c("id", "name")
md@data <- cbind(md@data, m)  
all(md$NAME_2 == md$name); all(md$ID_2 == md$id)
md@data <- md@data[, -c(2:4)]

# plot to make sure districts didn't get reordered at all
par(mar = rep(0, 4))
plot(sa.shp)
plot(md, add = TRUE)
plot(md[md$id == 214, ], col = "red", add = TRUE)
plot(md[md$id == 327, ], col = "red", add = TRUE)
plot(md[md$id == 263, ], col = "red", add = TRUE)
plot(md[md$name == "Barberton", ], col = "red", add = TRUE)
```
<a href="#top">Back to top</a>

### Calculate individual crop fractions

Rasterize MDs and calculate individual crop fractions (for maize), for both GTI and adjusted LC cropland fractions
```{r, eval = FALSE}
mdr <- rasterize(md, provr, field = "id",  
                 filename = "external/ext_data/mag_dist.tif", overwrite = TRUE)
# mdr <- raster("external/ext_data/mag_dist.tif")
mds <- stack(list("md" = crop(mdr, gti_agg$f1$gti), "gti" = gti_agg$f1$gti,
                  lc_adj))  # use gti and lc_adj  
# v <- values((lc_agg$f1$modmu > 1) * 1); sum(v[!is.na(v)])

# data.table calculations
md_dt <- as.data.table.raster(mds, xy = TRUE)
setkey(md_dt, md)
mdarea <- md_dt[, .N, by = md]  # how many km2 in each MD
setnames(mdarea, "N", "mdkm2") 
mdarea <- mdarea[!is.na(md)]  # remove NAs  
setkey(mdarea, md)  
md_dt <- na.omit(md_dt)  # remove NAs from fraction data.tables
varea <- md_dt[, .N, by = md]  # area of non-NA data in MDs
setnames(varea, "N", "vkm2")
setkey(varea, md)
# mdcf <- md_dt[, list("carea" = round(sum(f), 2), "parea" = length(f), 
#                         "cf" = round(sum(f) / length(f), 4)), by = md]
mdva <- mdarea[varea][, fmd:= round(vkm2 / mdkm2, 2)]  # md valid area
# mdcf[, fmd := mdarea[varea][, round(vkm2 / mdkm2, 2)]]  # fraction non-NA in md

# yield data for MDs
ydt <- data.table(md@data[, c("id", "mzha", "mzyld", "whha", "whyld")])
setnames(ydt, "id", "md")
setkey(ydt, "md")
mdvay <- mdva[ydt][is.na(mzha), c("mzha", "mzyld") := 0]
mdvay[is.na(whha), c("whha", "whyld") := 0]
mdvay[, c("mdkm2", "vkm2") := NULL]
# mdcfy[, mfrac := round(((mzha * fmd) / 100) / carea, 4)]  # maize fraction

# calculate crop fractions for each dataset
md_dta <- copy(md_dt)
# md_dt[, lapply(.SD, mean), by = md, .SDcols = c("gti", snms)]
md_lcl <- lapply(c("gti", snms), function(j) {
    DT <- md_dta[, list("carea" = round(sum(get(j)), 2),  # crop area
                        "parea" = length(get(j)),  # non-NA area in MD
                        "cf" = round(sum(get(j)) / length(get(j)), 4)),
                   by = md]
    # calculate crop fraction, adjusting ha first by how much non-NA area there
    # is in MD
    DTy <- DT[mdvay][, mfrac := round(((mzha * fmd) / 100) / carea, 4)]
    DTy[is.na(mfrac), mfrac := 0]  # set mfrac to 0 in 0 cropland areas
    # print(length(which(DTy$mfrac > 1)))  # 5, 4, 15, 9, 4
    DTy[mfrac > 1, mfrac := 1]  # set to 1 - mostly missing data causing > 1
    DTy
})
names(md_lcl) <- c("gti", snms)
# lapply(md_lcl, function(x) x[, sum(mzha)])
#length(which(is.na(md_lcl$modmu$mfrac)))
#length(which(md_lcl$sa30$mfrac > 1))

par(mfrow = c(2, 3))
for(i in 1:5) hist(md_lcl[[i]]$mfrac)
for(i in 1:5) hist(md_lcl[[i]]$carea)
plot(md_lcl[[1]]$carea, md_lcl[[3]]$carea)
```
<a href="#top">Back to top</a>

### Assign back crop fraction and yield to grid 
```{r, eval=FALSE}
# first in data.tables
crop_ay <- lapply(c("gti", snms), function(j) {
  #j <- snms[3]
  DTya <- copy(md_dt)[, list(x, y, md, "f" = get(j))]
  DTya <- DTya[md_lcl[[j]][, list(md, fmd, mfrac, mzyld)]]
  DTya[, fa := f * mfrac]  # this is Monfreda et al equation (pg. 10)
  #print(DTya[which(round(DTya$fa, 4) > 1)[1:4], ])  # check line
  DTya[, mzya := mzyld]  # adjusted yield variable
  DTya[fa == 0, mzya := 0] # set to 0 in pixels having no crop
  DTya
})
sapply(crop_ay, function(x) x[, sum(mfrac), by = md][, V1])

# then back to rasters for aggregation
crop_ayr <- lapply(crop_ay, function(x) {
  dt_to_raster(x[, list(x, y, fa, mzya)], CRSobj = sa.shp@proj4string)
})

# redo NA mask because of some slighting shifts
namask <- calc(stack(lapply(crop_ayr, function(x) x$fa)), sumna)
namask[namask > 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(is.na(gti), sum); cellStats(namask2, sum)
area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
area_wgts <- lapply(area_wgts, function(x) x[[1]])  # unlist on inner loop

# Aggregate yields and crop area to coarser resolutions
carea <- lapply(1:5, function(x) crop_ayr[[x]]$fa)
names(carea) <- c("gti", snms)
cyld <- lapply(1:5, function(x) crop_ayr[[x]]$mzya)
names(cyld) <- c("gti", snms)

# check Monfreda's yield data to see how it deals with null areas
monf <- full_path("external/ext_data/monfreda_data/maize_HarvAreaYield_NetCDF",
                  "maize_AreaYieldProduction.nc")
monfmz <- brick(monf, level = 2)
plot(crop(monfmz, spTransform(sa.shp, monfmz@crs)), col = bpy.colors(20))
plot(crop(monfmz, spTransform(sa.shp, monfmz@crs)) > 0)  # almost all SA > 0

# aggregate
# first define a function that removes zero yield areas when calculating mean 
# yield--this will allow for aggregation of yields only in those areas having 
# maize, according to Monfreda et al methods
nzmean <- function(x, na.rm) {
  # x <- c(0, 1:5, NA, 10, NA)
  x <- x[(x > 0) & !is.na(x)]
  if(length(x) == 0) {
    m <- 0
  } else if(length(x) > 0) {
    m <- mean(x, na.rm = na.rm)
  }
  return(m)
}
# sumha <- function(x, na.rm) sum(x, na.rm)
carea_agg <- aggregate_rast_list(fact, carea) # crop area, mean aggregation
#carea_agg2 <- aggregate_rast_list(fact, carea, sum) # crop area, sum agg
cyld_agg <- aggregate_rast_list(fact, cyld, mean) # yld raster normal agg
cyld_agg2 <- aggregate_rast_list(fact, cyld, nzmean) # correct yld agg: drop 0s

#plot((carea_agg$f50$gti * tareas$f50) - (carea_agg2$f50$gti * 100))  # equiv
# ta <- rep(1, 50 * 50)
# a <- runif(50 * 50, 0, 1)
# (sum(a) * 100) / (sum(ta) * 100)
# mean(a) * (sum(ta) * 100)

# area of each pixel at each aggregation scale
agg_res <- sapply(carea_agg, function(x) res(x$gti)[1]^2 / 10000)  
tareas <- lapply(area_wgts, function(x) x[[1]] * 100)

# crop production estimates per dataset per aggregation scale
# with straight mean (incorrect) yield aggregation
cpagg1 <- lapply(1:length(cyld_agg), function(x) {
  p <- lapply(1:length(cyld_agg[[x]]), function(y) {
    tareas[[x]] * cyld_agg[[x]][[y]]
  })
  named_out(p, c("gti", snms))
})
names(cpagg1) <- names(agg_res)

# with correct (0 removed) yield aggregation
cpagg2 <- lapply(1:length(cyld_agg2), function(x) {
  p <- lapply(1:length(cyld_agg2[[x]]), function(y) {
    tareas[[x]] * cyld_agg2[[x]][[y]]
  })
  named_out(p, c("gti", snms))
})
names(cpagg2) <- names(agg_res)

# Type 2 aggregation: crop production differences when aggregated from 1 km
# calculate production at 1 km
cprod2 <- lapply(1:length(cyld), function(x) {
  cyld[[x]] * (carea[[x]] * 100)
})
names(cprod2) <- c("gti", snms)
# aggregate production estimates
cpagg3 <- aggregate_rast_list(fact, cprod2, "sum") # crop area rasters
sapply(cprod2, function(x) cellStats(x, sum))
sapply(carea, function(x) cellStats(x, sum))

# calculate differences between them
# pdiff <- function(x, y) (x - y) / x * 100
# Production differences
ilist <- list(names(cpagg1[[1]])[-1], names(cpagg1), "gti")
list1 <- lapply(cpagg1, function(x) x[1])
list2 <- lapply(cpagg1, function(x) x[-1])
p_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1a agg
list1 <- lapply(cpagg2, function(x) x[1])
list2 <- lapply(cpagg2, function(x) x[-1])
p_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1b agg
list1 <- lapply(cpagg3, function(x) x[1])
list2 <- lapply(cpagg3, function(x) x[-1])
p_diff3 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 2 agg

# yield differences
# type 1 (incorrect)
list1 <- lapply(cyld_agg, function(x) x[1])
list2 <- lapply(cyld_agg, function(x) x[-1])
y_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")
# type 2 (correct)
list1 <- lapply(cyld_agg2, function(x) x[1])
list2 <- lapply(cyld_agg2, function(x) x[-1])
y_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")
plot(y_diff2$sa30$f25$gti)

# disaggregate for plotting
lev <- names(y_diff2[[1]])
p_diff1d <- disaggregate_rast_list(snms, lev, p_diff1, namask)
p_diff2d <- disaggregate_rast_list(snms, lev, p_diff2, namask)
p_diff3d <- disaggregate_rast_list(snms, lev, p_diff3, namask)
y_diff1d <- disaggregate_rast_list(snms, lev, y_diff1, namask)
y_diff2d <- disaggregate_rast_list(snms, lev, y_diff2, namask)

# stats for plotting
dstats <- function(dlist) {
  statsd <- lapply(dlist, function(x) {
    sapply(x, function(y) {
      c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
    })
  })
}
stats_pd1 <- dstats(p_diff1d)
stats_pd2 <- dstats(p_diff2d)
stats_pd3 <- dstats(p_diff3d)
stats_yd1 <- dstats(y_diff1d)
stats_yd2 <- dstats(y_diff2d)

```
<a href="#top">Back to top</a>

### Plot difference maps
```{r, eval=FALSE}
dnms <- ls()[grep("[y_|p_]diff*.*d", ls())]
stnms <- ls()[grep("stats_", ls())]
disaggl <- lapply(dnms, function(x) get(x))
statsl <- lapply(stnms, function(x) get(x))
rng <- lapply(statsl, function(x) {
  ilims <- c(ceiling(min(sapply(x, function(y) y[3, ]))), 
             floor(max(sapply(x, function(y) y[21, ]))))
  olims <- range(sapply(x, function(y) range(y)))
  c(olims[1], ilims, olims[2])
})

fnms <- c(paste0("prod_bias_map_", 1:3, ".pdf"), 
          paste0("yld_bias_map_", 1:2, ".pdf"))  # output names
legtext <- c(rep("tons", 3), rep("tons/ha", 2))
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev <- names(p_diff1d[[1]])[-c(2:3)]
lev2 <- c("1 km", "25 km", "50 km", "100 km")

rnder <- function(x) {
  dig <- nchar(max(round(x)))
  rmat <- cbind(1:8, c(1, 5, 10, 20, 50, 100, 100, 1000))
  rnd <- rmat[dig, 2]
  round(c(floor(x[1] / rnd), ceiling(x[2] / rnd))) * rnd 
}
div_breaks <- function(olim, ilim, ibrk) {
  il <- rnder(ilim) 
  ol <- c(floor(olim[1]), ceiling(olim[2]))
  brks <- c(ol[1], seq(il[1], -ibrk, abs(floor(il[1]) - -ibrk) / 4), 
            seq(ibrk, il[2], abs(ceiling(il[2]) - ibrk) / 4), ol[2])
  brks
}

for(i in 1:length(statsl)) {
  # i <- 1
  lim <- rng[[i]]
  brks <- div_breaks(olim = lim[c(1, 4)], ilim = lim[2:3], 
                     ibrk = 0.1 * min(abs(lim[2:3])))
  if(i < 4) brks <- round(brks)
  if(i >= 4) brks <- round(brks, 1)
  cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)
  pdf(full_path(p_fig, fnms[i]), height = 6, width = 7)
  par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
  for(j in 1:length(snms)) {
    print(snms[j])
    for(k in 1:length(lev)) {
      rl <- disaggl[[i]]
      print(lev[k])
      plot(sa.shp, lty = 0)
      plot(rl[[snms[j]]][[lev[k]]], add = TRUE, col = cols, 
           breaks = brks, legend = FALSE)
      if(k == 1) mtext(mcap[j], side = 2, line = 1, cex = cx)
      if(j == 1) mtext(lev2[k], side = 3, line = 0, cex = cx)
    }
  }
  flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
              legend.vals = brks, longdims = c(0.2, 0.8), 
              shortdims = c(0.1, 0.01), 
              colvec = cols, #(length(brks) - 1), 
              srt = c(270, 0), horiz = TRUE, textside = "bottom", 
              legend.pos = c(4, 5), leg.adj = list(c(0.1, 0.3), c(0, -0.5)), 
              cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
  dev.off()
}

```
<a href="#top">Back to top</a>

From the above it is clear that the nature in which aggregation of yields are done has an impact on both yield and production estimates. There are two types of yield aggregation that can be done:

  1. __Type 1__: or straight averaging, which lowers yields by averaging in 0 yield areas.

  2. __Type 2__: or zero-removed averaging, where only areas having fraction of that crop contribute to the aggregated average. This is the correct way to do it.

There is all two ways of aggregating production estimates:

  1.  __Type 1__: Aggregating yield and crop area separately, then multiplying.

  2.  __Type 2__: Calculating production at the base resolution, then aggregating by sum.

Why would someone do a yield aggregation, when you have the gridded estimates?  To calculate values and compare them to a coarser resolution product, for instance, or to get a country-level average.

Causes for bias concern to examine:

  + Bias in yield gap estimates, and how much closing that could affect contribution of closing gaps to boosting production.

  + Country-level statistics might cancel out bias, but if bias of one type of sign is spatially correlated with the driver of values in yield gaps, then that would bias larger-scale estimates of gap closure potential.

      + ___Note: investigate this by look at correlation between rainfall and yield bias patterns.___

### Bias statistics

Following methods developed in [carbon-bias](carbon-bias.html)
```{r, eval=FALSE}
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)

# create mask for non-cropland areas, unioning GTI and LC areas with maize
# estimates -- probably not needed if doing production estimates.  
lev <- names(y_diff1[[1]])

# first check aggregation to make yield aggregation variants have common area,
# so that just one lc_union is needed
for(i in lev) {
  for(j in snms) {
    print(cellStats(cyld_agg[[i]][[j]] > 0 & cyld_agg2[[i]][[j]] == 0, sum))
    #print(cellStats(carea_agg[[i]][[j]] > 0 & cyld_agg[[i]][[j]] == 0, sum))
  }
}  # all zeros

lc_union <- lapply(lev, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(carea_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(carea_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lc_union) <- lev
# plot(lc_union$f50$modmu)
# plot(gti_agg$f50$gti)

# selection variables
nms1 <- c(colnames(ctabo), "wgt")
nms2 <- colnames(ctabo)
bvals <- "mu"

```

```{r, eval=FALSE, echo=FALSE}
# looking into pycnophylactic interpolation of correction factors
# install.packages("pycno")
require(pycno)

# pycno package example
nc_sids <- readOGR(system.file("shapes/sids.shp", package="maptools")[1], 
                   layer = "sids")
nc_sids@proj4string <- CRS("+proj=longlat +ellps=clrk66")
births74 <- pycno(nc_sids, nc_sids$BIR74, 0.05, converge = 1)
plot(raster(as(births74, "SpatialPixelsDataFrame")))
plot(nc_sids,add=TRUE)

# on out data
provr <- rasterize(prov, y = gti_agg$f10$gti, field = "id_1")
prov2 <- rasterToPolygons(provr, dissolve = TRUE)
colnames(prov2@data) <- "provid"
prov2@data <- cbind(prov2@data, do.call(cbind, pcf))

snms <- colnames(prov2@data)[2:5]
cfrs <- lapply(snms, function(x) {
  #x <- snms[1]
  cfr <- pycno(prov2, prov2@data[, x], 10000, converge = 1)
  cfr_spp <- as(cfr, "SpatialPixelsDataFrame")
  cfrr <- raster(cfr_spp)
  cfrrs <- resample(x = cfrr, provr, method = "ngb")
})
names(cfrs) <- snms
plot(cfrs$sa30)

reclmat <- cbind(1:9, pcf$sa30)  # reclass matrix
cfr <- reclassify(provr, reclmat)  # assign cf values to rasterized
tst <- (lc_agg$f10$sa30 * cfr)

# this interpolation does not make sense in terms of leading to a correction 
# that ends up with cropland fractions matching the provincial estimates of 
# cropland area, at least not given the equations provided by Ramankutty et al
# (2008). The provincial level correction factor should not be disaggregated so # that the summed correction factors add back up to the original correction
# factor.

cft <- 18 / sum(rep(2, 10))
sum(rep(2, 10) * cft)

```

### 2015-07-15 R/as.data.table.raster.R

@@ -1,0 +1,63 @@
+#' Transform raster to data.table
+#' 
+#' @param x  Raster* object
+#' @param row.names	`NULL` or character vector giving row names for data frame. 
+#' Missing values are not allowed
+#' @param optional	logical. If `TRUE`, setting row names and converting column 
+#' names (to syntactic names: see make.names) is optional
+#' @param xy logical. If `TRUE`, also return the spatial coordinates
+#' @param centroids	logical. If TRUE return the centroids instead of all spatial 
+#' coordinates (only relevant if xy=TRUE)
+#' @param sepNA	logical. If TRUE the parts of the spatial objects are separated 
+#' by lines that are NA (only if xy=TRUE and, for polygons, if centroids=FALSE
+#' @param ...	 Additional arguments (none) passed to `raster::as.data.frame`
+#' 
+#' @return returns a data.table object
+#' @author etiennebr, from https://gist.github.com/etiennebr/9515738/download#
+#' @note This has been amended to fix the blocksize parameter, which was set to
+#' 2. 
+#' @examples
+#' logo <- brick(system.file("external/rlogo.grd", package="raster"))
+#' v <- as.data.table(logo)
+#' @export
+
+as.data.table.raster <- function(x, row.names = NULL, optional = FALSE, 
+                                 xy = FALSE, inmem = canProcessInMemory(x, 2), 
+                                 ...) {
+  stopifnot(require("data.table"))
+  if(inmem) {
+    v <- as.data.table(raster::as.data.frame(x, row.names=row.names, 
+                                             optional=optional, xy=xy, ...))
+    coln <- names(x)
+    if(xy) coln <- c("x", "y", coln)
+    setnames(v, coln)
+  } else {
+    tr <- blockSize(x)
+    l <- lapply(1:tr$n, function(i) {
+      DT <- as.data.table(as.data.frame(getValues(x, row = tr$row[i], 
+                                                  nrows = tr$nrows[i]), ...))  
+      if(xy == TRUE) {
+        cells <- cellFromRowCol(x, c(tr$row[i], tr$row[i] + tr$nrows[i] - 1),
+                                c(1, ncol(br)))
+        coords <- xyFromCell(x, cell = cells[1]:cells[2])
+        DT[, c("x", "y") := data.frame(xyFromCell(x, cell = cells[1]:cells[2]))]
+     } 
+     DT
+    })
[…]

### 2015-07-15 R/dt_to_raster.R

@@ -1,0 +1,23 @@
+#' Transform data.table to raster*
+#' @param dt data.table with x, y coordinates
+#' @param CRSobj crs object
+#' @param filename NULL, else raster* will be written to specified file
+#' @return returns a data.table object
+#' @export
+dt_to_raster <- function(dt, CRSobj, filename = NULL) {
+  if(any(!c("x", "y") %in% colnames(dt))) {
+    stop("dt needs planar coordinates in x, y format")
+  }
+  coordinates(dt) = ~x+y
+  proj4string(dt) <- CRSobj
+  gridded(dt) = TRUE
+  if(ncol(dt) > 1) {
+    dtr <- brick(dt)
+  } else {
+    dtr <- brick(dt)
+  }
+  if(!is.null(filename)) dtr <- writeRaster(dtr, filename)
+  return(dtr)
+}
+
+

### 2015-07-15 R/landcover-functions.R

@@ -42,6 +42,36 @@ aggregate_rast_list <- function(fact, rlist, fun = mean) {
   return(aggs2)
 }
 
+#' Disaggregates a raster or list of rasters
+#' 
+#' @param names1 names or indices for outer loop of 2 level raster list
+#' @param names2 inner list names denoting aggregation level (see details)
+#' @param rlist List of rasters to be aggregated
+#' @param mask Masking raster to be applied
+#' @return List of rasters disaggregated back to original resolution
+#' @details This is essentially the reverse of aggregate_rast_list, thus has a
+#' structure that is highly specific to this project. The names 2 vector should
+#' be of the form c("f1", "f5", "f10"), etc, where the f stands for factor and 
+#' the number gives the factor by which the raster being disaggregated was 
+#' aggregated from its original resolution.  
+#' @export
+disaggregate_rast_list <- function(names1, names2, rlist, rmask) {
+  disagg <- lapply(names1, function(x) {
+    l1 <- lapply(names2, function(y) {
+      fact <- as.numeric(gsub("f", "", y))
+      if(fact == 1) {
+        r <- rlist[[x]][[y]][[1]]
+      } else {
+        r <- disaggregate(rlist[[x]][[y]][[1]], fact = fact)
+        r <- mask(crop(r, rmask), rmask)
+      }
+    })
+    named_out(l1, names2)
+  })
+  named_out(disagg, names1)
+}
+
+
 #' @title Apply math on two lists
 #' @description Apply arbitrary math to two objects within two lists 
 #' @param ilist A 3 element list of list indices (names or index numbers). See details. 

### 2015-08-17 Rmd/yield-bias.Rmd
Major update, mostly based on undertaking and completing full version of yield-bias.Rmd, but also some changes to library functions, including new stats estimating functions and removal of data.table/raster functions to to new package, which is used here.

@@
 
-1.
 Use SA provinces with GTI to create the amount of agricultural land estimates per province, akin to provincial statistics used by Ramankutty et al (2008).
-Note: they mentioned SA has 11 administrative units they used, but the publications they cited ([here](http://www.statssa.gov.za/publications/DiscussLandAccounts/DiscussLandAccounts.pdf)) has just 9 provinces, so maybe they mistakenly counted Lesotho and Swaziland?
+Note: they mentioned SA has 11 administrative units they used, but the publications they cited [here](http://www.statssa.gov.za/publications/DiscussLandAccounts/DiscussLandAccounts.pdf) has just 9 provinces, so maybe they mistakenly counted Lesotho and Swaziland?
 
@@
 
-# Create namask and area_wgts, removing all NA areas across datasets
+# Create namask to remove all NA areas across datasets
 sumna <- function(x) sum(x, na.rm = FALSE)
@@
 fact <- c(5, 10, 25, 50, 100)
-# area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
-# area_wgts <- lapply(area_wgts, function(x) x[[1]])  # unlist on inner loop
-# plot(area_wgts$f25)
-
 ```
@@
 ## Cropland fractions
-
 ### Get provincial $cf$ factors from 1 km GTI dataset
@@
 # rasterize provinces and stack with gti raster, convert to data.table
-provr <- rasterize(prov, y = gti, field = "id_1", 
-                   filename = "external/ext_data/provinces.tif")
-# provr <- raster("external/ext_data/provinces.tif")
+# provr <- rasterize(prov, y = gti, field = "id_1", 
+#                    filename = "external/ext_data/provinces.tif")
+provr <- raster("external/ext_data/provinces.tif")
 provr <- mask(provr, namask)
@@
 # calculate fractions using raster::extract with provinces, to compare speed
-prov_est <- extract(gti, prov, progress = "text")
-prov_estNA <- lapply(prov_est, function(x) x[!is.na(x)])
-cftest <- cbind("carea" = sapply(prov_estNA, sum), 
-                "parea" = sapply(prov_estNA, length), 
-                "cf" = sapply(prov_estNA, sum) / sapply(prov_estNA, length))
+# prov_est <- extract(gti, prov, progress = "text")
+# prov_estNA <- lapply(prov_est, function(x) x[!is.na(x)])
+# cftest <- cbind("carea" = sapply(prov_estNA, sum), 
+#                 "parea" = sapply(prov_estNA, length), 
+#                 "cf" = sapply(prov_estNA, sum) / sapply(prov_estNA, length))
 
@@
 #pct_diff <- function(x, y) (x - y) / x * 100
-cf_pct_diff <- lapply(1:length(gti_agg), function(x) {
-  dif <- lapply(1:length(lc_agg[[x]]), function(y) {
-    d <- gti_agg[[x]][[1]] - lc_agg[[x]][[y]]
-  })
-  named_out(dif, snms)
-})
-names(cf_pct_diff) <- names(lc_agg)
 
@@
 lev2 <- c("1 km", "25 km", "50 km", "100 km")
-pdf(full_path(p_fig, "cropland_adj_bias_map.pdf"), height = 6, width = 7)
+pdf(full_path(p_fig, "cropland_adj_bias_map2.pdf"), height = 6, width = 7)
 par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
@@
 ```{r, eval = FALSE}
-mdr <- rasterize(md, provr, field = "id",  
-                 filename = "external/ext_data/mag_dist.tif", overwrite = TRUE)
-# mdr <- raster("external/ext_data/mag_dist.tif")
+# mdr <- rasterize(md, provr, field = "id",  
+#                  filename = "external/ext_data/mag_dist.tif", overwrite = TRUE)
+mdr <- raster("external/ext_data/mag_dist.tif")
 mds <- stack(list("md" = crop(mdr, gti_agg$f1$gti), "gti" = gti_agg$f1$gti,
@@
     # print(length(which(DTy$mfrac > 1)))  # 5, 4, 15, 9, 4
-    DTy[mfrac > 1, mfrac := 1]  # set to 1 - mostly missing data causing > 1
+    #DTy[mfrac > 1, mfrac := 1]  # set to 1 - mostly missing data causing > 1
     DTy
@@
 plot(md_lcl[[1]]$carea, md_lcl[[3]]$carea)
+for(i in 1:5) print(length(which(md_lcl[[i]]$mfrac > 1)))
 ```
 <a href="#top">Back to top</a>
+
+Note: when assigning crop fractions, previously we set all fractions > 1 to 1, but now turned this off because Monfreda et al (2010) allow for double-cropping.
+This is likely not correct, but we are doing it here to be consistent.
 
@@
 # first in data.tables
+asnms <- c("gti", snms)
 crop_ay <- lapply(c("gti", snms), function(j) {
@@
 })
-sapply(crop_ay, function(x) x[, sum(mfrac), by = md][, V1])
+names(crop_ay) <- asnms
+sapply(crop_ay, function(x) x[, round(sum(fa), 4), by = md][, V1])
 
@@
 # cellStats(is.na(gti), sum); cellStats(namask2, sum)
-area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
-area_wgts <- lapply(area_wgts, function(x) x[[1]])  # unlist on inner loop
+awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
+awgts <- lapply(awgts, function(x) x[[1]])  # unlist on inner loop
+# sumna <- function(x, na.rm = na.rm) sum(x, na.rm = na.rm)
+# amsk <- aggregate_rast_list(fact, list(namask2), fun = sumna)
+# amsk <- lapply(amsk, function(x) x[[1]])  # unlist on inner loop
 
-# Aggregate yields and crop area to coarser resolutions
-carea <- lapply(1:5, function(x) crop_ayr[[x]]$fa)
-names(carea) <- c("gti", snms)
-cyld <- lapply(1:5, function(x) crop_ayr[[x]]$mzya)
-names(cyld) <- c("gti", snms)
+```
+<a href="#top">Back to top</a>
+
+### Aggregate yields and crop area to coarser resolutions
+```{r, eval=FALSE}
+# Yield aggregation - aggregating with weighting by crop fraction
+# using a custom data.table function to allow this 
+lev <- names(disagg[[1]])
+wmfun <- parse(text = "sum((mzya * fa) / sum(fa, na.rm = TRUE), na.rm = TRUE)")
+sr <- sa.shp@proj4string
+d <- c(dim(gti_agg[[1]]$gti)[1:2], xres(gti_agg[[1]]$gti))  # dimensions
+cyld <- lapply(lev[-1], function(i) {
+  print(i)
+  f <- as.integer(gsub("f", "", i))
+  il <- lapply(c("gti", snms), function(j) {
+    print(paste("...", j))
+    rdt <- as.data.table(crop_ayr[[j]], xy = TRUE)
+    o <- dt_aggregate(rdt, d[1], d[2], f, wmfun, d[3], "yw")
+  })
+  named_out(il, asnms)
+})
+names(cyld) <- lev[-1]
+
+# convert back to rasters
+cyldr <- lapply(cyld, function(i) lapply(i, function(j) dt_to_raster(j, sr)))
+cyldf1 <- lapply("f1", function(i) {
+  f1yld <- lapply(asnms, function(j) crop_ayr[[j]]$mzya)
+  named_out(f1yld, asnms)
+})
+names(cyldf1) <- lev[1]
+cyld_agg <- c(cyldf1, cyldr)
+
+plot(cyld_agg$f10$gti)
 
@@
 
-# aggregate
+# not implausible that someone would do a direct aggregation of these yield 
+# values, without weighting by crop distribution, so this incorrect variant 
+# would be: 
+cyld_agg2 <- aggregate_rast_list(fact, cyld_agg$f1, mean) # incorrect yld agg
+
+# third variant - aggregate and remove zeros
 # first define a function that removes zero yield areas when calculating mean 
@@
 nzmean <- function(x, na.rm) {
-  # x <- c(0, 1:5, NA, 10, NA)
   x <- x[(x > 0) & !is.na(x)]
@@
 }
+cyld_agg3 <- aggregate_rast_list(fact, cyld_agg$f1, nzmean) # no zero agg
+
+# plot(cyld_agg$f25$modmu - cyld_agg2$f25$modmu)
+# plot(cyld_agg$f25$modmu - cyld_agg3$f25$modmu)
+wm <- function(x, w) stats::weighted.mean(x, w)
+
+## crop areas
+carea <- lapply(1:5, function(x) crop_ayr[[x]]$fa)
+names(carea) <- c("gti", snms)
+
 # sumha <- function(x, na.rm) sum(x, na.rm)
@@
 #carea_agg2 <- aggregate_rast_list(fact, carea, sum) # crop area, sum agg
-cyld_agg <- aggregate_rast_list(fact, cyld, mean) # yld raster normal agg
-cyld_agg2 <- aggregate_rast_list(fact, cyld, nzmean) # correct yld agg: drop 0s
+# plot(carea_agg$f25$globmu)
+# plot(carea_agg2$f25$globmu)
 
+# area of each pixel at each aggregation scale
+agg_res <- sapply(carea_agg, function(x) res(x$gti)[1]^2 / 10000)  
+tareas <- lapply(awgts, function(x) x[[1]] * 100)
+
+# plot(tareas$f50)
 #plot((carea_agg$f50$gti * tareas$f50) - (carea_agg2$f50$gti * 100))  # equiv
@@
 
-# area of each pixel at each aggregation scale
-agg_res <- sapply(carea_agg, function(x) res(x$gti)[1]^2 / 10000)  
-tareas <- lapply(area_wgts, function(x) x[[1]] * 100)
+yat <- list(cyld_agg, cyld_agg2, cyld_agg3)
+## check values
+# check yields
+for(i in lev[-1]) {
+  print(paste("factor", i))
+  for(j in asnms) {
+    print(paste("...", j))
+    for(k in 1:3) {
+      #ck <- cellStats(cyld_agg$f1$gti, nzmean)
+      print(paste("...... type", k, "agg"))
+      ck <- crop_ay[[j]][, wm(mzyld, fa)]  # weighted by fraction
+      ck2 <- crop_ay$gti[, wm(mzyld, fa)]  # weighted by fraction
+      w <- values(carea_agg[[i]][[j]])  # aggregated crop fraction
+      x <- values(yat[[k]][[i]][[j]])  # yield values
+      ind <- which(x > 0)  # which are non-zero
+      x <- x[ind]  # select non zeros
+      w <- w[ind]  # select non-zeros
+      #w <- w / max(w)
+      v <- weighted.mean(x, w)
+      d <- round((ck - v) / ck, 2)
+      d2 <- round((ck2 - v) / ck2, 2)
+      print(c(d, d2))
+    }
+  }  # type 1 and type 3 aggregations consistent between datasets, across scales
+}
+rm(yat)
 
-# crop production estimates per dataset per aggregation scale
-# with straight mean (incorrect) yield aggregation
+# check aggregated maize fractions
+for(i in lev) {
+  for(j in snms) {
+    ck <- crop_ay[[j]][, sum(fa)] * 100
+    d <- round((ck - cellStats(tareas[[i]] * carea_agg[[i]][[j]], sum)) / ck, 2)
+    print(d)
+  }
+}  # all equal
+```
+<a href="#top">Back to top</a>
+
+### Crop production estimates at different aggregation scales
+```{r, eval=FALSE}
+# Type A aggregation: multiplying aggregated yields by aggregated fractions
+# with crop-fraction weighted aggregate yields
 cpagg1 <- lapply(1:length(cyld_agg), function(x) {
   p <- lapply(1:length(cyld_agg[[x]]), function(y) {
-    tareas[[x]] * cyld_agg[[x]][[y]]
+    (tareas[[x]] * carea_agg[[x]][[y]]) * cyld_agg[[x]][[y]]
   })
-  named_out(p, c("gti", snms))
+  named_out(p, asnms)
 })
@@
 
-# with correct (0 removed) yield aggregation
-cpagg2 <- lapply(1:length(cyld_agg2), function(x) {
-  p <- lapply(1:length(cyld_agg2[[x]]), function(y) {
-    tareas[[x]] * cyld_agg2[[x]][[y]]
+# with 0-removed yield aggregation (type 3 yield aggregation)
+cpagg3 <- lapply(1:length(cyld_agg3), function(x) {
+  p <- lapply(1:length(cyld_agg3[[x]]), function(y) {
+    (tareas[[x]] * carea_agg[[x]][[y]]) * cyld_agg3[[x]][[y]]
   })
-  named_out(p, c("gti", snms))
+  named_out(p, asnms)
 })
-names(cpagg2) <- names(agg_res)
+names(cpagg3) <- names(agg_res)
 
-# Type 2 aggregation: crop production differences when aggregated from 1 km
-# calculate production at 1 km
-cprod2 <- lapply(1:length(cyld), function(x) {
-  cyld[[x]] * (carea[[x]] * 100)
+# Type B (formerly II) aggregation: crop production differences when aggregated
+# from 1 km calculate production at 1 km
+cprod <- lapply(1:length(cyldf1$f1), function(x) {
+  cyldf1$f1[[x]] * (carea[[x]] * 100)
 })
-names(cprod2) <- c("gti", snms)
-# aggregate production estimates
-cpagg3 <- aggregate_rast_list(fact, cprod2, "sum") # crop area rasters
-sapply(cprod2, function(x) cellStats(x, sum))
-sapply(carea, function(x) cellStats(x, sum))
+names(cprod) <- asnms
+cpaggII <- aggregate_rast_list(fact, cprod, "sum") # aggregate production
+
+# check production estimates for consistency
+pat <- list(cpagg1, cpagg3, cpaggII)
+ptype <- c("1", "3", "II")
+# check production estimates
+for(i in lev[-1]) {
+  print(paste("factor", i))
+  for(j in asnms) {
+    print(paste("...", j))
+    for(k in 1:3) {
+      print(paste("...... type", ptype[k], "agg"))
+      ck <- cellStats(cprod[[j]], sum)  # weighted by fraction
+      ck2 <- cellStats(cprod$gti, sum) # weighted by fraction
+      v <- cellStats(pat[[k]][[i]][[j]], sum)
+      d <- round((ck - v) / ck, 2)
+      d2 <- round((ck2 - v) / ck2, 2)
+      print(c(d, d2))
+    }
+  }  # all types consistent across scales
+}
+
+# sapply(cprod, function(x) cellStats(x, sum))
+# sapply(carea, function(x) cellStats(x, sum))
+
+# Mean production in producing cells at different levels of aggregation
+mup <- sapply(lev, function(i) {
+  msk <- cpaggII[[i]]$gti > 0
+  msk[msk == 0] <- NA
+  cpmsk <- msk * cpaggII[[i]]$gti
+  cellStats(cpmsk, mean)
+})
+# sapply(lev, function(i) {
+#   msk <- cpagg1[[i]]$gti > 0
+#   msk[msk == 0] <- NA
+#   cpmsk <- msk * cpagg1[[i]]$gti
+#   cellStats(cpmsk, mean)
+# })
+
+# mean yield in producing cells at different levels of aggregation
+muy <- sapply(lev, function(i) {
+  w <- values(carea_agg[[i]]$gti)  # aggregated crop fraction
+  x <- values(cyld_agg[[i]]$gti)  # yield values
+  ind <- which(x > 0)  # which are non-zero
+  x <- x[ind]  # select non zeros
+  w <- w[ind]  # select non-zeros
+  v <- weighted.mean(x, w, na.rm = TRUE, )
+  v
+})
 
 # calculate differences between them
-# pdiff <- function(x, y) (x - y) / x * 100
+# check first that cpagg1 and II are equivalent, make cpaggII master
+for(i in lev) {
+  for(j in snms) {
+    a <- (cpagg1[[i]]$gti - cpagg1[[i]][[j]]) - 
+      (cpaggII[[i]]$gti - cpagg1[[i]][[j]])
+    print(round(cellStats(a, sum), 5))
+  }
+}
 # Production differences
 ilist <- list(names(cpagg1[[1]])[-1], names(cpagg1), "gti")
-list1 <- lapply(cpagg1, function(x) x[1])
+list1 <- lapply(cpaggII, function(x) x[[1]])
 list2 <- lapply(cpagg1, function(x) x[-1])
-p_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1a agg
-list1 <- lapply(cpagg2, function(x) x[1])
-list2 <- lapply(cpagg2, function(x) x[-1])
-p_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1b agg
-list1 <- lapply(cpagg3, function(x) x[1])
+p_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1 agg
 list2 <- lapply(cpagg3, function(x) x[-1])
-p_diff3 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 2 agg
+p_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 3 agg
+list2 <- lapply(cpaggII, function(x) x[-1])
+p_diff3 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type II agg
+# pmu <- as.data.table.raster(cpagg3$f1$globmu)[layer > 0, mean(layer)]
 
 # yield differences
-# type 1 (incorrect)
 list1 <- lapply(cyld_agg, function(x) x[1])
 list2 <- lapply(cyld_agg, function(x) x[-1])
-y_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")
-# type 2 (correct)
-list1 <- lapply(cyld_agg2, function(x) x[1])
-list2 <- lapply(cyld_agg2, function(x) x[-1])
-y_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")
-plot(y_diff2$sa30$f25$gti)
+y_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1 
+# type 2 (non-zero)
+list2 <- lapply(cyld_agg3, function(x) x[-1]) # reuse cyld_agg for a, b/c correct
+y_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 2
+
+# convert them to percent differences relative to the mean pixel-wise production
+# plot(y_diff2$sa30$f25$gti)
+pdiffs <- lapply(list(p_diff1, p_diff2, p_diff3), function(x) {
+  l1 <- lapply(snms, function(y) {
+    l2 <- lapply(lev, function(z) {
+      x[[y]][[z]]$gti / mup[z] * 100
+    })
+    named_out(l2, lev)
+  })  
+  named_out(l1, snms)
+})
+names(pdiffs) <- paste0("pd", 1:3)
+
+ydiffs <- lapply(list(y_diff1, y_diff2), function(x) {
+  l1 <- lapply(snms, function(y) {
+    l2 <- lapply(lev, function(z) {
+      x[[y]][[z]]$gti / muy[z] * 100
+    })
+    named_out(l2, lev)
+  })  
+  named_out(l1, snms)
+})
+names(ydiffs) <- paste0("yd", 1:2)
 
@@
 lev <- names(y_diff2[[1]])
-p_diff1d <- disaggregate_rast_list(snms, lev, p_diff1, namask)
-p_diff2d <- disaggregate_rast_list(snms, lev, p_diff2, namask)
-p_diff3d <- disaggregate_rast_list(snms, lev, p_diff3, namask)
-y_diff1d <- disaggregate_rast_list(snms, lev, y_diff1, namask)
-y_diff2d <- disaggregate_rast_list(snms, lev, y_diff2, namask)
+p_diff1d <- disaggregate_rast_list(snms, lev, pdiffs$pd1, namask)
+p_diff2d <- disaggregate_rast_list(snms, lev, pdiffs$pd2, namask)
+p_diff3d <- disaggregate_rast_list(snms, lev, pdiffs$pd3, namask)
+y_diff1d <- disaggregate_rast_list(snms, lev, ydiffs$yd1, namask)
+y_diff2d <- disaggregate_rast_list(snms, lev, ydiffs$yd2, namask)
 
@@
 stats_yd2 <- dstats(y_diff2d)
-
+# cbind(stats_yd1$geow[, 4], stats_yd2$geow[, 4])
 ```
@@
           paste0("yld_bias_map_", 1:2, ".pdf"))  # output names
-legtext <- c(rep("tons", 3), rep("tons/ha", 2))
+# legtext <- c(rep("tons", 3), rep("tons/ha", 2))
+legtext <- rep("%", 5)
 cx <- 1.4
@@
 
+rng2 <- lapply(rng, function(x) {
+  x[2:3] <- c(-100, 100)
+  x
+})
+brksl <- lapply(rng2, function(x) {
+  c(floor(x[1:2]), c(-50, -25, -10, -1, 1, 10, 25, 50), ceiling(x[3:4]))
+})
+
 for(i in 1:length(statsl)) {
   # i <- 1
-  lim <- rng[[i]]
-  brks <- div_breaks(olim = lim[c(1, 4)], ilim = lim[2:3], 
-                     ibrk = 0.1 * min(abs(lim[2:3])))
-  if(i < 4) brks <- round(brks)
-  if(i >= 4) brks <- round(brks, 1)
+  #lim <- rng[[i]]
+  # lim <- rng2[[i]]
+  # brks <- div_breaks(olim = lim[c(1, 4)], ilim = lim[2:3], 
+  #                   ibrk = 0.01 * min(abs(lim[2:3])))
+  # if(i < 4) brks <- round(brks)
+  # if(i >= 4) brks <- round(brks, 1)
+  print(paste("figure", i, "of", length(statsl)))
+  brks <- brksl[[i]]
   cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)
+  cols[c(1, length(cols))] <- c("darkred", "purple3")
   pdf(full_path(p_fig, fnms[i]), height = 6, width = 7)
@@
 
+lev <- names(p_diff1d[[1]])  # reset levs
 ```
@@
 
-From the above it is clear that the nature in which aggregation of yields are done has an impact on both yield and production estimates. There are two types of yield aggregation that can be done:
+From the above it is clear that the nature in which aggregation of yields are done has some impact on both yield and production estimates. There are three types of yield aggregation that can be done:
 
-  1. __Type 1__: or straight averaging, which lowers yields by averaging in 0 yield areas.
+  1.
+__Type 1__: weighted mean averaging, with the weights provided by crop fractions.
+This produces correct country-level average yields (<1-2% different) across scales and between datasets, provided that the averaging is weighted by the cropland fractions that were aggregated to the level from which the country-level yield estimate is being calculated
 
-  2. __Type 2__: or zero-removed averaging, where only areas having fraction of that crop contribute to the aggregated average. This is the correct way to do it.
+  2. __Type 2__: Straight averaging, which results in diluted country-level yield estimates. Incorrect, and should be obvious that this wouldn't be done, so we haven't pursued it here.
 
-There is all two ways of aggregating production estimates:
+  3.
+__Type 3__: or zero-removed averaging, where only areas having fraction of that crop contribute to the aggregated average.
+As with __Type 1__, this produces correct country level estimates when weighting by cropland fraction, but it does not produce correct production estimates, probably because it inflates the value of high-yielding irrigated areas out towards the dry west, where the crop fractions are small.
+(see below)
 
-  1.  __Type 1__: Aggregating yield and crop area separately, then multiplying.
+There are also two ways of aggregating production estimates:
 
-  2.  __Type 2__: Calculating production at the base resolution, then aggregating by sum.
+  1.
+__Type A__: Aggregating yield and crop area separately, then multiplying.
+This is correct with __Type 1__ yield aggregation, and producing identical production estimates to __Type II__, but not with the other two types of yield aggregation.
+
+  2.  __Type B__: Calculating production at the base resolution, then aggregating by sum. This should be the correct way of doing it.
 
@@
 
+  + These concerns apply to all resolutions.
+
       + ___Note: investigate this by look at correlation between rainfall and yield bias patterns.___
 
-### Bias statistics
+## Bias statistics
 
@@
   for(j in snms) {
-    print(cellStats(cyld_agg[[i]][[j]] > 0 & cyld_agg2[[i]][[j]] == 0, sum))
+    print(cellStats(cyld_agg[[i]][[j]] > 0 & cyld_agg3[[i]][[j]] == 0, sum))
     #print(cellStats(carea_agg[[i]][[j]] > 0 & cyld_agg[[i]][[j]] == 0, sum))
@@
 
-lc_union <- lapply(lev, function(x) {
+snms <- names(p_diff1)
+lcu <- lapply(lev, function(x) {
   lcb <- lapply(snms, function(y) {
@@
 })
-names(lc_union) <- lev
+names(lcu) <- lev
 # plot(lc_union$f50$modmu)
@@
 
+# cropland cover bins, for looking at error as a function of cover
+binv <- seq(0, 1, 0.05)
+bins <- lapply(gti_agg, function(x) {
+  cut(x$gti, breaks = binv, include.lowest = TRUE)
+})
+# save(bins, awgts, lcu, p_diff1, wm, wma, p_diff1, cyld_agg, cyld_agg3, 
+#      carea_agg, cpagg1, cpagg3, cpaggII, p_diff2, p_diff3, y_diff1, 
+#      y_diff2, snms, lev, file = "external/ext_data/yield-bias.rda")
+# load("external/ext_data/yield-bias.rda")
+# reval <- p_diff1[[2]][[1]][[1]]  # hack in case there is a 3rd level
+# revall <- p_diff1
+# rmask <- lcu[[2]][[1]]
+# rmaskl <- lcu
+# rwgts <- awgts[[2]]
+# rwgtsl <- awgts
+# rbins <- bins[[2]]
+# rbinsl <- bins
+# fun <- sum
+# stnm <- "sum"
+# evalnm <- "bias"
+# rnd = 2; trim_wgt = TRUE; silent = FALSE; weighted = TRUE
+# wm <- function(x, w) stats::weighted.mean(x, w) 
+# tst <- data.table("a" = runif(20, 0, 1), 
+#                   "w" = sample(20:100, 20, replace = TRUE))
+# tst$w <- 0
+# tst[, lapply(.SD, sum, w), .SDcols = 1:2]
+# weighted.mean(tst$a, tst$w)
+# mean(tst$a)
+# sum(tst$a)
+
 # selection variables
-nms1 <- c(colnames(ctabo), "wgt")
-nms2 <- colnames(ctabo)
-bvals <- "mu"
+a <- bias_stats_list(bins, awgts, lcu, p_diff1, wm, "mu", "bias", TRUE)
+b <- bias_stats_list(bins, awgts, lcu, p_diff1, wma, "mua", "bias", TRUE)
+d <- bias_stats_list(bins, awgts, lcu, p_diff1, sum, "sum", "bias", FALSE)
+pd1_st <- rbind(a, b, d)
+a <- bias_stats_list(bins, awgts, lcu, p_diff2, wm, "mu", "bias", TRUE)
+b <- bias_stats_list(bins, awgts, lcu, p_diff2, wma, "mua", "bias", TRUE)
+d <- bias_stats_list(bins, awgts, lcu, p_diff2, sum, "sum", "bias", FALSE)
+pd2_st <- rbind(a, b, d)
+a <- bias_stats_list(bins, awgts, lcu, p_diff3, wm, "mu", "bias", TRUE)
+b <- bias_stats_list(bins, awgts, lcu, p_diff3, wma, "mua", "bias", TRUE)
+d <- bias_stats_list(bins, awgts, lcu, p_diff3, sum, "sum", "bias", FALSE)
+pd3_st <- rbind(a, b, d)
+a <- bias_stats_list(bins, awgts, lcu, y_diff1, wm, "mu", "bias", TRUE)
+b <- bias_stats_list(bins, awgts, lcu, y_diff1, wma, "mua", "bias", TRUE)
+yd1_st <- rbind(a, b)
+a <- bias_stats_list(bins, awgts, lcu, y_diff2, wm, "mu", "bias", TRUE)
+b <- bias_stats_list(bins, awgts, lcu, y_diff2, wma, "mua", "bias", TRUE)      
+yd2_st <- rbind(a, b)
+
+# Output statistics
+p1mu <- extract_stat(lev, snms, "all", "mu", "bias", pd1_st)  # identical to p3
+p1ma <- extract_stat(lev, snms, "all", "mua", "bias", pd1_st)  # identical to p3
+p1sm <- extract_stat(lev, snms, "all", "sum", "bias", pd1_st)  # identical to p3
+p2mu <- extract_stat(lev, snms, "all", "mu", "bias", pd2_st)
+p2ma <- extract_stat(lev, snms, "all", "mua", "bias", pd2_st)
+p2sm <- extract_stat(lev, snms, "all", "sum", "bias", pd2_st)
+p3mu <- extract_stat(lev, snms, "all", "mu", "bias", pd3_st)
+p3ma <- extract_stat(lev, snms, "all", "mua", "bias", pd3_st)
+p3sm <- extract_stat(lev, snms, "all", "sum", "bias", pd3_st)
+y1mu <- extract_stat(lev, snms, "all", "mu", "bias", yd1_st)
+y1ma <- extract_stat(lev, snms, "all", "mua", "bias", yd1_st)
+y2mu <- extract_stat(lev, snms, "all", "mu", "bias", yd2_st)
+y2ma <- extract_stat(lev, snms, "all", "mua", "bias", yd2_st)
+
+# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
+# version is finding correct results
+i <- y_diff1 #pdiff_3 # p_diff1
+jdt <- yd1_st # pd3_st # pd1_st
+bv <- "mu"
+for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
+                c("globmu", "f1"))) {
+  x <- chk[1] 
+  y <- chk[2] 
+  print(paste(".", x, "..", y))
+  rs <- lcu[[y]][[x]] 
+  rs[rs == 0] <- NA
+  l3 <- i[[x]][[y]][[1]]  
+  o <- rs * l3
+  w <- awgts[[y]][[1]] * rs
+  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
+  print(jdt[ol == y & il == x & bvals == bv & bin == "all", bias] ==
+          round(wmu, 2))
+}  # check
 
 ```
+<a href="#top">Back to top</a>
+
+### Bias plots
+```{r, eval=FALSE}
+alph <- c(225, 40)
+#Ccols <- brewer.pal(n = 7, name = "Greys")
+# LC2 <- c(LC[-grep("second|grass", LC)], "all")
+# xax <- seq(0, 7, 6 / (length(yax) - 2)); length(xax)
+x <- c(0, 3, 6, 9, 12, 15)
+w <- 3 / 8
+xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
+#x + xo[j] + o[k]
+o <- c(0, w)
+xa <- sapply(x, function(x) x + xo)
+cx <- c(2, 1.25, 1, 1)
+g1 <- "grey90"
+
+# mul <- c("p1mu", "p2mu", "p3mu", "y1mu", "y2mu")
+mul <- c("p2mu", "p3mu", "y1mu", "y2mu")
+#mal <- c("p1ma", "p2ma", "p3ma", "y1ma", "y2ma")
+mal <- c("p2ma", "p3ma", "y1ma", "y2ma")
+malmul <- list(mal, mul)
+#pyl <- c("mup", "mup", "mup", "muy", "muy")
+pyl <- c("mup", "mup", "muy", "muy")
+
+xl <- c(-0.5, 18)
+yl <- c(-70, 80)
+shd <- c(4.5, 10.5, 16.5)
+cols <- c("red", "orange3", "green4", "blue")
+lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
+pchs <- c("*", "+", "o", "x", "-")
+
+#yl <- range(mu[, 3:8, with = FALSE], mua[, 3:8, with = FALSE])
+#yl <- round(yl / 10) * 10
+yax <- seq(yl[1], yl[2], 10)
+xax <- seq(1, 6, 5 / (length(yax) - 1))
+
+pdf(full_path(p_fig, "yield_bias_scale2.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")#,
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+grid(NA, length(yax), lwd = 1, lty = 1)
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
+for(ii in 1:length(lev)) {
+  lv <- lev[ii]
+  for(i in 1:length(snms)) {
+    pchs1 <- pchs
+    nm <- snms[i]
+    for(k in 1:length(malmul)) {
+      mm <- malmul[[k]]
+      v <- sapply(1:length(mm), function(j) {
+        round(get(mm[j])[ol == lv & il == nm,  bias] / get(pyl[j])[lv] * 100, 1)
+      })
+      pcol <- makeTransparent(cols[i], alpha = alph[k])
+      polyfunc2(xa[i, ii] + o[k], range(v), w = w, col = pcol, bcol = pcol)
+      xx <- rep(x[ii] + xo[i] + o[k], length(v))
+      points(xx, v, pch = pchs, cex = cx)
+    }
+  }
+}
+axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "% bias", outer = TRUE, line = 1)
+legend(x = 12.4, y = -40, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = -40, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = -42, labels = "Absolute", srt = 45, adj = c(0, 0), cex = 0.8)
+text(x = 12.2, y = -42, labels = "Actual", srt = 45, adj = c(0, 0), cex = 0.8)
+legend(x = 12.1, y = 65, legend = c("Prod I", "Prod II", "Yield 1", "Yield 2"), 
+       pch = pchs, pt.cex = c(2.5, 1.5, 1.5, 1.5), bty = "n", cex = 0.8)
+dev.off()
+```
+<a href="#top">Back to top</a>
+
+## The impacts of spatial patterns in bias
+
+Using yield gap estimates as an indicator.
+First we'll bring in mean rainfall as a means of calculating a spatial correlation in yield gap estimates.
+Using data from the Princeton Global Forcing dataset, downloaded from the Africa Flood and Drought Monitor.
+```{r, eval=FALSE}
+# rainfall layer prepared for SA crop subsidy analysis
+load("external/ext_data/climate/mag_dist_clim.rda")
+pre <- calc(pre2, mean)  # 31 year mean rainfall for Africa
+p4s <- projection(raster(nrow = 10, ncol = 10)) # GCS
+presa <- crop(pre, spTransform(sa.shp, p4s))  # crop to SA
+projection(presa) <- p4s  
+presa <- projectRaster(presa, lcu$f25$sa30)  # project to Albers
+presa <- mask(presa, awgts$f25, maskvalue = 0)  # mask
+presa1k <- disaggregate(presa, fact = 25)
+
+# reclassify rainfalls to nearest 100 mm, tweak to catch lower and upper bound
+rclmat <- cbind(seq(50, 750, 100), seq(150, 850, 100), seq(100, 800, 100))
+rclmat[1] <- 15
+rclmat[8, 2] <- 875
+prebin <- crop(reclassify(presa1k, rclmat), namask)  # rainfall bins
+
+# Assume a yield gap over-estimated by 100% in a particular location, say the 
+# 300-400 mm isohyet
+prez <- (prebin == 300) | (prebin == 400)
+notprez <- (prebin < 300) | (prebin > 400)
+g1 <- 0
+g2 <- 1
+
+ygap_bias <- sapply(lev, function(x) {
+  if(x != "f1") { 
+    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun = modal)
+  } else {
+    preagg <- prebin
+  }
+  prez <- (preagg == 300) | (preagg == 400)
+  notprez <- (preagg < 300) | (preagg > 400)
+  sapply(snms, function(y) {
+    yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
+    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]] * prez)
+    pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
+    pg2 <- tareas[[x]] * carea_agg[[x]][[y]] * yg2
+    cellStats(pg - pg2, sum) / cellStats(pg, sum) * 100
+  })
+})
+
+# Assume a yield gap over-estimated by 100% in 400 mm isohyet
+ygap_bias2 <- sapply(lev, function(x) {
+  if(x != "f1") { 
+    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun = modal)
+  } else {
+    preagg <- prebin
+  }
+  prez <- preagg == 400
+  notprez <- preagg != 400
+  sapply(snms, function(y) {
+    yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
+    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]] * prez)
+    pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
+    pg2 <- tareas[[x]] * carea_agg[[x]][[y]] * yg2
+    cellStats(pg - pg2, sum) / cellStats(pg, sum) * 100
+  })
+})
+
+# Assume a yield gap over-estimated by 100% in 500 mm isohyet
+ygap_bias3 <- sapply(lev, function(x) {
+  if(x != "f1") { 
+    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun = modal)
+  } else {
+    preagg <- prebin
+  }
+  prez <- preagg == 500
+  notprez <- preagg != 500
+  sapply(snms, function(y) {
+    yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
+    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]] * prez)
+    pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
+    pg2 <- tareas[[x]] * carea_agg[[x]][[y]] * yg2
+    cellStats(pg - pg2, sum) / cellStats(pg, sum) * 100
+  })
+})
+
+# Print out combined results to latex table
+scen <- c(rep("300-400 mm", nrow(ygap_bias)), rep("400 mm", nrow(ygap_bias)),
+          rep("500 mm", nrow(ygap_bias)))
+ygaps <- cbind.data.frame(scen, "sensor" = rep(snms, 3), 
+                          rbind(ygap_bias, ygap_bias2, ygap_bias3))
+ygap_xtab <- xtable::xtable(ygaps, digits = 1)
+print(ygap_xtab, type = "latex", 
+      file = "paper/figures/ygap-spat-bias.tex")
+
+```
+<a href="#top">Back to top</a>
+
+The actual impact of the spatial bias is fairly low, being largely cancelled out by biases in other direction in other areas.
+This illustrates the advantage of a statistically constrained approach to crop area estimates.
 

### 2015-08-17 R/as.data.table.raster.R

@@ -1,63 +1,0 @@
-#' Transform raster to data.table
-#' 
-#' @param x  Raster* object
-#' @param row.names	`NULL` or character vector giving row names for data frame. 
-#' Missing values are not allowed
-#' @param optional	logical. If `TRUE`, setting row names and converting column 
-#' names (to syntactic names: see make.names) is optional
-#' @param xy logical. If `TRUE`, also return the spatial coordinates
-#' @param centroids	logical. If TRUE return the centroids instead of all spatial 
-#' coordinates (only relevant if xy=TRUE)
-#' @param sepNA	logical. If TRUE the parts of the spatial objects are separated 
-#' by lines that are NA (only if xy=TRUE and, for polygons, if centroids=FALSE
-#' @param ...	 Additional arguments (none) passed to `raster::as.data.frame`
-#' 
-#' @return returns a data.table object
-#' @author etiennebr, from https://gist.github.com/etiennebr/9515738/download#
-#' @note This has been amended to fix the blocksize parameter, which was set to
-#' 2. 
-#' @examples
-#' logo <- brick(system.file("external/rlogo.grd", package="raster"))
-#' v <- as.data.table(logo)
-#' @export
-
-as.data.table.raster <- function(x, row.names = NULL, optional = FALSE, 
-                                 xy = FALSE, inmem = canProcessInMemory(x, 2), 
-                                 ...) {
-  stopifnot(require("data.table"))
-  if(inmem) {
-    v <- as.data.table(raster::as.data.frame(x, row.names=row.names, 
-                                             optional=optional, xy=xy, ...))
-    coln <- names(x)
-    if(xy) coln <- c("x", "y", coln)
-    setnames(v, coln)
-  } else {
-    tr <- blockSize(x)
-    l <- lapply(1:tr$n, function(i) {
-      DT <- as.data.table(as.data.frame(getValues(x, row = tr$row[i], 
-                                                  nrows = tr$nrows[i]), ...))  
-      if(xy == TRUE) {
-        cells <- cellFromRowCol(x, c(tr$row[i], tr$row[i] + tr$nrows[i] - 1),
-                                c(1, ncol(br)))
-        coords <- xyFromCell(x, cell = cells[1]:cells[2])
-        DT[, c("x", "y") := data.frame(xyFromCell(x, cell = cells[1]:cells[2]))]
-     } 
-     DT
-    })
[…]

### 2015-08-17 R/bias-stats.R

@@ -1,0 +1,129 @@
+#' Calculate weighted bias statistics from raster
+#' @param rbins Raster of bins within which statistic should be calculated
+#' @param rwgts Raster defining weighting of pixel towards calculation of stat
+#' @param rmask Binary masking raster 
+#' @param reval Raster containing the bias statistics to evaluate
+#' @param fun Function to evaluate--should be able to handle weights
+#' @param stnm Name of statistic (for output column header)
+#' @param evalnm Name of value encoded in reval's cells (for column header)
+#' @param weighted TRUE or FALSE, whether statistic is weighted or not
+#' @param rnd Rounding value to apply to statistic
+#' @param trim_wgt Remove column of weights from output table
+#' @return A data.table of output statistics
+#' @details The statistic of interest is calculated across i) the entire masked 
+#' in area of the raster, ii) the entire non-NA part of the raster (masked in 
+#' and masked-out), iii) within each bin in the masked in-area, iv) within each 
+#' bin for the masked and un-masked area. There are thus two output statistics 
+#' for each category (whole raster, and for each bin), the second of which is 
+#' indicated by the suffix ".nm", meaning it was calculated for all non-NA areas.
+#' If an unweighted statistic is required, then for a mean value rwgts should be
+#' a raster of zeros
+#' @import data.table
+#' @export
+bias_stats <- function(rbins, rwgts, rmask, reval, fun, stnm, evalnm, weighted, 
+                       rnd = 2, trim_wgt = TRUE) {
+  # stack raster bins, cropland bins, and landcover set
+  if(fun(1:2) == 3) rwgts[] <- 0  # rwgts to 0 for sum
+  if(weighted == FALSE & fun(1:2) == 1.5) rwgts[] <- 1  # rwgts = 1 for mean   
+  s <- stack(list(rbins, rwgts, rmask, reval))
+  names(s) <- c("bin", "wgt", "mask", evalnm)
+  DT <- na.omit(as.data.table.raster(s, xy = TRUE))
+  setkey(DT, "bin")
+  fr1 <- data.table("bvals" = c(stnm, paste0(stnm, ".nm")))
+  fr2 <- data.table("bin" = "all", "N" = nrow(DT[mask == 1, ]))  # unmasked N
[…]

### 2015-08-17 R/dt_to_raster.R

@@ -1,23 +1,0 @@
-#' Transform data.table to raster*
-#' @param dt data.table with x, y coordinates
-#' @param CRSobj crs object
-#' @param filename NULL, else raster* will be written to specified file
-#' @return returns a data.table object
-#' @export
-dt_to_raster <- function(dt, CRSobj, filename = NULL) {
-  if(any(!c("x", "y") %in% colnames(dt))) {
-    stop("dt needs planar coordinates in x, y format")
-  }
-  coordinates(dt) = ~x+y
-  proj4string(dt) <- CRSobj
-  gridded(dt) = TRUE
-  if(ncol(dt) > 1) {
-    dtr <- brick(dt)
-  } else {
-    dtr <- brick(dt)
-  }
-  if(!is.null(filename)) dtr <- writeRaster(dtr, filename)
-  return(dtr)
-}
-
-

### 2015-08-20 Rmd/et-bias.Rmd
ET bias analysis added (et-bias.Rmd), nearly complete. Updated abm district selection also, changed to 2011 dataset rather than 2007.  Redid figure for supplemental

---
title: "ET bias"
author: "Lyndon Estes"
date: "August 17, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Overview

Bias in PET estimates as calculated using different version of the cropland datasets to determine the vegetation properties in VIC.

# Analysis
## Prepare datasets

Bring in cropland datasets, ET grids, reproject the latter to SA Albers, mask, convert to monthly total ET, etc. 
```{r, eval = FALSE}
library(SAcropland)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
p_et <- full_path(p_root, "external/ext_data/vic/")

# cropland data
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
anms <- c("gti", snms)
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100)

# check
# plot(lc_list[[1]])
# plot(gti)
# tst <- raster("external/ext_data/geowikisa_masked.tif")  # check
# plot(round(tst / 100 - lc_list[[4]], 4)) # okay

# aggregate rasters
fact <- 25
lc_agg <- aggregate_rast_list(fact, lc_list)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, list("gti" = gti))  # GTI rasters

# Create namask to remove all NA areas across datasets
sumna <- function(x) sum(x, na.rm = FALSE)
namask <- calc(stack(gti, stack(lc_list)), sumna)
namask2 <- !is.na(namask)  # set NAs to zero
# # namaskt <- raster(full_path(p_data, "namask.tif")) # load in NA mask
# # namaskt2 <- !is.na(namaskt)  # set NAs to zero
# cellStats(is.na(gti), sum); cellStats(namask2, sum); cellStats(namask2t, sum)
# cellStats(namaskt, sum)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts$f25, function(x) {
  r <- x[[1]]
  projection(r) <- sa.shp@proj4string
  r
})  # reshape a bit into list for subsequent use
names(awgts) <- "f25"
mask25k <- awgts$f25 > 0  # mask for et datasets

# read in ET estimates, but drop first year because of VIC problem in January
etf <- dir(p_et, pattern = "nc", full.names = TRUE)
etb <- lapply(etf, function(x) brick(x))
etb <- lapply(etb, function(x) dropLayer(x, i = 1:12))

# reproject them to Albers
etba <- lapply(etb, function(x) {
  b <- projectRaster(x, awgts$f25)
})

# rename layers and convert to total month mm
dts <- seq.Date(as.Date("1981/1/1"), as.Date("2008/12/31"), "days")
allmos <- substr(dts, 6, 7)
allyrs <- substr(dts, 1, 4)
ndays <- unlist(lapply(unique(allyrs), function(x) {
  unname(sapply(unique(allmos), function(y) {
    length(which(allyrs == x & allmos == y))
  }))
}))  # n days in each month in time series

mos <- format(ISOdatetime(2000,1:12,1,0,0,0),"%b")
lnames <- sapply(1981:2008, function(x) paste0(mos, "_", substr(x, 3, 4)))
lnames <- lnames[1:length(lnames)]
# length(lnames) == nlayers(etba[[1]])
etba <- lapply(etba, function(x) {
  b <- mask(x, mask25k, maskvalue = 0)
  names(b) <- lnames
  b
})
names(etba) <- anms

# check that mm/month multiplies through correctly
tst <- etba[[1]] * ndays
plot(tst[[15]] - etba[[1]][[15]] * ndays[15])

# convert to monthly total mm
etba <- lapply(etba, function(x) x * ndays)
# plot(etba[[1]][[1]])

# year indices
yrs <- gsub("*.*_", "", lnames)
mos <- gsub("_.*", "", lnames)
yi <- t(sapply(unique(yrs), function(x) range(which(yrs == x))))
mi <- sapply(unique(mos), function(x) which(mos == x))
```
<a href="#top">Back to top</a>

## Process ET time series

Figure out which months produce maximum ET values, calculate monthly means, overall country-wide time series, etc. 
```{r, eval=FALSE}
# Calculate mean, median, mode of dates in time series when PET max occurs
et_max <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    which.max(x[[ind]])
  }))
  s <- stack(calc(bi, max), calc(bi, modal), calc(bi, mean), calc(bi, median))
  names(s) <- c("maxmax", "mode", "mean", "median")
  s
})

# plot to see which statistic picks out PET peak date most effectively
plot(et_max$gti)  # all the same, not coherent
plot(et_max$gti %in% 6:9)  # median looks most sensible and coherent 
plot(et_max$gti - et_max$sa30)  # some differences in date of max
plot(et_max$gti - et_max$globmu) 
plot(et_max$gti - et_max$modmu)  
plot(et_max$gti - et_max$geow)

# Monthly mean values
et_momu <- lapply(etba, function(x) {
  bi <- stack(lapply(1:ncol(mi), function(y) {
    ind <- mi[, y]
    calc(x[[ind]], mean)
  }))
  names(bi) <- unique(mos)
  bi
})

# do a focal max to separate out noisy areas
et_medsm <- focal(et_max$gti$median, w = matrix(1, 3, 3), fun = modal, pad = 3)
buf <- (!is.na(et_max$gti$median) - !is.na(et_medsm)) * et_max$gti$median
et_medsm[is.na(et_medsm)] <- 0
et_medsm <- buf + et_medsm  # fill back in areas edged out by moving window
# plot((et_medsm > 0) - (et_momu$gti[[1]] > 0))

# 1 month before and after peak
et_meds_m1 <- et_medsm - 1
# plot(et_medsm - 1)  # a few zero areas
et_meds_m1[et_meds_m1 < 1] <- 12  # set to 12 (December) if median - 1 = 0
et_meds_p1 <- et_medsm + 1
# plot(et_medsm + 1)  # nothing greater than 12

# calculate 3 month peak ET
gs_peak <- lapply(et_momu, function(x) {
  et1 <- stackSelect(x, et_meds_m1)
  et2 <- stackSelect(x, et_medsm)
  et3 <- stackSelect(x, et_meds_p1)
  stack(et1, et2, et3)
})

# Annual mean values
et_amu <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    calc(x[[ind]], sum)
  }))
  et_ts <- cellStats(bi, mean)
  names(et_ts) <- 1981:2008
  list("gridmu" = calc(bi, mean), "cntry_ts" = et_ts)
})

# plot annual mean values
rbcol <- rainbow(4)
plot(1981:2008, et_amu$gti$cntry_ts, type = "l", las = 2)
for(i in 2:5) lines(1981:2008, et_amu[[i]]$cntry_ts, col = rbcol[i])

# figure out min_max years
mm_yr <- sapply(list(which.min, which.max), function(x) x(et_amu$gti$cntry_ts))

# annual totals
et_mutot <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    calc(x[[ind]], sum)
  }))
})
et_mm <- lapply(et_mutot, function(x) x[[mm_yr]])
```
<a href="#top">Back to top</a>

## Calculate ET differences
```{r, eval=FALSE}
# difference rasters and convert to percent
# between time series annual mean
ilist <- lapply(et_amu[2:5], function(x) x[[1]])
a <- et_amu$gti$gridmu
amu_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between time series mins
ilist <- lapply(et_mm[2:5], function(x) x[[1]])
a <- et_mm$gti[[1]]
mumin_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between time series maxs
ilist <- lapply(et_mm[2:5], function(x) x[[2]])
a <- et_mm$gti[[2]]
mumax_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between mean peak growing season ET (3 month total)
ilist <- lapply(gs_peak[2:5], function(x) calc(x, sum))
a <- calc(gs_peak$gti, sum)
gs3_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between median max growing season ET (3 month total)
ilist <- lapply(gs_peak[2:5], function(x) x[[2]])
a <- gs_peak$gti[[2]]
gsmx_diff <- lapply(ilist, function(b) {
  bl <- (a - b) / a * 100
})
```
<a href="#top">Back to top</a>

## Plot difference maps
### Reshape and disaggregate lists
```{r, eval = FALSE}
# function to reshape lists for use in disaggregate_rast_list
disaggl <- function(rl, lev = "f25") {
  bo <- lapply(rl, function(x) {
    b <- list(x)
    names(b) <- lev
    b
  })
  bo  
}

# reshape in list
l2dag <- list(amu_diff, mumax_diff, mumin_diff, gs3_diff, gsmx_diff)
names(l2dag) <- c("mu", "mumx", "mumn", "gs", "gspk")
l2dag <- lapply(l2dag, disaggl)

# disaggregate
l2dagd <- lapply(l2dag, function(x) {
  disaggregate_rast_list(snms, "f25", x, namask)
})

# stats for plotting
dstats <- function(dlist) {
  statsd <- lapply(dlist, function(x) {
    sapply(x, function(y) {
      c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
    })
  })
}
```
<a href="#top">Back to top</a>

### Plot maps
```{r, eval=FALSE}
# calculate single set of breakranges for all sets
statsl <- lapply(l2dag, dstats)
rng <- do.call(cbind, lapply(statsl, function(x) {
  ilims <- c(ceiling(min(sapply(x, function(y) y[3, ]))), 
             floor(max(sapply(x, function(y) y[21, ]))))
  olims <- range(sapply(x, function(y) range(y)))
  c(olims[1], ilims, olims[2])
}))
rngdt <- data.table(t(rng))
rng <- unname(unlist(rngdt[, lapply(.SD, function(x) x[which.max(abs(x))])]))

fnm <- "et_bias_map.pdf"
legtext <- rep("%", 5)
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
ecap <- c("Annual mean", "TS max", "TS min", "Peak 3", "Peak")

rnder <- function(x) {
  dig <- nchar(max(round(x)))
  rmat <- cbind(1:8, c(1, 5, 10, 20, 50, 100, 100, 1000))
  rnd <- rmat[dig, 2]
  round(c(floor(x[1] / rnd), ceiling(x[2] / rnd))) * rnd 
}

brks <- c(seq(floor(rng[1]), -5, by = 5), -1, 1, seq(5, ceiling(rng[4]), 5))
cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)
cols[c(1, length(cols))] <- c("darkred", "purple3")
pdf(full_path(p_fig, fnm), height = 6, width = 7)
par(mfcol = c(4, 5), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(j in 1:length(l2dagd)) {
  print(names(l2dagd)[j])
  for(k in 1:length(snms)) {
    rl <- l2dagd[[j]][[k]]$f25
    print(snms[k])
    plot(sa.shp, lty = 0)
    image(rl, add = TRUE, col = cols, breaks = brks, legend = FALSE)
    if(j == 1) mtext(mcap[k], side = 2, line = 1, cex = cx)
    if(k == 1) mtext(ecap[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
            legend.vals = brks, longdims = c(0.2, 0.8), 
            shortdims = c(0.06, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.1, 0.3), c(-1, -0.5)), 
            cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
dev.off()
```
<a href="#top">Back to top</a>

## Calculate bias statistics

### Mean percent bias in ET estimates
```{r, eval=FALSE}
lev_vec <- "f25"
# create mask for non-cropland areas, unioning GTI and each LC, filtering out
# areas of no-cropland (<1/2% total cover)
lcu <- lapply(lev_vec, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(round(gti_agg[[x]]$gti * 100) > 0)
    lc_gt0 <- Which(round(lc_agg[[x]][[y]] * 100) > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev_vec

# cropland cover bins, for looking at error as a function of cover
binv <- seq(0, 1, 0.05)
bins <- lapply(gti_agg[[2]], function(x) {
  cut(x$gti, breaks = binv, include.lowest = TRUE)
})
names(bins) <- lev_vec

# weighted mean functions
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)

# calculate bias stats
vnms <- names(l2dag)
bstats <- lapply(vnms, function(x) {
  a <- bias_stats_list(bins, awgts, lcu, l2dag[[x]], wm, "mu", "bias", TRUE)
  b <- bias_stats_list(bins, awgts, lcu, l2dag[[x]], wma, "mua", "bias", TRUE)
  stat <- rbind(a, b)
})
names(bstats) <- vnms

# checks to see if stats correct
# bias_stats* functions correct
chk <- bias_stats(bins$f25, awgts$f25, lcu$f25$globmu, l2dag$mu$globmu$f25, wm,
                  "mu", "bias", TRUE)
all(chk[bvals == "mu", bias] == bstats$mu[il == "globmu" & bvals == "mu", bias])
# compared to raster
for(i in snms) {
  msk <- lcu$f25[[i]]
  msk[msk == 0] <- NA
  a <- l2dag$mu[[i]]$f25 * msk
  b <- awgts$f25 * msk
  d <- weighted.mean(values(a), values(b), na.rm = TRUE)
  print(round(d, 2) == bstats$mu[il == i & bvals == "mu" & bin == "all", bias])
}  # check

# extract stats
mu <- lapply(bstats, function(x) {
  extract_stat(lev_vec, snms, "all", "mu", "bias", x)
})
mua <- lapply(bstats, function(x) {
  extract_stat(lev_vec, snms, "all", "mua", "bias", x)
})

# reshape
mus <- rbindlist(lapply(names(mu), function(x) cbind(x, mu[[x]])))
mus[, ol := NULL]
setnames(mus, c("x", "il"), c("ETvar", "dataset"))
setkeyv(mus, c("dataset", "ETvar"))
muas <- rbindlist(lapply(names(mua), function(x) cbind(x, mua[[x]])))
muas[, ol := NULL]
setnames(muas, c("x", "il", "bias"), c("ETvar", "dataset", "absbias"))
setkeyv(muas, c("dataset", "ETvar"))
mutab <- mus[muas]
mutab[, lapply(.SD, function(x) max(abs(x))), .SDcols = 3:4]

mutab_xtab <- xtable::xtable(mutab, digits = 1)
print(mutab_xtab, type = "latex", 
      file = "paper/figures/et-bias.tex")
```
<a href="#top">Back to top</a>

The largest (in an absolute sense) mean actual bias is -0.6%, while the largest mean absolute bias is 1.43%, so not much here.

### Bias in relation to rainfall

We'll end by having a look at the relationship between bias and rainfall

```{r, eval = FALSE}
# rainfall layer prepared for SA crop subsidy analysis
load("external/ext_data/climate/mag_dist_clim.rda")
pre <- calc(pre2, mean)  # 31 year mean rainfall for Africa
p4s <- projection(raster(nrow = 10, ncol = 10)) # GCS
presa <- crop(pre, spTransform(sa.shp, p4s))  # crop to SA
projection(presa) <- p4s  
presa <- projectRaster(presa, mask25k)  # project to Albers
presa <- mask(presa, awgts$f25, maskvalue = 0)  # mask

plot(presa)
plot(abs(l2dag$mu$globmu$f25))
plot(presa, abs(l2dag$mu$globmu$f25))

```

### 2015-08-20 Rmd/magisterial-district-ABM-select.Rmd
ET bias analysis added (et-bias.Rmd), nearly complete. Updated abm district selection also, changed to 2011 dataset rather than 2007.  Redid figure for supplemental

@@
     toc: yes 
+    number_sections: true
 ---
 
-## Overview
+# Overview
 
@@
 
-## Data
+# Data
 ```{r, eval=FALSE}
@@
 load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
-paths <- full_path(p_data, dir(p_data, paste0("cover*.*07sum_mask*")))
-gti <- raster(paths)  # gti 2007 rather than 2011, to match census data
+paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
+gti <- raster(paths[2])  # gti 2011
 
@@
 snms <- c("sa30", "globmu", "modmu", "geow")
-dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2007))
+dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2011))
 lc_list <- gti - dlist_1km
@@
 mdfilt <- mu[mu$na < 20 & mu$f > 5 & mu$N < 1500, ]  
-md@data[mdsel, 7:8]
 mdsel <- as.numeric(row.names(mdfilt))
@@
 
-### Selection
+# Selection
 
@@
 
+# Deprecated: earlier used when running 2007 rather than 2011
 # Let's compare the fractions between the GTI 2007 and GTI 2011 datasets for 
 # interest's sake
-gti_abm <- extract(gti, abm_md)
-abm_mu <- data.frame(t(sapply(gti_abm, function(x) {
-  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
-})))
-colnames(abm_mu) <- c("N", "na", "f")
-abm_mu$f - mu[mdsel[mdsel %in% c(269, 274:275, 288)], "f"]
+# gti_abm <- extract(gti, abm_md)
+# abm_mu <- data.frame(t(sapply(gti_abm, function(x) {
+#   c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
+# })))
+# colnames(abm_mu) <- c("N", "na", "f")
+# abm_mu$f - mu[mdsel[mdsel %in% c(269, 274:275, 288)], "f"]
  # -1.0  0.0 -0.4 -0.2  # tinf differences!
@@
 
-### Extraction
+## Extraction
 ```{r, eval=FALSE}
@@
 lc4 <- crop(lc_list, extent(abm_md))
-cb <- brick(stack(gti4, lc4), filename = "external/ext_data/cropland-md4.tif")
+cb <- brick(stack(gti4, lc4), filename = "external/ext_data/cropland-md4.tif", 
+            overwrite = TRUE)
 ```
 
-### Visuals
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-library(raster)
-library(rgdal)
-library(lmisc)
+## Plots
+```{r, eval = FALSE}
+brks1 <- seq(0, 100, 10)
+cols1 <- rev(terrain.colors(length(brks1) - 1))
+brks2 <- c(-100, -80, -60, -40, -20, -10, -5, -1)
+brks2 <- c(brks2, rev(abs(brks2)))
+cols2 <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks2) - 1)
+cols2[c(1, length(cols2))] <- c("darkred", "purple3")
+cx <- 1.1
+lcol <- "black"
+mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
 
-setwd(full_path(proj_root("SAcropland"), "SAcropland"))
-md <- readOGR("external/ext_data/mag_dists.sqlite", layer = "mag_dists")
-abm_md <- readOGR("external/ext_data/md4.sqlite", layer = "md4")
-cb <- brick("external/ext_data/cropland-md4.tif")
+# layout.show()
+# pdf(full_path(p_fig, "abm-selected-districts.pdf"), height = 7, width = 7)
+png(full_path(p_fig, "abm-selected-districts.png"), height = 700, width = 700)
+par(mfrow = c(3, 2), mar = c(0, 0, 0, 0), oma = c(2, 0, 0, 0))
+plot(md, col = "grey", border = "transparent")
+plot(abm_md, col = "red", border = "grey30", add = TRUE)
+par(mar = c(3, 0, 0, 0))
+plot(abm_md)
+plot(cb[[1]], axes = FALSE, box = FALSE, breaks = brks1, col = cols1, add = TRUE, 
+     legend = FALSE)
+plot(abm_md, add = TRUE)
+for(i in 2:5) {
+  plot(abm_md)
+  plot(cb[[1]] - cb[[i]], axes = FALSE, box = FALSE, breaks = brks, col = cols, 
+       add = TRUE, legend = FALSE)
+  plot(abm_md, add = TRUE, lwd = 2)
+  mtext(mcap[i - 1], side = 1, line = -4, adj = 0.7)
+}
+nct <- length(brks1) - 1
+flex_legend(ncuts = nct, legend.text = "% cropland", legend.vals = rep("", nct),
+            longdims = c(0.6, 0.9), shortdims = c(0.72, 0.012), colvec = cols1, 
+            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
+            legend.pos = c(4, 3), leg.adj = list(c(4.2, 0), c(-0.5, -0.5)), 
+            cex.val = 1.1, textcol = lcol, bordercol = lcol)
+tc <- rect_coords(0.6, 0.9, nct, constEWorNS = 0.72, resEWorNS = 0.02)
+xs <- c(tc[[1]][1, 1], sapply(tc, function(x) x[2, 1]))
+ys <- sapply(tc, function(x) x[1, 2])
+text(xs, ys, labels = brks1, srt = 270, adj = c(-0.15, 0.5), cex = 1.1)
+nct <- length(brks2) - 1
+flex_legend(ncuts = nct, legend.text = "% bias", legend.vals = rep("", nct),
+            longdims = c(0.2, 0.8), shortdims = c(0.05, 0.015), colvec = cols2, 
+            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
+            legend.pos = c(4, 6), leg.adj = list(c(4.2, 0), c(-0.5, -0.5)), 
+            cex.val = 1.1, textcol = lcol, bordercol = lcol)
+tc <- rect_coords(0.2, 0.8, nct, constEWorNS = 0.047, resEWorNS = 0.012)
+xs <- c(tc[[1]][1, 1], sapply(tc, function(x) x[2, 1]))
+ys <- sapply(tc, function(x) x[1, 2])
+text(xs, ys, labels = brks2, srt = 270, adj = c(-0.1, 0.5), cex = 1.1)
+dev.off()
 ```
 
-```{r}
-par(mar = c(0, 0, 0, 0))
-plot(md, col = "grey", border = "transparent")
-plot(abm_md, col = "red", border = "grey30", add = TRUE)
-```
-
-The four selected districts.
-
-```{r}
-par(mar = c(0, 0, 0, 0))
-plot(cb[[1]], axes = FALSE, box = FALSE)
-plot(abm_md, add = TRUE)
-```
-
-2007 cropland fraction in the four districts.
-
-```{r}
-par(mar = c(0, 0, 0, 0), mfrow = c(2, 2))
-for(i in 2:5) {
-  plot(cb[[i]], axes = FALSE, box = FALSE, zlim = c(0, 101))
-  plot(abm_md, add = TRUE)
-}
-```
-
-Cropland fractions estimated by South Africa's landcover product, GLobCover, MODIS, and Geo-Wiki for the same 4 districts. Note the substantial variability.
+![figure](../paper/figures/abm-selected-districts.png)

### 2015-08-20 Rmd/yield-bias.Rmd
ET bias analysis added (et-bias.Rmd), nearly complete. Updated abm district selection also, changed to 2011 dataset rather than 2007.  Redid figure for supplemental

@@
     yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
-    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]] * prez)
+    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2*cyld_agg[[x]][[y]] * prez)
     pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
@@
     yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
-    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]] * prez)
+    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]]*prez)
     pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
@@
   if(x != "f1") { 
-    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun = modal)
+    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun=modal)
   } else {

### 2015-08-20 Rmd/agent-allocation-bias.Rmd
Renamed ABM file

--- 2015-08-20 Rmd/magisterial-district-ABM-select.Rmd
@@
 ---
-title: "Districts for ABM"
+title: "Bias in Allocation of ABM Agents"
 author: "Lyndon Estes"
@@
 
-Selection of magisterial districts and associated landcover sets for use in ABM example.
+Selection of magisterial districts and associated landcover sets for use in ABM example, and analysis of bias after initial allocation of agents performed based on available farmland estimated from 100 m downscaled versions of the landcover data.
 

### 2015-08-21 Rmd/agent-allocation-bias.Rmd
Updated agent bias analysis and figures

@@
 library(lmisc)
+library(readxl)
 
@@
 ```
+<a href="#top">Back to top</a>
 
-# Selection
+## Magisterial District selection
 
@@
 ```
+<a href="#top">Back to top</a>
 
-## Extraction
+### Extract cropland percentage from selected districts
 ```{r, eval=FALSE}
@@
 ```
+<a href="#top">Back to top</a>
 
-## Plots
+### Map selected districts, cropland percentage, and bias
 ```{r, eval = FALSE}
@@
 # layout.show()
-# pdf(full_path(p_fig, "abm-selected-districts.pdf"), height = 7, width = 7)
-png(full_path(p_fig, "abm-selected-districts.png"), height = 700, width = 700)
+pdf(full_path(p_fig, "abm-selected-districts.pdf"), height = 7, width = 7)
+# png(full_path(p_fig, "abm-selected-districts.png"), height = 700, width = 700)
 par(mfrow = c(3, 2), mar = c(0, 0, 0, 0), oma = c(2, 0, 0, 0))
@@
 ```
+<a href="#top">Back to top</a>
 
 ![figure](../paper/figures/abm-selected-districts.png)
+
+<a href="#top">Back to top</a>
+
+# Downscaling and agent allocation
+
+Performed separately and described in the methods.
+
+# Analysis of agent allocation bias
+```{r, eval = FALSE}
+abias <- data.table(read_excel(full_path(p_data, "abm/agent-bias.xlsx")))
+abias <- abias[Alg == 1]
+setkey(abias, MD)
+hhs <- data.table(read_excel(full_path(p_data, "abm/agent-bias.xlsx"), 
+                             sheet = 2))
+setkey(hhs, MD)
+abias <- abias[hhs]
+abias
+
+# calculate district level bias as percent
+pctbias <- abias[var == "ag_area", 
+                 lapply(list(sa30, globmu, modmu, geow), function(x) {
+                   (gti - x) / gti * 100
+                 })]
+
+# HHs without land as percent total households
+noland <- abias[var == "HHs_no_land", 
+                lapply(list(sa30, globmu, modmu, geow), function(x) {
+                  x / HHs * 100
+                })]
+
+# unallocated farmland
+unalloc <- round(abias[var == "ag_area_unalloc", snms, with = FALSE] / 
+  abias[var == "ag_area", snms, with = FALSE] * 100, 2)
+
+cexs <- seq(1, 2, 1 / 3)
+pdf(full_path(p_fig, "agent-bias.pdf"), height = 3.5, width = 7)
+cx <- c(0.7, 0.8, 0.6)
+# png(full_path(p_fig, "agent-bias.png"), height = 350, width = 700)
+par(mfrow = c(1, 2), mar = c(3, 2.5, 3, 1), mgp = c(1.5, 0.25, 0), tcl = -0.2, 
+    oma = c(0, 0, 0, 4))
+scols <- c("red", "orange3", "green4", "blue")
+cols <- c(rep("red", 4), rep("orange3", 4), rep("green4", 4), rep("blue", 4))
+plot(unlist(pctbias), unlist(noland), col = cols, cex = cexs, pch = 20, 
+     ylab = "% HH without farms", xlab = "% cropland bias", cex.lab = cx[2], 
+     cex.axis = cx[1])
+plot(unlist(pctbias), unlist(unalloc), col = cols, cex = cexs, pch = 20, 
+     ylab = "% land unallocated", xlab = "% cropland bias", cex.lab = cx[2], 
+     cex.axis = cx[1])
+par(xpd = NA)
+xst <- seq(0.86, 0.96, 0.07 / 3)
+xy <- cbind(grconvertX(xst, from = "ndc"), 
+            grconvertY(rep(0.6, 5), from = "ndc"))
+for(i in 2:length(xst)) {
+  if(i == 2) {
+    leg <- 1:4
+  } else {
+    leg <- rep("", 4)
+  }
+  legend(x = xy[i, 1], y = xy[i, 2], legend = leg, pch = 20, cex = cx[3], 
+         adj = c(4.5, 0.75),  
+         col = scols[i - 1], pt.cex = cexs, bty = "n")#cex = cexs,)
+}
+xy <- cbind(grconvertX(xst, from = "ndc"), 
+            grconvertY(rep(0.58, 5), from = "ndc"))
+text(x = xy[, 1], y = xy[, 2], labels = c("District", mcap), srt = 90, #pos = 2, 
+     adj = c(0, 1.5), col = c("black", scols), cex = cx[1])
+dev.off()
+
+```
+
+<a href="#top">Back to top</a>
+
+![figure](../paper/figures/agent-bias.png)
+
+<a href="#top">Back to top</a>

### 2015-08-24 paper/main.tex
Paper main manuscript .tex and .bib added, up through intro.

%% PNAStwoS.tex
%% Sample file to use for PNAS articles prepared in LaTeX
%% For two column PNAS articles
%% Version1: Apr 15, 2008
%% Version2: Oct 04, 2013

%% BASIC CLASS FILE
\documentclass{pnastwo}

%% ADDITIONAL OPTIONAL STYLE FILES Font specification

%\usepackage{pnastwoF}
\usepackage{cite}
\usepackage{lineno}
\setlength\linenumbersep{4pt}

%% OPTIONAL MACRO DEFINITIONS
\def\s{\sigma}
%%%%%%%%%%%%
%% For PNAS Only:
\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
\copyrightyear{2008}
\issuedate{Issue Date}
\volume{Volume}
\issuenumber{Issue Number}
%\setcounter{page}{2687} %Set page number here if desired
%%%%%%%%%%%%

\begin{document}

\title{Quantifying the impacts of bias in landcover data on global change analyses}

\author{Lyndon Estes\affil{1}{Princeton University, Princeton, NJ USA},
Peng Chen\affil{2}{Indiana University, Bloomington, IN USA},
Stephanie Debats\affil{1}{Princeton University, Princeton, NJ USA},
Tom Evans\affil{2}{Indiana University, Bloomington, IN USA},
Fanie Ferreira\affil{3}{GeoTerraImage, Pretoria, RSA},
Gabrielle Ragazzo\affil{1}{Princeton University, Princeton, NJ USA},
Justin Sheffield\affil{1}{Princeton University, Princeton, NJ USA}
\and
Kelly Caylor\affil{1}{Princeton University, Princeton, NJ USA}}

\contributor{Submitted to Proceedings of the National Academy of Sciences
of the United States of America}

%%%Newly updated.
%%% If significance statement need, then can use the below command otherwise just delete it.
%\significancetext{LDE developed the concept of the study, conducted the analysis, data interpretation and drafted the manuscript. \color{red}{To be completed}}

\maketitle

\begin{article}
\begin{abstract}
{Blah blah.}
\end{abstract}

\keywords{landcover | bias | remote sensing | agriculture | crop yield | harvested area | carbon | agent-based model | landscape}

\abbreviations{GTI, GeoTerraImage}
\linenumbers

\dropcap{T}he nature and distribution of landcover is a fundamental determinant of many environmental and social processes that drive or are affected by global change \cite{lambin_modelling_1997}, such as agricultural production and food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010,gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015,luoto_predicting_2004}, or demographic changes \cite{linard_assessing_2010}. Landcover maps are therefore critical for understanding the nature and impact of such changes \cite{see_improved_2015}, and they need to be accurate at the finest scales at which the underlying processes operate. For example, agricultural productivity and nutrient loadings can vary greatly between neighboring fields, and field sizes are often $<$2 hectares in regions where smallholder farming still dominates \cite{jain_mapping_2013, debats_generalized_????}. To understand agriculturally driven processes, it is thus necessary to accurately delineate fields at their smallest grain size, and to do so at regional to global scales to have a consistent set of maps.

Landcover data can only be developed with satellite imaging, but often the average size class of the cover type of interest is smaller than the sensor resolution, or spectrally indistinct from other neighboring covers, which propagates classification error \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}. The result is that landcover datasets are generally inaccurate at finer scales and greatly differ between one another, particularly in those parts of the world undergoing the most rapid land use changes, where the aforementioned sources of bias tend to be most pronounced \cite{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011}.

These errors are well-known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of efforts underway to improve landcover maps, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}. What is less known is the degree to which these errors bias measurements built upon the distributional and areal information in landcover. An impediment to this understanding is that the errors are hard to quantify because spatially extensive reference data are not available for most regions of the world--particularly over Africa and other developing regions. Errors assessment therefore typically rely on a small number of ground truth points or survey data aggregated to political boundaries. For this reason, we have a better understanding of the biases between landcover datasets or in relation to country-level statistics (e.g. \cite{fritz_comparison_2010,fritz_cropland_2011}) than we do of how error changes over spatial gradients or as a function of aggregation scale.

Not being able to full quantify the errors in landcover maps of course makes it difficult, if not impossible, to quantify their impact on downstream analyses. There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, and carbon flux \cite{quaife_impact_2008} and human population estimates \cite{linard_assessing_2010}, but these use either simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}. The exception is \cite{schmit_limitations_2006}, who use a high quality, ground-collected reference map detailing farm land use parcels in central Belgium, but the number of sites and region were both fairly restricted, and the parcels were not spatially contiguous.

There is thus an urgent need in global change science to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.  We address this need in this study, using a unique, high accuracy agricultural landcover map for South Africa to quantify the errors in several latest generation landcover maps that are broadly used in global change studies.  We use this information to examine how i) landcover properties and related classification schemes influence error, ii) how these errors change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and 3) how these errors propagate through several different forms of downstream analyses that broadly represent the global change research focus areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.

\section{}

\subsection{Data Sources}
%large gradient ($\sim \displaystyle{ \inlinefrac{1}{\delta}}$).

\section{Discussion}
\subsection{Blather}
% $\theta$ changes from 0 to 1. (see Figure \ref{afoto}).That means we are considering $\theta$  of the form
%For these solutions we have the following

%\begin{remark}
%Note that equation \eqref{theta} specifies  the function $\varphi$
%up to an error of order $\delta$. Theorem 1 provides an evolution
%equation for the function $\varphi$ up to an error of order
%$\delta |log \delta|$.
%\end{remark}

%(see \eqref{weaksol}). 
\subsection{More blather}

\begin{materials}
\section{Methods} 
Perhaps it is right {\it SI Materials and Methods}.

\section{Digital RCD Analysis}

\end{materials}

\appendix[App 1]

\appendix
This is an example of an appendix without a title.

\begin{acknowledgments}
I thank everyone tearfully. 
\end{acknowledgments}

% bib solution from here
% http://tex.stackexchange.com/questions/167650/is-there-a-more-recent-bibliography-style-file-bst-for-pnas
% https://github.com/jburon/pnas2011.bst
\bibliographystyle{pnas2011} 
\bibliography{bias}

%\begin{thebibliography}{10}
%\bibitem{BN}
%M.~Belkin and P.~Niyogi, {\em Using manifold structure for partially
%  labelled classification}, Advances in NIPS, 15 (2003).
%
%\bibitem{BBG:EmbeddingRiemannianManifoldHeatKernel}
%P.~B\'erard, G.~Besson, and S.~Gallot, {\em Embedding {R}iemannian
%  manifolds by their heat kernel}, Geom. and Fun. Anal., 4 (1994),
%  pp.~374--398.
%\end{thebibliography}

\end{article}

%\begin{figure}
%\centerline{\includegraphics[width=.4\textwidth]{figsamp.eps}}
%\caption{LKB1 phosphorylates Thr-172 of AMPK$\alpha$ \textit{in vitro}
%and activates its kinase activity.}\label{afoto}
%\end{figure}
%
%\begin{figure*}[ht]
%\begin{center}
%\centerline{\includegraphics[width=.7\textwidth]{figsamp.eps}}
%\caption{LKB1 phosphorylates Thr-172 of AMPK$\alpha$ \textit{in vitro}
%and activates its kinase activity.}\label{afoto2}
%\end{center}
%\end{figure*}
%
%\begin{table}[h]
%\caption{Repeat length of longer allele by age of onset class.
%This is what happens when the text continues.}
%\begin{tabular}{@{\vrule height 10.5pt depth4pt  width0pt}lrcccc}
%&\multicolumn5c{Repeat length}\\
%\noalign{\vskip-11pt}
%Age of onset,\\
%\cline{2-6}
%\vrule depth 6pt width 0pt years&\multicolumn1c{\it n}&Mean&SD&Range&Median\\
%\hline
%Juvenile, 2$-$20&40&60.15& 9.32&43$-$86&60\\
%Typical, 21$-$50&377&45.72&2.97&40$-$58&45\\
%Late, $>$50&26&41.85&1.56&40$-$45&42\tablenote{The no. of wells for all samples was 384. Genotypes were
%determined by mass spectrometric assay. The $m_t$ value indicates the
%average number of wells positive for the over represented allele.}
%\\
%\hline
%\end{tabular}
%\end{table}
%
%
%\begin{table*}[ht]
%\caption{Summary of the experimental results}
%\begin{tabular*}{\hsize}
%{@{\extracolsep{\fill}}rrrrrrrrrrrrr}
%\multicolumn{3}{l}{Parameters}&
%\multicolumn{5}{c}{Averaged Results}&
%\multicolumn{5}{c}{Comparisons}\cr
%\hline
%\multicolumn1c{$n$}&\multicolumn1c{$S^*_{MAX}$}&
%\multicolumn1c{$t_1$}&\multicolumn1c{\ $r_1$}&
%\multicolumn1c{\ $m_1$}&\multicolumn1c{$t_2$}&
%\multicolumn1c{$r_2$}&\multicolumn1c{$m_2$}
%&\multicolumn1c{$t_{lb}$}&\multicolumn1c{\ \ $t_1/t_2$}&
%$r_1/r_2$&$m_1/m_2$&
%$t_1/t_{lb}$\cr
%\hline
%10\tablenote{Stanford Synchrotron Radiation Laboratory (Stanford University,
%Stanford, CA)}&1\quad &4&.0007&4&4&.0020&4&4&1.000&.333&1.000&1.000\cr
%10\tablenote{$R_{\rm FREE}=R$ factor for the $\sim 5$\% of the randomly
%chosen unique ref\/lections not used in the ref\/inement.}&5\quad &50&.0008&8&50&.0020&12&49&.999&.417&.698&1.020\cr
%100\tablenote{Calculated for all observed data}&20\quad &2840975&.0423&95&2871117&.1083&521&---&
%.990&.390&.182&---\ \ \cr
%\hline
%\end{tabular*}
%\end{table*}

\end{document}

### 2015-08-25 paper/main.tex
Results outline added
Drafted up to Results.

@@
 
-\abbreviations{GTI, GeoTerraImage}
+\abbreviations{GTI, GeoTerraImage; SSA, sub-Saharan Africa}
 \linenumbers
@@
 Errors assessment therefore typically rely on a small number of ground truth points or survey data aggregated to political boundaries.
-For this reason, we have a better understanding of the biases between landcover datasets or in relation to country-level statistics (e.g.
-\cite{fritz_comparison_2010,fritz_cropland_2011}) than we do of how error changes over spatial gradients or as a function of aggregation scale.
+For this reason, we have a better understanding of the biases between landcover datasets or in relation to country-level statistics \cite{fritz_comparison_2010,fritz_cropland_2011,kaptue_tchuente_comparison_2011} than we do of how error changes over spatial gradients or as a function of aggregation scale.
 
-Not being able to full quantify the errors in landcover maps of course makes it difficult, if not impossible, to quantify their impact on downstream analyses.
-There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, and carbon flux \cite{quaife_impact_2008} and human population estimates \cite{linard_assessing_2010}, but these use either simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}.
+Being unable to fully quantify the errors in landcover maps of course makes it difficult, if not impossible, to quantify their impact on downstream analyses.
+There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, and carbon flux \cite{quaife_impact_2008} and human population estimates \cite{linard_assessing_2010}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}.
 The exception is \cite{schmit_limitations_2006}, who use a high quality, ground-collected reference map detailing farm land use parcels in central Belgium, but the number of sites and region were both fairly restricted, and the parcels were not spatially contiguous.
 
-There is thus an urgent need in global change science to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.
+There is thus an urgent need to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.
 We address this need in this study, using a unique, high accuracy agricultural landcover map for South Africa to quantify the errors in several latest generation landcover maps that are broadly used in global change studies.
-We use this information to examine how i) landcover properties and related classification schemes influence error, ii) how these errors change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and 3) how these errors propagate through several different forms of downstream analyses that broadly represent the global change research focus areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.
+We use this information to examine how i) landcover properties and related classification schemes influence error, ii) how these errors change with aggregation scale, with the specific goal of determining ``safe'' scales for making area-based calculations, and 3) how these errors propagate through several different forms of downstream analyses that broadly represent the global change research focus areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.
 
-\section{}
+\vspace{-0.5 cm}
+\section{Study area and landcover data}
+Our study focused on South Africa, which comprises nearly 6\% of sub-Saharan Africa's (SSA) landmass, and has a large, diverse agricultural sector, ranging from large commercial operations to smallholder farms \cite{hardy_rainfed_2011,estes_using_2014}.
+This diversity suggests that the country's agricultural landcover spans the range of types that are found throughout the rest of SSA.
 
-\subsection{Data Sources}
+The South African government commissioned a whole country cropland boundary map in order to stratifying the annual aerial crop type census used to calculate harvested area estimates \cite{fourie_better_2009}.
+The map was made by trained workers who visually interpreted high resolution satellite imagery and manually digitized field boundaries following a standardized mapping protocol.
+The resulting vectorized field maps, which were made in 2007 and updated in 2011, provide a unique, high accuracy reference dataset of both crop field distribution and size classes.
+We converted the vector data into a rasterized estimated of cropland percentage at 1 km resolution (henceforth the ``reference map''), which was 97\% accurate in distinguishing cropped from non-cropped areas.
+
+We compared our reference percent cropland estimates to those created from four satellite-derived landcover datasets.
+We obtained South Africa's 30 m resolution National Landcover map (SA-LC) for 2009 \cite{sanbi_national_2009}, the 500 m resolution MODIS Landcover for 2011 \cite{land_processes_distributed_active_archive_center_lp_daac_modis_2011, friedl_modis_2010}, the 300 m resolution GlobCover 2009 \cite{arino_global_2012}, and the new 1 km Geo-wiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}.
+We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite{fry_completion_2009},  MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013,shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and is the current state of the art for agricultural landcover maps.
+We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates (hereafter simply ``cropland maps''), resulting in 4 maps to compare to the reference.
+
+\vspace{-0.5 cm}
+\section{Quantifying Error}
+We used these maps to first quantify error in cropland area estimates.
+We calculated error as the difference between the reference and cropland map at different scales of aggregation (1 to 100 km), to determine the extent of bias in each map and how it varies with scale.
+Next, we assessed how map error relates to cover patterns in agricultural  landscapes, by examining how map error over cropped pixels correlates with the actual amount of cropland.
+
+We undertook five further analyses to investigate how map error can impact assessments that are founded on landcover maps.
+These include first-order analyses, in which values for a variable of interest are mapped to  particular cover type(s), and second-order analyses, in which a process model draws on the cover types' values to calculate an output value.
+We created four datasets to represent second order analyses.
+The first was a series of maps of vegetated carbon stocks created following the methodology of Ruesch and Gibbs' \cite{ruesch_new_2008}.
+The second was cropland percentage maps, which, following Ramankutty et al \cite{ramankutty_farming_2008} were adjusted so that their total areas matched the reported provincial-level cropland area.
+Using these adjusted cropland percentage maps, we followed the methods of Monfreda et al \cite{monfreda_farming_2008} to disaggregate census-reported maize harvested area and yields.
+We then compared differences between total carbon stock estimates calculated from the reference map with those from the four cropland maps, and again examined how these differences changed as a function of aggregation scale.
+We made the same comparisons for total maize harvested area, average yield, and total production.
+
+For the second-order analyses, we examined how cropland cover errors influence 25 km resolution monthly evapotranspiration estimates produced using the Variable Infiltration Capacity \cite{liang_simple_1994}.
+For this example, we used the cropland maps to adjust the seasonally varying, landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
+In the second example, we examined how these errors can impact the parameterization of a computationally intensive, spatially explicit agent-based model of food security.
+In this case, we used the cropland maps to allocate farmland to model agents representing individual households in political districts, with each agent's initial holdings assigned as a function of the total cropland available in the district, and the proximity of cropland within a specified distance of agent's location.
+
+%\vspace{-0.5 cm}
+\section{Results}
+
+\subsection{Percent cropland estimates}
+
+\subsection{Error as a function of cropland density}
+
+\subsection{Potential bias in harvested areas, yield, and production estimates}
+
+\subsection{Potential bias in estimates of carbon stocks}
+
+\subsection{Potential bias in harvested areas, yield, and production estimates}
+
+\subsection{Impacts on evapotranspiration estimates}
+
+\subsection{Initialization errors in spatial agent-based models}
+
 %large gradient ($\sim \displaystyle{ \inlinefrac{1}{\delta}}$).

### 2015-08-27 Rmd/compare-landcover.Rmd
Into results in manuscript. Updated cropland bias maps, bias statistics plots. Added Rmd for miscellaneous supporting calculations.

--- 2015-06-02 Rmd/compare-landcover.Rmd
@@
     toc: yes
+    number_sections: true
 ---
 
-## Overview
+# Overview
 This section of the analysis calculates the differences between the GTI cropland percentages and those from the various landcover products.
-It draws from results generated by the [cropland pre-processing](https://github.com/PrincetonUniversity/SAcropland/blob/master/vignettes/sa-cropland-analysis.Rmd).
-(Note to self here: I want to have a transportable link between chapters here, rather than a direct link to the other RMD document in
-github).
+It draws from results generated by the [cropland pre-processing](cropland-preprocess.Rmd).
+(Note: html needed).
 
-## Analyses
-### Load in data 
-1. Set up paths
+# Analyses
+## Load in data 
+### Set up paths
 ```{r, eval = FALSE}
@@
 
-2. Load cropland grids 
-  + Derived from global landcover products (note: it turns out GLC-Share and geowiki are the same thing)
+### Load cropland grids
+
+Derived from global landcover products (note: it turns out GLC-Share and geowiki are the same thing, so GLC-Share dropped)
 
@@
 salc <- raster(full_path(p_data, "sa_ag_masked.tif")) * 100
-globsa_list_sum <- lapply(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
-modsa_list_sum <- lapply(full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
-glcsa <- raster(full_path(p_data, "glcsa_masked.tif"))
+ds <- full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)])
+globsa_list_sum <- lapply(ds, raster)
+fnames <- full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)])
+modsa_list_sum <- lapply(fnames, raster)
+# glcsa <- raster(full_path(p_data, "glcsa_masked.tif"))
 geow <- raster(full_path(p_data, "geowikisa_masked.tif"))
-lclist <- c(salc, globsa_list_sum, modsa_list_sum, glcsa, geow)  # into list
-names(lclist) <- c("sa30", "globmin", "globmu", "globmax", "modmin", "modmu", "modmax", "glcsh", "geow")
+lclist <- c(salc, globsa_list_sum, modsa_list_sum, geow)  # into list
+names(lclist) <- c("sa30", "globmin", "globmu", "globmax", "modmin", "modmu",
+                   "modmax", "geow")
 ```
 
-  + SA landcover data
+### SA landcover data
 ```{r, eval = FALSE}
 load(full_path(p_data, "MZshapes.Rdata"))  
-paths <- full_path(p_data, dir(p_data, paste("cover*.*sum_mask*", sep = "")))
+paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
 gti <- lapply(paths, raster)
@@
 ```
-[[back to top]][Overview]
+<a href="#top">Back to top</a>
 
-### Calculate differences at different levels of aggregation
+## Calculate differences at different levels of aggregation
 This steps draws on functions I wrote to calculate the actual and absolute differences between a given "master" grid and lists of other grids.
@@
 # Aggregate rasters
-#fact <- c(5, 10, 25, 50, 100, 200, 300)
-fact <- 25
+fact <- c(5, 10, 25, 50, 100)
+#fact <- 25
 lc_agg <- aggregate_rast_list(fact, lclist)   # landcover rasters 
@@
 # Calculate differences
+# actual
 dlist_act <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)),
-                            gti_agg, lc_agg, "a - b", silent = FALSE)  # actual
+                            gti_agg, lc_agg, "a - b", silent = FALSE)  
+# absolute
 dlist_abs <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)), 
-                            gti_agg, lc_agg, "abs(a - b)", silent = FALSE)  # absolute
+                            gti_agg, lc_agg, "abs(a - b)", silent = FALSE)  
 dlist_gti <- rast_list_math(list("g2007", names(gti_agg), "g2011"), 
@@
 ```
-[[back to top]][Overview]
-### Bias as function of cropcover
-+ First create NA mask at 1 km resolution, then weighting raster based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc)
+<a href="#top">Back to top</a>
+
+## Bias as function of cropcover
+### First create NA mask at 1 km resolution, then weighting raster based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc)
 ```{r, eval = FALSE}
@@
 binv <- seq(0, 100, 5)
-gti_bins <- lapply(gti_agg, function(x) lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE)))
+gti_bins <- lapply(gti_agg, function(x) {
+  lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE))
+})
 ```
 
-+ Calculate biases, pooling MODIS and GlobCover across all version
+### Calculate biases, pooling MODIS and GlobCover across all version
 Step 1. Reshape lists, so that all MODIS and GlobCover variants are pooled by scale
@@
 lev_vec <- paste("f", fact_all, sep = "") 
-nm_rt <- c("sa30", "glob", "mod", "glcsh", "geow")
+nm_rt <- c("sa30", "glob", "mod", "geow")
 
@@
 
-+ Then calculate biases across datasets
+### Then calculate biases across datasets
+
+Note: this uses an older and slower version of the code. Subsequent analyses use the faster version based on data.tables.  
 ```{r, eval = FALSE}
@@
 ```
-[[back to top]][Overview]
 
-+ Finally, differences between gti2007 and 2011
+<a href="#top">Back to top</a>
+
+## Differences between gti2007 and 2011
 ```{r, eval = FALSE}
@@
 ```
+<a href="#top">Back to top</a>
+
 Checks: compared number of non-NA cells coming out of each GTI version at each level of aggregation (fine); compared cell counts between previous committed version and this one (difference exists, but due to fact that gtimu is no longer considered; checked that previous versions of difference rasters and reshapes of them were referencing the correct GTI-landcover bias rasters (i.e.
@@
 Compared results from original roughed out version of approach for calculating bias statistics with results from new outputs list (using geowiki at 20 km aggregation as the comparison)
-
-## Next steps
-Plot bias results.  
-[[back to top]][Overview]

### 2015-08-27 Rmd/landcover-bias-plots.Rmd
Into results in manuscript. Updated cropland bias maps, bias statistics plots. Added Rmd for miscellaneous supporting calculations.

@@
 
-## Plotting bias figures
-### Biases in space
-Load in necessary datasets
+# Plotting bias figures
+# Spatial biases
+## Datasets
 
@@
 
+## Plots
 Plot maps of biases at different resolutions, for publication output (and display here).
@@
 
-Plot of selected rasters and levels 
+## Plot of selected rasters and levels 
 ```{r, eval = FALSE}
@@
 cols <- colorRampPalette(c("red", "grey80", "blue4"))
-brks <- c(-100.1, -75, -50, -25, -10, 0, 10, 25, 50, 75, 100.1)
-n_cols <- length(brks) - 1
+brks <- c(-100.1, -75, -50, -25, -10, -5, -1, 1, 5, 10, 25, 50, 75, 100.1)
+brklen <- length(brks) - 1
+cols <- cols(brklen)
+cols[c(1, length(cols))] <- c("darkred", "purple3")
 legtext <- "% Difference"
@@
 lev2 <- c("1 km", "25 km", "50 km", "100 km")
-pdf(full_path(p_fig, "bias_map.pdf"), height = 6, width = 7)
-par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
+pdf(full_path(p_fig, "bias_map.pdf"), height = 7, width = 7)
+par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(3.7, 1.5, 1.5, 0))
 for(i in 1:length(snms)) {
@@
     plot(sa.shp, lty = 0)
-    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols(n_cols), 
-         breaks = brks, legend = FALSE)
-  if(j == 1) mtext(mcap[i], side = 2, line = 1, cex = cx)
-  if(i == 1) mtext(lev2[j], side = 3, line = 0, cex = cx)
+    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols, breaks = brks,
+         legend = FALSE)
+  if(j == 1) mtext(mcap[i], side = 2, line = -0.25, cex = cx)
+  if(i == 1) mtext(lev2[j], side = 3, line = -0.5, cex = cx)
   }
 }
-flex_legend(ncuts = length(brks) - 1, legend.text = legtext, 
-            legend.vals = round(brks), 
-            longdims = c(0.2, 0.8), shortdims = c(0.06, 0.01), 
-            colvec = cols(length(brks) - 1), 
-            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
-            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0.5, -0.5)), 
-            cex.val = cx, textcol = lcol, bordercol = lcol)
+flex_legend(ncuts = brklen, legend.text = legtext, 
+            legend.vals = round(brks), longdims = c(0.2, 0.8), 
+            shortdims = c(0.05, 0.01), colvec = cols, srt = c(270, 0), 
+            horiz = TRUE, textside = "bottom", legend.pos = c(4, 5), 
+            leg.adj = list(c(0.2, 0.35), c(0.5, -0.5)), cex.val = cx, 
+            textcol = lcol, bordercol = lcol)
 dev.off()
@@
 
-[[back to top]][Plotting bias figures]
+<a href="#top">Back to top</a>
 
-### Bias in relation to field density
-1. Supplementary figures - bias plots at 1 km with boxplots.
+# Supplementary figures: bias in relation to field density
+## Bias plots at 1 km with boxplots.
 
@@
            "yl" = list(c(-100, 100), c(0, 100)), 
-           #"yl2" = list(c(-40, 40), c(0, 60)), "xl" = c(1, 20),
            "yl2" = c(-20, 60), "xl" = c(1, 20),
@@
            "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
-           #"yax2" = list(
-            # unlist(lapply(seq(-40, 40, 20), function(x) c(" ", x)))[-1], 
-            # unlist(lapply(seq(0, 60, 20), function(x) c(" ", x)))[-1]),
-           #"yaxp" = list(-100, 0), 
            "ycol" = c("grey70"))  # plot parameters
@@
 
-2. Bias with scale
+## Bias with scale
 ```{r, eval=FALSE}
-lev <- lev[!lev %in% c("f5", "f300")]
+#lev <- lev[!lev %in% c("f5", "f300")]
 levnms <- paste(gsub("f", "", lev), "km")
@@
 
-pdf(full_path(p_fig, "biases_1-200km.pdf"), width = 7.5, height = 7)
+pdf(full_path(p_fig, "biases_1-100km.pdf"), width = 7.5, height = 7)
 l <- layout(m3)
+# layout.show(l)
 # i <- 1; j <- 1
@@
 
+```
+
+```{r, eval=FALSE, echo=FALSE}
 DF <- do.call(rbind.data.frame, lapply(snms[4], function(x) {

### 2015-08-27 Rmd/misc.Rmd
Into results in manuscript. Updated cropland bias maps, bias statistics plots. Added Rmd for miscellaneous supporting calculations.

---
title: "Miscallaneous calculations"
author: "Lyndon Estes"
date: "August 24, 2015"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

Various small calculations made for the analysis.

```{r, eval=FALSE}
library(rgdal)
library(SAcropland)
library(rgeos)

p_dat <- full_path(full_path(proj_root("SAcropland"), "SAcropland"), 
                   "external/ext_data")

```

Area of sub-Saharan Africa and South Africa's share of that area
```{r, eval=FALSE}
eco <- readOGR(dsn = full_path(p_dat, "africa_ecofloristic_zones.sqlite"), 
               layer = "africa_ecofloristic_zones")
eco@proj4string <- CRS(projection(raster(nrow = 1, ncol = 1)))
af <- readOGR(dsn = full_path(p_dat, "africa_countries_alb.sqlite"), 
               layer = "africa_countries_alb")
af@data <- af@data[, c(1, 3:4)]
ecoalb <- spTransform(eco, af@proj4string)

# cut down countries
nms <- af$cntry_name[af$region == "Northern Africa"]
nms <- nms[nms != "Sudan"]  # keep Sudan
ssa <- af[!af$cntry_name %in% nms, ]
ssa <- ssa[-grep("Island|Sao|Verde|Mauritius|Helena|Comoros|Portugal|Seychell",
                 as.character(ssa@data$cntry_name)), ]
ssa <- gBuffer(ssa, width = 0)
# plot(ssa)

desert <- ecoalb[grep("desert", ecoalb$gez_term), ]
# plot(ecoalb[names(which.max(gArea(desert, byid = TRUE) / 10000 / 100)), ])
sahara <- as.numeric(names(which.max(gArea(desert, byid = TRUE))))

nosahara <- ecoalb[-sahara, ]
plot(nosahara, col = "red")

ovs <- which(!is.na(over(nosahara, ssa)))
ssaeco <- nosahara[ovs, ]
ssaeco <- ssaeco[-1, ]  # drop a couple Saharan mountains
ssaeco <- ssaeco[-1, ]
# plot(ssaeco[ssaeco@data$gez_term == "Tropical mountain system", ][1, ])
plot(ssa)
plot(ssaeco, col = "red", add = TRUE)

# calculate area of remainder
afarea <- gArea(ssaeco) / 10000
# plot(af[af$cntry_name == "South Africa", ])
saarea <- gArea(af[af$cntry_name == "South Africa", ]) / 10000
saarea / afarea

```

```{r, eval=FALSE}
a <- sample(1:100, size = 20)
b <- sample(1:80, size = 20)
mean(a - b)
sum(a) / sum(b)

```

### 2015-08-27 Rmd/yield-bias.Rmd
Into results in manuscript. Updated cropland bias maps, bias statistics plots. Added Rmd for miscellaneous supporting calculations.

@@
   if(x != "f1") { 
-    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun = modal)
+    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), 
+                        fun = modal)
   } else {

### 2015-08-27 paper/main.tex
Into results in manuscript. Updated cropland bias maps, bias statistics plots. Added Rmd for miscellaneous supporting calculations.

@@
 %\usepackage{pnastwoF}
-\usepackage{cite}
+%\usepackage{cite}
+\usepackage[numbers,round,compress]{natbib}
+\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
 \usepackage{lineno}
-\setlength\linenumbersep{4pt}
+\setlength\linenumbersep{2pt}
 
@@
 We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite{fry_completion_2009},  MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013,shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and is the current state of the art for agricultural landcover maps.
-We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates (hereafter simply ``cropland maps''), resulting in 4 maps to compare to the reference.
+We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates (hereafter simply the ``landcover maps''), resulting in 4 maps to compare to our reference cropland map (the ``reference map'').
 
@@
 We used these maps to first quantify error in cropland area estimates.
-We calculated error as the difference between the reference and cropland map at different scales of aggregation (1 to 100 km), to determine the extent of bias in each map and how it varies with scale.
-Next, we assessed how map error relates to cover patterns in agricultural  landscapes, by examining how map error over cropped pixels correlates with the actual amount of cropland.
+We calculated error as the difference between the reference and landcover maps at different scales of aggregation (1 to 100 km), in order to estimate bias and how it varies with scale.
+Next, we assessed how bias correlates with the amount of cropland cover in agricultural landscapes, to gain insight into how landscape patterns may affect error.
 
-We undertook five further analyses to investigate how map error can impact assessments that are founded on landcover maps.
+We undertook five further analyses to investigate how map error can impact assessments founded on landcover maps.
 These include first-order analyses, in which values for a variable of interest are mapped to particular cover type(s), and second-order analyses, in which a process model draws on the cover types' values to calculate an output value.
@@
 The first was a series of maps of vegetated carbon stocks created following the methodology of Ruesch and Gibbs' \cite{ruesch_new_2008}.
-The second was cropland percentage maps, which, following Ramankutty et al \cite{ramankutty_farming_2008} were adjusted so that their total areas matched the reported provincial-level cropland area.
-Using these adjusted cropland percentage maps, we followed the methods of Monfreda et al \cite{monfreda_farming_2008} to disaggregate census-reported maize harvested area and yields.
+The second was constrained cropland percentage maps, which, following Ramankutty et al \cite{ramankutty_farming_2008} were adjusted so that their total cropland areas matched provincial-level reported cropland totals.
+Using these adjusted cropland percentage maps, we disaggregated district-reported maize harvested area and yields \cite[following][]{monfreda_farming_2008} .
 We then compared differences between total carbon stock estimates calculated from the reference map with those from the four cropland maps, and again examined how these differences changed as a function of aggregation scale.
@@
 
-For the second-order analyses, we examined how cropland cover errors influence 25 km resolution monthly evapotranspiration estimates produced using the Variable Infiltration Capacity \cite{liang_simple_1994}.
+For the second-order analyses, we examined how cropland cover errors influence 25 km resolution monthly evapotranspiration estimates produced using the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model.
 For this example, we used the cropland maps to adjust the seasonally varying, landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
-In the second example, we examined how these errors can impact the parameterization of a computationally intensive, spatially explicit agent-based model of food security.
-In this case, we used the cropland maps to allocate farmland to model agents representing individual households in political districts, with each agent's initial holdings assigned as a function of the total cropland available in the district, and the proximity of cropland within a specified distance of agent's location.
+In the second example, we examined how these errors can impact the parameterization of an agent-based food security model \cite{chen_dependency_2013}.
+Spatially-explicit, agent-based models are frequently employed in land change science, and require an initialization step to assign landscape resources to model agents  \cite[e.g.][]{manson_agent-based_2007,evans_multi-scale_2004,kelley_relative_2011}.
+In this case, we used the cropland maps to allocate farmland to agents representing individual households in political districts, with the model assigning each household its initial cropland holdings using a function that considers total district cropland and how much cropland is near the agent's location.
 
-%\vspace{-0.5 cm}
+\vspace{-0.5 cm}
 \section{Results}
+\subsection{Bias and its correlates}
 
-\subsection{Percent cropland estimates}
+We created the 1 km reference and landcover maps, removing all croplands marked as communal or smallholder farmland in the reference vector maps (individual fields were not mapped with the same precision) or as plantation forestry (SI), and then aggregated each map to 5, 10, 25, 50, and 100 km resolutions.
+We then subtracted each landcover map from the reference map at each scale of aggregation (Fig.
+1).
+
+\vspace{-0.25 cm}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=.5\textwidth]{figures/bias_mapc.pdf}}
+\caption{Differences in percent cropland estimates between the reference map and each of the four landcover maps.
+Rows indicate the landcover map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation.
+White indicates areas with no data where communal farmlands or plantation forests were removed.}\label{afoto}
+\end{figure}
+
+The spatial patterns of bias are most pronounced in the MODIS and GlobCover maps, with both substantially underestimating the amount of cropland in most regions of the country, particularly in the central part of the country where the bulk of the country's cropland is found (blue areas in Fig.
+1).
+The mean of these differences, which is the landcover map's bias, are 21\% and 34\% at 1 km for MODIS and GlobCover, respectively (Fig.
+S1), meaning that each map underestimates cropland by that amount at this resolution.
+MODIS bias drops to 8\% at 50 km of aggregation, whereas GlobCover bias is still 24\% at 100 km (Fig.
+S1).
+
+The SA-LC and GeoWiki datasets show much spatial variability and overall bias.
 

### 2015-08-27 Rmd/landcover-bias-plots.Rmd
Results up through landscape characteristics and bias.

@@
 
-# Supplementary figures: bias in relation to field density
-## Bias plots at 1 km with boxplots.
+# Bias in relation to field density
+
+For supplementary data.
+
+## Bias plots at 1 km with boxplots
 
@@
 ```
+
+<a href="#top">Back to top</a>
 
@@
 ```
+<a href="#top">Back to top</a>
+
+# Total bias
+
+Calculate how much cropland estimates differ between datasets
+```{r, eval = FALSE}
+paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
+gti <- lapply(paths, raster)
+names(gti) <- c("g2007", "g2011")
+snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
+
+# recreate cropland percentages
+lc2007 <- lapply(snms, function(x) gti$g2007 - dlist_act[[x]]$f1$g2007)
+lc2011 <- lapply(snms, function(x) gti$g2011 - dlist_act[[x]]$f1$g2011)
+# plot(lc2007[[2]] - lc2011[[2]])  # same thing, as it should be
+
+# calculate km2 cropland in LC datasets
+lckm2 <- sapply(lc2011, function(x) cellStats(x / 100, sum))
+refkm2 <- sapply(gti, function(x) cellStats(x / 100, sum))
+refkm2[2] / refkm2[1]  # 3.6 percent more in 2011 than 2007
+
+# bias in total cropland relative to GTI
+stats2007 <- round((refkm2[2] - lckm2) / refkm2[2] * 100, 1)
+stats2011 <- round((refkm2[1] - lckm2) / refkm2[1] * 100, 1)
+
+# N km^2 in country
+#sapply(lc2007, function(x) cellStats(!is.na(x), sum))
+# sapply(lc2011, function(x) cellStats(!is.na(x) , sum))
+# sapply(gti, function(x) cellStats(!is.na(x) , sum))
+totkm2 <- cellStats(!is.na(gti$g2011) , sum)
+save(stats2007, stats2011, totkm2, refkm2, 
+     file = full_path(p_data, "total-cropland.rda"))
+```
+<a href="#top">Back to top</a>
+
+Total square kilometers of South Africa being evaluated
+```{r, echo = FALSE}
+library(lmisc)
+p_root <- proj_root("SAcropland")
+p_data <- full_path(p_root, "SAcropland/external/ext_data/")
+load(full_path(p_data, "total-cropland.rda"))
+totkm2
+```
+
+Total square kilometers of cropland according to 2007 and 2011 GTI data
+```{r, echo = FALSE}
+refkm2
+```
+
+Percent bias in total cropland area estimates relative to 2007 GTI data
+```{r, echo = FALSE}
+stats2007
+```
+
+And relative to 2011 GTI data
+```{r, echo = FALSE}
+stats2011
+```
+
+<a href="#top">Back to top</a>
 

### 2015-08-27 Rmd/magisterial-district-bias.Rmd
Initial versions of carbon bias estimation complete. Added utility functions used in various analyses to library. Minor edit to magisterial-district-bias.Rmd
Results up through landscape characteristics and bias.

@@
     toc: yes 
+    number_sections: true
 ---
 
-## Bias in relation to typical crop cover
+# Bias as a function of typical crop cover
 
-### Data
-```{r, eval = FALSE}
+## Data
+```{r, message=FALSE, results='hold'}
 library(raster)
@@
 load(full_path(p_data, "d_grid_abs.rda"))  # actual diffence grids
-paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask*")))
+paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
 gti <- raster(paths)  # gti 2011
@@
 md <- md[md$ID_2 != 313, ]
+```
+
+Mean area of magisterial district.
+```{r}
 round(mean(rgeos::gArea(md, byid = TRUE) / 1000000))
@@
 
-### Analyses
+<a href="#top">Back to top</a>
 
-1. Extract the cropland percentages for magisterial district, as well as the absolute biases. 
+# Analyses
+
+## Extract data for magisterial districts
+
+Both cropland percentages and absolute biases. 
 ```{r, eval=FALSE}
@@
 med0$w <- nonzero  # weight - N contributing pixels (non-zeros)
+
 ```
+<a href="#top">Back to top</a>
 
-2. Run regressions and check for spatial autocorrelation
+## Run regressions and check for spatial autocorrelation
 ```{r, eval=FALSE}
@@
 ```
+<a href="#top">Back to top</a>
 
-Plot the fitted LM models
+### Plot the fitted LM models
 ```{r, eval=FALSE}
@@
 ```
+<a href="#top">Back to top</a>
 
-Check for spatial autocorrelation in residuals
+### Check for spatial autocorrelation in residuals
 ```{r, eval=FALSE}
@@
 ```
+<a href="#top">Back to top</a>
 
@@
 
-3. A generalized least squares (GLS) with spatial autocovariance terms.
+### A generalized least squares (GLS) with spatial autocovariance terms.
 
@@
 
-3. Use a GAM instead
+### Use a GAM instead
 
@@
 
-### Plot final results
+## Plot final results
 
-1. Helper functions
+Helper functions
 ```{r, eval = FALSE}
@@
 
-makeTransparent<-function(someColor, alpha = 100) {
-  newColor <- col2rgb(someColor)
-  apply(newColor, 2, function(x) {
-    rgb(red = x[1], green = x[2], blue = x[3],alpha = alpha, maxColorValue=255)
-  })
-}
 ```
+<a href="#top">Back to top</a>
 
-2. GAM fits
+### GAM fits
 ```{r, eval = FALSE}
@@
 ```
+<a href="#top">Back to top</a>
 
-3. GAM plots
+### GAM plots
 ```{r, eval = FALSE}
-plot(xl, yl2, pch = "", ylab = "% bias)", xlab = "% cropland", xaxs = "i",
-     yaxs = "i", yaxt = "n")
-
-round(log(seq(1, 90, 10)), 2)
-exp(seq(0, 4.5, 0.5))
+# plot(xl, yl2, pch = "", ylab = "% bias)", xlab = "% cropland", xaxs = "i",
+#      yaxs = "i", yaxt = "n")
+# round(log(seq(1, 90, 10)), 2)
+# exp(seq(0, 4.5, 0.5))
 
@@
 ```
+<a href="#top">Back to top</a>
 
-Final plot of magisterial districts, for supplementals
+### Final plot of magisterial districts, for supplementals
 ```{r, eval=FALSE}
@@
 dev.off()
+
 ```
+<a href="#top">Back to top</a>

### 2015-08-30 Rmd/agent-allocation-bias.Rmd
Main paper: Half of results done, bullet points on rest through Discussion (but not yet organized).  Updated cropland bias, carbon bias, yield bias, agent-bias, as worked through results. Earlier results reconfirmed.

@@
 cexs <- seq(1, 2, 1 / 3)
-pdf(full_path(p_fig, "agent-bias.pdf"), height = 3.5, width = 7)
+pdf(full_path(p_fig, "agent-bias.pdf"), height = 5, width = 3)
 cx <- c(0.7, 0.8, 0.6)
 # png(full_path(p_fig, "agent-bias.png"), height = 350, width = 700)
-par(mfrow = c(1, 2), mar = c(3, 2.5, 3, 1), mgp = c(1.5, 0.25, 0), tcl = -0.2, 
-    oma = c(0, 0, 0, 4))
+par(mfrow = c(2, 1), mar = c(0.5, 2.5, 0.5, 1), mgp = c(1, 0.05, 0), 
+    tcl = -0.2, oma = c(3.5, 0, 0, 0))
 scols <- c("red", "orange3", "green4", "blue")
@@
 par(xpd = NA)
-xst <- seq(0.86, 0.96, 0.07 / 3)
-xy <- cbind(grconvertX(xst, from = "ndc"), 
-            grconvertY(rep(0.6, 5), from = "ndc"))
+
+yst <- seq(0.03, 0.13, 0.07 / 3)
+xy <- cbind(grconvertX(rep(0.5, 5), from = "ndc"), 
+            grconvertY(yst, from = "ndc"))
 for(i in 2:length(xst)) {
-  if(i == 2) {
-    leg <- 1:4
+  if(i == 5) {
+    leg <- c("District", 1:4)
   } else {
@@
   }
-  legend(x = xy[i, 1], y = xy[i, 2], legend = leg, pch = 20, cex = cx[3], 
-         adj = c(4.5, 0.75),  
+  legend(x = xy[i, 1], y = xy[i, 2], legend = rep("",4), #legend = leg, 
+         pch = 20, cex = cx[3], 
+         horiz = TRUE, adj = c(4, -5.75),  
          col = scols[i - 1], pt.cex = cexs, bty = "n")#cex = cexs,)
 }
-xy <- cbind(grconvertX(xst, from = "ndc"), 
-            grconvertY(rep(0.58, 5), from = "ndc"))
-text(x = xy[, 1], y = xy[, 2], labels = c("District", mcap), srt = 90, #pos = 2, 
-     adj = c(0, 1.5), col = c("black", scols), cex = cx[1])
+yst <- seq(0.005, 0.105, 0.07 / 3)
+xy <- cbind(grconvertX(rep(0.55, 5), from = "ndc"), 
+            grconvertY(yst, from = "ndc"))
+text(x = xy[2:5, 1], y = xy[2:5, 2], labels = c(mcap),# "District"), 
+     #srt = 90, #pos = 2, 
+     #adj = c(-1, 1), 
+     pos = 2, col = c(scols), cex = cx[1])
 dev.off()

### 2015-08-30 Rmd/carbon-bias.Rmd
Major update, mostly based on undertaking and completing full version of yield-bias.Rmd, but also some changes to library functions, including new stats estimating functions and removal of data.table/raster functions to to new package, which is used here.
Main paper: Half of results done, bullet points on rest through Discussion (but not yet organized).  Updated cropland bias, carbon bias, yield bias, agent-bias, as worked through results. Earlier results reconfirmed.

@@
 
-## Bias in carbon stock estimates
+# Bias in carbon stock estimates
 Based on the Ruesch & Gibbs (2008) approach for estimating carbon density.
 
-## Data
-
-### Paths
+# Data
 ```{r, eval=FALSE}
@@
 library(RColorBrewer)
+library(xtable)
 
@@
 
-### Carbon values
+## Carbon values
 
@@
 
-## Analysis
-### Prepare raster data
+# Analysis
+## Prepare raster data
 ```{r, eval=FALSE}
@@
 # need NA mask for weighting of aggregated value
-namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
-namask2 <- !is.na(namask)  # set NAs to zero
+sumna <- function(x) sum(x, na.rm = FALSE)
+namask <- calc(stack(gti, stack(lc_list)), sumna)
+namask[namask > 0] <- 1
+namask2 <- !is.na(namask) 
+# namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
+# namask2 <- !is.na(namask)  # set NAs to zero
 area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
@@
 
-### Create cropland carbon estimates
+## Create cropland carbon estimates
 ```{r, eval=FALSE}
 # Apply carbon estimates
-cc <- ctab2[ctab2$class == 16, 2]
+# Function for pixel-wise carbon density from crop & non-crop fractions
 carbon <- function(fcrop, cropC, noncropC) {
@@
   s
-})
+})  # gti
 
@@
   lc
-})
+})  # landcover datasets
 
@@
               sum)
-}) # all zeroes
+}) # should be all zeroes
 
@@
 
-### Difference the carbon datasets
+## Difference the carbon datasets
 ```{r, eval=FALSE}
@@
 # # disggregate selected rasters at selected levels for plotting
-namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
+# namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
 lev <- names(c_pct_diff_mu) 
@@
 
-## Outputs
-### Carbon bias maps
+# Outputs
+## Carbon bias maps
 ```{r, eval=FALSE}
@@
 
-### Carbon bias statistics
-#### Calculate weighted bias statistics
+## Carbon bias statistics
+### Calculate how much country-lvel carbon estimates differ between datasets
+```{r, eval = FALSE}
+lev_vec <- names(c_pct_diff)
+# created mask for non-cropland areas, unioning GTI and each LC, filtering out
+# areas of no-cropland (<1/2% total cover)
+lc_union <- lapply(lev_vec, function(x) {
+  lcb <- lapply(snms, function(y) {
+    gti_gt0 <- Which(round(gti_agg[[x]]$gti * 100) > 0)
+    lc_gt0 <- Which(round(lc_agg[[x]][[y]] * 100) > 0)
+    all_gt0 <- gti_gt0 + lc_gt0
+    all_gt0[all_gt0 > 0] <- 1
+    all_gt0
+  })
+  named_out(lcb, snms)
+})
+names(lc_union) <- lev_vec
+# plot(lc_union$f50$globmu)
+# plot(gti_agg$f50$gti)
+
+# calculate for whole country
+tareas <- lapply(area_wgts, function(x) x[[1]] * 100)
+gti_ctot <- sapply(1:length(gti_c), function(x) {
+  cellStats(gti_c[[x]] * tareas[[x]], sum)
+})  # gti total carbon stock
+lc_c2 <- lapply(snms, function(x) {
+  sapply(names(lc_c), function(y) lc_c[[y]][[x]])
+})  # reshape lc_c2 -> landcover in outer list
+names(lc_c2) <- snms
+lc_ctot <- lapply(lc_c2, function(x) {
+  sapply(1:length(x), function(y) {
+    cellStats(x[[y]] * tareas[[y]], sum)
+  })
+})  # landcover total carbon stocks
+totc_cntry <- lapply(lc_ctot, function(x) {
+  stats <- (gti_ctot - x) / gti_ctot * 100
+  colnames(stats) <- names(lc_c2$sa30)
+  rownames(stats) <- names(lc_c2$sa30$f1)
+  stats
+})  # country-level percent differences: gti versus landcover
+names(totc_cntry) <- snms
+
+# for cropped areas only - note here we are masking on cropland fraction only,
+# not on union of gti and each landcover map, which is needed below for mean 
+# bias estimates
+tot_area <- sum(freq(namask)[1:2, ])
+crop_areas <- lapply(lc_union$f1, freq)
+sapply(crop_areas, function(x) x[2, 2] / tot_area)  # 29, 31, 33, 53 (Glob)%
+
+lc_ag2 <- lapply(snms, function(x) {
+  sapply(names(lc_agg), function(y) lc_agg[[y]][[x]])
+})  # reshape lc_ag2 -> landcover fractions in outer list
+gti_ctot2 <- sapply(1:length(gti_c), function(x) {
+  msk <- gti_agg[[x]]$gti > 0.005  # farmland > 0.05% mask
+  msked <- mask(gti_c[[x]] * tareas[[x]], msk, maskvalue = 0)
+  cellStats(msked, sum)
+})  # gti 
+lc_ctot2 <- lapply(1:length(lc_c2), function(x) {
+  xx <- lc_c2[[x]]  # recycle reshaped lc_c2 list
+  jj <- lc_ag2[[x]]  # recycle reshaped lc_c2 list
+  sapply(1:length(xx), function(y) {
+    msk <- jj[[y]] > 0.005  # farmland > 0.05% mask
+    msked <- mask(xx[[y]] * tareas[[y]], msk, maskvalue = 0)
+    cellStats(msked, sum)
+  })
+})  # landcover carbon estimates
+
+# but discrepancy will only be relevant at 1 km scale, because the carbon total 
+# keeps increasing when cropland areas are the only ones being considered
+totc_crop <- lapply(lc_ctot2, function(x) {
+  stats <- (gti_ctot2 - x) / gti_ctot2 * 100
+  colnames(stats) <- names(lc_c2$sa30)
+  rownames(stats) <- names(lc_c2$sa30$f1)
+  stats
+})
+names(totc_crop) <- snms
+
+# Combine tables for output to supplementals, 1 km % differences for country
+# and agricultural levels
+totc_out <- rbind(t(sapply(totc_cntry, function(x) x[, 1])),
+                  t(sapply(totc_crop, function(x) x[, 1])))
+pnms <- c("Forest", "Secondary", "Shrubland", "Grassland", "Sparse")
+knms <- rep(mcap, 2)
+totc_out <- cbind.data.frame(knms, round(unname(totc_out), 2))
+colnames(totc_out) <- c("map", pnms)
+totc_out <- cbind(region = c(rep("Country", 4), rep("Agricultural", 4)), 
+                  totc_out)
+caption <- paste("Percent differences in total carbon stock estimates", 
+                 "calculated from the reference maps and from each of the four",
+                 "cropland maps. Differences are evaluated for total carbon",
+                 "estimates either at the country scale or over just the",
+                 "agricultural regions (cropland $>$0.05\\%), using",
+                 "the carbon densities of 5 different cover types to provide", 
+                 "the values for the non-agricultural portions of each pixel",  
+                 "(cover types indicated by column names).")
+totc_xtab <- xtable(totc_out, caption = caption, digits = 1)
+print(totc_xtab, type = "latex", caption.placement = "top",
+      file = "paper/figures/totC-bias.tex", include.rownames = FALSE)
+
+```
+
+### Calculate weighted bias statistics
 ```{r, eval=FALSE}
@@
 (areas * 0.005)  # 1/2 to 5000 ha
-
-lev_vec <- names(c_pct_diff)
-# created mask for non-cropland areas, unioning GTI and each LC, filtering out
-# areas of no-cropland (<1/2% total cover)
-lc_union <- lapply(lev_vec, function(x) {
-  lcb <- lapply(snms, function(y) {
-    gti_gt0 <- Which(round(gti_agg[[x]]$gti * 100) > 0)
-    lc_gt0 <- Which(round(lc_agg[[x]][[y]] * 100) > 0)
-    all_gt0 <- gti_gt0 + lc_gt0
-    all_gt0[all_gt0 > 0] <- 1
-    all_gt0
-  })
-  named_out(lcb, snms)
-})
-names(lc_union) <- lev_vec
-# plot(lc_union$f50$globmu)
-# plot(gti_agg$f50$gti)
 
@@
 
-extract_stat <- function(namevec, stats, i1, i2) {
-  estats <- do.call(rbindlist(lapply(namevec, function(i) {
-    data.table(sapply(stats, function(x) {
-      unlist(sapply(x, function(y) {
+namevec <- LC
+stats <- cb_stats
+i1 <- "all"
+i2 <- "m"
+
+extract_stat0 <- function(namevec, stats, i1, i2) {
+  estats <- do.call(rbind, lapply(namevec, function(i) {
+      dat <- sapply(stats, function(x) {
+      vals <- unlist(sapply(x, function(y) {
         v <- y[bin == i1 & bvals == i2, i, with = FALSE]
@@
       }))
-    }))
+    })#)
   }))
@@
                data.table(round(estats, 2)))
-  setnames(out, old = names(mu), new = )
+  #setnames(out, old = names(mu), new = "")
   return(out)
@@
 
-#### Plot means against scale, supplemental version 
+### Plot means against scale, supplemental version 
 ```{r, eval=FALSE}
-a <- extract_stat(LC, cb_stats, "all", "m")  # mean abs crop only
-b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+a <- extract_stat0(LC, cb_stats, "all", "m")  # mean cropland only 
+b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
 setcolorder(a, c(2:1, 3:ncol(a)))
@@
 
-a <- extract_stat(LC, cb_stats, "all", "ma")  # mean abs crop only
-b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+a <- extract_stat0(LC, cb_stats, "all", "ma")  # mean abs cropland only
+b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
 setcolorder(b, c(2:1, 3:ncol(b)))
@@
 
-a <- extract_stat(LC, cb_stats, "all", "m0")  # mean abs crop only
-b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+a <- extract_stat0(LC, cb_stats, "all", "m0")  # mean whole country
+b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
 setcolorder(b, c(2:1, 3:ncol(b)))
@@
 
-a <- extract_stat(LC, cb_stats, "all", "ma0")  # mean abs crop only
-b <- cbind(V2 = "all", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+a <- extract_stat0(LC, cb_stats, "all", "ma0")  # mean abs whole country
+b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
 setcolorder(b, c(2:1, 3:ncol(b)))
@@
 # retained
+LC2 <- c(LC, "all")
 pdf(full_path(p_fig, "carbon_veg_scale.pdf"), height = 7, width = 5)
@@
     oma = c(6, 4, 1, 0))
-for(f in 1:length(LC)) {
+for(f in 1:length(LC2)) {
   plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
@@
   for(i in 1:4) {
-    lines(1:6, mu[V1 == snms[i] & V2 == LC[f], lev_vec, with = FALSE], 
+    lines(1:6, mu[V1 == snms[i] & V2 == LC2[f], lev_vec, with = FALSE], 
           col = cols[i], lty = 2, lwd = lw)
@@
   for(i in 1:4) {
-    lines(1:6, mua[V1 == snms[i] & V2 == LC[f], lev_vec, with = FALSE], 
+    lines(1:6, mua[V1 == snms[i] & V2 == LC2[f], lev_vec, with = FALSE], 
           col = cols[i], lty = 1, lwd = lw)
@@
 }
-plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
-     ylab = "", xlab = "", 
-     panel.first = grid(NA, length(yax), lwd = 1, lty = 1))
-for(i in 1:4) {
- lines(1:6, all_mu[V1 == snms[i], lev_vec, with = FALSE], col = cols[i], 
-       lty = 2, lwd = lw)
-}
-for(i in 1:4) {
-  lines(1:6, all_mua[V1 == snms[i], lev_vec, with = FALSE], col = cols[i],
-        lty = 1, lwd = lw)
-}
+# plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+#      ylab = "", xlab = "", 
+#      panel.first = grid(NA, length(yax), lwd = 1, lty = 1))
+# for(i in 1:4) {
+#  lines(1:6, mu[V1 == snms[i], lev_vec, with = FALSE], col = cols[i], 
+#        lty = 2, lwd = lw)
+# }
+# for(i in 1:4) {
+#   lines(1:6, mua[V1 == snms[i], lev_vec, with = FALSE], col = cols[i],
+#         lty = 1, lwd = lw)
+# }
 mtext(side = 3, text = "Vegetation mean", line = 0)
@@
 
-#### Main text version of mean carbon bias against scale
+### Main text version of mean carbon bias against scale
 ```{r, eval=FALSE}
@@
 #Ccols <- brewer.pal(n = 7, name = "Greys")
-LC2 <- c(LC[-grep("second|grass", LC)], "all")
+LC2 <- c(LC[-grep("second|grass", LC)], "All")
 # xax <- seq(0, 7, 6 / (length(yax) - 2)); length(xax)
@@
 pchs <- c("*", "+", "o", "-")
-cx <- c(1, 0.7, 0.6, 1)
+# cx <- c(1, 0.7, 0.6, 1)
+cx <- c(2, 1.25, 1, 1)
 g1 <- "grey90"
@@
 
-pdf(full_path(p_fig, "carbon_veg_scale2.pdf"), height = 7, width = 7)
+pdf(full_path(p_fig, "carbon_veg_scale3.pdf"), height = 7, width = 7)
 par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
@@
 
-#### Supplementary mean bias tables
+### Supplementary mean bias tables
 ```{r, eval=FALSE}
 # Statistics from non-croplands areas included in means, for supplementals
-nmrepl <- cbind(snms, lcnms)
-mu0[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mu0a[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mu0[, V1 := nmrepl[match(V1, nmrepl), 2]]
-mu0a[, V1 := nmrepl[match(V1, nmrepl), 2]]
-mu0s <- rbind(cbind(stat = "actual", mu0), cbind(stat = "absolute", mu0a))
-mu0s_xtab <- xtable::xtable(mu0s)
-print(mu0s_xtab, type = "latex", file = "paper/figures/C-bias-mean-all-SA.tex")
+# mu0[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+# mu0a[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mu0s <- copy(rbind(cbind(Bias = "Actual", mu0), 
+                   cbind(Bias = "Absolute", mu0a)))
+for(i in 1:length(LC)) mu0s[V2 == LC[i], V2 := pnms[i]]
+for(i in 1:length(snms)) mu0s[V1 == snms[i], V1 := mcap[i]]
+setnames(mu0s, c("V1", "V2", lev_vec), 
+         c("Map", "Cover", paste(c(1, fact), "km")))
+caption <- paste("Mean actual and absolute biases for each of the cropland", 
+                 "maps across aggregation scales and each possible landcover", 
+                 "type sharing the pixel with cropland. Means were calculated",
+                 "across the entire country.")
+mu0s_xtab <- xtable(mu0s, digits = 1, caption = caption)
+print(mu0s_xtab, type = "latex", file = "paper/figures/C-bias-mean-all-SA.tex", 
+      tabular.environment = "longtable", floating = FALSE, 
+      caption.placement = "top", include.rownames = FALSE)
 
 # Statistics from non-croplands areas included in means, for supplementals
-nmrepl <- cbind(snms, lcnms)
-mu[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mua[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mu[, V1 := nmrepl[match(V1, nmrepl), 2]]
-mua[, V1 := nmrepl[match(V1, nmrepl), 2]]
-mus <- rbind(cbind(stat = "actual", mu), cbind(stat = "absolute", mua))
-mus_xtab <- xtable::xtable(mus)
-print(mu0s_xtab, type = "latex", 
-      file = "paper/figures/C-bias-mean-cropland-SA.tex")
+# mu[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+# mua[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+mus <- copy(rbind(cbind(Bias = "Actual", mu), 
+                  cbind(Bias = "Absolute", mua)))
+for(i in 1:length(LC)) mus[V2 == LC[i], V2 := pnms[i]]
+for(i in 1:length(snms)) mus[V1 == snms[i], V1 := mcap[i]]
+setnames(mus, c("V1", "V2", lev_vec), 
+         c("Map", "Cover", paste(c(1, fact), "km")))
+caption <- paste("Mean actual and absolute biases for each of the cropland", 
+                 "maps across aggregation scales and each possible landcover", 
+                 "type sharing the pixel with cropland. Means were calculated",
+                 "across the union of agricultural areas", 
+                 "(cropland $>$0.05\\%) identified by the reference map and", 
+                 "the cropland map to which it was being compared")
+mus_xtab <- xtable::xtable(mus, digits = 1, caption = caption)
+print(mus_xtab, type = "latex", 
+      file = "paper/figures/C-bias-mean-cropland-SA.tex", 
+      tabular.environment = "longtable", floating = FALSE, 
+      caption.placement = "top", include.rownames = FALSE)
 ```

### 2015-08-30 Rmd/yield-bias.Rmd
Main paper: Half of results done, bullet points on rest through Discussion (but not yet organized).  Updated cropland bias, carbon bias, yield bias, agent-bias, as worked through results. Earlier results reconfirmed.

@@
 
-# redo NA mask because of some slighting shifts
+# redo NA mask because of some slight shifts
 namask <- calc(stack(lapply(crop_ayr, function(x) x$fa)), sumna)
@@
 cyld_agg <- c(cyldf1, cyldr)
-
-plot(cyld_agg$f10$gti)
+# plot(cyld_agg$f10$gti)
 
@@
     ck <- crop_ay[[j]][, sum(fa)] * 100
-    d <- round((ck - cellStats(tareas[[i]] * carea_agg[[i]][[j]], sum)) / ck, 2)
+    d <- round((ck - cellStats(tareas[[i]] * carea_agg[[i]][[j]], sum)) / ck,2)
     print(d)
@@
 }
+
 # Production differences
@@
 # type 2 (non-zero)
-list2 <- lapply(cyld_agg3, function(x) x[-1]) # reuse cyld_agg for a, b/c correct
+list2 <- lapply(cyld_agg3, function(x) x[-1]) # cyld_agg for a, b/c correct
 y_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 2
 
-# convert them to percent differences relative to the mean pixel-wise production
+# convert them to percent differences relative to mean pixel-wise production
+# here only for map plotting.
 # plot(y_diff2$sa30$f25$gti)
@@
 
-# selection variables
+# Spatial bias stats calculation on yield and production differences, not on
+# percentages
 a <- bias_stats_list(bins, awgts, lcu, p_diff1, wm, "mu", "bias", TRUE)
@@
 p1mu <- extract_stat(lev, snms, "all", "mu", "bias", pd1_st)  # identical to p3
-p1ma <- extract_stat(lev, snms, "all", "mua", "bias", pd1_st)  # identical to p3
-p1sm <- extract_stat(lev, snms, "all", "sum", "bias", pd1_st)  # identical to p3
+p1ma <- extract_stat(lev, snms, "all", "mua", "bias", pd1_st)  # ident to p3
+p1sm <- extract_stat(lev, snms, "all", "sum", "bias", pd1_st)  # ident to p3
 p2mu <- extract_stat(lev, snms, "all", "mu", "bias", pd2_st)
@@
 
-pdf(full_path(p_fig, "yield_bias_scale2.pdf"), height = 7, width = 7)
+pdf(full_path(p_fig, "yield_bias_scale.pdf"), height = 7, width = 7)
 par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
@@
 for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+# grid(NA, NULL, lwd = 1, lty = 1)
+abline(h = yax, v = NULL, col = "grey80", lty = 1)
 polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
-grid(NA, length(yax), lwd = 1, lty = 1)
 lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
@@
       mm <- malmul[[k]]
+      # fetch bias stats, convert to percentage relative to mean
+      # e.g. GTI mean yield
       v <- sapply(1:length(mm), function(j) {
-        round(get(mm[j])[ol == lv & il == nm,  bias] / get(pyl[j])[lv] * 100, 1)
+        round(get(mm[j])[ol == lv & il == nm,  bias] / get(pyl[j])[lv] * 100,1)
       })
@@
        cex = 0.8)
-text(x = 12.7, y = -42, labels = "Absolute", srt = 45, adj = c(0, 0), cex = 0.8)
+text(x = 12.7, y = -42, labels = "Absolute", srt = 45, adj = c(0, 0), cex= 0.8)
 text(x = 12.2, y = -42, labels = "Actual", srt = 45, adj = c(0, 0), cex = 0.8)
-legend(x = 12.1, y = 65, legend = c("Prod I", "Prod II", "Yield 1", "Yield 2"), 
-       pch = pchs, pt.cex = c(2.5, 1.5, 1.5, 1.5), bty = "n", cex = 0.8)
+legend(x = 12.1, y = 65, legend = c("Prod I", "Prod II", "Yield 1", "Yield 2"),        pch = pchs, pt.cex = c(2.5, 1.5, 1.5, 1.5), bty = "n", cex = 0.8)
 dev.off()
+
+lapply(1:length(lev), function(ii) {
+  lv <- lev[ii]
+  lapply(1:length(snms), function(i) {
+    nm <- snms[i]
+    lapply(1:length(malmul), function(k) {
+      mm <- malmul[[k]]
+      v <- sapply(1:length(mm), function(j) {
+        round(get(mm[j])[ol == lv & il == nm,  bias] / get(pyl[j])[lv] * 100,1)
+      })
+    })
+  })
+})
+
 ```

### 2015-08-30 paper/figures/et-bias.tex
Main paper: Half of results done, bullet points on rest through Discussion (but not yet organized).  Updated cropland bias, carbon bias, yield bias, agent-bias, as worked through results. Earlier results reconfirmed.

% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Wed Aug 19 16:18:01 2015
\begin{table}[ht]
\centering
\begin{tabular}{rllrr}
  \hline
 & ETvar & dataset & bias & absbias \\ 
  \hline
1 & gs & geow & 0.2 & 0.8 \\ 
  2 & gspk & geow & 0.2 & 0.8 \\ 
  3 & mu & geow & 0.3 & 0.7 \\ 
  4 & mumn & geow & 0.3 & 0.6 \\ 
  5 & mumx & geow & 0.3 & 0.8 \\ 
  6 & gs & globmu & 0.1 & 1.2 \\ 
  7 & gspk & globmu & 0.2 & 1.4 \\ 
  8 & mu & globmu & -0.1 & 1.0 \\ 
  9 & mumn & globmu & -0.2 & 0.9 \\ 
  10 & mumx & globmu & 0.2 & 1.2 \\ 
  11 & gs & modmu & -0.5 & 0.9 \\ 
  12 & gspk & modmu & -0.4 & 1.0 \\ 
  13 & mu & modmu & -0.6 & 0.8 \\ 
  14 & mumn & modmu & -0.6 & 0.7 \\ 
  15 & mumx & modmu & -0.5 & 0.8 \\ 
  16 & gs & sa30 & 0.3 & 0.7 \\ 
  17 & gspk & sa30 & 0.2 & 0.8 \\ 
  18 & mu & sa30 & 0.5 & 0.6 \\ 
  19 & mumn & sa30 & 0.4 & 0.5 \\ 
  20 & mumx & sa30 & 0.4 & 0.6 \\ 
   \hline
\end{tabular}
\end{table}

### 2015-08-30 paper/main.tex
Main paper: Half of results done, bullet points on rest through Discussion (but not yet organized).  Updated cropland bias, carbon bias, yield bias, agent-bias, as worked through results. Earlier results reconfirmed.

@@
 \section{Results}
-\subsection{Bias and its correlates}
+\subsection{Landcover bias}
 
-We created the 1 km reference and landcover maps, removing all croplands marked as communal or smallholder farmland in the reference vector maps (individual fields were not mapped with the same precision) or as plantation forestry (SI), and then aggregated each map to 5, 10, 25, 50, and 100 km resolutions.
-We then subtracted each landcover map from the reference map at each scale of aggregation (Fig.
+We created the 1 km reference and landcover maps, removing all croplands marked as communal or smallholder farmland in the reference vector maps (individual fields were not mapped with the same precision), as horticulture, or plantation forestry (SI), and calculated the total cropland extent in the remaining area (1,081,000 km$^2$, or 90\% of South Africa).
+The 2011 reference map showed a cropland area of 104,310 km$^2$, which the SA-LC and GeoWiki maps overestimated by 26 and 5.8\%, respectively, and GlobCover and MODIS underestimated by 21 and 26.1\%.
+
+We then aggregated each map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from the reference map at each scale of aggregation to asses the spatial patterns of bias (Fig.
 1).
@@
 \begin{figure}[ht]
-\centerline{\includegraphics[width=.5\textwidth]{figures/bias_mapc.pdf}}
+\centerline{\includegraphics[width=.5\textwidth]{figures/bias_map.pdf}}
 \caption{Differences in percent cropland estimates between the reference map and each of the four landcover maps.
@@
 
-The spatial patterns of bias are most pronounced in the MODIS and GlobCover maps, with both substantially underestimating the amount of cropland in most regions of the country, particularly in the central part of the country where the bulk of the country's cropland is found (blue areas in Fig.
+The most pronounced extreme biases are in the MODIS and GlobCover maps, with both underestimating cropland extent by 10-75\% in the center of the country (blue areas in Fig.
+1, and the dominant production region), and overestimating along the eastern to northern margins (red areas in Fig.
 1).
-The mean of these differences, which is the landcover map's bias, are 21\% and 34\% at 1 km for MODIS and GlobCover, respectively (Fig.
-S1), meaning that each map underestimates cropland by that amount at this resolution.
-MODIS bias drops to 8\% at 50 km of aggregation, whereas GlobCover bias is still 24\% at 100 km (Fig.
+The average biases for GlobCover and MODIS are 21\% and 34\% at 1 km, meaning that each map underestimates cropland by that amount at this resolution.
+MODIS mean bias drops to 8\% at 50 km of aggregation, whereas GlobCover bias is still 24\% at 100 km (SI Appendix, Fig.
 S1).
 
-The SA-LC and GeoWiki datasets show much spatial variability and overall bias.
+The SA-LC map uniformly overestimates cropland throughout the country (Fig.
+1), but the magnitude of mean bias is much lower at -8\% at 1 km to -6\% at 100 km  (Fig.
+S1).
+The GeoWiki map has a very heterogeneous pattern of bias, having a slight underestimation tendency (5\% mean bias) at 1 km, which changes to a smaller overestimation bias (-2 \%) at 100 km (Fig.
+S1).
 
-\subsection{Error as a function of cropland density}
+\subsection{The role of landscape characteristics in shaping bias}
+To investigate the degree to which landscape features influence landcover map bias, we extract all pixels in agricultural areas ($>$0.5\% cropland) of the 1 km reference map using the the boundaries of 354 magisterial districts (South Africa's finest administrative unit, which average 3,445 km$^2$ in size; SI Appendix, Fig.
+S3), and averaged the values of these pixels within each district.
+These provided an estimate of the intensity of cropping within the farmed parts of the landscape (with districts defining the landscape scale), as well as the degree of mixing between cropland and other land covers.
+We then extracted the cropland map biases for the same locations, and calculated the mean absolute bias within each district.
+Absolute bias describes the magnitude of bias but not its direction, and is more informative about the likelihood of the map being biased at any given point in the landscape than actual bias, as positive and negative biases can cancel each other out within small areas.
 
-\subsection{Potential bias in harvested areas, yield, and production estimates}
+A generalized additive model fit to district-level absolute bias (log-transformed) shows that absolute bias peaks at 50\% cropland cover for all but the GlobCover map (which continued to increase with cropland cover), and is lowest when the landscape is dominated either by cropland or other cover types (Fig.
+2).
+In other words, bias is highest when cropland cover is mixed evenly with other cover types.
 
-\subsection{Potential bias in estimates of carbon stocks}
+\vspace{-1 cm}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=.4\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
+\caption{The relationship between the mean absolute bias in cropland maps and cropland cover within agricultural areas (reference map pixels having $>$0.5\% cropland), averaged within the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
+Prediction curves are color-coded to the different cropland maps, with the solid line indicating  predicted absolute bias, and the lighter shading the standard error of the coefficients.
+Models were also fit to the mean absolute biases across all four maps (black curve), and from all maps exlcuding GlobCover (grey surve).}
+\label{afoto}
+\end{figure}
 
-\subsection{Potential bias in harvested areas, yield, and production estimates}
+\subsection{The impact of bias on calculating carbon stocks}
+We used the data of \cite{ruesch_new_2008} to calculate carbon densities for African forests, shrublands, croplands, grasslands, and sparse habitats.
+We assigned the cropland carbon values to map cells in proportion to their cropland cover.
+For the non-cropland proportions, we assigned the carbon value from each of the other types, such that we created four  different carbon maps for each landcover map at each aggregation scale (Fig.
+S4), which allowed us to test how carbon estimates vary as a function of i) bias in the cropland estimates and ii) the characteristics of adjacent cover types.
 
-\subsection{Impacts on evapotranspiration estimates}
+At the 1 km scale, the differences between country-wide total carbon stocks made using any of the cropland maps and the reference were within +/-3\%, regardless of which cover type was adjacent to cropland (Table S1), because differences between total stock estimates are diluted by the large area ($\sim$50-70\%) of the country having no cropland.
+Comparing the carbon stocks from each map's agricultural area reveals much greater differences (Table S1).
+SA-LC overestimates carbon stocks by just 2\% when the adjacent cover type is forest, and up to 15\% when it is spare cover.
+MODIS ranges from negligible overestimates in denser carbon classes (forest, secondary forest, and shrublands) to 8-13\% underestimates for the grassland and sparse classes.
+GeoWiki underestimates for all types, from $<$1\% for sparse cover to 8\% for forest.
+GlobCover shows the largest total carbon stock bias, with over-estimates ranging from for sparse lands to 162\% for forest.
+The magnitude of this bias is due to false positives--GlobCover identified cropland in nearly 50\% of pixels, compared to 30\% for the other three cropland maps.
+
+The spatial patterns of carbon bias (Fig.
+S4) are a reflection of the cropland biases (Fig.
+1).
+Where cropland is underestimated and the surrounding cover type is of higher carbon density than cropland, carbon density is overestimated.
+For lower density (grassland and sparse vegetation), there may be slight underestimates.
+These tendencies are reflected in the average of each map's spatial biases over agricultural areas (Fig.
+3).
+For example, MODIS and GlobCover overestimate carbon by $\sim$50\% on average at 1 km resolution when forest is the cropland-adjacent cover (stars in semi-transparent green and gold bars, Fig.
+3; Table S2).
+When it is sparse vegetation (open circles in Fig.
+3), MODIS underestimates by nearly 3\% at all scales, whereas maps that overestimate cropland (e.g.
+SA-LC, semi-transparent red) overestimate carbon density for this cover.
+Overall, GeoWiki has the lowest average spatial biases, for all cover types and all resolutions.
+It's maximum bias is a 12\% overestimate at 1 km for forest cover, thereafter all biases are less than 10\%, and converge to +/-2\% above 1 km aggregation (Fig.
+3, Table S2).
+All maps' biases are within +/-10\% bias after aggregation to 25 km.
+
+\vspace{-0.9 cm}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=.4\textwidth]{figures/carbon_veg_scale3.pdf}}
+\caption{The average spatial biases in carbon density estimates based on cropland maps, expressed as a percent difference relative to the reference map.
+The values of actual bias (incorporating bias direction and magnitude) and absolute bias (magnitude only) are presented in semi-transparent and solid colors, respectively, with colors denoting the specific cropland map and symbols indicating which cover type was used to calculate cropland adjacent carbon density.
+The bar represents the mean biases calculated across each of the 5 cover types.
+Shrubland and grassland bias values are near zero and not shown for display clarity.}
+\label{afoto}
+\end{figure}
+
+The patterns of mean absolute spatial bias (solid colored bars in Fig.
+3) generally follow that of actual bias, but with somewhat higher magnitudes and a few notable differences.
+The most notable is that Geo-Wiki has fairly large absolute bias at 1 km, averaging 27\% across across cover types (line in solid blue bar, Fig.
+3, Table S2), which is close to the 36-37\% for GlobCover and MODIS.
+SA-LC has the lowest absolute bias across scales, averaging (across cover types) from 14\% at 1 km to 3\% at 100 km.
+(Fig.
+3, Table S2).
+The increase in Geo-Wiki's absolute spatial bias relative to SA-LC's can be attributed to the spatially heterogeneous nature of its cropland biases (Fig.
+1) compared to the other three cropland maps.
+
+\subsection{Bias in harvested areas, yield, and production estimates}
+
+\begin{itemize}
+  \item Used provincial boundaries to extract total cropland from reference data.
+Cropland totals used to adjust cropland fractions falling within those boundaries from each cropland map so cropland map totals equal the provincial-level total from the reference (SI).
+  \item Disaggregated magisterial district level maize yield and maize harvested area onto the adjusted cropland fractions, and onto the reference map.
+  \item Calculated differences in spatial biases of yields and total production estimates at different scales of aggregation. 
+  \item Yields aggregated to higher scale using two methods: 1) weighted mean aggregation, where harvested area fractions used to weight yields at each aggregation scale, technically more difficult; 2) mean of any harvested area contributing to aggregation, without weighting.
+Production aggregated by either calculating production at 1 km and then aggregating to each scale and differencing, or aggregating harvested fractions and type 2 yield, and calculating production at each level.
+  \item Average bias normalized as a percentage of mean pixel-level yield or production calculated from the reference map.   
+ \end{itemize}
+
+\vspace{-1 cm}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=.4\textwidth]{figures/yield_bias_scale.pdf}}
+\caption{The mean spatial biases in disaggregated maize yield estimates, and the bias in total crop production estimates calculated from these yields and constrained crop area estimates.
+The values of actual bias (incorporating bias direction and magnitude) and absolute bias (magnitude only) are presented in semi-transparent and solid colors, respectively, with colors denoting the specific cropland map and symbols indicating whether the bias is related to one of two yield aggregation methods and associated crop production estimates.}
+\label{afoto}
+\end{figure}
+
+\begin{itemize}
+ \item GlobCover average bias shows large yield overestimates (nearly 60\%) at 1 km, declining with aggregation scales.
+Interestingly, MODIS and GeoWiki start with small yield underestimates that grow to 30\% yield under-estimates at 10 km, declining thereafter.
+SA-LC has lowest, +/-5\%.
+(Fig.
+4)
+ \item No production biases.
+ \item Absolute biases in both yield and production are large.
+All $>$40\% for both production and yield at 1 km, except SA-LC which is between 15-20\%.
+The biases decline with aggregation, but remain above 10\% for yield for most datasets at 100 km.
+
+\end{itemize}
+
+\subsection{Bias in evapotranspiration estimates}
+\begin{itemize}
+ \item Spatial biases most pronounced for VIC run with MODIS and GlobCover.
+These two overestimate ET along the southern margin, but typically by less than 5-10\%.
+(Fig.
+5).
+In the interior, VIC-GlobCover switches between under-  and over-estimation and then back again moving from West to East, but the bias is on the order of 5\% magnitude.
+MODIS shows a similar pattern.
+ \item GeoWiki and SA-LC shows much less discrepancy in interior. Greatest tendency is to underestimate along the southern coast. 
+ \item Difference in country-level ET estimates negligible ($<$2\%).
+ \item \textcolor{red}{\textbf{Justin}, can you give some thoughts on reasons for these patterns?
+The tendency seems to be that where cropland is underestimated (blue areas in Fig.
+1) then ET is overestimated (red areas in Fig.
+5).
+Here I calculate all difference as (ref - lc) / ref * 100, with lc being one of the four landcover maps.
+My expectation was that crop ET was higher than most surrounding vegetation types, so one would expect ET to be underestimated where cropland is underestimated?}
+\end{itemize}
+
+\vspace{-0.5 cm}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
+\caption{Differences in evapotranpiration estimates from a 29 year run of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four cropland maps.
+Columns indicate differences between ET estimates from different portions of the model's time series: annual mean ET; highest annual ET value in the time series (TS max); smallest annual ET value in the time series (TS min); total of mean monthly ET for the three months having the highest ET during the year (Peak 3); the mean monthly ET of the peak ET month.}\label{afoto}
+\end{figure}
 
 \subsection{Initialization errors in spatial agent-based models}
+
+\begin{itemize}
+  \item We selected four magisterial districts with similar climates, and differing abundances of cropland ranging from X-Y\%, and differing patterns of map bias (Fig. S8). 
+  \item Using household survey statistics from a project in Zambia that provides information on the average field area of farming households (X ha, SD = Y ha), we assigned each district a number of households determined by the total area of reference cropland in each district divided by mean farm size.
+\textcolor{red}{\textbf{Tom, Peng}, please correct as needed--maybe there is some variability to account for agent heterogeneity drawn by resampling from the distribution?}.
+This number of households represents the ``true'' number of households in the district.
+  \item Disaggregated cropland to 100 m resolution, which is required by the model. 
+  \item An agent-based model was then initialized wherein the household locations were seeded throughout the district, and assigned the required amount of cropland with a radius of X from the household location.
+  \item These steps were repeated for each cropland map using the same number of household agents. 
+  \item We then examined the differences that occurred between the agent initialization from the reference map and each of the four cropland maps, in terms of the percent of households that did not have cropland allocated, and the amount of cropland left unallocated in each district (Fig.
+6).
+  \item The percent households without farms is 0 when cropland is overestimated (because there is plenty of land to choose from), but increases in a 1:1 relationship with the amount of cropland underestimation (positive values on X-axis in Fig.
+6).
+\end{itemize}
+
+\vspace{-0.5 cm}
+\begin{figure}[ht]
+\centerline{\includegraphics[width=.3\textwidth]{figures/agent-bias.pdf}}
+\caption{Biases in agent-based model initialization relative to cropland map biases, measured in terms of the percent of households having no cropland allocated (top) and the percent of cropland left unallocated (bottom).
+Dot sizes represent a district number (smallest = district 1; largest = district 4), colors represent the landcover map.
+\textcolor{red}{to fix: add x-axis legend, add district number to legend.}}
+\label{afoto}
+\end{figure}
+
+\begin{itemize}
+  \item The percent of land unallocated, however, shows a U-shaped relationship with cropland biases, declining to near zero when cropland bias is between zero and up to $\sim$50\% under-estimated, thereafter increasing.
+Thus in cases where cropland bias insufficient farmland for households, a signficant share of available cropland can go unallocated (nearly 15\%).
+\textcolor{red}{\textbf{Tom, Peng}, I assume this has to do something with seeding and the spatial pattern of cropland, but not sure exactly what.
+Model not sure which HH to give the land too because there are too many for the amount of cropland available, so it gives up?
+}
+\end{itemize}
 
@@
 \section{Discussion}
-\subsection{Blather}
+
+Some points to make thrown out now because out of time. Others will be added. Please add any important implications/considerations you see from the results.
+
+Some points from my notes last year: 
+\begin{itemize}
+  \item Bias as a function of scale
+    \begin{enumerate}
+      \item At 1 km resolution all landcover products are still fairly biased.  
+       \item Bias drops to acceptable levels quickly for geowiki--at 5X5 km, mean bias is just 1\% (overestimated).
+The absolute bias for this dataset is 10\% or lower from 10X10 km resolution and coarser.
+       \item The SA dataset's bias is fairly consistent but low across all levels of aggregation, amounting to no more than an 8\% overestimate of cropland with absolute bias of similar magnitude. 
+       \item MODIS and GlobCover biases (mostly of underestimation) do not dissipate until the higher levels of aggregation.
+MODIS's actual bias (under-estimation) falls below 10\% at 20 km resolution, but the absolute bias remains above 10\% until more than a 100-fold aggregation is done ($>$100 km resolution).
+For GlobCover, it is still too high.
+    \end{enumerate}
+  \item Bias as a function of cropland cover
+    \begin{enumerate}
+      \item Classification algorithms are thus more error-prone where landcover is mixed/heterogenous.
+      \item The exception to this lies in the GlobCover dataset, where bias primarily increases as a function of cropland cover.
+The reason for this is that GlobCover's cropland classes do not provide for 100\% cropland cover, so aggregation tends to exacerbate underestimates.
+      \item Thus caution is needed when aggregating a mixed pixel class.  
+        \begin{itemize} 
+          \item An example illustrats this: take 4 1 km pixels, 2 of which are 100\% cropland, 2 of which are other cover types.
+Imagine a landcover product classifies 3 of these as cropland (2 correct, 1 an error of commission), using a cropland class that is defined as 50\% cropland.
+Aggregating the actual fraction by a factor of 4 will result in a new 4 km pixel having 50\% cropland, whereas aggregating the landcover product's pixels will give just 38\% cropland, even when factoring in the incorrect classification.)
+          \end{itemize}
+    \end{enumerate}
+  \item Bias as a function of method
+      \begin{enumerate}
+        \item Higher resolution and ensemble-based approaches have less bias.
+        \item geowiki represents a fusion of multiple coarse resolution data sources that has undergone extensive validation using a crowdsourcing approach
+        \item the SALC dataset is based on 30 m landsat data, but incorporates a range of ancillary data and expert judgement
+        \item MODIS and GlobCover data are effectively single source/single algorithm.
+        \item \textbf{Newer points begin here}
+        \item Statistically constrained constrained landcover estimation approaches provide accurate area-based inferences when aggregated.
+But spatial errors are still high, as seen with GeoWiki and production/yield estimates.
+Using these to identify yield gaps at specific map locations is inappropriate, or even for a larger location if it does not coincide with the geographic boundaries of the statistical unit.
+        \item Constrained estimates are also dependent on the accuracy of the statistics.  
+    \end{enumerate}
+  \item Fix above to have section on bias for global change studies
+    \begin{enumerate}
+      \item Scales at which it is safe to estimate values of say carbon stocks. 
+      \item Above point about bias in disaggregated yield estimates - no point mapping these out.
+A new approach might be to take these statistically reported yields and then combine them with satellite data to estimate yield variability within the district.
+That way would have meaningful reason for disaggregating yields, and would be pegged to real yield values, which would help minimize errors in remote sensing of yields.
+      \item Something on ET - doesn't seem to matter much, but land-atmosphere interactions can make these discrepancies meaningful, particularly since biases occur in arid areas where a lot of irrigation happens--can cause significant impacts on regional climate.
+etc.
+etc.
+Also we didn't change out land cover types, and the vegetation in SA around the cropland will have reasonably similar LAI and ET responses (I think), thus impact more muted than it might be elsewhere (e.g.
+in forested landscapes).
+      \item Agent-based models. \textcolor{red}{Tom, Peng, something of significance/implications of this, please}
+    \end{enumerate}
+  \item Will need a section on way forward for data, etc.
+Key role of accurate landcover, particularly agricultural.
+New methods, vectorized field boundaries seem to be highly valuable, Mapping Africa, Stephanie's paper, Geo-Wiki, etc are the way ahead.
+
+\end{itemize}
+
+Mixed landscapes increase the chances of omission and commission errors by increasing the number of cover classes, or because such landscapes are less spectrally distinct \cite{estes_diylandcover:_2015}
 % $\theta$ changes from 0 to 1. (see Figure \ref{afoto}).That means we are considering $\theta$  of the form
@@
 Perhaps it is right {\it SI Materials and Methods}.
+
+Describe weighted mean bias reasons.
 

### 2015-09-09 Rmd/yield-bias.Rmd
Drafted through yield/production biases. Output tables for supp for yield biases. Various.

@@
 library(SAcropland)
+library(xtable)
 p_root <- full_path(proj_root("SAcropland"), "SAcropland")
@@
 md <- md[md$ID_2 != 313, ]
-# round(mean(rgeos::gArea(md, byid = TRUE) / 1000000))
 md@data <- md@data[, c("PID", "ID_2", "NAME_1", "NAME_2")]  # trim down columns
@@
 md@data[md$ID_2 == 86, "NAME_2"] <- "HeidelbergG"
+
+# mean MD area
+round(mean(rgeos::gArea(md, byid = TRUE) / 1000000))  # 3445 km square
 
@@
 y_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1
+
+# check grids (output in QGIS and examine for sensible results--they are)
+# for(i in lev) {
+#   writeRaster(y_diff1$globmu[[i]]$gti, 
+#               file = full_path(p_data, paste0("test/globdiff", i, ".tif")))
+# }
+# for(i in lev) {
+#   writeRaster(list1[[i]]$gti, 
+#               file = full_path(p_data, paste0("test/gtiyldtst", i, ".tif")))
+# }
+# for(i in lev) {
+#   writeRaster(list2[[i]]$globmu, 
+#               file = full_path(p_data, paste0("test/globyld2tst", i, ".tif")))
+# }
+
 # type 2 (non-zero)
@@
 xa <- sapply(x, function(x) x + xo)
-cx <- c(2, 1.25, 1, 1)
+# cx <- c(2, 1.25, 1, 1)
+cx <- c(1.25, 1)
 g1 <- "grey90"
@@
 # mul <- c("p1mu", "p2mu", "p3mu", "y1mu", "y2mu")
-mul <- c("p2mu", "p3mu", "y1mu", "y2mu")
+# mul <- c("p2mu", "p3mu", "y1mu", "y2mu")
+mul <- c("p1mu", "y1mu")
 #mal <- c("p1ma", "p2ma", "p3ma", "y1ma", "y2ma")
-mal <- c("p2ma", "p3ma", "y1ma", "y2ma")
+# mal <- c("p2ma", "p3ma", "y1ma", "y2ma")
+mal <- c("p1ma", "y1ma")
 malmul <- list(mal, mul)
 #pyl <- c("mup", "mup", "mup", "muy", "muy")
-pyl <- c("mup", "mup", "muy", "muy")
+# pyl <- c("mup", "mup", "muy", "muy")
+pyl <- c("mup", "muy")
 
@@
 lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
-pchs <- c("*", "+", "o", "x", "-")
+# pchs <- c("*", "+", "o", "x", "-")
+pchs <- c("+", "o")
 
@@
 text(x = 12.2, y = -42, labels = "Actual", srt = 45, adj = c(0, 0), cex = 0.8)
-legend(x = 12.1, y = 65, legend = c("Prod I", "Prod II", "Yield 1", "Yield 2"),        pch = pchs, pt.cex = c(2.5, 1.5, 1.5, 1.5), bty = "n", cex = 0.8)
+# legend(x = 12.1, y = 65, legend = c("Prod I", "Prod II", "Yield 1", "Yield 2"),        pch = pchs, pt.cex = c(2.5, 1.5, 1.5, 1.5), bty = "n", cex = 0.8)
+legend(x = 12.1, y = 65, legend = c("Production", "Yield"), pch = pchs, 
+       pt.cex = c(1.5, 1.5), bty = "n", cex = 0.8)
+
 dev.off()
@@
 })
+
+do.call(cbind, lapply(lev, function(x) {
+  rbind(cbind.data.frame("Map" = y1mu[ol == x, il], 
+                         "v" = round(y1mu[ol == x, bias] / muy[x] * 100, 1)),
+        cbind.data.frame("Map" = p1mu[ol == x, il], 
+                         "v" = round(p1mu[ol == x, bias] / mup[x] * 100, 1)))
+}))
+
+vnms <- c(rep("Yield", length(snms)), rep("Production", length(snms)))
+aa <- c("Actual", "Absolute")
+out_tab <- do.call(rbind, lapply(1:2, function(x) {
+  out <- do.call(cbind, lapply(1:length(lev), function(ii) {
+    lv <- lev[ii]  
+    mm <- malmul[2:1][[x]] 
+    v <- do.call(rbind, lapply(length(mm):1, function(j) {
+      v2 <- do.call(rbind, lapply(1:length(snms), function(i) {
+        nm <- snms[i]
+        round(get(mm[j])[ol == lv & il == nm,  bias] / get(pyl[j])[lv] * 100,1)
+      }))
+      rownames(v2) <- snms
+      v2
+    }))
+  }))
+  out <- cbind.data.frame(aa[x], mcap, "Variable" = vnms, unname(out))
+  colnames(out) <- c("Bias", "Map", "Variable", paste(c(1, fact), "km"))
+  out
+}))
+
+caption <- paste("Mean actual and absolute biases for yield and", 
+                 "production estimates disaggregated from magisterial", 
+                 "district level farm census data onto adjusted cropland", 
+                 "maps. Biases were calculated as percent differences", 
+                 "relative to the mean value calculated from the reference", 
+                 "map.")    
+out_xtab <- xtable(out_tab, digits = 1, caption = caption)
+print(out_xtab, type = "latex", file = "paper/figures/yldprod-bias.tex", 
+      tabular.environment = "longtable", floating = FALSE, 
+      caption.placement = "top", include.rownames = FALSE)
 

### 2015-09-09 paper/main.tex
Drafted through yield/production biases. Output tables for supp for yield biases. Various.

@@
 \vspace{-0.5 cm}
-\section{Results}
+\section{Bias}
 \subsection{Landcover bias}
@@
 The patterns of mean absolute spatial bias (solid colored bars in Fig.
-3) generally follow that of actual bias, but with somewhat higher magnitudes and a few notable differences.
+3) generally follow that of actual bias, but with somewhat higher magnitudes and a few important differences.
 The most notable is that Geo-Wiki has fairly large absolute bias at 1 km, averaging 27\% across across cover types (line in solid blue bar, Fig.
@@
 \subsection{Bias in harvested areas, yield, and production estimates}
+We used the reference dataset to calculate total cropland area within South Africa's nine provinces, and then adjusted the the four cropland maps so that they each summed to the same area within each province \cite[per][]{ramankutty_farming_2008}.
+Despite this statistical constraint, the updated cropland maps still had substantial spatial biases similar in pattern (Fig.
+S5) to those in the unadjusted maps (Fig.
+1), and we evaluated how these biases affected crop production estimates.
+We derived maize yield and harvested area estimates by disaggregating district-level  (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} following Monfreda \emph{et al.} \cite{monfreda_farming_2008}, and used that to calculate spatial biases in yield and production estimates at each aggregation scale.
 
-\begin{itemize}
-  \item Used provincial boundaries to extract total cropland from reference data.
-Cropland totals used to adjust cropland fractions falling within those boundaries from each cropland map so cropland map totals equal the provincial-level total from the reference (SI).
-  \item Disaggregated magisterial district level maize yield and maize harvested area onto the adjusted cropland fractions, and onto the reference map.
-  \item Calculated differences in spatial biases of yields and total production estimates at different scales of aggregation. 
-  \item Yields aggregated to higher scale using two methods: 1) weighted mean aggregation, where harvested area fractions used to weight yields at each aggregation scale, technically more difficult; 2) mean of any harvested area contributing to aggregation, without weighting.
-Production aggregated by either calculating production at 1 km and then aggregating to each scale and differencing, or aggregating harvested fractions and type 2 yield, and calculating production at each level.
-  \item Average bias normalized as a percentage of mean pixel-level yield or production calculated from the reference map.   
- \end{itemize}
+The yields disaggregated onto the cropland maps were markedly different to those on the reference map, particularly in the lower density cropland areas in the center of the country, where GlobCover overestimated yields and MODIS and GeoWiki (to a lesser extent) underestimated them at 1 km resolution (Fig.
+S6).
+However, the mean spatial bias in yields at this resolution was pronounced only for GlobCover, equivalent to nearly 60\% of the mean reference yield of 3.4 tons ha$^{-1}$, compared to +/-5\% for the other maps (Fig.
+5).
+Interestingly, GeoWiki and MODIS biases increased with aggregation, peaking at 10 km where both datasets underestimated yields by 30\%, thereafter declining to 10\% at 100 km, whereas GlobCover's mean yield bias declined linearly with aggregation (Fig.
+5).
 
-\vspace{-1 cm}
-\begin{figure}[ht]
+%\vspace{-1 cm}
+\begin{figure}[!ht]
 \centerline{\includegraphics[width=.4\textwidth]{figures/yield_bias_scale.pdf}}
@@
 
-\begin{itemize}
- \item GlobCover average bias shows large yield overestimates (nearly 60\%) at 1 km, declining with aggregation scales.
-Interestingly, MODIS and GeoWiki start with small yield underestimates that grow to 30\% yield under-estimates at 10 km, declining thereafter.
-SA-LC has lowest, +/-5\%.
-(Fig.
-4)
- \item No production biases.
- \item Absolute biases in both yield and production are large.
-All $>$40\% for both production and yield at 1 km, except SA-LC which is between 15-20\%.
-The biases decline with aggregation, but remain above 10\% for yield for most datasets at 100 km.
+In contrast to yields, production errors were unbiased in aggregate (Fig.
+5).
+The statistical constraints on harvested and cropland areas led to the canceling out of spatial errors in production estimates, which is evident in the checkerboard-like pattern in maps of production biases (Figure S7).
+However, this pattern of bias means that the average absolute biases in  production estimates are large, between 40 to 55\% for GeoWiki, MODIS, and GlobCover at 1 km, and remaining generally high ($<$10\%) for these three datasets even up to 50-100 km of aggregation (Fig.
+5).
+SA-LC production biases were lowest across all spatial scales (20\% at 1 km to 2\% in 100 km).
 
-\end{itemize}
+Absolute mean biases in yield were also substantial, and generally 10-15\% larger than production biases across all spatial scales, except for GlobCover where absolute production biases exceeded yield bias at 5-100 km of aggregation .
 

### 2015-09-11 Rmd/agent-allocation-bias.Rmd
Finished draft through results. Updated Figure 6 (agent-bias), updated ET bias map, added Appendix to repo

--- 2015-08-21 Rmd/agent-allocation-bias.Rmd
@@
 ```{r, eval=FALSE}
+library(SAcropland)
 library(raster)
@@
 library(readxl)
+library(rgeos)
 
@@
  # -1.0  0.0 -0.4 -0.2  # tinf differences!
+rgeos::gArea(abm_md, byid = TRUE) / 10000 / 100
 ```
@@
             overwrite = TRUE)
+cpct <- extract(gti4, abm_md)
+dcrop_pct <- round(sapply(cpct, sum, na.rm = TRUE) / sapply(cpct, length), 1)
 ```
+Cropland percentages:
+
+  + Clocolan (269) - 45%
+  + Ficksburg (274) - 39%
+  + Fouriesburg (275) - 29%
+  + Marquard (288) - 44%
+
 <a href="#top">Back to top</a>
@@
 plot(abm_md)
-plot(cb[[1]], axes = FALSE, box = FALSE, breaks = brks1, col = cols1, add = TRUE, 
-     legend = FALSE)
+plot(cb[[1]], axes = FALSE, box = FALSE, breaks = brks1, col = cols1, 
+     add = TRUE, legend = FALSE)
 plot(abm_md, add = TRUE)
+polygonsLabel(abm_md, labels = paste0(abm_md$name_2, " (", 1:4, ")"), 
+              method = "buffer", cex = 0.9)
 for(i in 2:5) {
   plot(abm_md)
-  plot(cb[[1]] - cb[[i]], axes = FALSE, box = FALSE, breaks = brks, col = cols, 
-       add = TRUE, legend = FALSE)
+  plot(cb[[1]] - cb[[i]], axes = FALSE, box = FALSE, breaks = brks2, 
+       col = cols2, add = TRUE, legend = FALSE)
   plot(abm_md, add = TRUE, lwd = 2)
+
   mtext(mcap[i - 1], side = 1, line = -4, adj = 0.7)
@@
 cexs <- seq(1, 2, 1 / 3)
-pdf(full_path(p_fig, "agent-bias.pdf"), height = 3.5, width = 7)
+# pdf(full_path(p_fig, "agent-bias.pdf"), height = 5, width = 3)
+pdf(full_path(p_fig, "agent-bias.pdf"), height = 3, width = 5)
 cx <- c(0.7, 0.8, 0.6)
 # png(full_path(p_fig, "agent-bias.png"), height = 350, width = 700)
-par(mfrow = c(1, 2), mar = c(3, 2.5, 3, 1), mgp = c(1.5, 0.25, 0), tcl = -0.2, 
-    oma = c(0, 0, 0, 4))
+# par(mfrow = c(2, 1), mar = c(0.5, 2.5, 0.5, 1), mgp = c(1, 0.05, 0), 
+#     tcl = -0.2, oma = c(3.5, 0, 0, 0))
+par(mfrow = c(1, 2), mar = c(2, 2, 0.5, 0), mgp = c(0.75, 0.05, 0), 
+    tcl = -0.2, oma = c(3.5, 0, 0, 0.1))
 scols <- c("red", "orange3", "green4", "blue")
@@
 par(xpd = NA)
-xst <- seq(0.86, 0.96, 0.07 / 3)
-xy <- cbind(grconvertX(xst, from = "ndc"), 
-            grconvertY(rep(0.6, 5), from = "ndc"))
-for(i in 2:length(xst)) {
-  if(i == 2) {
-    leg <- 1:4
+
+yst <- seq(0.05, 0.23, 0.18 / 4)
+xy <- cbind(grconvertX(rep(0.5, 5), from = "ndc"), 
+            grconvertY(yst, from = "ndc"))
+for(i in 2:length(yst)) {
+  if(i == 5) {
+    leg <- c("District", 1:4)
   } else {
@@
   }
-  legend(x = xy[i, 1], y = xy[i, 2], legend = leg, pch = 20, cex = cx[3], 
-         adj = c(4.5, 0.75),  
+  legend(x = xy[i, 1], y = xy[i, 2], legend = rep("",4), #legend = leg, 
+         pch = 20, cex = cx[3], 
+         horiz = TRUE, adj = c(4, -5.75),  
          col = scols[i - 1], pt.cex = cexs, bty = "n")#cex = cexs,)
 }
+yst <- seq(0.05, 0.23, 0.18 / 4)
+xy <- cbind(grconvertX(rep(0.515, 5), from = "ndc"), 
+            grconvertY(yst, from = "ndc"))
+text(x = xy[, 1], y = xy[, 2], labels = c(mcap, "District"), 
+     #srt = 90, #pos = 2, 
+     #adj = c(-1, 1), 
+     pos = 2, col = c(scols, "black"), cex = cx[1])
+xst <- seq(0.545, 0.65, 0.105 / 3)
 xy <- cbind(grconvertX(xst, from = "ndc"), 
-            grconvertY(rep(0.58, 5), from = "ndc"))
-text(x = xy[, 1], y = xy[, 2], labels = c("District", mcap), srt = 90, #pos = 2, 
-     adj = c(0, 1.5), col = c("black", scols), cex = cx[1])
+            grconvertY(rep(yst[length(yst)], 4), from = "ndc"))
+text(x = xy[, 1], y = xy[, 2], labels = 1:4, pos = 2, cex = cx[1])
 dev.off()

### 2015-09-11 Rmd/et-bias.Rmd
Finished draft through results. Updated Figure 6 (agent-bias), updated ET bias map, added Appendix to repo

@@
 library(SAcropland)
+library(RColorBrewer)
 p_root <- full_path(proj_root("SAcropland"), "SAcropland")
@@
 
-# year indices
+# date indices
 yrs <- gsub("*.*_", "", lnames)
 mos <- gsub("_.*", "", lnames)
-yi <- t(sapply(unique(yrs), function(x) range(which(yrs == x))))
-mi <- sapply(unique(mos), function(x) which(mos == x))
+yi <- t(sapply(unique(yrs), function(x) range(which(yrs == x))))  # year index
+mi <- sapply(unique(mos), function(x) which(mos == x))  # month index
 ```
@@
 # reshape in list
-l2dag <- list(amu_diff, mumax_diff, mumin_diff, gs3_diff, gsmx_diff)
-names(l2dag) <- c("mu", "mumx", "mumn", "gs", "gspk")
+# l2dag <- list(amu_diff, mumax_diff, mumin_diff, gs3_diff, gsmx_diff)
+# names(l2dag) <- c("mu", "mumx", "mumn", "gs", "gspk")
+l2dag <- list(amu_diff, mumax_diff, mumin_diff, gs3_diff)
+names(l2dag) <- c("mu", "mumx", "mumn", "gs")
 l2dag <- lapply(l2dag, disaggl)
@@
 
-fnm <- "et_bias_map.pdf"
+fnm <- "et_bias_map4.pdf"
 legtext <- rep("%", 5)
@@
 mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
-ecap <- c("Annual mean", "TS max", "TS min", "Peak 3", "Peak")
+# ecap <- c("Annual mean", "TS max", "TS min", "Peak 3", "Peak")
+ecap <- c("Annual mean", "TS max", "TS min", "Peak")
 
@@
 
-brks <- c(seq(floor(rng[1]), -5, by = 5), -1, 1, seq(5, ceiling(rng[4]), 5))
-cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)
-cols[c(1, length(cols))] <- c("darkred", "purple3")
+# For Supplementals
+# brks <- c(seq(floor(rng[1]), -5, by = 5), -1, 1, seq(5, ceiling(rng[4]), 5))
+brks <- c(-25, -15, -10, -5, -3, -1, 1, 3, 5, 10, 15, 25)
+# cvec <- c("red", "orange", "grey80", "green", "blue4")
+# cols <- colorRampPalette(cvec)(length(brks) - 1)
+# cols[c(1, length(cols))] <- c("darkred", "purple3")
+cols <- brewer.pal(length(brks) - 1, name = "RdBu")
+cols[6] <- "grey80"
 pdf(full_path(p_fig, fnm), height = 6, width = 7)
-par(mfcol = c(4, 5), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
+par(mfcol = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
 for(j in 1:length(l2dagd)) {
@@
     rl <- l2dagd[[j]][[k]]$f25
-    print(snms[k])
+    print(paste("...", snms[k]))
     plot(sa.shp, lty = 0)
@@
 dev.off()
+
+fnm <- "et_bias_map.pdf"
+pdf(full_path(p_fig, fnm), height = 6, width = 7)
+par(mfcol = c(2, 2), mar = c(0, 0, 0, 0), oma = c(4, 0, 1, 0))
+for(j in 1) {
+  print(names(l2dagd)[j])
+  for(k in 1:length(snms)) {
+    rl <- l2dagd[[j]][[k]]$f25
+    print(paste("...", snms[k]))
+    plot(sa.shp, lty = 0)
+    image(rl, add = TRUE, col = cols, breaks = brks, legend = FALSE)
+    mtext(mcap[k], side = 3, line = -1, cex = cx)
+  }
+}
+flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
+            legend.vals = brks, longdims = c(0.2, 0.8), 
+            shortdims = c(0.07, 0.01), 
+            colvec = cols, #(length(brks) - 1), 
+            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
+            legend.pos = c(4, 5), leg.adj = list(c(0.3, 0.3), c(-1, -0.75)), 
+            cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
+dev.off()
+
 ```
@@
 mus[, ol := NULL]
-setnames(mus, c("x", "il"), c("ETvar", "dataset"))
-setkeyv(mus, c("dataset", "ETvar"))
+setnames(mus, c("x", "il", "bias"), c("Variable", "Map", "Bias"))
+setkeyv(mus, c("Map", "Variable"))
 muas <- rbindlist(lapply(names(mua), function(x) cbind(x, mua[[x]])))
 muas[, ol := NULL]
-setnames(muas, c("x", "il", "bias"), c("ETvar", "dataset", "absbias"))
-setkeyv(muas, c("dataset", "ETvar"))
+setnames(muas, c("x", "il", "bias"), c("Variable", "Map", "Abs Bias"))
+setkeyv(muas, c("Map", "Variable"))
 mutab <- mus[muas]
 mutab[, lapply(.SD, function(x) max(abs(x))), .SDcols = 3:4]
+setkey(mutab, "Variable")
+vnmr <- c("Annual Mean", "29-year Max", "29-year Min", "Peak")
+for(i in 1:length(vnms)) mutab[Variable == vnms[i], Variable := vnmr[i]]
+setkey(mutab, "Map")
+for(i in 1:length(snms)) mutab[Map == snms[i], Map := mcap[i]]
+mutab[order(-rank(Map))]
 
-mutab_xtab <- xtable::xtable(mutab, digits = 1)
-print(mutab_xtab, type = "latex", 
-      file = "paper/figures/et-bias.tex")
+caption <- paste("Mean actual and absolute biases (as \\%) for",              
+                 "evapotranspiration variables derived from a 29-year time",
+                 "series calculated by the VIC model, including the average", 
+                 "total ET for the", 
+                 "3 months of the year when ET is highest, the annual mean", 
+                 "and the minimum and maximum annual ETs in the time series.")
+mutab_xtab <- xtable::xtable(mutab, digits = 1, caption = caption)
+print(mutab_xtab, type = "latex", file = "paper/figures/et-bias.tex",
+      tabular.environment = "longtable", floating = FALSE, 
+      caption.placement = "top", include.rownames = FALSE)
 ```
@@
 
-The largest (in an absolute sense) mean actual bias is -0.6%, while the largest mean absolute bias is 1.43%, so not much here.
+The largest (in an absolute sense) mean actual bias is -0.6%, while the largest mean absolute bias is 1.25%, so not much here.
 
@@
 
-plot(presa)
+plot(presa, breaks = seq(0, 1000, 100), col = topo.colors(9))
 plot(abs(l2dag$mu$globmu$f25))

### 2015-09-11 paper/app_s1.tex
Finished draft through results. Updated Figure 6 (agent-bias), updated ET bias map, added Appendix to repo

\RequirePackage{lineno}  % http://www-d0.fnal.gov/Run2Physics/WWW/templates/lineno.html
\documentclass[12pt]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
% parameters from other files
\newcommand{\citetapos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}  % apostrophe's in cites
%\usepackage[square,numbers,sort&compress]{natbib}
\def\linenumberfont{\normalfont\small\sffamily}
\usepackage{graphicx}
\usepackage{epstopdf}
% From link below: next two lines allow caption to fill the full figure box, and caption formats to be changed (e.g. font)
% http://tex.stackexchange.com/questions/107350/caption-below-the-figure-and-aligned-with-left-side-of-figure
% http://ctan.mackichan.com/macros/latex/contrib/caption/caption-eng.pdf
\usepackage{caption}
\captionsetup[figure]{slc=off, font=footnotesize}
%\usepackage{amsmath}
\usepackage{longtable}
\usepackage{placeins}

\begin{document}
\bibliographystyle{unsrt}

\setpagewiselinenumbers
%\modulolinenumbers[5]
%\linenumbers
\section*{\large Appendix S1: Supplementary Results}
%\begin{table}[ht]
%\centering
%\caption{Confidence classes for the forcing data based on bias-correction weights.}
%\begin{tabular}{cc}
%  \hline
%  Confidence class & Bias-correction weight \\ 
%  \hline
%High & 0.75-1.00 \\ 
%Medium-High & 0.50-0.75\\ 
%Medium-Low & 0.25-0.50\\ 
% Low & 0.05-0.25 \\ 
% None & $<$0.05 \\ 
%   \hline
%\end{tabular}
%\end{table}  
\section*{Landcover map bias}

\begin{figure}[ht]
  \centering
     \includegraphics[width = 16cm]{figures/biases_1-100km.pdf} 
%     \vspace{-3cm}
      \caption{Mean biases for each of the landcover maps as a function of cropland density (calculated using the 2011 reference maps) and aggregation scales. Rows present biases by map product, columns by aggregation scale.  Dash lines indicate mean actual bias at each level of cropland density, calculated in bins spanning 5\% of density (e.g. 0-5\% cropland cover, 5-10\%, etc.), while solid lines indicate the mean absolute bias.  The black numbers in each plot area present the overall means of actual and absolute bias for each sensor-scale combination (actual/absolute). }
      \label{fig:default}
\end{figure}

%\vspace{0.5 cm}
%\begin{equation}
%\hspace*{1.5 cm} 
%   \label{metric}
%     R_n  = \frac{(1 - \alpha)dswr + dlwr - \sigma * T^4}{\lambda}
%\end{equation}

\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/biases_1km.pdf} 
%     \vspace{-3cm}
      \caption{Mean biases for each of the landcover maps at 1 km resolution, as a function of cropland density. Colored lines (color-coded to map product name) show the mean bias at each level of cropland density, calculated in bins spanning 5\% (e.g. 0-5\% cropland cover, 5-10\%, etc.). Box plots show the variability of bias in each bin (whiskers = 2.5 and 97.5 percentiles, box the inter-quartile, and grey bar in box the median). Actual biases are presented in the top row, and absolute biases in the bottom row.}
      \label{fig:default}
\end{figure}

\begin{figure}[ht]
  \centering
     \includegraphics[width = 8cm]{figures/md_map.pdf} 
%     \vspace{-3cm}
      \caption{South Africa's magisterial districts.}
      \label{fig:default}
\end{figure}

%\FloatBarrier
\section*{Carbon bias}
\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/carbon_bias_map.pdf} 
%     \vspace{-3cm}
      \caption{Spatial patterns of mean bias (across four different possible cover types adjacent to cropland) in carbon stock estimates. }
      \label{fig:default}
\end{figure}

\input{figures/totC-bias.tex}

\FloatBarrier
\input{figures/C-bias-mean-cropland-SA.tex}

\input{figures/C-bias-mean-all-SA.tex}

\section*{Yield and Harvested Area Bias}
\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/cropland_adj_bias_map2.pdf} 
      \caption{Biases in cropland maps adjusted using provincial level cropland area estimates.}
      \label{fig:default}
\end{figure}

\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/yld_bias_map_2.pdf} 
      \caption{Biases in disaggregated maize yield estimates.}
      \label{fig:default}
\end{figure}

\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/prod_bias_map_2.pdf} 
      \caption{Biases in production estimates calculated from disaggregated maize yield and harvested area estimates.}
      \label{fig:default}
\end{figure}

\FloatBarrier
\input{figures/yldprod-bias.tex}

\section*{Evapotranspiration bias}
\input{figures/et-bias.tex}

\section*{Agent allocation bias}

\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/abm-selected-districts.pdf} 
      \caption{The location of the four selected magisterial districts (top left) used in evaluating agent allocation bias, the reference levels of cropland cover within those districts (top right), and the difference in cropland percentage between the reference and each of the four cropland maps (lower four panels). }
      \label{fig:default}
\end{figure}

\FloatBarrier
\section*{References}
\bibliography{}
\end{document}

### 2015-09-11 paper/figures/et-bias.tex
Finished draft through results. Updated Figure 6 (agent-bias), updated ET bias map, added Appendix to repo

@@
 % latex table generated in R 3.2.1 by xtable 1.7-4 package
-% Wed Aug 19 16:18:01 2015
-\begin{table}[ht]
-\centering
-\begin{tabular}{rllrr}
+% Wed Sep  9 13:32:11 2015
+\begin{longtable}{llrr}
+\caption{Mean actual and absolute biases (as \%) for evapotranspiration variables derived from a 29-year time series calculated by the VIC model, including the average total ET for the 3 months of the year when ET is highest, the annual mean and the minimum and maximum annual ETs in the time series.} \\
   \hline
- & ETvar & dataset & bias & absbias \\ 
+Variable & Map & Bias & Abs Bias \\ 
   \hline
-1 & gs & geow & 0.2 & 0.8 \\ 
-  2 & gspk & geow & 0.2 & 0.8 \\ 
-  3 & mu & geow & 0.3 & 0.7 \\ 
-  4 & mumn & geow & 0.3 & 0.6 \\ 
-  5 & mumx & geow & 0.3 & 0.8 \\ 
-  6 & gs & globmu & 0.1 & 1.2 \\ 
-  7 & gspk & globmu & 0.2 & 1.4 \\ 
-  8 & mu & globmu & -0.1 & 1.0 \\ 
-  9 & mumn & globmu & -0.2 & 0.9 \\ 
-  10 & mumx & globmu & 0.2 & 1.2 \\ 
-  11 & gs & modmu & -0.5 & 0.9 \\ 
-  12 & gspk & modmu & -0.4 & 1.0 \\ 
-  13 & mu & modmu & -0.6 & 0.8 \\ 
-  14 & mumn & modmu & -0.6 & 0.7 \\ 
-  15 & mumx & modmu & -0.5 & 0.8 \\ 
-  16 & gs & sa30 & 0.3 & 0.7 \\ 
-  17 & gspk & sa30 & 0.2 & 0.8 \\ 
-  18 & mu & sa30 & 0.5 & 0.6 \\ 
-  19 & mumn & sa30 & 0.4 & 0.5 \\ 
-  20 & mumx & sa30 & 0.4 & 0.6 \\ 
+Peak & GeoWiki & 0.2 & 0.8 \\ 
+  Annual Mean & GeoWiki & 0.3 & 0.7 \\ 
+  29-year Min & GeoWiki & 0.3 & 0.6 \\ 
+  29-year Max & GeoWiki & 0.3 & 0.8 \\ 
+  Peak & GlobCover & 0.1 & 1.2 \\ 
+  Annual Mean & GlobCover & -0.1 & 1.0 \\ 
+  29-year Min & GlobCover & -0.2 & 0.9 \\ 
+  29-year Max & GlobCover & 0.2 & 1.2 \\ 
+  Peak & MODIS & -0.5 & 0.9 \\ 
+  Annual Mean & MODIS & -0.6 & 0.8 \\ 
+  29-year Min & MODIS & -0.6 & 0.7 \\ 
+  29-year Max & MODIS & -0.5 & 0.8 \\ 
+  Peak & SA-LC & 0.3 & 0.7 \\ 
+  Annual Mean & SA-LC & 0.5 & 0.6 \\ 
+  29-year Min & SA-LC & 0.4 & 0.5 \\ 
+  29-year Max & SA-LC & 0.4 & 0.6 \\ 
    \hline
-\end{tabular}
-\end{table}
+\hline
+\end{longtable}

### 2015-09-11 paper/main.tex
Finished draft through results. Updated Figure 6 (agent-bias), updated ET bias map, added Appendix to repo

@@
 \setlength\linenumbersep{2pt}
+\usepackage{placeins}
 
@@
 These include first-order analyses, in which values for a variable of interest are mapped to particular cover type(s), and second-order analyses, in which a process model draws on the cover types' values to calculate an output value.
-We created four datasets to represent second order analyses.
+We created four datasets to represent first order analyses.
 The first was a series of maps of vegetated carbon stocks created following the methodology of Ruesch and Gibbs' \cite{ruesch_new_2008}.
@@
 
-Absolute mean biases in yield were also substantial, and generally 10-15\% larger than production biases across all spatial scales, except for GlobCover where absolute production biases exceeded yield bias at 5-100 km of aggregation .
+Absolute mean biases in yield were also substantial, and generally 10-15\% larger than production biases across all spatial scales, except for GlobCover where absolute production biases exceeded yield bias at 5-100 km of aggregation.
 
 \subsection{Bias in evapotranspiration estimates}
-\begin{itemize}
- \item Spatial biases most pronounced for VIC run with MODIS and GlobCover.
-These two overestimate ET along the southern margin, but typically by less than 5-10\%.
-(Fig.
-5).
-In the interior, VIC-GlobCover switches between under-  and over-estimation and then back again moving from West to East, but the bias is on the order of 5\% magnitude.
-MODIS shows a similar pattern.
- \item GeoWiki and SA-LC shows much less discrepancy in interior. Greatest tendency is to underestimate along the southern coast. 
- \item Difference in country-level ET estimates negligible ($<$2\%).
- \item \textcolor{red}{\textbf{Justin}, can you give some thoughts on reasons for these patterns?
-The tendency seems to be that where cropland is underestimated (blue areas in Fig.
-1) then ET is overestimated (red areas in Fig.
-5).
-Here I calculate all difference as (ref - lc) / ref * 100, with lc being one of the four landcover maps.
-My expectation was that crop ET was higher than most surrounding vegetation types, so one would expect ET to be underestimated where cropland is underestimated?}
-\end{itemize}
+Compared to carbon and crop related examples, actual and absolute biases in evapotranspiration calculated using the VIC model were small and averaged to less than +/-1\%.
+There are several hotspots of discrepancy, however, which are evident in maps of spatial biases (Fig.
+6).
+The most pronounced of these are the 5-15\% overestimates by VIC initialized with MODIS and GlobCover in the center of the country, and along the southern and western coasts, particularly the latter where the ET overestimates reach 25\%.
+These locations correspond primarily to margins of major crop production regions--in the center is the westernmost boundary of the summer rainfall growing region, where maize is the primary crop, while the bias hotspot up the west coast marks one of the two boundaries of the winter rainfall region where wheat farming dominates\cite{hardy_rainfed_2011}.
+These two boundaries correspond respectively to approximately 400 and 200 mm of growing season rainfall.
 
-\vspace{-0.5 cm}
-\begin{figure}[ht]
+SA-LC and GeoWiki also produced biased ET estimates along the southern and western coasts, but here the tendency was to underestimate ET, while biases in the center of the country were either negligible to absent.
+All but MODIS underestimated ET by 5-15\% in the northern tip of the country.
+
+\FloatBarrier
+\begin{figure}[!ht]
 \centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
-\caption{Differences in evapotranpiration estimates from a 29 year run of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four cropland maps.
-Columns indicate differences between ET estimates from different portions of the model's time series: annual mean ET; highest annual ET value in the time series (TS max); smallest annual ET value in the time series (TS min); total of mean monthly ET for the three months having the highest ET during the year (Peak 3); the mean monthly ET of the peak ET month.}\label{afoto}
+\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four cropland maps.}\label{afoto}
 \end{figure}
 
-\subsection{Initialization errors in spatial agent-based models}
+\subsection{Initialization biases in agent-based models}
+We used an agent-based model of food security that represents the interactions between hundreds of individual farming households over multiple seasons \cite{chen_dependency_2013}.
+The model is thus computationally intensive, and, like many spatial ABMs, run over smaller geographic domains (e.g.
+districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) that are needed to represent the different land units of single farmers.
+To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig.
+S8) in the eastern part of the country, having between 28-45\% of their areas devoted to cropland, according to the reference map.
 
-\begin{itemize}
-  \item We selected four magisterial districts with similar climates, and differing abundances of cropland ranging from X-Y\%, and differing patterns of map bias (Fig. S8). 
-  \item Using household survey statistics from a project in Zambia that provides information on the average field area of farming households (X ha, SD = Y ha), we assigned each district a number of households determined by the total area of reference cropland in each district divided by mean farm size.
-\textcolor{red}{\textbf{Tom, Peng}, please correct as needed--maybe there is some variability to account for agent heterogeneity drawn by resampling from the distribution?}.
-This number of households represents the ``true'' number of households in the district.
-  \item Disaggregated cropland to 100 m resolution, which is required by the model. 
-  \item An agent-based model was then initialized wherein the household locations were seeded throughout the district, and assigned the required amount of cropland with a radius of X from the household location.
-  \item These steps were repeated for each cropland map using the same number of household agents. 
-  \item We then examined the differences that occurred between the agent initialization from the reference map and each of the four cropland maps, in terms of the percent of households that did not have cropland allocated, and the amount of cropland left unallocated in each district (Fig.
-6).
-  \item The percent households without farms is 0 when cropland is overestimated (because there is plenty of land to choose from), but increases in a 1:1 relationship with the amount of cropland underestimation (positive values on X-axis in Fig.
-6).
-\end{itemize}
+We disaggregated the cropland percentages in all maps to binary cropland/non-cropland cover types with 100 m resolution, which matches the typical field size (1 ha) for smallholder farmers in household survey data (collected in Zambia) used in developing the agent-based model \cite{chen_dependency_2013}.
+These surveys found mean household crop field area to be 2 ha, which we divided into reference cropland areas to estimate the total households within each district.
+We then initialized the model by assigning each household agent two cropland pixels.
+In order to emulate the natural groupings of communities, the model only assigns a household fields that are within 1.5 km of other agents' fields, and if those pixels have not already been allocated to another agent.
+The model thus iteratively grows ``communities'' until all households are assigned cropland, or all available cropland is allocated.
 
-\vspace{-0.5 cm}
+We used each of the five maps to initialize the model, and assessed how bias in cropland cover estimates impacted the initialization process, in terms of how many agents were not assigned fields, and how many fields were left unallocated.
+For the first metric, there was a one-to-one relationship between the percent by which cropland area underestimated and the percent of households left without and farmland (Fig.
+6, left panel), the most extreme examples occurring when MODIS cropland was used to initialize the model in districts 1 and 2, where $\sim$85\% of agents could not be allocated farmland because MODIS underestimated cropland by that amount $\sim$85\%.
+No agents were left without fields where maps overestimated cropland extent (GeoWiki, SA-LC), but in these cases the percent of cropland left unallocated matched the size of the overestimate (e.g.
+$\sim$20\% for SA-LC, Fig.
+6 right panel).
+Interestingly, the overall relationship between land allocation and cropland bias was U-shaped, as the model also failed to give land to agents when cropland was underestimated by 50\% or more (Fig.
+6, right panel), as exemplified by the MODIS result for districts 1 and 2, where 7-12\% of cropland was left unallocated despite the fact that only 85\% of agents had no land.
+This curious relationship occurred because cropland tends to cluster, and when it is underestimated, the size of these clusters is small, resulting in islands of cropland that fall outside of the search radius (which is constrained by an absolute distance and the proximity of other agents) within which cropland is sought when agents are seeded onto the landscape.
+
+%randomly siting 100 household agents within each district, and allocating the nearest two cropland pixels to each household.
+The remaining agents are then iteratively assigned unallocated cropland pixels within a 1.5 km radius of existing agents' fields, and this process continues until all agents are assigned cropland, or all available cropland is allocated.
+This initialization process
+
 \begin{figure}[ht]
-\centerline{\includegraphics[width=.3\textwidth]{figures/agent-bias.pdf}}
-\caption{Biases in agent-based model initialization relative to cropland map biases, measured in terms of the percent of households having no cropland allocated (top) and the percent of cropland left unallocated (bottom).
-Dot sizes represent a district number (smallest = district 1; largest = district 4), colors represent the landcover map.
-\textcolor{red}{to fix: add x-axis legend, add district number to legend.}}
+\centerline{\includegraphics[width=.5\textwidth]{figures/agent-bias.pdf}}
+\caption{Biases in agent-based model initialization relative to cropland map biases, measured in terms of the percent of households having no cropland allocated (left) and the percent of cropland left unallocated (right).
+Dot sizes correspond to district numbers, colors represent the landcover map.}
 \label{afoto}
 \end{figure}
-
-\begin{itemize}
-  \item The percent of land unallocated, however, shows a U-shaped relationship with cropland biases, declining to near zero when cropland bias is between zero and up to $\sim$50\% under-estimated, thereafter increasing.
-Thus in cases where cropland bias insufficient farmland for households, a signficant share of available cropland can go unallocated (nearly 15\%).
-\textcolor{red}{\textbf{Tom, Peng}, I assume this has to do something with seeding and the spatial pattern of cropland, but not sure exactly what.
-Model not sure which HH to give the land too because there are too many for the amount of cropland available, so it gives up?
-}
-\end{itemize}
-
-%large gradient ($\sim \displaystyle{ \inlinefrac{1}{\delta}}$).
 

### 2015-09-22 Rmd/agent-allocation-bias.Rmd
Updated through Results with AW comments.

@@
 plot(unlist(pctbias), unlist(noland), col = cols, cex = cexs, pch = 20, 
-     ylab = "% HH without farms", xlab = "% cropland bias", cex.lab = cx[2], 
-     cex.axis = cx[1])
+     ylab = "% HH without farms", xlab = "% error in cropland area", 
+     cex.lab = cx[2], cex.axis = cx[1])
 plot(unlist(pctbias), unlist(unalloc), col = cols, cex = cexs, pch = 20, 
-     ylab = "% land unallocated", xlab = "% cropland bias", cex.lab = cx[2], 
-     cex.axis = cx[1])
+     ylab = "% land unallocated", xlab = "% error in cropland area", 
+     cex.lab = cx[2], cex.axis = cx[1])
 par(xpd = NA)

### 2015-09-22 Rmd/magisterial-district-bias.Rmd
Updated through Results with AW comments.

@@
 axis(2, at = 0:4, round(exp(c(-10, 1:4))))
-for(i in 1:length(gam_mods)) {
+# for(i in 1:length(gam_mods)) {
+for(i in 1:4) {
   m <- gam_mods[[i]]
@@
 }
-legend("topleft", c(mcap, "All 4", "No GlobCover"), lty = 1, lwd = 1.5,
+# legend("topleft", c(mcap, "All 4", "No GlobCover"), lty = 1, lwd = 1.5,
+#        col = cols, bty = "n", xpd = NA)#, title = "Fit")
+legend("topleft", mcap, lty = 1, lwd = 1.5,
        col = cols, bty = "n", xpd = NA)#, title = "Fit")
@@
 axis(2, at = 0:4, round(exp(c(-10, 1:4))))
-for(i in 1:length(gam_mods)) {
+#for(i in 1:length(gam_mods)) {
+for(i in 1:4) {
   m <- gam_mods[[i]]

### 2015-09-22 paper/main.tex
Updated through Results with AW comments.

@@
 %\usepackage{cite}
-\usepackage[numbers,round,compress]{natbib}
+\usepackage[numbers,round,sort,compress]{natbib}
 \bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
@@
 \usepackage{placeins}
+\usepackage{float}
 
@@
 
-\title{Quantifying the impacts of bias in landcover data on global change analyses}
+\title{The Impact of Bias in Landcover Maps}
+%\title{Our Understanding of Global Change is Built on a Shaky Foundation}
 
@@
 Justin Sheffield\affil{1}{Princeton University, Princeton, NJ USA}
+Adam Wolf\affil{1}{Princeton University, Princeton, NJ USA}
 \and
@@
 
-\dropcap{T}he nature and distribution of landcover is a fundamental determinant of many environmental and social processes that drive or are affected by global change \cite{lambin_modelling_1997}, such as agricultural production and food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010,gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015,luoto_predicting_2004}, or demographic changes \cite{linard_assessing_2010}.
-Landcover maps are therefore critical for understanding the nature and impact of such changes \cite{see_improved_2015}, and they need to be accurate at the finest scales at which the underlying processes operate.
-For example, agricultural productivity and nutrient loadings can vary greatly between neighboring fields, and field sizes are often $<$2 hectares in regions where smallholder farming still dominates \cite{jain_mapping_2013, debats_generalized_????}.
-To understand agriculturally driven processes, it is thus necessary to accurately delineate fields at their smallest grain size, and to do so at regional to global scales to have a consistent set of maps.
+\dropcap{T}he nature and distribution of landcover is an indicator that gives significant insight into socio-econonomic processes \cite{lambin_modelling_1997}, because human endeavors are so closely tied to how we transform land, whether it be the felling of ancient forests for farmland or erecting skycrapers.
+The vastness of our alteration of Earth’s landscapes suggests that landcover is a fundamental mediator of many environmental and social processes that drive or are affected by global change \cite{lambin_modelling_1997}, such as agricultural production and food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010,gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015,luoto_predicting_2004}, and changes in human demography \cite{linard_assessing_2010}.
+Like any view into nature, resolution and fidelity at fine scales are the keys to unlocking more granular and mechanistic insights into these processes \cite{see_improved_2015}.
+It is therefore unsurprising to see the explosive growth in private sector initiatives to develop new Earth observing capabilities, which range from small hobbyist drones\footnote{e.g.
+3DRobotics, DJIA} to satellite arrays\footnote{Planet Labs, Skybox}, in order to add value to industries such as agriculture, mining, and construction.
+This rapid growth in fine-scale landcover mapping capability is creating new opportunities to develop actionable information for traditionally public-sector concerns, such as agricultural development\footnote{USAID's Feed the Future}, drought and flood adaptation\footnote{Global Index Insurance Facility, www.indexinsuranceforum.org}, and carbon cycle management\footnote{United Nations REDD+, www.un-redd.org/aboutredd}.
+But while the demand for more nuanced, landcover-based insights is growing, there is only now the opportunity to use finer-scaled imagery to comprehensively interrogate the accuracy and biases in the landcover products that have become ubiquitous in global change research.
 
-Landcover data can only be developed with satellite imaging, but often the average size class of the cover type of interest is smaller than the sensor resolution, or spectrally indistinct from other neighboring covers, which propagates classification error \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}.
-The result is that landcover datasets are generally inaccurate at finer scales and greatly differ between one another, particularly in those parts of the world undergoing the most rapid land use changes, where the aforementioned sources of bias tend to be most pronounced \cite{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011}.
+Global landcover data can only practically be derived from satellite imaging, but in many regions the average size class of the cover type of interest is smaller than the sensor resolution, or spectrally indistinct from other neighboring covers, which propagates classification error \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}.
+The result is that landcover maps are generally inaccurate at finer scales and disagree greatly, particularly in those parts of the world undergoing the most rapid land use changes, where these sources of bias tend to be most pronounced \cite{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011}.
 
-These errors are well-known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of efforts underway to improve landcover maps, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}.
-What is less known is the degree to which these errors bias measurements built upon the distributional and areal information in landcover.
-An impediment to this understanding is that the errors are hard to quantify because spatially extensive reference data are not available for most regions of the world--particularly over Africa and other developing regions.
-Errors assessment therefore typically rely on a small number of ground truth points or survey data aggregated to political boundaries.
-For this reason, we have a better understanding of the biases between landcover datasets or in relation to country-level statistics \cite{fritz_comparison_2010,fritz_cropland_2011,kaptue_tchuente_comparison_2011} than we do of how error changes over spatial gradients or as a function of aggregation scale.
+Errors in landcover products are widely-acknowledged \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of efforts underway to improve landcover maps, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}.
+What is less known is the degree to which these errors bias analyses derived from the distributional and areal information in landcover.
+Errors are hard to quantify because spatially extensive reference data are not available for most regions of the world--particularly over Africa and other developing regions.
+Error assessments therefore typically rely on a small number of ground truth points for a bottom-up assessment or aggregated survey data for a top-down sanity check.
+For this reason, we have a better understanding of discrepancies between landcover datasets in relation to country-level statistics \cite{fritz_comparison_2010,fritz_cropland_2011,kaptue_tchuente_comparison_2011}, which offers little direction for how to arrive at a true number.
 
@@
 There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, and carbon flux \cite{quaife_impact_2008} and human population estimates \cite{linard_assessing_2010}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}.
-The exception is \cite{schmit_limitations_2006}, who use a high quality, ground-collected reference map detailing farm land use parcels in central Belgium, but the number of sites and region were both fairly restricted, and the parcels were not spatially contiguous.
+One exception is a study  \cite{schmit_limitations_2006} that used a high quality, ground-collected reference map detailing farm land use parcels in central Belgium, but the number of sites and region were both fairly restricted, and the parcels were not spatially contiguous.
 
+Just as a tall edifice cannot be built on a shaky foundation, global change science needs to be based on sound landcover data.
 There is thus an urgent need to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.
-We address this need in this study, using a unique, high accuracy agricultural landcover map for South Africa to quantify the errors in several latest generation landcover maps that are broadly used in global change studies.
-We use this information to examine how i) landcover properties and related classification schemes influence error, ii) how these errors change with aggregation scale, with the specific goal of determining ``safe'' scales for making area-based calculations, and 3) how these errors propagate through several different forms of downstream analyses that broadly represent the global change research focus areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.
+We address this need in this study, using a unique, high accuracy agricultural landcover map for the entire country of South Africa to quantify the errors in several latest generation landcover maps that are widely used in global change studies.
+We use this information to examine how i) landcover properties and related classification schemes influence error, ii) how these errors change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and iii) how these errors propagate through several different forms of downstream analyses that broadly represent the global change research focus areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.
 
@@
 \section{Study area and landcover data}
-Our study focused on South Africa, which comprises nearly 6\% of sub-Saharan Africa's (SSA) landmass, and has a large, diverse agricultural sector, ranging from large commercial operations to smallholder farms \cite{hardy_rainfed_2011,estes_using_2014}.
+South Africa comprises nearly 6\% of sub-Saharan Africa's (SSA) landmass, and has a large, diverse agricultural sector, ranging from large commercial operations to smallholder farms \cite{hardy_rainfed_2011,estes_using_2014}.
 This diversity suggests that the country's agricultural landcover spans the range of types that are found throughout the rest of SSA.
 
-The South African government commissioned a whole country cropland boundary map in order to stratifying the annual aerial crop type census used to calculate harvested area estimates \cite{fourie_better_2009}.
-The map was made by trained workers who visually interpreted high resolution satellite imagery and manually digitized field boundaries following a standardized mapping protocol.
-The resulting vectorized field maps, which were made in 2007 and updated in 2011, provide a unique, high accuracy reference dataset of both crop field distribution and size classes.
-We converted the vector data into a rasterized estimated of cropland percentage at 1 km resolution (henceforth the ``reference map''), which was 97\% accurate in distinguishing cropped from non-cropped areas.
+The South African government commissioned a whole country cropland boundary map to enhance its annual collection of agricultural statistics \cite{fourie_better_2009}.
+The map was made by trained workers who visually interpreted high resolution satellite imagery ($<$5 m SPOT imagery) and manually digitized field boundaries following a standardized mapping protocol.
+The resulting vectorized field maps, which were made in 2007 and updated in 2011, provide a unique, high quality reference dataset describing crop field distributions and size classes, and are 97\% accurate in distinguishing cropland from non-cropland at 200 m resolution.
+We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create a gridded cropland reference map.
 
-We compared our reference percent cropland estimates to those created from four satellite-derived landcover datasets.
+We compared the reference map with similar maps derived from four existing landcover datasets.
 We obtained South Africa's 30 m resolution National Landcover map (SA-LC) for 2009 \cite{sanbi_national_2009}, the 500 m resolution MODIS Landcover for 2011 \cite{land_processes_distributed_active_archive_center_lp_daac_modis_2011, friedl_modis_2010}, the 300 m resolution GlobCover 2009 \cite{arino_global_2012}, and the new 1 km Geo-wiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}.
-We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite{fry_completion_2009},  MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013,shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and is the current state of the art for agricultural landcover maps.
-We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates (hereafter simply the ``landcover maps''), resulting in 4 maps to compare to our reference cropland map (the ``reference map'').
+We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite[e.g.][]{fry_completion_2009}, MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013,shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
+We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates (hereafter simply the ``cropland maps''), resulting in 4 maps to compare to our reference map.
 
 \vspace{-0.5 cm}
-\section{Quantifying Error}
-We used these maps to first quantify error in cropland area estimates.
-We calculated error as the difference between the reference and landcover maps at different scales of aggregation (1 to 100 km), in order to estimate bias and how it varies with scale.
-Next, we assessed how bias correlates with the amount of cropland cover in agricultural landscapes, to gain insight into how landscape patterns may affect error.
+\section{Quantifying Error and Bias}
+We first quantified the errors in cropland area estimates based on the pixel-wise differences between the reference map and each of the four cropland maps.
+We calculated these errors for five different scales of aggregation, from the original 1 km up to 100 km, in order to calculate bias--the average pixel-wise error--and how it varies with scale.
+Next, we assessed the degree to which the pattern of agricultural landscapes, expressed in terms of the mean cropland cover, impacts bias.
 
-We undertook five further analyses to investigate how map error can impact assessments founded on landcover maps.
-These include first-order analyses, in which values for a variable of interest are mapped to particular cover type(s), and second-order analyses, in which a process model draws on the cover types' values to calculate an output value.
-We created four datasets to represent first order analyses.
-The first was a series of maps of vegetated carbon stocks created following the methodology of Ruesch and Gibbs' \cite{ruesch_new_2008}.
-The second was constrained cropland percentage maps, which, following Ramankutty et al \cite{ramankutty_farming_2008} were adjusted so that their total cropland areas matched provincial-level reported cropland totals.
-Using these adjusted cropland percentage maps, we disaggregated district-reported maize harvested area and yields \cite[following][]{monfreda_farming_2008} .
-We then compared differences between total carbon stock estimates calculated from the reference map with those from the four cropland maps, and again examined how these differences changed as a function of aggregation scale.
-We made the same comparisons for total maize harvested area, average yield, and total production.
+We then investigated how these landcover map biases propagate through data that are built upon them.
+We began by examining the biases in three different kinds of landcover-based data that we consider to be ``first-order'', in that they are derived by mapping a variable of interest to a landcover type(s).
+The first of these was to assess potential biases in Tier-1 carbon accounting maps that we constructed from the cropland maps, following \cite{ruesch_new_2008}.
+The second was an assessment of gridded maize yields derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}.
+We used these same data to calculate gridded maize production estimates, which we used to conduct the third first order analysis.
 
-For the second-order analyses, we examined how cropland cover errors influence 25 km resolution monthly evapotranspiration estimates produced using the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model.
-For this example, we used the cropland maps to adjust the seasonally varying, landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
+Finally, we conducted two second-order analyses, in which a process model draws on the cover types' values to calculate an output value.
+For the first of these, we examined how cropland map biases influence 25 km resolution monthly evapotranspiration calculated by the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model.
+For this example, we used the cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
 In the second example, we examined how these errors can impact the parameterization of an agent-based food security model \cite{chen_dependency_2013}.
 Spatially-explicit, agent-based models are frequently employed in land change science, and require an initialization step to assign landscape resources to model agents  \cite[e.g.][]{manson_agent-based_2007,evans_multi-scale_2004,kelley_relative_2011}.
-In this case, we used the cropland maps to allocate farmland to agents representing individual households in political districts, with the model assigning each household its initial cropland holdings using a function that considers total district cropland and how much cropland is near the agent's location.
+In this case, we used cropland maps to provide the model the location and abundance of cropland, which is used to  allocate an initial share of cropland to each farm household (agent) in the model.
 
@@
 
-We created the 1 km reference and landcover maps, removing all croplands marked as communal or smallholder farmland in the reference vector maps (individual fields were not mapped with the same precision), as horticulture, or plantation forestry (SI), and calculated the total cropland extent in the remaining area (1,081,000 km$^2$, or 90\% of South Africa).
+We created the 1 km reference map after removing all field types classified as communal/smallholder agriculture (individual fields in this category were not mapped, thus they were removed to prevent potential cropland area overestimates) or permanent tree-crops (SI), and calculated the total cropland extent in the remaining area (1,081,000 km$^2$, or 90\% of South Africa).
 The 2011 reference map showed a cropland area of 104,310 km$^2$, which the SA-LC and GeoWiki maps overestimated by 26 and 5.8\%, respectively, and GlobCover and MODIS underestimated by 21 and 26.1\%.
 
-We then aggregated each map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from the reference map at each scale of aggregation to asses the spatial patterns of bias (Fig.
+We then aggregated the reference and each cropland map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from the reference map at each scale of aggregation to assess patterns of differences (Fig.
 1).
+Negative pixels here represent overestimation error, while positive values indicate underestimates.
 
@@
 
-The most pronounced extreme biases are in the MODIS and GlobCover maps, with both underestimating cropland extent by 10-75\% in the center of the country (blue areas in Fig.
-1, and the dominant production region), and overestimating along the eastern to northern margins (red areas in Fig.
+The most pronounced differences were in the MODIS and GlobCover maps, which both underestimated cropland extent by 10-75\% in the center of the country (blue areas in Fig.
+1, and the dominant production region), and overestimated along the eastern to northern margins (red areas in Fig.
 1).
-The average biases for GlobCover and MODIS are 21\% and 34\% at 1 km, meaning that each map underestimates cropland by that amount at this resolution.
-MODIS mean bias drops to 8\% at 50 km of aggregation, whereas GlobCover bias is still 24\% at 100 km (SI Appendix, Fig.
+The differences between the reference and these two maps at 1 km resolution were respectively 21\% (MODIS) and 34\% (GlobCover), on average.
+This mean pixel-wise error expresses the amount of \emph{bias} in a map, that is its tendency to err in a particular direction.
+For MODIS, bias drops to 8\% at 50 km of aggregation, whereas GlobCover bias is 24\% at 100 km (SI Appendix, Fig.
 S1).
 
-The SA-LC map uniformly overestimates cropland throughout the country (Fig.
-1), but the magnitude of mean bias is much lower at -8\% at 1 km to -6\% at 100 km  (Fig.
+The SA-LC map uniformly overestimated cropland throughout the country (Fig.
+1), but was less biased, ranging from -8\% at 1 km to -6\% at 100 km  (Fig.
 S1).
-The GeoWiki map has a very heterogeneous pattern of bias, having a slight underestimation tendency (5\% mean bias) at 1 km, which changes to a smaller overestimation bias (-2 \%) at 100 km (Fig.
+The GeoWiki map has a strongly heterogenous pattern of differences (Fig.
+1) and the smallest magnitude of bias, which changed between slight tendencies to underestimate (5\% at 1 km) and overestimate (-2 \% at 100 km,  Fig.
 S1).
@@
 \subsection{The role of landscape characteristics in shaping bias}
-To investigate the degree to which landscape features influence landcover map bias, we extract all pixels in agricultural areas ($>$0.5\% cropland) of the 1 km reference map using the the boundaries of 354 magisterial districts (South Africa's finest administrative unit, which average 3,445 km$^2$ in size; SI Appendix, Fig.
-S3), and averaged the values of these pixels within each district.
-These provided an estimate of the intensity of cropping within the farmed parts of the landscape (with districts defining the landscape scale), as well as the degree of mixing between cropland and other land covers.
-We then extracted the cropland map biases for the same locations, and calculated the mean absolute bias within each district.
+To investigate the degree to which landscape features influence landcover map bias, we extracted all pixels in agricultural areas ($>$0\% cropland) of the 1 km reference map using the boundaries of 354 magisterial districts (South Africa's finest administrative unit, which average 3,445 km$^2$ in size; SI Appendix, Fig.
+S3), and calculated the district-wise mean for these pixels to provide a measure of cropland density, and information on the degree of mixing between cropland and other land covers.
+We then extracted the cropland map errors for the same pixels, and calculated the absolute value of bias for each district.
 Absolute bias describes the magnitude of bias but not its direction, and is more informative about the likelihood of the map being biased at any given point in the landscape than actual bias, as positive and negative biases can cancel each other out within small areas.
@@
 
-\vspace{-1 cm}
+%\vspace{-1 cm}
 \begin{figure}[ht]
@@
 \subsection{The impact of bias on calculating carbon stocks}
-We used the data of \cite{ruesch_new_2008} to calculate carbon densities for African forests, shrublands, croplands, grasslands, and sparse habitats.
-We assigned the cropland carbon values to map cells in proportion to their cropland cover.
+We used the data of \cite{ruesch_new_2008} to calculate carbon densities for African forests, shrublands, croplands, grasslands, and sparse habitats (semi-arid grasslands and low shrublands), assigning cropland carbon values to map cells in proportion to their cropland cover.
 For the non-cropland proportions, we assigned the carbon value from each of the other types, such that we created four different carbon maps for each landcover map at each aggregation scale (Fig.
-S4), which allowed us to test how carbon estimates vary as a function of i) bias in the cropland estimates and ii) the characteristics of adjacent cover types.
+S4), which allowed us to test how carbon estimates vary as a function of i) cropland map bias and ii) the characteristics of adjacent cover types.
 
-At the 1 km scale, the differences between country-wide total carbon stocks made using any of the cropland maps and the reference were within +/-3\%, regardless of which cover type was adjacent to cropland (Table S1), because differences between total stock estimates are diluted by the large area ($\sim$50-70\%) of the country having no cropland.
-Comparing the carbon stocks from each map's agricultural area reveals much greater differences (Table S1).
-SA-LC overestimates carbon stocks by just 2\% when the adjacent cover type is forest, and up to 15\% when it is spare cover.
-MODIS ranges from negligible overestimates in denser carbon classes (forest, secondary forest, and shrublands) to 8-13\% underestimates for the grassland and sparse classes.
-GeoWiki underestimates for all types, from $<$1\% for sparse cover to 8\% for forest.
-GlobCover shows the largest total carbon stock bias, with over-estimates ranging from for sparse lands to 162\% for forest.
-The magnitude of this bias is due to false positives--GlobCover identified cropland in nearly 50\% of pixels, compared to 30\% for the other three cropland maps.
+The difference between total carbon stocks for the country made using any of the cropland maps were within +/-3\% of those based on the reference map, regardless of which cover type was adjacent to cropland (Table S1), because the large of non-cropland in the country ($\sim$50-70\%, Fig.
+1) dilutes any map errors.
+Comparing total stocks between maps for just the agricultural area (30-50\% of the country) reveals much greater differences (Table S1).
+SA-LC overestimated carbon stocks by just 2\% when the adjacent cover type was forest, and up to 15\% when it was sparse cover.
+MODIS ranged from negligible differences in denser carbon classes (forest, secondary forest, and shrublands) to 8-13\% underestimates for the grassland and sparse classes.
+GeoWiki underestimated for all types, from $<$1\% for sparse cover to 8\% for forest.
+GlobCover grossly overestimated total carbon stocks for agricultural areas, varying from 64\% for sparse lands to 162\% for forest.
+The magnitude of this bias was due to false positives--GlobCover identified cropland in nearly 50\% of pixels, compared to 30\% for the other three cropland maps.
 
-The spatial patterns of carbon bias (Fig.
-S4) are a reflection of the cropland biases (Fig.
+The spatial patterns of errors in carbon estimates (Fig.
+S4) reflect those of cropland biases (Fig.
 1).
-Where cropland is underestimated and the surrounding cover type is of higher carbon density than cropland, carbon density is overestimated.
-For lower density (grassland and sparse vegetation), there may be slight underestimates.
-These tendencies are reflected in the average of each map's spatial biases over agricultural areas (Fig.
+Where cropland was underestimated and the surrounding cover type was of higher carbon density than cropland, carbon density was overestimated.
+For lower density cover (grassland and sparse vegetation), carbon stocks were underestimated, but by small magnitudes.
+These tendencies were reflected in each map's biases, calculated over agricultural areas (Fig.
 3).
-For example, MODIS and GlobCover overestimate carbon by $\sim$50\% on average at 1 km resolution when forest is the cropland-adjacent cover (stars in semi-transparent green and gold bars, Fig.
+For example, MODIS and GlobCover bias was $\sim$-50\% (overestimation) at 1 km resolution when forest was the cropland-adjacent cover (stars in semi-transparent green and gold bars, Fig.
 3; Table S2).
-When it is sparse vegetation (open circles in Fig.
-3), MODIS underestimates by nearly 3\% at all scales, whereas maps that overestimate cropland (e.g.
-SA-LC, semi-transparent red) overestimate carbon density for this cover.
-Overall, GeoWiki has the lowest average spatial biases, for all cover types and all resolutions.
-It's maximum bias is a 12\% overestimate at 1 km for forest cover, thereafter all biases are less than 10\%, and converge to +/-2\% above 1 km aggregation (Fig.
+For sparse vegetation (open circles in Fig.
+3), MODIS bias was 3\% at all scales, whereas maps that overestimated cropland (e.g.
+SA-LC, semi-transparent red) overestimated carbon density for this cover type, because cropland has a higher carbon density \cite{ruesch_new_2008}.
+Overall, GeoWiki had the lowest bias, for all cover types and all resolutions.
+It worst bias was a tendency to overestimate by 12\% at 1 km when forest was adjacent, but at coarser scales this bias reduced to just a few percent (Fig.
 3, Table S2).
@@
 
-\vspace{-0.9 cm}
-\begin{figure}[ht]
+%\vspace{-0.9 cm}
+\begin{figure}[!ht]
 \centerline{\includegraphics[width=.4\textwidth]{figures/carbon_veg_scale3.pdf}}
-\caption{The average spatial biases in carbon density estimates based on cropland maps, expressed as a percent difference relative to the reference map.
+\caption{The biases in carbon density estimates based on cropland maps, expressed as a percent difference relative to the reference map.
 The values of actual bias (incorporating bias direction and magnitude) and absolute bias (magnitude only) are presented in semi-transparent and solid colors, respectively, with colors denoting the specific cropland map and symbols indicating which cover type was used to calculate cropland adjacent carbon density.
@@
 
-The patterns of mean absolute spatial bias (solid colored bars in Fig.
-3) generally follow that of actual bias, but with somewhat higher magnitudes and a few important differences.
-The most notable is that Geo-Wiki has fairly large absolute bias at 1 km, averaging 27\% across across cover types (line in solid blue bar, Fig.
+Absolute bias in carbon maps (solid colored bars in Fig.
+3) generally followed the same patterns, but with somewhat higher magnitudes and a few important differences.
+The most notable is that Geo-Wiki had fairly large absolute bias at 1 km, averaging 27\% across across cover types (line in solid blue bar, Fig.
 3, Table S2), which is close to the 36-37\% for GlobCover and MODIS.
-SA-LC has the lowest absolute bias across scales, averaging (across cover types) from 14\% at 1 km to 3\% at 100 km.
+SA-LC had the lowest absolute bias across scales, averaging (across cover types) from 14\% at 1 km to 3\% at 100 km.
 (Fig.
 3, Table S2).
-The increase in Geo-Wiki's absolute spatial bias relative to SA-LC's can be attributed to the spatially heterogeneous nature of its cropland biases (Fig.
-1) compared to the other three cropland maps.
+The increase in Geo-Wiki's absolute bias relative to SA-LC's can be attributed to the highly heterogeneous nature of its cropland errors (Fig.
+1), which varied between positive and negative over much shorter distances than the other three cropland maps.
 
 \subsection{Bias in harvested areas, yield, and production estimates}
-We used the reference dataset to calculate total cropland area within South Africa's nine provinces, and then adjusted the the four cropland maps so that they each summed to the same area within each province \cite[per][]{ramankutty_farming_2008}.
-Despite this statistical constraint, the updated cropland maps still had substantial spatial biases similar in pattern (Fig.
+The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived agricultural area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}.
+To be consistent with this methodology, we adjusted our cropland maps using the same procedure, but we used the reference map to calculate total cropland area for each of South Africa's nine provinces, and then updated the pixel-wise cropland percentages in the four cropland maps so that the province-wise sums matched the reference areas \cite[][and see SI]{ramankutty_farming_2008}.
+Despite this statistical constraint, the updated cropland maps still had large spatial errors that were similar in pattern (Fig.
 S5) to those in the unadjusted maps (Fig.
-1), and we evaluated how these biases affected crop production estimates.
-We derived maize yield and harvested area estimates by disaggregating district-level  (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} following Monfreda \emph{et al.} \cite{monfreda_farming_2008}, and used that to calculate spatial biases in yield and production estimates at each aggregation scale.
+1), and we evaluated how these residuals affected gridded estimates of the yield and production of maize, South Africa's largest crop \cite{estes_comparing_2013}.
+To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.
 
@@
 S6).
-However, the mean spatial bias in yields at this resolution was pronounced only for GlobCover, equivalent to nearly 60\% of the mean reference yield of 3.4 tons ha$^{-1}$, compared to +/-5\% for the other maps (Fig.
+However, only GlobCover showed a notable bias in yields at this resolution, which was equivalent to nearly 60\% of the mean reference yield of 3.4 tons ha$^{-1}$.
+All other maps had biases of just +/-5\% at 1 km (Fig.
 5).
-Interestingly, GeoWiki and MODIS biases increased with aggregation, peaking at 10 km where both datasets underestimated yields by 30\%, thereafter declining to 10\% at 100 km, whereas GlobCover's mean yield bias declined linearly with aggregation (Fig.
+Interestingly, GeoWiki and MODIS biases increased with aggregation, peaking at 10 km where both had underestimation biases of 30\%, thereafter declining to 10\% at 100 km.
+In contrast, GlobCover's yield bias declined linearly with aggregation (Fig.
 5).
@@
 %\vspace{-1 cm}
-\begin{figure}[!ht]
+\begin{figure}[ht]
 \centerline{\includegraphics[width=.4\textwidth]{figures/yield_bias_scale.pdf}}
-\caption{The mean spatial biases in disaggregated maize yield estimates, and the bias in total crop production estimates calculated from these yields and constrained crop area estimates.
+\caption{The biases in disaggregated maize yield and production estimates.
 The values of actual bias (incorporating bias direction and magnitude) and absolute bias (magnitude only) are presented in semi-transparent and solid colors, respectively, with colors denoting the specific cropland map and symbols indicating whether the bias is related to one of two yield aggregation methods and associated crop production estimates.}
@@
 
-In contrast to yields, production errors were unbiased in aggregate (Fig.
+Production errors were completely unbiased (Fig.
 5).
-The statistical constraints on harvested and cropland areas led to the canceling out of spatial errors in production estimates, which is evident in the checkerboard-like pattern in maps of production biases (Figure S7).
-However, this pattern of bias means that the average absolute biases in  production estimates are large, between 40 to 55\% for GeoWiki, MODIS, and GlobCover at 1 km, and remaining generally high ($<$10\%) for these three datasets even up to 50-100 km of aggregation (Fig.
+The statistical constraints on harvested and cropland areas resulted in the canceling out of spatial errors in production estimates, which is evident in the checkerboard-like pattern in maps of production biases (Figure S7).
+However, this pattern of bias means that absolute bias in production estimates were large, between 40 to 55\% for GeoWiki, MODIS, and GlobCover at 1 km, and remained generally high (10-28\%) even up to 50 km of aggregation (Fig.
 5).
@@
 Compared to carbon and crop related examples, actual and absolute biases in evapotranspiration calculated using the VIC model were small and averaged to less than +/-1\%.
-There are several hotspots of discrepancy, however, which are evident in maps of spatial biases (Fig.
+However, there are several hotspots of discrepancy evident in the error maps (Fig.
 6).
-The most pronounced of these are the 5-15\% overestimates by VIC initialized with MODIS and GlobCover in the center of the country, and along the southern and western coasts, particularly the latter where the ET overestimates reach 25\%.
-These locations correspond primarily to margins of major crop production regions--in the center is the westernmost boundary of the summer rainfall growing region, where maize is the primary crop, while the bias hotspot up the west coast marks one of the two boundaries of the winter rainfall region where wheat farming dominates\cite{hardy_rainfed_2011}.
-These two boundaries correspond respectively to approximately 400 and 200 mm of growing season rainfall.
+The most pronounced of these are the 5-15\% overestimates in the center of the country that resulted when VIC was initialized with MODIS and GlobCover, while overestimates along the southern and western coasts reached 25\%.
+These locations correspond primarily to the margins of major crop production regions--in the center is the westernmost boundary of the summer rainfall growing region, where maize is the primary crop, while the hotspot on the west coast marks the western boundary of the wheat-dominated winter rainfall region \cite{hardy_rainfed_2011}.
+These two boundaries respectively correspond to the 400 and 200 mm isohyets of growing season rainfall.
 
@@
 \FloatBarrier
-\begin{figure}[!ht]
+\begin{figure}[ht]
 \centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
@@
 \subsection{Initialization biases in agent-based models}
-We used an agent-based model of food security that represents the interactions between hundreds of individual farming households over multiple seasons \cite{chen_dependency_2013}.
-The model is thus computationally intensive, and, like many spatial ABMs, run over smaller geographic domains (e.g.
+We used an agent-based model (ABM) of food security that represents the interactions between hundreds of individual farming households over multiple seasons \cite{chen_dependency_2013}.
+Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g.
 districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) that are needed to represent the different land units of single farmers.
@@
 We then initialized the model by assigning each household agent two cropland pixels.
-In order to emulate the natural groupings of communities, the model only assigns a household fields that are within 1.5 km of other agents' fields, and if those pixels have not already been allocated to another agent.
+In order to emulate the natural groupings of communities, the model only assigns a household fields that are within 1.5 km of other agents' fields, provided those pixels were not previously allocated to another agent.
 The model thus iteratively grows ``communities'' until all households are assigned cropland, or all available cropland is allocated.
 
-We used each of the five maps to initialize the model, and assessed how bias in cropland cover estimates impacted the initialization process, in terms of how many agents were not assigned fields, and how many fields were left unallocated.
-For the first metric, there was a one-to-one relationship between the percent by which cropland area underestimated and the percent of households left without and farmland (Fig.
-6, left panel), the most extreme examples occurring when MODIS cropland was used to initialize the model in districts 1 and 2, where $\sim$85\% of agents could not be allocated farmland because MODIS underestimated cropland by that amount $\sim$85\%.
-No agents were left without fields where maps overestimated cropland extent (GeoWiki, SA-LC), but in these cases the percent of cropland left unallocated matched the size of the overestimate (e.g.
+We used the reference map and each cropland map to separately initialize the model, and compared the agent allocation results to assess how cropland map errors impacted the initialization process.
+[insert importance here].
+We examined two metrics, the first being the number of agents that were not assigned fields.
+Here there was a one-to-one relationship between the percentage of cropland area underestimation and the percentage of households left without farmland (Fig.
+6, left panel).
+The most extreme examples occurred when MODIS cropland initialized the ABM in districts 1 and 2, where $\sim$85\% of agents did not receive cropland.
+
+All households were assigned fields when maps overestimated cropland extent (GeoWiki, SA-LC), but in these cases the percent of cropland left unallocated--our second metric for assessing cropland error impacts--matched the size of the overestimate (e.g.
 $\sim$20\% for SA-LC, Fig.
 6 right panel).
-Interestingly, the overall relationship between land allocation and cropland bias was U-shaped, as the model also failed to give land to agents when cropland was underestimated by 50\% or more (Fig.
-6, right panel), as exemplified by the MODIS result for districts 1 and 2, where 7-12\% of cropland was left unallocated despite the fact that only 85\% of agents had no land.
+Interestingly, the overall relationship between the percent of cropland allocated and percent cropland error was U-shaped, as the model also failed to give land to households when cropland was underestimated by more than 50\% (Fig.
+6, right panel).
+MODIS again provided the most pronounced results in districts 1 and 2, where 7-12\% of cropland was left unallocated despite the fact that 85\% of agents had no land.
 This curious relationship occurred because cropland tends to cluster, and when it is underestimated, the size of these clusters is small, resulting in islands of cropland that fall outside of the search radius (which is constrained by an absolute distance and the proximity of other agents) within which cropland is sought when agents are seeded onto the landscape.
@@
 \centerline{\includegraphics[width=.5\textwidth]{figures/agent-bias.pdf}}
-\caption{Biases in agent-based model initialization relative to cropland map biases, measured in terms of the percent of households having no cropland allocated (left) and the percent of cropland left unallocated (right).
+\caption{Biases in agent-based model initialization relative to the district-wise errors (as a percent) in total cropland area, measured in terms of the percent of households having no cropland allocated (left), and the percent of cropland left unallocated (right).
 Dot sizes correspond to district numbers, colors represent the landcover map.}
@@
 \section{Discussion}
+
+\begin{itemize}
+  \item First large area quantification of spatial biases
+  \item How large those biases are, for one of the most widely spread (spreading landcovers)
+  \item Insight into causes of bias, and thus some understanding of where biases are likely to be greater or smaller
+  \item How much progress made in reducing it
+  \item Class type and bias
+\end{itemize}
 

### 2015-09-24 Rmd/carbon-bias.Rmd
Updated through results: changed terminology for absolute errors, updated all figures and Rmds creating them accordingly.

@@
 totc_out <- cbind.data.frame(knms, round(unname(totc_out), 2))
-colnames(totc_out) <- c("map", pnms)
-totc_out <- cbind(region = c(rep("Country", 4), rep("Agricultural", 4)), 
+colnames(totc_out) <- c("Map", pnms)
+totc_out <- cbind(Region = c(rep("Country", 4), rep("Agricultural", 4)), 
                   totc_out)
@@
 mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 1.5)
-mtext(side = 2, text = "% bias", outer = TRUE, line = 2)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 2)
 par(xpd = NA)
@@
 y <- grconvertY(0.075, from = "ndc", to = "user")
-legend(x = x, y = y, legend = c("actual bias", "absolute bias"), lty = c(2, 1), 
+legend(x = x, y = y, legend = c("Bias", "MAE"), lty = c(2, 1), 
        lwd = 2, bty = "n", cex = 1.2)
@@
 
-pdf(full_path(p_fig, "carbon_veg_scale3.pdf"), height = 7, width = 7)
+pdf(full_path(p_fig, "carbon_veg_scale.pdf"), height = 7, width = 7)
 par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
@@
 mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
-mtext(side = 2, text = "% bias", outer = TRUE, line = 1)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
 legend(x = 12.4, y = -40, legend = lcnms, pch = 15, col = cols, adj = 0,
@@
        cex = 0.8)
-text(x = 12.7, y = -42, labels = "Absolute", srt = 45, adj = c(0, 0), cex = 0.8)
-text(x = 12.2, y = -42, labels = "Actual", srt = 45, adj = c(0, 0), cex = 0.8)
+text(x = 12.7, y = -42, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
+text(x = 12.2, y = -42, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
 legend(x = 12.1, y = 65, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
@@
 # mu0a[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mu0s <- copy(rbind(cbind(Bias = "Actual", mu0), 
-                   cbind(Bias = "Absolute", mu0a)))
-for(i in 1:length(LC)) mu0s[V2 == LC[i], V2 := pnms[i]]
-for(i in 1:length(snms)) mu0s[V1 == snms[i], V1 := mcap[i]]
+mu0s <- copy(rbind(cbind(Metric = "Bias", mu0), 
+                   cbind(Metric = "MAE", mu0a)))
 setnames(mu0s, c("V1", "V2", lev_vec), 
          c("Map", "Cover", paste(c(1, fact), "km")))
-caption <- paste("Mean actual and absolute biases for each of the cropland", 
+setkeyv(mu0s, c("Map", "Cover"))
+for(i in 1:length(LC)) mu0s[Cover == LC[i], Cover := pnms[i]]
+setkey(mu0s, "Map")
+for(i in 1:length(snms)) mu0s[Map == snms[i], Map := mcap[i]]
+
+caption <- paste("Biases and mean absolute errors for each of the cropland", 
                  "maps across aggregation scales and each possible landcover", 
@@
                  "across the entire country.")
-mu0s_xtab <- xtable(mu0s, digits = 1, caption = caption)
+mu0s_xtab <- xtable(mu0s[order(Metric, Cover)], digits = 1, caption = caption)
 print(mu0s_xtab, type = "latex", file = "paper/figures/C-bias-mean-all-SA.tex", 
@@
 # mua[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-mus <- copy(rbind(cbind(Bias = "Actual", mu), 
-                  cbind(Bias = "Absolute", mua)))
-for(i in 1:length(LC)) mus[V2 == LC[i], V2 := pnms[i]]
-for(i in 1:length(snms)) mus[V1 == snms[i], V1 := mcap[i]]
+mus <- copy(rbind(cbind(Metric = "Bias", mu), 
+                  cbind(Metric = "MAE", mua)))
 setnames(mus, c("V1", "V2", lev_vec), 
          c("Map", "Cover", paste(c(1, fact), "km")))
-caption <- paste("Mean actual and absolute biases for each of the cropland", 
+setkeyv(mus, c("Map", "Cover"))
+for(i in 1:length(LC)) mus[Cover == LC[i], Cover := pnms[i]]
+setkey(mus, "Map")
+for(i in 1:length(snms)) mus[Map == snms[i], Map := mcap[i]]
+
+caption <- paste("Biases and mean absolute errors for each of the cropland", 
                  "maps across aggregation scales and each possible landcover", 
-                 "type sharing the pixel with cropland. Means were calculated",
+                 "type sharing the pixel with cropland. Metrics were calculated",
                  "across the union of agricultural areas", 
@@
                  "the cropland map to which it was being compared")
-mus_xtab <- xtable::xtable(mus, digits = 1, caption = caption)
+mus_xtab <- xtable::xtable(mus[order(Metric, Cover)], digits = 1, 
+                           caption = caption)
 print(mus_xtab, type = "latex", 

### 2015-09-24 Rmd/et-bias.Rmd
Updated through results: changed terminology for absolute errors, updated all figures and Rmds creating them accordingly.

@@
 fnm <- "et_bias_map4.pdf"
-legtext <- rep("%", 5)
+legtext <- rep("% Difference", 5)
 cx <- 1.4
@@
             srt = c(270, 0), horiz = TRUE, textside = "bottom", 
-            legend.pos = c(4, 5), leg.adj = list(c(0.1, 0.3), c(-1, -0.5)), 
+            legend.pos = c(4, 4), leg.adj = list(c(0.1, 0.3), c(-0.25, -0.5)), 
             cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
@@
             srt = c(270, 0), horiz = TRUE, textside = "bottom", 
-            legend.pos = c(4, 5), leg.adj = list(c(0.3, 0.3), c(-1, -0.75)), 
+            legend.pos = c(4, 4), leg.adj = list(c(0.3, 0.3), c(-0.25, -0.75)), 
             cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
@@
 
-caption <- paste("Mean actual and absolute biases (as \\%) for",              
+caption <- paste("Biases and mean absolute errors (as \\%) for",              
                  "evapotranspiration variables derived from a 29-year time",

### 2015-09-24 Rmd/landcover-bias-plots.Rmd
Updated through results: changed terminology for absolute errors, updated all figures and Rmds creating them accordingly.

@@
 library(R.utils)
+library(RColorBrewer)
 
@@
 # Plotting colors
-cols <- colorRampPalette(c("red", "grey80", "blue4"))
+# cols <- colorRampPalette(c("red", "grey80", "blue4"))
 brks <- c(-100.1, -75, -50, -25, -10, -5, -1, 1, 5, 10, 25, 50, 75, 100.1)
+cols <- c(rev(brewer.pal(length(brks) / 2, "Reds")[-1]), "grey80", 
+          brewer.pal(length(brks) / 2, "Blues")[-1])
 brklen <- length(brks) - 1
-cols <- cols(brklen)
-cols[c(1, length(cols))] <- c("darkred", "purple3")
+# cols <- cols(brklen)
+# cols[c(1, length(cols))] <- c("darkred", "purple3")
 legtext <- "% Difference"
@@
             shortdims = c(0.05, 0.01), colvec = cols, srt = c(270, 0), 
-            horiz = TRUE, textside = "bottom", legend.pos = c(4, 5), 
+            horiz = TRUE, textside = "bottom", legend.pos = c(4, 7), 
             leg.adj = list(c(0.2, 0.35), c(0.5, -0.5)), cex.val = cx, 
@@
     lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = 3)
-    if(i == 1) mtext(ifelse(bt == 1, "% bias", "% bias (absolute)"), side = 2, 
-                     line = 2.2, cex = 0.8)
+    if(i == 1) mtext(ifelse(bt == 1, "Bias (%)", "Mean absolute error (%)"), 
+                     side = 2, line = 2.2, cex = 0.8)
     if(bt == 1) mtext(lcnms[i], side = 3, line = 0, col = cols[i])
@@
 }
-mtext("% cropland", side = 1, outer = TRUE, line = -1, cex = 0.8)
+mtext("% Cropland", side = 1, outer = TRUE, line = -1, cex = 0.8)
 dev.off()
@@
 # i <- 1; j <- 1
-par(mar = c(0.5, 0, 1, 0.3), oma = c(5, 4, 2, 2))
+par(mar = c(0.5, 0, 1, 0.3), oma = c(5, 5, 2, 2))
 for(i in 1:length(snms)) {
@@
     if(j == 1) axis(2, at = yl, labels = yl, las = 2, mgp = c(1, 0.6, 0))
-    if(j == 1) mtext(lcnms[i], side = 2, line = 2.8, cex = 1, col = cols[i])
+    if(j == 1) mtext(lcnms[i], side = 2, line = 3, cex = 1, col = cols[i])
     if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 1.2)
@@
 }
-mtext("% bias", side = 2, line = 2.1, cex = 0.8, outer = TRUE, adj = 0.5)
-mtext("% cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
+mtext("% Bias/MAE", side = 2, line = 2, cex = 0.8, outer = TRUE, adj = 0.5)
+mtext("% Cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
 par(xpd = NA)
@@
 y <- grconvertY(0.07, from = "ndc", to = "user")
-legend(x = x, y = y, legend = c("actual bias", "absolute bias"), lty = lt, 
+legend(x = x, y = y, legend = c("Bias", "Mean absolute error"), lty = lt, 
        lwd = ylwd, bty = "n", cex = 1.2)

### 2015-09-24 Rmd/magisterial-district-bias.Rmd
Updated through results: changed terminology for absolute errors, updated all figures and Rmds creating them accordingly.

--- 2015-08-27 Rmd/magisterial-district-bias.Rmd
@@
 par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
-plot(xl, yl2, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i",
-     mgp = c(1.5, 0.5, 0), yaxs = "i", yaxt = "n")
+plot(xl, yl2, pch = "", ylab = "Mean Absolute Error (%)", 
+     xlab = "% Cropland", xaxs = "i", mgp = c(1.5, 0.5, 0), yaxs = "i", 
+     yaxt = "n")
 axis(2, at = 0:4, round(exp(c(-10, 1:4))))
-for(i in 1:length(gam_mods)) {
+# for(i in 1:length(gam_mods)) {
+for(i in 1:4) {
   m <- gam_mods[[i]]
@@
 par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2))
-plot(xl, yl2, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i",
-     mgp = c(1.5, 0.5, 0), yaxs = "i", yaxt = "n")
+plot(xl, yl2, pch = "", ylab = "Mean Absolute Error (%)", 
+     xlab = "% Cropland", xaxs = "i", mgp = c(1.5, 0.5, 0), yaxs = "i", 
+     yaxt = "n")
 axis(2, at = 0:4, round(exp(c(-10, 1:4))))
-for(i in 1:length(gam_mods)) {
+#for(i in 1:length(gam_mods)) {
+for(i in 1:4) {
   m <- gam_mods[[i]]

### 2015-09-24 Rmd/yield-bias.Rmd
Updated through results: changed terminology for absolute errors, updated all figures and Rmds creating them accordingly.

@@
 library(xtable)
+library(RColorBrewer)
 p_root <- full_path(proj_root("SAcropland"), "SAcropland")
@@
 # legtext <- c(rep("tons", 3), rep("tons/ha", 2))
-legtext <- rep("%", 5)
+legtext <- rep("% Difference", 5)
 cx <- 1.4
@@
   brks <- brksl[[i]]
-  cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)
-  cols[c(1, length(cols))] <- c("darkred", "purple3")
+  cols <- c(rev(brewer.pal(length(brks) / 2, "Reds")[-1]), "grey80", 
+          brewer.pal(length(brks) / 2, "Blues")[-1])
+  brklen <- length(brks) - 1
+  #cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)
+  #cols[c(1, length(cols))] <- c("darkred", "purple3")
   pdf(full_path(p_fig, fnms[i]), height = 6, width = 7)
@@
   flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
-              legend.vals = brks, longdims = c(0.2, 0.8), 
-              shortdims = c(0.1, 0.01), 
+              legend.vals = brks, longdims = c(0.15, 0.8), 
+              shortdims = c(0.07, 0.01), 
               colvec = cols, #(length(brks) - 1), 
               srt = c(270, 0), horiz = TRUE, textside = "bottom", 
-              legend.pos = c(4, 5), leg.adj = list(c(0.1, 0.3), c(0, -0.5)), 
-              cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
+              legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0, -0.5)),
+              cex.val = cx, textcol = lcol, bordercol = lcol)
   dev.off()
@@
 alph <- c(225, 40)
-#Ccols <- brewer.pal(n = 7, name = "Greys")
-# LC2 <- c(LC[-grep("second|grass", LC)], "all")
-# xax <- seq(0, 7, 6 / (length(yax) - 2)); length(xax)
 x <- c(0, 3, 6, 9, 12, 15)
@@
 xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
-#x + xo[j] + o[k]
 o <- c(0, w)
 xa <- sapply(x, function(x) x + xo)
-# cx <- c(2, 1.25, 1, 1)
 cx <- c(1.25, 1)
@@
        cex = 0.8)
-text(x = 12.7, y = -42, labels = "Absolute", srt = 45, adj = c(0, 0), cex= 0.8)
-text(x = 12.2, y = -42, labels = "Actual", srt = 45, adj = c(0, 0), cex = 0.8)
-# legend(x = 12.1, y = 65, legend = c("Prod I", "Prod II", "Yield 1", "Yield 2"),        pch = pchs, pt.cex = c(2.5, 1.5, 1.5, 1.5), bty = "n", cex = 0.8)
+text(x = 12.7, y = -42, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
+text(x = 12.2, y = -42, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
 legend(x = 12.1, y = 65, legend = c("Production", "Yield"), pch = pchs, 
@@
 vnms <- c(rep("Yield", length(snms)), rep("Production", length(snms)))
-aa <- c("Actual", "Absolute")
+aa <- c("Bias", "MAE")
 out_tab <- do.call(rbind, lapply(1:2, function(x) {
@@
   out <- cbind.data.frame(aa[x], mcap, "Variable" = vnms, unname(out))
-  colnames(out) <- c("Bias", "Map", "Variable", paste(c(1, fact), "km"))
+  colnames(out) <- c("Metric", "Map", "Variable", paste(c(1, fact), "km"))
   out
@@
 
-caption <- paste("Mean actual and absolute biases for yield and", 
-                 "production estimates disaggregated from magisterial", 
-                 "district level farm census data onto adjusted cropland", 
-                 "maps. Biases were calculated as percent differences", 
-                 "relative to the mean value calculated from the reference", 
-                 "map.")    
+caption <- paste("Biases and mean absolute errors (MAE) in disaggregated maize",                  "yield and production (calculated from disaggregated yield",                    "and harvested area estimates)",
+                 "maps. Bias and MAE were normalized to their", 
+                 "respective mean values calculated from reference maps.")    
 out_xtab <- xtable(out_tab, digits = 1, caption = caption)
@@
 
+save(list = ls(), file = full_path(p_data, "yield-bias.rda"))
 ```

### 2015-09-24 paper/app_s1.tex
Updated through results: changed terminology for absolute errors, updated all figures and Rmds creating them accordingly.

@@
 %     \vspace{-3cm}
-      \caption{Mean biases for each of the landcover maps as a function of cropland density (calculated using the 2011 reference maps) and aggregation scales.
+      \caption{Biases and mean absolute errors (MAE) for each of the cropland maps as a function of cropland density (calculated using the 2011 reference maps) and aggregation scales.
 Rows present biases by map product, columns by aggregation scale.
-Dash lines indicate mean actual bias at each level of cropland density, calculated in bins spanning 5\% of density (e.g.
-0-5\% cropland cover, 5-10\%, etc.), while solid lines indicate the mean absolute bias.
-The black numbers in each plot area present the overall means of actual and absolute bias for each sensor-scale combination (actual/absolute).
+Dash lines indicate bias at each level of cropland density, calculated in bins spanning 5\% of density (e.g.
+0-5\% cropland cover, 5-10\%, etc.), while solid lines indicate the mean absolute error.
+The black numbers in each plot area present the overall means of bias/MAE for each sensor-scale combination.
 }
@@
 %     \vspace{-3cm}
-      \caption{Mean biases for each of the landcover maps at 1 km resolution, as a function of cropland density.
-Colored lines (color-coded to map product name) show the mean bias at each level of cropland density, calculated in bins spanning 5\% (e.g.
+      \caption{Biases and mean absolute errors (MAE) for each of the cropland maps at 1 km resolution, as a function of cropland density.
+Colored lines (color-coded to map product name) show the bias/MAE at each level of cropland density, calculated in bins spanning 5\% (e.g.
 0-5\% cropland cover, 5-10\%, etc.).
 Box plots show the variability of bias in each bin (whiskers = 2.5 and 97.5 percentiles, box the inter-quartile, and grey bar in box the median).
-Actual biases are presented in the top row, and absolute biases in the bottom row.}
+Biases are presented in the top row, and MAEs in the bottom row.
+Statistics are calculated from pooled map errors calculated from differences between the 2007 reference map and each cropland map, and the 2011 reference map and each cropland map.}
       \label{fig:default}
@@
 
-%\FloatBarrier
+\FloatBarrier
 \section*{Carbon bias}
@@
 %     \vspace{-3cm}
-      \caption{Spatial patterns of mean bias (across four different possible cover types adjacent to cropland) in carbon stock estimates. }
+      \caption{Spatial patterns of error (averaged across four different possible cover types adjacent to cropland) in carbon stock estimates. }
       \label{fig:default}
@@
      \includegraphics[width = 12cm]{figures/cropland_adj_bias_map2.pdf} 
-      \caption{Biases in cropland maps adjusted using provincial level cropland area estimates.}
+      \caption{Errors in cropland maps adjusted using provincial cropland area statistics.}
       \label{fig:default}
@@
      \includegraphics[width = 12cm]{figures/yld_bias_map_2.pdf} 
-      \caption{Biases in disaggregated maize yield estimates.}
+      \caption{Errors (normalized to the reference-derived country mean) in disaggregated maize yield estimates.}
       \label{fig:default}
@@
      \includegraphics[width = 12cm]{figures/prod_bias_map_2.pdf} 
-      \caption{Biases in production estimates calculated from disaggregated maize yield and harvested area estimates.}
+      \caption{Errors (normalized to reference-derived country mean) production estimates calculated from disaggregated maize yield and harvested area estimates.}
       \label{fig:default}

### 2015-09-24 paper/figures/et-bias.tex
Updated through results: changed terminology for absolute errors, updated all figures and Rmds creating them accordingly.

@@
 % latex table generated in R 3.2.1 by xtable 1.7-4 package
-% Wed Sep  9 13:32:11 2015
+% Thu Sep 24 11:01:34 2015
 \begin{longtable}{llrr}
-\caption{Mean actual and absolute biases (as \%) for evapotranspiration variables derived from a 29-year time series calculated by the VIC model, including the average total ET for the 3 months of the year when ET is highest, the annual mean and the minimum and maximum annual ETs in the time series.} \\
+\caption{Biases and mean absolute errors (as \%) for evapotranspiration variables derived from a 29-year time series calculated by the VIC model, including the average total ET for the 3 months of the year when ET is highest, the annual mean and the minimum and maximum annual ETs in the time series.} \\
   \hline

### 2015-09-24 paper/figures/totC-bias.tex
Updated through results: changed terminology for absolute errors, updated all figures and Rmds creating them accordingly.

% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Wed Sep 23 22:19:49 2015
\begin{table}[ht]
\centering
\caption{Percent differences in total carbon stock estimates calculated from the reference maps and from each of the four cropland maps. Differences are evaluated for total carbon estimates either at the country scale or over just the agricultural regions (cropland $>$0.05\%), using the carbon densities of 5 different cover types to provide the values for the non-agricultural portions of each pixel (cover types indicated by column names).} 
\begin{tabular}{llrrrrr}
  \hline
Region & Map & Forest & Secondary & Shrubland & Grassland & Sparse \\ 
  \hline
Country & SA-LC & 2.6 & 2.5 & 2.5 & 0.1 & -2.1 \\ 
  Country & GlobCover & -2.1 & -2.0 & -2.0 & -0.1 & 1.7 \\ 
  Country & MODIS & -2.6 & -2.5 & -2.5 & -0.1 & 2.1 \\ 
  Country & GeoWiki & 0.6 & 0.5 & 0.5 & 0.0 & -0.5 \\ 
  Agricultural & SA-LC & -2.0 & -2.7 & -2.8 & -10.6 & -14.9 \\ 
  Agricultural & GlobCover & -161.9 & -156.3 & -155.5 & -95.9 & -63.6 \\ 
  Agricultural & MODIS & -1.6 & -0.8 & -0.7 & 8.4 & 13.3 \\ 
  Agricultural & GeoWiki & 7.7 & 7.3 & 7.2 & 2.9 & 0.5 \\ 
   \hline
\end{tabular}
\end{table}

### 2015-09-24 paper/main.tex
Updated through results: changed terminology for absolute errors, updated all figures and Rmds creating them accordingly.

@@
 
-\title{The Impact of Bias in Landcover Maps}
+\title{A Full Accounting of Landcover Map Error and Bias and Their Impacts on Assessments of Global Change}
 %\title{Our Understanding of Global Change is Built on a Shaky Foundation}
@@
 There is thus an urgent need to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.
-We address this need in this study, using a unique, high accuracy agricultural landcover map for the entire country of South Africa to quantify the errors in several latest generation landcover maps that are widely used in global change studies.
-We use this information to examine how i) landcover properties and related classification schemes influence error, ii) how these errors change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and iii) how these errors propagate through several different forms of downstream analyses that broadly represent the global change research focus areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.
+We address this need in this study, using a unique, high accuracy agricultural landcover map for the entire country of South Africa to conduct a spatially comprehensive, bottom-up quantification of error in several latest generation landcover maps that are widely used in global change studies.
+We use these errors to assess the extent of bias in i) landcover data, ii) how  landcover properties influence this bias, iii) how these biases change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and iv) how these biases propagate through several different forms of downstream analyses that broadly represent the global change research focus areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.
 
@@
 We obtained South Africa's 30 m resolution National Landcover map (SA-LC) for 2009 \cite{sanbi_national_2009}, the 500 m resolution MODIS Landcover for 2011 \cite{land_processes_distributed_active_archive_center_lp_daac_modis_2011, friedl_modis_2010}, the 300 m resolution GlobCover 2009 \cite{arino_global_2012}, and the new 1 km Geo-wiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}.
-We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite[e.g.][]{fry_completion_2009}, MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013,shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
+We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite[e.g.][]{fry_completion_2009}, MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013, shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
 We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates (hereafter simply the ``cropland maps''), resulting in 4 maps to compare to our reference map.
@@
 We first quantified the errors in cropland area estimates based on the pixel-wise differences between the reference map and each of the four cropland maps.
-We calculated these errors for five different scales of aggregation, from the original 1 km up to 100 km, in order to calculate bias--the average pixel-wise error--and how it varies with scale.
-Next, we assessed the degree to which the pattern of agricultural landscapes, expressed in terms of the mean cropland cover, impacts bias.
+We calculated these errors for five different scales of aggregation, from the original 1 km up to 100 km, in order to calculate bias (the mean pixel-wise error) and mean absolute error (which measures the average error magnitude), and how these vary with scale.
+Next, we assessed the degree to which the average cropland cover in agricultural landscapes, which relates to their pattern, impacts error magnitude.
 
-We then investigated how these landcover map biases propagate through data that are built upon them.
-We began by examining the biases in three different kinds of landcover-based data that we consider to be ``first-order'', in that they are derived by mapping a variable of interest to a landcover type(s).
-The first of these was to assess potential biases in Tier-1 carbon accounting maps that we constructed from the cropland maps, following \cite{ruesch_new_2008}.
-The second was an assessment of gridded maize yields derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}.
-We used these same data to calculate gridded maize production estimates, which we used to conduct the third first order analysis.
+We then investigated how landcover map error propagates through ``downstream'' analyses that are built on landcover data, and which are typical of those performed in global change assessments.
+We began by examining the biases and mean absolute errors in three ``first-order'' landcover-based analyses, in which a variable of interest is mapped onto a landcover type(s) using a simple empirical relationship.
+The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}.
+The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates.
+Maps based on these analyses underpin many assessments of crop productivity and production \cite[e.g.][]{foley_solutions_2011,licker_mind_2010}.
 
-Finally, we conducted two second-order analyses, in which a process model draws on the cover types' values to calculate an output value.
-For the first of these, we examined how cropland map biases influence 25 km resolution monthly evapotranspiration calculated by the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model.
-For this example, we used the cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
-In the second example, we examined how these errors can impact the parameterization of an agent-based food security model \cite{chen_dependency_2013}.
+Finally, we examined errors resulting from two second-order analyses, in which a process model draws on the cover types' values to calculate an output value.
+For the first of these, we used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model to calculate monthly evapotranspiration, using the reference and cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
+In the second example, we examined how these map errors impact the parameterization of an agent-based food security model \cite{chen_dependency_2013}.
 Spatially-explicit, agent-based models are frequently employed in land change science, and require an initialization step to assign landscape resources to model agents  \cite[e.g.][]{manson_agent-based_2007,evans_multi-scale_2004,kelley_relative_2011}.
@@
 \section{Bias}
-\subsection{Landcover bias}
+\subsection{Landcover error and bias}
 
@@
 
-We then aggregated the reference and each cropland map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from the reference map at each scale of aggregation to assess patterns of differences (Fig.
+We then aggregated the reference and each cropland map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from the reference map at each scale of aggregation to assess error patterns (Fig.
 1).
@@
 \centerline{\includegraphics[width=.5\textwidth]{figures/bias_map.pdf}}
-\caption{Differences in percent cropland estimates between the reference map and each of the four landcover maps.
+\caption{Errors in the percent cropland estimates resulting from each of the four cropland maps relative to the reference map at different scale of pixel aggegation.
 Rows indicate the landcover map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation.
@@
 
-The most pronounced differences were in the MODIS and GlobCover maps, which both underestimated cropland extent by 10-75\% in the center of the country (blue areas in Fig.
+The most pronounced errors were in the MODIS and GlobCover maps, which both underestimated cropland extent by 10-75\% in the center of the country (blue areas in Fig.
 1, and the dominant production region), and overestimated along the eastern to northern margins (red areas in Fig.
 1).
-The differences between the reference and these two maps at 1 km resolution were respectively 21\% (MODIS) and 34\% (GlobCover), on average.
-This mean pixel-wise error expresses the amount of \emph{bias} in a map, that is its tendency to err in a particular direction.
-For MODIS, bias drops to 8\% at 50 km of aggregation, whereas GlobCover bias is 24\% at 100 km (SI Appendix, Fig.
+These average of these were 21\% for MODIS and 34\% for GlobCover at 1 km resolution (Fig.
+S1), meaning that each map had a strong underestimation bias.
+Both maps's biases decreased with pixel aggregation, with MODIS biases falling to 8\% at 50 km, but GlobCover's was still 24\% at 100 km (SI Appendix, Fig.
 S1).
@@
 The SA-LC map uniformly overestimated cropland throughout the country (Fig.
-1), but was less biased, ranging from -8\% at 1 km to -6\% at 100 km  (Fig.
+1), but its overall bias was relatively small, ranging from -8\% at 1 km to -6\% at 100 km  (Fig.
 S1).
-The GeoWiki map has a strongly heterogenous pattern of differences (Fig.
+The GeoWiki map had a strongly heterogenous pattern of differences (Fig.
 1) and the smallest magnitude of bias, which changed between slight tendencies to underestimate (5\% at 1 km) and overestimate (-2 \% at 100 km,  Fig.
@@
 
-\subsection{The role of landscape characteristics in shaping bias}
-To investigate the degree to which landscape features influence landcover map bias, we extracted all pixels in agricultural areas ($>$0\% cropland) of the 1 km reference map using the boundaries of 354 magisterial districts (South Africa's finest administrative unit, which average 3,445 km$^2$ in size; SI Appendix, Fig.
-S3), and calculated the district-wise mean for these pixels to provide a measure of cropland density, and information on the degree of mixing between cropland and other land covers.
-We then extracted the cropland map errors for the same pixels, and calculated the absolute value of bias for each district.
-Absolute bias describes the magnitude of bias but not its direction, and is more informative about the likelihood of the map being biased at any given point in the landscape than actual bias, as positive and negative biases can cancel each other out within small areas.
+\subsection{Landscape characteristics and error magnitude}
+To investigate the degree to which landscape features influence landcover map errors, we extracted all pixels in agricultural areas ($>$0\% cropland) of the 1 km reference map using the boundaries of 354 magisterial districts (South Africa's finest administrative unit, which average 3,445 km$^2$ in size; SI Appendix, Fig.
+S3), and calculated the district-wise mean for these pixels, providing a measure of cropland density that was informative of the degree of mixing between cropland and other land covers at landscape scales.
+We then extracted the cropland map errors for the same pixels, and calculated their mean absolute errors for each district.
 
-A generalized additive model fit to district-level absolute bias (log-transformed) shows that absolute bias peaks at 50\% cropland cover for all but the GlobCover map (which continued to increase with cropland cover), and is lowest when the landscape is dominated either by cropland or other cover types (Fig.
+A generalized additive model fit to district-level mean absolute error (log-transformed) shows that error peaks at 50\% cropland cover for all but the GlobCover map (which continued to increase with cropland cover), and is lowest when the landscape is dominated either by cropland or other cover types (Fig.
 2).
 In other words, bias is highest when cropland cover is mixed evenly with other cover types.
+The reason that GlobCover bias continued to increase with cropland density was because this dataset's dominant class identified for South African cropland was a mixture of 50-70\% cropland and other vegetation types, which resulted in the capping of GlobCover-derived cropland cover estimates at the intermediate densities defined by this class.
 
@@
 \begin{figure}[ht]
-\centerline{\includegraphics[width=.4\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
-\caption{The relationship between the mean absolute bias in cropland maps and cropland cover within agricultural areas (reference map pixels having $>$0.5\% cropland), averaged within the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
-Prediction curves are color-coded to the different cropland maps, with the solid line indicating  predicted absolute bias, and the lighter shading the standard error of the coefficients.
-Models were also fit to the mean absolute biases across all four maps (black curve), and from all maps exlcuding GlobCover (grey surve).}
+\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
+\caption{The relationship between the mean absolute error (error magnitude) in cropland maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
+Prediction curves are color-coded to the different cropland maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}
 \label{afoto}
@@
 
-\subsection{The impact of bias on calculating carbon stocks}
-We used the data of \cite{ruesch_new_2008} to calculate carbon densities for African forests, shrublands, croplands, grasslands, and sparse habitats (semi-arid grasslands and low shrublands), assigning cropland carbon values to map cells in proportion to their cropland cover.
-For the non-cropland proportions, we assigned the carbon value from each of the other types, such that we created four different carbon maps for each landcover map at each aggregation scale (Fig.
+\subsection{Map error and Tier-1 carbon estimates}
+Using the methods provided by \cite{ruesch_new_2008}, we calculated average carbon densities for African forests, secondary forests, shrublands, croplands, grasslands, and sparse habitats (semi-arid grasslands and low shrublands), and assigned cropland carbon values to map cells in proportion to their cropland cover.
+For the non-cropland proportions, we assigned the carbon value from each of the other types, creating five different carbon maps for each landcover map at each aggregation scale (Fig.
 S4), which allowed us to test how carbon estimates vary as a function of i) cropland map bias and ii) the characteristics of adjacent cover types.
@@
 For lower density cover (grassland and sparse vegetation), carbon stocks were underestimated, but by small magnitudes.
-These tendencies were reflected in each map's biases, calculated over agricultural areas (Fig.
+These tendencies were reflected in each map's biases, as calculated over the cropped areas of the country as jointly defined by the reference and each cropland map (Fig.
 3).
@@
 Overall, GeoWiki had the lowest bias, for all cover types and all resolutions.
-It worst bias was a tendency to overestimate by 12\% at 1 km when forest was adjacent, but at coarser scales this bias reduced to just a few percent (Fig.
+Its worst bias was a tendency to overestimate by 12\% at 1 km when forest was adjacent, but at coarser scales this bias reduced to just a few percent (Fig.
 3, Table S2).
@@
 \begin{figure}[!ht]
-\centerline{\includegraphics[width=.4\textwidth]{figures/carbon_veg_scale3.pdf}}
-\caption{The biases in carbon density estimates based on cropland maps, expressed as a percent difference relative to the reference map.
-The values of actual bias (incorporating bias direction and magnitude) and absolute bias (magnitude only) are presented in semi-transparent and solid colors, respectively, with colors denoting the specific cropland map and symbols indicating which cover type was used to calculate cropland adjacent carbon density.
+\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scale.pdf}}
+\caption{Biases and mean absolute errors in carbon densities derived from cropland maps, calculated as percents relative to the reference map.
+Bias estimates (represented by symbols) fall within the semi-transparent bars, while mean absolute errors are contained in the solid bars.
+Bar colors are coded to specific cropland map, while the symbols indicate which cover type was used to calculate cropland adjacent carbon density.
 The bar represents the mean biases calculated across each of the 5 cover types.
-Shrubland and grassland bias values are near zero and not shown for display clarity.}
+Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2).}
 \label{afoto}
@@
 
-Absolute bias in carbon maps (solid colored bars in Fig.
-3) generally followed the same patterns, but with somewhat higher magnitudes and a few important differences.
-The most notable is that Geo-Wiki had fairly large absolute bias at 1 km, averaging 27\% across across cover types (line in solid blue bar, Fig.
+The mean absolute error in carbon maps (solid colored bars in Fig.
+3) generally followed the same patterns, but with higher magnitudes and a few important differences.
+The most notable is that Geo-Wiki, despite relatively low bias, had large mean absolute errors at 1 km, averaging 27\% across across cover types (line in solid blue bar, Fig.
 3, Table S2), which is close to the 36-37\% for GlobCover and MODIS.
-SA-LC had the lowest absolute bias across scales, averaging (across cover types) from 14\% at 1 km to 3\% at 100 km.
+SA-LC had the lowest mean absolute error across scales, averaging (across cover types) 14\% at 1 km to 3\% at 100 km.
 (Fig.
@@
 The increase in Geo-Wiki's absolute bias relative to SA-LC's can be attributed to the highly heterogeneous nature of its cropland errors (Fig.
-1), which varied between positive and negative over much shorter distances than the other three cropland maps.
+1), which traded between between positive and negative errors of high magnitude over shorter distances than the other three cropland maps.
 
-\subsection{Bias in harvested areas, yield, and production estimates}
-The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived agricultural area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}.
-To be consistent with this methodology, we adjusted our cropland maps using the same procedure, but we used the reference map to calculate total cropland area for each of South Africa's nine provinces, and then updated the pixel-wise cropland percentages in the four cropland maps so that the province-wise sums matched the reference areas \cite[][and see SI]{ramankutty_farming_2008}.
-Despite this statistical constraint, the updated cropland maps still had large spatial errors that were similar in pattern (Fig.
+\subsection{Map error and gridded agricultural data}
+The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived cropland area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}.
+To be consistent with this methodology, we first adjusted our cropland maps according to this procedure, using the reference map to calculate total cropland area for each of South Africa's nine provinces, then updating the pixel-wise cropland percentages in the four cropland maps so that the province-wise sums matched the reference areas \cite[][and see SI]{ramankutty_farming_2008}.
+Despite this statistical constraint, the updated cropland maps still had substantial errors that were similar in pattern (Fig.
 S5) to those in the unadjusted maps (Fig.
 1), and we evaluated how these residuals affected gridded estimates of the yield and production of maize, South Africa's largest crop \cite{estes_comparing_2013}.
-To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.
+To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize (South Africa's largest crop by area, \cite{estes_comparing_2013}) yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.
 
@@
 %\vspace{-1 cm}
-\begin{figure}[ht]
-\centerline{\includegraphics[width=.4\textwidth]{figures/yield_bias_scale.pdf}}
-\caption{The biases in disaggregated maize yield and production estimates.
-The values of actual bias (incorporating bias direction and magnitude) and absolute bias (magnitude only) are presented in semi-transparent and solid colors, respectively, with colors denoting the specific cropland map and symbols indicating whether the bias is related to one of two yield aggregation methods and associated crop production estimates.}
+\begin{figure}[!ht]
+\centerline{\includegraphics[width=.5\textwidth]{figures/yield_bias_scale.pdf}}
+\caption{Biases and mean absolute errors in disaggregated maize yield and production estimates.
+Bias estimates (represented by symbols) fall within the semi-transparent bars, while mean absolute errors are contained in the solid bars.
+Bar colors are coded to specific cropland map, while the symbols indicate which the values for the different variables (production and yield).}
 \label{afoto}
@@
 The statistical constraints on harvested and cropland areas resulted in the canceling out of spatial errors in production estimates, which is evident in the checkerboard-like pattern in maps of production biases (Figure S7).
-However, this pattern of bias means that absolute bias in production estimates were large, between 40 to 55\% for GeoWiki, MODIS, and GlobCover at 1 km, and remained generally high (10-28\%) even up to 50 km of aggregation (Fig.
+However, this reduction in bias comes at the cost of higher error magnitude, as the mean absolute error in production estimates were large, between 40 to 55\% for GeoWiki, MODIS, and GlobCover at 1 km, and remained generally high (10-28\%) even up to 50 km of aggregation (Fig.
 5).
-SA-LC production biases were lowest across all spatial scales (20\% at 1 km to 2\% in 100 km).
+SA-LC production biases were lowest across all spatial scales (20\% at 1 km, dropping linearly to 2\% by 100 km).
 
-Absolute mean biases in yield were also substantial, and generally 10-15\% larger than production biases across all spatial scales, except for GlobCover where absolute production biases exceeded yield bias at 5-100 km of aggregation.
+Absolute mean errors in yield were also substantial, and generally 10-15\% larger than production biases across all aggregation scales, except for GlobCover where absolute production biases exceeded yield bias at 5-100 km of aggregation.
 
-\subsection{Bias in evapotranspiration estimates}
-Compared to carbon and crop related examples, actual and absolute biases in evapotranspiration calculated using the VIC model were small and averaged to less than +/-1\%.
-However, there are several hotspots of discrepancy evident in the error maps (Fig.
+\subsection{Map error and evapotranspiration}
+Compared to carbon and crop related examples, bias and mean absolute errors in evapotranspiration (ET) calculated using the VIC model were small and averaged to less than +/-1\%.
+However, there were several hotspots of discrepancy evident in the error maps (Fig.
 6).
-The most pronounced of these are the 5-15\% overestimates in the center of the country that resulted when VIC was initialized with MODIS and GlobCover, while overestimates along the southern and western coasts reached 25\%.
-These locations correspond primarily to the margins of major crop production regions--in the center is the westernmost boundary of the summer rainfall growing region, where maize is the primary crop, while the hotspot on the west coast marks the western boundary of the wheat-dominated winter rainfall region \cite{hardy_rainfed_2011}.
-These two boundaries respectively correspond to the 400 and 200 mm isohyets of growing season rainfall.
+The most pronounced of these are the 5-15\% overestimates in the center resulting from VIC when initialized with MODIS and GlobCover, while overestimates along the southern and western coasts reached 25\%.
+These locations correspond primarily to the margins of major crop production regions--in the center is the westernmost boundary of the summer rainfall growing region, marked approximately by the 400 mm isohyet, where maize is the primary crop.
+The west coast hotspot falls at the western edge of the wheat-dominated winter rainfall region \cite{hardy_rainfed_2011}, where growing season rainfall is approximately 200 mm.
 
-SA-LC and GeoWiki also produced biased ET estimates along the southern and western coasts, but here the tendency was to underestimate ET, while biases in the center of the country were either negligible to absent.
+SA-LC and GeoWiki also resulted in ET errors estimates along the southern and western coasts, but here the tendency was to underestimate ET, while biases in the center of the country were either negligible to absent.
 All but MODIS underestimated ET by 5-15\% in the northern tip of the country.
 
-\FloatBarrier
-\begin{figure}[ht]
-\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
-\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four cropland maps.}\label{afoto}
-\end{figure}
-
-\subsection{Initialization biases in agent-based models}
+\subsection{Initialization errors in agent-based models}
 We used an agent-based model (ABM) of food security that represents the interactions between hundreds of individual farming households over multiple seasons \cite{chen_dependency_2013}.
@@
 This initialization process
+%\FloatBarrier
+\begin{figure}[!ht]
+\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
+\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four cropland maps.}\label{afoto}
+\end{figure}
 
@@
 
+Main points:
+
+1. Previous studies have quantified disagreement over large areas, where
+
+What we found, significance of study 
 \begin{itemize}
@@
 
-Some points to make thrown out now because out of time. Others will be added. Please add any important implications/considerations you see from the results.
+\begin{itemize}
+  \item Bias decreases as function of scale
+  \item General bias patterns, appropriate use of landcover products, which landcover products
+  \item Appropriate scales of inference, by type of product - 
+  \item Aggregation improves results for landcover, generally, 
+  \item Sensor resolution, statistical resolution, and merging products have high value
+   \begin{itemize}
+     \item But don't remove spatial bias - absolute bias matters. Statistical constraints seems to just compress spatial biases to higher rates of turnover. Geo-wiki
+     \item But of course these types of data are then dependent on how accurate the statistical data are defining the constraint (cite emperor has no data)
+    \end{itemize}
+   \item Mixed landscapes increase the chances of omission and commission errors by increasing the number of cover classes, or because such landscapes are less spectrally distinct \cite{estes_diylandcover:_2015}
+   \item caveats: Only single country in South Africa.
+More commercial farming than many other countries, but results are still instructive.
+Analysis of error as function of landscape type suggests that areas where cropland is more mixed with natural vegetation have higher errors.
+These sorts of landscapes quite common in smallholder-dominated systems, thus suggests that biases may be even higher elsewhere on the continent.
+\end{itemize}
 
-Some points from my notes last year: 
+Implications for understanding global change and policy: \\
+
+Increasing awareness of need to have spatial assessments in global change analyses.
+Do things such as identify areas where yield gaps are high, or how much carbon or biodiversity will be lost to changes in land use, in order to try prioritize development \cite{searchinger_high_2015,newbold_global_2015}, or to understand coupled human-social dynamics, etc.
+\\
+
+Our finding suggest:
+
 \begin{itemize}
-  \item Bias as a function of scale
-    \begin{enumerate}
-      \item At 1 km resolution all landcover products are still fairly biased.  
-       \item Bias drops to acceptable levels quickly for geowiki--at 5X5 km, mean bias is just 1\% (overestimated).
-The absolute bias for this dataset is 10\% or lower from 10X10 km resolution and coarser.
-       \item The SA dataset's bias is fairly consistent but low across all levels of aggregation, amounting to no more than an 8\% overestimate of cropland with absolute bias of similar magnitude. 
-       \item MODIS and GlobCover biases (mostly of underestimation) do not dissipate until the higher levels of aggregation.
-MODIS's actual bias (under-estimation) falls below 10\% at 20 km resolution, but the absolute bias remains above 10\% until more than a 100-fold aggregation is done ($>$100 km resolution).
-For GlobCover, it is still too high.
-    \end{enumerate}
-  \item Bias as a function of cropland cover
-    \begin{enumerate}
-      \item Classification algorithms are thus more error-prone where landcover is mixed/heterogenous.
-      \item The exception to this lies in the GlobCover dataset, where bias primarily increases as a function of cropland cover.
+  \item Area-based estimates only safe at coarser scales of aggregation for most types of global change analyses, and primarily with constrained products. 
+  \item 50-100 km scale of aggregation reduces bias sufficiently.
+  \item Not so with unconstrained products
+  \item Assessments of spatial variability unsafe, for all products, bar one - finer country-scale product.
+Here you look at absolute bias.
+This is high in many products even at higher scales of aggregation.
+  \item This suggests that disaggregation approaches or paint by numbers approaches are nice maps, but can't give clear guidance about differences between grid cells, even when highly aggregated.
+[work on this]
+  \item can lead to misinformed policies
+  \begin{itemize}
+    \item E.g.
+Efforts to identify area where yield gaps are most pronounced and/or concentrated are likely to be highly misleading, leading to ineffective targeting of resources.
+Most informative simply to look at these areas at the political boundary resolution
+    \item Comparing carbon stocks against potential yield for tradeoff analysis, which may be done with conservation planning to find areas with high benefit/low-cost. Also misleading. 
+    \item Looking at land availability for cropland or biofuels expansion (look at biofuel paper for example)--land might not be as available as people think. Can lead to formulation of bad policy
+    \end{itemize}
+  \item Analyses of higher order interactions, biogeochemistry, human decision-making, also misleading (maybe pair this with yield example). 
+  \begin{itemize}
+     \item Our example here, ET estimates not heavily biased, but in marginal areas of low rainfall some pronounced differences.
+These are areas where irrigation is more common, but VIC doesn't simulate this, so absolute bias in those zones likely to be underestimated, and such regions can have substantial impacts on altering climate \cite{estes_changing_2014, sacks_effects_2008}.
+    \item Can skew understanding of more advanced attempts to understand the human factors that go into driving agricultural productivity. Examples here
+    \end{itemize}
+\end{itemize}
+
+Way forward
+\begin{itemize}
+  \item For now, use latest generation products fusion products or more detailed country-level products
+  \item Avoid change detection based on landcover products, e.g. MODIS. 
+  \item But moving forward key will be developing new approaches to map landcover with much greater fidelity, e.g.
+scaling out approach that led to this dataset, combining with latest computer vision algorithms, etc.
+\end{itemize}
+
+%Some points from my notes last year: 
+%\begin{itemize}
+%  \item Bias as a function of scale
+%  \item Bias as a function of cropland cover
+%    \begin{enumerate}
+%      \item Classification algorithms are thus more error-prone where landcover is mixed/heterogenous.
+%      \item The exception to this lies in the GlobCover dataset, where bias primarily increases as a function of cropland cover.
 The reason for this is that GlobCover's cropland classes do not provide for 100\% cropland cover, so aggregation tends to exacerbate underestimates.
-      \item Thus caution is needed when aggregating a mixed pixel class.  
-        \begin{itemize} 
-          \item An example illustrats this: take 4 1 km pixels, 2 of which are 100\% cropland, 2 of which are other cover types.
+%      \item Thus caution is needed when aggregating a mixed pixel class.  
+%        \begin{itemize} 
+%          \item An example illustrates this: take 4 1 km pixels, 2 of which are 100\% cropland, 2 of which are other cover types.
 Imagine a landcover product classifies 3 of these as cropland (2 correct, 1 an error of commission), using a cropland class that is defined as 50\% cropland.
 Aggregating the actual fraction by a factor of 4 will result in a new 4 km pixel having 50\% cropland, whereas aggregating the landcover product's pixels will give just 38\% cropland, even when factoring in the incorrect classification.)
-          \end{itemize}
-    \end{enumerate}
-  \item Bias as a function of method
-      \begin{enumerate}
-        \item Higher resolution and ensemble-based approaches have less bias.
-        \item geowiki represents a fusion of multiple coarse resolution data sources that has undergone extensive validation using a crowdsourcing approach
-        \item the SALC dataset is based on 30 m landsat data, but incorporates a range of ancillary data and expert judgement
-        \item MODIS and GlobCover data are effectively single source/single algorithm.
-        \item \textbf{Newer points begin here}
-        \item Statistically constrained constrained landcover estimation approaches provide accurate area-based inferences when aggregated.
+%          \end{itemize}
+%    \end{enumerate}
+%  \item Bias as a function of method
+%      \begin{enumerate}
+%        \item Higher resolution and ensemble-based approaches have less bias.
+%        \item geowiki represents a fusion of multiple coarse resolution data sources that has undergone extensive validation using a crowdsourcing approach
+%        \item the SALC dataset is based on 30 m landsat data, but incorporates a range of ancillary data and expert judgement
+%        \item MODIS and GlobCover data are effectively single source/single algorithm.
+%        \item \textbf{Newer points begin here}
+%        \item Statistically constrained constrained landcover estimation approaches provide accurate area-based inferences when aggregated.
 But spatial errors are still high, as seen with GeoWiki and production/yield estimates.
 Using these to identify yield gaps at specific map locations is inappropriate, or even for a larger location if it does not coincide with the geographic boundaries of the statistical unit.
-        \item Constrained estimates are also dependent on the accuracy of the statistics.  
-    \end{enumerate}
-  \item Fix above to have section on bias for global change studies
-    \begin{enumerate}
-      \item Scales at which it is safe to estimate values of say carbon stocks. 
-      \item Above point about bias in disaggregated yield estimates - no point mapping these out.
+%        \item Constrained estimates are also dependent on the accuracy of the statistics.  
+%    \end{enumerate}
+%  \item Fix above to have section on bias for global change studies
+%    \begin{enumerate}
+%      \item Scales at which it is safe to estimate values of say carbon stocks. 
+%      \item Above point about bias in disaggregated yield estimates - no point mapping these out.
 A new approach might be to take these statistically reported yields and then combine them with satellite data to estimate yield variability within the district.
 That way would have meaningful reason for disaggregating yields, and would be pegged to real yield values, which would help minimize errors in remote sensing of yields.
-      \item Something on ET - doesn't seem to matter much, but land-atmosphere interactions can make these discrepancies meaningful, particularly since biases occur in arid areas where a lot of irrigation happens--can cause significant impacts on regional climate.
+%      \item Something on ET - doesn't seem to matter much, but land-atmosphere interactions can make these discrepancies meaningful, particularly since biases occur in arid areas where a lot of irrigation happens--can cause significant impacts on regional climate.
 etc.
@@
 in forested landscapes).
-      \item Agent-based models. \textcolor{red}{Tom, Peng, something of significance/implications of this, please}
-    \end{enumerate}
-  \item Will need a section on way forward for data, etc.
+%      \item Agent-based models. \textcolor{red}{Tom, Peng, something of significance/implications of this, please}
+%    \end{enumerate}
+%  \item Will need a section on way forward for data, etc.
 Key role of accurate landcover, particularly agricultural.
 New methods, vectorized field boundaries seem to be highly valuable, Mapping Africa, Stephanie's paper, Geo-Wiki, etc are the way ahead.
+%  
+%    
+%\end{itemize}
 
-\end{itemize}
-
-Mixed landscapes increase the chances of omission and commission errors by increasing the number of cover classes, or because such landscapes are less spectrally distinct \cite{estes_diylandcover:_2015}
 % $\theta$ changes from 0 to 1. (see Figure \ref{afoto}).That means we are considering $\theta$  of the form
@@
 %(see \eqref{weaksol}).
-\subsection{More blather}
 
@@
 \bibliographystyle{pnas2011} 
-\bibliography{bias}
+{\footnotesize \bibliography{bias}}
 

### 2015-10-01 Rmd/agent-allocation-bias.Rmd
A number of methods revised and redone. Mostly described in new changelog file.

--- 2015-09-11 Rmd/agent-allocation-bias.Rmd
@@
 gti <- raster(paths[2])  # gti 2011
+namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
+gti <- mask(gti, namask)  # apply mask to GTI
+# cellStats(!is.na(gti), sum)
 
@@
 snms <- c("sa30", "globmu", "modmu", "geow")
+dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
 dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2011))
-lc_list <- gti - dlist_1km
+lc_list <- lapply(dlist_1km, function(x) gti - x) 
+# lc_list <- gti - dlist_1km 
+# cellStats(!is.na(lc_list$sa30), sum)
+lc_list <- stack(lc_list)
 
@@
 ```{r, eval = FALSE}
-abias <- data.table(read_excel(full_path(p_data, "abm/agent-bias.xlsx")))
+abias <- data.table(read_excel(full_path(p_data, "abm/agent-bias2.xlsx")))
 abias <- abias[Alg == 1]
 setkey(abias, MD)
-hhs <- data.table(read_excel(full_path(p_data, "abm/agent-bias.xlsx"), 
+hhs <- data.table(read_excel(full_path(p_data, "abm/agent-bias2.xlsx"), 
                              sheet = 2))
@@
 
+# land deficit farmland
+land_def <- round(abias[var == "land_def", snms, with = FALSE] / 
+  abias[var == "ag_area", snms, with = FALSE] * 100, 2)
+
 cexs <- seq(1, 2, 1 / 3)
 # pdf(full_path(p_fig, "agent-bias.pdf"), height = 5, width = 3)
-pdf(full_path(p_fig, "agent-bias.pdf"), height = 3, width = 5)
-cx <- c(0.7, 0.8, 0.6)
+pdf(full_path(p_fig, "agent-bias.pdf"), height = 2.5, width = 5)
+cx <- c(0.9, 1, 0.6)
 # png(full_path(p_fig, "agent-bias.png"), height = 350, width = 700)
@@
 #     tcl = -0.2, oma = c(3.5, 0, 0, 0))
-par(mfrow = c(1, 2), mar = c(2, 2, 0.5, 0), mgp = c(0.75, 0.05, 0), 
+par(mfrow = c(1, 3), mar = c(4, 2, 0.5, 0), mgp = c(0.9, 0.1, 0), 
     tcl = -0.2, oma = c(3.5, 0, 0, 0.1))
@@
 plot(unlist(pctbias), unlist(noland), col = cols, cex = cexs, pch = 20, 
-     ylab = "% HH without farms", xlab = "% cropland bias", cex.lab = cx[2], 
-     cex.axis = cx[1])
+     ylab = "% HH without farms", xlab = "% error in cropland area", 
+     cex.lab = cx[2], cex.axis = cx[1])
 plot(unlist(pctbias), unlist(unalloc), col = cols, cex = cexs, pch = 20, 
-     ylab = "% land unallocated", xlab = "% cropland bias", cex.lab = cx[2], 
-     cex.axis = cx[1])
+     ylab = "% land unallocated", xlab = "% error in cropland area", 
+     cex.lab = cx[2], cex.axis = cx[1])
+plot(unlist(pctbias), unlist(land_def), col = cols, cex = cexs, pch = 20, 
+     ylab = "% land deficit", xlab = "% error in cropland area", 
+     cex.lab = cx[2], cex.axis = cx[1])
 par(xpd = NA)
@@
 }
-yst <- seq(0.05, 0.23, 0.18 / 4)
-xy <- cbind(grconvertX(rep(0.515, 5), from = "ndc"), 
+yst <- seq(0.06, 0.24, 0.18 / 4)
+xy <- cbind(grconvertX(rep(0.505, 5), from = "ndc"), 
             grconvertY(yst, from = "ndc"))
@@
      pos = 2, col = c(scols, "black"), cex = cx[1])
-xst <- seq(0.545, 0.65, 0.105 / 3)
+xst <- seq(0.53, 0.60, 0.07 / 3)
 xy <- cbind(grconvertX(xst, from = "ndc"), 

### 2015-10-01 Rmd/carbon-bias.Rmd
A number of methods revised and redone. Mostly described in new changelog file.

@@
 gti <- raster(paths) / 100  # gti 2011
+namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
+gti <- mask(gti, namask)  # apply mask to GTI
+# cellStats(!is.na(gti), sum)
 
@@
 plot(lc_list[[1]])
+# cellStats(!is.na(lc_list[[1]]), sum)
 plot(gti)
@@
 # need NA mask for weighting of aggregated value
-sumna <- function(x) sum(x, na.rm = FALSE)
-namask <- calc(stack(gti, stack(lc_list)), sumna)
-namask[namask > 0] <- 1
+# sumna <- function(x) sum(x, na.rm = FALSE)
+# namask <- calc(stack(stack(gti), stack(lclist)), sumna)
+# namask[namask > 0] <- 1
 namask2 <- !is.na(namask)
@@
 # bias estimates
-tot_area <- sum(freq(namask)[1:2, ])
+tot_area <- sum(freq(namask)[1, 2])  # sum of just the non-NA area
 crop_areas <- lapply(lc_union$f1, freq)  
-sapply(crop_areas, function(x) x[2, 2] / tot_area)  # 29, 31, 33, 53 (Glob)%
+sapply(crop_areas, function(x) x[2, 2] / tot_area)  # 29, 53, 33, 31
 
@@
   msk <- gti_agg[[x]]$gti > 0.005  # farmland > 0.05% mask
+  # msk <- gti_agg[[x]]$gti > 0.05  # farmland > 0.05% mask
   msked <- mask(gti_c[[x]] * tareas[[x]], msk, maskvalue = 0)
@@
     msk <- jj[[y]] > 0.005  # farmland > 0.05% mask
+    # msk <- jj[[y]] > 0  # farmland > 0.05% mask
     msked <- mask(xx[[y]] * tareas[[y]], msk, maskvalue = 0)

### 2015-10-01 Rmd/compare-landcover.Rmd
A number of methods revised and redone. Mostly described in new changelog file.

@@
 
+### Apply NA masks
+```{r, eval = FALSE}
+sumna <- function(x) sum(x, na.rm = FALSE)
+namask <- calc(stack(stack(gti), stack(lclist)), sumna)
+namask[namask >= 0] <- 1
+namask2 <- !is.na(namask)  # set NAs to zero
+# cellStats(namask2, sum)
+namask[namask == 0] <- NA
+# cellStats(namask, sum)
+writeRaster(namask, filename = full_path(p_data, "namask.tif"), 
+            overwrite = TRUE)
+
+# mask out all joint NAs
+gti <- lapply(gti, function(x) mask(x, namask))
+lclist <- lapply(lclist, function(x) mask(x, namask))
+
+```
+
 ## Calculate differences at different levels of aggregation
@@
 #fact <- 25
+
+area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)  # cells/pixel
 lc_agg <- aggregate_rast_list(fact, lclist)   # landcover rasters 
@@
 ## Bias as function of cropcover
-### First create NA mask at 1 km resolution, then weighting raster based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc)
+### Weighting rasters based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc) 
 ```{r, eval = FALSE}
-# NA mask and weights
-namask <- !is.na(gti[[1]])
-area_wgts <- aggregate_rast_list(fact, list(namask), fun = sum)
-namask[namask == 0] <- NA
-writeRaster(namask, filename = full_path(p_data, "namask.tif"), 
-            overwrite = TRUE)
 
@@
 dang <- Sys.time()
+# i <- dlists[[1]]; x <- names(i)[1]; y <- lev_vec[5]; j <- 14; z <- 1; k <- 1
 dstats_all <- lapply(dlists, function(i) {
@@
         } else if((nrow(d2) > 0) & (nrow(d2) < length(i[[x]][[y]]) * 20)) {
-          qs <- c(rep(NA, 6), Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], 
-                                              na.rm = FALSE))
+          qs <- c(nrow(d2), rep(NA, 5), 
+                  Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], na.rm = FALSE))
         } else {
@@
       } else if((nrow(oof) > 0) & (nrow(oof) < 10)) {
-        qs <- c(rep(NA, 6), 
+        qs <- c(nrow(oof), rep(NA, 5),  
                 Hmisc::wtd.mean(oof[, 2], weights = oof[, 3], na.rm = FALSE))
@@
 Compared results from original roughed out version of approach for calculating bias statistics with results from new outputs list (using geowiki at 20 km aggregation as the comparison)
+
+## 2011 differences
+Whole country and agricultural areas only
+```{r, eval = FALSE}
+load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
+
+# Subset difference grids, for 2011
+snms <- c("sa30", "globmu", "modmu", "geow")
+lc_pct_diff <- lapply(dlist_act[snms], function(x) {
+  sapply(x, function(y) list(y$g2011))
+})
+
+# construct agricultural area union mask
+lev <- names(lc_pct_diff[[1]])
+lcu <- lapply(lev, function(x) {
+  print(paste(".", x))
+  lcb <- lapply(snms, function(y) {
+    print(paste("...", y))
+    gti_gt0 <- Which(gti_agg[[x]][[1]] > 0)
+    lc_gt0 <- Which(lc_agg[[x]][[y]] > 0)
+    all_gt0 <- gti_gt0 + lc_gt0
+    all_gt0[all_gt0 > 0] <- 1
+    all_gt0
+  })
+  named_out(lcb, snms)
+})
+names(lcu) <- lev
+
+bins <- lapply(gti_bins, function(x) x$g2011)
+awgts <- lapply(area_wgts, function(x) x[[1]])
+
+# calculate means for 2011
+wm <- function(x, w) stats::weighted.mean(x, w)
+wma <- function(x, w) stats::weighted.mean(abs(x), w)
+a <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wm, "mu", "bias", TRUE)
+b <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wma, "mua", "bias", TRUE)
+mu_dt <- rbind(a, b)
+mu <- extract_stat(lev, snms, "all", "mu", "bias", mu_dt)  # ag area means
+mua <- extract_stat(lev, snms, "all", "mua", "bias", mu_dt)  # ag area abs mean
+mu_nm <- extract_stat(lev, snms, "all", "mu.nm", "bias", mu_dt)  # country mean
+mua_nm <- extract_stat(lev, snms, "all", "mua.nm", "bias", mu_dt)  # country abs
+# across bin mean and abs mean
+mu_bin <- mu_dt[!like(bin, "all") & bvals == "mu", mean(bias), by = .(ol, il)]
+setnames(mu_bin, "V1", "bias")
+mua_bin <- mu_dt[!like(bin, "all") & bvals == "mua", mean(bias), by = .(ol, il)]
+setnames(mua_bin, "V1", "bias")
+# mu_dt[il == "sa30" & bin != "all" & ol == "f1" & bvals == "mua", mean(bias)]
+
+# check
+# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
+# version is finding correct results
+bv <- "mua"
+for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
+                c("globmu", "f1"))) {
+  x <- chk[1] 
+  y <- chk[2] 
+  print(paste(".", x, "..", y))
+  rs <- lcu[[y]][[x]] 
+  rs[rs == 0] <- NA
+  l3 <- abs(lc_pct_diff[[x]][[y]])
+  o <- rs * l3
+  w <- awgts[[y]][[1]] * rs
+  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
+  print(b[ol == y & il == x & bvals == bv & bin == "all", bias] ==
+          round(wmu, 2))
+}  # check
+
+svec <- c("mu_nm", "mua_nm", "mu", "mua", "mu_bin", "mua_bin")
+aa <- c(rep("Country", 2), rep("Agricultural", 2), rep("Density independent", 2))
+bb <- c("Bias", "MAE", "Bias", "MAE", "Bias", "MAE")
+mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
+
+lcf_out11 <- do.call(rbind, lapply(1:length(svec), function(i) {
+  ao <- do.call(rbind, lapply(snms, function(x) {
+    ai <- sapply(lev, function(y) {
+      get(svec[i])[ol == y & il == x, bias]
+    })
+    named_out(ai, paste(c(1, fact), "km"))
+  }))
+  #cbind.data.frame("Region" = aa[i], "Map" = mcap, a)
+  cbind.data.frame(Region = aa[i], Metric = bb[i], "Map" = mcap, ao)
+}))
+
+caption <- paste("Biases and mean absolute errors (MAE) in cropland maps",
+                 "relative to the 2011 reference map for each aggregation scale",
+                 "calculated over the entire country, for the union of",
+                 "agricultural regions (cropland $>$ 0), and as density", 
+                 "independent means, wherein the mean bias/MAE values",
+                 "for each of 20 cropland cover classes",
+                 "(representing 5\\% increments of cover 0\\% to 100\\% defined",
+                 "by the reference map) were calculated and then averaged.")
+lcfout_xtab <- xtable(lcf_out11, digits = 1, caption = caption)
+print(lcfout_xtab, type = "latex", file = "paper/figures/cropland2011-bias.tex", 
+      tabular.environment = "longtable", floating = FALSE, 
+      caption.placement = "top", include.rownames = FALSE)
+
+# do.call(rbind, lapply(snms, function(i) {
+#   mu_dt[!like(bin, "all") & ol == "f1" & il == i & bvals == "mu.nm", N]  
+# }))
+mubias_2011 <- mu_dt
+save(lcf_out11, mubias_2011, file = full_path(p_data, "lcf_out2011.rda"))
+
+```
+
+## 2007 differences
+Whole country and agricultural areas only
+```{r, eval = FALSE}
+# Subset difference grids, for 2011
+snms <- c("sa30", "globmu", "modmu", "geow")
+lc_pct_diff <- lapply(dlist_act[snms], function(x) {
+  sapply(x, function(y) list(y$g2007))
+})
+
+# construct agricultural area union mask
+lev <- names(lc_list[[1]])
+lcu <- lapply(lev, function(x) {
+  print(paste(".", x))
+  lcb <- lapply(snms, function(y) {
+    print(paste("...", y))
+    gti_gt0 <- Which(gti_agg[[x]][[1]] > 0)
+    lc_gt0 <- Which(lc_agg[[x]][[y]] > 0)
+    all_gt0 <- gti_gt0 + lc_gt0
+    all_gt0[all_gt0 > 0] <- 1
+    all_gt0
+  })
+  named_out(lcb, snms)
+})
+names(lcu) <- lev
+
+bins <- lapply(gti_bins, function(x) x$g2007)
+awgts <- lapply(area_wgts, function(x) x[[1]])
+
+# calculate means for 2007
+wm <- function(x, w) stats::weighted.mean(x, w)
+wma <- function(x, w) stats::weighted.mean(abs(x), w)
+a <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wm, "mu", "bias", TRUE)
+b <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wma, "mua", "bias", TRUE)
+mu_dt <- rbind(a, b)
+mu <- extract_stat(lev, snms, "all", "mu", "bias", mu_dt)
+mua <- extract_stat(lev, snms, "all", "mua", "bias", mu_dt)
+mu_nm <- extract_stat(lev, snms, "all", "mu.nm", "bias", mu_dt)
+mua_nm <- extract_stat(lev, snms, "all", "mua.nm", "bias", mu_dt)
+# across bin mean and abs mean
+mu_bin <- mu_dt[!like(bin, "all") & bvals == "mu", mean(bias), by = .(ol, il)]
+setnames(mu_bin, "V1", "bias")
+mua_bin <- mu_dt[!like(bin, "all") & bvals == "mua", mean(bias), by = .(ol, il)]
+setnames(mua_bin, "V1", "bias")
+
+# check
+# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
+# version is finding correct results
+bv <- "mua"
+for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
+                c("globmu", "f1"))) {
+  x <- chk[1] 
+  y <- chk[2] 
+  print(paste(".", x, "..", y))
+  rs <- lcu[[y]][[x]] 
+  rs[rs == 0] <- NA
+  l3 <- abs(lc_pct_diff[[x]][[y]])
+  o <- rs * l3
+  w <- awgts[[y]][[1]] * rs
+  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
+  print(b[ol == y & il == x & bvals == bv & bin == "all", bias] ==
+          round(wmu, 2))
+}  # check
+
+svec <- c("mu_nm", "mua_nm", "mu", "mua", "mu_bin", "mua_bin")
+aa <- c(rep("Country", 2), rep("Agricultural", 2), rep("Density independent", 2))
+bb <- c("Bias", "MAE", "Bias", "MAE", "Bias", "MAE")
+mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
+
+lcf_out07 <- do.call(rbind, lapply(1:length(svec), function(i) {
+  ao <- do.call(rbind, lapply(snms, function(x) {
+    ai <- sapply(lev, function(y) {
+      get(svec[i])[ol == y & il == x, bias]
+    })
+    named_out(ai, paste(c(1, fact), "km"))
+  }))
+  #cbind.data.frame("Region" = aa[i], "Map" = mcap, a)
+  cbind.data.frame(Region = aa[i], Metric = bb[i], "Map" = mcap, ao)
+}))
+
+caption <- paste("Biases and mean absolute errors (MAE) in cropland maps",
+                 "relative to the 2007 reference map for each aggregation scale",
+                 "calculated over the entire country, for the union of",
+                 "agricultural regions, and as density", 
+                 "independent means, wherein the mean bias/MAE values",
+                 "for each of 20 cropland cover classes",
+                 "(representing 5\\% increments of cover 0\\% to 100\\% defined",
+                 "by the reference map) were calculated and then averaged.")
+lcfout_xtab <- xtable(lcf_out07, digits = 1, caption = caption)
+print(lcfout_xtab, type = "latex", file = "paper/figures/cropland2007-bias.tex", 
+      tabular.environment = "longtable", floating = FALSE, 
+      caption.placement = "top", include.rownames = FALSE)
+
+```

### 2015-10-01 Rmd/et-bias.Rmd
A number of methods revised and redone. Mostly described in new changelog file.

@@
 lc_list <- lapply(dlist_1km, function(x) gti - x / 100) 
+# cellStats(!is.na(lc_list$sa30), sum)
 
@@
 # Create namask to remove all NA areas across datasets
-sumna <- function(x) sum(x, na.rm = FALSE)
-namask <- calc(stack(gti, stack(lc_list)), sumna)
+# sumna <- function(x) sum(x, na.rm = FALSE)
+# namask <- calc(stack(gti, stack(lc_list)), sumna)
 namask2 <- !is.na(namask)  # set NAs to zero
+# cellStats(namask2, sum)
 # # namaskt <- raster(full_path(p_data, "namask.tif")) # load in NA mask
@@
 bins <- lapply(gti_agg[[2]], function(x) {
-  cut(x$gti, breaks = binv, include.lowest = TRUE)
+  cut(x[[1]], breaks = binv, include.lowest = TRUE)
 })
 names(bins) <- lev_vec
+# bins <- cut(gti_agg$f25$gti, breaks = binv, include.lowest = TRUE)
 

### 2015-10-01 Rmd/landcover-bias-plots.Rmd
A number of methods revised and redone. Mostly described in new changelog file.

@@
 load(full_path(p_data, "bias_stats_gti.rda"))  # bias statistics
+# cellStats(!is.na(dlist_act$modmu$f1$g2011), sum)
 ```
@@
 namask <- raster(full_path(p_data, "namask.tif"))
+# cellStats(namask, sum)
 snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
@@
 # disggregate selected rasters at selected levels
+# x <- snms[3]; y <- lev[3]
 disagg <- lapply(snms, function(x) {
@@
 ## Bias with scale
+
+Note: the mean values presented in the figure here is a straight mean across 5 percentile bins, which gives bias independent of cropland density. Supplementary tables report the bin-weighted biases. 
 ```{r, eval=FALSE}
@@
 
+wm <- function(x, w) stats::weighted.mean(x, w, na.rm = TRUE)
+# wma <- function(x, w) stats::weighted.mean(abs(x), w)
+
 pdf(full_path(p_fig, "biases_1-100km.pdf"), width = 7.5, height = 7)
@@
 # layout.show(l)
-# i <- 1; j <- 1
+# i <- 1; j <- 1; bt <- 1
 par(mar = c(0.5, 0, 1, 0.3), oma = c(5, 5, 2, 2))
@@
     muv <- rep(NA, 2)
+    muv2 <- rep(NA, 2)
     for(bt in 1:2) {
@@
             lty = lt[bt])
-      muv[bt] <- round(mean(d[, 8], na.rm = TRUE))
+      muv[bt] <- round(mean(d[, 8], na.rm = TRUE)) # unweighted means
+      muv2[bt] <- round(wm(d[, 8], d[, 2]))  # weighted means
       toohigh <- which(d[, "mu"] > pp$yl2[2])
@@
     }
-    text(5, pp$yl2[2] - 14, labels = paste(muv, collapse = " / "), cex = 1.1, 
+    text(5, pp$yl2[2] - 5, labels = paste(muv, collapse = " / "), cex = 1.1, 
+         col = "grey10", font = 4)
+    text(5, pp$yl2[2] - 16, labels = paste(muv2, collapse = " / "), cex = 1.1, 
          col = "grey10")
@@
 names(gti) <- c("g2007", "g2011")
+gti <- lapply(gti, function(x) mask(x, namask))  # apply namask
 snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
@@
 
+# Cropland error and bias plots
+
+In relation to calculation method (across country, agricultural region only, or bin-wise (cropland density classes) mean.
+```{r, eval = FALSE}
+load(full_path(p_data, "lcf_out2011.rda"))
+lcf <- data.table(lcf_out11)
+
+fact <- c(1, 5, 10, 25, 50, 100)
+alph <- c(225, 40)
+x <- c(0, 3, 6, 9, 12, 15)
+w <- 3 / 8
+xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
+o <- c(0, w)
+xa <- sapply(x, function(x) x + xo)
+cx <- c(1.25, 1, 1)
+g1 <- "grey90"
+reg <- c("Country", "Agricultural", "Density independent")
+lcnms <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
+mutype <- c("MAE", "Bias")
+svec <- colnames(lcf)[grep("km", colnames(lcf))]
+xl <- c(-0.5, 18)
+yl <- c(-40, 40)
+shd <- c(4.5, 10.5, 16.5)
+cols <- c("red", "orange3", "green4", "blue")
+colnms <- names(lcf)[4:length(names(lcf))]
+pchs <- c("+", "o", "x")
+yax <- seq(yl[1], yl[2], 5)
+xax <- seq(1, 6, 5 / (length(yax) - 1))
+
+pdf(full_path(p_fig, "cropland_bias_scale.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")#,
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+abline(h = yax, v = NULL, col = "grey80", lty = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
+# i <- 1; j <- 1; k <- 1
+for(i in 1:length(svec)) {
+  lv <- colnms[i]
+  for(j in 1:length(snms)) {
+    pchs1 <- pchs
+    nm <- lcnms[j]
+    for(k in 1:length(mutype)) {
+      #dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
+      dat <- lcf[Metric == mutype[k] & Map == nm, lv, with = FALSE]
+      pcol <- makeTransparent(cols[j], alpha = alph[k])
+      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
+      for(z in 1:length(reg)) {
+        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
+               pch = pchs[z], type = "pt")
+      }
+    }
+  }
+}
+
+axis(1, at = seq(1.5, 18.5, 3), labels = fact)
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
+legend(x = 12.4, y = -19.9, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = -19.9, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = -20.5, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
+text(x = 12.2, y = -20.5, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
+legend(x = 12.1, y = 40, 
+       legend = c("Country", "Agricultural", "Density independent"), pch = pchs, 
+       pt.cex = c(1.5, 1.5), bty = "n", cex = 0.8)
+dev.off()
+
+```
+<a href="#top">Back to top</a>
+
+## Bin density plots
+
+To see how many values contribute to bin-wise means.
+```{r, eval = FALSE}
+
+bin_size <- do.call(rbind, lapply(lev, function(i) {
+  # i <- "f50"
+  st <- mubias_2011[bin != "all" & ol == i & il == "sa30" & bvals == "mu.nm", 
+                    .(bin, N)]
+  st[, bin := as.integer(bin)]  # convert bin to integers
+  setkey(st, bin)  
+  st[data.table(bin = 1:20), N]  # merge to bin key to allow NAs
+}))
+
+pdf(full_path(p_fig, "cropland_bins.pdf"), height = 7, width = 7)
+par(mfrow = c(3, 2), mar = c(2, 2, 2, 2), oma = c(3, 3, 0, 0), 
+    mgp = c(1, 0.25, 0), tcl = -0.1)
+for(i in 1:nrow(bin_size)) {
+  barplot(bin_size[i, ], las = 2, col = "grey70", names.arg = seq(5, 100, 5))
+  # axis(1, at = 1:20, labels = seq(5, 100, 5), las = 2)
+  mtext(side = 3, text = paste(fact[i], "km"))
+}
+mtext(side = 1, text = "% Cropland", outer = TRUE)
+mtext(side = 2, text = "Bin count", outer = TRUE, line = 1)
+dev.off()
+
+```
+
 ```{r, eval=FALSE, echo=FALSE}

### 2015-10-01 Rmd/magisterial-district-bias.Rmd
A number of methods revised and redone. Mostly described in new changelog file.

--- 2015-08-27 Rmd/magisterial-district-bias.Rmd
@@
 library(fitdistrplus)
+library(SAcropland)
 
@@
 gti@crs <- sa.shp@proj4string
+namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
+gti <- mask(gti, namask)  # apply mask to GTI
+# cellStats(!is.na(gti), sum)
 
@@
 
-# how many contribute when non-cropland removed
+# how many contribute when non-cropland removed, and which are non-zero
 nonzero <- sapply(gti_md, function(x) length(x[!is.na(x) & (x > 0)]))
+whichnonzero <- sapply(gti_md, function(x) which(!is.na(x) & (x > 0)))
 
@@
 dlist_abs1km <- stack(lapply(dlist_abs[snms], function(x) x$f1$g2011))
+# cellStats(!is.na(dlist_abs1km), sum)
 dlist_abs1km@crs <- sa.shp@proj4string
@@
 # mean no zero
-dlist_mu0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
-  tab <- round(x)
-  round(apply(tab, 2, function(x) mean(x[x > 0], na.rm = TRUE)), 1)
+# dlist_mu0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
+#   tab <- round(x)
+#   round(apply(tab, 2, function(x) mean(x[x > 0], na.rm = TRUE)), 1)
+# })))
+dlist_mu0 <- cbind(nas, t(sapply(1:length(dlist_abs1km_md), function(x) {
+  tab <- round(dlist_abs1km_md[[x]])
+  sel <- whichnonzero[[x]]  # select out reference cropland pixels only
+  round(apply(tab, 2, function(x) mean(x, na.rm = TRUE)), 1)
 })))
+
+# plot(dlist_mu0[, 2], dlist_mu02[, 2])
+# i <- 2
+# plot(gti_mu0[, 3], dlist_mu02[, i])
+# plot(gti_mu0[, 3], dlist_mu0[, i])
 
@@
 
-dlist_median0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
-  tab <- round(x)
-  round(apply(tab, 2, function(x) median(x[x > 0], na.rm = TRUE)), 1)
+# dlist_median0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
+#   tab <- round(x)
+#   round(apply(tab, 2, function(x) median(x[x > 0], na.rm = TRUE)), 1)
+# })))
+dlist_median0 <- cbind(nas, t(sapply(1:length(dlist_abs1km_md), function(x) {
+  tab <- round(dlist_abs1km_md[[x]])
+  sel <- whichnonzero[[x]]  # select out reference cropland pixels only
+  round(apply(tab, 2, function(x) median(x, na.rm = TRUE)), 1)
 })))
@@
 # quartz()
-dat <- mu0[, "sa30"]
+# dat <- mu0[, "sa30"]
+dat <- mu[, "sa30"]
 dat <- dat[!is.na(dat)]
+dat[dat == 0] <- dat[dat == 0] + 0.1
 descdist(dat, discrete = FALSE)  # check potential distributions
@@
   dat <- dat[!is.na(dat)] / 100
+  dat[dat == 0] <- dat[dat == 0] + 0.1
   fits <- lapply(dists, function(x) fitdist(dat, x))
@@
 names(fit_distrs) <- snms
-lapply(fit_distrs, function(x) sort(sapply(x, function(y) y$aic)))
-plot(fit_distrs[[1]]$lnorm)  # lognorm shows up as best fit
+# lapply(fit_distrs, function(x) sort(sapply(x, function(y) y$aic)))
+plot(fit_distrs[[2]]$lnorm)  # lognorm shows up as best fit
 
@@
 # dataset
-# dat <- proc_df(x = mu0, nm = snms[4])  # interactively change response  
-dat <- proc_df(x = med0, nm = snms[1])  # and input data
+dat <- proc_df(x = mu, nm = snms[4])  # interactively change response  
+# dat <- proc_df(x = med0, nm = snms[1])  # and input data
 dat <- dat[!is.na(dat$pv), ]
+dat[, 4] <- dat[, 4] + 0.1  # add 1 to allow for log transform
 
@@
 dgam3 <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
-            family = Gamma(link = "log"), method = "ML")
+            family = Gamma(link = "identity"), method = "ML")
 summary(dgam3)
@@
 # above, with smooth fit instead of quadratic
-dgam4b <- gam(log(dv) ~ s(pv, k = 3) + s(x, y), weights = w, data = dat, 
+dgam4b <- gam(log(dv + 0.0001) ~ s(pv, k = 3) + s(x, y), weights = w, data = dat,
               method = "ML")
+plot(dat$pv, dat$dv)
 summary(dgam4b)
@@
 ```{r, eval = FALSE}
-
 predfunc <- function(model, newdat) {
@@
   dat1 <- dat1[!is.na(dat1$pv), ]
-  dat2 <- proc_df(x = med0, nm = i)
+  dat1$dv <- dat1$dv + 0.1  # adding 0.1 to allow log transform
+  dat2 <- proc_df(x = mu, nm = i)
   dat2 <- dat2[!is.na(dat2$pv), ]
+  dat2$dv <- dat2$dv + 0.1  # adding 0.1 to allow log transform
   dgam1 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat1, 
@@
   pdat2 <- predfunc(dgam2, newdat)
+  # pdat1 <- termfunc(dgam1, newdat)
+  # pdat2 <- termfunc(dgam2, newdat)
   return(list("mumod" = dgam1, "medmod" = dgam2, "pdat1" = pdat1, 
@@
 lw <- c(rep(2, 4), 3, 3)
-yl1 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat1$fit))), 1))
-yl2 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat2$fit))), 1))
-pdf(full_path(p_fig, "biases_md_lnorm_gam_mu0.pdf"), width = 7, height = 7)
+yl1 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat1$fit) * 1.1)), 1))
+yl2 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat2$fit) * 1.1)), 1))
+pdf(full_path(p_fig, "biases_md_lnorm_gam_mu0.pdf"), width = 7, height = 4.5)
 par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
-plot(xl, yl2, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i",
-     mgp = c(1.5, 0.5, 0), yaxs = "i", yaxt = "n")
+plot(xl, yl1, pch = "", ylab = "Mean Absolute Error (%)", 
+     xlab = "% Cropland", xaxs = "i", mgp = c(1.5, 0.5, 0), yaxs = "i", 
+     yaxt = "n")
 axis(2, at = 0:4, round(exp(c(-10, 1:4))))
-for(i in 1:length(gam_mods)) {
+# for(i in 1:length(gam_mods)) {
+for(i in 1:4) {
   m <- gam_mods[[i]]
@@
 
-pdf(full_path(p_fig, "biases_md_lnorm_gam_med0.pdf"), width = 7, height = 7)
-par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2))
-plot(xl, yl2, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i",
-     mgp = c(1.5, 0.5, 0), yaxs = "i", yaxt = "n")
+pdf(full_path(p_fig, "biases_md_lnorm_gam_mu.pdf"), width = 7, height = 4.5)
+par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
+plot(xl, yl2, pch = "", ylab = "Mean Absolute Error (%)", 
+     xlab = "% Cropland", xaxs = "i", mgp = c(1.5, 0.5, 0), yaxs = "i", 
+     yaxt = "n")
 axis(2, at = 0:4, round(exp(c(-10, 1:4))))
-for(i in 1:length(gam_mods)) {
+#for(i in 1:length(gam_mods)) {
+for(i in 1:4) {
   m <- gam_mods[[i]]

### 2015-10-01 Rmd/yield-bias.Rmd
A number of methods revised and redone. Mostly described in new changelog file.

@@
 #p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")
+# load(full_path(p_data, "yield-bias.rda"))
 
@@
 sumna <- function(x) sum(x, na.rm = FALSE)
-namask <- calc(stack(gti, stack(lc_list)), sumna)
-namask[namask > 0] <- 1
+# namask <- calc(stack(stack(gti), stack(lclist)), sumna)
+# namask[namask >= 0] <- 1
 namask2 <- !is.na(namask)  # set NAs to zero
-# cellStats(is.na(gti), sum); cellStats(namask2, sum)
+# cellStats(!is.na(gti), sum); cellStats(namask2, sum)
 fact <- c(5, 10, 25, 50, 100)
@@
 
+# set up masks and weights for assessing output bias and absolute errors
+namask <- calc(stack(gtic, lc_agg$f1), sumna)
+# cellStats(!is.na(gtic), sum); cellStats(!is.na(lc_agg$f1$sa30), sum)
+namask[namask >= 0] <- 1
+namask2 <- !is.na(namask)  # set NAs to zero
+# cellStats(!is.na(namask), sum); cellStats(namask2, sum)
+fact <- c(5, 10, 25, 50, 100)
+awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
+awgts <- lapply(awgts, function(x) x[[1]])  # unlist on inner loop
+
 # compare extents to make sure no offsets - look at in QGIS
@@
 # plot(tst * 100)
-
 #pct_diff <- function(x, y) (x - y) / x * 100
 
+# calculate differences between adjusted fraction and reference
 cf_pct_diff <- lapply(snms, function(x) {
@@
 
+# same, but as percent
+cf_pct_diff2 <- lapply(snms, function(x) {
+  dif <- lapply(names(lc_agg), function(y) {
+    d <- (gti_agg[[y]][[1]] - lc_agg[[y]][[x]]) * 100
+  })
+  named_out(dif, names(lc_agg))
+})
+names(cf_pct_diff2) <- snms
+
+# cropland cover bins, for looking at error as a function of cover, based on 
+# original extent of gti data, for ease
+binv <- seq(0, 1, 0.05)
+bins <- lapply(gti_agg, function(x) {
+  cut(x$gti, breaks = binv, include.lowest = TRUE)
+})
+
+# snms <- names(p_diff1)
+# lcnu <- lapply(lev, function(x) {
+#   lcb <- lapply(snms, function(y) {
+#     all_gt0 <- !is.na(bins[[x]])
+#     all_gt0
+#   })
+#   named_out(lcb, snms)
+# })
+# names(lcnu) <- lev
+lev <- names(cf_pct_diff[[1]])
+snms <- names(p_diff1)
+lcu <- lapply(lev, function(x) {
+  lcb <- lapply(snms, function(y) {
+    gti_gt0 <- Which(gti_agg[[x]][[1]] > 0)
+    lc_gt0 <- Which(lc_agg[[x]][[y]] > 0)
+    all_gt0 <- gti_gt0 + lc_gt0
+    all_gt0[all_gt0 > 0] <- 1
+    all_gt0
+  })
+  named_out(lcb, snms)
+})
+names(lcu) <- lev
+
+# calculate means
+wm <- function(x, w) stats::weighted.mean(x, w)
+wma <- function(x, w) stats::weighted.mean(abs(x), w)
+a <- bias_stats_list(bins, awgts, lcu, cf_pct_diff2, wm, "mu", "bias", TRUE)
+b <- bias_stats_list(bins, awgts, lcu, cf_pct_diff2, wma, "mua", "bias", TRUE)
+# b <- bias_stats_list(bins, awgts, lcnu, cf_pct_diff2, wma, "mua", "bias", TRUE)
+# b[ol == "f1" & il == "modmu" & bin == "all"]
+
+# check
+# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
+# version is finding correct results
+bv <- "mua"
+for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
+                c("globmu", "f1"))) {
+  x <- chk[1] 
+  y <- chk[2] 
+  print(paste(".", x, "..", y))
+  rs <- lcu[[y]][[x]] 
+  rs[rs == 0] <- NA
+  l3 <- abs(cf_pct_diff2[[x]][[y]])
+  o <- rs * l3
+  w <- awgts[[y]][[1]] * rs
+  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
+  print(b[ol == y & il == x & bvals == bv & bin == "all", bias] ==
+          round(wmu, 2))
+}  # check
+
+lcfmua <- extract_stat(lev, snms, "all", "mua", "bias", b)
+lcfmua_nm <- extract_stat(lev, snms, "all", "mua.nm", "bias", b) 
+# b[ol == "f10" & il == "globmu" & bin == "all"]
+
+svec <- c("mua.nm", "mua")
+aa <- c("Country", "Agricultural")
+mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
+
+lcf_out <- do.call(rbind, lapply(1:length(svec), function(i) {
+  a <- do.call(rbind, lapply(snms, function(x) {
+    aa <- sapply(lev, function(y) {
+      b[ol == y & il == x & bin == "all" & bvals == svec[i], bias]
+    })
+    named_out(aa, paste(c(1, fact), "km"))
+  }))
+  cbind.data.frame("Region" = aa[i], "Map" = mcap, a)
+}))
+
+caption <- paste("Mean absolute errors (MAE) in statistically constrained", 
+                 "cropland maps, calculated over the entire country, and for", 
+                 "the union of agricultural regions (cropland $>$ 0) defined ", 
+                 "by the reference and corresponding adjusted cropland",
+                 "maps")    
+lcfout_xtab <- xtable(lcf_out, digits = 1, caption = caption)
+print(lcfout_xtab, type = "latex", file = "paper/figures/cropadj-mae.tex", 
+      tabular.environment = "longtable", floating = FALSE, 
+      caption.placement = "top", include.rownames = FALSE)
+
 # disaggregate for plotting
-namaskc <- crop(namask, gtic)
-lev <- names(cf_pct_diff[[1]])
-disagg <- disaggregate_rast_list(snms, lev, cf_pct_diff, crop(namask, gtic))
+# namaskc <- crop(namask, gtic)
+disagg <- disaggregate_rast_list(snms, lev, cf_pct_diff, namask)
 
@@
 namask <- calc(stack(lapply(crop_ayr, function(x) x$fa)), sumna)
-namask[namask > 0] <- 1
+namask[namask >= 0] <- 1
 namask2 <- !is.na(namask)  # set NAs to zero
-# cellStats(is.na(gti), sum); cellStats(namask2, sum)
+# cellStats(!is.na(gti), sum); cellStats(namask2, sum)
 awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
@@
 # amsk <- lapply(amsk, function(x) x[[1]])  # unlist on inner loop
-
 ```
@@
 # would be: 
-cyld_agg2 <- aggregate_rast_list(fact, cyld_agg$f1, mean) # incorrect yld agg
+# cyld_agg2 <- aggregate_rast_list(fact, cyld_agg$f1, mean) # incorrect yld agg
 
@@
 # maize, according to Monfreda et al methods
-nzmean <- function(x, na.rm) {
-  x <- x[(x > 0) & !is.na(x)]
-  if(length(x) == 0) {
-    m <- 0
-  } else if(length(x) > 0) {
-    m <- mean(x, na.rm = na.rm)
-  }
-  return(m)
-}
-cyld_agg3 <- aggregate_rast_list(fact, cyld_agg$f1, nzmean) # no zero agg
+# nzmean <- function(x, na.rm) {
+#   x <- x[(x > 0) & !is.na(x)]
+#   if(length(x) == 0) {
+#     m <- 0
+#   } else if(length(x) > 0) {
+#     m <- mean(x, na.rm = na.rm)
+#   }
+#   return(m)
+# }
+# cyld_agg3 <- aggregate_rast_list(fact, cyld_agg$f1, nzmean) # no zero agg
 
@@
 
-yat <- list(cyld_agg, cyld_agg2, cyld_agg3)
-## check values
-# check yields
-for(i in lev[-1]) {
-  print(paste("factor", i))
-  for(j in asnms) {
-    print(paste("...", j))
-    for(k in 1:3) {
-      #ck <- cellStats(cyld_agg$f1$gti, nzmean)
-      print(paste("...... type", k, "agg"))
-      ck <- crop_ay[[j]][, wm(mzyld, fa)]  # weighted by fraction
-      ck2 <- crop_ay$gti[, wm(mzyld, fa)]  # weighted by fraction
-      w <- values(carea_agg[[i]][[j]])  # aggregated crop fraction
-      x <- values(yat[[k]][[i]][[j]])  # yield values
-      ind <- which(x > 0)  # which are non-zero
-      x <- x[ind]  # select non zeros
-      w <- w[ind]  # select non-zeros
-      #w <- w / max(w)
-      v <- weighted.mean(x, w)
-      d <- round((ck - v) / ck, 2)
-      d2 <- round((ck2 - v) / ck2, 2)
-      print(c(d, d2))
-    }
-  }  # type 1 and type 3 aggregations consistent between datasets, across scales
-}
-rm(yat)
+# yat <- list(cyld_agg, cyld_agg2, cyld_agg3)
+# ## check values
+# # check yields
+# for(i in lev[-1]) {
+#   print(paste("factor", i))
+#   for(j in asnms) {
+#     print(paste("...", j))
+#     for(k in 1:3) {
+#       #ck <- cellStats(cyld_agg$f1$gti, nzmean)
+#       print(paste("...... type", k, "agg"))
+#       ck <- crop_ay[[j]][, wm(mzyld, fa)]  # weighted by fraction
+#       ck2 <- crop_ay$gti[, wm(mzyld, fa)]  # weighted by fraction
+#       w <- values(carea_agg[[i]][[j]])  # aggregated crop fraction
+#       x <- values(yat[[k]][[i]][[j]])  # yield values
+#       ind <- which(x > 0)  # which are non-zero
+#       x <- x[ind]  # select non zeros
+#       w <- w[ind]  # select non-zeros
+#       #w <- w / max(w)
+#       v <- weighted.mean(x, w)
+#       d <- round((ck - v) / ck, 2)
+#       d2 <- round((ck2 - v) / ck2, 2)
+#       print(c(d, d2))
+#     }
+#   }  # type 1 and type 3 aggregations consistent between datasets, across scales
+# }
+# rm(yat)
 
@@
 # with 0-removed yield aggregation (type 3 yield aggregation)
-cpagg3 <- lapply(1:length(cyld_agg3), function(x) {
-  p <- lapply(1:length(cyld_agg3[[x]]), function(y) {
-    (tareas[[x]] * carea_agg[[x]][[y]]) * cyld_agg3[[x]][[y]]
-  })
-  named_out(p, asnms)
-})
-names(cpagg3) <- names(agg_res)
+# cpagg3 <- lapply(1:length(cyld_agg3), function(x) {
+#   p <- lapply(1:length(cyld_agg3[[x]]), function(y) {
+#     (tareas[[x]] * carea_agg[[x]][[y]]) * cyld_agg3[[x]][[y]]
+#   })
+#   named_out(p, asnms)
+# })
+# names(cpagg3) <- names(agg_res)
 
@@
 # check production estimates for consistency
-pat <- list(cpagg1, cpagg3, cpaggII)
-ptype <- c("1", "3", "II")
+# pat <- list(cpagg1, cpagg3, cpaggII)
+# ptype <- c("1", "3", "II")
+pat <- list(cpagg1, cpaggII)
+ptype <- c("1", "II")
+
 # check production estimates
@@
     print(paste("...", j))
-    for(k in 1:3) {
+    for(k in 1:length(ptype)) {
       print(paste("...... type", ptype[k], "agg"))
@@
 mup <- sapply(lev, function(i) {
-  msk <- cpaggII[[i]]$gti > 0
+  msk <- cpagg1[[i]]$gti > 0
+  #msk <- cpaggII[[i]]$gti > 0
   msk[msk == 0] <- NA
-  cpmsk <- msk * cpaggII[[i]]$gti
+  cpmsk <- msk * cpagg1[[i]]$gti
   cellStats(cpmsk, mean)
@@
   v
-})
+})  # yields across scales and datasets all 3.4-3.5 kg/ha
 
@@
 ilist <- list(names(cpagg1[[1]])[-1], names(cpagg1), "gti")
-list1 <- lapply(cpaggII, function(x) x[[1]])
+list1 <- lapply(cpagg1, function(x) x[1])
 list2 <- lapply(cpagg1, function(x) x[-1])
 p_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1 agg
-list2 <- lapply(cpagg3, function(x) x[-1])
-p_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 3 agg
-list2 <- lapply(cpaggII, function(x) x[-1])
-p_diff3 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type II agg
+# list2 <- lapply(cpagg3, function(x) x[-1])
+# p_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 3 agg
+# list2 <- lapply(cpaggII, function(x) x[-1])
+# p_diff3 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type II agg
 # pmu <- as.data.table.raster(cpagg3$f1$globmu)[layer > 0, mean(layer)]
@@
 # type 2 (non-zero)
-list2 <- lapply(cyld_agg3, function(x) x[-1]) # cyld_agg for a, b/c correct
-y_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 2
+# list2 <- lapply(cyld_agg3, function(x) x[-1]) # cyld_agg for a, b/c correct
+# y_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 2
 
@@
 # here only for map plotting.
-# plot(y_diff2$sa30$f25$gti)
-pdiffs <- lapply(list(p_diff1, p_diff2, p_diff3), function(x) {
+pdiffs <- lapply(list(p_diff1), function(x) {
   l1 <- lapply(snms, function(y) {
@@
 })
-names(pdiffs) <- paste0("pd", 1:3)
-
-ydiffs <- lapply(list(y_diff1, y_diff2), function(x) {
+# names(pdiffs) <- paste0("pd", 1:3)
+names(pdiffs) <- "pd1"
+
+# ydiffs <- lapply(list(y_diff1, y_diff2), function(x) {
+ydiffs <- lapply(list(y_diff1), function(x) {
   l1 <- lapply(snms, function(y) {
@@
 })
-names(ydiffs) <- paste0("yd", 1:2)
+# names(ydiffs) <- paste0("yd", 1:2)
+names(ydiffs) <- "yd1"
 
 # disaggregate for plotting
-lev <- names(y_diff2[[1]])
+lev <- names(y_diff1[[1]])
 p_diff1d <- disaggregate_rast_list(snms, lev, pdiffs$pd1, namask)
-p_diff2d <- disaggregate_rast_list(snms, lev, pdiffs$pd2, namask)
-p_diff3d <- disaggregate_rast_list(snms, lev, pdiffs$pd3, namask)
+# p_diff2d <- disaggregate_rast_list(snms, lev, pdiffs$pd2, namask)
+# p_diff3d <- disaggregate_rast_list(snms, lev, pdiffs$pd3, namask)
 y_diff1d <- disaggregate_rast_list(snms, lev, ydiffs$yd1, namask)
-y_diff2d <- disaggregate_rast_list(snms, lev, ydiffs$yd2, namask)
+# y_diff2d <- disaggregate_rast_list(snms, lev, ydiffs$yd2, namask)
 
@@
 stats_pd1 <- dstats(p_diff1d)
-stats_pd2 <- dstats(p_diff2d)
-stats_pd3 <- dstats(p_diff3d)
+# stats_pd2 <- dstats(p_diff2d)
+# stats_pd3 <- dstats(p_diff3d)
 stats_yd1 <- dstats(y_diff1d)
-stats_yd2 <- dstats(y_diff2d)
+# stats_yd2 <- dstats(y_diff2d)
 # cbind(stats_yd1$geow[, 4], stats_yd2$geow[, 4])
@@
 
-fnms <- c(paste0("prod_bias_map_", 1:3, ".pdf"), 
-          paste0("yld_bias_map_", 1:2, ".pdf"))  # output names
+# fnms <- c(paste0("prod_bias_map_", 1:3, ".pdf"), 
+#           paste0("yld_bias_map_", 1:2, ".pdf"))  # output names
+fnms <- c("prod_bias_map_1.pdf", "yld_bias_map_1.pdf")  # output names
 # legtext <- c(rep("tons", 3), rep("tons/ha", 2))
@@
 
-From the above it is clear that the nature in which aggregation of yields are done has some impact on both yield and production estimates. There are three types of yield aggregation that can be done:
+From the above it is clear that the nature in which aggregation of yields are done has some impact on both yield and production estimates.
+There are three types of yield aggregation that can be done (most commented out now after initial checking):
 
@@
 
+Only Types 1 and A are included in further investigation. The other variants are left as accuracy checks.
+
 ## Bias statistics
@@
 # so that just one lc_union is needed
-for(i in lev) {
-  for(j in snms) {
-    print(cellStats(cyld_agg[[i]][[j]] > 0 & cyld_agg3[[i]][[j]] == 0, sum))
-    #print(cellStats(carea_agg[[i]][[j]] > 0 & cyld_agg[[i]][[j]] == 0, sum))
-  }
-}  # all zeros
+# for(i in lev) {
+#   for(j in snms) {
+#     print(cellStats(cyld_agg[[i]][[j]] > 0 & cyld_agg3[[i]][[j]] == 0, sum))
+#     #print(cellStats(carea_agg[[i]][[j]] > 0 & cyld_agg[[i]][[j]] == 0, sum))
+#   }
+# }  # all zeros
 
@@
 names(lcu) <- lev
-# plot(lc_union$f50$modmu)
+# plot(lcu$f50$modmu)
 # plot(gti_agg$f50$gti)
@@
 pd1_st <- rbind(a, b, d)
-a <- bias_stats_list(bins, awgts, lcu, p_diff2, wm, "mu", "bias", TRUE)
-b <- bias_stats_list(bins, awgts, lcu, p_diff2, wma, "mua", "bias", TRUE)
-d <- bias_stats_list(bins, awgts, lcu, p_diff2, sum, "sum", "bias", FALSE)
-pd2_st <- rbind(a, b, d)
-a <- bias_stats_list(bins, awgts, lcu, p_diff3, wm, "mu", "bias", TRUE)
-b <- bias_stats_list(bins, awgts, lcu, p_diff3, wma, "mua", "bias", TRUE)
-d <- bias_stats_list(bins, awgts, lcu, p_diff3, sum, "sum", "bias", FALSE)
-pd3_st <- rbind(a, b, d)
+# a <- bias_stats_list(bins, awgts, lcu, p_diff2, wm, "mu", "bias", TRUE)
+# b <- bias_stats_list(bins, awgts, lcu, p_diff2, wma, "mua", "bias", TRUE)
+# d <- bias_stats_list(bins, awgts, lcu, p_diff2, sum, "sum", "bias", FALSE)
+# pd2_st <- rbind(a, b, d)
+# a <- bias_stats_list(bins, awgts, lcu, p_diff3, wm, "mu", "bias", TRUE)
+# b <- bias_stats_list(bins, awgts, lcu, p_diff3, wma, "mua", "bias", TRUE)
+# d <- bias_stats_list(bins, awgts, lcu, p_diff3, sum, "sum", "bias", FALSE)
+# pd3_st <- rbind(a, b, d)
 a <- bias_stats_list(bins, awgts, lcu, y_diff1, wm, "mu", "bias", TRUE)
@@
 yd1_st <- rbind(a, b)
-a <- bias_stats_list(bins, awgts, lcu, y_diff2, wm, "mu", "bias", TRUE)
-b <- bias_stats_list(bins, awgts, lcu, y_diff2, wma, "mua", "bias", TRUE)      
-yd2_st <- rbind(a, b)
+# a <- bias_stats_list(bins, awgts, lcu, y_diff2, wm, "mu", "bias", TRUE)
+# b <- bias_stats_list(bins, awgts, lcu, y_diff2, wma, "mua", "bias", TRUE)    
+# yd2_st <- rbind(a, b)
 
@@
 p1ma <- extract_stat(lev, snms, "all", "mua", "bias", pd1_st)  # ident to p3
-p1sm <- extract_stat(lev, snms, "all", "sum", "bias", pd1_st)  # ident to p3
-p2mu <- extract_stat(lev, snms, "all", "mu", "bias", pd2_st)
-p2ma <- extract_stat(lev, snms, "all", "mua", "bias", pd2_st)
-p2sm <- extract_stat(lev, snms, "all", "sum", "bias", pd2_st)
-p3mu <- extract_stat(lev, snms, "all", "mu", "bias", pd3_st)
-p3ma <- extract_stat(lev, snms, "all", "mua", "bias", pd3_st)
-p3sm <- extract_stat(lev, snms, "all", "sum", "bias", pd3_st)
+# p1sm <- extract_stat(lev, snms, "all", "sum", "bias", pd1_st)  # ident to p3
+# p2mu <- extract_stat(lev, snms, "all", "mu", "bias", pd2_st)
+# p2ma <- extract_stat(lev, snms, "all", "mua", "bias", pd2_st)
+# p2sm <- extract_stat(lev, snms, "all", "sum", "bias", pd2_st)
+# p3mu <- extract_stat(lev, snms, "all", "mu", "bias", pd3_st)
+# p3ma <- extract_stat(lev, snms, "all", "mua", "bias", pd3_st)
+# p3sm <- extract_stat(lev, snms, "all", "sum", "bias", pd3_st)
 y1mu <- extract_stat(lev, snms, "all", "mu", "bias", yd1_st)
 y1ma <- extract_stat(lev, snms, "all", "mua", "bias", yd1_st)
-y2mu <- extract_stat(lev, snms, "all", "mu", "bias", yd2_st)
-y2ma <- extract_stat(lev, snms, "all", "mua", "bias", yd2_st)
+# y2mu <- extract_stat(lev, snms, "all", "mu", "bias", yd2_st)
+# y2ma <- extract_stat(lev, snms, "all", "mua", "bias", yd2_st)
 
@@
 jdt <- yd1_st # pd3_st # pd1_st
-bv <- "mu"
+bv <- "mua" # bv <- "mu"
 for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
@@
   rs[rs == 0] <- NA
-  l3 <- i[[x]][[y]][[1]]  
+  l3 <- abs(i[[x]][[y]][[1]])
+  # l3 <- i[[x]][[y]][[1]]
   o <- rs * l3
@@
 
-caption <- paste("Biases and mean absolute errors (MAE) in disaggregated maize",                  "yield and production (calculated from disaggregated yield",                    "and harvested area estimates)",
+caption <- paste("Biases and mean absolute errors (MAE) in disaggregated maize",
+                 "yield and production (calculated from disaggregated yield",
+                 "and harvested area estimates)",
                  "maps. Bias and MAE were normalized to their", 

### 2015-10-01 paper/app_s1.tex
A number of methods revised and redone. Mostly described in new changelog file.

@@
 
+\input{figures/cropland2011-bias.tex}
+
+\begin{figure}[ht]
+  \centering
+     \includegraphics[width = 16cm]{figures/cropland_bins.pdf} 
+%     \vspace{-3cm}
+      \caption{Number of cells within each cropland density bin at each scale of aggregation, where bins represent 5\% increment of cropland cover from 0 to 100\% (values on x-axis provide the upper limit of each bin).}
+      \label{fig:default}
+\end{figure}
+
+\input{figures/cropland2007-bias.tex}
+
 \begin{figure}[ht]
@@
 The black numbers in each plot area present the overall means of bias/MAE for each sensor-scale combination.
+The bin-wise and overall mean statistics were calculated from pooled map errors calculated from differences between the 2007 reference map and each cropland map (including all three variants--high, medium, and low--of the MODIS and GlobCover-derived cropland maps), and the 2011 reference map and each cropland map.
 }
@@
 Biases are presented in the top row, and MAEs in the bottom row.
-Statistics are calculated from pooled map errors calculated from differences between the 2007 reference map and each cropland map, and the 2011 reference map and each cropland map.}
+Statistics were calculated from pooled map errors calculated from differences between the 2007 reference map and each cropland map (including all three variants--high, medium, and low--of the MODIS and GlobCover-derived cropland maps), and the 2011 reference map and each cropland map.
+}
       \label{fig:default}
@@
 
+\FloatBarrier
+\input{figures/cropadj-mae.tex}
+
+\clearpage
 \begin{figure}[ht]
   \centering
-     \includegraphics[width = 12cm]{figures/yld_bias_map_2.pdf} 
+     \includegraphics[width = 12cm]{figures/yld_bias_map_1.pdf} 
       \caption{Errors (normalized to the reference-derived country mean) in disaggregated maize yield estimates.}
@@
   \centering
-     \includegraphics[width = 12cm]{figures/prod_bias_map_2.pdf} 
+     \includegraphics[width = 12cm]{figures/prod_bias_map_1.pdf} 
       \caption{Errors (normalized to reference-derived country mean) production estimates calculated from disaggregated maize yield and harvested area estimates.}

### 2015-10-01 paper/main.tex
A number of methods revised and redone. Mostly described in new changelog file.

@@
 
-\dropcap{T}he nature and distribution of landcover is an indicator that gives significant insight into socio-econonomic processes \cite{lambin_modelling_1997}, because human endeavors are so closely tied to how we transform land, whether it be the felling of ancient forests for farmland or erecting skycrapers.
-The vastness of our alteration of Earth’s landscapes suggests that landcover is a fundamental mediator of many environmental and social processes that drive or are affected by global change \cite{lambin_modelling_1997}, such as agricultural production and food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010,gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015,luoto_predicting_2004}, and changes in human demography \cite{linard_assessing_2010}.
+\dropcap{T}he nature and distribution of landcover provides significant insight into econonomic processes \cite{lambin_modelling_1997} because human endeavors are so closely tied to how we transform land, whether it be the felling of ancient forests for farmland or converting that farmland into office parks.
+The vastness of our alteration of Earth’s landscapes suggests that landcover is a prime mediator of many environmental and social processes that drive or are affected by global change \cite{lambin_modelling_1997}, such as agricultural production and food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010,gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015,luoto_predicting_2004}, and changes in human demography \cite{linard_assessing_2010}.
 Like any view into nature, resolution and fidelity at fine scales are the keys to unlocking more granular and mechanistic insights into these processes \cite{see_improved_2015}.
@@
 
-Global landcover data can only practically be derived from satellite imaging, but in many regions the average size class of the cover type of interest is smaller than the sensor resolution, or spectrally indistinct from other neighboring covers, which propagates classification error \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}.
-The result is that landcover maps are generally inaccurate at finer scales and disagree greatly, particularly in those parts of the world undergoing the most rapid land use changes, where these sources of bias tend to be most pronounced \cite{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011}.
+Global landcover data can only practically be derived from satellite imaging, but in many regions the cover types of interest are smaller, on average, than the sensor resolution, or spectrally indistinct from other neighboring covers, and these factors propagate classification errors \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}.
+The result is that landcover maps are generally inaccurate at finer scales and disagree substantially with one another, particularly in those parts of the world undergoing the most rapid land use changes \cite{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011}.
 
@@
 
-Just as a tall edifice cannot be built on a shaky foundation, global change science needs to be based on sound landcover data.
+Just as a building needs a solid foundation, global change science needs to be based on sound landcover data.
 There is thus an urgent need to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.
@@
 We calculated these errors for five different scales of aggregation, from the original 1 km up to 100 km, in order to calculate bias (the mean pixel-wise error) and mean absolute error (which measures the average error magnitude), and how these vary with scale.
-Next, we assessed the degree to which the average cropland cover in agricultural landscapes, which relates to their pattern, impacts error magnitude.
+Next, we assessed the degree to which the average cropland cover in agricultural landscapes, a descriptor of landcover pattern, impacts error magnitude.
 
@@
 
-We created the 1 km reference map after removing all field types classified as communal/smallholder agriculture (individual fields in this category were not mapped, thus they were removed to prevent potential cropland area overestimates) or permanent tree-crops (SI), and calculated the total cropland extent in the remaining area (1,081,000 km$^2$, or 90\% of South Africa).
+We created the 1 km reference map after removing all field types classified as communal/smallholder agriculture (individual fields in this category were not delineated, thus they were removed to prevent potential cropland area overestimates, SI) or permanent tree-crops (SI), and calculated the total cropland extent in the remaining area (1,081,000 km$^2$, or 90\% of South Africa).
 The 2011 reference map showed a cropland area of 104,310 km$^2$, which the SA-LC and GeoWiki maps overestimated by 26 and 5.8\%, respectively, and GlobCover and MODIS underestimated by 21 and 26.1\%.
 
-We then aggregated the reference and each cropland map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from the reference map at each scale of aggregation to assess error patterns (Fig.
+We then aggregated the reference and each cropland map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from each reference (2007 and 2011) map at each scale of aggregation to assess error patterns (Fig.
 1).
@@
 1).
-These average of these were 21\% for MODIS and 34\% for GlobCover at 1 km resolution (Fig.
+Averaging these errors across the whole country shows resulted
+
+The average of these were 21\% for MODIS and 34\% for GlobCover at 1 km resolution (Fig.
 S1), meaning that each map had a strong underestimation bias.
-Both maps's biases decreased with pixel aggregation, with MODIS biases falling to 8\% at 50 km, but GlobCover's was still 24\% at 100 km (SI Appendix, Fig.
+Both maps's biases decreased with pixel aggregation, with MODIS biases falling to 8\% at 50 km, whereas GlobCover's was still 24\% at 100 km (SI Appendix, Fig.
 S1).
@@
 \section{Discussion}
+This spatially comprehensive, bottom-up assessment of error and biases in landcover maps provides unique insight into their extent, causes, and consequences for understanding global change processes, made possible by a unique, high accuracy dataset that provides the truest measure of total cropland area and distribution that is currently available for this region.
+This dataset is of course not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission), while some of the cropland map error we found may have been caused by the slight temporal mismatches between the reference data and the original landcover datasets we used.
+However, our assessment (SI) suggests that these errors are small, and do not have an appreciable impact on our general findings about the degree of error and bias in landcover data, which is bolstered by previous work showing large disagreements between landcover map-based cropland area estimates and national inventory data \cite{fritz_comparison_2010}.
+
+These results suggest several guidelines for selecting and using landcover data, and contain some important implications for how understanding of global change processes based on the data, and associated policy decisions, may be affected.
+In terms of developing a base landcover map, the first important rule of thumb is that standard landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, appear to be too biased to be useful without substantial aggregation.
+If we use the standard that bias within +/-10\% is acceptable, then at least 25-100 km of aggregation is needed to sufficiently cancel out the errors in the base landcover data and subsequent first order estimates built on them (Fig.
+3 \& 4).
+The upper range of aggregation scale is necessary if a mixed pixel class becomes dominant, as in the case with GlobCover, because these lead to underestimation bias that will persist until the pixel size becomes substantially greater than the average area of landscapes that are dominated by the cover type of interest, which can be $>$1000 km$^2$ in some of South Africa's farming regions.
+
+Maps derived from higher resolution sensors, such as the SA-LC dataset, if carefully done, do not have this mixed class problem, and  sufficiently unbiased for most applications with just 1-5 km of aggregation.
+However, such datasets are typically developed for specific countries, using varying methods, and can be hard to obtain.
+For broader scale analyses, the best option is to use newer generation maps such as GeoWiki (and the GLC-Share \footnote{GLC-Share.
+www.glcn.org} datasets for other cover types) which is relatively unbiased at 1 km resolution.
+GeoWiki's lower bias comes from its process of evaluating consensus between several landcover datasets (including the other three in this study), resulting in cropland probabilities that are converted to percentages by calibrating to statistical data \cite{fritz_cropland_2011,fritz_mapping_2015}.
+This method mirrors the ensemble averaging used by other fields (e.g.
+crop \cite{asseng_uncertainty_2013}, climate \cite{giorgi_calculation_2002}, and ecological modeling \cite{araujo_ensemble_2007}) to increase prediction confidence.
+
+The statistical constraint procedure is similar to the one we used \cite[following][]{ramankutty_farming_2008}, which resulted in maize unbiased production estimates (Fig.
+3) because it eliminated bias in the adjusted cropland and harvested maps that they were built upon.
+This result, together with GeoWiki's low bias, indicate the value of fusing inventory data with remote sensing.
+However, this approach depends on the quality of inventory data, which can be suspect in many countries, particularly in Africa \cite{carletto_emperor_2013,fao_action_2013}.
+The statistical constraint also does not greatly improve map \emph{accuracy}, as evidenced by GeoWiki's 23\% mean absolute error in 1 km cropland percentage estimates (Fig.
+S1), which is only slightly more accurate than MODIS (31\%) but worse than SA-LC (11\%).
+GeoWiki is definitely most accurate among the large scale landcover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
+Statistical constraints only adjust map values at locations where cropland is identified, so their use it .
+
+Map accuracy is perhaps more important than bias for landcover maps.
+
+Broader regional implications - error higher elsewhere
 
 Main points:
-
-1. Previous studies have quantified disagreement over large areas, where
 

### 2015-10-10 Rmd/agent-allocation-bias.Rmd
October 10th, 2015 commit. Edits in changelog.txt.

@@
 dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
-dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2011))
+# dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2011))
 lc_list <- lapply(dlist_1km, function(x) gti - x) 
@@
 
+# households couldn't find enough land
+not_enough <- round(abias[var == "HHs_cant_find", snms, with = FALSE] / 
+                      (abias[var == "HHs_no_land", 
+                             lapply(.SD, function(x) HH_gen - x), 
+                             .SDcol = c(snms, "HH_gen")][, snms, with = FALSE]), 
+                    2)
+
 cexs <- seq(1, 2, 1 / 3)

### 2015-10-10 Rmd/carbon-bias.Rmd
October 10th, 2015 commit. Edits in changelog.txt.

@@
 
-extract_stat0 <- function(namevec, stats, i1, i2) {
+cb_stats$f25$geow[bvals == "m" & bin != "all", mean("forest")lapply(.SD, mean), 
+                   .SDcols = LC]
+cb_stats$f25$geow[bvals == "m", ]
+cb_stats$f25$geow[bvals == "m" & bin != "all", lapply(.SD, mean), 
+                  .SDcols = "forest"]
+
+extract_stat0 <- function(namevec, stats, i1, i2, type = "density") {
   estats <- do.call(rbind, lapply(namevec, function(i) {
@@
       vals <- unlist(sapply(x, function(y) {
+        if(type == "density") {
           v <- y[bin == i1 & bvals == i2, i, with = FALSE]
+        } else if(type == "nodensity") {
+          v <- y[bin != i1 & bvals == i2, lapply(.SD, mean), .SDcols = i]
+        }
         if(nrow(v) == 0) v <- rbindlist(list(v, as.list(NA)))
@@
 ```{r, eval=FALSE}
+# cropland area only
 a <- extract_stat0(LC, cb_stats, "all", "m")  # mean cropland only 
@@
 setcolorder(a, c(2:1, 3:ncol(a)))
-mu <- rbind(a, b)
+mu <- rbind(a, b)  # bias
 
@@
 setcolorder(b, c(2:1, 3:ncol(b)))
-mua <- rbind(a, b)
-
+mua <- rbind(a, b)  # MAE
+
+# whole country
 a <- extract_stat0(LC, cb_stats, "all", "m0")  # mean whole country
@@
 setcolorder(b, c(2:1, 3:ncol(b)))
-mu0 <- rbind(a, b)
+mu0 <- rbind(a, b)  # bias
 
@@
 setcolorder(b, c(2:1, 3:ncol(b)))
-mu0a <- rbind(a, b)
-
+mu0a <- rbind(a, b)  # MAE
+
+# density independent
+a <- extract_stat0(LC, cb_stats, "all", "m0", type = "nodensity")
+b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+setcolorder(a, c(2:1, 3:ncol(a)))
+mund <- rbind(a, b)  # bias
+
+# check
+all(round(sapply(lev_vec, function(x) {
+  mean(cb_stats[[x]]$modmu[bvals == "m0" & bin != "all", forest])
+}), 2) == mund[V2 == "forest" & V1 == "modmu", lev_vec, with = FALSE])
+
+a <- extract_stat0(LC, cb_stats, "all", "ma0", type = "nodensity")
+b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
+setcolorder(b, c(2:1, 3:ncol(b)))
+munda <- rbind(a, b)  # MAE
+
+# check
+all(round(sapply(lev_vec, function(x) {
+  mean(cb_stats[[x]]$modmu[bvals == "ma0" & bin != "all", forest])
+}), 2) == munda[V2 == "forest" & V1 == "modmu", lev_vec, with = FALSE])
+
+```
+<a href="#top">Back to top</a>
+
+### Main text version of mean carbon bias against scale
+
+#### Agricultural area bias and accuracy
+
+```{r, eval=FALSE}
 #yr <- range(all_l[, 3:8, with = FALSE], all_u[, 3:8, with = FALSE])
@@
 
+alph <- c(225, 40)
+#Ccols <- brewer.pal(n = 7, name = "Greys")
+LC2 <- c(LC[-grep("second|grass", LC)], "All")
+# xax <- seq(0, 7, 6 / (length(yax) - 2)); length(xax)
+x <- c(0, 3, 6, 9, 12, 15)
+w <- 3 / 8
+xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
+#x + xo[j] + o[k]
+xa <- sapply(x, function(x) x + xo)
+mutype <- c("mua", "mu")
+
+o <- c(0, w)
+pchs <- c("*", "+", "o", "-")
+# cx <- c(1, 0.7, 0.6, 1)
+cx <- c(2, 1.25, 1, 1)
+g1 <- "grey90"
+# pchShow()
+# ctabo
+xl <- c(-0.5, 18)
+yl <- c(-60, 70)
+shd <- c(4.5, 10.5, 16.5)
+
+pdf(full_path(p_fig, "carbon_veg_scale.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")#,
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+grid(NA, 13, lwd = 1, lty = 1)
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey50")
+for(i in 1:length(lev_vec)) {
+  for(j in 1:length(snms)) {
+    for(k in 1:length(mutype)) {
+      dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
+      pcol <- makeTransparent(cols[j], alpha = alph[k])
+      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
+      for(z in 1:length(LC2)) {
+        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
+               pch = pchs[z], type = "pt")
+      }
+    }
+  }
+}
+axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
+legend(x = 12.4, y = -40, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = -40, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = -42, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
+text(x = 12.2, y = -42, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
+legend(x = 12.1, y = 65, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
+       pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)
+dev.off()
+
+```
+<a href="#top">Back to top</a>
+
+#### Density independent bias and accuracy
+
+```{r, eval=FALSE}
+cols <- c("red", "orange3", "green4", "blue")
+lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
+pnms <- c("Forest", "Secondary", "Shrubland", "Grassland", "Sparse")
+lw <- 1.5
+
+yl <- range(mund[, 3:8, with = FALSE], munda[, 3:8, with = FALSE])
+yl <- round(yl / 10) * 10
+yax <- seq(yl[1], yl[2], 10)
+xax <- seq(1, 6, 5 / (length(yax) - 1))
+
+mutype <- c("munda", "mund")
+
+o <- c(0, w)
+pchs <- c("*", "+", "o", "-")
+cx <- c(2, 1.25, 1, 1)
+g1 <- "grey90"
+xl <- c(-0.5, 18)
+shd <- c(4.5, 10.5, 16.5)
+
+pdf(full_path(p_fig, "carbon_veg_scale_nd.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")#,
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+grid(NA, length(yax) - 1, lwd = 1, lty = 1)
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey50")
+for(i in 1:length(lev_vec)) {
+  for(j in 1:length(snms)) {
+    for(k in 1:length(mutype)) {
+      dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
+      pcol <- makeTransparent(cols[j], alpha = alph[k])
+      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
+      for(z in 1:length(LC2)) {
+        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
+               pch = pchs[z], type = "pt")
+      }
+    }
+  }
+}
+yll <- -120
+axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
+legend(x = 12.4, y = yll, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = yll, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = yll - 2, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
+text(x = 12.2, y = yll - 2, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
+legend(x = 12.1, y = 165, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
+       pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)
+dev.off()
+
+```
+<a href="#top">Back to top</a>
+
+```{r, eval = FALSE, echo=FALSE}
 # Version 1 of mean carbon bias against scale - probably for supplementals, if 
@@
 dev.off()
-
-```
-<a href="#top">Back to top</a>
-
-### Main text version of mean carbon bias against scale
-```{r, eval=FALSE}
-alph <- c(225, 40)
-#Ccols <- brewer.pal(n = 7, name = "Greys")
-LC2 <- c(LC[-grep("second|grass", LC)], "All")
-# xax <- seq(0, 7, 6 / (length(yax) - 2)); length(xax)
-x <- c(0, 3, 6, 9, 12, 15)
-w <- 3 / 8
-xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
-#x + xo[j] + o[k]
-xa <- sapply(x, function(x) x + xo)
-mutype <- c("mua", "mu")
-
-o <- c(0, w)
-pchs <- c("*", "+", "o", "-")
-# cx <- c(1, 0.7, 0.6, 1)
-cx <- c(2, 1.25, 1, 1)
-g1 <- "grey90"
-# pchShow()
-# ctabo
-xl <- c(-0.5, 18)
-yl <- c(-60, 70)
-shd <- c(4.5, 10.5, 16.5)
-
-pdf(full_path(p_fig, "carbon_veg_scale.pdf"), height = 7, width = 7)
-par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
-plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
-     ylab = "", xlab = "")#,
-for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
-polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
-grid(NA, 13, lwd = 1, lty = 1)
-lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey50")
-for(i in 1:length(lev_vec)) {
-  for(j in 1:length(snms)) {
-    for(k in 1:length(mutype)) {
-      dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
-      pcol <- makeTransparent(cols[j], alpha = alph[k])
-      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
-      for(z in 1:length(LC2)) {
-        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
-               pch = pchs[z], type = "pt")
-      }
-    }
-  }
-}
-axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
-axis(2, at = yax, labels = yax, las = 2)
-mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
-mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
-legend(x = 12.4, y = -40, legend = lcnms, pch = 15, col = cols, adj = 0,
-       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
-legend(x = 11.9, y = -40, legend = rep("", 4), pch = 15, adj = 0, 
-       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
-       cex = 0.8)
-text(x = 12.7, y = -42, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
-text(x = 12.2, y = -42, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
-legend(x = 12.1, y = 65, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
-       pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)
-dev.off()
-
-```
-<a href="#top">Back to top</a>
+```
 
@@
       caption.placement = "top", include.rownames = FALSE)
+
+# still to do: tables for density independent biases
 ```

### 2015-10-10 Rmd/yield-bias.Rmd
October 10th, 2015 commit. Edits in changelog.txt.

@@
 lev <- names(cf_pct_diff[[1]])
-snms <- names(p_diff1)
 lcu <- lapply(lev, function(x) {
@@
 # b[ol == "f10" & il == "globmu" & bin == "all"]
+lcfmua_nd <- extract_stat(lev, snms, 1:20, "mua.nm", "bias", b, 
+                          type = "nodensity")
+b[ol == "f1" & il == "sa30" & bin != "all" & bvals == "mua.nm", mean(bias)]
 

### 2015-10-10 paper/main.tex
October 10th, 2015 commit. Edits in changelog.txt.

@@
 \usepackage{float}
+%\usepackage{subfigure}
+\usepackage{wrapfig}
+%\setlength\belowcaptionskip{-2cm}
+%\setlength{\textfloatsep}{5pt plus 1.0pt minus 2.0pt}
 
@@
 There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, and carbon flux \cite{quaife_impact_2008} and human population estimates \cite{linard_assessing_2010}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}.
-One exception is a study  \cite{schmit_limitations_2006} that used a high quality, ground-collected reference map detailing farm land use parcels in central Belgium, but the number of sites and region were both fairly restricted, and the parcels were not spatially contiguous.
+One exception is a Belgian study  \cite{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover map errors bias measurements of agricultural land use patterns, but the study extent was fairly small and the validation data were discontiguous.
 
@@
 There is thus an urgent need to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.
-We address this need in this study, using a unique, high accuracy agricultural landcover map for the entire country of South Africa to conduct a spatially comprehensive, bottom-up quantification of error in several latest generation landcover maps that are widely used in global change studies.
-We use these errors to assess the extent of bias in i) landcover data, ii) how  landcover properties influence this bias, iii) how these biases change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and iv) how these biases propagate through several different forms of downstream analyses that broadly represent the global change research focus areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.
+We address this need here, using a unique, high accuracy agricultural landcover map for the entire country of South Africa to conduct a spatially comprehensive, bottom-up quantification of error in several latest generation landcover maps that are widely used in global change studies.
+We use these errors to assess the biases and accuracies of i) these landcover data, ii) how landcover properties influence map errors, iii) how bias and accuracy change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and iv) how landcover error propagates through ``downstream'' analyses that represent several major global change research areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.
 
@@
 
-The South African government commissioned a whole country cropland boundary map to enhance its annual collection of agricultural statistics \cite{fourie_better_2009}.
+The South African government commissioned a whole-country cropland boundary map to enhance its annual collection of agricultural statistics \cite{fourie_better_2009}.
 The map was made by trained workers who visually interpreted high resolution satellite imagery ($<$5 m SPOT imagery) and manually digitized field boundaries following a standardized mapping protocol.
-The resulting vectorized field maps, which were made in 2007 and updated in 2011, provide a unique, high quality reference dataset describing crop field distributions and size classes, and are 97\% accurate in distinguishing cropland from non-cropland at 200 m resolution.
+The resulting vectorized field maps, provide a unique, high quality reference dataset describing South African crop field distributions and size classes for the period 2009-2011, and are 97\% accurate in distinguishing cropland from non-cropland at 200 m resolution.
 We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create a gridded cropland reference map.
@@
 We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite[e.g.][]{fry_completion_2009}, MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013, shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
-We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates (hereafter simply the ``cropland maps''), resulting in 4 maps to compare to our reference map.
+We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates, resulting in 4 maps (hereafter simply the ``cropland maps'') to compare to our reference map.
 
@@
 \section{Quantifying Error and Bias}
-We first quantified the errors in cropland area estimates based on the pixel-wise differences between the reference map and each of the four cropland maps.
-We calculated these errors for five different scales of aggregation, from the original 1 km up to 100 km, in order to calculate bias (the mean pixel-wise error) and mean absolute error (which measures the average error magnitude), and how these vary with scale.
-Next, we assessed the degree to which the average cropland cover in agricultural landscapes, a descriptor of landcover pattern, impacts error magnitude.
+We first calculated errors in pixel-wise estimated cropland percentages by differencing the reference map with each cropland map at each of the five aggregation scales (1-100 km).
+We used these errors to calculate map bias (the mean pixel-wise error) and accuracy (the mean absolute error, and how these vary with scale.
+Next, we assessed the degree to which the average cropland cover in agricultural landscapes, a descriptor of landcover pattern, impacts map accuracy.
 
-We then investigated how landcover map error propagates through ``downstream'' analyses that are built on landcover data, and which are typical of those performed in global change assessments.
-We began by examining the biases and mean absolute errors in three ``first-order'' landcover-based analyses, in which a variable of interest is mapped onto a landcover type(s) using a simple empirical relationship.
+To investigate how landcover map error can impact downstream global change research, we quantified the biases and accuracies in five landcover-based analyses built upon our reference and cropland maps.
+The first three represented ``first-order'' analyses, in which a variable of interest is mapped onto a landcover type(s) using a simple empirical relationship.
 The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}.
@@
 
-Finally, we examined errors resulting from two second-order analyses, in which a process model draws on the cover types' values to calculate an output value.
+The other two analyses can be considered second-order, wherein a process model draws on the cover types' values to calculate an output value.
 For the first of these, we used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model to calculate monthly evapotranspiration, using the reference and cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
-In the second example, we examined how these map errors impact the parameterization of an agent-based food security model \cite{chen_dependency_2013}.
-Spatially-explicit, agent-based models are frequently employed in land change science, and require an initialization step to assign landscape resources to model agents  \cite[e.g.][]{manson_agent-based_2007,evans_multi-scale_2004,kelley_relative_2011}.
-In this case, we used cropland maps to provide the model the location and abundance of cropland, which is used to  allocate an initial share of cropland to each farm household (agent) in the model.
+In the second example, we examined how these map errors impact the land allocation process of an agent-based food security model \cite{chen_dependency_2013}.
 
 \vspace{-0.5 cm}
-\section{Bias}
-\subsection{Landcover error and bias}
+\section{Landcover maps}
+\subsection{Bias and accuracy}
 
 We created the 1 km reference map after removing all field types classified as communal/smallholder agriculture (individual fields in this category were not delineated, thus they were removed to prevent potential cropland area overestimates, SI) or permanent tree-crops (SI), and calculated the total cropland extent in the remaining area (1,081,000 km$^2$, or 90\% of South Africa).
-The 2011 reference map showed a cropland area of 104,310 km$^2$, which the SA-LC and GeoWiki maps overestimated by 26 and 5.8\%, respectively, and GlobCover and MODIS underestimated by 21 and 26.1\%.
+The 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.
 
 We then aggregated the reference and each cropland map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from each reference (2007 and 2011) map at each scale of aggregation to assess error patterns (Fig.
-1).
+1A).
 Negative pixels here represent overestimation error, while positive values indicate underestimates.
 
-\vspace{-0.25 cm}
-\begin{figure}[ht]
-\centerline{\includegraphics[width=.5\textwidth]{figures/bias_map.pdf}}
-\caption{Errors in the percent cropland estimates resulting from each of the four cropland maps relative to the reference map at different scale of pixel aggegation.
+%\vspace{-0.5 cm}
+\begin{figure*}[ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\centerline{\includegraphics[width=0.99\textwidth]{figures/figure1.pdf}}
+\caption{(A) Errors in the percent cropland estimates resulting from each of the four cropland maps relative to the reference map at different scale of pixel aggegation.
 Rows indicate the landcover map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation.
-White indicates areas with no data where communal farmlands or plantation forests were removed.}\label{afoto}
-\end{figure}
+White indicates areas with no data where communal farmlands or plantation forests were removed from analysis.
+(B) The bias (mean error) and accuracy (mean absolute error [MAE]) of each cropland map at each scale of aggregation.
+Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.
+The symbols indicate the method of calculating each metric, either by averaging across the entire country, within the union of  agricultural areas (cropland $>$0.05\%) for each reference-cropland map combination, or as the average of the averages calculated per 5\% bin of cropland cover, a measure of error that is independent of cropland density.}\label{afoto}
+\end{figure*}
+%\end{wrapfigure}
 
+%\begin{figure*}[htp]
+%\centering
+%    \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}\quad
+%     \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}
+%\end{figure*}
+
+\linenumbers
 The most pronounced errors were in the MODIS and GlobCover maps, which both underestimated cropland extent by 10-75\% in the center of the country (blue areas in Fig.
-1, and the dominant production region), and overestimated along the eastern to northern margins (red areas in Fig.
-1).
-Averaging these errors across the whole country shows resulted
+1A, and the dominant production region), and overestimated along the eastern to northern margins (red areas in Fig.
+1A).
+All maps have less than +/-5\% bias (crosses in semi-transparent bars, Fig.
+1B), at all aggregation scales, when errors are averaged across the entire country, due to the extensive non-agricultural areas in the country that were effectively discerned by all maps.
+The whole country mean absolute error (the measure of map accuracy) were a few percent higher values in most cases, but still $<$10\%, except for GlobCover ($\sim$11\%) at 1 km resolution.
+Biases increase only slightly when errors are averaged within the combined agricultural regions (areas with greater than 0.05\% cropland, indicated by circles in transparent bars in Fig.
+1B) of the reference and each cropland map, but accuracy decreases substantially, with mean absolute errors increasing to 10\% for SA-LC and around 20\% for MODIS, GlobCover, and GeoWiki at 1 km resolution (circles in solid bars in Fig.
+1B), and remain between 5-10\% for all three datasets even at 25 km of aggregation.
 
-The average of these were 21\% for MODIS and 34\% for GlobCover at 1 km resolution (Fig.
-S1), meaning that each map had a strong underestimation bias.
-Both maps's biases decreased with pixel aggregation, with MODIS biases falling to 8\% at 50 km, whereas GlobCover's was still 24\% at 100 km (SI Appendix, Fig.
-S1).
+As a third measure, we calculated bias and mean absolute error by averaging their mean per-ventile values (i.e.
+the average bias or accuracy value for areas have 0-5\% cropland, 5-10\% cropland, etc.).
+These provide density independent measures of map performance because they remove the dominance any particular level of cover has on map accuracy, which in this case was the 0-5\% cover class where map performance was best (SI).
+Calculated this way, bias and accuracy were substantially worse.
+The average bias jumped to 21\% for MODIS and 34\% for GlobCover at 1 km resolution (Fig.
+1B), meaning that each map had a strong tendency to underestimate cropland.
+This decreased with pixel aggregation, falling to 8\% at 50 km for MODIS, but remaining as high as 14\% at 100 km for GlobCover.
+SA-LC bias was smaller in magnitude, but tended towards cropland overestimates of 5-10\% at all aggregation scales.
+GeoWiki alone was relatively unbiased, but showed substantial inaccuracy with a density-independent mean absolute error of 23\%, which dropped below 10\% only above 25 km of aggregation.
+GlobCover was by far the most inaccurate, having $>$35\% mean absolute error at 1 km and 17\% at 100 km, following by MODIS (over 30\% at 1 km and 11\% at 100 km).
+SA-LC was the most accurate of all four maps with the density-independent measure, ranging from just above 10\% mean absolute error at 1 km to 5\% at 100 km.
 
-The SA-LC map uniformly overestimated cropland throughout the country (Fig.
-1), but its overall bias was relatively small, ranging from -8\% at 1 km to -6\% at 100 km  (Fig.
-S1).
-The GeoWiki map had a strongly heterogenous pattern of differences (Fig.
-1) and the smallest magnitude of bias, which changed between slight tendencies to underestimate (5\% at 1 km) and overestimate (-2 \% at 100 km,  Fig.
-S1).
+\subsection{Landscape characteristics and error}
+The density-independent measures indicate that error and bias increase with cropland density, which suggests that map error varies as a function of the typical abundance of cover in the landscape being imaged.
+To investigate the relationship between cover density and map accuracy, we calculated the mean density for agricultural pixels ($>$0.05\% cropland) in the 1 km reference map within the boundaries of 354 magisterial districts (South Africa's finest administrative unit, averaging 3,445 km$^2$; SI Figure S3), providing a landscape-scale measure representing typical cropland density.
+We then extracted the cropland map errors for these same pixels, and calculated their district-wise mean absolute errors.
 
-\subsection{Landscape characteristics and error magnitude}
-To investigate the degree to which landscape features influence landcover map errors, we extracted all pixels in agricultural areas ($>$0\% cropland) of the 1 km reference map using the boundaries of 354 magisterial districts (South Africa's finest administrative unit, which average 3,445 km$^2$ in size; SI Appendix, Fig.
-S3), and calculated the district-wise mean for these pixels, providing a measure of cropland density that was informative of the degree of mixing between cropland and other land covers at landscape scales.
-We then extracted the cropland map errors for the same pixels, and calculated their mean absolute errors for each district.
-
-A generalized additive model fit to district-level mean absolute error (log-transformed) shows that error peaks at 50\% cropland cover for all but the GlobCover map (which continued to increase with cropland cover), and is lowest when the landscape is dominated either by cropland or other cover types (Fig.
+A generalized additive model fit to district-level mean absolute error (log-transformed) shows that error peaks at 50-60\% cropland cover for all but the GlobCover map (which continued to increase with cropland cover), and is lowest when the landscape is dominated either by cropland or other cover types (Fig.
 2).
-In other words, bias is highest when cropland cover is mixed evenly with other cover types.
-The reason that GlobCover bias continued to increase with cropland density was because this dataset's dominant class identified for South African cropland was a mixture of 50-70\% cropland and other vegetation types, which resulted in the capping of GlobCover-derived cropland cover estimates at the intermediate densities defined by this class.
+In other words, accuracy is generally lowest when cropland cover is mixed evenly with other cover types.
+GlobCover's accuracy continued to decrease with cropland density because the dominant cover class contributing to the percentage cropland estimate was a mixed class defined as 50-70\% crops mingled with other vegetation, thus the maximum percentage was constrained by this mixture range.
 
@@
 \centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
-\caption{The relationship between the mean absolute error (error magnitude) in cropland maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
+\caption{The relationship between map accuracy (the mean absolute error) in cropland maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
 Prediction curves are color-coded to the different cropland maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}
@@
 
-\subsection{Map error and Tier-1 carbon estimates}
+\section{First-order analyses}
+\subsection{Tier-1 carbon estimates}
 Using the methods provided by \cite{ruesch_new_2008}, we calculated average carbon densities for African forests, secondary forests, shrublands, croplands, grasslands, and sparse habitats (semi-arid grasslands and low shrublands), and assigned cropland carbon values to map cells in proportion to their cropland cover.
@@
 The spatial patterns of errors in carbon estimates (Fig.
-S4) reflect those of cropland biases (Fig.
+S4) reflect those of the cropland maps (Fig.
 1).
@@
 
-%\vspace{-0.9 cm}
-\begin{figure}[!ht]
-\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scale.pdf}}
-\caption{Biases and mean absolute errors in carbon densities derived from cropland maps, calculated as percents relative to the reference map.
+%%\vspace{-0.9 cm}
+%\begin{figure}[!ht]
+%\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scale.pdf}}
+%\caption{Biases and mean absolute errors in carbon densities derived from cropland maps, calculated as percents relative to the reference map.
 Bias estimates (represented by symbols) fall within the semi-transparent bars, while mean absolute errors are contained in the solid bars.
@@
 Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2).}
-\label{afoto}
-\end{figure}
+%\label{afoto}
+%\end{figure}
 
@@
 3) generally followed the same patterns, but with higher magnitudes and a few important differences.
-The most notable is that Geo-Wiki, despite relatively low bias, had large mean absolute errors at 1 km, averaging 27\% across across cover types (line in solid blue bar, Fig.
+The most notable is that GeoWiki, despite relatively low bias, was highly inaccurate at 1 km, with 27\% mean absolute error across across cover types (line in solid blue bar, Fig.
 3, Table S2), which is close to the 36-37\% for GlobCover and MODIS.
@@
 3, Table S2).
-The increase in Geo-Wiki's absolute bias relative to SA-LC's can be attributed to the highly heterogeneous nature of its cropland errors (Fig.
-1), which traded between between positive and negative errors of high magnitude over shorter distances than the other three cropland maps.
+The decrease in GeoWiki's accuracy relative to SA-LC's can be attributed to the highly heterogeneous nature of its cropland errors (Fig.
+1), which alternated between high magnitude positive and negative errors over relative short distances.
 
-\subsection{Map error and gridded agricultural data}
+\subsection{Gridded yield and production estimates}
 The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived cropland area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}.
@@
 %\vspace{-1 cm}
-\begin{figure}[!ht]
-\centerline{\includegraphics[width=.5\textwidth]{figures/yield_bias_scale.pdf}}
-\caption{Biases and mean absolute errors in disaggregated maize yield and production estimates.
-Bias estimates (represented by symbols) fall within the semi-transparent bars, while mean absolute errors are contained in the solid bars.
-Bar colors are coded to specific cropland map, while the symbols indicate which the values for the different variables (production and yield).}
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=1\textwidth]{figures/figure3.pdf}}
+\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in (A) carbon densities derived from cropland maps and (B) in disaggregated maize yield and production estimates.
+Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.
+For the carbon maps (A), symbols indicate the cover type used to calculate cropland-adjacent carbon density: the line represents the mean bias/MAE across all 5 cover types; shrubland and grassland values (near zero) and secondary forest (close to forest values) are not shown for clarity (see SI).
+For (B) symbols code the different variables (production and yield), normalized to their respective means.
+Values in A and B were calculated within the union of agricultural areas (cropland $>$0.05\%) for each reference-cropland map combination.}
 \label{afoto}
-\end{figure}
+\end{figure*}
 
@@
 
-\subsection{Map error and evapotranspiration}
-Compared to carbon and crop related examples, bias and mean absolute errors in evapotranspiration (ET) calculated using the VIC model were small and averaged to less than +/-1\%.
-However, there were several hotspots of discrepancy evident in the error maps (Fig.
+\section{Second-order analyses}
+\subsection{Evapotranspiration estimates}
+Compared to carbon and crop related examples, bias and accuracy in evapotranspiration (ET) calculated using the VIC model were small and averaged to less than +/-1\%.
+However, there were several error hotspots in the resulting ET difference maps (Fig.
 6).
@@
 
-\subsection{Initialization errors in agent-based models}
-We used an agent-based model (ABM) of food security that represents the interactions between hundreds of individual farming households over multiple seasons \cite{chen_dependency_2013}.
+\begin{figure}[!ht]
+\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
+\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four cropland maps.}\label{afoto}
+\end{figure}
+
+\subsection{Initializing an agent-based model}
+Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}.
+A common feature of such ABMs is that they need to be calibrated against data describing the characteristics of land users, including an initialization step to assign land resources to ``agents'' representing the land users, wherein the simulated landscape pattern and distribution of agent resources matches those in the real world.
+In our example, we used an ABM of household food security that simulate the interactions between many individual farming households (the agents) and their environment over multiple seasons \cite{chen_dependency_2013}.
+We used cropland maps to provide the model the location and abundance of cropland, which is used to allocate an initial share of cropland to each simulated household.
 Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g.
-districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) that are needed to represent the different land units of single farmers.
+districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers.
 To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig.
-S8) in the eastern part of the country, having between 28-45\% of their areas devoted to cropland, according to the reference map.
-
-We disaggregated the cropland percentages in all maps to binary cropland/non-cropland cover types with 100 m resolution, which matches the typical field size (1 ha) for smallholder farmers in household survey data (collected in Zambia) used in developing the agent-based model \cite{chen_dependency_2013}.
-These surveys found mean household crop field area to be 2 ha, which we divided into reference cropland areas to estimate the total households within each district.
-We then initialized the model by assigning each household agent two cropland pixels.
-In order to emulate the natural groupings of communities, the model only assigns a household fields that are within 1.5 km of other agents' fields, provided those pixels were not previously allocated to another agent.
-The model thus iteratively grows ``communities'' until all households are assigned cropland, or all available cropland is allocated.
+S8) in the eastern part of the country with 28-45\% of their areas devoted to cropland.
+The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity.
+The number of households and the cropland area per household is derived from survey data of communities where all cropland is owned.
+The model is thus considered adequately initialized when all households are allocated their appropriate area of cropland, and all cropland is occupied.
 
 We used the reference map and each cropland map to separately initialize the model, and compared the agent allocation results to assess how cropland map errors impacted the initialization process.
-[insert importance here].
-We examined two metrics, the first being the number of agents that were not assigned fields.
-Here there was a one-to-one relationship between the percentage of cropland area underestimation and the percentage of households left without farmland (Fig.
+We examined three variable, the first being the number of agents that were not assigned fields, the second the amount of cropland left unallocated, and the third the area of land deficit, or the amount of land that should have been assigned to households but wasn't.
+For the first variable, there was a one-to-one relationship between the percentages by which cropland was underestimated and households that could not be assigned fields (Fig.
 6, left panel).
 The most extreme examples occurred when MODIS cropland initialized the ABM in districts 1 and 2, where $\sim$85\% of agents did not receive cropland.
-
-All households were assigned fields when maps overestimated cropland extent (GeoWiki, SA-LC), but in these cases the percent of cropland left unallocated--our second metric for assessing cropland error impacts--matched the size of the overestimate (e.g.
+All households were assigned fields when total cropland area was overestimated (GeoWiki, SA-LC), but in these cases the area of cropland allocated to no one (the second variable) was proportional to the size of the overestimate (e.g.
 $\sim$20\% for SA-LC, Fig.
@@
 
+The last measure, land deficit, increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2--and would become infinite in the case of a 100\% underestimate.
+
 %randomly siting 100 household agents within each district, and allocating the nearest two cropland pixels to each household.
@@
 %\FloatBarrier
+
 \begin{figure}[!ht]
-\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
-\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four cropland maps.}\label{afoto}
-\end{figure}
-
-\begin{figure}[ht]
 \centerline{\includegraphics[width=.5\textwidth]{figures/agent-bias.pdf}}
@@
 \section{Discussion}
-This spatially comprehensive, bottom-up assessment of error and biases in landcover maps provides unique insight into their extent, causes, and consequences for understanding global change processes, made possible by a unique, high accuracy dataset that provides the truest measure of total cropland area and distribution that is currently available for this region.
+This spatially comprehensive, bottom-up assessment of landcover map bias and inaccuracy and provides unique insight into their extent, causes, and consequences for understanding global change processes, made possible by a unique, high accuracy dataset that likely provides the truest measure of total cropland area and distribution that is currently available for this region.
 This dataset is of course not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission), while some of the cropland map error we found may have been caused by the slight temporal mismatches between the reference data and the original landcover datasets we used.
-However, our assessment (SI) suggests that these errors are small, and do not have an appreciable impact on our general findings about the degree of error and bias in landcover data, which is bolstered by previous work showing large disagreements between landcover map-based cropland area estimates and national inventory data \cite{fritz_comparison_2010}.
+However, our assessment (SI) suggests that these errors are small, and do not appreciably impact our findings, which is bolstered by previous work showing the large scale of disagreements between landcover map-based cropland area estimates and national inventory data \cite{fritz_comparison_2010}.
 
-These results suggest several guidelines for selecting and using landcover data, and contain some important implications for how understanding of global change processes based on the data, and associated policy decisions, may be affected.
-In terms of developing a base landcover map, the first important rule of thumb is that standard landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, appear to be too biased to be useful without substantial aggregation.
+Our results suggest several guidelines for using landcover data in global change research, and contain some important implications for how understanding of global change processes based on the data, and associated policy decisions, may be affected.
+In terms of developing a base landcover map, the first rule of thumb is that standard landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, appear to be too biased to be useful without substantial aggregation.
 If we use the standard that bias within +/-10\% is acceptable, then at least 25-100 km of aggregation is needed to sufficiently cancel out the errors in the base landcover data and subsequent first order estimates built on them (Fig.
@@
 
-Maps derived from higher resolution sensors, such as the SA-LC dataset, if carefully done, do not have this mixed class problem, and  sufficiently unbiased for most applications with just 1-5 km of aggregation.
+Maps derived from higher resolution sensors, such as the SA-LC dataset, if carefully done, do not have this mixed class problem, and are sufficiently unbiased for most applications with just 1-5 km of aggregation.
 However, such datasets are typically developed for specific countries, using varying methods, and can be hard to obtain.
@@
 
-The statistical constraint procedure is similar to the one we used \cite[following][]{ramankutty_farming_2008}, which resulted in maize unbiased production estimates (Fig.
-3) because it eliminated bias in the adjusted cropland and harvested maps that they were built upon.
+GeoWiki's statistical constraint procedure is similar to the one we used \cite[following][]{ramankutty_farming_2008}, which produced unbiased maize production estimates (Fig.
+3) by eliminating bias in the adjusted cropland and harvested area maps that they were built upon.
 This result, together with GeoWiki's low bias, indicate the value of fusing inventory data with remote sensing.
-However, this approach depends on the quality of inventory data, which can be suspect in many countries, particularly in Africa \cite{carletto_emperor_2013,fao_action_2013}.
+However, this method depends on the quality of inventory data, which are often suspect, particularly in Africa \cite{carletto_emperor_2013,fao_action_2013}.
 The statistical constraint also does not greatly improve map \emph{accuracy}, as evidenced by GeoWiki's 23\% mean absolute error in 1 km cropland percentage estimates (Fig.
@@
 
+We disaggregated the cropland percentages in all maps to binary cropland/non-cropland cover types with 100 m resolution, which matches the typical field size (1 ha) for smallholder farmers in household survey data (collected in Zambia) used in developing the agent-based model \cite{chen_dependency_2013}.
+The surveys provided the mean cropland area per household (2.2 ha) and frequencies distribution of cropland area holdings across households (e.g.
+how many households have 1 ha, 2 ha, etc.).
+We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area, and preserved the cropland area distributions by multiplying the total number of households by the frequencies.
+We then initialized the model, which takes a weighted (by cropland area frequency) random draw of 100 households and places these within the district, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household.
+This process is iterated until all households are assigned cropland, or all available cropland is allocated.
+
 \section{Digital RCD Analysis}

### 2015-10-10 R/bias-stats.R

@@ -103,27 +103,4 @@ bias_stats_list <- function(rbinsl, rwgtsl, rmaskl, revall, fun, stnm, evalnm,
   return(DTO)
 }
 
-#' Extract specific statistics from bias_stats data.table
-#' @param ols vector with value(s) to subset from outer list (ol) variable
-#' @param ils vector with value(s) to subset from inner list (il) variable
-#' @param bins vector naming the statistical aggregation level to subset
-#' @param stat statistic to subset
-#' @param val Column header containing the value of interest (N or bias stat)
-#' @param dt Name of the bias_stat produced data.table
-#' @return Subset of statistics from bias_stat data.table
-#' @export
-extract_stat <- function(ols, ils, bins, stat, val, dt) {
-  ii <- rbindlist(lapply(ols, function(i) {
-    jj <- rbindlist(lapply(ils, function(j) {
-      v <- dt[ol == i & il == j & bin == bins & bvals == stat, val, 
-              with = FALSE]
-      if(nrow(v) == 0) v <- rbindlist(list(v, as.list(NA)))
-      vd <- data.table(v)
-    }))
-    jjdt <- cbind("ol" = i, "il" = ils, jj)
-  }))
-  setkey(ii, "ol")
-  return(ii)
-}
-
 

### 2015-10-10 R/bias-statsw.R

@@ -1,0 +1,115 @@
+
+#' Calculate bias statistics from simple list of rasters of spatial errors
+#' @description A variant of bias_stats in which error can be doubly weighted, 
+#' singly weighted, or not at all
+#' @param ref Raster containing reference weights
+#' @param awgts Raster defining area weights, for aggregation
+#' @param rerror List of error rasters
+#' @param evalnm Name of value encoded in reval's cells (for column header)
+#' @param fun Function to evaluate--should be able to handle weights
+#' @param rweight Apply reference weighting or not
+#' @param aweight Apply area weighting or not
+#' @param rnd Rounding value to apply to statistic
+#' @param trim_wgt Remove column of weights from output table
+#' @return A data.table of output statistics under evalnm
+#' @details This can calculate weighted or unweighted statistics
+#' @examples 
+#' double weighting, proof
+#' a <- c(1, 0.5)
+#' w1 <- c(0.5, 1) 
+#' w2 <- c(1, 0.5)
+#' sum(a * (w1 * w2) / sum(w1 * w2)) == mean(a)  # equal to straight mean
+#' 
+#' # equal to weighted.mean
+#' a <- c(1, 0.5)
+#' w1 <- c(0.5, 1) 
+#' # w2 <- c(1, 1)
+#' w2 <- c(2, 2)
+#' sum(a * (w1 * w2) / sum(w1 * w2)) == weighted.mean(a, w1)
+#' 
+#' # toy data
+#' ref <- rerror <- awgts <- raster(nrow = 10, ncol = 10)
+#' ref[] <- runif(ncell(ref))
+#' rerror[] <- runif(ncell(ref))
+#' rerror <- rerror - ref
+#' names(rerror) <- "rerror"
+#' awgts[] <- sample(5:10, size = ncell(ref), replace = TRUE)
+#' wm <- function(x, w) stats::weighted.mean(x, w)
+#' # no weight by ref, check against raster version
+#' bias_statsw(ref, awgts, rerror, fun = wm, evalnm = "rerror", 
+#'            rweight = FALSE, rnd = 4) ==
+#'  round(weighted.mean(getValues(rerror), getValues(awgts)), 4)
+#' 
+#' # no weight by area, check against raster version
+#' bias_statsw(ref, awgts, rerror, fun = wm, evalnm = "rerror", aweight = FALSE, 
+#'             rnd = 4) == 
[…]

### 2015-10-10 R/extract_stat.R

@@ -1,0 +1,31 @@
+#' Extract specific statistics from bias_stats data.table
+#' @param ols vector with value(s) to subset from outer list (ol) variable
+#' @param ils vector with value(s) to subset from inner list (il) variable
+#' @param bins vector naming the statistical aggregation level to subset
+#' @param stat statistic to subset
+#' @param val Column header containing the value of interest (N or bias stat)
+#' @param dt Name of the bias_stat produced data.table
+#' @param type "density" for normal mode, "nodensity" for density independent
+#' @return Subset of statistics from bias_stat data.table
+#' @details The type parameter allows for density independent and density 
+#' dependent statistics to be extracted. If density independent, the mean is 
+#' calculated as the mean of bin-wise means.  
+#' @export
+extract_stat <- function(ols, ils, bins, stat, val, dt, type = "density") {
+  ii <- rbindlist(lapply(ols, function(i) {
+    jj <- rbindlist(lapply(ils, function(j) {
+      if(type == "density") {
+        v <- dt[ol == i & il == j & bin == bins & bvals == stat, val, 
+                with = FALSE]
+      } else if(type == "nodensity") {
+        v <- dt[ol == i & il == j & bin != "all" & bvals == stat, 
+                lapply(.SD, mean), .SDcols = val]
+      }
+      if(nrow(v) == 0) v <- rbindlist(list(v, as.list(NA)))
+      vd <- data.table(v)
+    }))
+    jjdt <- cbind("ol" = i, "il" = ils, jj)
+  }))
+  setkey(ii, "ol")
+  return(ii)
+}

### 2015-10-17 Rmd/carbon-bias.Rmd
17/10/2015 changes, see changelog

@@
 ### Calculate weighted bias statistics
+
+#### Primary method
+
+Weighted by __actual__ cropland percentage, and then, with aggregation, by number of pixels being aggregated.
+
 ```{r, eval=FALSE}
@@
 # }
+
 wm <- function(x, w) stats::weighted.mean(x, w)
@@
 
-# check constancy of < 1/2% cropland being excluded
-areas <- sapply(lc_agg, function(x) res(x[[1]])[1]^2 / 10000)
-plot(areas, (areas * 0.005))  # scales
-(areas * 0.005)  # 1/2 to 5000 ha
-
+# Bias
+cb_statsw_mu <- rbindlist(lapply(lev_vec, function(x) {
+  il <- rbindlist(lapply(snms, function(y) {
+    ref <- gti_agg[[x]][[1]]
+    awgts <- area_wgts[[x]][[1]]
+    rerror <- c_pct_diff[[x]][[y]]
+    bstats <- bias_statsw(ref, awgts, rerror, LC, wm, "Bias", rnd = 2, 
+                          rweight = TRUE, aweight = TRUE, trim_wgt = TRUE)
+    cbind("map" = y, bstats)
+  }))
+  # named_out(ol, snms)
+  cbind("ol" = x, il)
+}))
+
+# calculate mean across cover types
+cb_statsw_mu <- cbind(cb_statsw_mu, 
+                      "All" = apply(cb_statsw_mu[, LC, with = FALSE], 1, mean))
+
+# Accuracy
+cb_statsw_mua <- rbindlist(lapply(lev_vec, function(x) { # x <- "f1"
+  il <- rbindlist(lapply(snms, function(y) { # y <- "sa30"
+    ref <- gti_agg[[x]][[1]]
+    awgts <- area_wgts[[x]][[1]]
+    rerror <- c_pct_diff[[x]][[y]]
+    bstats <- bias_statsw(ref, awgts, rerror, LC, wma, "MAE", rnd = 2, 
+                          rweight = TRUE, aweight = TRUE, trim_wgt = TRUE)
+    cbind("map" = y, bstats)
+  }))
+  # named_out(ol, snms)
+  cbind("ol" = x, il)
+}))
+
+# calculate mean across cover types
+cb_statsw_mua <- cbind(cb_statsw_mua, 
+                       "All" = apply(cb_statsw_mua[, LC, with=FALSE], 1, mean))
+# rowMeans(as.data.frame(cb_statsw_mua[1, ])[, LC])
+
+# check error stats - do they match alternate approaches?
+for(i in c("f1", "f25", "f10")) { # i <- "f1"
+  for(j in c("sa30", "modmu", "geow")) { # j <- "modmu"
+    for(k in c("forest", "shrub", "sparse", "second")) {  # k <- "second"
+      print(paste("cross-checking calculations in", i, j, k))
+      ref <- gti_agg[[i]][[1]]
+      awgts <- area_wgts[[i]][[1]]
+      rerror <- c_pct_diff[[i]][[j]]
+      a1 <- bias_statsw(ref, awgts, rerror, LC, wm, "mu", 
+                        aweight = FALSE)[, get(k)]
+      a2 <- bias_statsw(ref, awgts, rerror, LC, wm, "mu", 
+                        rweight = FALSE, aweight = FALSE)[, get(k)]
+      a3 <- bias_statsw(ref, awgts, rerror, LC, wm, "mu")[, get(k)]
+      c1 <- getValues(c_pct_diff[[i]][[j]][[k]])
+      w1 <- getValues(gti_agg[[i]][[1]])
+      w2 <- getValues(area_wgts[[i]][[1]])
+      print("...non-area weighted mean matches?")
+      print(a1 == round(weighted.mean(c1, w1, na.rm = TRUE), 2))
+      print("...totally unweighted mean matches?")
+      print(a2 == round(cellStats(rerror[[k]], mean), 2))
+      print("...double weighted mean matches?")
+      print(a3 == round(weighted.mean(c1, w1 * w2, na.rm = TRUE), 2))
+    }
+  }
+}
+
+````
+
+#### Secondary methods
+
+```{r, eval = TRUE}
 # selection variables
@@
 bvals <- "mu"
+
+# check constancy of < 1/2% cropland being excluded
+areas <- sapply(lc_agg, function(x) res(x[[1]])[1]^2 / 10000)
+plot(areas, (areas * 0.005))  # scales
+(areas * 0.005)  # 1/2 to 5000 ha
 
@@
 
-cb_stats$f25$geow[bvals == "m" & bin != "all", mean("forest")lapply(.SD, mean), 
-                   .SDcols = LC]
+cb_stats$f25$geow[bvals == "m" & bin != "all", lapply(.SD, mean), .SDcols = LC]
 cb_stats$f25$geow[bvals == "m", ]
@@
 
-### Plot means against scale, supplemental version 
+### Extract statistics for supplemental version 
 ```{r, eval=FALSE}
@@
 
-### Main text version of mean carbon bias against scale
-
-#### Agricultural area bias and accuracy
-
-```{r, eval=FALSE}
-#yr <- range(all_l[, 3:8, with = FALSE], all_u[, 3:8, with = FALSE])
-# yl <- range(all_mus[, 3:8, with = FALSE])
-
+### Plot mean carbon bias and accuracy against scale
+
+#### Bias and accuracy weighted by agricultural area
+
+```{r, eval = FALSE}
 cols <- c("red", "orange3", "green4", "blue")
@@
 
-yl <- range(mu[, 3:8, with = FALSE], mua[, 3:8, with = FALSE])
-yl <- round(yl / 10) * 10
-yax <- seq(yl[1], yl[2], 10)
+# yl <- range(cb_statsw_mu[, LC, with = FALSE], 
+#             cb_statsw_mua[, LC, with = FALSE])
+# yl <- round(yl / 10) * 10
+yl <- c(-290, 290)
+yax <- seq(yl[1], yl[2], 20)
 xax <- seq(1, 6, 5 / (length(yax) - 1))
@@
 alph <- c(225, 40)
-#Ccols <- brewer.pal(n = 7, name = "Greys")
 LC2 <- c(LC[-grep("second|grass", LC)], "All")
-# xax <- seq(0, 7, 6 / (length(yax) - 2)); length(xax)
 x <- c(0, 3, 6, 9, 12, 15)
@@
 xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
-#x + xo[j] + o[k]
 xa <- sapply(x, function(x) x + xo)
-mutype <- c("mua", "mu")
+mutype <- c("cb_statsw_mua", "cb_statsw_mu")
 
@@
 pchs <- c("*", "+", "o", "-")
-# cx <- c(1, 0.7, 0.6, 1)
 cx <- c(2, 1.25, 1, 1)
 g1 <- "grey90"
-# pchShow()
+
 # ctabo
+xl <- c(-0.5, 18)
+# yl <- c(-200, 200)
+shd <- c(4.5, 10.5, 16.5)
+
+pdf(full_path(p_fig, "carbon_veg_scalew.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")#,
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+abline(h = yax, v = NULL, col = "grey80", lty = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey50")
+for(i in 1:length(lev_vec)) { # i <- 1; j <- 1; k <- 1
+  for(j in 1:length(snms)) {
+    for(k in 1:length(mutype)) {
+      dat <- get(mutype[k])[ol == lev_vec[i] & map == snms[j], LC2, with=FALSE]
+      pcol <- makeTransparent(cols[j], alpha = alph[k])
+      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
+      for(z in 1:length(LC2)) {
+        lpfunc(x[i] + xo[j] + o[k], dat[, z, with = FALSE], col = "black", 
+               size = cx[z], pch = pchs[z], type = "pt")
+      }
+    }
+  }
+}
+axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
+legend(x = 12.4, y = -210, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = -210, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = -212, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
+text(x = 12.2, y = -212, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
+legend(x = 12.1, y = 260, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
+       pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)
+dev.off()
+
+```
+
+#### Agricultural area bias and accuracy
+
+```{r, eval=FALSE}
+#yr <- range(all_l[, 3:8, with = FALSE], all_u[, 3:8, with = FALSE])
+# yl <- range(all_mus[, 3:8, with = FALSE])
+
+# yl <- range(mu[, 3:8, with = FALSE], mua[, 3:8, with = FALSE])
+# yl <- round(yl / 10) * 10
+yax <- seq(yl[1], yl[2], 10)
+xax <- seq(1, 6, 5 / (length(yax) - 1))
+
+mutype <- c("mua", "mu")
+
 xl <- c(-0.5, 18)
@@
 ```{r, eval=FALSE}
-cols <- c("red", "orange3", "green4", "blue")
-lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
-pnms <- c("Forest", "Secondary", "Shrubland", "Grassland", "Sparse")
-lw <- 1.5
-
 yl <- range(mund[, 3:8, with = FALSE], munda[, 3:8, with = FALSE])
@@
 
-```{r, eval = FALSE, echo=FALSE}
-# Version 1 of mean carbon bias against scale - probably for supplementals, if 
-# retained
-LC2 <- c(LC, "all")
-pdf(full_path(p_fig, "carbon_veg_scale.pdf"), height = 7, width = 5)
-par(mfrow = c(3, 2), mar = c(1, 0, 1, 1.25), mgp = c(1, 0.5, 0), tcl = -0.3,  
-    oma = c(6, 4, 1, 0))
-for(f in 1:length(LC2)) {
-  plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
-       ylab = "", xlab = "", 
-       panel.first = grid(NA, length(yax), lwd = 1, lty = 1))
-  #grid(NA, NULL, lwd = 1, lty = 1)
-  mtext(side = 3, text = pnms[f], line = 0)
-  if(f == 5) axis(1, at = 1:6, labels = c(1, fact))
-  if(f %in% c(1, 3, 5)) axis(2, at = yax, labels = yax, las = 2)
-  if(f %in% c(2, 4)) axis(2, at = yax, labels = rep("", length(yax)), las = 2)
-  for(i in 1:4) {
-    lines(1:6, mu[V1 == snms[i] & V2 == LC2[f], lev_vec, with = FALSE], 
-          col = cols[i], lty = 2, lwd = lw)
-  }
-  for(i in 1:4) {
-    lines(1:6, mua[V1 == snms[i] & V2 == LC2[f], lev_vec, with = FALSE], 
-          col = cols[i], lty = 1, lwd = lw)
-  }
-}
-# plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
-#      ylab = "", xlab = "", 
-#      panel.first = grid(NA, length(yax), lwd = 1, lty = 1))
-# for(i in 1:4) {
-#  lines(1:6, mu[V1 == snms[i], lev_vec, with = FALSE], col = cols[i], 
-#        lty = 2, lwd = lw)
-# }
-# for(i in 1:4) {
-#   lines(1:6, mua[V1 == snms[i], lev_vec, with = FALSE], col = cols[i],
-#         lty = 1, lwd = lw)
-# }
-mtext(side = 3, text = "Vegetation mean", line = 0)
-axis(1, at = 1:6, labels = c(1, fact))
-axis(2, at = yax, labels = rep("", length(yax)), las = 2)
-mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 1.5)
-mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 2)
-par(xpd = NA)
-x <- grconvertX(0.7, from = "ndc", to = "user")
-y <- grconvertY(0.075, from = "ndc", to = "user")
-legend(x = x, y = y, legend = c("Bias", "MAE"), lty = c(2, 1), 
-       lwd = 2, bty = "n", cex = 1.2)
-x <- grconvertX(0.1, from = "ndc", to = "user")
-y <- grconvertY(0.11, from = "ndc", to = "user")
-legend(x = x, y = y, legend = mcap, lty = 1, col = cols, lwd = 2, bty = "n",
-       cex = 1.2)
-dev.off()
-```
-
-### Supplementary mean bias tables
+### Supplementary bias/accuracy tables
 ```{r, eval=FALSE}
+# Statistics from cropland area-weighted measures, reshaped
+cb_stats_s <- lapply(rev(mutype), function(i) {
+  i0 <- do.call(rbind, lapply(snms, function(x) {
+    i1 <- do.call(rbind, lapply(c(LC, "All"), function(y) {
+      i2 <- do.call(cbind, lapply(lev_vec, function(z) {
+        get(i)[ol == z & map == x, get(y)]
+      }))
+      cbind.data.frame(y, i2)
+    }))
+    i1 <- cbind.data.frame(x, i1)
+    colnames(i1) <- c("Map", "Cover", lev_vec)
+    data.table(i1)
+  }))
+})
+names(cb_stats_s) <- c("Bias", "Accuracy")
+muws <- copy(rbind(cbind(Metric = "Bias", cb_stats_s$Bias), 
+                   cbind(Metric = "MAE", cb_stats_s$Accuracy)))
+setnames(muws, lev_vec, paste(c(1, fact), "km"))
+setkeyv(muws, c("Map", "Cover"))
+for(i in 1:length(LC)) muws[Cover == LC[i], Cover := pnms[i]]
+setkey(muws, "Map")
+for(i in 1:length(snms)) muws[Map == snms[i], Map := mcap[i]]
+
 # Statistics from non-croplands areas included in means, for supplementals
@@
 # still to do: tables for density independent biases
+
+# Bind all three tables together
+fulls <- rbind(cbind("Region" = "Density weighted", muws), 
+               cbind("Region" = "Country", mu0s),
+               cbind("Region" = "Agricultural", mus))
+
+caption <- paste("Biases and mean absolute errors for each of the cropland", 
+                 "maps across aggregation scales and each possible landcover", 
+                 "type sharing the pixel with cropland. Three variants", 
+                 "of the bias/accuracy values are provided, indicated by the:",
+                 "Region column: 1) weighted by reference cropland density", 
+                 "(Density weighted); 2) across the entire country (Country);",
+                 "3) for agricultural areas only (Agricultural, i.e.", 
+                 "$>$0.05\\% cropland).")
+fulls_xtab <- xtable(fulls[order(Region, Metric, Cover)], digits = 1, 
+                     caption = caption)
+print(fulls_xtab, type = "latex", file = "paper/figures/C-bias-accuracy.tex", 
+      tabular.environment = "longtable", floating = FALSE, 
+      caption.placement = "top", include.rownames = FALSE)
+
 ```

### 2015-10-17 Rmd/compare-landcover.Rmd
17/10/2015 changes, see changelog

@@
 This section of the analysis calculates the differences between the GTI cropland percentages and those from the various landcover products.
-It draws from results generated by the [cropland pre-processing](cropland-preprocess.Rmd).
-(Note: html needed).
+It draws from results generated by the [cropland pre-processing](cropland-preprocess.html).
+Bias statistics calculated here are primarily for the supplementals.
 
@@
 caption <- paste("Biases and mean absolute errors (MAE) in cropland maps",
-                 "relative to the 2011 reference map for each aggregation scale",
-                 "calculated over the entire country, for the union of",
+                 "relative to the 2011 reference map for each aggregation",
+                 "scale calculated over the entire country, for the union of",
                  "agricultural regions (cropland $>$ 0), and as density", 
                  "independent means, wherein the mean bias/MAE values",
-                 "for each of 20 cropland cover classes",
-                 "(representing 5\\% increments of cover 0\\% to 100\\% defined",
+                 "for each of 20 cropland cover classes (representing",
+                 "5\\% increments of cover 0\\% to 100\\% defined",
                  "by the reference map) were calculated and then averaged.")

### 2015-10-17 Rmd/cropland-bias.Rmd
17/10/2015 changes, see changelog

--- 2015-10-01 Rmd/landcover-bias-plots.Rmd
@@
 ---
-title: "Plotting cropland biases"
+title: "Cropland Bias and Accuracy"
 author: "Lyndon Estes"
@@
 
-# Plotting bias figures
+# Overview
+
+The following provide further analyses and plots of bias and accuracy metrics calculated from spatial error maps created in [compare-landcover](compare-landcover.html) (which includes some supplementary metrics),
+
 # Spatial biases
@@
 Plot maps of biases at different resolutions, for publication output (and display here).
-To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and then masked before plotting.
-```{r, eval = FALSE}
+To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and masked before plotting.
+```{r, ethen val = FALSE}
 namask <- raster(full_path(p_data, "namask.tif"))
@@
 wm <- function(x, w) stats::weighted.mean(x, w, na.rm = TRUE)
-# wma <- function(x, w) stats::weighted.mean(abs(x), w)
+wma <- function(x, w) stats::weighted.mean(abs(x), w)
 
@@
 
-In relation to calculation method (across country, agricultural region only, or bin-wise (cropland density classes) mean.
+## Primary method
+
+Weighted by __actual__ cropland percentage, and then, with aggregation, by number of pixels being aggregated.
+
+```{r, eval = FALSE}
+# Create weights maks
+fact <- c(5, 10, 25, 50, 100)
+namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
+# cellStats(!is.na(gti$g2011), sum)
+# cellStats(namask, sum)
+
+namask2 <- !is.na(namask)
+awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
+awgts <- lapply(awgts, function(x) x[[1]])
+
+# Subset difference grids, for 2011
+snms <- c("sa30", "globmu", "modmu", "geow")
+err2011[[1]]
+err2011 <- lapply(lev, function(x) {
+  sapply(dlist_act[snms], function(y) list(y[[x]]$g2011))
+})
+names(err2011) <- lev
+
+gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters
+gti_agg11 <- lapply(gti_agg, function(x) x$g2011)
+
+# calculate bias/MAE for data
+a <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wm, "mu")
+b <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wma, "mua")
+bacc2011 <- rbind(a, b)
+
+# check error stats - do they match alternate approaches?
+for(i in c("f1", "f25", "f10")) {
+  for(j in c("sa30", "modmu", "geow")) { 
+    print(paste("cross-checking calculations in", i, j))
+    a1 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
+                      "mu", aweight = FALSE)[, j, with = FALSE]
+    a2 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
+                      "mu", rweight = FALSE, aweight = FALSE)[,j, with=FALSE]
+    a3 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
+                      "mu")[, j, with = FALSE]
+    c1 <- getValues(err2011[[i]][[j]])
+    w1 <- getValues(gti_agg11[[i]])
+    w2 <- getValues(awgts[[i]])
+    print("non-area weighted mean matches?")
+    print(a1 == round(weighted.mean(c1, w1, na.rm = TRUE), 2))
+    print("totally unweighted mean matches?")
+    print(a2 == round(cellStats(err2011[[i]][[j]], mean), 2))
+    print("double weighted mean matches?")
+    print(a3 == round(weighted.mean(c1, w1 * w2, na.rm = TRUE), 2))
+  }
+}
+
+# Subset difference grids, for 2007
+snms <- c("sa30", "globmu", "modmu", "geow")
+err2007 <- lapply(lev, function(x) {
+  sapply(dlist_act[snms], function(y) list(y[[x]]$g2007))
+})
+names(err2007) <- lev
+gti_agg07 <- lapply(gti_agg, function(x) x$g2007)
+
+a <- bias_statsw_list(gti_agg07, awgts, err2007, snms, wm, "mu")
+b <- bias_statsw_list(gti_agg07, awgts, err2007, snms, wma, "mua")
+bacc2007 <- rbind(a, b)
+
+fact <- c(1, 5, 10, 25, 50, 100)
+alph <- c(225, 40)
+x <- c(0, 3, 6, 9, 12, 15)
+w <- 3 / 8
+xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
+o <- c(0, w)
+xa <- sapply(x, function(x) x + xo)
+cx <- c(1.25, 1, 1)
+g1 <- "grey90"
+reg <- c("Country", "Agricultural", "Density independent")
+lcnms <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
+mutype <- c("mua", "mu")
+# svec <- paste(gsub("f", "", bacc2011[stnm == "mu", ol]), "km")
+svec <- bacc2011[stnm == "mu", ol]
+xl <- c(-0.5, 18)
+yl <- c(-30, 50)
+shd <- c(4.5, 10.5, 16.5)
+cols <- c("red", "orange3", "green4", "blue")
+colnms <- snms
+pchs <- c("+", "o", "x")
+yax <- seq(yl[1], yl[2], 5)
+xax <- seq(1, 6, 5 / (length(yax) - 1))
+
+pdf(full_path(p_fig, "cropland_bias_mainw.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")#,
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+abline(h = yax, v = NULL, col = "grey80", lty = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
+# i <- 1; j <- 1; k <- 1
+for(i in 1:length(svec)) {
+  lv <- svec[i]
+  for(j in 1:length(snms)) {
+    # pchs1 <- pchs
+    nm <- snms[j]
+    for(k in 1:length(mutype)) {
+      #dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
+      #dat <- lcf[Metric == mutype[k] & Map == nm, lv, with = FALSE]
+      dat <- bacc2011[ol == lv & stnm == mutype[k], nm, with = FALSE][[1]]
+      pcol <- makeTransparent(cols[j], alpha = alph[k])
+      #polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
+      polyfunc2(xa[j, i] + o[k], c(0, dat), w = w, col = pcol, bcol = pcol)
+      # for(z in 1:length(reg)) {
+        # lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
+               # pch = pchs[z], type = "pt")
+      # }
+    }
+  }
+}
+axis(1, at = seq(1.5, 18.5, 3), labels = fact)
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
+legend(x = 12.4, y = -19.9, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = -19.9, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = -20.5, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
+text(x = 12.2, y = -20.5, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
+# legend(x = 12.1, y = 40, 
+#        legend = c("Country", "Agricultural", "Density independent"), pch = pchs, 
+#        pt.cex = c(1.5, 1.5), bty = "n", cex = 0.8)
+dev.off()
+
+fact <- c(1, 5, 10, 25, 50, 100)
+alph <- c(225, 40)
+x <- c(0, 3, 6, 9, 12, 15)
+w <- 3 / 8
+xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
+o <- c(0, w)
+xa <- sapply(x, function(x) x + xo)
+cx <- c(1.25, 1, 1)
+g1 <- "grey90"
+reg <- c("Country", "Agricultural", "Density independent")
+lcnms <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
+mutype <- c("MAE", "Bias")
+svec <- paste(gsub("f", "", bacc2011[stnm == "mu", ol]), "km")
+xl <- c(-0.5, 18)
+yl <- c(-40, 40)
+shd <- c(4.5, 10.5, 16.5)
+cols <- c("red", "orange3", "green4", "blue")
+colnms <- snms
+pchs <- c("+", "o", "x")
+yax <- seq(yl[1], yl[2], 5)
+xax <- seq(1, 6, 5 / (length(yax) - 1))
+
+pdf(full_path(p_fig, "cropland_bias_main.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")#,
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+abline(h = yax, v = NULL, col = "grey80", lty = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
+# i <- 1; j <- 1; k <- 1
+for(i in 1:length(svec)) {
+  lv <- colnms[i]
+  for(j in 1:length(snms)) {
+    pchs1 <- pchs
+    nm <- lcnms[j]
+    for(k in 1:length(mutype)) {
+      #dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
+      dat <- lcf[Metric == mutype[k] & Map == nm, lv, with = FALSE]
+      pcol <- makeTransparent(cols[j], alpha = alph[k])
+      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
+      for(z in 1:length(reg)) {
+        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
+               pch = pchs[z], type = "pt")
+      }
+    }
+  }
+}
+
+axis(1, at = seq(1.5, 18.5, 3), labels = fact)
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
+legend(x = 12.4, y = -19.9, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = -19.9, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = -20.5, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
+text(x = 12.2, y = -20.5, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
+legend(x = 12.1, y = 40, 
+       legend = c("Country", "Agricultural", "Density independent"), pch = pchs, 
+       pt.cex = c(1.5, 1.5), bty = "n", cex = 0.8)
+dev.off()
+```
+
+## Other calculations methods
+
+Across country, agricultural region only, or bin-wise (cropland density classes) mean.
 ```{r, eval = FALSE}

### 2015-10-17 R/bias-statsw.R

@@ -7,6 +7,7 @@
 #' @param rerror List of error rasters
 #' @param evalnm Name of value encoded in reval's cells (for column header)
 #' @param fun Function to evaluate--should be able to handle weights
+#' @param stnm Name of statistic (for output column header)
 #' @param rweight Apply reference weighting or not
 #' @param aweight Apply area weighting or not
 #' @param rnd Rounding value to apply to statistic
@@ -61,7 +62,7 @@
 #'   round(cellStats(rerror, sum), 4)
 #' @import data.table
 #' @export
-bias_statsw <- function(ref, awgts, rerror, evalnm, fun, rnd = 2, 
+bias_statsw <- function(ref, awgts, rerror, evalnm, fun, stnm = "st", rnd = 2, 
                         rweight = TRUE, aweight = TRUE, trim_wgt = TRUE) {
   if(rweight == FALSE) ref[!is.na(ref)] <- 1  # to turn of ref weights
   if(aweight == FALSE) awgts[!is.na(awgts)] <- 1  # turn off area weights
@@ -75,6 +76,7 @@ bias_statsw <- function(ref, awgts, rerror, evalnm, fun, rnd = 2,
   DT[, wgt := ref * awgts]  # calculate weights from ref and area weights
   nms <- c(evalnm, "wgt")
   odt <- round(DT[, lapply(.SD, fun, wgt), .SDcols = nms], rnd)
+  odt <- cbind(stnm, odt)
   if(trim_wgt == TRUE) odt[, wgt := NULL]
   odt
 }
@@ -85,7 +87,9 @@ bias_statsw <- function(ref, awgts, rerror, evalnm, fun, rnd = 2,
 #' @param rerrorl 2-level list of spatial error rasters 
 #' @param evalnm Name of value encoded in reval's cells (for column header)
 #' @param fun Function to evaluate--should be able to handle weights
+#' @param stnm Name of statistic (for output column header)
 #' @param rnd Rounding value to apply to statistic
+
 #' @param rweight Apply reference weighting or not
 #' @param aweight Apply area weighting or not
 #' @param rnd Rounding value to apply to statistic
@@ -93,7 +97,7 @@ bias_statsw <- function(ref, awgts, rerror, evalnm, fun, rnd = 2,
 #' @return A list of data.table output statistics
 #' @details This function runs the bias_statsw function over a list of rasters
[…]

### 2015-10-17 Rmd/carbon-bias.Rmd
17/10/2015, second update, see changelog.txt.

@@
 
-```{r, eval = TRUE}
+```{r, eval=FALSE}
 # selection variables
@@
     setkey(DT, "bin")
-    # indexing columns
-    #front1 <- data.table("bin" = rep("all", 6), "N" = rep(nrow(DT), 6))
-    #front2 <- data.table("bin" = rep("abs", 6), "N" = rep(nrow(DT), 6))
-    # stats, overall means
-    #m <- cbind(bvals, DT[mask == 1, lapply(.SD, bfn, wgt), .SDcols = nms1])
-    #ma <- cbind(bvals, DT[mask == 1, lapply(.SD, bfna, wgt), .SDcols = nms1])
-    # stats per bin
-    #b <- cbind(bvals, 
-    #           DT[mask == 1, lapply(.SD, bfn, wgt), by = bin, .SDcols = nms1])
-    #ba <- cbind(bvals, 
-    #            DT[mask == 1, lapply(.SD, bfna, wgt), by = bin, .SDcols = nms1])    
-    # output
-    #odt <- rbind(cbind(front1, mu), cbind(front2, muabs), 
-    #             binl[bin][, bin := paste0("a", bin)], 
-    #             binl[binabs][, bin := paste0("b", bin)])[, wgt := NULL]
+
+    # Potentially useful material deleted here: check repo prior to 17/10 if
+    # needed
 
@@
 
-### Extract statistics for supplemental version 
+### Extract statistics for secondary methods
+
+For supplementals
 ```{r, eval=FALSE}
@@
 
-#### Bias and accuracy weighted by agricultural area
+#### Density-weighted bias and accuracy
 
@@
 xl <- c(-0.5, 18)
-# yl <- c(-200, 200)
 shd <- c(4.5, 10.5, 16.5)
@@
 dev.off()
-
 ```
 
-#### Agricultural area bias and accuracy
-
-```{r, eval=FALSE}
-#yr <- range(all_l[, 3:8, with = FALSE], all_u[, 3:8, with = FALSE])
-# yl <- range(all_mus[, 3:8, with = FALSE])
-
-# yl <- range(mu[, 3:8, with = FALSE], mua[, 3:8, with = FALSE])
-# yl <- round(yl / 10) * 10
-yax <- seq(yl[1], yl[2], 10)
-xax <- seq(1, 6, 5 / (length(yax) - 1))
-
-mutype <- c("mua", "mu")
-
-xl <- c(-0.5, 18)
-yl <- c(-60, 70)
-shd <- c(4.5, 10.5, 16.5)
-
-pdf(full_path(p_fig, "carbon_veg_scale.pdf"), height = 7, width = 7)
-par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
-plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
-     ylab = "", xlab = "")#,
-for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
-polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
-grid(NA, 13, lwd = 1, lty = 1)
-lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey50")
-for(i in 1:length(lev_vec)) {
-  for(j in 1:length(snms)) {
-    for(k in 1:length(mutype)) {
-      dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
-      pcol <- makeTransparent(cols[j], alpha = alph[k])
-      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
-      for(z in 1:length(LC2)) {
-        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
-               pch = pchs[z], type = "pt")
-      }
-    }
-  }
-}
-axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
-axis(2, at = yax, labels = yax, las = 2)
-mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
-mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
-legend(x = 12.4, y = -40, legend = lcnms, pch = 15, col = cols, adj = 0,
-       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
-legend(x = 11.9, y = -40, legend = rep("", 4), pch = 15, adj = 0, 
-       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
-       cex = 0.8)
-text(x = 12.7, y = -42, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
-text(x = 12.2, y = -42, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
-legend(x = 12.1, y = 65, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
-       pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)
-dev.off()
-
-```
-<a href="#top">Back to top</a>
-
-#### Density independent bias and accuracy
-
-```{r, eval=FALSE}
-yl <- range(mund[, 3:8, with = FALSE], munda[, 3:8, with = FALSE])
-yl <- round(yl / 10) * 10
-yax <- seq(yl[1], yl[2], 10)
-xax <- seq(1, 6, 5 / (length(yax) - 1))
-
-mutype <- c("munda", "mund")
-
-o <- c(0, w)
-pchs <- c("*", "+", "o", "-")
-cx <- c(2, 1.25, 1, 1)
-g1 <- "grey90"
-xl <- c(-0.5, 18)
-shd <- c(4.5, 10.5, 16.5)
-
-pdf(full_path(p_fig, "carbon_veg_scale_nd.pdf"), height = 7, width = 7)
-par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
-plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
-     ylab = "", xlab = "")#,
-for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
-polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
-grid(NA, length(yax) - 1, lwd = 1, lty = 1)
-lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey50")
-for(i in 1:length(lev_vec)) {
-  for(j in 1:length(snms)) {
-    for(k in 1:length(mutype)) {
-      dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
-      pcol <- makeTransparent(cols[j], alpha = alph[k])
-      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
-      for(z in 1:length(LC2)) {
-        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
-               pch = pchs[z], type = "pt")
-      }
-    }
-  }
-}
-yll <- -120
-axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
-axis(2, at = yax, labels = yax, las = 2)
-mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
-mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
-legend(x = 12.4, y = yll, legend = lcnms, pch = 15, col = cols, adj = 0,
-       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
-legend(x = 11.9, y = yll, legend = rep("", 4), pch = 15, adj = 0, 
-       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
-       cex = 0.8)
-text(x = 12.7, y = yll - 2, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
-text(x = 12.2, y = yll - 2, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
-legend(x = 12.1, y = 165, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
-       pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)
-dev.off()
-
+```{r, echo=FALSE, eval=FALSE}
+# Agricultural area bias and accuracy, deleted here. Values in supplmentary table
+# Check repo before 17/10/15 to retrieve
 ```
@@
 
-caption <- paste("Biases and mean absolute errors for each of the cropland", 
-                 "maps across aggregation scales and each possible landcover", 
-                 "type sharing the pixel with cropland. Means were calculated",
-                 "across the entire country.")
-mu0s_xtab <- xtable(mu0s[order(Metric, Cover)], digits = 1, caption = caption)
-print(mu0s_xtab, type = "latex", file = "paper/figures/C-bias-mean-all-SA.tex", 
-      tabular.environment = "longtable", floating = FALSE, 
-      caption.placement = "top", include.rownames = FALSE)
-
-# Statistics from non-croplands areas included in means, for supplementals
-# mu[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
-# mua[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
+# Statistics from agricultural areas
 mus <- copy(rbind(cbind(Metric = "Bias", mu), 
@@
 
-caption <- paste("Biases and mean absolute errors for each of the cropland", 
-                 "maps across aggregation scales and each possible landcover", 
-                 "type sharing the pixel with cropland. Metrics were calculated",
-                 "across the union of agricultural areas", 
-                 "(cropland $>$0.05\\%) identified by the reference map and", 
-                 "the cropland map to which it was being compared")
-mus_xtab <- xtable::xtable(mus[order(Metric, Cover)], digits = 1, 
-                           caption = caption)
-print(mus_xtab, type = "latex", 
-      file = "paper/figures/C-bias-mean-cropland-SA.tex", 
-      tabular.environment = "longtable", floating = FALSE, 
-      caption.placement = "top", include.rownames = FALSE)
-
-# still to do: tables for density independent biases
-
 # Bind all three tables together
-fulls <- rbind(cbind("Region" = "Density weighted", muws), 
+fulls <- rbind(cbind("Region" = "Density", muws), 
                cbind("Region" = "Country", mu0s),
@@
       caption.placement = "top", include.rownames = FALSE) 
+#       add.to.row = list(pos = list(0), 
+#                         command = "\\hline \\endhead"))
 
@@
 ```{r, eval=FALSE, echo=FALSE}
-# Plot means against cropland density (to possibly be run - for now, 
-# too much information)
-mub <- do.call(rbind, lapply(1:20, function(x) {
-  cbind("bin" = x, extract_stat(LC, cb_stats, x, "m"))
-}))
-muba <- do.call(rbind, lapply(1:20, function(x) {
-  cbind("bin" = x, extract_stat(LC, cb_stats, x, "ma"))
-}))
-n <- do.call(rbind, lapply(1:20, function(x) {
-  cbind("bin" = x, N(LC, cb_stats, x, "m"))
-}))
-#n[, mean(f50), by = bin]
-
-pp <- list("x" = c(-2, 21), "y" = c(0, 0), 
-           "x21" = list(c(0, 9), c(9, 21)),
-           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), 
-                       seq(10, 100, 10)), 
-           "y1lwd" = list(2, 1),  
-           "y2lwd" = 0.5, 
-           "y2lc" = list(c("black", rep("grey80", 4), "black", rep("grey80", 4),
-                           rep("grey80", 4), "black", rep("grey80", 4),"black"),
-                         rep("grey80", 10)),
-           "yl" = list(c(-100, 100), c(0, 100)), 
-           "yl2" = c(-100, 100), "xl" = c(1, 20),
-           "xaxl" = unlist(lapply(seq(5, 100, 10), function(x) c(x, " "))),
-           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), 
-                            " "),
-           "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
-           "ycol" = c("grey70"))  # plot parameters
-
-yax <- seq(-100, 100, 10)
-xax <- seq(0, 22, 20 / (length(yax) - 2))
-
-pdf(full_path(p_fig, "carbon_biases_1-100km_v1.pdf"), width = 7.5, height = 7)
-par(mfrow = c(6, 4), mar = c(0.5, 0, 0.5, 0.5), mgp = c(1, 0.5, 0), tcl = -0.3,
-    oma = c(6, 4, 1, 0))
-for(f in 1:length(LC)) {
-  for(j in c(1, 4:6)) {#j <- 1; f <- 1
-      plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
-       ylab = "", xlab = "", 
-       panel.first = grid(NA, length(yax) - 1, lwd = 1, lty = 1))
-      lines(pp$x, pp$y, lwd = 1)
-    for(i in 1:4) {
-      mub[V1 == snms[i] & V2 == LC[f], 
-          lines(get(lev_vec[j]), col = cols[i], lty = 2, lwd = lw)]
-      muba[V1 == snms[i] & V2 == LC[f], 
-           lines(get(lev_vec[j]), col = cols[i], lty = 1, lwd = lw)]
-    }
-  }
-}
-dev.off()
-
-ys <- c(-100, 100)
-yax <- seq(ys[1], ys[2], 10)
-xax <- seq(0, 22, 20 / (length(yax) - 2))
-
-#yl <- list(c(-100, 100), c(0, 100))
-yl <- seq(yax[1], yax[length(yax)], 10)
-yll <- as.character(yl)
-yll[seq(2, length(yll) - 1, 2)] <- ""
-Ccols <- brewer.pal(n = 6, name = "Dark2") 
-xaxl = ifelse(seq(5, 100, 5) %in% c(5, 25, 50, 75, 100), seq(5, 100, 5), " ")
-x <- mub[V1 == snms[f] & V2 == LC[j], get(lev_vec[i])]
-y <- 100
-toohigh <- function(x, y, col = "black", yadj = 0, cex = 1.1) {
-  h <- which(abs(x) > abs(y))
-  if(length(h) > 0) {
-    xi <- h[which.max(h)]
-    text(xi, y = y + yadj, labels = round(x[xi]), cex = cex, col = col, 
-         xpd = NA)
-  }
-}
-
-pdf(full_path(p_fig, "carbon_biases_1-100km_v2.pdf"), width = 7.5, height = 7)
-par(mfrow = c(4, 4), mar = c(0.5, 0, 0.5, 0.5), mgp = c(1, 0.5, 0), tcl = -0.3,
-    oma = c(6, 4, 1, 0))
-for(f in 1:length(snms)) {
-  for(i in c(1, 4:6)) {#i <- 1; j <- 1; f <- 2
-      plot(xax, yax, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
-           ylab = "", xlab = "", 
-           panel.first = grid(NA, (length(yax) - 1) / 2, lwd = 1, lty = 1))
-      lines(pp$x, pp$y, lwd = 1)
-    for(j in 1:length(LC)) {
-      e <- parse(text = "V1 == snms[f] & V2 == LC[j]")
-      mub[eval(e), 
-          lines(get(lev_vec[i]), col = Ccols[j], lty = 2, lwd = lw)]
-
-      mub[V1 == snms[f] & V2 == LC[j], 
-          lines(get(lev_vec[i]), col = Ccols[j], lty = 2, lwd = lw)]
-      mub[V1 == snms[f] & V2 == LC[j], toohigh(get(lev_vec[i]), ys[1])]
-      muba[V1 == snms[f] & V2 == LC[j], 
-           lines(get(lev_vec[i]), col = Ccols[j], lty = 1, lwd = lw)]
-      axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
-           tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[1])
-      axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
-           tcl = -0.3, pos = pp$yl2[1])
-      if(f == length(snms)) {
-        axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
-             tcl = -0.3, pos = pp$yl2[1])
-      }
-      if(i == 1) axis(2, at = yl, labels = yll, las = 2, mgp = c(1, 0.6, 0))
-      if(i == 1 & j == 1) {
-        mtext(lcnms[f], side = 2, line = 2.8, cex = 1, col = cols[f])
-      }
-    }
-  }
-}
-dev.off()
-
-i1 <- "sa30"; i2 <- "forest"; s1 <- "f1"
-mub[V1 == i1 & V2 == i2, ]
-mub[V1 == i1 & V2 == i2, plot(bin, f1, type = "l")]
-mubin[V1 == i1 & V2 == i2, plot(bin, f1, type = "l")]
+# Plot means against cropland density, deleted because not used. Check repo
+# prior to 17/10/15 if needed. 
 ```
-
-```{r, eval=FALSE, echo=FALSE}
-# Checks on data values (embedded code)
-# Carbon values and carbon biases - random selection of cells from c_tab_diff
-# compared to recalculated values for those points
-i <- c_pct_diff
-for(chk in list(c("modmu", "f25", "forest"), c("geow", "f10", "shrub"), 
-                c("sa30", "f50", "sparse"))) {
-  x <- chk[1] #names(i[[1]])[3]
-  y <- chk[2] #names(i)[4]
-  tp <- chk[3]
-  r <- i[[y]][[x]][[tp]]
-  rid <- which(values(Which(r != 0)))
-  sid <- sample(rid, 20)
-  v1 <- r[sid]  # value to check
-  sav <- gti_agg[[y]]$gti[sid]
-  lcv <- lc_agg[[y]][[x]][sid]
-  c1 <- carbon(sav, cropC = 5, noncropC = ctabo[1, tp])
-  c2 <- carbon(lcv, cropC = 5, noncropC = ctabo[1, tp])
-  #round((sav * 5) + (ctabo[1, tp] * (1 - sav)), 2) == c1
-  #round((lcv * 5) + (ctabo[1, tp] * (1 - lcv)), 2) == c2
-  print(all(v1 == (c1 - c2) / c1 * 100))
-}  # all true
-
-# Carbon bias by bins 
-# DT-derived values
-mu0a <- do.call(rbind, lapply(1:20, function(x) {
-  cbind("bin" = x, extract_stat(LC, cb_stats, x, "ma0"))
-}))
-cb_stats$f1$sa30
-cb_stats$f10$sa30[get("bin") == 19, ]
-
-# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
-# version is finding correct results
-for(chk in list(c("modmu", "f25", "forest"), c("geow", "f10", "shrub"), 
-                c("sa30", "f50", "sparse"))) {
-  x <- chk[1] #names(i[[1]])[3]
-  y <- chk[2] #names(i)[4]
-  tp <- chk[3]
-  print(paste(".", x, "..", y, "..", tp))
-  d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(j) {
-    #print(paste("......bin", j))
-    d2 <- do.call(rbind, lapply(tp, function(z) {#:length(i[[y]][[x]]),..
-      #print(paste("........variant", z))
-      l3 <- abs(i[[y]][[x]][[z]])  # i[[y]][[x]][[z]]
-      #print(paste("........vegetation:", names(l3)))
-      d3 <- do.call(rbind, lapply(1, function(k) { #:length(l3)...
-        rs <- gti_bins[[y]][[1]] == j #[[k]] == j    
-        rs[rs == 0] <- NA
-        o <- rs * l3 #[[k]]
-        w <- area_wgts[[y]][[1]] * rs
-        oo <- cbind.data.frame("lc" = rep(paste0(x, y, z), length(w)), 
-                               getValues(o), getValues(w))
-      }))
-      d3f <- d3[!is.na(d3[, 2]), ]
-    }))
-    if(nrow(d2) > 0) { #(length(i[[x]][[y]]) * 10)) {
-      qs <- c(nrow(d2), box_stats(d2[, 2], weighted = TRUE, 
-                                  weight.opts = list("weights" = d2[, 3])))
-    } else if((nrow(d2) > 0) & (nrow(d2) < length(i[[x]][[y]]) * 20)) {
-      qs <- c(rep(NA, 6), Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], 
-                                          na.rm = FALSE))
-    } else {
-      qs <- rep(NA, 7)
-    }
-  }))
-  #print(round(d1[, "mu"], 2) == mubin[V1 == x & V2 == tp, y, with = FALSE], 
-  #          na.rm = TRUE)
-  #print(all(round(d1[, "mu"], 2) == mubin[V1 == x & V2 == tp, y, with = FALSE], 
-  #          na.rm = TRUE))
-  print(all(round(d1[, "mu"], 2) == 
-              mu0a[V1 == x & V2 == tp, y, with = FALSE], na.rm = TRUE))
-}
-# all true
-
-# another check of DT-derived stats against a raster-calculated variant
-tst <- sapply(names(c_pct_diff)[1], function(x) {
-  lapply(names(c_pct_diff[[1]]), function(y) {
-    cellStats(c_pct_diff[[x]][[y]]$forest, mean)
-  })
-})
-mean(round(unlist(tst)), 2) == mean(round(mu0[V2 == "forest", f1]), 2)
-
-gti_gt0 <- Which(gti_agg$f1$gti > 0)
-lc_gt0 <- calc(stack(lapply(lc_agg$f1, function(x) Which(x > 0))), sum)
-all_gt0 <- gti_gt0 + lc_gt0
-all_gt0 <- all_gt0 / all_gt0
-
-filts <- sapply(names(c_pct_diff), function(x) {
-  gti_gt0 <- Which(gti_agg[[x]]$gti > 0)
-  lc_gt0 <- calc(stack(lapply(lc_agg[[x]], function(j) Which(j > 0))), sum)
-  all_gt0 <- gti_gt0 + lc_gt0
-  all_gt0 <- all_gt0 / all_gt0
-})
-
-tst <- sapply(names(c_pct_diff), function(x) {
-  lapply(names(c_pct_diff[[x]]), function(y) {
-    cellStats(filts[[x]] * c_pct_diff[[x]][[y]]$forest, mean)
-  })
-})
-
-tstabs <- sapply(names(c_pct_diff), function(x) {
-  lapply(names(c_pct_diff[[x]]), function(y) {
-    cellStats(abs(filts[[x]] * c_pct_diff[[x]][[y]]$forest), mean)
-  })
-})
-
-plot(c(1, 6), range(tst, tstabs), pch = "", xlab = )
-for(i in 1:nrow(tst)) lines(1:6, tst[i, ], lty = 2)
-for(i in 1:nrow(tstabs)) lines(1:6, tstabs[i, ], col = "red")
-
-cellStats(all_gt0 * c_pct_diff$f1$sa30$sparse, mean)
-cellStats(all_gt0 * c_pct_diff$f1$sa30$forest, median)
-cellStats(c_pct_diff$f1$sa30$forest, mean)
-cellStats(all_gt0 * c_pct_diff$f1$sa30$forest, mean)
-
-```

### 2015-10-18 Rmd/cropland-bias.Rmd
18/10/2015: Updates through yield/production analysis. See changelog for specifics.

--- 2015-10-01 Rmd/landcover-bias-plots.Rmd
@@
 Plot maps of biases at different resolutions, for publication output (and display here).
-To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and then masked before plotting.
+To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and masked before plotting.
 ```{r, eval = FALSE}
@@
 # Total bias
-
 Calculate how much cropland estimates differ between datasets
@@
 
-In relation to calculation method (across country, agricultural region only, or bin-wise (cropland density classes) mean.
+## Primary method
+
+Weighted by __actual__ cropland percentage, and then, with aggregation, by number of pixels being aggregated.
+
+```{r, eval = FALSE}
+# Create weights maks
+fact <- c(5, 10, 25, 50, 100)
+namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
+# cellStats(!is.na(gti$g2011), sum)
+# cellStats(namask, sum)
+
+namask2 <- !is.na(namask)
+awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
+awgts <- lapply(awgts, function(x) x[[1]])
+
+# Subset difference grids, for 2011
+snms <- c("sa30", "globmu", "modmu", "geow")
+err2011 <- lapply(lev, function(x) {
+  sapply(dlist_act[snms], function(y) list(y[[x]]$g2011))
+})
+names(err2011) <- lev
+
+gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters
+gti_agg11 <- lapply(gti_agg, function(x) x$g2011)
+
+# calculate bias/MAE for data
+a <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wm, "mu")
+b <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wma, "mua")
+bacc2011 <- rbind(a, b)
+
+# check error stats - do they match alternate approaches?
+for(i in c("f1", "f25", "f10")) {
+  for(j in c("sa30", "modmu", "geow")) { 
+    print(paste("cross-checking calculations in", i, j))
+    a1 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
+                      "mu", aweight = FALSE)[, j, with = FALSE]
+    a2 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
+                      "mu", rweight = FALSE, aweight = FALSE)[,j, with=FALSE]
+    a3 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
+                      "mu")[, j, with = FALSE]
+    c1 <- getValues(err2011[[i]][[j]])
+    w1 <- getValues(gti_agg11[[i]])
+    w2 <- getValues(awgts[[i]])
+    print("non-area weighted mean matches?")
+    print(a1 == round(weighted.mean(c1, w1, na.rm = TRUE), 2))
+    print("totally unweighted mean matches?")
+    print(a2 == round(cellStats(err2011[[i]][[j]], mean), 2))
+    print("double weighted mean matches?")
+    print(a3 == round(weighted.mean(c1, w1 * w2, na.rm = TRUE), 2))
+  }
+}
+
+# Subset difference grids, for 2007
+snms <- c("sa30", "globmu", "modmu", "geow")
+err2007 <- lapply(lev, function(x) {
+  sapply(dlist_act[snms], function(y) list(y[[x]]$g2007))
+})
+names(err2007) <- lev
+gti_agg07 <- lapply(gti_agg, function(x) x$g2007)
+
+a <- bias_statsw_list(gti_agg07, awgts, err2007, snms, wm, "mu")
+b <- bias_statsw_list(gti_agg07, awgts, err2007, snms, wma, "mua")
+bacc2007 <- rbind(a, b)
+
+fact <- c(1, 5, 10, 25, 50, 100)
+alph <- c(225, 40)
+x <- c(0, 3, 6, 9, 12, 15)
+w <- 3 / 8
+xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
+o <- c(0, w)
+xa <- sapply(x, function(x) x + xo)
+cx <- c(1.25, 1, 1)
+g1 <- "grey90"
+reg <- c("Country", "Agricultural", "Density independent")
+lcnms <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
+mutype <- c("mua", "mu")
+# svec <- paste(gsub("f", "", bacc2011[stnm == "mu", ol]), "km")
+svec <- bacc2011[stnm == "mu", ol]
+xl <- c(-0.5, 18)
+yl <- c(-30, 50)
+shd <- c(4.5, 10.5, 16.5)
+cols <- c("red", "orange3", "green4", "blue")
+colnms <- snms
+pchs <- c("+", "o", "x")
+yax <- seq(yl[1], yl[2], 5)
+xax <- seq(1, 6, 5 / (length(yax) - 1))
+
+pdf(full_path(p_fig, "cropland_bias_main.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")#,
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
+abline(h = yax, v = NULL, col = "grey80", lty = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
+# i <- 1; j <- 1; k <- 1
+for(i in 1:length(svec)) {
+  lv <- svec[i]
+  for(j in 1:length(snms)) {
+    # pchs1 <- pchs
+    nm <- snms[j]
+    for(k in 1:length(mutype)) {
+      dat <- bacc2011[ol == lv & stnm == mutype[k], nm, with = FALSE][[1]]
+      pcol <- makeTransparent(cols[j], alpha = alph[k])
+      polyfunc2(xa[j, i] + o[k], c(0, dat), w = w, col = pcol, bcol = pcol)
+    }
+  }
+}
+axis(1, at = seq(1.5, 18.5, 3), labels = fact)
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
+legend(x = 12.4, y = -19.9, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = -19.9, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = -20.5, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
+text(x = 12.2, y = -20.5, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
+dev.off()
+
+```
+
+## Other calculations methods
+
+Across country, agricultural region only, or bin-wise (cropland density classes) mean.
 ```{r, eval = FALSE}
@@
 
-pdf(full_path(p_fig, "cropland_bias_scale.pdf"), height = 7, width = 7)
+pdf(full_path(p_fig, "cropland_bias_region.pdf"), height = 7, width = 7)
 par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
@@
 
+# check bin frequencies starting a bit higher: 2.5% cover
+binv <- seq(2.5, 97.5, 5)
+gti_bins <- lapply(gti_agg, function(x) {
+  lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE))
+})
+bins_25 <- lapply(gti_bins, function(x) freq(x$g2011))
+barplot(bins_25$f1[1:19, 2])
 ```

### 2015-10-18 Rmd/yield-bias.Rmd
18/10/2015: Updates through yield/production analysis. See changelog for specifics.

@@
 ---
-title: "Yield bias"
+title: "Yield and production bias/accuracy"
 author: "Lyndon Estes"
@@
 1.
-How does these processes create bias if they are not aggregated by the administrative units to which they are scaled?
+How do these processes create bias if they are not aggregated by the administrative units to which they are scaled?
 i.e.
@@
 
+# calculate bias/MAE for data
 # calculate means
@@
 wma <- function(x, w) stats::weighted.mean(abs(x), w)
-a <- bias_stats_list(bins, awgts, lcu, cf_pct_diff2, wm, "mu", "bias", TRUE)
-b <- bias_stats_list(bins, awgts, lcu, cf_pct_diff2, wma, "mua", "bias", TRUE)
-# b <- bias_stats_list(bins, awgts, lcnu, cf_pct_diff2, wma, "mua", "bias", TRUE)
-# b[ol == "f1" & il == "modmu" & bin == "all"]
+
+# reshape error raster for density-weighted bias/accuracy calculations
+cf_pct_resh <- lapply(lev, function(x) {
+  sapply(cf_pct_diff2, function(y) list(y[[x]]))
+})
+names(cf_pct_resh) <- lev
+
+# bias/accuracy
+a <- bias_statsw_list(gti_agg, awgts, cf_pct_resh, snms, wm, "mu")
+b <- bias_statsw_list(gti_agg, awgts, cf_pct_resh, snms, wma, "mua")
+lcf_err <- rbind(a, b)
 
@@
 # version is finding correct results
-bv <- "mua"
-for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
-                c("globmu", "f1"))) {
-  x <- chk[1] 
-  y <- chk[2] 
-  print(paste(".", x, "..", y))
-  rs <- lcu[[y]][[x]] 
-  rs[rs == 0] <- NA
-  l3 <- abs(cf_pct_diff2[[x]][[y]])
-  o <- rs * l3
-  w <- awgts[[y]][[1]] * rs
-  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
-  print(b[ol == y & il == x & bvals == bv & bin == "all", bias] ==
-          round(wmu, 2))
-}  # check
-
-lcfmua <- extract_stat(lev, snms, "all", "mua", "bias", b)
-lcfmua_nm <- extract_stat(lev, snms, "all", "mua.nm", "bias", b) 
-# b[ol == "f10" & il == "globmu" & bin == "all"]
-lcfmua_nd <- extract_stat(lev, snms, 1:20, "mua.nm", "bias", b, 
-                          type = "nodensity")
-b[ol == "f1" & il == "sa30" & bin != "all" & bvals == "mua.nm", mean(bias)]
-
-svec <- c("mua.nm", "mua")
-aa <- c("Country", "Agricultural")
+# check error stats - do they match alternate approaches?
+for(i in c("f1", "f25", "f10")) {
+  for(j in c("sa30", "modmu", "geow")) { 
+    print(paste("cross-checking calculations in", i, j))
+    a1 <- bias_statsw(gti_agg[[i]]$gti, awgts[[i]], cf_pct_resh[[i]], snms, wm, 
+                      "mu", aweight = FALSE)[, j, with = FALSE]
+    a2 <- bias_statsw(gti_agg[[i]]$gti, awgts[[i]], cf_pct_resh[[i]], snms, wm, 
+                      "mu", rweight = FALSE, aweight = FALSE)[,j, with=FALSE]
+    a3 <- bias_statsw(gti_agg[[i]]$gti, awgts[[i]], cf_pct_resh[[i]], snms, wm, 
+                      "mu")[, j, with = FALSE]
+    c1 <- getValues(cf_pct_resh[[i]][[j]])
+    w1 <- getValues(gti_agg[[i]]$gti)
+    w2 <- getValues(awgts[[i]])
+    print("non-area weighted mean matches?")
+    print(a1 == round(weighted.mean(c1, w1, na.rm = TRUE), 2))
+    print("totally unweighted mean matches?")
+    print(a2 == round(cellStats(cf_pct_resh[[i]][[j]], mean), 2))
+    print("double weighted mean matches?")
+    print(a3 == round(weighted.mean(c1, w1 * w2, na.rm = TRUE), 2))
+  }
+}
+
+# Removed whole-country and agricultural area accuracy measures. See repo prior
+# to 18/10 if needed. 
+svec <- c("mu", "mua")
+aa <- c("Bias", "Accuracy")
 mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
 
+# Reshape for output table
 lcf_out <- do.call(rbind, lapply(1:length(svec), function(i) {
-  a <- do.call(rbind, lapply(snms, function(x) {
+  a <- do.call(rbind.data.frame, lapply(snms, function(x) {
     aa <- sapply(lev, function(y) {
-      b[ol == y & il == x & bin == "all" & bvals == svec[i], bias]
-    })
-    named_out(aa, paste(c(1, fact), "km"))
+      lcf_err[ol == y & stnm == svec[i], x, with = FALSE]
+    })
+    named_out(c(x, aa), c("Map", paste(c(1, fact), "km")))
   }))
-  cbind.data.frame("Region" = aa[i], "Map" = mcap, a)
+  named_out(cbind.data.frame(aa[i], a), 
+            c("Metric", "Map", paste(c(1, fact), "km")))
 }))
 
-caption <- paste("Mean absolute errors (MAE) in statistically constrained", 
-                 "cropland maps, calculated over the entire country, and for", 
-                 "the union of agricultural regions (cropland $>$ 0) defined ", 
-                 "by the reference and corresponding adjusted cropland",
-                 "maps")    
-lcfout_xtab <- xtable(lcf_out, digits = 1, caption = caption)
-print(lcfout_xtab, type = "latex", file = "paper/figures/cropadj-mae.tex", 
+# rename landcover sets for output
+lcf_out <- as.data.table(lcf_out)
+setkey(lcf_out, "Map")
+for(i in 1:length(snms)) lcf_out[Map == snms[i], Map := mcap[i]]
+
+# Output table
+caption <- paste("Bias and mean absolute errors (MAE) in statistically", 
+                 "constrained cropland maps across aggregation scales,",
+                 "weighted by density of cropland", 
+                 "cover in the reference map. ")    
+lcfout_xtab <- xtable(lcf_out[order(Metric)], digits = 1, caption = caption)
+print(lcfout_xtab, type = "latex", 
+      file = "paper/figures/cropadj-bias-accuracy.tex", 
       tabular.environment = "longtable", floating = FALSE, 
@@
 
-# not implausible that someone would do a direct aggregation of these yield 
-# values, without weighting by crop distribution, so this incorrect variant 
-# would be: 
-# cyld_agg2 <- aggregate_rast_list(fact, cyld_agg$f1, mean) # incorrect yld agg
-
-# third variant - aggregate and remove zeros
-# first define a function that removes zero yield areas when calculating mean 
-# yield--this will allow for aggregation of yields only in those areas having 
-# maize, according to Monfreda et al methods
-# nzmean <- function(x, na.rm) {
-#   x <- x[(x > 0) & !is.na(x)]
-#   if(length(x) == 0) {
-#     m <- 0
-#   } else if(length(x) > 0) {
-#     m <- mean(x, na.rm = na.rm)
-#   }
-#   return(m)
-# }
-# cyld_agg3 <- aggregate_rast_list(fact, cyld_agg$f1, nzmean) # no zero agg
-
-# plot(cyld_agg$f25$modmu - cyld_agg2$f25$modmu)
-# plot(cyld_agg$f25$modmu - cyld_agg3$f25$modmu)
-wm <- function(x, w) stats::weighted.mean(x, w)
+# Removed other versions of yield aggregation. Refer to repo prior to 18/10 to 
+# to recover
 
@@
 
-# yat <- list(cyld_agg, cyld_agg2, cyld_agg3)
-# ## check values
-# # check yields
-# for(i in lev[-1]) {
-#   print(paste("factor", i))
-#   for(j in asnms) {
-#     print(paste("...", j))
-#     for(k in 1:3) {
-#       #ck <- cellStats(cyld_agg$f1$gti, nzmean)
-#       print(paste("...... type", k, "agg"))
-#       ck <- crop_ay[[j]][, wm(mzyld, fa)]  # weighted by fraction
-#       ck2 <- crop_ay$gti[, wm(mzyld, fa)]  # weighted by fraction
-#       w <- values(carea_agg[[i]][[j]])  # aggregated crop fraction
-#       x <- values(yat[[k]][[i]][[j]])  # yield values
-#       ind <- which(x > 0)  # which are non-zero
-#       x <- x[ind]  # select non zeros
-#       w <- w[ind]  # select non-zeros
-#       #w <- w / max(w)
-#       v <- weighted.mean(x, w)
-#       d <- round((ck - v) / ck, 2)
-#       d2 <- round((ck2 - v) / ck2, 2)
-#       print(c(d, d2))
-#     }
-#   }  # type 1 and type 3 aggregations consistent between datasets, across scales
-# }
-# rm(yat)
+# Removed yat check. Refer to repo prior to 18/10 to recover
 
@@
 
-# with 0-removed yield aggregation (type 3 yield aggregation)
-# cpagg3 <- lapply(1:length(cyld_agg3), function(x) {
-#   p <- lapply(1:length(cyld_agg3[[x]]), function(y) {
-#     (tareas[[x]] * carea_agg[[x]][[y]]) * cyld_agg3[[x]][[y]]
-#   })
-#   named_out(p, asnms)
-# })
-# names(cpagg3) <- names(agg_res)
+# Removed production aggregation with 0-removed yieldsother. Refer to repo prior
+# to 18/10 to recover
 
@@
 
-# Mean production in producing cells at different levels of aggregation
+DT <- na.omit(as.data.table.raster(s, xy = TRUE))
+DT[, wgt := ref * awgts]  # calculate weights from ref and area weights
+
+# Mean production in producing cells at different levels of aggregation. 
+# No need to weight by cropland density, because production value already factors
+# in area - just mask out zero production areas
 mup <- sapply(lev, function(i) {
   msk <- cpagg1[[i]]$gti > 0
-  #msk <- cpaggII[[i]]$gti > 0
   msk[msk == 0] <- NA
@@
 })
-# sapply(lev, function(i) {
-#   msk <- cpagg1[[i]]$gti > 0
-#   msk[msk == 0] <- NA
-#   cpmsk <- msk * cpagg1[[i]]$gti
-#   cellStats(cpmsk, mean)
-# })
-
-# mean yield in producing cells at different levels of aggregation
+
+# mean yield in producing cells at different levels of aggregation, use density
+# weighting here
 muy <- sapply(lev, function(i) {
-  w <- values(carea_agg[[i]]$gti)  # aggregated crop fraction
-  x <- values(cyld_agg[[i]]$gti)  # yield values
-  ind <- which(x > 0)  # which are non-zero
-  x <- x[ind]  # select non zeros
-  w <- w[ind]  # select non-zeros
-  v <- weighted.mean(x, w, na.rm = TRUE, )
-  v
-})  # yields across scales and datasets all 3.4-3.5 kg/ha
+  # i <- lev[2]; print(i)
+  s <- stack(cyld_agg[[i]]$gti, carea_agg[[i]]$gti, awgts[[i]])
+  names(s) <- c("yld", "ref", "awgts")
+  DT <- na.omit(as.data.table.raster(s, xy = TRUE))
+  DT[, wgt := ref * awgts]  # calculate weights from ref and area weights
+  DT[, weighted.mean(yld, wgt)]
+})  # yield is 3.361 kg/ha at all scales
+
+# Removed original mup and muy. Refer to repo prior to 18/10 to recover
 
@@
 # Production differences
+# first standardize
+# std_raster <- function(x) {
+#   (x - cellStats(x, min)) / diff(cellStats(x, range))
+# }
+# cpagg_std <- lapply(cpagg1, function(x) lapply(x, function(y) std_raster(y)))
 ilist <- list(names(cpagg1[[1]])[-1], names(cpagg1), "gti")
+# list1 <- lapply(cpagg1, function(x) x[1])
+# list2 <- lapply(cpagg1, function(x) x[-1])
 list1 <- lapply(cpagg1, function(x) x[1])
@@
 p_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1 agg
-# list2 <- lapply(cpagg3, function(x) x[-1])
-# p_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 3 agg
-# list2 <- lapply(cpaggII, function(x) x[-1])
-# p_diff3 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type II agg
-# pmu <- as.data.table.raster(cpagg3$f1$globmu)[layer > 0, mean(layer)]
+
+# Removed original p_diff2/3. Refer to repo prior to 18/10 to recover
 
@@
 
-# type 2 (non-zero)
-# list2 <- lapply(cyld_agg3, function(x) x[-1]) # cyld_agg for a, b/c correct
-# y_diff2 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 2
+# Removed original y_diff2. Refer to repo prior to 18/10 to recover
 
@@
 # here only for map plotting.
+
 pdiffs <- lapply(list(p_diff1), function(x) {
@@
 })
-# names(pdiffs) <- paste0("pd", 1:3)
 names(pdiffs) <- "pd1"
 
-# ydiffs <- lapply(list(y_diff1, y_diff2), function(x) {
 ydiffs <- lapply(list(y_diff1), function(x) {
@@
 })
-# names(ydiffs) <- paste0("yd", 1:2)
 names(ydiffs) <- "yd1"
@@
 p_diff1d <- disaggregate_rast_list(snms, lev, pdiffs$pd1, namask)
-# p_diff2d <- disaggregate_rast_list(snms, lev, pdiffs$pd2, namask)
-# p_diff3d <- disaggregate_rast_list(snms, lev, pdiffs$pd3, namask)
 y_diff1d <- disaggregate_rast_list(snms, lev, ydiffs$yd1, namask)
-# y_diff2d <- disaggregate_rast_list(snms, lev, ydiffs$yd2, namask)
 
@@
 stats_pd1 <- dstats(p_diff1d)
-# stats_pd2 <- dstats(p_diff2d)
-# stats_pd3 <- dstats(p_diff3d)
 stats_yd1 <- dstats(y_diff1d)
-# stats_yd2 <- dstats(y_diff2d)
-# cbind(stats_yd1$geow[, 4], stats_yd2$geow[, 4])
 ```
@@
 
-# fnms <- c(paste0("prod_bias_map_", 1:3, ".pdf"), 
-#           paste0("yld_bias_map_", 1:2, ".pdf"))  # output names
-fnms <- c("prod_bias_map_1.pdf", "yld_bias_map_1.pdf")  # output names
-# legtext <- c(rep("tons", 3), rep("tons/ha", 2))
+fnms <- c("prod_bias_map.pdf", "yld_bias_map.pdf")  # output names
 legtext <- rep("% Difference", 5)
@@
 brksl <- lapply(rng2, function(x) {
-  c(floor(x[1:2]), c(-50, -25, -10, -1, 1, 10, 25, 50), ceiling(x[3:4]))
-})
+  c(floor(x[1:2]), c(-50, -25, -10, -5, -1, 1, 5, 10, 25, 50), ceiling(x[3:4]))
+})
+# brksl <- list(c(-1, -0.75, -0.5, -0.25, -0.1, -0.05, -0.01, 0.01, 0.05, 0.1, 
+#                 0.25, 0.5, 0.75, 1), 
+#               c(-10.7, -7, -5, -3, -2, -1, -0.1, 0.1, 1, 2, 3, 5, 7, 10.7))
 
 for(i in 1:length(statsl)) {
-  # i <- 1
-  #lim <- rng[[i]]
-  # lim <- rng2[[i]]
-  # brks <- div_breaks(olim = lim[c(1, 4)], ilim = lim[2:3], 
-  #                   ibrk = 0.01 * min(abs(lim[2:3])))
-  # if(i < 4) brks <- round(brks)
-  # if(i >= 4) brks <- round(brks, 1)
+  # i <- 2
   print(paste("figure", i, "of", length(statsl)))
@@
   brklen <- length(brks) - 1
-  #cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)
-  #cols[c(1, length(cols))] <- c("darkred", "purple3")
   pdf(full_path(p_fig, fnms[i]), height = 6, width = 7)
@@
 
-From the above it is clear that the nature in which aggregation of yields are done has some impact on both yield and production estimates.
+Several versions of yield aggregation were tested in this analysis, each of which has a different impact on yield and production impacts.
 There are three types of yield aggregation that can be done (most commented out now after initial checking):
@@
 
-      + ___Note: investigate this by look at correlation between rainfall and yield bias patterns.___
-
-Only Types 1 and A are included in further investigation. The other variants are left as accuracy checks.
+Type 2 and 3 yield aggregation are incorrect methods, and are not retained here (but see earlier commits of code, specifically prior to 18/10/2015.
+The other variants were used for comparison and for initial methods checks only.
 
@@
 
-Following methods developed in [carbon-bias](carbon-bias.html)
+### Primary method
+
+Density weighted, following methods developed in [cropland-bias](croland-bias.html), but here metrics are based on residuals rather than percent errors, because of infinite values (harvested area/yield disaggregation methods mean that there are gridded yields/production estimates in some areas where no reference value exists, leading to infinite values if converting to percent)
 ```{r, eval=FALSE}
-wm <- function(x, w) stats::weighted.mean(x, w)
-wma <- function(x, w) stats::weighted.mean(abs(x), w)
-
+
+# yield bias/accuracy
+# reshape
+yerr <- lapply(lev, function(x) {
+  sapply(y_diff1, function(y) list(y[[x]]$gti))
+})
+names(yerr) <- lev
+
+# carea_gti <- lapply(carea_agg, function(x) x$gti)
+a <- bias_statsw_list(gti_agg, awgts, yerr, snms, wm, "mu")
+b <- bias_statsw_list(gti_agg, awgts, yerr, snms, wma, "mua")
+yld_bacc <- rbind(a, b)
+
+# production bias/accuracy
+# reshape
+perr <- lapply(lev, function(x) {
+  sapply(p_diff1, function(y) list(y[[x]]$gti))
+})
+names(perr) <- lev
+
+# carea_gti <- lapply(carea_agg, function(x) x$gti)
+a <- bias_statsw_list(gti_agg, awgts, perr, snms, wm, "mu")
+b <- bias_statsw_list(gti_agg, awgts, perr, snms, wma, "mua")
+prod_bacc <- rbind(a, b)
+
+# check error stats - do they match alternate approach?
+ref <- lapply(gti_agg, function(x) x$gti)
+test_err <- "perr" #"yerr"
+evst <- "wma" # "wm"
+for(i in c("f1", "f25", "f10")) {
+  for(j in c("sa30", "modmu", "geow")) { 
+    print(paste("cross-checking calculations in", i, j))
+    a1 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms,  
+                      get(evst), aweight = FALSE)[, j, with = FALSE]
+    a2 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms, 
+                      get(evst), rweight = FALSE, 
+                      aweight = FALSE)[, j, with = FALSE]
+    a3 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms, 
+                      get(evst))[, j, with = FALSE]
+    s <- stack(get(test_err)[[i]][[j]], ref[[i]], awgts[[i]])
+    sv <- getValues(s)
+    sv <- sv[which(!is.na(rowSums(sv))), ]  # remove NAs
+    if(evst == "wm") {
+      print(a1 == round(weighted.mean(sv[, 1], sv[, 2], na.rm = TRUE), 2))
+      print(a2 == round(mean(sv[, 1]), 2))
+      print(a3 == round(weighted.mean(sv[, 1], sv[, 2] * sv[, 3]), 2))
+    } else if(evst == "wma"){
+      print(a1 == round(weighted.mean(abs(sv[, 1]), sv[, 2], na.rm = TRUE), 2))
+      print(a2 == round(mean(abs(sv[, 1])), 2))
+      print(a3 == round(weighted.mean(abs(sv[, 1]), sv[, 2] * sv[, 3]), 2))
+    }
+  }
+}
+
+```
+
+### Secondary method
+
+Bias and accuracy within agricultural areas
+```{r, eval=FALSE}
 # create mask for non-cropland areas, unioning GTI and LC areas with maize
@@
 names(lcu) <- lev
-# plot(lcu$f50$modmu)
-# plot(gti_agg$f50$gti)
 
@@
 })
-# save(bins, awgts, lcu, p_diff1, wm, wma, p_diff1, cyld_agg, cyld_agg3, 
-#      carea_agg, cpagg1, cpagg3, cpaggII, p_diff2, p_diff3, y_diff1, 
-#      y_diff2, snms, lev, file = "external/ext_data/yield-bias.rda")
-# load("external/ext_data/yield-bias.rda")
-# reval <- p_diff1[[2]][[1]][[1]]  # hack in case there is a 3rd level
-# revall <- p_diff1
-# rmask <- lcu[[2]][[1]]
-# rmaskl <- lcu
-# rwgts <- awgts[[2]]
-# rwgtsl <- awgts
-# rbins <- bins[[2]]
-# rbinsl <- bins
-# fun <- sum
-# stnm <- "sum"
-# evalnm <- "bias"
-# rnd = 2; trim_wgt = TRUE; silent = FALSE; weighted = TRUE
-# wm <- function(x, w) stats::weighted.mean(x, w) 
-# tst <- data.table("a" = runif(20, 0, 1), 
-#                   "w" = sample(20:100, 20, replace = TRUE))
-# tst$w <- 0
-# tst[, lapply(.SD, sum, w), .SDcols = 1:2]
-# weighted.mean(tst$a, tst$w)
-# mean(tst$a)
-# sum(tst$a)
+
+# Many commented out checks removed here. Look at pre 18/10 commits to restore
 
@@
 # percentages
+# Many stats for alternate yield/production metrics removed here. 
+# Look at pre 18/10 commits to restore
 a <- bias_stats_list(bins, awgts, lcu, p_diff1, wm, "mu", "bias", TRUE)
@@
 pd1_st <- rbind(a, b, d)
-# a <- bias_stats_list(bins, awgts, lcu, p_diff2, wm, "mu", "bias", TRUE)
-# b <- bias_stats_list(bins, awgts, lcu, p_diff2, wma, "mua", "bias", TRUE)
-# d <- bias_stats_list(bins, awgts, lcu, p_diff2, sum, "sum", "bias", FALSE)
-# pd2_st <- rbind(a, b, d)
-# a <- bias_stats_list(bins, awgts, lcu, p_diff3, wm, "mu", "bias", TRUE)
-# b <- bias_stats_list(bins, awgts, lcu, p_diff3, wma, "mua", "bias", TRUE)
-# d <- bias_stats_list(bins, awgts, lcu, p_diff3, sum, "sum", "bias", FALSE)
-# pd3_st <- rbind(a, b, d)
+
 a <- bias_stats_list(bins, awgts, lcu, y_diff1, wm, "mu", "bias", TRUE)
@@
 yd1_st <- rbind(a, b)
-# a <- bias_stats_list(bins, awgts, lcu, y_diff2, wm, "mu", "bias", TRUE)
-# b <- bias_stats_list(bins, awgts, lcu, y_diff2, wma, "mua", "bias", TRUE)    
-# yd2_st <- rbind(a, b)
 
@@
 p1ma <- extract_stat(lev, snms, "all", "mua", "bias", pd1_st)  # ident to p3
-# p1sm <- extract_stat(lev, snms, "all", "sum", "bias", pd1_st)  # ident to p3
-# p2mu <- extract_stat(lev, snms, "all", "mu", "bias", pd2_st)
-# p2ma <- extract_stat(lev, snms, "all", "mua", "bias", pd2_st)
-# p2sm <- extract_stat(lev, snms, "all", "sum", "bias", pd2_st)
-# p3mu <- extract_stat(lev, snms, "all", "mu", "bias", pd3_st)
-# p3ma <- extract_stat(lev, snms, "all", "mua", "bias", pd3_st)
-# p3sm <- extract_stat(lev, snms, "all", "sum", "bias", pd3_st)
 y1mu <- extract_stat(lev, snms, "all", "mu", "bias", yd1_st)
 y1ma <- extract_stat(lev, snms, "all", "mua", "bias", yd1_st)
-# y2mu <- extract_stat(lev, snms, "all", "mu", "bias", yd2_st)
-# y2ma <- extract_stat(lev, snms, "all", "mua", "bias", yd2_st)
 
@@
 
-### Bias plots
-```{r, eval=FALSE}
-alph <- c(225, 40)
+### Bias/accuracy plots
+
+#### Density-weighted bias and accuracy 
+```{r, eval = FALSE}
+alph <- c(225, 60)
 x <- c(0, 3, 6, 9, 12, 15)
@@
 
-# mul <- c("p1mu", "p2mu", "p3mu", "y1mu", "y2mu")
-# mul <- c("p2mu", "p3mu", "y1mu", "y2mu")
-mul <- c("p1mu", "y1mu")
-#mal <- c("p1ma", "p2ma", "p3ma", "y1ma", "y2ma")
-# mal <- c("p2ma", "p3ma", "y1ma", "y2ma")
-mal <- c("p1ma", "y1ma")
-malmul <- list(mal, mul)
-#pyl <- c("mup", "mup", "mup", "muy", "muy")
-# pyl <- c("mup", "mup", "muy", "muy")
+# get stats data into shape
+mulw <- list(prod_bacc[stnm == "mu"], yld_bacc[stnm == "mu"])
+malw <- list(prod_bacc[stnm == "mua"], yld_bacc[stnm == "mua"])
+malmulw <- list(malw, mulw)  # density-weighted stats, reordered
 pyl <- c("mup", "muy")
@@
 xl <- c(-0.5, 18)
-yl <- c(-70, 80)
+yl <- c(-30, 110)  # try it out, play with it 
 shd <- c(4.5, 10.5, 16.5)
@@
 lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
-# pchs <- c("*", "+", "o", "x", "-")
 pchs <- c("+", "o")
 
-#yl <- range(mu[, 3:8, with = FALSE], mua[, 3:8, with = FALSE])
-#yl <- round(yl / 10) * 10
 yax <- seq(yl[1], yl[2], 10)
@@
 
-pdf(full_path(p_fig, "yield_bias_scale.pdf"), height = 7, width = 7)
+pdf(full_path(p_fig, "yield_prod_bias.pdf"), height = 7, width = 7)
 par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
@@
 for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
-# grid(NA, NULL, lwd = 1, lty = 1)
+abline(h = yax, v = NULL, col = "grey80", lty = 1)
+polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
+lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
+for(ii in 1:length(lev)) { # ii <- 1; i <- 1; k <- 1
+  lv <- lev[ii]
+  for(i in 1:length(snms)) {
+    pchs1 <- pchs
+    nm <- snms[i]
+    for(k in 1:length(malmulw)) {
+      mm <- malmulw[[k]]
+      # fetch bias stats, convert to percentage relative to mean
+      # e.g. GTI mean yield
+      v <- sapply(1:length(mm), function(j) {  # j <- 1
+        round(mm[[j]][ol == lv, get(nm)] / get(pyl[j])[lv] * 100, 1)
+      })
+      pcol <- makeTransparent(cols[i], alpha = alph[k])
+      polyfunc2(xa[i, ii] + o[k], range(v), w = w, col = pcol, bcol = pcol)
+      xx <- rep(x[ii] + xo[i] + o[k], length(v))
+      points(xx, v, pch = pchs, cex = cx)
+    }
+  }
+}
+axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
+axis(2, at = yax, labels = yax, las = 2)
+mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
+mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
+legend(x = 12.4, y = -10, legend = lcnms, pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
+legend(x = 11.9, y = -10, legend = rep("", 4), pch = 15, adj = 0, 
+       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
+       cex = 0.8)
+text(x = 12.7, y = -12, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
+text(x = 12.2, y = -12, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
+legend(x = 12.1, y = 100, legend = c("Production", "Yield"), pch = pchs, 
+       pt.cex = c(1.5, 1.5), bty = "n", cex = 0.8)
+dev.off()
+```
+
+#### Agriultural area bias and accuracy
+
+Supplemental
+```{r, eval=FALSE}
+alph <- c(225, 60)
+
+mula <- c("p1mu", "y1mu")
+mala <- c("p1ma", "y1ma")
+malmula <- list(mala, mula)
+pyl <- c("mup", "muy")
+
+yl <- c(-70, 80)
+yax <- seq(yl[1], yl[2], 10)
+
+pdf(full_path(p_fig, "yield_prod_bias_agric.pdf"), height = 7, width = 7)
+par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
+plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
+     ylab = "", xlab = "")
+for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
 abline(h = yax, v = NULL, col = "grey80", lty = 1)
@@
     nm <- snms[i]
-    for(k in 1:length(malmul)) {
-      mm <- malmul[[k]]
+    for(k in 1:length(malmula)) {
+      mm <- malmula[[k]]
       # fetch bias stats, convert to percentage relative to mean
@@
     nm <- snms[i]
-    lapply(1:length(malmul), function(k) {
-      mm <- malmul[[k]]
+    lapply(1:length(malmula), function(k) {
+      mm <- malmula[[k]]
       v <- sapply(1:length(mm), function(j) {
@@
 }))
-
+```
+
+#### Supplemental tables
+```{r, eval = FALSE}
 vnms <- c(rep("Yield", length(snms)), rep("Production", length(snms)))
 aa <- c("Bias", "MAE")
-out_tab <- do.call(rbind, lapply(1:2, function(x) {
+
+# Density-weighted
+out_tabw <- do.call(rbind, lapply(1:2, function(x) {  # x <- 1 
+  out <- do.call(cbind, lapply(1:length(lev), function(ii) {  # ii <- 1
+    lv <- lev[ii]  
+    mm <- malmulw[2:1][[x]]  
+    v <- do.call(rbind, lapply(length(mm):1, function(j) {  # j <- 1
+      v2 <- do.call(rbind, lapply(1:length(snms), function(i) { # i <- 1
+        nm <- snms[i]
+        round(mm[[j]][ol == lv, get(nm)] / get(pyl[j])[lv] * 100, 1)
+      }))
+      rownames(v2) <- snms
+      v2
+    }))
+  }))
+  out <- cbind.data.frame(aa[x], mcap, "Variable" = vnms, unname(out))
+  colnames(out) <- c("Metric", "Map", "Variable", paste(c(1, fact), "km"))
+  out
+}))
+
+# Agricultural area
+out_taba <- do.call(rbind, lapply(1:2, function(x) {
   out <- do.call(cbind, lapply(1:length(lev), function(ii) {
     lv <- lev[ii]  
-    mm <- malmul[2:1][[x]] 
+    mm <- malmula[2:1][[x]] 
     v <- do.call(rbind, lapply(length(mm):1, function(j) {
@@
 
+# Join them
+out_tab <- rbind(cbind("Region" = "Density", out_tabw), 
+                 cbind("Region" = "Agricultural", out_taba))
+
 caption <- paste("Biases and mean absolute errors (MAE) in disaggregated maize",
@@
                  "and harvested area estimates)",
-                 "maps. Bias and MAE were normalized to their", 
+                 "maps. Values for both density-weighted and agricultural", 
+                 "areas bias and accuracy are presennted.",
+                 "Bias and MAE were normalized to their", 
                  "respective mean values calculated from reference maps.")    
@@
 Using data from the Princeton Global Forcing dataset, downloaded from the Africa Flood and Drought Monitor.
+Note: currently this is only explored here.
 ```{r, eval=FALSE}

### 2015-10-18 paper/app_s1.tex
18/10/2015: Updates through yield/production analysis. See changelog for specifics.

@@
 \FloatBarrier
-\input{figures/C-bias-mean-cropland-SA.tex}
-
-\input{figures/C-bias-mean-all-SA.tex}
+\input{figures/C-bias-accuracy.tex}
 
@@
 \FloatBarrier
-\input{figures/cropadj-mae.tex}
+\input{figures/cropadj-bias-accuracy.tex}
 
@@
   \centering
-     \includegraphics[width = 12cm]{figures/yld_bias_map_1.pdf} 
+     \includegraphics[width = 12cm]{figures/yld_bias_map.pdf} 
       \caption{Errors (normalized to the reference-derived country mean) in disaggregated maize yield estimates.}
@@
   \centering
-     \includegraphics[width = 12cm]{figures/prod_bias_map_1.pdf} 
+     \includegraphics[width = 12cm]{figures/prod_bias_map.pdf} 
       \caption{Errors (normalized to reference-derived country mean) production estimates calculated from disaggregated maize yield and harvested area estimates.}

### 2015-10-18 paper/figures/cropadj-bias-accuracy.tex
18/10/2015: Updates through yield/production analysis. See changelog for specifics.

% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Sat Oct 17 22:47:55 2015
\begin{longtable}{llrrrrrr}
\caption{Bias and mean absolute errors (MAE) in statistically constrained cropland maps across aggregation scales, weighted by density of cropland cover in the reference map. } \\ 
  \hline
Metric & Map & 1 km & 5 km & 10 km & 25 km & 50 km & 100 km \\ 
  \hline
Bias & GeoWiki & 9.7 & 1.1 & 0.6 & 0.4 & 0.5 & 0.1 \\ 
  Bias & GlobCover & 34.5 & 18.3 & 14.5 & 10.6 & 7.6 & 4.6 \\ 
  Bias & MODIS & 17.8 & 5.5 & 3.2 & 1.3 & 0.1 & -1.3 \\ 
  Bias & SA-LC & 6.6 & 2.7 & 2.1 & 1.6 & 1.1 & 0.6 \\ 
  Accuracy & GeoWiki & 23.8 & 12.6 & 9.4 & 6.8 & 4.8 & 3.0 \\ 
  Accuracy & GlobCover & 42.3 & 27.3 & 23.3 & 18.8 & 15.6 & 11.2 \\ 
  Accuracy & MODIS & 33.8 & 21.5 & 18.4 & 15.3 & 12.7 & 10.6 \\ 
  Accuracy & SA-LC & 11.4 & 6.0 & 4.7 & 3.7 & 2.8 & 1.9 \\ 
   \hline
\hline
\end{longtable}

### 2015-10-18 Rmd/et-bias.Rmd
18/10/2015, second commit, ET updated. See changelog

@@
 
-## Calculate bias statistics
+## Calculate bias/accuracy statistics
 
-### Mean percent bias in ET estimates
+### Primary method
+
+Density dependent method
+
+```{r, eval=FALSE}
+# weighted mean functions
+wm <- function(x, w) stats::weighted.mean(x, w)
+wma <- function(x, w) stats::weighted.mean(abs(x), w)
+snms <- c("sa30", "globmu", "modmu", "geow")
+
+# Quick look at density dependent
+# reshape for ET mu
+err2011 <- lapply("f25", function(x) {
+  sapply(l2dag$mu, function(y) list(y[[x]]))
+})
+names(err2011) <- "f25"
+
+# calculate bias/MAE for data 
+a <- bias_statsw(gti_agg$f25, awgts, err2011$f25, snms, wm, "mu")
+b <- bias_statsw(gti_agg$f25, awgts, err2011$f25, snms, wma, "mua")
+et_muw <- rbind(a, b)
+
+# reshape for ET mu
+err2011 <- lapply("f25", function(x) {
+  sapply(l2dag$gs, function(y) list(y[[x]]))
+})
+names(err2011) <- "f25"
+
+# calculate bias/MAE for data
+a <- bias_statsw(gti_agg$f25, awgts, err2011$f25, snms, wm, "mu")
+b <- bias_statsw(gti_agg$f25, awgts, err2011$f25, snms, wma, "mua")
+et_gsw <- rbind(a, b)
+```
+
+Agricultural area method
 ```{r, eval=FALSE}
@@
 # bins <- cut(gti_agg$f25$gti, breaks = binv, include.lowest = TRUE)
-
-# weighted mean functions
-wm <- function(x, w) stats::weighted.mean(x, w)
-wma <- function(x, w) stats::weighted.mean(abs(x), w)
 
@@
 muas[, ol := NULL]
-setnames(muas, c("x", "il", "bias"), c("Variable", "Map", "Abs Bias"))
+setnames(muas, c("x", "il", "bias"), c("Variable", "Map", "MAE"))
 setkeyv(muas, c("Map", "Variable"))

### 2015-10-22 Rmd/carbon-bias.Rmd
22/10/2015 revision. Revisions through results. See changelog.

@@
 # yl <- round(yl / 10) * 10
-yl <- c(-290, 290)
-yax <- seq(yl[1], yl[2], 20)
+yl <- c(-150, 150)
+yax <- seq(yl[1], yl[2], 10)
 xax <- seq(1, 6, 5 / (length(yax) - 1))

### 2015-10-22 paper/main.tex
22/10/2015 revision. Revisions through results. See changelog.

@@
 \setlength\linenumbersep{2pt}
-\usepackage{placeins}
-\usepackage{float}
+%\usepackage{placeins}
+%\usepackage{float}
+%\restylefloat*{figure}
+%\usepackage{dblfloatfix}  % http://tex.stackexchange.com/questions/235623/placing-a-figure-in-the-bottom-of-a-page-spanning-the-two-columns-of-an-ieee-doc 
 %\usepackage{subfigure}
-\usepackage{wrapfig}
+%\usepackage{wrapfig}
+%\usepackage{caption}
 %\setlength\belowcaptionskip{-2cm}
 %\setlength{\textfloatsep}{5pt plus 1.0pt minus 2.0pt}
+\setlength{\parskip}{1cm plus4mm minus3mm}
 
@@
 
-\title{A Full Accounting of Landcover Map Error and Bias and Their Impacts on Assessments of Global Change}
+\title{Landcover Data Limits Our Understanding of Earth System Processes}
 %\title{Our Understanding of Global Change is Built on a Shaky Foundation}
@@
 
-\dropcap{T}he nature and distribution of landcover provides significant insight into econonomic processes \cite{lambin_modelling_1997} because human endeavors are so closely tied to how we transform land, whether it be the felling of ancient forests for farmland or converting that farmland into office parks.
-The vastness of our alteration of Earth’s landscapes suggests that landcover is a prime mediator of many environmental and social processes that drive or are affected by global change \cite{lambin_modelling_1997}, such as agricultural production and food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010,gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015,luoto_predicting_2004}, and changes in human demography \cite{linard_assessing_2010}.
-Like any view into nature, resolution and fidelity at fine scales are the keys to unlocking more granular and mechanistic insights into these processes \cite{see_improved_2015}.
-It is therefore unsurprising to see the explosive growth in private sector initiatives to develop new Earth observing capabilities, which range from small hobbyist drones\footnote{e.g.
-3DRobotics, DJIA} to satellite arrays\footnote{Planet Labs, Skybox}, in order to add value to industries such as agriculture, mining, and construction.
-This rapid growth in fine-scale landcover mapping capability is creating new opportunities to develop actionable information for traditionally public-sector concerns, such as agricultural development\footnote{USAID's Feed the Future}, drought and flood adaptation\footnote{Global Index Insurance Facility, www.indexinsuranceforum.org}, and carbon cycle management\footnote{United Nations REDD+, www.un-redd.org/aboutredd}.
-But while the demand for more nuanced, landcover-based insights is growing, there is only now the opportunity to use finer-scaled imagery to comprehensively interrogate the accuracy and biases in the landcover products that have become ubiquitous in global change research.
+\dropcap{T}he functioning of the Earth System\footnote{give definition here} is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface.
+Human endeavors are both strongly governed by and shape landcover, whether it be felling ancient forests for timber production or burning savannas to stimulate a green flush for livestock, while landcover is a primary driver of climate and biogeochemical processes \cite{lambin_dynamics_2003}.
+The vastness of our modification of the Earth's surface \cite{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover.
+To fully understand these processes and the nature of global change, it is therefore essential to know the nature and distribution of landcover.
 
-Global landcover data can only practically be derived from satellite imaging, but in many regions the cover types of interest are smaller, on average, than the sensor resolution, or spectrally indistinct from other neighboring covers, and these factors propagate classification errors \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}.
-The result is that landcover maps are generally inaccurate at finer scales and disagree substantially with one another, particularly in those parts of the world undergoing the most rapid land use changes \cite{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011}.
+This importance is understood by a growing number of social, economic, and physical scientists, who increasingly use landcover data to advance understanding in Earth System research areas ranging from food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010} to carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, and demographic shifts \cite{linard_assessing_2010}.
 
-Errors in landcover products are widely-acknowledged \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of efforts underway to improve landcover maps, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}.
-What is less known is the degree to which these errors bias analyses derived from the distributional and areal information in landcover.
-Errors are hard to quantify because spatially extensive reference data are not available for most regions of the world--particularly over Africa and other developing regions.
-Error assessments therefore typically rely on a small number of ground truth points for a bottom-up assessment or aggregated survey data for a top-down sanity check.
-For this reason, we have a better understanding of discrepancies between landcover datasets in relation to country-level statistics \cite{fritz_comparison_2010,fritz_cropland_2011,kaptue_tchuente_comparison_2011}, which offers little direction for how to arrive at a true number.
+The validity of the knowledge resulting from such studies depends on the veracity of the landcover data upon which they are based, much as a building requires a solid foundation if it is to remain standing.
+Unfortunately, the evidence so far suggests that Earth System science is being built on a shaky foundation.
+The reason for this is that landcover data can only practically be derived from satellite imagery, but in many regions the cover types of interest are smaller \cite[e.g.
+smallholder's farms][]{jain_mapping_2013} than the sensor resolution, or spectrally indistinct from neighboring covers, which translates into substantial mapping errors \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}.
+The result is that landcover maps are generally inaccurate at finer scales and disagree substantially with one another, particularly in those parts of the world undergoing the most rapid land use changes \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}.
+These errors mean that we are still unable to obtain the granular, mechanistic understanding of global change processes that we need.
 
-Being unable to fully quantify the errors in landcover maps of course makes it difficult, if not impossible, to quantify their impact on downstream analyses.
-There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, and carbon flux \cite{quaife_impact_2008} and human population estimates \cite{linard_assessing_2010}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}.
-One exception is a Belgian study  \cite{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover map errors bias measurements of agricultural land use patterns, but the study extent was fairly small and the validation data were discontiguous.
+These problems with landcover products are known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}.
+What remains an open question is exactly how much the maps Earth System researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of Earth System processes.
+Answering this question depends on having spatially comprehensive ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}.
+Our understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data.
+This allows us to quantify between map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.
 
-Just as a building needs a solid foundation, global change science needs to be based on sound landcover data.
-There is thus an urgent need to more precisely quantify landcover map errors and how these vary over large regions, particularly for the regions where landcover is changing most rapidly yet is most poorly known.
-We address this need here, using a unique, high accuracy agricultural landcover map for the entire country of South Africa to conduct a spatially comprehensive, bottom-up quantification of error in several latest generation landcover maps that are widely used in global change studies.
-We use these errors to assess the biases and accuracies of i) these landcover data, ii) how landcover properties influence map errors, iii) how bias and accuracy change with aggregation scale, with the specific goal of determining ``safe'' scales for drawing area-based inferences, and iv) how landcover error propagates through ``downstream'' analyses that represent several major global change research areas, including biogeochemical and land use change studies, food security assessments, land surface hydrology and climatology, and human geography.
+Being unable to fully quantify the errors in landcover maps makes it difficult, if not impossible, to gauge their impact on downstream analyses.
+There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, carbon flux measurements \cite{quaife_impact_2008}, and human population estimates \cite{linard_assessing_2010}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}.
+One exception is a Belgian study \cite{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover errors bias measurements of agricultural land use patterns, but the study extent was fairly small and the validation data were discontiguous.
+
+Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g.
+3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \cite{fritz_geo-wiki:_2012,estes_projected_2013,debats_generalized_????}, are finally providing the means to comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \cite{searchinger_high_2015}.
+
+In this study, we take advantage of these recent advances to address the urgent need to more thoroughly and precisely quantify landcover map errors and how they might impact our understanding of Earth System processes in the world's most dynamic regions.
+We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through ``downstream'' studies investigating into both the physical and socioeconomic components of the Earth System.
+Our objective is to provide scientists and policy-makers who use landcover data with a better, more up-to-date, understanding of their appropriate uses and limitations.
 
 \vspace{-0.5 cm}
-\section{Study area and landcover data}
-South Africa comprises nearly 6\% of sub-Saharan Africa's (SSA) landmass, and has a large, diverse agricultural sector, ranging from large commercial operations to smallholder farms \cite{hardy_rainfed_2011,estes_using_2014}.
-This diversity suggests that the country's agricultural landcover spans the range of types that are found throughout the rest of SSA.
+\section{Overview of study area and analyses}
+In the late 2000s, the South African government commissioned a cropland map that was made by manually interpreting and digitizing fields visible within high resolution satellite imagery \cite{fourie_better_2009}.
+The resulting vectorized field boundaries provide highly accurate data on field sizes and distribution for the period 2009-2011.
+This dataset is particularly valuable because South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \cite{searchinger_high_2015}, yet is notably lacking trustworthy maps of existing agriculture land \cite{fritz_comparison_2010}.
+Moreover, South Africa's large agricultural sector represents the diversity of farming systems found throughout SSA, ranging from large commercial operations to smallholder farms \cite{hardy_rainfed_2011,estes_using_2014}.
 
-The South African government commissioned a whole-country cropland boundary map to enhance its annual collection of agricultural statistics \cite{fourie_better_2009}.
-The map was made by trained workers who visually interpreted high resolution satellite imagery ($<$5 m SPOT imagery) and manually digitized field boundaries following a standardized mapping protocol.
-The resulting vectorized field maps, provide a unique, high quality reference dataset describing South African crop field distributions and size classes for the period 2009-2011, and are 97\% accurate in distinguishing cropland from non-cropland at 200 m resolution.
-We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create a gridded cropland reference map.
+We used this dataset as a reference layer for evaluating four landcover products representative of the type commonly used in Earth Systems research.
+The first was South Africa's own 30 m resolution 2009 National Landcover map (SA-LC)\cite{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are typically available only for individual countries \cite[e.g.][]{fry_completion_2009}.
+The second and third were respectively the 300 m GlobCover 2009 \cite{arino_global_2012} and 500 m resolution 2011 MODIS Landcover products, which are widely used global-scale products \cite[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
+The fourth dataset was the new 1 km GeoWiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}, which incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
+For comparison, we converted all five datasets into 1 km gridded maps of cropland density, expressed as the pixel-wise percent cover.
 
-We compared the reference map with similar maps derived from four existing landcover datasets.
-We obtained South Africa's 30 m resolution National Landcover map (SA-LC) for 2009 \cite{sanbi_national_2009}, the 500 m resolution MODIS Landcover for 2011 \cite{land_processes_distributed_active_archive_center_lp_daac_modis_2011, friedl_modis_2010}, the 300 m resolution GlobCover 2009 \cite{arino_global_2012}, and the new 1 km Geo-wiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}.
-We chose these particular datasets because they are nearly contemporaneous with our reference data, and represent the major types of landcover products used by researchers: SA-LC typifies the higher resolution, Landsat-derived maps that are developed individually for many countries \cite[e.g.][]{fry_completion_2009}, MODIS and GlobCover are widely used global-scale products \cite{gross_monitoring_2013, shackelford_conservation_2015}, while Geo-Wiki incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
-We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates, resulting in 4 maps (hereafter simply the ``cropland maps'') to compare to our reference map.
+In our first set of comparisons, we evaluated map quality by subtracting each of the four landcover product-derived percent cropland maps (hereafter we refer to these as the ``the test maps'') from the reference map, so that we could calculate each map's bias (the mean pixel-wise error) and accuracy (the mean absolute pixel-wise error, where a lower value indicates higher accuracy).
+We performed this analysis for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km, in order to evaluate how error and bias changes with scale.
+We also assessed how landcover pattern impacts map performance by modeling the correlation between map accuracy and cropland density.
+
+We then used these maps to conduct four further analyses typical of Earth System science research.
+The first two related to physical processes, namely the calculation of vegetative carbon stocks and the simulation of evapotranspiration by a hydrological model.
+The second two were socio-economic in nature: the estimation of agriculture yield and production, and an agent-based model assessment of household food security.
+The first and third analyses were relatively simple, in that the variable(s) of interest were mapped onto landcover using empirical relationships.
+The second and fourth relied on more complex numerical methods, where landcover was one of several variables needed to run each model.
+For the simpler analyses, we examined how results were influenced by map aggregation, while for the more complex cases, our assessments were confined to each numerical model's standard output resolution.
 
 \vspace{-0.5 cm}
-\section{Quantifying Error and Bias}
-We first calculated errors in pixel-wise estimated cropland percentages by differencing the reference map with each cropland map at each of the five aggregation scales (1-100 km).
-We used these errors to calculate map bias (the mean pixel-wise error) and accuracy (the mean absolute error, and how these vary with scale.
-Next, we assessed the degree to which the average cropland cover in agricultural landscapes, a descriptor of landcover pattern, impacts map accuracy.
-
-To investigate how landcover map error can impact downstream global change research, we quantified the biases and accuracies in five landcover-based analyses built upon our reference and cropland maps.
-The first three represented ``first-order'' analyses, in which a variable of interest is mapped onto a landcover type(s) using a simple empirical relationship.
-The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}.
-The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates.
-Maps based on these analyses underpin many assessments of crop productivity and production \cite[e.g.][]{foley_solutions_2011,licker_mind_2010}.
-
-The other two analyses can be considered second-order, wherein a process model draws on the cover types' values to calculate an output value.
-For the first of these, we used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model to calculate monthly evapotranspiration, using the reference and cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
-In the second example, we examined how these map errors impact the land allocation process of an agent-based food security model \cite{chen_dependency_2013}.
-
-\vspace{-0.5 cm}
-\section{Landcover maps}
+\section{Map quality}
 \subsection{Bias and accuracy}
 
-We created the 1 km reference map after removing all field types classified as communal/smallholder agriculture (individual fields in this category were not delineated, thus they were removed to prevent potential cropland area overestimates, SI) or permanent tree-crops (SI), and calculated the total cropland extent in the remaining area (1,081,000 km$^2$, or 90\% of South Africa).
-The 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.
+Our reference map indicated that crop fields covered 104,304 km$^2$, or nearly 10\%, of the total mapped area in the 2009-2011 time period.
+The test maps derived from SA-LC and GeoWiki overestimated this area by 31 and 10\%, respectively, while GlobCover and MODIS underestimated it by 18 and 23\%.
+Subtracting each test map from the reference maps created pixel-wise residuals, where negative and positive values respectively represent overestimates and underestimates by the test map (Fig.
+1A).
 
-We then aggregated the reference and each cropland map to 5, 10, 25, 50, and 100 km resolutions, and subtracted the four landcover maps from each reference (2007 and 2011) map at each scale of aggregation to assess error patterns (Fig.
-1A).
-Negative pixels here represent overestimation error, while positive values indicate underestimates.
-
-%\vspace{-0.5 cm}
-\begin{figure*}[ht]
+\begin{figure*}[t]
 %\begin{wrapfigure}{c}{1\textwidth}
-\centerline{\includegraphics[width=0.99\textwidth]{figures/figure1.pdf}}
-\caption{(A) Errors in the percent cropland estimates resulting from each of the four cropland maps relative to the reference map at different scale of pixel aggegation.
-Rows indicate the landcover map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation.
-White indicates areas with no data where communal farmlands or plantation forests were removed from analysis.
-(B) The bias (mean error) and accuracy (mean absolute error [MAE]) of each cropland map at each scale of aggregation.
-Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.
-The symbols indicate the method of calculating each metric, either by averaging across the entire country, within the union of  agricultural areas (cropland $>$0.05\%) for each reference-cropland map combination, or as the average of the averages calculated per 5\% bin of cropland cover, a measure of error that is independent of cropland density.}\label{afoto}
+\centerline{\includegraphics[width=0.95\textwidth]{figures/figure1.pdf}}
+\vspace{-0.15 cm}
+\caption{(A) Errors in the percent cropland estimates resulting from each of the four test maps relative to the reference map at different scale of pixel aggregation.
+Rows indicate the test map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation.
+White indicates areas where areas under communal farmlands or permanent tree crops were removed from analysis.
+(B) The bias (mean error) and accuracy (mean absolute error [MAE]) of each test map at each scale of aggregation, weighted by the percentage cropland in each cell of the reference map.
+Bias estimates are indicated by the semi-transparent bars, accuracy (lower is more accurate) by the solid bars, with bar colors coded to specific cropland maps.}
+\label{afoto1}
 \end{figure*}
@@
 \linenumbers
-The most pronounced errors were in the MODIS and GlobCover maps, which both underestimated cropland extent by 10-75\% in the center of the country (blue areas in Fig.
-1A, and the dominant production region), and overestimated along the eastern to northern margins (red areas in Fig.
-1A).
-All maps have less than +/-5\% bias (crosses in semi-transparent bars, Fig.
-1B), at all aggregation scales, when errors are averaged across the entire country, due to the extensive non-agricultural areas in the country that were effectively discerned by all maps.
-The whole country mean absolute error (the measure of map accuracy) were a few percent higher values in most cases, but still $<$10\%, except for GlobCover ($\sim$11\%) at 1 km resolution.
-Biases increase only slightly when errors are averaged within the combined agricultural regions (areas with greater than 0.05\% cropland, indicated by circles in transparent bars in Fig.
-1B) of the reference and each cropland map, but accuracy decreases substantially, with mean absolute errors increasing to 10\% for SA-LC and around 20\% for MODIS, GlobCover, and GeoWiki at 1 km resolution (circles in solid bars in Fig.
-1B), and remain between 5-10\% for all three datasets even at 25 km of aggregation.
+The most pronounced errors were in the MODIS and GlobCover maps, which showed large positive residuals in the center of the country where cropland is most concentrated (blue areas in Fig.
+1A), and negative residuals (red areas) along the eastern and northern margins.
+These patterns translated into substantial map bias (Fig.
+1.B), with GlobCover and MODIS mean error (weighted by the reference cropland density) exceeding 45\% and 25\% respectively at 1 km resolution, meaning that each map tends to underestimate cropland by that amount at that resolution.
+This bias declined with each level of map aggregation, being reduced to nearly 15\% for GlobCover and 5\% for MODIS at 100 km.
+The magnitude of (density weighted) mean absolute error (MAE) was somewhat higher in all cases.
+The GeoWiki map, in contrast, was the least biased overall, showing just a ~7\% bias at 1 km and near 0 for all other scales of aggregation, although its accuracy (23\% MAE) was only half as good as SA-LC's at 1 km (11\% MAE), which despite its uniform overestimation bias (Fig.
+1A) was the most accurate map at aggregation scales $<10$km.
+Above this, GeoWiki became slightly more accurate, having $<$5\% MAE at 100 km resolution.
+The reason GeoWiki had relatively poor accuracy at 1 km resolution was due to the heterogeneity of residuals, which traded between positive and negative residuals over short distances, thereby inflating MAE at this scale.
 
-As a third measure, we calculated bias and mean absolute error by averaging their mean per-ventile values (i.e.
-the average bias or accuracy value for areas have 0-5\% cropland, 5-10\% cropland, etc.).
-These provide density independent measures of map performance because they remove the dominance any particular level of cover has on map accuracy, which in this case was the 0-5\% cover class where map performance was best (SI).
-Calculated this way, bias and accuracy were substantially worse.
-The average bias jumped to 21\% for MODIS and 34\% for GlobCover at 1 km resolution (Fig.
-1B), meaning that each map had a strong tendency to underestimate cropland.
-This decreased with pixel aggregation, falling to 8\% at 50 km for MODIS, but remaining as high as 14\% at 100 km for GlobCover.
-SA-LC bias was smaller in magnitude, but tended towards cropland overestimates of 5-10\% at all aggregation scales.
-GeoWiki alone was relatively unbiased, but showed substantial inaccuracy with a density-independent mean absolute error of 23\%, which dropped below 10\% only above 25 km of aggregation.
-GlobCover was by far the most inaccurate, having $>$35\% mean absolute error at 1 km and 17\% at 100 km, following by MODIS (over 30\% at 1 km and 11\% at 100 km).
-SA-LC was the most accurate of all four maps with the density-independent measure, ranging from just above 10\% mean absolute error at 1 km to 5\% at 100 km.
+\subsection{Error and landcover density}
+Given that landcover characteristics can influence landcover map quality \cite{see_improved_2015, estes_diylandcover:_2015, debats_generalized_????}, we set out to elucidate the relationship between cropland configuration and map error.
+We first calculated the average 1 km reference cropland percentage within each of South Africa's 354 magisterial districts (the finest administrative unit, averaging 3,445 km$^2$; SI), which yielded a landscape-scaled metric of characteristic cropland density, and similarly calculated the district-wise MAE for each test map.
 
-\subsection{Landscape characteristics and error}
-The density-independent measures indicate that error and bias increase with cropland density, which suggests that map error varies as a function of the typical abundance of cover in the landscape being imaged.
-To investigate the relationship between cover density and map accuracy, we calculated the mean density for agricultural pixels ($>$0.05\% cropland) in the 1 km reference map within the boundaries of 354 magisterial districts (South Africa's finest administrative unit, averaging 3,445 km$^2$; SI Figure S3), providing a landscape-scale measure representing typical cropland density.
-We then extracted the cropland map errors for these same pixels, and calculated their district-wise mean absolute errors.
-
-A generalized additive model fit to district-level mean absolute error (log-transformed) shows that error peaks at 50-60\% cropland cover for all but the GlobCover map (which continued to increase with cropland cover), and is lowest when the landscape is dominated either by cropland or other cover types (Fig.
+We then used a generalized additive model to evaluate the shape of the relationship between district MAE and cropland density.
+The model shows that map accuracy is typically lowest at intermediate levels of cropland density (50-60\% cover) for all but the GlobCover map (where accuracy continues to decline with cropland cover), and is highest when the landscape is dominated either by cropland or by another type (Fig.
 2).
 In other words, accuracy is generally lowest when cropland cover is mixed evenly with other cover types.
-GlobCover's accuracy continued to decrease with cropland density because the dominant cover class contributing to the percentage cropland estimate was a mixed class defined as 50-70\% crops mingled with other vegetation, thus the maximum percentage was constrained by this mixture range.
+GlobCover's accuracy continued to decrease with cropland density because the dominant agricultural cover class contributing to the test map was defined as 50-70\% crops mingled with other vegetation, thus the maximum percentage was constrained by this mixture range.
+
+\vspace{-0.75 cm}
+\begin{figure}[h]
+\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
+\caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
+Prediction curves are color-coded to the different test maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}\label{afoto2}
+\end{figure}
+
+\section{The impact of map error on physical analyses}
+\subsection{Estimating vegetative carbon stocks}
+To understand the carbon cycle and climate forcing due to land use change, it is important to have accurate, high resolution maps of vegetative carbon stocks \cite[][]{searchinger_high_2015}.
+One widely used vegetative carbon dataset is that of \cite{ruesch_new_2008}, who mapped estimated carbon density values for different vegetation types to the classes of a global landcover product.
+The resulting data were intended to provide a baseline for climate policy by the Intergovernmental Panel on Climate Change (IPCC), as well as input to other land use and biogeochemical analyses \cite{ruesch_new_2008}.
 
 %\vspace{-1 cm}
+We followed this method to create vegetative carbon maps for South Africa.
+Since our maps represented cropland percentage, we developed several variants by assigning the carbon densities of different vegetation types (forest, secondary forest, shrubland, grassland, and sparse vegetation \cite{ruesch_new_2008}) to the non-cropland fraction of our maps.
+These hypothetical maps represented the range in potential carbon densities, and allowed us to investigate how carbon estimates can vary as a function of i) test map errors and ii) the properties of neighboring cover types.
+To assess carbon estimation error, we subtracted test map-derived carbon maps from those based on the reference map, and calculated bias and accuracy scores.
+
+The spatial patterns of test map errors transmitted into carbon estimation errors, but the sign varied as a function of the density of carbon adjacent to croplands (SI).
+Where cropland was underestimated and the surrounding cover type was more carbon dense than cropland, carbon density was overestimated, but when the cover type was less dense than croplands (e.g.
+sparse vegetation), then carbon density was underestimated.
+The inverse was true where cropland was overestimated.
+
+The magnitude of carbon errors varied as a function of the carbon density of surrounding cover, as demonstrated by the bias statistics (Fig.
+3).
+Bias was near zero when grassland was the adjacent cover type (SI), as its carbon density is nearly the same as cropland.
+However, when forest was adjacent then bias was a three- to five-fold multiple of cropland map bias (Fig.
+1B).
+At the most extreme, GlobCover's bias was -276\% at 1 km, but even SA-LC and GeoWiki had biases of 22\% and -50\%, respectively.
+Bias could be substantial even for the least carbon dense vegetation type (sparse), as evidenced by the 15-25\% bias at 1 km for MODIS and GlobCover under this class.
+The mean bias across the different potential adjacent vegetation classes ranged between -20 for GeoWiki and -123\% for GlobCover at 1 km (with MODIS in between these), while SA-LC's average bias was 11\%.
+Biases declined fairly rapidly with aggregation, with all datasets having an average (across cover types) bias magnitude of less than 10\% by 25 km of aggregation, except for GlobCover, which was -12\% at 100 km (SI).
+As with cropland percentages, GeoWiki produced the least biased carbon density estimates above above 1 km resolution.
+
+In terms of accuracy, MAE values were essentially the same as bias magnitudes, except for GeoWiki's, which were twice as large.
+The average MAE across vegetation classes was 47\% at 1 km, dropping to $<$10 only with 25 km of aggregation.
+In contrast, SA-LC's carbon estimates were twice as accurate at 1 km, and was slightly more accurate up to 25 km of aggregation were GeoWiki achieved parity.
+
+\vspace{-0.75 cm}
 \begin{figure}[ht]
-\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
-\caption{The relationship between map accuracy (the mean absolute error) in cropland maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
-Prediction curves are color-coded to the different cropland maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}
+\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scalew.pdf}}
+\caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map.
+Bias estimates (represented by symbols) fall within the semi-transparent floating bars, while accuracies are contained in the solid bars.
+Bar colors are coded to specific cropland map, symbols indicate which cover type was used to calculate cropland-adjacent carbon density.
+The bar represents the mean biases calculated across each of the 5 cover types.
+Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2).
+MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.}
 \label{afoto}
@@
 
-\section{First-order analyses}
-\subsection{Tier-1 carbon estimates}
-Using the methods provided by \cite{ruesch_new_2008}, we calculated average carbon densities for African forests, secondary forests, shrublands, croplands, grasslands, and sparse habitats (semi-arid grasslands and low shrublands), and assigned cropland carbon values to map cells in proportion to their cropland cover.
-For the non-cropland proportions, we assigned the carbon value from each of the other types, creating five different carbon maps for each landcover map at each aggregation scale (Fig.
-S4), which allowed us to test how carbon estimates vary as a function of i) cropland map bias and ii) the characteristics of adjacent cover types.
+\subsection{Evapotranspiration estimates}
+Accurate estimation of hydrological fluxes is critical to understanding how land-atmosphere interactions impact the climate system and runoff \cite{liang_simple_1994}.
+Land surface hydrological models, such as the Variable Infiltration Capacity \cite{liang_simple_1994}, are used to simulate these processes, and depend on landcover maps to provide information on the characteristics of vegetation and other materials covering the surface, as these govern the rates of runoff, infiltration, and evapo-transpiration.
+We used the VIC model to generate 25 km estimates of monthly evapotranspiration throughout South Africa, and examined how these were impacted by error in the test maps, which were used to determine the landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
 
-The difference between total carbon stocks for the country made using any of the cropland maps were within +/-3\% of those based on the reference map, regardless of which cover type was adjacent to cropland (Table S1), because the large of non-cropland in the country ($\sim$50-70\%, Fig.
-1) dilutes any map errors.
-Comparing total stocks between maps for just the agricultural area (30-50\% of the country) reveals much greater differences (Table S1).
-SA-LC overestimated carbon stocks by just 2\% when the adjacent cover type was forest, and up to 15\% when it was sparse cover.
-MODIS ranged from negligible differences in denser carbon classes (forest, secondary forest, and shrublands) to 8-13\% underestimates for the grassland and sparse classes.
-GeoWiki underestimated for all types, from $<$1\% for sparse cover to 8\% for forest.
-GlobCover grossly overestimated total carbon stocks for agricultural areas, varying from 64\% for sparse lands to 162\% for forest.
-The magnitude of this bias was due to false positives--GlobCover identified cropland in nearly 50\% of pixels, compared to 30\% for the other three cropland maps.
-
-The spatial patterns of errors in carbon estimates (Fig.
-S4) reflect those of the cropland maps (Fig.
-1).
-Where cropland was underestimated and the surrounding cover type was of higher carbon density than cropland, carbon density was overestimated.
-For lower density cover (grassland and sparse vegetation), carbon stocks were underestimated, but by small magnitudes.
-These tendencies were reflected in each map's biases, as calculated over the cropped areas of the country as jointly defined by the reference and each cropland map (Fig.
-3).
-For example, MODIS and GlobCover bias was $\sim$-50\% (overestimation) at 1 km resolution when forest was the cropland-adjacent cover (stars in semi-transparent green and gold bars, Fig.
-3; Table S2).
-For sparse vegetation (open circles in Fig.
-3), MODIS bias was 3\% at all scales, whereas maps that overestimated cropland (e.g.
-SA-LC, semi-transparent red) overestimated carbon density for this cover type, because cropland has a higher carbon density \cite{ruesch_new_2008}.
-Overall, GeoWiki had the lowest bias, for all cover types and all resolutions.
-Its worst bias was a tendency to overestimate by 12\% at 1 km when forest was adjacent, but at coarser scales this bias reduced to just a few percent (Fig.
-3, Table S2).
-All maps' biases are within +/-10\% bias after aggregation to 25 km.
-
-%%\vspace{-0.9 cm}
-%\begin{figure}[!ht]
-%\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scale.pdf}}
-%\caption{Biases and mean absolute errors in carbon densities derived from cropland maps, calculated as percents relative to the reference map.
-Bias estimates (represented by symbols) fall within the semi-transparent bars, while mean absolute errors are contained in the solid bars.
-Bar colors are coded to specific cropland map, while the symbols indicate which cover type was used to calculate cropland adjacent carbon density.
-The bar represents the mean biases calculated across each of the 5 cover types.
-Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2).}
-%\label{afoto}
-%\end{figure}
-
-The mean absolute error in carbon maps (solid colored bars in Fig.
-3) generally followed the same patterns, but with higher magnitudes and a few important differences.
-The most notable is that GeoWiki, despite relatively low bias, was highly inaccurate at 1 km, with 27\% mean absolute error across across cover types (line in solid blue bar, Fig.
-3, Table S2), which is close to the 36-37\% for GlobCover and MODIS.
-SA-LC had the lowest mean absolute error across scales, averaging (across cover types) 14\% at 1 km to 3\% at 100 km.
-(Fig.
-3, Table S2).
-The decrease in GeoWiki's accuracy relative to SA-LC's can be attributed to the highly heterogeneous nature of its cropland errors (Fig.
-1), which alternated between high magnitude positive and negative errors over relative short distances.
-
-\subsection{Gridded yield and production estimates}
-The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived cropland area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}.
-To be consistent with this methodology, we first adjusted our cropland maps according to this procedure, using the reference map to calculate total cropland area for each of South Africa's nine provinces, then updating the pixel-wise cropland percentages in the four cropland maps so that the province-wise sums matched the reference areas \cite[][and see SI]{ramankutty_farming_2008}.
-Despite this statistical constraint, the updated cropland maps still had substantial errors that were similar in pattern (Fig.
-S5) to those in the unadjusted maps (Fig.
-1), and we evaluated how these residuals affected gridded estimates of the yield and production of maize, South Africa's largest crop \cite{estes_comparing_2013}.
-To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize (South Africa's largest crop by area, \cite{estes_comparing_2013}) yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.
-
-The yields disaggregated onto the cropland maps were markedly different to those on the reference map, particularly in the lower density cropland areas in the center of the country, where GlobCover overestimated yields and MODIS and GeoWiki (to a lesser extent) underestimated them at 1 km resolution (Fig.
-S6).
-However, only GlobCover showed a notable bias in yields at this resolution, which was equivalent to nearly 60\% of the mean reference yield of 3.4 tons ha$^{-1}$.
-All other maps had biases of just +/-5\% at 1 km (Fig.
-5).
-Interestingly, GeoWiki and MODIS biases increased with aggregation, peaking at 10 km where both had underestimation biases of 30\%, thereafter declining to 10\% at 100 km.
-In contrast, GlobCover's yield bias declined linearly with aggregation (Fig.
-5).
-
-%\vspace{-1 cm}
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=1\textwidth]{figures/figure3.pdf}}
-\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in (A) carbon densities derived from cropland maps and (B) in disaggregated maize yield and production estimates.
-Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.
-For the carbon maps (A), symbols indicate the cover type used to calculate cropland-adjacent carbon density: the line represents the mean bias/MAE across all 5 cover types; shrubland and grassland values (near zero) and secondary forest (close to forest values) are not shown for clarity (see SI).
-For (B) symbols code the different variables (production and yield), normalized to their respective means.
-Values in A and B were calculated within the union of agricultural areas (cropland $>$0.05\%) for each reference-cropland map combination.}
-\label{afoto}
-\end{figure*}
-
-Production errors were completely unbiased (Fig.
-5).
-The statistical constraints on harvested and cropland areas resulted in the canceling out of spatial errors in production estimates, which is evident in the checkerboard-like pattern in maps of production biases (Figure S7).
-However, this reduction in bias comes at the cost of higher error magnitude, as the mean absolute error in production estimates were large, between 40 to 55\% for GeoWiki, MODIS, and GlobCover at 1 km, and remained generally high (10-28\%) even up to 50 km of aggregation (Fig.
-5).
-SA-LC production biases were lowest across all spatial scales (20\% at 1 km, dropping linearly to 2\% by 100 km).
-
-Absolute mean errors in yield were also substantial, and generally 10-15\% larger than production biases across all aggregation scales, except for GlobCover where absolute production biases exceeded yield bias at 5-100 km of aggregation.
-
-\section{Second-order analyses}
-\subsection{Evapotranspiration estimates}
-Compared to carbon and crop related examples, bias and accuracy in evapotranspiration (ET) calculated using the VIC model were small and averaged to less than +/-1\%.
-However, there were several error hotspots in the resulting ET difference maps (Fig.
-6).
-The most pronounced of these are the 5-15\% overestimates in the center resulting from VIC when initialized with MODIS and GlobCover, while overestimates along the southern and western coasts reached 25\%.
+Compared to the carbon analysis, the bias and accuracy in evapotranspiration (ET) calculated using the VIC model was negligible, averaging less than than +/-2\%.
+However, there were several error hotspots in the resulting ET residual maps (Fig.
+4).
+The most pronounced of these were the 5-15\% overestimates in the center of the country caused when VIC was initialized with MODIS and GlobCover, while overestimates along the southern and western coasts reached 25\%.
 These locations correspond primarily to the margins of major crop production regions--in the center is the westernmost boundary of the summer rainfall growing region, marked approximately by the 400 mm isohyet, where maize is the primary crop.
@@
 
-\begin{figure}[!ht]
+%\begin{figure*}[!b]
+%\centering
+%  \begin{minipage}[b]{.45\linewidth}
+%    \centerline{\includegraphics[scale=0.51]{figures/carbon_veg_scalew.pdf}}
+%    \caption{Caption}
+%    \label{afoto3}
+%  \end{minipage}\hfill%\qquad
+%  \begin{minipage}[b]{.45\linewidth}
+%    \centerline{\includegraphics[scale=0.51]{figures/et_bias_map.pdf}}
+%    \caption{Caption 2}
+%    \label{afoto4}
+%  \end{minipage}
+%\end{figure*}
+
+\vspace{-0.75 cm}
+\begin{figure}[h]
 \centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
-\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four cropland maps.}\label{afoto}
+\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
+\end{figure}
+
+\section{Socio-economic analyses}
+\subsection{Gridded crop yield and production data}
+The spatial variability of crop productivity and production is critical for understanding a host of social, economic, and environmental issues, such as food security, trade, and the potential for agricultural expansion and intensification \cite{licker_mind_2010,monfreda_farming_2008}.
+The most reliable source of such data are national to sub-national agricultural statistics, which are provided for relatively coarse-scaled administrative boundaries.
+To obtain higher spatial resolutions, efforts have been made to disaggregate these statistics using gridded landcover data \cite{ramankutty_farming_2008,monfreda_farming_2008}.
+These disaggregated datasets, which are constrained to match the agricultural statistics within the boundaries of the areas for which they are reported, have seen increasing use because they are considered to be more accurate than single source methods, and provide consistent data on which to base studies of global change \cite{ramankutty_farming_2008, see_improved_2015}
+
+We used these same methods  \cite{ramankutty_farming_2008,monfreda_farming_2008} to disaggregate maize harvested area (South Africa's largest crop \cite{estes_projected_2013}) on top of our reference and cropland maps, followed by yields, which were assigned to cells have harvested areas greater than zero.
+We then used these two layers to calculate maize production, and further aggregated the yield and production grids to 5, 10, 25, 50, and 100 km resolutions before quantifying the bias and accuracy of each test map's yield and production values.
+In this case, we could not convert cell-wise errors into percentages of the reference map values (because many cells had zero values for one map but not the other), so calculated bias and accuracy from the map residuals and then normalized their values to the reference map means.
+
+Maize yields disaggregated onto the test maps showed some marked differences relative to the reference map, but only at the margins of the major crop production areas where cropland is sparser (SI).
+These differences resulted when a yield value was mapped onto a grid cell where the reference map had no harvested area, and thus zero yield.
+In more densely cropped areas, such discrepancies were less frequent because both the reference and test maps were both likely to have some maize harvested area, and therefore a yield value.
+Yield biases were thus fairly low (and accuracy high), with the largest being 20\% for MODIS at 1 km, following by GlobCover with 10\% (Fig.
+5).
+These dropped to below 10\% with aggregation.
+
+Production biases were generally slightly higher, but still low, for most datasets, with the exception of GlobCover, which had an gigantic underestimation bias of $>$60\% (relative to mean production) at 1 km, which remained above 10\% even at 100 km of aggregation.
+MODIS production bias was above 20\% at 1 km, but declined to below 10\% at higher levels of aggregation.
+
+The accuracy of production estimates was another story.
+Here all datasets but SA-LC had MAE values of 30\% or higher below 25 km of aggregation (Fig.
+5), reaching as high at 100, 65, and 45\% at 1 km for GlobCover, MODIS, and GeoWiki, respectively.
+Only GeoWiki's MAE fell below 10\% with 100 km of aggregation.
+SA-LC estimated production was most accurate, having between 10-20\% MAE between 1 and 10 km, and $<$10\% at 25 km and higher.
+This low accuracy relative to the gridded yield measures relates to the disaggregation process for harvested area, which allocates a fractional value to each pixel, which is itself a fraction.
+The process of adjusting the gridded values so that their total match the statistics from which they are derived does not adjust map errors relative to the reference map, and the constraint in fact appears to shorten the distance between negative and positive residuals (SI), thereby increasing absolute errors.
+
+%\vspace{-0.5 cm}
+\begin{figure}[!h]
+\vspace{-0.75 cm}
+\centerline{\includegraphics[width=0.5\textwidth]{figures/yield_prod_bias.pdf}}
+\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates.
+Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.
+Symbols code the different variables (production and yield), normalized to their respective means.}
+\label{afoto}
 \end{figure}
@@
 
+The South African government commissioned a whole-country cropland boundary map to enhance its annual collection of agricultural statistics \cite{fourie_better_2009}.
+The map was made by trained workers who visually interpreted high resolution satellite imagery ($<$5 m SPOT imagery) and manually digitized field boundaries following a standardized mapping protocol.
+The resulting vectorized field maps, provide a unique, high quality reference dataset describing South African crop field distributions and size classes for the period 2009-2011, and are 97\% accurate in distinguishing cropland from non-cropland at 200 m resolution.
+We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create a gridded cropland reference map.
+
+We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates, resulting in 4 maps (hereafter simply the ``cropland maps'') to compare to our reference map.
+
+The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}.
+The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates.
+Maps based on these analyses underpin many assessments of crop productivity and production \cite[e.g.][]{foley_solutions_2011,licker_mind_2010}.
+
+For the first of these, we used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model to calculate monthly evapotranspiration, using the reference and cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
+In the second example, we examined how these map errors impact the land allocation process of an agent-based food security model \cite{chen_dependency_2013}.
+
+The reference dataset covered all of South Africa's field crop estate, from which we filtered out areas of communally owned farmlands and permanent tree crops to have a common basis of comparison across all landcover products, leaving us with map, after conversion to cropland percentage, that covered or 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.
+
+from agricultural pixels ($>$0.05\% cropland)
+
+The difference between total carbon stocks for the country made using any of the cropland maps were within +/-3\% of those based on the reference map, regardless of which cover type was adjacent to cropland (Table S1), as the large uncropped area of South Africa (Fig.
+1) dilutes the errors within the
+
+The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived cropland area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}.
+To be consistent with this methodology, we first adjusted our cropland maps according to this procedure, using the reference map to calculate total cropland area for each of South Africa's nine provinces, then updating the pixel-wise cropland percentages in the four cropland maps so that the province-wise sums matched the reference areas \cite[][and see SI]{ramankutty_farming_2008}.
+Despite this statistical constraint, the updated cropland maps still had substantial errors that were similar in pattern (Fig.
+S5) to those in the unadjusted maps (Fig.
+1), and we evaluated how these residuals affected gridded estimates of the yield and production of maize, South Africa's largest crop \cite{estes_comparing_2013}.
+To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize (South Africa's largest crop by area, \cite{estes_comparing_2013}) yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.
+
 \section{Digital RCD Analysis} 
-
 \end{materials}

### 2015-11-02 Rmd/agent-allocation-bias.Rmd
2/11/2015 commit. See changelog

@@
 # Examine, including in QGIS
-plot(md)
-plot(md[mdsel[mdsel < 314], ], col = "red", add = TRUE)  # filter out Cape ones
+# plot(md)
+# plot(md[mdsel[mdsel < 314], ], col = "red", add = TRUE)  # filter out Cape ones
 writeOGR(md[mdsel[mdsel < 314], ], dsn = "external/ext_data/md_abms.sqlite",
@@
             overwrite = TRUE)
-cpct <- extract(gti4, abm_md)
+cpct <- raster::extract(gti4, abm_md)
 dcrop_pct <- round(sapply(cpct, sum, na.rm = TRUE) / sapply(cpct, length), 1)
@@
 
+# Select modeled yields for agent-based model
+
+From earlier simulation, to give idea about variability
+```{r, eval = FALSE}
+dssat <- fread(full_path(p_data, "abm/dssat-yields.csv"))
+dssat_red <- copy(dssat[])
+
+dts <- c(2007288, 2007303, 2007318, 2007333, 2007348, 2007363, 2008013, 2008028)
+dssat[PDAT %in% dts, ]
+
+plot(1:31, 1:31, ylim = range(dssat$HWAH), pch = "", xlab = "YEAR", 
+     ylab = "HWAH")
+points(dssat[PDAT %in% dts & FNAM != "ZASP0001", HWAH])
+boxplot(dssat[PDAT %in% dts & FNAM == "ZASP0001", HWAH])
+boxplot(dssat[PDAT %in% dts & FNAM == "ZASP0002", HWAH])
+boxplot(dssat[PDAT %in% dts & FNAM == "ZASP0003", HWAH])
+dssat[PDAT %in% dts, mean(HWAH), by = FNAM]
+dssat[PDAT %in% dts, mean(HWAH), by = .(CU, FNAM)]
+
+plot(1:8, 1:8, ylim = range(dssat$HWAH), pch = "", xlab = "YEAR", 
+     ylab = "HWAH")
+for(i in unique(dssat$FNAM)) {
+  dssat[PDAT %in% dts & CU == 1 & FNAM == i, lines(HWAH)]
+}
+
+# provincial ag census statistics from 2007
+prodyld <- cbind.data.frame("md" = as.character(abm_md$name_2), 
+                            "pha" = c(6219, 6203, 6391, 9404), 
+                            "prod" = c(14781, 17434, 24366, 25578))
+prodyld$yld <- prodyld$prod / prodyld$pha
+
+outylds <- copy(dssat[PDAT %in% c(2007303, 2007348, 2008013) & 
+                   FNAM != "ZASP0001", ])
+outylds[PDAT %in% dts, mean(HWAH), by = .(FNAM)]
+outylds[PDAT %in% dts, mean(HWAH), by = CU]
+outylds[PDAT %in% dts, mean(HWAH), by = .(CU, FNAM)]
+write.csv(outylds, file = full_path(p_data, "abm/abm-yields.csv"))
+```
+
+<a href="#top">Back to top</a>
+
 # Downscaling and agent allocation
@@
 ```{r, eval = FALSE}
-abias <- data.table(read_excel(full_path(p_data, "abm/agent-bias2.xlsx")))
-abias <- abias[Alg == 1]
+abias <- data.table(read_excel(full_path(p_data, "abm/stats.xlsx")))
+# abias <- abias[Alg == 1]
 setkey(abias, MD)
@@
 
+# production deficits
+hhprod <- abias[var == "prod", gti / HHs]
+mdprodlc <- abias[var == "prod", 
+                  lapply(.(sa30, globmu, modmu, geow), function(x) x / HHs)]
+prdef <- abias[var == "prod", lapply(.(sa30, globmu, modmu, geow), function(x) {
+  (gti - x) / gti * 100
+})]  # district total deficit
+hhprodlc <- abias[var == "prod", 
+                  lapply(.(sa30, globmu, modmu, geow), function(x) x / HHs)]
+hhprod_def <- (hhprod - hhprodlc) / hhprod * 100  # household deficit
+
 cexs <- seq(1, 2, 1 / 3)
@@
 cols <- c(rep("red", 4), rep("orange3", 4), rep("green4", 4), rep("blue", 4))
-plot(unlist(pctbias), unlist(noland), col = cols, cex = cexs, pch = 20, 
-     ylab = "% HH without farms", xlab = "% error in cropland area", 
-     cex.lab = cx[2], cex.axis = cx[1])
 plot(unlist(pctbias), unlist(unalloc), col = cols, cex = cexs, pch = 20, 
@@
      ylab = "% land deficit", xlab = "% error in cropland area", 
+     cex.lab = cx[2], cex.axis = cx[1])
+plot(unlist(pctbias), unlist(hhprod_def), col = cols, cex = cexs, pch = 20, 
+     ylab = "% food deficit", xlab = "% error in cropland area", 
      cex.lab = cx[2], cex.axis = cx[1])

### 2015-11-02 Rmd/carbon-bias.Rmd
2/11/2015 commit. See changelog

@@
 cols <- c("red", "orange3", "green4", "blue")
-lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
+lcnms <- c("SA-LC", "GlobCover", "MODIS", "Geowiki")
 pnms <- c("Forest", "Secondary", "Shrubland", "Grassland", "Sparse")
@@
 mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
-legend(x = 12.4, y = -210, legend = lcnms, pch = 15, col = cols, adj = 0,
+legend(x = 12.4, y = -110, legend = lcnms, pch = 15, col = cols, adj = 0,
        pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
-legend(x = 11.9, y = -210, legend = rep("", 4), pch = 15, adj = 0, 
+legend(x = 11.9, y = -110, legend = rep("", 4), pch = 15, adj = 0, 
        col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
        cex = 0.8)
-text(x = 12.7, y = -212, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
-text(x = 12.2, y = -212, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
-legend(x = 12.1, y = 260, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
+text(x = 12.7, y = -115, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
+text(x = 12.2, y = -115, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
+legend(x = 12.1, y = 120, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
        pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)

### 2015-11-02 Rmd/cropland-bias.Rmd
2/11/2015 commit. See changelog

@@
 names(disagg) <- snms
+
+# writeRaster(disagg$globmu$f100, 
+#             filename = "external/ext_data/test/disagg100km.tif")
 ```
@@
 dev.off()
+
+lev <- names(gti_agg)
 ```

### 2015-11-02 paper/main.tex
2/11/2015 commit. See changelog

@@
 
-\title{Landcover Data Limits Our Understanding of Earth System Processes}
+\title{Landcover Data Limits Our Understanding of Global Change}
 %\title{Our Understanding of Global Change is Built on a Shaky Foundation}
@@
 \begin{abstract}
-{Blah blah.}
+{Guidelines: landcover maps based on coarser resolution individual sensors (e.g.
+MODIS) should be aggregated to 50-100 km resolution to minimize map bias and accuracy.
+Map accuracy is lowest at intermediate levels of cover, thus in highly mixed landscapes, such as small-holder farming systems or ecotones, even coarser aggregation may be necessary.
+Finer resolutions can be used with Landsat-based maps, if available, or for global analysis from new improved fusion products, particularly GeoWiki.
+The widely used datasets for carbon, crop area, and crop yield should be rebuilt on these updated datasets.
+Newer approaches to improving landcover maps, using interactive training with expert judgement and next generation classifiers are needed to create more reliable landcover, and in doing so will improve our understanding of global change.
+}
 \end{abstract}
@@
 
-\dropcap{T}he functioning of the Earth System\footnote{give definition here} is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface.
+\dropcap{T}he functioning of the Earth System is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface.
 Human endeavors are both strongly governed by and shape landcover, whether it be felling ancient forests for timber production or burning savannas to stimulate a green flush for livestock, while landcover is a primary driver of climate and biogeochemical processes \cite{lambin_dynamics_2003}.
@@
 
-This importance is understood by a growing number of social, economic, and physical scientists, who increasingly use landcover data to advance understanding in Earth System research areas ranging from food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010} to carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, and demographic shifts \cite{linard_assessing_2010}.
+This importance is understood by a growing number of social, economic, and physical scientists, who increasingly use landcover data to advance understanding of global change in areas ranging from food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010} to carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, and demographic shifts \cite{linard_assessing_2010}.
 
-The validity of the knowledge resulting from such studies depends on the veracity of the landcover data upon which they are based, much as a building requires a solid foundation if it is to remain standing.
-Unfortunately, the evidence so far suggests that Earth System science is being built on a shaky foundation.
+The value of the insights resulting from such studies depends on the veracity of the landcover data upon which they are based, much as a house requires a solid foundation in order to remain standing.
+Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations.
 The reason for this is that landcover data can only practically be derived from satellite imagery, but in many regions the cover types of interest are smaller \cite[e.g.
@@
 These problems with landcover products are known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}.
-What remains an open question is exactly how much the maps Earth System researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of Earth System processes.
+What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes.
 Answering this question depends on having spatially comprehensive ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}.
@@
 
-In this study, we take advantage of these recent advances to address the urgent need to more thoroughly and precisely quantify landcover map errors and how they might impact our understanding of Earth System processes in the world's most dynamic regions.
-We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through ``downstream'' studies investigating into both the physical and socioeconomic components of the Earth System.
+In this study, we take advantage of these recent advances to address the urgent need to more thoroughly and precisely quantify landcover map errors and how they might impact our understanding of change in the world's most dynamic regions.
+We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through ``downstream'' studies investigating into both the physical and socioeconomic components of global change.
 Our objective is to provide scientists and policy-makers who use landcover data with a better, more up-to-date, understanding of their appropriate uses and limitations.
@@
 
-\vspace{-0.75 cm}
-\begin{figure}[h]
-\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
-\caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
-Prediction curves are color-coded to the different test maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}\label{afoto2}
-\end{figure}
-
 \section{The impact of map error on physical analyses}
@@
 The resulting data were intended to provide a baseline for climate policy by the Intergovernmental Panel on Climate Change (IPCC), as well as input to other land use and biogeochemical analyses \cite{ruesch_new_2008}.
+
+%\vspace{-0.75 cm}
+\begin{figure}[!h]
+\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
+\caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
+Prediction curves are color-coded to the different test maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}\label{afoto2}
+\end{figure}
 
@@
 
-\vspace{-0.75 cm}
-\begin{figure}[ht]
+%\vspace{-0.75 cm}
+\begin{figure}[!h]
 \centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scalew.pdf}}
@@
 Land surface hydrological models, such as the Variable Infiltration Capacity \cite{liang_simple_1994}, are used to simulate these processes, and depend on landcover maps to provide information on the characteristics of vegetation and other materials covering the surface, as these govern the rates of runoff, infiltration, and evapo-transpiration.
-We used the VIC model to generate 25 km estimates of monthly evapotranspiration throughout South Africa, and examined how these were impacted by error in the test maps, which were used to determine the landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
+We used the VIC model to generate 25 km estimates of monthly evapotranspiration throughout South Africa, and examined how these were impacted by error in the test maps, which were used to determine landcover-specific values for leaf area index (LAI), plant rooting depth, aerodynamic roughness, and several other variables that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
 
@@
 
-\vspace{-0.75 cm}
-\begin{figure}[h]
+%\vspace{-0.75 cm}
+\begin{figure}[!h]
 \centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
@@
 
+\subsection{Agent-based simulation of food security}
+Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}.
+To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistic represents the social and biophysical features of the study region \cite{berger_creating_2006}.
+In our example, we used an ABM of household food security that simulates the food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}.
+The model is initialized such that each household is allocated its designated share of cropland, and then produces a crop over the course of one to many seasons.
+Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g.
+districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers.
+To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig.
+S8) in the eastern part of the country with 28-45\% of their areas devoted to cropland.
+The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity.
+The number of households and the cropland area per household is derived from survey data of communities where all cropland is owned.
+The model is thus considered to well calibrated when all households are allocated their appropriate area of cropland, and all cropland is occupied.
+
 %\vspace{-0.5 cm}
-\begin{figure}[!h]
-\vspace{-0.75 cm}
+\begin{figure}[!hb]
+%\vspace{-0.75 cm}
 \centerline{\includegraphics[width=0.5\textwidth]{figures/yield_prod_bias.pdf}}
@@
 
-\subsection{Initializing an agent-based model}
-Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}.
-A common feature of such ABMs is that they need to be calibrated against data describing the characteristics of land users, including an initialization step to assign land resources to ``agents'' representing the land users, wherein the simulated landscape pattern and distribution of agent resources matches those in the real world.
-In our example, we used an ABM of household food security that simulate the interactions between many individual farming households (the agents) and their environment over multiple seasons \cite{chen_dependency_2013}.
-We used cropland maps to provide the model the location and abundance of cropland, which is used to allocate an initial share of cropland to each simulated household.
-Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g.
-districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers.
-To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig.
-S8) in the eastern part of the country with 28-45\% of their areas devoted to cropland.
-The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity.
-The number of households and the cropland area per household is derived from survey data of communities where all cropland is owned.
-The model is thus considered adequately initialized when all households are allocated their appropriate area of cropland, and all cropland is occupied.
+We used the reference map and each cropland map to separately initialize the model, and examined how map errors impacted the land allocation process and household food production estimates.
+We examined three variables: the percent of unallocated cropland, land deficit, and food deficit.
+The percent of unallocated cropland measures how effective the model was in matching household agents to available cropland resources.
+Where error was negative (indicating a cropland overestimate by the test maps), the amount of land left unallocated had a straight one-to-one relationship with the percentage of overestimation (Fig.
+6A).
+When cropland was underestimated, no land was left unallocated until the underestimation exceeded 50\%.
+The MODIS-based simulation for districts 1 and 2 was most pronounced for this tendency, where 5-10\% of cropland remaining unallocated despite the fact that cropland with such a high cropland underestimate (85\%) the majority of households were not allocated cropland.
+This curious relationship occurred because croplands tend to cluster, and when underestimated clusters tend to be small, resulting in islands of cropland that fall outside of the search radius within which cropland is sought when agents are seeded onto the landscape.
 
-We used the reference map and each cropland map to separately initialize the model, and compared the agent allocation results to assess how cropland map errors impacted the initialization process.
-We examined three variable, the first being the number of agents that were not assigned fields, the second the amount of cropland left unallocated, and the third the area of land deficit, or the amount of land that should have been assigned to households but wasn't.
-For the first variable, there was a one-to-one relationship between the percentages by which cropland was underestimated and households that could not be assigned fields (Fig.
-6, left panel).
-The most extreme examples occurred when MODIS cropland initialized the ABM in districts 1 and 2, where $\sim$85\% of agents did not receive cropland.
-All households were assigned fields when total cropland area was overestimated (GeoWiki, SA-LC), but in these cases the area of cropland allocated to no one (the second variable) was proportional to the size of the overestimate (e.g.
-$\sim$20\% for SA-LC, Fig.
-6 right panel).
-Interestingly, the overall relationship between the percent of cropland allocated and percent cropland error was U-shaped, as the model also failed to give land to households when cropland was underestimated by more than 50\% (Fig.
-6, right panel).
-MODIS again provided the most pronounced results in districts 1 and 2, where 7-12\% of cropland was left unallocated despite the fact that 85\% of agents had no land.
-This curious relationship occurred because cropland tends to cluster, and when it is underestimated, the size of these clusters is small, resulting in islands of cropland that fall outside of the search radius (which is constrained by an absolute distance and the proximity of other agents) within which cropland is sought when agents are seeded onto the landscape.
-
-The last measure, land deficit, increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2--and would become infinite in the case of a 100\% underestimate.
+The second measure, land deficit, describes the amount of land that the model should have assigned but couldn't due to the mismatch between the cropland map and the statistical estimate of cropland holdings.
+Land deficit increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2 (Fig.
+6B)--and would become infinite in the case of a 100\% underestimate.
+This contrasted with the third measure, food deficit, the shortfall in the average amount of food that should have been produced by each household but wasn't because of the land deficit; this particular deficit increased linearly with the percentage of cropland underestimate (Fig.
+6C).
 
@@
 \begin{figure}[!ht]
-\centerline{\includegraphics[width=.5\textwidth]{figures/agent-bias.pdf}}
-\caption{Biases in agent-based model initialization relative to the district-wise errors (as a percent) in total cropland area, measured in terms of the percent of households having no cropland allocated (left), and the percent of cropland left unallocated (right).
+\centerline{\includegraphics[width=.5\textwidth]{figures/figure6.pdf}}
+\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the percent of cropland that should have been allocated to households but wasn't, and C) the food deficit, or the average food production across households resulting from inadequate cropland allocation.
 Dot sizes correspond to district numbers, colors represent the landcover map.}
@@
 \section{Discussion}
-This spatially comprehensive, bottom-up assessment of landcover map bias and inaccuracy and provides unique insight into their extent, causes, and consequences for understanding global change processes, made possible by a unique, high accuracy dataset that likely provides the truest measure of total cropland area and distribution that is currently available for this region.
-This dataset is of course not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission), while some of the cropland map error we found may have been caused by the slight temporal mismatches between the reference data and the original landcover datasets we used.
-However, our assessment (SI) suggests that these errors are small, and do not appreciably impact our findings, which is bolstered by previous work showing the large scale of disagreements between landcover map-based cropland area estimates and national inventory data \cite{fritz_comparison_2010}.
+This spatially comprehensive, bottom-up assessment of landcover map bias and accuracy provides new insight into their extent, causes, and consequences for understanding global change, made possible by a unique, high accuracy dataset that likely provides the truest measure of cropland area and distribution currently available for this region.
+This dataset is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission), while some of the error we identified may be due to slight temporal mismatches between the reference and test datasets.
+However, our assessment (SI) suggests that these errors are small, and do not appreciably impact our findings, which is bolstered by previous work showing the substantial errors and inconsistencies between landcover maps \cite{fritz_comparison_2010,gross_monitoring_2013}.
+These previous studies, which were conducted over broader extents than ours, also indicate that our findings should be generalizable beyond South Africa's borders, particularly to more rapidly developing, data poor regions.
 
-Our results suggest several guidelines for using landcover data in global change research, and contain some important implications for how understanding of global change processes based on the data, and associated policy decisions, may be affected.
-In terms of developing a base landcover map, the first rule of thumb is that standard landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, appear to be too biased to be useful without substantial aggregation.
-If we use the standard that bias within +/-10\% is acceptable, then at least 25-100 km of aggregation is needed to sufficiently cancel out the errors in the base landcover data and subsequent first order estimates built on them (Fig.
-3 \& 4).
-The upper range of aggregation scale is necessary if a mixed pixel class becomes dominant, as in the case with GlobCover, because these lead to underestimation bias that will persist until the pixel size becomes substantially greater than the average area of landscapes that are dominated by the cover type of interest, which can be $>$1000 km$^2$ in some of South Africa's farming regions.
+\subsection{General guidelines for selecting and using landcover data}
+Our results suggest several guidelines for selecting and using landcover data in global change research.
+First, landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions.
+Researchers might need to use these datasets for a variety of reasons, for instance, because they provide consistent, multiple landcover classes at a global scale, or allow changes to be tracked across large scales over time \cite{luoto_predicting_2004}.
+The appropriate scale of aggregation then depends on which error property the user wants to minimize.
+For instance, if the objective is to minimize systematic errors in order to calculate an unbiased mean (e.g.
+average carbon density), then at least 50-100 km of aggregation is needed to reduce map bias to within +/-10\% (Fig.
+1B, 3).
+If the user wants to maximize accuracy (accuracy is important for comparing differences between locations on a single map, or between dates at fixed locations), then even coarser aggregations are needed to reduce mean absolute error below 10\% (Fig.
+1, 3, 5).
 
-Maps derived from higher resolution sensors, such as the SA-LC dataset, if carefully done, do not have this mixed class problem, and are sufficiently unbiased for most applications with just 1-5 km of aggregation.
-However, such datasets are typically developed for specific countries, using varying methods, and can be hard to obtain.
-For broader scale analyses, the best option is to use newer generation maps such as GeoWiki (and the GLC-Share \footnote{GLC-Share.
-www.glcn.org} datasets for other cover types) which is relatively unbiased at 1 km resolution.
-GeoWiki's lower bias comes from its process of evaluating consensus between several landcover datasets (including the other three in this study), resulting in cropland probabilities that are converted to percentages by calibrating to statistical data \cite{fritz_cropland_2011,fritz_mapping_2015}.
-This method mirrors the ensemble averaging used by other fields (e.g.
+The characteristics of cover at the landscape scale may also require greater levels of aggregation to achieve sufficient accuracy.
+We found that accuracy is lowest where mixing of cover types is greatest (Fig.
+2), a result echoed in an Africa-wide change detection study made with MODIS \cite{gross_monitoring_2013}.
+Outside of South Africa, where smallholder-driven agriculture predominates \cite{lambin_estimating_????} such mixed landscapes are more common, thus greater levels of aggregation may be needed over these areas.
+Coarser aggregations are also necessary if the cover type of interest belongs primarily to a mixed class (as with GlobCover).
+This leads to underestimation bias (Fig.
+2) and high inaccuracy, which will persist until the aggregated pixel exceeds the characteristic size of the geographical area within which the cover type dominates.
+In South Africa's farming regions, this can exceed 1000 km$^2$.
+
+Maps derived from higher resolution sensors, such as the SA-LC dataset, do not have this mixture problem because the pixels are fine enough to be assigned to a single cover type.
+This characteristic leads to higher accuracy, which is what we observed here: the SA-LC dataset needed the least aggregation--just 5-10 km--to achieve $<$10\% MAE in most of the example applications we tested (note, however, another guideline in this result: even with high resolution maps some aggregation is needed to achieve higher accuracy).
+
+%needed This observation leads to a second rule of thumb: users should select datasets from higher resolution sensors where possible (assuming they were carefully developed), as these needed much less aggregation to return accurate estimates for most applications we tested here.
+For $<$10\% MAE, just 5-10 km of aggregation were needed (however, this is still much coarser than the 30 m resolution of the sensor itself).
+
+Unfortunately, Landsat-scaled maps are typically developed for specific countries, using varying methods, and can be hard to obtain (an exception to this is the newly released GLC30, which has a reported classification accuracy of 80\% \cite{chen_global_2015}).
+Our results also suggest that, despite their higher accuracy, they are not necessarily the least biased maps.
+Except for a few results (notably the 1 km carbon and maize production estimates), GeoWiki-based maps had the least amount of bias across most scales.
+GeoWiki's lower bias results from its process of calibrating the cropland percentages to reported agricultural statistics \cite{fritz_cropland_2011,fritz_mapping_2015}.
+The procedure is similar to the one we used in disaggregating maize harvested areas \cite{ramankutty_farming_2008, monfreda_farming_2008}, which led to fairly unbiased estimates of maize production and yield (Fig.
+5) above 1 km (except for GlobCover's production estimates).
+This result, together with GeoWiki's unbiased cropland estimates (Fig 1.B), indicates the value of fusing inventory data with remote sensing to reduce map bias.
+
+%The consensus method mirrors the ensemble averaging used by other fields (e.g.
 crop \cite{asseng_uncertainty_2013}, climate \cite{giorgi_calculation_2002}, and ecological modeling \cite{araujo_ensemble_2007}) to increase prediction confidence.
 
-GeoWiki's statistical constraint procedure is similar to the one we used \cite[following][]{ramankutty_farming_2008}, which produced unbiased maize production estimates (Fig.
-3) by eliminating bias in the adjusted cropland and harvested area maps that they were built upon.
-This result, together with GeoWiki's low bias, indicate the value of fusing inventory data with remote sensing.
-However, this method depends on the quality of inventory data, which are often suspect, particularly in Africa \cite{carletto_emperor_2013,fao_action_2013}.
-The statistical constraint also does not greatly improve map \emph{accuracy}, as evidenced by GeoWiki's 23\% mean absolute error in 1 km cropland percentage estimates (Fig.
-S1), which is only slightly more accurate than MODIS (31\%) but worse than SA-LC (11\%).
-GeoWiki is definitely most accurate among the large scale landcover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
-Statistical constraints only adjust map values at locations where cropland is identified, so their use it .
+Of course, for the constraint to be effective, the inventory data themselves must be faithful to reality, but in many places, particularly Africa, agricultural statistics are suspect \cite{carletto_emperor_2013, fao_action_2013}, which is a vulnerability of the fusion approach \cite{see_improved_2015}.
+Beyond this limitation, this method does not improve map accuracy, which is evident in GeoWiki's relatively high MAE for cropland percentage at 1 km (23\%, Fig.
+1B), and by all maps' (except SA-LC's) highly inaccurate maize production estimates (Fig.
+5).
+The reason that accuracy is not improved is because the statistical constraint cannot correct the omission and commission errors within the landcover maps.
 
-Map accuracy is perhaps more important than bias for landcover maps.
+These last observations suggest two further guidelines: 1) for high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \cite{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share.
+www.glcn.org} datasets for other cover types).
+But in either case some aggregation--5-25 km, depending on application and whether low bias or high accuracy is the objective--will still likely be necessary to achieve sufficiently low error.
+Additionally, these higher accuracy, lower bias product are made infrequently, thus applications requiring high frequency landcover data will depend on MODIS and confreres, as previously mentioned, and therefore much coarser aggregations.
 
-Broader regional implications - error higher elsewhere
+%GeoWiki was the most accurate among the large scale landcover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
 
-Main points:
+\subsection{Implications for global change analyses}
+Besides the preceding landcover data selection and usage guidelines, our findings also have broader implications for global change analyses and associated policy decisions.
+The simpler, most direct uses of landcover data have substantial to skew our understanding of global change processes.
+For example, the global carbon carbon density map of \cite{ruesch_new_2008}, of which we created facsimiles here, has been widely used, but our results suggest that any analyses of the spatial variability of carbon stocks, or of mean carbon densities, can be highly misleading if the map is not first sufficiently aggregated.
+For example, a study of deforestation risks in the Democratic Republic of the Congo estimated forest rents as a function of carbon densities based on this map.
+Although the map was aggregated to $\sim$50 km, we found that mean absolute errors can still reach 40\% at this scale (Fig.
+3).
+The same carbon map provided country-specific mean forest carbon densities for an evaluation of climate mitigation policies \cite{cattaneo_international_2010}, which our results suggest could be underestimated by 70\% (Fig.
+3).
+In contrast to these, an assessment of the congruence between carbon and biodiversity \cite{strassburg_global_2010} aggregated the carbon data to $\sim$110 km resolution, where accuracy is much higher.
 
-What we found, significance of study 
-\begin{itemize}
-  \item First large area quantification of spatial biases
-  \item How large those biases are, for one of the most widely spread (spreading landcovers)
-  \item Insight into causes of bias, and thus some understanding of where biases are likely to be greater or smaller
-  \item How much progress made in reducing it
-  \item Class type and bias
-\end{itemize}
+A related example can be seen in studies that create high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \cite[e.g.
+Figure 3 in][]{foley_solutions_2011}.
+Such maps give an inflated sense of the precision with which such locations can be identified.
+Although the study in this example is careful to report aggregated statistics, in which we can have greater confidence, it is not inconceivable that the maps themselves are used to inform policy.
+For this reason, it is advisable to degrade maps until their accuracy is of an acceptable standard (i.e.
+$\geq$100 km).
 
-\begin{itemize}
-  \item Bias decreases as function of scale
-  \item General bias patterns, appropriate use of landcover products, which landcover products
-  \item Appropriate scales of inference, by type of product - 
-  \item Aggregation improves results for landcover, generally, 
-  \item Sensor resolution, statistical resolution, and merging products have high value
-   \begin{itemize}
-     \item But don't remove spatial bias - absolute bias matters. Statistical constraints seems to just compress spatial biases to higher rates of turnover. Geo-wiki
-     \item But of course these types of data are then dependent on how accurate the statistical data are defining the constraint (cite emperor has no data)
-    \end{itemize}
-   \item Mixed landscapes increase the chances of omission and commission errors by increasing the number of cover classes, or because such landscapes are less spectrally distinct \cite{estes_diylandcover:_2015}
-   \item caveats: Only single country in South Africa.
-More commercial farming than many other countries, but results are still instructive.
-Analysis of error as function of landscape type suggests that areas where cropland is more mixed with natural vegetation have higher errors.
-These sorts of landscapes quite common in smallholder-dominated systems, thus suggests that biases may be even higher elsewhere on the continent.
-\end{itemize}
+Aggregating until acceptable accuracy is achieved will be particularly important in growing efforts to find tradeoffs between agricultural development and ecological protection \cite[e.g.][]{searchinger_high_2015,west_trading_2010}.
+Finding win-wins fundamentally requires accurate, fine-scale data in order to identify ``win-win'' solutions, which are comprised of areas where there is some deviation from an expected relationship; for example, an area of unusually low carbon density but high crop fertility \cite{searchinger_high_2015}.
+Without sufficient aggregation, it is difficult to know whether such efficiencies are due to real biophysical properties, or simply landcover error.
 
-Implications for understanding global change and policy: \\
+What then, of the impacts of landcover data on more complex analysis?
+Here the results appear to be mixed.
+Map error appears to have had little impact on the overall bias and accuracy of VIC's evapotranspiration (ET) estimates, which contrasts with a related study that examined how landcover maps impact rainfall simulations \cite{ge_impacts_2007}.
+There are two reasons for this.
+First, map accuracy increased nearly two-fold by aggregating to 25 km, the resolution of VIC.
+Second, VIC's landcover-derived variables (e.g.
+LAI curves, effective rooting depth) could dampen the effect of map errors on ET calculations when the values of these variables in the adjacent landcovers are similar to that of cropland, which is the case in much of VIC's landcover scheme for South Africa.
+In other regions, where the matrix landcover's LAI values depart substantially from cropland's (e.g.
+in forest), the ET errors would likely be much larger.
+The carbon example illustrates how cover properties can mute or amplify map error.
 
-Increasing awareness of need to have spatial assessments in global change analyses.
-Do things such as identify areas where yield gaps are high, or how much carbon or biodiversity will be lost to changes in land use, in order to try prioritize development \cite{searchinger_high_2015,newbold_global_2015}, or to understand coupled human-social dynamics, etc.
-\\
+Despite the low values of the overall error metrics, the hotspots of ET errors (Fig.
+4) are important, because they occurred at the climatic margins of the major crop growing regions, where a substantial share of farms use irrigation.
+Since VIC does not simulate irrigation, the magnitude of these errors is likely to have been underestimated, because irrigation substantially increases latent heat flux \cite{sacks_effects_2008}.
+This also suggests the potential for map error to magnify in a coupled land-atmosphere model (which could explain the difference between this result and that of Ge et al.
+\cite{ge_impacts_2007}.), and thereby give misleading insight into the climatic effects of land use.
 
-Our finding suggest:
+Relative to the ET example, map errors had a much larger impact on the agent-based model's results.
+For the model's most important metric of food security--household-level crop production--the result was fairly straightforward for a class of models where the interactions of multiple agents often produce analytically intractable outcomes \cite{janssen_empirically_2006}.
+In this case, the model was only sensitive to cropland underestimates, which had the perfectly predictable effect of lowering average household food production, which would in turn overstate the degree of food \emph{in}security.
+Overestimates did not matter in this case, because the number of households remained fixed, and their cropland holdings were constrained to match the survey statistics (but this would matter for models that allow these parameters to vary).
+Less predictable was the finding that the spatial distribution of cover can impact model parameterization, with land being left unallocated even when needed, thus such models also require spatial, not just statistical, accuracy in their base landcover maps.
 
-\begin{itemize}
-  \item Area-based estimates only safe at coarser scales of aggregation for most types of global change analyses, and primarily with constrained products. 
-  \item 50-100 km scale of aggregation reduces bias sufficiently.
-  \item Not so with unconstrained products
-  \item Assessments of spatial variability unsafe, for all products, bar one - finer country-scale product.
-Here you look at absolute bias.
-This is high in many products even at higher scales of aggregation.
-  \item This suggests that disaggregation approaches or paint by numbers approaches are nice maps, but can't give clear guidance about differences between grid cells, even when highly aggregated.
-[work on this]
-  \item can lead to misinformed policies
-  \begin{itemize}
-    \item E.g.
-Efforts to identify area where yield gaps are most pronounced and/or concentrated are likely to be highly misleading, leading to ineffective targeting of resources.
-Most informative simply to look at these areas at the political boundary resolution
-    \item Comparing carbon stocks against potential yield for tradeoff analysis, which may be done with conservation planning to find areas with high benefit/low-cost. Also misleading. 
-    \item Looking at land availability for cropland or biofuels expansion (look at biofuel paper for example)--land might not be as available as people think. Can lead to formulation of bad policy
-    \end{itemize}
-  \item Analyses of higher order interactions, biogeochemistry, human decision-making, also misleading (maybe pair this with yield example). 
-  \begin{itemize}
-     \item Our example here, ET estimates not heavily biased, but in marginal areas of low rainfall some pronounced differences.
-These are areas where irrigation is more common, but VIC doesn't simulate this, so absolute bias in those zones likely to be underestimated, and such regions can have substantial impacts on altering climate \cite{estes_changing_2014, sacks_effects_2008}.
-    \item Can skew understanding of more advanced attempts to understand the human factors that go into driving agricultural productivity. Examples here
-    \end{itemize}
-\end{itemize}
+The errors in the cropland data we tested were large, and it is unlikely that current ABM-based studies, mostly of small regions, are based on such inaccurate landcover data.
+However, as computing power increases, so, too, does the feasible application extent of ABMs, which then become more reliant on larger-scaled, more erroneous, landcover data, with the attendant risk that insights into socio-ecological systems arising from these models could be highly misleading.
 
-Way forward
-\begin{itemize}
-  \item For now, use latest generation products fusion products or more detailed country-level products
-  \item Avoid change detection based on landcover products, e.g. MODIS. 
-  \item But moving forward key will be developing new approaches to map landcover with much greater fidelity, e.g.
-scaling out approach that led to this dataset, combining with latest computer vision algorithms, etc.
-\end{itemize}
+Beyond these examples, there are other applications where landcover map errors could alter understanding and policy.
+One of these lies in assessments of land availability for agricultural expansion, which could be particularly erroneous if they use a ``residual approach'' \cite{lambin_estimating_????}, where existing cropland and other occupied cover types are used to filter a map of potential suitability.
+Our results suggest that the pronounced underestimation biases (Fig.
+1) we found could inflate land availability estimates, and therefore increase the risk of unjust land allocation policies \cite{rulli_global_2013}.
+Another example is the growing efforts to minimize the environmental and climate impacts on new agricultural expansion frontiers, particularly Africa's savannas, which requires identifying areas that provide the highest agricultural benefit for the lowest environmental cost \cite{searchinger_high_2015}.
+In order to be effective, such analyses must be undertaken at the fine scales--hectares, rather than square kilometers--at which land use decisions are made, thus their value for informing land use policy is contingent on the accuracy of their underlying landcover/land use data \cite{searchinger_high_2015}.
 
-%Some points from my notes last year: 
-%\begin{itemize}
-%  \item Bias as a function of scale
-%  \item Bias as a function of cropland cover
-%    \begin{enumerate}
-%      \item Classification algorithms are thus more error-prone where landcover is mixed/heterogenous.
-%      \item The exception to this lies in the GlobCover dataset, where bias primarily increases as a function of cropland cover.
-The reason for this is that GlobCover's cropland classes do not provide for 100\% cropland cover, so aggregation tends to exacerbate underestimates.
-%      \item Thus caution is needed when aggregating a mixed pixel class.  
-%        \begin{itemize} 
-%          \item An example illustrates this: take 4 1 km pixels, 2 of which are 100\% cropland, 2 of which are other cover types.
-Imagine a landcover product classifies 3 of these as cropland (2 correct, 1 an error of commission), using a cropland class that is defined as 50\% cropland.
-Aggregating the actual fraction by a factor of 4 will result in a new 4 km pixel having 50\% cropland, whereas aggregating the landcover product's pixels will give just 38\% cropland, even when factoring in the incorrect classification.)
-%          \end{itemize}
-%    \end{enumerate}
-%  \item Bias as a function of method
-%      \begin{enumerate}
-%        \item Higher resolution and ensemble-based approaches have less bias.
-%        \item geowiki represents a fusion of multiple coarse resolution data sources that has undergone extensive validation using a crowdsourcing approach
-%        \item the SALC dataset is based on 30 m landsat data, but incorporates a range of ancillary data and expert judgement
-%        \item MODIS and GlobCover data are effectively single source/single algorithm.
-%        \item \textbf{Newer points begin here}
-%        \item Statistically constrained constrained landcover estimation approaches provide accurate area-based inferences when aggregated.
-But spatial errors are still high, as seen with GeoWiki and production/yield estimates.
-Using these to identify yield gaps at specific map locations is inappropriate, or even for a larger location if it does not coincide with the geographic boundaries of the statistical unit.
-%        \item Constrained estimates are also dependent on the accuracy of the statistics.  
-%    \end{enumerate}
-%  \item Fix above to have section on bias for global change studies
-%    \begin{enumerate}
-%      \item Scales at which it is safe to estimate values of say carbon stocks. 
-%      \item Above point about bias in disaggregated yield estimates - no point mapping these out.
-A new approach might be to take these statistically reported yields and then combine them with satellite data to estimate yield variability within the district.
-That way would have meaningful reason for disaggregating yields, and would be pegged to real yield values, which would help minimize errors in remote sensing of yields.
-%      \item Something on ET - doesn't seem to matter much, but land-atmosphere interactions can make these discrepancies meaningful, particularly since biases occur in arid areas where a lot of irrigation happens--can cause significant impacts on regional climate.
-etc.
-etc.
-Also we didn't change out land cover types, and the vegetation in SA around the cropland will have reasonably similar LAI and ET responses (I think), thus impact more muted than it might be elsewhere (e.g.
-in forested landscapes).
-%      \item Agent-based models. \textcolor{red}{Tom, Peng, something of significance/implications of this, please}
-%    \end{enumerate}
-%  \item Will need a section on way forward for data, etc.
-Key role of accurate landcover, particularly agricultural.
-New methods, vectorized field boundaries seem to be highly valuable, Mapping Africa, Stephanie's paper, Geo-Wiki, etc are the way ahead.
-%  
-%    
-%\end{itemize}
+\subsection{The way forward}
+Our analysis demonstrates that the landcover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on how accuracy or low bias.
+We provide some basic guidelines for selecting and using landcover data that can help increase confidence in the conclusions drawn from landcover-based studies.
+
+Beyond these guidelines, we recommend updating several important, global-scale datasets using the latest generation landcover maps.
+For example, rebuilding the widely used cropland distribution and yield maps \cite{monfreda_farming_2008, ramankutty_farming_2008} using the GeoWiki map could substantially improve our knowledge about current global agricultural distribution and productivity.
+Similarly, reconstructing the carbon map of \cite{ruesch_new_2008} using the GLC-Share datasets (in which GeoWiki provides the cropland layer) would improve our confidence in regional to global assessments of vegetative carbon stocks.
+Alternatively, the new high-resolution, Landsat-based GLC30 \cite{chen_global_2015}, aggregated and converted to fractional cover classes at 1 km--and fused with high quality national landcover maps, where available \cite[similar to][]{fritz_mapping_2015,fritz_cropland_2011}--may provide a preferable base map for rebuilding these datasets.
+
+Even more promising are newer approaches that fuse human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power.
+The GeoWiki project and the new high resolution global forest change map \cite{hansen_high-resolution_2013} provide the most prominent examples of these latest methods.
+Another approach is that of the Mapping Africa project\footnote{http://mappingafrica.princeton.edu}, which, inspired by quality of the reference map used in this study, enlists internet-based workers to manually digitize cropland boundaries \cite{estes_diylandcover:_2015}.
+Merging this approach with newer computer vision-based mapping algorithms \cite{debats_generalized_????} is a near-term goal of this project, which will combine the superior capacity of humans to delineate discrete cover types in noisy images \cite{estes_diylandcover:_2015} with the algorithm's speed and ability to classify highly dimensioned image features.
+These ongoing projects represent the next generation of landcover maps, and offer to greatly improve our understanding of global change.
 
@@
 To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize (South Africa's largest crop by area, \cite{estes_comparing_2013}) yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.
+
+(which is constrained by an absolute distance and the proximity of other agents)
 

### 2015-11-03 Rmd/overview.Rmd
Full draft completed, ready for some edits.

@@
     theme: spacelab
-    toc: yes
 ---
 
-# Introduction
+## Overview
 
-## Sidebar
-I am just learning how to use Rmarkdown with Rmd files to document analyses, primarily analyses that are more interactive in nature and notrepeated many times.
-Code that will be re-used multiple times will go into a library.
+The following files contain the code used to conduct each portion of the analysis described in the accompanying paper __Landcover Data Limits Our Understanding of Global Change__.
 
-One things I am noticing is that the knit process for compiling the notebook does not find functions within code chunks unless the libraries are called within the code chunks themselves.
-It seems like each code chunk is its own environment.
-The solution to this is to do something like `raster::raster()` when you want a function, or run this where chunks are set to {r, eval = FALSE}, and just run the code chunk as if you were in a normal R script working interactively.
-I have already set up the library for SAcropland to call the packages its needs so they are loaded up in the global environment.
+#### [Preparing reference maps](cropland-preprocess.html)
+#### [Downloading and preparing cropland maps](fetch-land-cover.html)
+#### [Cropland map bias and accuracy](compare-landcover.html)
+#### [Accuracy versus cropland density](magisterial-district-bias.html)
+#### [Carbon bias and accuracy](carbon-bias.html)
+#### [Evapotranspiration bias and accuracy](et-bias.html)
+#### [Yield and production bias and accuracy](yield-bias.html)
+#### [Agent-based model sensitivity](agent-allocation-bias.html)

### 2015-11-03 paper/main.tex
Full draft completed, ready for some edits.

@@
 \begin{abstract}
-{Guidelines: landcover maps based on coarser resolution individual sensors (e.g.
-MODIS) should be aggregated to 50-100 km resolution to minimize map bias and accuracy.
-Map accuracy is lowest at intermediate levels of cover, thus in highly mixed landscapes, such as small-holder farming systems or ecotones, even coarser aggregation may be necessary.
-Finer resolutions can be used with Landsat-based maps, if available, or for global analysis from new improved fusion products, particularly GeoWiki.
-The widely used datasets for carbon, crop area, and crop yield should be rebuilt on these updated datasets.
-Newer approaches to improving landcover maps, using interactive training with expert judgement and next generation classifiers are needed to create more reliable landcover, and in doing so will improve our understanding of global change.
+{Our knowledge of global change processes is built upon landcover data, but we lack the means to comprehensively assess landcover maps, particularly in rapidly developing regions, and thus do not fully comprehend how map quality impacts our understanding of global change.
+We used a unique, high quality, national-scale cropland map to assess the bias and accuracy in a) current generation landcover products, and b) analyses built on these.
+Maps have high bias and low accuracy, particularly at finer resolutions (1-10 km), which transfers into, and is often amplified, in downstream analyses of vegetative carbon stock, gridded crop production, and process-based food security models.
+Substantial map aggregation is needed to avoid potential misunderstandings of global change processes or misguided policy.
+Maps derived from coarse resolution sensors (such as MODIS and GlobCover) should be aggregated to at least 50-100 km resolution (and sometimes more).
+Map accuracy is lowest at intermediate levels of cover, thus coarser aggregation may be necessary in highly mixed landscapes.
+Higher resolution imaging systems (e.g.
+Landsat) produce maps that require less aggregation (1-25 km) to achieve high accuracy and low bias.
+For regional to global analyses, new fusion products, such as the GeoWiki cropland map, need just 5 km of aggregation to produce unbiased maps for most applications, but can require $\geq$25 km of aggregation to have sufficient accuracy.
+Widely used global datasets for carbon, crop area, and crop yield should be rebuilt on GeoWiki and similar maps, while new approaches methods that fuse human judgement, the latest computer vision-based algorithms, and growing high resolution image archives are needed to produce the next generation of landcover maps.
 }
@@
 \dropcap{T}he functioning of the Earth System is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface.
-Human endeavors are both strongly governed by and shape landcover, whether it be felling ancient forests for timber production or burning savannas to stimulate a green flush for livestock, while landcover is a primary driver of climate and biogeochemical processes \cite{lambin_dynamics_2003}.
-The vastness of our modification of the Earth's surface \cite{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover.
-To fully understand these processes and the nature of global change, it is therefore essential to know the nature and distribution of landcover.
+Human endeavors are strongly governed by and shape landcover, whether it be felling forests for timber or burning savannas to rejuvenate forage, and landcover is a primary factor in many climate and biogeochemical processes \cite{lambin_dynamics_2003}.
+Our increasing modification of the Earth's surface \cite{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover.
+To fully understand these processes and the nature of global change, it is essential to know the nature and distribution of landcover.
 
-This importance is understood by a growing number of social, economic, and physical scientists, who increasingly use landcover data to advance understanding of global change in areas ranging from food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010} to carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, and demographic shifts \cite{linard_assessing_2010}.
+This importance is understood by a growing number of social, economic, and natural scientists, who are using landcover data to advance understanding of food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, demographic shifts \cite{linard_assessing_2010}, and other important facets of global change.
 
-The value of the insights resulting from such studies depends on the veracity of the landcover data upon which they are based, much as a house requires a solid foundation in order to remain standing.
+The value of the insights resulting from such studies depends upon the veracity of their underlying landcover data, much as a house requires a solid foundation in order to remain standing.
 Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations.
-The reason for this is that landcover data can only practically be derived from satellite imagery, but in many regions the cover types of interest are smaller \cite[e.g.
+The reason for this is that landcover data can only practically be derived from satellite imagery, and in many regions the cover types of interest are smaller \cite[e.g.
 smallholder's farms][]{jain_mapping_2013} than the sensor resolution, or spectrally indistinct from neighboring covers, which translates into substantial mapping errors \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}.
-The result is that landcover maps are generally inaccurate at finer scales and disagree substantially with one another, particularly in those parts of the world undergoing the most rapid land use changes \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}.
+Landcover maps are therefore often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}.
 These errors mean that we are still unable to obtain the granular, mechanistic understanding of global change processes that we need.
@@
 Our understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data.
-This allows us to quantify between map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.
+This allows us to quantify between-map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.
 
@@
 \begin{figure}[!h]
-\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scalew.pdf}}
+\centerline{\includegraphics[width=.5\textwidth]{figures/figure3.pdf}}
 \caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map.
@@
 To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig.
-S8) in the eastern part of the country with 28-45\% of their areas devoted to cropland.
+S8) in the eastern part of the country with similar climate and 28-45\% of their areas devoted to cropland.
 The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity.
-The number of households and the cropland area per household is derived from survey data of communities where all cropland is owned.
-The model is thus considered to well calibrated when all households are allocated their appropriate area of cropland, and all cropland is occupied.
+The number of households and the cropland area per household was derived from surveys of farming communities where all cropland is used.
+The model was thus considered to well calibrated when all households were allocated their appropriate area of cropland, and all cropland was allocated.
 
@@
 
-Even more promising are newer approaches that fuse human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power.
-The GeoWiki project and the new high resolution global forest change map \cite{hansen_high-resolution_2013} provide the most prominent examples of these latest methods.
+Fusing human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power appears to offer the greatest promise for improving the quality of landcover maps.
+The GeoWiki project and the new high resolution global forest change map \cite{hansen_high-resolution_2013} are the most advanced examples of these newer methods.
 Another approach is that of the Mapping Africa project\footnote{http://mappingafrica.princeton.edu}, which, inspired by quality of the reference map used in this study, enlists internet-based workers to manually digitize cropland boundaries \cite{estes_diylandcover:_2015}.
-Merging this approach with newer computer vision-based mapping algorithms \cite{debats_generalized_????} is a near-term goal of this project, which will combine the superior capacity of humans to delineate discrete cover types in noisy images \cite{estes_diylandcover:_2015} with the algorithm's speed and ability to classify highly dimensioned image features.
-These ongoing projects represent the next generation of landcover maps, and offer to greatly improve our understanding of global change.
+Merging this crowdsourced approach to vector map creation with newer computer vision-based mapping algorithms \cite{debats_generalized_????} is a near-term goal of this project, as the former can interactively provide high quality training and test data (by using humans' superior capacity to delineate discrete cover types in noisy images \cite{estes_diylandcover:_2015}) for the latter, which has the advantage of speed and the ability to handle high-dimensional datasets.
+Approaches such as these will be needed to provide the next generation of landcover maps, so that we can gain a clearer, unbiased understanding of global change.
 
@@
 \section{Methods} 
-Perhaps it is right {\it SI Materials and Methods}.
+To establish the accuracy of the cropland boundary reference map, we randomly selected 609 cell from a 1 km$^2$ grid covering South Africa, weighted by the probability of cropland occurrence \cite{estes_diylandcover:_2015}.
+We sub-divided each cells into 200 m sub-cells, and visually interpreted whether crop fields presence/absence within these against Google Earth imagery, and compared these results to the digitized field polygons to determine the reference map's sensitivity (0.89) and specificity (0.99), overall map accuracy (97\%), and its true skill statistic \cite[0.88;][]{allouche_assessing_2006}.
+We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create the gridded cropland reference map.
+In the resulting grid, we masked out areas classified as communal farmland, because only their outer perimeters were digitized \cite{fourie_better_2009}, which posed a risk of overestimating cropland extent.
+We also excluded the permanent tree crop class, to avoid confusion with landcover maps that include these in their cropland classes.
+For the same reason, we used other high resolution landcover maps for selected provinces of South Africa to mask out commercially afforested areas and sugarcane plantations, which were also not included in the reference map.
+The resulting masked reference grid covered 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ was cropland.
+We extracted the cropland classes from SA-LC, MODIS, and GlobCover, and converted these into percent cropland estimates at 1 km resolution (GeoWiki is already a cropland percentage map).
+Both MODIS and GlobCover had mixed pixel classes, thus we used the upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages.
+We used the mean map for the main analysis, but estimated error variability using all three versions (SI).\\
+\indent To uncover the relationship between map error and cover density, we used magisterial district boundaries (n=354, mean area=3,445 km$^2$, see SI) to provide a landscape-scaled unit for calculating characteristic cover density.
+We filtered out pixels with $<$0.05\% (0.5 ha) cropland, to prevent the much larger areas of non-agricultural districts from drowning out any relationship.
+We extracted the absolute values of test map errors and corresponding reference cropland percentages, calculated their district-wide means, and then modeled the relationship between mean absolute error (response) and cropland density (predictor) as a polynomial function within a generalized additive model.
+We accounted for potential spatial autocorrelation by fitting a two-dimensional smoothing spline to each district's centroid coordinates, and weighted the model by the number of agricultural pixels per district \cite{wood_mgcv:_2001}.\\
+\indent To develop the carbon maps, we followed the methods used by Ruesch and Gibbs \cite{ruesch_new_2008} to create their vegetative carbon map.
+We took the average of the biome specific carbon density values they provided for African forests, secondary forests, shrublands, grasslands, sparse habitats, and croplands.
+We multiplied cropland densities by cropland fractions and added these to each of the five other densities multiplied by the residual non-cropland fractions to create five different carbon density maps.
+We aggregated each carbon map up to the five coarser resolutions for scaling comparisons.
+\\
+\indent We used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model run with the Africa Flood and Drought Monitor's meteorological data \cite{sheffield_drought_2013} to produce monthly gridded evapotranspiration estimates for South Africa for the years 1979-2010.
+We adjusted VIC's native landcover scheme so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25${\circ}$ resolution).
+We ran one instance of VIC for each of the five landcover permutations, and compared the mean annual ET produced by the reference map variant with those from the test maps.
+\\
+\indent To create the gridded maize yield and production maps, we followed \cite{ramankutty_farming_2008} to first calibrate the test cropland maps against cropland areas reported by administrative-level agricultural censuses.
+In place of census reports, we used the reference map to calculate total cropland areas for South Africa's nine provinces, then adjusted the pixel-wise cropland percentages in the four test maps so that their province-wise sums matched these totals (SI).
+We then followed Monfreda et al's procedure \cite{monfreda_farming_2008} to disaggregate maize \cite[South Africa's largest crop][]{estes_comparing_2013} planted area and yields onto the reference and adjusted test maps.
+Per \cite{monfreda_farming_2008}, the gridded planted area and yield estimates were obtained from magisterial district-level agricultural censuses, in this case for the most recent census of 2007 \cite{statistics_south_africa_commercial_2007}.
+We then aggregated gridded harvested areas and yields (weighted by the corresponding harvested area) to each coarser resolution, and multiplied yield by harvested area to calculate production.\\
+\indent The agent-based model of food security used here calculates average household-level food production over multiple seasons based on the biophysical environment (soil properties, weather), management inputs (planting decisions, labor availability), and farm assets (total field area) \cite{chen_dependency_2013}.
+The frequency distributions of household cropland holdings were derived from household survey data obtained from Zambia (no equivalent data were available for South Africa).
+We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area (2.2 ha), and preserved the cropland area distributions by multiplying the total number of households by the frequencies.
+To create cropland surfaces for the, we disaggregated all five cropland maps to binary cropland/non-cropland cover types with 100 m resolution, in order to match the survey-recorded average field size of 1 ha.
+In initializing, the model takes a weighted (by cropland area frequency) random draw of 100 households and places these within a district, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household.
+This process is iterated until all households are assigned cropland, or all available cropland is allocated.
+The model then runs separately for each district, with the household agents planting and managing a maize crop throughout the growing season, with possible final maize yields drawn from a look-up table that relates yield to particular climatic outcomes and management decisions.
+We ran the model for a single season, once per district and when initialized by each of the cropland maps (20 simulations total).
+\\
+\indent In calculating map error statistics, and with the exception of the agent-based model results, we use the reference cropland percentages to calculate the weighted mean map error (bias) and mean absolute map error (accuracy).
+For the comparisons between aggregated map values, we applied a further weight--the number of pixels contributing to each aggregated pixel, to prevent pixels close to national boundaries or where non-target cover types were masked out from having outsize influence on the statistics.
 
-Describe weighted mean bias reasons.
-
-We disaggregated the cropland percentages in all maps to binary cropland/non-cropland cover types with 100 m resolution, which matches the typical field size (1 ha) for smallholder farmers in household survey data (collected in Zambia) used in developing the agent-based model \cite{chen_dependency_2013}.
-The surveys provided the mean cropland area per household (2.2 ha) and frequencies distribution of cropland area holdings across households (e.g.
-how many households have 1 ha, 2 ha, etc.).
-We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area, and preserved the cropland area distributions by multiplying the total number of households by the frequencies.
-We then initialized the model, which takes a weighted (by cropland area frequency) random draw of 100 households and places these within the district, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household.
-This process is iterated until all households are assigned cropland, or all available cropland is allocated.
-
-The South African government commissioned a whole-country cropland boundary map to enhance its annual collection of agricultural statistics \cite{fourie_better_2009}.
-The map was made by trained workers who visually interpreted high resolution satellite imagery ($<$5 m SPOT imagery) and manually digitized field boundaries following a standardized mapping protocol.
-The resulting vectorized field maps, provide a unique, high quality reference dataset describing South African crop field distributions and size classes for the period 2009-2011, and are 97\% accurate in distinguishing cropland from non-cropland at 200 m resolution.
-We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create a gridded cropland reference map.
-
-We extracted the cropland classes from the first three datasets and converted these to 1 km resolution percent cropland estimates, resulting in 4 maps (hereafter simply the ``cropland maps'') to compare to our reference map.
-
-The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}.
+%The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}.
 The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates.
@@
 
-For the first of these, we used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model to calculate monthly evapotranspiration, using the reference and cropland maps to adjust landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
-In the second example, we examined how these map errors impact the land allocation process of an agent-based food security model \cite{chen_dependency_2013}.
+%The reference dataset covered all of South Africa's field crop estate, from which we filtered out areas of communally owned farmlands and permanent tree crops to have a common basis of comparison across all landcover products, leaving us with map, after conversion to cropland percentage, that covered or 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.
 
-The reference dataset covered all of South Africa's field crop estate, from which we filtered out areas of communally owned farmlands and permanent tree crops to have a common basis of comparison across all landcover products, leaving us with map, after conversion to cropland percentage, that covered or 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.
+%The difference between total carbon stocks for the country made using any of the cropland maps were within +/-3\% of those based on the reference map, regardless of which cover type was adjacent to cropland (Table S1), as the large uncropped area of South Africa (Fig.
+1) dilutes the errors within the
+%(which is constrained by an absolute distance and the proximity of other agents)
 
-from agricultural pixels ($>$0.05\% cropland)
-
-The difference between total carbon stocks for the country made using any of the cropland maps were within +/-3\% of those based on the reference map, regardless of which cover type was adjacent to cropland (Table S1), as the large uncropped area of South Africa (Fig.
-1) dilutes the errors within the
-
-The disaggregated yield and harvested area maps of \cite{monfreda_farming_2008} are built upon cropland fraction maps where the total area is adjusted to match survey-derived cropland area statistics reported for administrative districts \cite[provinces, in South Africa's case][]{ramankutty_farming_2008}.
-To be consistent with this methodology, we first adjusted our cropland maps according to this procedure, using the reference map to calculate total cropland area for each of South Africa's nine provinces, then updating the pixel-wise cropland percentages in the four cropland maps so that the province-wise sums matched the reference areas \cite[][and see SI]{ramankutty_farming_2008}.
-Despite this statistical constraint, the updated cropland maps still had substantial errors that were similar in pattern (Fig.
-S5) to those in the unadjusted maps (Fig.
-1), and we evaluated how these residuals affected gridded estimates of the yield and production of maize, South Africa's largest crop \cite{estes_comparing_2013}.
-To create these maps, we followed \cite{monfreda_farming_2008} by disaggregating district-level (n = 354, mean area = 3,445 km$^2$) agricultural census data \cite{statistics_south_africa_commercial_2007} for maize (South Africa's largest crop by area, \cite{estes_comparing_2013}) yield and harvested area, aggregated each set of maps, and multiplied the two to calculate production at each scale.
-
-(which is constrained by an absolute distance and the proximity of other agents)
-
-\section{Digital RCD Analysis} 
+%\section{Digital RCD Analysis} 
 \end{materials}
 
-\appendix[App 1]
+%\appendix[App 1]
 
-\appendix
-This is an example of an appendix without a title.
+%\appendix
+%This is an example of an appendix without a title.
 
 \begin{acknowledgments}
-I thank everyone tearfully. 
+This work was supported by funds from the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program
+(NNX15AC64G), and the National Science Foundation (SES-1360463 and BCS-1026776).
 \end{acknowledgments}

### 2015-11-25 paper/app_s1.tex
22/10/2015 revision. Revisions through results. See changelog.
Near-final version of main text, pending Tom's edits. Confusion matrix added to Appendix.

@@
 \RequirePackage{lineno}  % http://www-d0.fnal.gov/Run2Physics/WWW/templates/lineno.html
-\documentclass[12pt]{iopart}
-\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
+%\documentclass[12pt]{iopart}
+\documentclass[11pt, titlepage]{article}
+\usepackage[top=20mm, bottom=20mm, left=23mm, right=23mm]{geometry}
+%\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
 %Uncomment next line if AMS fonts required
@@
 \usepackage{placeins}
+\usepackage{multirow}
+\usepackage[numbers,round,sort,compress]{natbib}
+\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
 
 \begin{document}
-\bibliographystyle{unsrt}
+%\bibliographystyle{unsrt}
 
@@
 %\end{table}  
+\section*{Reference data accuracy}
+
+The accuracy of the reference vector dataset was assessed by Estes et al \cite{estes_platform_2015}.
+The assessment was undertaken using a sub-sample of 609 1 km$^2$ grid cells, which were selected using a weighted randomized sampling scheme.
+Weights were derived from a logistic regression model of cropland occurrence probability in South Africa, with 1 corresponding to the lowest quartile of probability, and 4 the highest.
+The accuracy assessment included all cropland classes mapped within the reference dataset, which ranged from communal/smallholder fields to commercial row crops to orchards and other types of horticulture.
+An visual accuracy assessment was conducted within each 1 km$^2$ grid cell, wherein each cell was divided into 25 smaller cells of 200 X 200 m (4 ha), and then the proportion of each cell filled with cropfields visible in underlying Google Maps imagery was then calculated to the nearest 5\% of coverage, using a finer 20-cell mesh overlaid on each sub-cell within which field presence/absence was recorded.
+The same procedure was then performed to assess coverage by reference map polygons, as well as the intersections and differences between cells determined to be occupied by visual assessment, and those by the reference polygons.
+These intersections and differences were used to calculate the area of true positives, false positives, true negatives, and false negatives within each grid cell, and from this obtain accuracy measures.
+In this case, accuracy was assessed using two landcover classes: cropland and non-cropland, thus a two class confusion matrix was constructed (Table 1) according to remote sensing classification accuracy guidelines set out by Oloffson et al.
+\cite{olofsson_good_2014}.
+In this case, the reference data were the visually interpreted crop field presences/absences, and the map data were the field boundary vectors being assessed (i.e.
+the data which provide the reference maps in this study), the class ``crop" refers to crop field presence and ``non-crop" to crop field absence, or all other landcover types.
+The total area of vectorized crop fields was used to calculate the proportion \emph{W} of South Africa's total area mapped as crop fields, and of the areas ``mapped'' (by omission) as having other cover types.
+These were then used to weight the proportions of each mapped class \emph{i} corresponding to reference class \emph{j} by the total area \emph{A} mapped for class \emph{i}.
+The total accuracy (0.97, or 97\%) was calculated by summing the diagonal (bold in Table 1), as well as the producer's accuracy \emph{P} for each reference class \emph{j}, and the user's accuracy \emph{U} for each mapped class \emph{i}.
+
+\begin{table}[ht]
+\centering
+\caption{Confusion matrix for the assessment of the hand-digitized crop field boundaries used to generate reference cropland cover percentage values.
+Here \emph{Reference} denotes a visual assessment of crop field presence in high resolution imagery within a sample of 609 sites, while \emph{Map} refers to the vectorized crop field boundaries.
+The assessment had two classes: crop fields and non-crop fields, and proportions corresponding to these two categories are the proportions of areas determined by class agreements and disagreement in the 609 sites (see text above), which are weighted by the proportion \emph{W} of each class' total mapped area \emph{A$_i$} in South Africa.
+\emph{P} and \emph{U} provide the Producer's and User's accuracies, respectively.
+}
+\begin{tabular}{rrrrrrr}
+  \hline
+  & & Reference (j) & & & & \\\cline{3-7}
+  & & Crop & Non-crop & \emph{W}$_i$ & \emph{A}$_i$ (ha) & \emph{U}$_i$ \\ 
+  \hline
+  Map (i) & Crop & \textbf{0.108} & 0.007 & 0.114 & 14018567 & 0.943 \\ 
+  & Non-crop & 0.021 & \textbf{0.865} & 0.886 & 108541733 & 0.977 \\ 
+  & Total & 0.128 & 0.872 & 1.000 & 122560300 &  \\ 
+  & \emph{P}$_j$ & 0.840 & 0.992 &  &  &  \\ 
+   \hline
+\end{tabular}
+\end{table}
+
 \section*{Landcover map bias}
@@
 \section*{Evapotranspiration bias}
-\input{figures/et-bias.tex}
+%\input{figures/et-bias.tex}
 
@@
 \FloatBarrier
-\section*{References}
-\bibliography{}
+%\section*{References}
+\bibliographystyle{pnas2011} 
+{\footnotesize \bibliography{/Users/lestes/Dropbox/publications/full.bib}}
 \end{document}

### 2015-11-25 paper/main.tex
Near-final version of main text, pending Tom's edits. Confusion matrix added to Appendix.

@@
 %\setlength{\textfloatsep}{5pt plus 1.0pt minus 2.0pt}
+\usepackage{multirow}
 \setlength{\parskip}{1cm plus4mm minus3mm}
+\usepackage{makecell}
+\usepackage{boldline}
+\setcellgapes{3pt}
+\usepackage[flushleft]{threeparttable}
 
@@
 \begin{document}
+%\hspace*{\fill}
 
@@
 
-\author{Lyndon Estes\affil{1}{Princeton University, Princeton, NJ USA},
-Peng Chen\affil{2}{Indiana University, Bloomington, IN USA},
-Stephanie Debats\affil{1}{Princeton University, Princeton, NJ USA},
-Tom Evans\affil{2}{Indiana University, Bloomington, IN USA},
-Fanie Ferreira\affil{3}{GeoTerraImage, Pretoria, RSA},
-Gabrielle Ragazzo\affil{1}{Princeton University, Princeton, NJ USA},
-Justin Sheffield\affil{1}{Princeton University, Princeton, NJ USA}
-Adam Wolf\affil{1}{Princeton University, Princeton, NJ USA}
+\author{Lyndon Estes{\affil{1}{Woodrow Wilson School, Princeton University, Princeton, NJ USA}\affil{2}{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}},
+Peng Chen\affil{3}{Indiana University, Bloomington, IN USA},
+Stephanie Debats\affil{2}{},
+Tom Evans\affil{3}{},
+Stefanus Ferreira\affil{4}{GeoTerraImage, Pretoria, RSA}{},
+Gabrielle Ragazzo\affil{5}{Ecology and Evolutionary Biology, Princeton University, Princeton, NJ USA},
+Justin Sheffield\affil{2}{},
+Adam Wolf\affil{5}{},
+Eric Wood\affil{2}{},
 \and
-Kelly Caylor\affil{1}{Princeton University, Princeton, NJ USA}}
+Kelly Caylor\affil{2}{}}
 
@@
 \begin{abstract}
-{Our knowledge of global change processes is built upon landcover data, but we lack the means to comprehensively assess landcover maps, particularly in rapidly developing regions, and thus do not fully comprehend how map quality impacts our understanding of global change.
-We used a unique, high quality, national-scale cropland map to assess the bias and accuracy in a) current generation landcover products, and b) analyses built on these.
-Maps have high bias and low accuracy, particularly at finer resolutions (1-10 km), which transfers into, and is often amplified, in downstream analyses of vegetative carbon stock, gridded crop production, and process-based food security models.
-Substantial map aggregation is needed to avoid potential misunderstandings of global change processes or misguided policy.
-Maps derived from coarse resolution sensors (such as MODIS and GlobCover) should be aggregated to at least 50-100 km resolution (and sometimes more).
-Map accuracy is lowest at intermediate levels of cover, thus coarser aggregation may be necessary in highly mixed landscapes.
-Higher resolution imaging systems (e.g.
-Landsat) produce maps that require less aggregation (1-25 km) to achieve high accuracy and low bias.
-For regional to global analyses, new fusion products, such as the GeoWiki cropland map, need just 5 km of aggregation to produce unbiased maps for most applications, but can require $\geq$25 km of aggregation to have sufficient accuracy.
-Widely used global datasets for carbon, crop area, and crop yield should be rebuilt on GeoWiki and similar maps, while new approaches methods that fuse human judgement, the latest computer vision-based algorithms, and growing high resolution image archives are needed to produce the next generation of landcover maps.
+{Research into global environmental change is often built upon landcover maps, which can have substantial errors.
+How these errors impact our understanding of global change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions.
+We used a high quality, national-scale cropland map to assess bias and accuracy in current generation landcover maps and four examples of ``downstream'' (landcover-dependent) analyses, two related to physical processes (estimates of vegetative carbon stocks and evapotranspiration) and two to socio-economic processes (gridded crop production estimates and agent-based simulation of household food security), at resolutions ranging from 1-100 km.
+We found that the landcover maps have substantial errors below 25 km resolution, with substantial biases and inaccuracies, particularly from maps derived from coarse resolution sensors.
+In some cases these error metrics remain high even when maps are aggregated to 100 km.
+These errors can be substantially amplified in downstream analyses (e.g.
+the carbon and crop production analyses), but in cases where the values of different cover types that underpin the analysis are similar (e.g.
+evapotranspiration estimates) the results can be relatively insensitive.
+To avoid these errors, and thus minimize the risk of misunderstanding global change processes or misinforming policy, substantial map aggregation of landcover maps from their base resolution is often needed.
+We provide several guidelines that can help landcover map users select appropriate datasets and  aggregation scales under different use cases.
+
+%To achieve low \emph{bias} in a map-wide statistic (e.g.
+average carbon density), then finer resolutions can be maintained, but if high \emph{accuracy} is desired (e.g.
+spatial comparisons of map values, or change detection), then coarser pixels are needed.
+%Map accuracy is lowest at intermediate levels of cover, thus if the mapped region has highly heterogenous cover (e.g.
+regions dominated by smallholder-farming), then greater aggregation is also necessary.
+From a dataset perspective, maps derived from coarse resolution sensors (e.g.
+MODIS) often need to be aggregated to at \textcolor{red}{50-100} km resolution for bias to be reduced to within +\-10\%, and \textcolor{red}{50-100+} km for mean absolute error (MAE), a measure of accuracy, to be <10\%.
+For comparable rates of bias and error, higher resolution imaging systems (e.g.
+Landsat) produce maps that require less aggregation (1-25 km).
+For regional to global analyses, new fusion products, such as the GeoWiki cropland map, which constrain maps so that the sum of their pixels match a total reported for a larger scale (e.g.
+government cropped area statistics for administrative districts), need just 5 km of aggregation to produce unbiased maps for most applications.
+However, this technique does less to improve accuracy, which may still require $\geq$25 km or more aggregation to achieve <10\% MAE.
+Widely used global datasets for carbon, crop area, and crop yield should be rebuilt on GeoWiki and similar maps, while new methods that fuse human judgement, the latest computer vision-based algorithms, and growing high resolution image archives can help produce the next generation of landcover maps.
 }
@@
 
-\abbreviations{GTI, GeoTerraImage; SSA, sub-Saharan Africa}
+%\abbreviations{GTI, GeoTerraImage; SSA, sub-Saharan Africa}
 \linenumbers
@@
 Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations.
-The reason for this is that landcover data can only practically be derived from satellite imagery, and in many regions the cover types of interest are smaller \cite[e.g.
-smallholder's farms][]{jain_mapping_2013} than the sensor resolution, or spectrally indistinct from neighboring covers, which translates into substantial mapping errors \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}.
-Landcover maps are therefore often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}.
-These errors mean that we are still unable to obtain the granular, mechanistic understanding of global change processes that we need.
+The reason for this is that landcover data can only practically be derived from satellite imaging, which has several important constraints--related to both physics and analytics--that propagate mapping errors.
+First, in many regions the cover types of interest are smaller than the sensor resolution \cite[e.g.
+smallholder's farms][]{jain_mapping_2013,debats_generalized_2015,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \cite{sweeney_mapping_2015}.
+Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties (e.g.
+agriculture \cite{debats_generalized_2015,estes_platform_2015}) and thus be difficult for the classifier to distinguish, while discretizing a continuous cover type (e.g.
+dividing a forest into different canopy cover classes) can also promote classification error, particularly near class boundaries \cite{foody_status_2002}, as well as confusion about the actual extent of the cover type \cite{sexton_conservation_2015}.
+Furthermore, class definitions often vary between maps, complicating inter-comparison \cite{kuemmerle_challenges_2013}.
+Third, landcover map series are often used to detect changes \cite[e.g.][]{gross_monitoring_2013}, or may use within-class seasonality as a means for discriminating between classes (e.g.
+phenology of cropping versus surrounding vegetation \cite{sweeney_mapping_2015}), but seasonal variability and landcover changes can be easily confused.
+Given these multiple sources of error, landcover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}.
+These errors limit our ability to obtain granular, mechanistic understanding of global change processes.
 
-These problems with landcover products are known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}.
+These problems with landcover products are well known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \cite[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}.
 What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes.
-Answering this question depends on having spatially comprehensive ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}.
-Our understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data.
-This allows us to quantify between-map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.
+Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}.
+Our current understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data.
+In terms of the former, there are well-developed, best-practice guidelines for selecting and using ground-truth samples to robustly quantify map error, but these are not commonly applied by the broader community of landcover map users (often due to lack of data), which makes it challenging to compare how error varies within and between maps \cite{stehman_global_2012, olofsson_making_2013,olofsson_good_2014,foody_status_2002}.
+The latter form of assessment are made at coarser administrative levels relative to reported statistics \cite[e.g.][]{fritz_comparison_2010}, which can also be suspect \cite{carletto_emperor_2013,fao_action_2013}.
 
-Being unable to fully quantify the errors in landcover maps makes it difficult, if not impossible, to gauge their impact on downstream analyses.
-There has been some work examining how such error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, carbon flux measurements \cite{quaife_impact_2008}, and human population estimates \cite{linard_assessing_2010}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008}.
+%Top-down accuracy assessments are also made from  "sanity checks" made in comparison to aggregated survey data. 
+%
+%
+%
+%This allows us to quantify between-map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.
+%
+%
+%
+%In practice, this means that we have an incomplete understanding of map error, and 
+%
+%
+%
+%
+%There are also well-defined guidelines for quantifying and accounting for map accuracy \cite{stehman_global_2012, olofsson_good_2014, olofsson_good_2014}.
+However, inadequate assessment of map error and their impacts on subsequent analyses remains quite common within the community of landcover map users \cite{olofsson_making_2013,olofsson_good_2014}.
+There is also an open question
+%
+%or from top-down "sanity checks" made in comparison to aggregated survey data.
+
+%This allows us to quantify between-map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.
+
+Since it is difficult to fully quantify landcover map errors, it is even more challenging to gauge their impact on downstream analyses, in which there is substantial risk of error amplification\cite{kuemmerle_challenges_2013}.
+There has been some work examining how map error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, carbon flux measurements \cite{quaife_impact_2008}, human population estimates \cite{linard_assessing_2010}, and species distribution models \cite{tuanmu_global_2014}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}.
 One exception is a Belgian study \cite{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover errors bias measurements of agricultural land use patterns, but the study extent was fairly small and the validation data were discontiguous.
@@
 Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g.
-3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \cite{fritz_geo-wiki:_2012,estes_projected_2013,debats_generalized_????}, are finally providing the means to comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \cite{searchinger_high_2015}.
+3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \cite{fritz_geo-wiki:_2012,estes_projected_2013, debats_generalized_2015}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \cite{searchinger_high_2015}.
 
-In this study, we take advantage of these recent advances to address the urgent need to more thoroughly and precisely quantify landcover map errors and how they might impact our understanding of change in the world's most dynamic regions.
-We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through ``downstream'' studies investigating into both the physical and socioeconomic components of global change.
-Our objective is to provide scientists and policy-makers who use landcover data with a better, more up-to-date, understanding of their appropriate uses and limitations.
+In this study, we take advantage of this recent growth in high resolution data to address the call to more thoroughly assess errors and in landcover maps \cite{kuemmerle_challenges_2013, olofsson_good_2014,olofsson_global_2012}, and further examine how these errors might impact our understanding of change in the world's most dynamic regions.
+We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through downstream studies investigating both the physical and socioeconomic components of global change.
+Our objective is to provide scientists and policy-makers who use landcover maps an up-to-date assessment of their appropriate uses and limitations.
 
@@
 The resulting vectorized field boundaries provide highly accurate data on field sizes and distribution for the period 2009-2011.
+Measured in terms of its ability to distinguish crop fields from other cover types, this dataset is 97\% percent accurate at 1 km resolution, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\%  \cite[][SI]{estes_platform_2015}.
+
 This dataset is particularly valuable because South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \cite{searchinger_high_2015}, yet is notably lacking trustworthy maps of existing agriculture land \cite{fritz_comparison_2010}.
@@
 The fourth dataset was the new 1 km GeoWiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}, which incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
-For comparison, we converted all five datasets into 1 km gridded maps of cropland density, expressed as the pixel-wise percent cover.
+Since GeoWiki provides a continuous bounded value of landcover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.
 
-In our first set of comparisons, we evaluated map quality by subtracting each of the four landcover product-derived percent cropland maps (hereafter we refer to these as the ``the test maps'') from the reference map, so that we could calculate each map's bias (the mean pixel-wise error) and accuracy (the mean absolute pixel-wise error, where a lower value indicates higher accuracy).
-We performed this analysis for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km, in order to evaluate how error and bias changes with scale.
+We first evaluated the quality of the landcover product-derived percent cropland maps (hereafter referred to as the ``test maps'').
+Instead of the standard confusion matrix-based accuracy assessment metrics \cite{olofsson_good_2014,olofsson_making_2013}, which apply to categorical landcover maps, we assessed the bias and accuracy of the test maps based on the gridded residuals that resulted when each test map was subtracted from the reference map.
+Here bias is the mean residual value, weighted by the density of reference cropland, and accuracy is the mean of the absolute values of residuals (also weighted by cropland density), thus accuracy is inversely related to this value.
+We calculated these metrics for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km, in order to evaluate how bias and accuracy changes with scale.
 We also assessed how landcover pattern impacts map performance by modeling the correlation between map accuracy and cropland density.
@@
 
-Our reference map indicated that crop fields covered 104,304 km$^2$, or nearly 10\%, of the total mapped area in the 2009-2011 time period.
-The test maps derived from SA-LC and GeoWiki overestimated this area by 31 and 10\%, respectively, while GlobCover and MODIS underestimated it by 18 and 23\%.
+Our reference map indicated that crop fields covered nearly 10\% of the total study area in the 2009-2011 time period.
 Subtracting each test map from the reference maps created pixel-wise residuals, where negative and positive values respectively represent overestimates and underestimates by the test map (Fig.
@@
 
-\begin{figure*}[t]
+%\end{wrapfigure}
+
+%\begin{figure*}[htp]
+%\centering
+%    \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}\quad
+%     \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}
+%\end{figure*}
+%\linenumbers
+The most pronounced errors were in the MODIS and GlobCover maps, which showed large positive residuals in the center of the country where cropland is most concentrated (blue areas in Fig.
+1A), and negative residuals (red areas) along the eastern and northern margins.
+These patterns translated into substantial map bias (Fig.
+1B), with GlobCover and MODIS mean error exceeding 45\% and 25\% respectively at 1 km resolution, meaning that each map tends to underestimate cropland by that amount at that resolution.
+This bias declined with each level of map aggregation, being reduced to nearly 15\% for GlobCover and 5\% for MODIS at 100 km.
+The magnitude of mean absolute error (MAE) was somewhat higher in all cases.
+The GeoWiki map, in contrast, was the least biased overall, showing just a ~7\% bias at 1 km and near 0 for all other scales of aggregation, although its accuracy (23\% MAE) was only half as good as SA-LC's at 1 km (11\% MAE), which despite its uniform overestimation bias (Fig.
+1A) was the most accurate map at aggregation scales $<10$km.
+Above this, GeoWiki became slightly more accurate, having $<$5\% MAE at 100 km resolution.
+The reason GeoWiki had relatively poor accuracy at 1 km resolution was due to the high heterogeneous error pattern, which traded between positive and negative residuals over short distances, thereby inflating MAE at this scale.
+
+\subsection{Error and landcover density}
+Given that landcover characteristics can influence landcover map quality \cite{see_improved_2015, estes_platform_2015, debats_generalized_2015}, we set out to elucidate the relationship between cropland configuration and map error.
+We first calculated the average 1 km reference cropland percentage within each of South Africa's 354 magisterial districts (the finest administrative unit, averaging 3,445 km$^2$; SI), which yielded a landscape-scaled metric of characteristic cropland density, and similarly calculated the district-wise MAE for each test map.
+
+\begin{figure*}[hb]
 %\begin{wrapfigure}{c}{1\textwidth}
@@
 \end{figure*}
-%\end{wrapfigure}
-
-%\begin{figure*}[htp]
-%\centering
-%    \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}\quad
-%     \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}
-%\end{figure*}
-
-\linenumbers
-The most pronounced errors were in the MODIS and GlobCover maps, which showed large positive residuals in the center of the country where cropland is most concentrated (blue areas in Fig.
-1A), and negative residuals (red areas) along the eastern and northern margins.
-These patterns translated into substantial map bias (Fig.
-1.B), with GlobCover and MODIS mean error (weighted by the reference cropland density) exceeding 45\% and 25\% respectively at 1 km resolution, meaning that each map tends to underestimate cropland by that amount at that resolution.
-This bias declined with each level of map aggregation, being reduced to nearly 15\% for GlobCover and 5\% for MODIS at 100 km.
-The magnitude of (density weighted) mean absolute error (MAE) was somewhat higher in all cases.
-The GeoWiki map, in contrast, was the least biased overall, showing just a ~7\% bias at 1 km and near 0 for all other scales of aggregation, although its accuracy (23\% MAE) was only half as good as SA-LC's at 1 km (11\% MAE), which despite its uniform overestimation bias (Fig.
-1A) was the most accurate map at aggregation scales $<10$km.
-Above this, GeoWiki became slightly more accurate, having $<$5\% MAE at 100 km resolution.
-The reason GeoWiki had relatively poor accuracy at 1 km resolution was due to the heterogeneity of residuals, which traded between positive and negative residuals over short distances, thereby inflating MAE at this scale.
-
-\subsection{Error and landcover density}
-Given that landcover characteristics can influence landcover map quality \cite{see_improved_2015, estes_diylandcover:_2015, debats_generalized_????}, we set out to elucidate the relationship between cropland configuration and map error.
-We first calculated the average 1 km reference cropland percentage within each of South Africa's 354 magisterial districts (the finest administrative unit, averaging 3,445 km$^2$; SI), which yielded a landscape-scaled metric of characteristic cropland density, and similarly calculated the district-wise MAE for each test map.
 
@@
 At the most extreme, GlobCover's bias was -276\% at 1 km, but even SA-LC and GeoWiki had biases of 22\% and -50\%, respectively.
-Bias could be substantial even for the least carbon dense vegetation type (sparse), as evidenced by the 15-25\% bias at 1 km for MODIS and GlobCover under this class.
+Bias could be substantial even for the least carbon dense vegetation type (sparse), as evidenced by the 15-25\% mean error at 1 km for MODIS and GlobCover under this class.
 The mean bias across the different potential adjacent vegetation classes ranged between -20 for GeoWiki and -123\% for GlobCover at 1 km (with MODIS in between these), while SA-LC's average bias was 11\%.
-Biases declined fairly rapidly with aggregation, with all datasets having an average (across cover types) bias magnitude of less than 10\% by 25 km of aggregation, except for GlobCover, which was -12\% at 100 km (SI).
-As with cropland percentages, GeoWiki produced the least biased carbon density estimates above above 1 km resolution.
+Biases declined fairly rapidly with aggregation, with all datasets having an average (across cover types) bias magnitude of $<$10\% at $\geq$25 km of aggregation, except for GlobCover, which was -12\% at 100 km (SI).
+As with cropland percentages, GeoWiki produced the least biased carbon density estimates above 1 km resolution.
 
@@
 The average MAE across vegetation classes was 47\% at 1 km, dropping to $<$10 only with 25 km of aggregation.
-In contrast, SA-LC's carbon estimates were twice as accurate at 1 km, and was slightly more accurate up to 25 km of aggregation were GeoWiki achieved parity.
+In contrast, SA-LC's carbon estimates were twice as accurate at 1 km, and were slightly more accurate up to 25 km of aggregation were GeoWiki achieved parity.
 
@@
 We then used these two layers to calculate maize production, and further aggregated the yield and production grids to 5, 10, 25, 50, and 100 km resolutions before quantifying the bias and accuracy of each test map's yield and production values.
-In this case, we could not convert cell-wise errors into percentages of the reference map values (because many cells had zero values for one map but not the other), so calculated bias and accuracy from the map residuals and then normalized their values to the reference map means.
+In this case, we could not convert cell-wise errors into percentages of the reference map values (because many cells had zero values for one map but not the other), so we calculated bias and accuracy from the map residuals and then normalized their values to the reference map means.
 
@@
 5).
-These dropped to below 10\% with aggregation.
+These dropped to $<$10\% with aggregation.
 
-Production biases were generally slightly higher, but still low, for most datasets, with the exception of GlobCover, which had an gigantic underestimation bias of $>$60\% (relative to mean production) at 1 km, which remained above 10\% even at 100 km of aggregation.
+Production biases were generally slightly higher, but still low, for most datasets, with the exception of GlobCover, which had a large underestimation bias of $>$60\% (relative to mean production) at 1 km, which remained above 10\% even at 100 km of aggregation.
 MODIS production bias was above 20\% at 1 km, but declined to below 10\% at higher levels of aggregation.
 
-The accuracy of production estimates was another story.
-Here all datasets but SA-LC had MAE values of 30\% or higher below 25 km of aggregation (Fig.
-5), reaching as high at 100, 65, and 45\% at 1 km for GlobCover, MODIS, and GeoWiki, respectively.
-Only GeoWiki's MAE fell below 10\% with 100 km of aggregation.
+In contrast, the accuracy of production estimates was poor.
+Here all datasets but SA-LC had MAE values of $\geq$30\% below 25 km of aggregation (Fig.
+5), reaching as high as 100\% for GlobCover at 1 km, followed by 65\% for MODIS and 45\% for GeoWiki.
 SA-LC estimated production was most accurate, having between 10-20\% MAE between 1 and 10 km, and $<$10\% at 25 km and higher.
 This low accuracy relative to the gridded yield measures relates to the disaggregation process for harvested area, which allocates a fractional value to each pixel, which is itself a fraction.
-The process of adjusting the gridded values so that their total match the statistics from which they are derived does not adjust map errors relative to the reference map, and the constraint in fact appears to shorten the distance between negative and positive residuals (SI), thereby increasing absolute errors.
+The process of adjusting the gridded values so that their totals match reported statistics does relatively little to correct the map's underlying commission or omission errors, and this constraint in fact appears to shorten the distance between negative and positive residuals (SI), thereby increasing absolute errors.
 
@@
 To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistic represents the social and biophysical features of the study region \cite{berger_creating_2006}.
-In our example, we used an ABM of household food security that simulates the food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}.
+In our example, we used an ABM of household food security that simulates food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}.
 The model is initialized such that each household is allocated its designated share of cropland, and then produces a crop over the course of one to many seasons.
@@
 districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers.
-To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig.
-S8) in the eastern part of the country with similar climate and 28-45\% of their areas devoted to cropland.
+To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, SI) in the eastern part of the country with similar climate and 28-45\% of their areas devoted to cropland.
 The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity.
 The number of households and the cropland area per household was derived from surveys of farming communities where all cropland is used.
-The model was thus considered to well calibrated when all households were allocated their appropriate area of cropland, and all cropland was allocated.
+The model was thus considered to be well-calibrated when all households were allocated their appropriate area of cropland, and all cropland was allocated to households.
 
@@
 The MODIS-based simulation for districts 1 and 2 was most pronounced for this tendency, where 5-10\% of cropland remaining unallocated despite the fact that cropland with such a high cropland underestimate (85\%) the majority of households were not allocated cropland.
-This curious relationship occurred because croplands tend to cluster, and when underestimated clusters tend to be small, resulting in islands of cropland that fall outside of the search radius within which cropland is sought when agents are seeded onto the landscape.
+This non-linear relationship occurred because croplands tend to cluster, and when underestimated clusters tend to be small, resulting in islands of cropland that fall outside of the search radius within which cropland is sought when agents are seeded onto the landscape.
 
@@
 \section{Discussion}
-This spatially comprehensive, bottom-up assessment of landcover map bias and accuracy provides new insight into their extent, causes, and consequences for understanding global change, made possible by a unique, high accuracy dataset that likely provides the truest measure of cropland area and distribution currently available for this region.
-This dataset is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission), while some of the error we identified may be due to slight temporal mismatches between the reference and test datasets.
-However, our assessment (SI) suggests that these errors are small, and do not appreciably impact our findings, which is bolstered by previous work showing the substantial errors and inconsistencies between landcover maps \cite{fritz_comparison_2010,gross_monitoring_2013}.
-These previous studies, which were conducted over broader extents than ours, also indicate that our findings should be generalizable beyond South Africa's borders, particularly to more rapidly developing, data poor regions.
+This spatially comprehensive, bottom-up landcover map assessment provides new insight into the extent and causes of map error, and its consequences for understanding global change.
+The analysis was made possible by a spatially comprehensive, high accuracy dataset that likely provides the truest measure of cropland area and distribution currently available for this region.
+Although this reference map is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission, SI), while temporal mismatches between the reference and test maps may account for some of the error we identified, our assessment (SI) suggests that such discrepancies do not appreciably impact our findings.
+Our results are also supported by previous, continental-scale assessments that revealed substantial errors and inconsistencies between landcover maps \cite{fritz_comparison_2010,gross_monitoring_2013}.
+These similarities, combined with the larger extents of these studies, also suggest that our findings are applicable beyond South Africa's borders.
 
 \subsection{General guidelines for selecting and using landcover data}
-Our results suggest several guidelines for selecting and using landcover data in global change research.
-First, landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions.
+Our findings regarding bias and accuracy suggest several guidelines for selecting and using landcover data in global change research (Table 1), and suggested aggregation scales.
+These guidelines vary according to the intended application, what error property the users wants to minimize or maximize, and what tradeoff the user is willing to make between error and resolution.
+For the purpose of clarity in these guidelines, we assume that a maximum error magnitude of 10\% is acceptable, but the results of our analyses (Fig.
+1, 3, 5) allow different thresholds to be  evaluated.
+
+The first general guideline is that landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions.
 Researchers might need to use these datasets for a variety of reasons, for instance, because they provide consistent, multiple landcover classes at a global scale, or allow changes to be tracked across large scales over time \cite{luoto_predicting_2004}.
 The appropriate scale of aggregation then depends on which error property the user wants to minimize.
-For instance, if the objective is to minimize systematic errors in order to calculate an unbiased mean (e.g.
-average carbon density), then at least 50-100 km of aggregation is needed to reduce map bias to within +/-10\% (Fig.
-1B, 3).
-If the user wants to maximize accuracy (accuracy is important for comparing differences between locations on a single map, or between dates at fixed locations), then even coarser aggregations are needed to reduce mean absolute error below 10\% (Fig.
+If the objective is to minimize systematic errors in order to calculate an unbiased mean (e.g.
+average carbon density), then at least 25-50 km of aggregation is needed to reduce map bias to within +/-10\% (Table 1, Fig.
+1B \& 3).
+If the user wants to maximize accuracy, which is important for comparing differences between locations on a single map, or between dates at fixed locations, then even coarser aggregations are needed to reduce mean absolute error below 10\% (Fig.
 1, 3, 5).
@@
 2), a result echoed in an Africa-wide change detection study made with MODIS \cite{gross_monitoring_2013}.
-Outside of South Africa, where smallholder-driven agriculture predominates \cite{lambin_estimating_????} such mixed landscapes are more common, thus greater levels of aggregation may be needed over these areas.
+Outside of South Africa, where smallholder-driven agriculture predominates \cite{lambin_estimating_????}, such mixed landscapes are more common, thus greater levels of aggregation may be needed over these areas.
 Coarser aggregations are also necessary if the cover type of interest belongs primarily to a mixed class (as with GlobCover).
@@
 Unfortunately, Landsat-scaled maps are typically developed for specific countries, using varying methods, and can be hard to obtain (an exception to this is the newly released GLC30, which has a reported classification accuracy of 80\% \cite{chen_global_2015}).
-Our results also suggest that, despite their higher accuracy, they are not necessarily the least biased maps.
-Except for a few results (notably the 1 km carbon and maize production estimates), GeoWiki-based maps had the least amount of bias across most scales.
-GeoWiki's lower bias results from its process of calibrating the cropland percentages to reported agricultural statistics \cite{fritz_cropland_2011,fritz_mapping_2015}.
-The procedure is similar to the one we used in disaggregating maize harvested areas \cite{ramankutty_farming_2008, monfreda_farming_2008}, which led to fairly unbiased estimates of maize production and yield (Fig.
+Our results also suggest that, despite their higher accuracy, they may not necessarily the least biased maps.
+Except for a few results (notably the 1 km carbon and maize production estimates), GeoWiki-based maps had the lowest bias across most scales, which is the largely the result of calibrating the cropland percentages to match reported agricultural statistics \cite{fritz_cropland_2011,fritz_mapping_2015}.
+This procedure is similar to the one we used in disaggregating maize harvested areas \cite{ramankutty_farming_2008, monfreda_farming_2008}, which led to fairly unbiased estimates of maize production and yield (Fig.
 5) above 1 km (except for GlobCover's production estimates).
@@
 The reason that accuracy is not improved is because the statistical constraint cannot correct the omission and commission errors within the landcover maps.
+To minimize these types of errors, different techniques are needed, such as creating a single consensus map from multiple landcover products \cite[e.g.][]{fritz_comparison_2010,tuanmu_global_2014}.
+This second kind of data fusion was used in the first stage of creating the GeoWiki map (before statistical adjustment), and it accounts for GeoWiki's higher accuracy scores relative to MODIS and GlobCover (Fig.
+1B, 3, \& 5).
+It may be particularly effective in minimizing underestimation errors that result when a mixed class type becomes dominant (as with GlobCover), or when dealing with substantial sub-pixel heterogeneity \cite{fritz_mapping_2015,tuanmu_global_2014}
 
-These last observations suggest two further guidelines: 1) for high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \cite{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share.
+These last few points suggest two final guidelines.
+For high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \cite{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share.
 www.glcn.org} datasets for other cover types).
-But in either case some aggregation--5-25 km, depending on application and whether low bias or high accuracy is the objective--will still likely be necessary to achieve sufficiently low error.
-Additionally, these higher accuracy, lower bias product are made infrequently, thus applications requiring high frequency landcover data will depend on MODIS and confreres, as previously mentioned, and therefore much coarser aggregations.
+But in either case some aggregation--1-25 km, depending on application and whether low bias or high accuracy is the objective--will still likely be necessary to achieve sufficiently low error.
+Additionally, these higher accuracy, lower bias products often represent just a single time period \cite[e.g.][]{fritz_mapping_2015,tuanmu_global_2014}, thus applications requiring higher frequency (e.g.
+annual), large area landcover maps will depend on MODIS or similar sensors, and therefore much coarser aggregations.
 
@@
 
+\begin{table*}[!h]
+  \begin{threeparttable}
+    \caption{Guidelines for selecting and using different classes of landcover product (based on the characteristics of the satellite sensor(s) and map-making methods), according to the general purpose of the application, the error minimization objective, and the range of aggregation needed to achieve the error objective (here we use a maximum acceptable error magnitude of 10\%).
+The preferred landcover product class for each application is indicated in boldface.
+}
+    \begin{center}
+      \begin{tabular}{p{4.5cm}lp{6cm}l}
+       \hline
+Application                     & Objective & Source of Landcover Map & Minimum aggregation* \\
+\Xhline{1pt}
+High frequency change           & Maximize accuracy   & Coarse resolution sensor (e.g. MODIS) & $>$100 km\\
+detection                                                               &                                  & \textbf{High resolution sensor} (e.g. Landsat)   &  1-25 km\\
+\hline
+Mean values         & Minimize bias            & Coarse resolution sensor  & 25-50 km \\
+(e.g.
+carbon density)                                                              &                                   &  \multicolumn{1}{r}{\emph{if mixed class dominates}}                                    & $>$100 km \\
+                                                              &                                   & \textbf{Fusion product: statistical constraint} & 1-5 km\\
+                                                              &                                   & High resolution sensor & 1-10 km\\
+\hline
+Spatial comparisons & Maximize accuracy  & Coarse resolution sensor & 25-100 km\\
+ (e.g.
+crop production)                                                               &                                   &  \multicolumn{1}{r}{\emph{if mixed class dominates}}                                    & $>$100 km \\
+                                                             &                                   & Fusion product: map consensus** & 10-50 km\\
+                                                              &                                   & \textbf{High resolution sensor} & 1-25 km\\
+\Xhline{1pt}                         
+       \end{tabular}
+       \begin{tablenotes}
+        \small
+        \item *Aggregation ranges account for variation in the results from the different map analyses.
+        \item **This assumes that consensus-based fusion products \cite[e.g.][]{fritz_cropland_2011,fritz_mapping_2015,tuanmu_global_2014} incorporate maps from coarse resolution sensors, making them less accurate at finer scales than products created using high resolution sensors.
+       \end{tablenotes}
+     \end{center}
+     \label{default}
+  \end{threeparttable}
+\end{table*}%
+
 \subsection{Implications for global change analyses}
 Besides the preceding landcover data selection and usage guidelines, our findings also have broader implications for global change analyses and associated policy decisions.
-The simpler, most direct uses of landcover data have substantial to skew our understanding of global change processes.
+The simpler, most direct uses of landcover data have substantial potential to skew our understanding of global change processes.
 For example, the global carbon carbon density map of \cite{ruesch_new_2008}, of which we created facsimiles here, has been widely used, but our results suggest that any analyses of the spatial variability of carbon stocks, or of mean carbon densities, can be highly misleading if the map is not first sufficiently aggregated.
@@
 3).
-The same carbon map provided country-specific mean forest carbon densities for an evaluation of climate mitigation policies \cite{cattaneo_international_2010}, which our results suggest could be underestimated by 70\% (Fig.
+The same carbon map provided country-specific mean forest carbon densities for an evaluation of climate mitigation policies \cite{cattaneo_international_2010}, which our results suggest could be substantially underestimated (Fig.
 3).
-In contrast to these, an assessment of the congruence between carbon and biodiversity \cite{strassburg_global_2010} aggregated the carbon data to $\sim$110 km resolution, where accuracy is much higher.
+In contrast to these, an assessment of the congruence between carbon and biodiversity \cite{strassburg_global_2010} aggregated the carbon densities to $\sim$110 km, where accuracy is much higher.
 
-A related example can be seen in studies that create high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \cite[e.g.
+A related example can be seen in high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \cite[e.g.
 Figure 3 in][]{foley_solutions_2011}.
 Such maps give an inflated sense of the precision with which such locations can be identified.
-Although the study in this example is careful to report aggregated statistics, in which we can have greater confidence, it is not inconceivable that the maps themselves are used to inform policy.
-For this reason, it is advisable to degrade maps until their accuracy is of an acceptable standard (i.e.
-$\geq$100 km).
+Although the study in this example is careful to report aggregated statistics, and the map-makers stress that their products should not be used at the grid cell scale\footnote{www.earthstat.org\/wp-content\/uploads\/METADATA\_HarvestedAreaYield175Crops.pdf}, it is not inconceivable that policy-makers might make decisions based on their visual interpretations of maps.
+For this reason, it is advisable to aggregate presented maps until their accuracy reaches an acceptable standard (Table 1).
 
-Aggregating until acceptable accuracy is achieved will be particularly important in growing efforts to find tradeoffs between agricultural development and ecological protection \cite[e.g.][]{searchinger_high_2015,west_trading_2010}.
-Finding win-wins fundamentally requires accurate, fine-scale data in order to identify ``win-win'' solutions, which are comprised of areas where there is some deviation from an expected relationship; for example, an area of unusually low carbon density but high crop fertility \cite{searchinger_high_2015}.
-Without sufficient aggregation, it is difficult to know whether such efficiencies are due to real biophysical properties, or simply landcover error.
-
-What then, of the impacts of landcover data on more complex analysis?
+What then, of the impacts of landcover map error on more complex analysis?
 Here the results appear to be mixed.
@@
 4) are important, because they occurred at the climatic margins of the major crop growing regions, where a substantial share of farms use irrigation.
-Since VIC does not simulate irrigation, the magnitude of these errors is likely to have been underestimated, because irrigation substantially increases latent heat flux \cite{sacks_effects_2008}.
+Since VIC does not simulate irrigation, the magnitude of these errors is likely to have been underestimated, because irrigation substantially increases latent heat flux and can substantially modify regional climate\cite{sacks_effects_2008,mueller_cooling_2015}.
 This also suggests the potential for map error to magnify in a coupled land-atmosphere model (which could explain the difference between this result and that of Ge et al.
@@
 Another example is the growing efforts to minimize the environmental and climate impacts on new agricultural expansion frontiers, particularly Africa's savannas, which requires identifying areas that provide the highest agricultural benefit for the lowest environmental cost \cite{searchinger_high_2015}.
-In order to be effective, such analyses must be undertaken at the fine scales--hectares, rather than square kilometers--at which land use decisions are made, thus their value for informing land use policy is contingent on the accuracy of their underlying landcover/land use data \cite{searchinger_high_2015}.
+This sort of analysis entails finding ``win-win'' land use solutions, which are essentially areas that deviate from an expected development cost-benefit relationship; for example, regions of unusually low carbon density but high crop fertility \cite{searchinger_high_2015}.
+The validity of such analysis are entirely dependent In order to be effective, such analyses must be undertaken at the fine scales--hectares, rather than square kilometers--at which land use decisions are made, thus their value for informing land use policy is contingent on the accuracy of their underlying landcover/land use data \cite{searchinger_high_2015}.
+The validity of such analyses, which must be undertaken at the fine-scales where land use decisions are made (hectares rather than square kilometers), depends on having high quality map data, in order to have confidence that the identified tradeoffs are based on real biophysical properties, not simply map error.
 
@@
 Our analysis demonstrates that the landcover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on how accuracy or low bias.
-We provide some basic guidelines for selecting and using landcover data that can help increase confidence in the conclusions drawn from landcover-based studies.
+We provide some basic guidelines (Table 1) for selecting and using landcover data that can help increase confidence in the conclusions drawn from landcover-based studies.
 
@@
 For example, rebuilding the widely used cropland distribution and yield maps \cite{monfreda_farming_2008, ramankutty_farming_2008} using the GeoWiki map could substantially improve our knowledge about current global agricultural distribution and productivity.
-Similarly, reconstructing the carbon map of \cite{ruesch_new_2008} using the GLC-Share datasets (in which GeoWiki provides the cropland layer) would improve our confidence in regional to global assessments of vegetative carbon stocks.
+Similarly, reconstructing the IPCC's Tier 1 vegetative carbon map \cite{ruesch_new_2008} using a combination of the GeoWiki and new global landcover consensus maps \cite[e.g.
+GLC-Share or][]{tuanmu_global_2014} the GLC-Share datasets would improve our confidence in regional to global assessments of vegetative carbon stocks.
 Alternatively, the new high-resolution, Landsat-based GLC30 \cite{chen_global_2015}, aggregated and converted to fractional cover classes at 1 km--and fused with high quality national landcover maps, where available \cite[similar to][]{fritz_mapping_2015,fritz_cropland_2011}--may provide a preferable base map for rebuilding these datasets.
@@
 Fusing human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power appears to offer the greatest promise for improving the quality of landcover maps.
-The GeoWiki project and the new high resolution global forest change map \cite{hansen_high-resolution_2013} are the most advanced examples of these newer methods.
-Another approach is that of the Mapping Africa project\footnote{http://mappingafrica.princeton.edu}, which, inspired by quality of the reference map used in this study, enlists internet-based workers to manually digitize cropland boundaries \cite{estes_diylandcover:_2015}.
-Merging this crowdsourced approach to vector map creation with newer computer vision-based mapping algorithms \cite{debats_generalized_????} is a near-term goal of this project, as the former can interactively provide high quality training and test data (by using humans' superior capacity to delineate discrete cover types in noisy images \cite{estes_diylandcover:_2015}) for the latter, which has the advantage of speed and the ability to handle high-dimensional datasets.
-Approaches such as these will be needed to provide the next generation of landcover maps, so that we can gain a clearer, unbiased understanding of global change.
+The GeoWiki project and new high resolution global forest change maps \cite{hansen_high-resolution_2013,sexton_conservation_2015} are prominent examples of these newer methods.
+Another approach is that of the Mapping Africa project\footnote{http://mappingafrica.princeton.edu}, which, inspired by quality of the reference map used in this study, enlists internet-based workers to manually digitize cropland boundaries \cite{estes_platform_2015}.
+Merging this crowdsourced approach for creating vector maps of discrete cover types with newer computer vision-based mapping algorithms \cite{debats_generalized_2015} is a near-term goal of this project, as the former can interactively provide high quality training and test data (by using humans' superior capacity to delineate discrete cover types in noisy images \cite{estes_platform_2015}) for the latter, which has the advantage of speed and the ability to handle high-dimensional datasets.
+Approaches such as these will be needed to create the next generation of landcover maps, so that we can gain a clearer, unbiased understanding of global change.
 
@@
 \section{Methods} 
-To establish the accuracy of the cropland boundary reference map, we randomly selected 609 cell from a 1 km$^2$ grid covering South Africa, weighted by the probability of cropland occurrence \cite{estes_diylandcover:_2015}.
-We sub-divided each cells into 200 m sub-cells, and visually interpreted whether crop fields presence/absence within these against Google Earth imagery, and compared these results to the digitized field polygons to determine the reference map's sensitivity (0.89) and specificity (0.99), overall map accuracy (97\%), and its true skill statistic \cite[0.88;][]{allouche_assessing_2006}.
-We intersected the field vectors with a 1 km grid, and calculated the percent of each cell occupied by fields to create the gridded cropland reference map.
+To develop the cropland reference map (see SI for accuracy assessment details), we intersected the original field boundary vector maps with a 1 km grid, and calculated the percent of each cell occupied by fields to create the gridded cropland reference map.
 In the resulting grid, we masked out areas classified as communal farmland, because only their outer perimeters were digitized \cite{fourie_better_2009}, which posed a risk of overestimating cropland extent.
-We also excluded the permanent tree crop class, to avoid confusion with landcover maps that include these in their cropland classes.
+We also excluded the permanent tree crop class, to avoid confusion with landcover maps that exclude these in their cropland classes.
 For the same reason, we used other high resolution landcover maps for selected provinces of South Africa to mask out commercially afforested areas and sugarcane plantations, which were also not included in the reference map.
-The resulting masked reference grid covered 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ was cropland.
-We extracted the cropland classes from SA-LC, MODIS, and GlobCover, and converted these into percent cropland estimates at 1 km resolution (GeoWiki is already a cropland percentage map).
-Both MODIS and GlobCover had mixed pixel classes, thus we used the upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages.
+The resulting masked reference grid covered 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ was cropland.\\
+\indent We extracted the cropland classes from SA-LC, MODIS, and GlobCover, and converted these into percent cropland estimates at 1 km resolution (GeoWiki is already a cropland percentage map).
+Both MODIS and GlobCover had mixed pixel classes, thus, following the approach used in creating the GeoWiki maps \cite{fritz_mapping_2015} we used the upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages.
 We used the mean map for the main analysis, but estimated error variability using all three versions (SI).\\
-\indent To uncover the relationship between map error and cover density, we used magisterial district boundaries (n=354, mean area=3,445 km$^2$, see SI) to provide a landscape-scaled unit for calculating characteristic cover density.
+\indent To uncover the relationship between map error and cover density, we used magisterial district boundaries (n=354, mean area=3,445 km$^2$; SI) to provide a landscape-scaled unit for calculating characteristic cover density.
 We filtered out pixels with $<$0.05\% (0.5 ha) cropland, to prevent the much larger areas of non-agricultural districts from drowning out any relationship.
 We extracted the absolute values of test map errors and corresponding reference cropland percentages, calculated their district-wide means, and then modeled the relationship between mean absolute error (response) and cropland density (predictor) as a polynomial function within a generalized additive model.
-We accounted for potential spatial autocorrelation by fitting a two-dimensional smoothing spline to each district's centroid coordinates, and weighted the model by the number of agricultural pixels per district \cite{wood_mgcv:_2001}.\\
+We accounted for potential spatial autocorrelation by fitting a two-dimensional smoothing spline to the coordinates of each district's centroid, and weighted the model by the number of agricultural pixels per district \cite{wood_mgcv:_2001}.\\
 \indent To develop the carbon maps, we followed the methods used by Ruesch and Gibbs \cite{ruesch_new_2008} to create their vegetative carbon map.
@@
 We multiplied cropland densities by cropland fractions and added these to each of the five other densities multiplied by the residual non-cropland fractions to create five different carbon density maps.
-We aggregated each carbon map up to the five coarser resolutions for scaling comparisons.
+We aggregated each carbon map up to the five coarser resolutions for scaling comparisons (SI).
 \\
 \indent We used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model run with the Africa Flood and Drought Monitor's meteorological data \cite{sheffield_drought_2013} to produce monthly gridded evapotranspiration estimates for South Africa for the years 1979-2010.
-We adjusted VIC's native landcover scheme so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25${\circ}$ resolution).
+We adjusted VIC's native landcover scheme (derived from AVHRR) so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25$^{\circ}$ resolution).
 We ran one instance of VIC for each of the five landcover permutations, and compared the mean annual ET produced by the reference map variant with those from the test maps.
 \\
-\indent To create the gridded maize yield and production maps, we followed \cite{ramankutty_farming_2008} to first calibrate the test cropland maps against cropland areas reported by administrative-level agricultural censuses.
+\indent To create the gridded maize yield and production maps, we followed Ramankutty et al.
+\cite{ramankutty_farming_2008} to first calibrate the test cropland maps against cropland areas reported by administrative-level agricultural censuses.
 In place of census reports, we used the reference map to calculate total cropland areas for South Africa's nine provinces, then adjusted the pixel-wise cropland percentages in the four test maps so that their province-wise sums matched these totals (SI).
-We then followed Monfreda et al's procedure \cite{monfreda_farming_2008} to disaggregate maize \cite[South Africa's largest crop][]{estes_comparing_2013} planted area and yields onto the reference and adjusted test maps.
-Per \cite{monfreda_farming_2008}, the gridded planted area and yield estimates were obtained from magisterial district-level agricultural censuses, in this case for the most recent census of 2007 \cite{statistics_south_africa_commercial_2007}.
+We then followed Monfreda et al's procedure \cite{monfreda_farming_2008} for disaggregating maize \cite[South Africa's largest crop][]{estes_comparing_2013} planted area and yields onto the reference and adjusted test maps.
+The necessary statistics were obtained from magisterial district-level agricultural censuses conducted for the year 2007 \cite{statistics_south_africa_commercial_2007}.
 We then aggregated gridded harvested areas and yields (weighted by the corresponding harvested area) to each coarser resolution, and multiplied yield by harvested area to calculate production.\\
@@
 \bibliographystyle{pnas2011} 
-{\footnotesize \bibliography{bias}}
+{\footnotesize \bibliography{/Users/lestes/Dropbox/publications/full.bib}}
 

### 2015-12-01 Rmd/c1-landcover-preprocess-1.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

---
title: "Pre-processing cropland datasets"
author: "Lyndon Estes"
date: "29 September 2014"
output:
  html_document:
    theme: spacelab
    toc: yes
---

# Pre-processing reference cropland data layer
<!-- This analysis was undertaken to determine the bias in satellite-derived landcover products at different
spatial scales, using GeoTerraImage's cropland maps for South Africa to assess this error.-->
The following describes the initial pre-processing of landcover datasets that was
undertaken for an analysis of error in landcover data.

## Gridding the SA landcover data
Much of the original analysis was undertaken on the Mapping Africa server, using postgis/postgres as well as pprepair to do much of the cleaning and conversion of the GTI (reference) dataset into a 1 km resolution cropland percentage cover dataset. The following steps were taken:

### Cleaning up fields in GTI data
This was done for both the 2007 and 2011 datasets, using a script that pulled out the fields, found the bad ones, read them into an R data object, ran pprepair over it, and then stuck the cleaned geometries back into a postgis table holding the full dataset. The intersection was then performed.

```{r, eval = FALSE}
source("R_ext/grid-fields-intersect-postgis.R")  # run on mapper.princeton.edu
```

### Calculating field areas

The next step was running the following script to calculate the area for different cropcover classes in each 1 km^2^ pixel.  This created two core gridded datasets, cover2007.tif and cover2011.tif.

```{r, eval = FALSE}
source("R_ext/grid-fields-intersect-analyze.R")  # run on mapper.princeton.edu
```

### 2015-12-01 Rmd/c2-landcover-preprocess-2.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

--- 2015-06-26 Rmd/fetch-land-cover.Rmd
@@
 
-### Introduction
+# Pre-processing of landcover datasets
 Fetching and pre-processing of global landcover datasets to be analyzed relative to the SA dataset: the MODIS 2011 landcover product for South Africa was downloaded, the GLC-SHARE dataset, the 2009 GlobCover product, South Africa's 2009 landcover product, and geo-wiki's cropland fusion map.
@@
 
-### Analyses
-#### Set-up
+## Analyses
+### Set-up
 ```{r, eval = FALSE}
@@
 
-#### Bring in landcover datasets
+### Bring in landcover datasets
 Let's start with the improved/fusion datasets, GLC share and geo-wiki's
@@
 
-Create mask to remove sugarcane growing regions
+### Mask out sugarcance
 ```{r, eval = FALSE}
@@
 
-Create low, medium, high cover fraction variants from MODIS classes 12 and 14 (from the IGBP-DIS scheme).
+### MODIS cropland variants
+Different variants of cropland fractions from MODIS classes 12 and 14 (from the IGBP-DIS scheme).
 Class 14 is a cropland mosaic class, with minimum of 10% and a max of 60%, so create variants of 10%, 35% (mean) and 60%.
@@
 
+### GlobCover cropland variants
 And the same for GlobCover.
@@
 
-Bring in GTI cropland fractions to set up masks
+### Further masking of reference maps
+
 ```{r, eval = FALSE}

### 2015-12-01 Rmd/c3-cropland-1.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

--- 2015-10-17 Rmd/compare-landcover.Rmd
@@
 ---
-title: "Calculating cropland biases"
+title: "Calculating cropland bias and accuracy"
 author: "Lyndon Estes"
@@
     toc: yes
+    toc_depth: 4
     number_sections: true
@@
 
-# Overview
-This section of the analysis calculates the differences between the GTI cropland percentages and those from the various landcover products.
+# Cropland errors, part 1
+This section of the analysis calculates the differences between the reference cropland percentages and those from the various landcover products.
 It draws from results generated by the [cropland pre-processing](cropland-preprocess.html).
@@
 
-# Analyses
-## Load in data 
-### Set up paths
+## Analyses
 ```{r, eval = FALSE}
@@
 
-### Load cropland grids
-
+### Load cropland test grids 
 Derived from global landcover products (note: it turns out GLC-Share and geowiki are the same thing, so GLC-Share dropped)
@@
 
-### SA landcover data
+#### Reference data
 ```{r, eval = FALSE}
@@
 
-### Apply NA masks
+#### Apply NA masks
 ```{r, eval = FALSE}
@@
 
-## Calculate differences at different levels of aggregation
-This steps draws on functions I wrote to calculate the actual and absolute differences between a given "master" grid and lists of other grids.
+### Calculate differences at different levels of aggregation
+This step draws on functions I wrote to calculate the actual and absolute differences between a given "master" grid and lists of other grids.
 This comparison is done at multiple aggregations, ranging from 1-600 km resolution cropland averages.
@@
 
-## Bias as function of cropcover
-### Weighting rasters based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc) 
+### Error versus cropland density
+#### Weighting rasters 
+Based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc) 
 ```{r, eval = FALSE}
@@
 
-### Calculate biases, pooling MODIS and GlobCover across all version
+#### Calculate bias/MAE, pooling MODIS and GlobCover across all versions
 Step 1. Reshape lists, so that all MODIS and GlobCover variants are pooled by scale
@@
 
-### Then calculate biases across datasets
+#### Then calculate bias/MAE across datasets
 
@@
 
-## Differences between gti2007 and 2011
+### Differences between 2007 and 2011 reference
 ```{r, eval = FALSE}
@@
 
-## 2011 differences
+### 2011 reference-test bias/accuracy
 Whole country and agricultural areas only
@@
 
-## 2007 differences
+### 2007 reference-test bias/accuracy
 Whole country and agricultural areas only

### 2015-12-01 Rmd/c4-cropland-2.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

--- 2015-11-02 Rmd/cropland-bias.Rmd
@@
 ---
-title: "Cropland Bias and Accuracy"
+title: "Cropland percentage bias and accuracy, part II"
 author: "Lyndon Estes"
@@
     toc: yes
+    depth: 4
     number_sections: true
@@
 
-# Overview
+# Cropland errors, part 2
 
@@
 
-# Spatial biases
-## Datasets
+## Spatial biases
+### Datasets
 
@@
 
-## Plots
-Plot maps of biases at different resolutions, for publication output (and display here).
+## Error maps
+Plot maps of biases at different resolutions.
 To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and masked before plotting.
@@
 
-## Plot of selected rasters and levels 
+### Plot of selected rasters and levels 
 ```{r, eval = FALSE}
@@
 
-lev <- names(gti_agg)
+lev <- names(dlist_abs$sa30)
 ```
@@
 
-# Bias in relation to field density
+## Bias/MAE in relation to field density
 
-For supplementary data.
+### 1 km Bias/MAE with boxplots
 
-## Bias plots at 1 km with boxplots
-
+For supplemental.
 ```{r, eval = FALSE}
@@
 
-## Bias with scale
+### Bias/MAE with scale
 
@@
 
-# Total bias
+### Total error
 Calculate how much cropland estimates differ between datasets
@@
 
-# Cropland error and bias plots
+## Cropland bias/MAE plots
 
-## Primary method
+### Primary method
 
@@
 
+# Difference between 2007 and 2011 bias and accuracy statistics
+baccdiff <- bacc2011[, snms, with = FALSE] -  bacc2007[, snms, with = FALSE]
+setnames(baccdiff, colnames(baccdiff), mcap)
+
+# baccdiff <- round(rbind(baccdiff, baccdiff[, lapply(.SD, mean)]), 2)
+baccdiff <- cbind("Resolution" = rep(paste(fact, "km"), 2), baccdiff)
+xtable::xtable(baccdiff)
 ```
 
-## Other calculations methods
+### Other calculations methods
 
@@
 
-## Bin density plots
+### Bin density plots
 
@@
 ```{r, eval = FALSE}
-
+# Using the union of agricultural areas, rather than whole country
 bin_size <- do.call(rbind, lapply(lev, function(i) {
   # i <- "f50"
-  st <- mubias_2011[bin != "all" & ol == i & il == "sa30" & bvals == "mu.nm", 
+  st <- mubias_2011[bin != "all" & ol == i & il == "sa30" & bvals == "mu", 
                     .(bin, N)]

### 2015-12-01 Rmd/c5-accuracy-vs-density.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

--- 2015-10-01 Rmd/magisterial-district-bias.Rmd
@@
 ---
-title: "District-level cropcover and bias"
+title: "District-level cropland density versus map accuracy"
 author: "Lyndon Estes"
@@
 
-# Bias as a function of typical crop cover
+# Accuracy in relation to typical crop cover
 
 ## Data
-```{r, message=FALSE, results='hold'}
+```{r, message=FALSE, results='hold', warning=FALSE}
 library(raster)
@@
 
-# Analyses
+## Analyses
 
-## Extract data for magisterial districts
+### Extract data for magisterial districts
 
-Both cropland percentages and absolute biases. 
+Both cropland percentages and absolute errors. 
 ```{r, eval=FALSE}
@@
 
-## Run regressions and check for spatial autocorrelation
+### Run regressions and check for spatial autocorrelation
 ```{r, eval=FALSE}
@@
 
-### Use a GAM instead
+### Generalized additive model
 
@@
 # above, with smooth fit instead of quadratic
-dgam4b <- gam(log(dv + 0.0001) ~ s(pv, k = 3) + s(x, y), weights = w, data = dat,
-              method = "ML")
+dgam4b <- gam(log(dv + 0.0001) ~ s(pv, k = 3) + s(x, y), weights = w, 
+              data = dat, method = "ML")
 plot(dat$pv, dat$dv)
@@
 
-## Plot final results
+### Plot final results
 
@@
 
-### GAM fits
+#### GAM fits
 ```{r, eval = FALSE}
@@
 
-### GAM plots
+#### GAM plots
 ```{r, eval = FALSE}
@@
 
-### Final plot of magisterial districts, for supplementals
+#### Final plot of magisterial districts
+
+For supplemental.
 ```{r, eval=FALSE}

### 2015-12-01 Rmd/c6-carbon.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

--- 2015-11-02 Rmd/carbon-bias.Rmd
@@
 ---
-title: "Carbon bias"
+title: "Vegetative carbon bias and accuracy"
 author: "Lyndon Estes"
@@
 
-# Bias in carbon stock estimates
+# Error, bias, and accuracy in carbon stock estimates
 Based on the Ruesch & Gibbs (2008) approach for estimating carbon density.
 
-# Data
+## Data
 ```{r, eval=FALSE}
@@
 
-## Carbon values
+### Carbon values
 
@@
 
-# Analysis
-## Prepare raster data
+## Analysis
+### Prepare raster data
 ```{r, eval=FALSE}
@@
 
-## Create cropland carbon estimates
+### Create cropland carbon estimates
 ```{r, eval=FALSE}
@@
 
-## Difference the carbon datasets
+### Difference the carbon datasets
 ```{r, eval=FALSE}
@@
 
-# Outputs
-## Carbon bias maps
+## Outputs
+### Carbon error maps
 ```{r, eval=FALSE}
@@
 
-## Carbon bias statistics
-### Calculate how much country-lvel carbon estimates differ between datasets
+### Carbon bias/accuracy
+#### Calculate how much country-level carbon estimates differ between datasets
 ```{r, eval = FALSE}
@@
 
-### Calculate weighted bias statistics
+### Calculate bias/MAE statistics
 
@@
 
-### Extract statistics for secondary methods
+#### Extract statistics for secondary methods
 
@@
 
-### Plot mean carbon bias and accuracy against scale
+### Plot mean carbon bias/MAE against scale
 
@@
 # Bind all three tables together
-fulls <- rbind(cbind("Region" = "Density", muws), 
-               cbind("Region" = "Country", mu0s),
-               cbind("Region" = "Agricultural", mus))
+fulls <- muws 
+# fulls <- rbind(cbind("Region" = "Density", muws), 
+#                cbind("Region" = "Country", mu0s),
+#                cbind("Region" = "Agricultural", mus))
 
-caption <- paste("Biases and mean absolute errors for each of the cropland", 
+caption <- paste("Biases and mean absolute errors, weighted by reference", 
+                 "cropland density, for each of the test", 
                  "maps across aggregation scales and each possible landcover", 
-                 "type sharing the pixel with cropland. Three variants", 
-                 "of the bias/accuracy values are provided, indicated by the:",
-                 "Region column: 1) weighted by reference cropland density", 
-                 "(Density weighted); 2) across the entire country (Country);",
-                 "3) for agricultural areas only (Agricultural, i.e.", 
-                 "$>$0.05\\% cropland).")
-fulls_xtab <- xtable(fulls[order(Region, Metric, Cover)], digits = 1, 
+                 "type sharing the pixel with cropland.")
+# fulls_xtab <- xtable(fulls[order(Region, Metric, Cover)], digits = 1, 
+fulls_xtab <- xtable(fulls[order(Metric, Cover)], digits = 1, 
                      caption = caption)

### 2015-12-01 Rmd/c7-et.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

--- 2015-10-18 Rmd/et-bias.Rmd
@@
 
-# Overview
+# Error, bias, and accuracy in evapotranspiration estimates
 
@@
 
-# Analysis
-## Prepare datasets
+## Analysis
+### Prepare datasets
 
@@
 
-## Process ET time series
+### Process ET time series
 
@@
 
-## Calculate ET differences
+### Calculate ET differences
 ```{r, eval=FALSE}
@@
 
-## Plot difference maps
-### Reshape and disaggregate lists
+### Plot difference maps
+#### Reshape and disaggregate lists
 ```{r, eval = FALSE}
@@
 
-### Plot maps
+#### Plot maps
 ```{r, eval=FALSE}
@@
 
-## Calculate bias/accuracy statistics
+### Calculate bias/accuracy statistics
 
-### Primary method
+#### Primary method
 

### 2015-12-01 Rmd/c8-yield-production.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

--- 2015-10-18 Rmd/yield-bias.Rmd
@@
 author: "Lyndon Estes"
-date: "July 6, 2015"
 output: 
@@
 
-# Methods to examine
-
-Ramankutty et al's (2008) procedure for calculating cropland area.
-
-1. Divide provincial cropland area estimates by total area in each province. 
-2. Linear model calculates cropland proportions from model of all satellite landcover classes, creating a 5 minute gridded estimate of cropland in each pixel.
-3.
-These grids are aggregated to administrative level (provinces in our case), and then compared with the provincial statistics.
-A correction factor is then applied, and _pycnophylactically_ interpolated, which is then used to adjust the gridded estimates.
-
-$$ fcropland_{x,y} = ucf_{x,y} * cf_{x,y}$$
-
-Where ucf is the correction factor and cf is the step 1 modeled cropland fraction estimate for a given pixel (x, y).
-
-Monfreda et al (2008) then use $fcropland_{x,y}$ in a three step process to calculate the area in administrative units for each crop type, in the following steps:
-
-1. They divide $fcropland_{x,y}$ by the area of the administrative unit to calculate the fraction of each unit taken up by a particular crop.
-
-2. They then disaggregate this fraction back onto the gridded data as follows:
-
-$$ fcrop_{x,y} = fcropland_{x,y}\frac{croparea_{pu}}{cropland_{pu}}$$
-
-They assign yields for each crop--obtained from the finest administrative unit available, uniformly across pixels having a non-zero proportion of that crop within the administrative unit.
-
-They rescaled both area and yield within sub-administrative units to match country totals derived from FAOStat from 1997-2003.
-
-## Key Questions
-
-1.
-How do these processes create bias if they are not aggregated by the administrative units to which they are scaled?
-i.e.
-If you aggregate by grid cells rather than by administrative units, how does this affect estimates of total production (and therefore things such as calories to be gained by closing yield gaps)
-
-2. Do any studies use it in this particular way?
-
-    + Foley et al (2011) estimate amount of calories that could be gained by closing yield gaps.  
-    + Find one or two others - maybe Roy with yield trends, or Mueller et al 2012.
-
-# Data
+# Error, bias, and accuracy of gridded crop yield & production data
+
+Based on Ramankutty et al's (2008) procedure for calculating cropland area, followed by Monfreda et al's (2008) to disaggregate yields and harvested areas.
+
+## Data
 
@@
 
-# Analyses
-## Cropland fractions
-### Get provincial $cf$ factors from 1 km GTI dataset
+## Analyses
+### Cropland fractions
+#### Get provincial $cf$ factors from 1 km GTI dataset
 ```{r, eval=FALSE}
@@
 
-### Calculate cropland fractions from 1 km landcover sets
+#### Calculate cropland fractions from 1 km landcover sets
 
@@
 
-### Re-scale landcover cropland fraction using $cf$ factors
+#### Re-scale landcover cropland fraction using $cf$ factors
 ```{r, eval=FALSE}
@@
 
-### Aggregate adjusted rasters, compare cropland fractions
+#### Aggregate adjusted rasters, compare cropland fractions
 ```{r, eval=FALSE}
@@
 
-# snms <- names(p_diff1)
-# lcnu <- lapply(lev, function(x) {
-#   lcb <- lapply(snms, function(y) {
-#     all_gt0 <- !is.na(bins[[x]])
-#     all_gt0
-#   })
-#   named_out(lcb, snms)
-# })
-# names(lcnu) <- lev
 lev <- names(cf_pct_diff[[1]])
@@
 
-Plot for supplementary data
+#### Plot for supplementary data
 ```{r, eval=FALSE}
@@
 
-## Yield disaggregation
+### Yield disaggregation
 
@@
 
-### Process magisterial districts and census data
+#### Process magisterial districts and census data
 ```{r, eval=FALSE}
@@
 
-### Calculate individual crop fractions
+#### Calculate individual crop fractions
 
@@
 
-### Assign back crop fraction and yield to grid 
+#### Assign back crop fraction and yield to grid 
 ```{r, eval=FALSE}
@@
 
-### Aggregate yields and crop area to coarser resolutions
+#### Aggregate yields and crop area to coarser resolutions
 ```{r, eval=FALSE}
@@
 
-### Crop production estimates at different aggregation scales
+#### Crop production estimates at different aggregation scales
+
+Including differences relative to reference versions.
 ```{r, eval=FALSE}
@@
 
-## Bias statistics
+## Bias/MAE statistics
 
@@
 
-### Bias/accuracy plots
-
-#### Density-weighted bias and accuracy 
+### Bias/MAE plots
+
+#### Density-weighted
 ```{r, eval = FALSE}
@@
 
-#### Agriultural area bias and accuracy
+#### Agricultural area
 
@@
 
-#### Supplemental tables
+### Supplemental tables
 ```{r, eval = FALSE}
@@
 
-## The impacts of spatial patterns in bias
-
+### The impacts of spatial patterns in bias
+
+(Supplmentary analysis not presented).
 Using yield gap estimates as an indicator.

### 2015-12-01 Rmd/code.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

---
title: 'Code Appendix'
output:
  pdf_document:
    toc: yes
    toc_depth: 4
    highlight: tango
    number_sections: yes
---

# Pre-processing reference cropland data layer
<!-- This analysis was undertaken to determine the bias in satellite-derived landcover products at different
spatial scales, using GeoTerraImage's cropland maps for South Africa to assess this error.-->
The following describes the initial pre-processing of landcover datasets that was
undertaken for an analysis of error in landcover data.

## Gridding the SA landcover data
Much of the original analysis was undertaken on the Mapping Africa server, using postgis/postgres as well as pprepair to do much of the cleaning and conversion of the GTI (reference) dataset into a 1 km resolution cropland percentage cover dataset. The following steps were taken:

### Cleaning up fields in GTI data
This was done for both the 2007 and 2011 datasets, using a script that pulled out the fields, found the bad ones, read them into an R data object, ran pprepair over it, and then stuck the cleaned geometries back into a postgis table holding the full dataset. The intersection was then performed.

```{r, eval = FALSE}
source("R_ext/grid-fields-intersect-postgis.R")  # run on mapper.princeton.edu
```

### Calculating field areas

The next step was running the following script to calculate the area for different cropcover classes in each 1 km^2^ pixel.  This created two core gridded datasets, cover2007.tif and cover2011.tif.

```{r, eval = FALSE}
source("R_ext/grid-fields-intersect-analyze.R")  # run on mapper.princeton.edu
```

# Pre-processing of landcover datasets
Fetching and pre-processing of global landcover datasets to be analyzed relative to the SA dataset: the MODIS 2011 landcover product for South Africa was downloaded, the GLC-SHARE dataset, the 2009 GlobCover product, South Africa's 2009 landcover product, and geo-wiki's cropland fusion map. These were converted to 1 km$^2$ percentage cropland estimates after making several versions of MODIS and GlobCover products to account for mixed cropland classes.  A further processing was applied to the two GTI derived cover images.

## Analyses
### Set-up
```{r, eval = FALSE}
library(SAcropland)
library(gdalUtils)
library(RCurl)
library(RPostgreSQL)
library(rgdal)
library(rgeos)
library(raster)
library(lmisc)
```

Paths
```{r, eval = FALSE}
#setwd("/u/lestes/analyses/SAcropland/")
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/external/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
```

Background data, projections, database connections, and SA grid
```{r, eval = FALSE}
# Get SA shape
if(!file.exists(full_path(p_data, "MZshapes.Rdata"))) {
  system("wget http://dl.dropboxusercontent.com/u/31890709/MZshapes.Rdata")
}
load(full_path(p_data, "MZshapes.Rdata"))
sa_buf <- gBuffer(sa.shp, width = 3000)

# Connection
drv <- dbDriver("PostgreSQL")
con <- dbConnect(drv, dbname = "SouthAfricaSandbox", user = "postgis", 
                 password = "P0stG1S")

# projections, etc
prjsrid <- 97490  # Albers
prj_sql <- paste0("select proj4text from spatial_ref_sys where srid=", prjsrid)
prjstr <- dbGetQuery(con, prj_sql)$proj4text
gcsid <- 4326  # GCS
gcs_sql <- paste0("select proj4text from spatial_ref_sys where srid=", gcsid)
gcsstr <- dbGetQuery(con, gcs_sql)$proj4text

# Convert SA to GCS for cropping extent
sa_buf_gcs <- spTransform(sa_buf, CRSobj = CRS(gcsstr))

# Create 1 km grid from sa1kilo to rectify grids against that 
if(!file.exists(full_path(p_data, "sagrid.tif"))) {  # already made? 
  sql <- paste("SELECT gid, id, ST_AsText(ST_Centroid(geom)) as center", 
               "FROM sa1kilo")
  geom.tab <- dbGetQuery(con, sql)
  coord.mat <- do.call(rbind, lapply(1:nrow(geom.tab), function(y) {
    strip <- gsub("POINT|\\(|\\)", "", geom.tab[y, 3])
    coord.mat <- do.call(rbind, strsplit(strsplit(strip, ",")[[1]], " "))
  }))
  class(coord.mat) <- "numeric"
  point.tab <- cbind(geom.tab[, 1:2], coord.mat)
  colnames(point.tab)[3:4] <- c("x", "y") 
  pointsXYZ <- point.tab[, c(3:4, 2)]
  sa_r <- rasterFromXYZ(pointsXYZ)
  projection(sa_r) <- sa.buf@proj4string
  sa_r <- writeRaster(sa_r, filename = full_path(p_data, "sagrid.tif"), 
                      overwrite = TRUE)
} else {
  sa_r <- raster(full_path(p_data, "sagrid.tif"))
}
```

### Bring in landcover datasets
Let's start with the improved/fusion datasets, GLC share and geo-wiki's
```{r, eval = FALSE}
if(!file.exists(full_path(p_data, "glcsa_masked.tif"))) {  # see if file exists,  
  # to avoid redoing step
  path <- "/u/lestes/spatial_data/glc_share/"
  url <- "http://www.fao.org/geonetwork/srv/en/resources.get?id=47948&fname=GlcShare_v10_02.zip&acces=private"
  download.file(url, method = "auto", destfile = "glc_share.zip")
  unzip("glc_share.zip", exdir = "glc_share")
  glc <- raster(full_path(path, "glc_shv10_02.Tif"))
  projection(glc) <- gcsstr
  glcSA <- crop(glc, y=sa_buf_gcs, file = full_path(path, "glcSA.tif"), 
                overwrite = TRUE)  # crop to SA
  gdalwarp(srcfile = glcSA@file@name, 
           dstfile = full_path(path, "glcSA_alb.tif"), t_srs = prjstr, 
           tr = c(1000, 1000), r = "bilinear")
  glcSA_alb <- raster(full_path(path, "glcSA_alb.tif"))

  # Warp to SA grid
  glcsa <- resample(glcSA_alb, sa_r, method = "bilinear", 
                    filename = full_path(path, "glcSA_alb_rect.tif"))
  glcsa_m <- mask(glcsa, sa_r, file = full_path(p_data, "glcsa_masked.tif"),
                  overwrite = TRUE)
} else {
  glcsa_m <- raster(full_path(p_data, "glcsa_masked.tif"))
}

if(!file.exists(full_path(p_data, "geowikisa_masked.tif"))) {  # see if file exists, to avoid redoing step 
   # first downloaded manually from password protected geo-wiki site
  path <- "/u/lestes/spatial_data/geowiki/"
  unzip("cropland_hybrid_14052014v8.zip", exdir= "geowiki") 
  geowiki <- raster(full_path(path, "Hybrid_14052014V8.img"))
  projection(geowiki) <- gcsstr
  geowikiSA <- crop(geowiki, y=sa_buf_gcs, 
                    file = full_path(path, "geowikiSA.tif"))  # crop to SA
  gdalwarp(srcfile = geowikiSA@file@name, 
           dstfile = full_path(path, "geowikiSA_alb.tif"), t_srs = prjstr, 
           tr = c(1000, 1000), r = "bilinear")
  geowikisa_alb <- raster(full_path(path, "geowikiSA_alb.tif"))
  geowikisa <- resample(geowikisa_alb, sa_r, method = "bilinear", # Warp to SA
                        filename = full_path(path, "geowikisa_alb_rect.tif"))    
  geowikisa_m <- mask(glcsa, sa_r, 
                      file = full_path(p_data, "geowikisa_masked.tif"), 
                      overwrite = TRUE)
} else {
  geowikisa_m <- raster(full_path(p_data, "geowikisa_masked.tif"))
}
```

And the standard products, Globcover 2009 and MODIS
```{r, eval = FALSE}
path <- "/u/lestes/spatial_data/globcover_2009/"
if(!file.exists(full_path(path, "globSA_alb.tif"))) {
  url <- "http://due.esrin.esa.int/globcover/LandCover2009/Globcover2009_V2.3_Global_.zip"
  download.file(url, method="auto", destfile = "globcover_2009.zip")
  unzip("globcover_2009.zip", exdir= "globcover_2009")
  glob <- raster(full_path(path, "GLOBCOVER_L4_200901_200912_V2.3.tif"))
  globSA <- crop(glob, y=sa_buf_gcs)  # crop to extent of SA
  gdalwarp(srcfile = globSA@file@name, 
           dstfile = full_path(path, "globSA_alb.tif"), t_srs = prjstr, 
           tr = c(300, 300))
  globSA_alb <- raster(full_path(path, "globSA_alb.tif"))
} else {
  globSA_alb <- raster(full_path(path, "globSA_alb.tif"))
}

# MODIS landcover 
moddir <- "/u/lestes/spatial_data/MCD12Q1/"
if(!file.exists(full_path(moddir, "modis_type_1_alb.tif"))) {
  tiles <- expand.grid("h" = c(19, 20), "v" = c(11, 12))
  tiles <- apply(tiles, 1, function(x) paste("h", x[1], "v", x[2], sep = ""))
  url <- "http://e4ftl01.cr.usgs.gov/MOTA/MCD12Q1.051/2011.01.01/"
  tile.names <- getURL(url, verbose=TRUE, dirlistonly = TRUE) 
  tile.names1 <- strsplit(tile.names, "alt")[[1]]
  tile.names2 <- unique(substr(tile.names1, 20, 64))
  modnames <- tile.names2[sapply(tiles, function(x) grep(x, tile.names2))]
  lapply(modnames, function(x) {
    url <- paste0("http://e4ftl01.cr.usgs.gov/MOTA/MCD12Q1.051/2011.01.01/", x)
    download.file(url, method="auto", destfile = x)
  })
  dir.create("MCD12Q1")
  file.copy(from = dir(pattern = "hdf"), 
            to = paste(getwd(), "MCD12Q1", dir(pattern = "hdf"), sep = "/"))
  file.remove(dir(pattern = "hdf"))

  modis <- dir(moddir)

  # Translate to geotiff
  lapply(1:5, function(x) {
    batch_gdal_translate(infiles = modis, outdir = moddir, 
                         outsuffix = paste("_LC_", x, ".tif", sep = ""), 
                         sd_index = x)  
  })
  sdnames <- lapply(1:5, function(x) dir(moddir, 
                                         pattern = paste0("LC_", x, ".tif")))

  # warp to albers, mosaic, and clip to SA extent
  lapply(1:5, function(x) {
    print(paste("processing type", x))
    gdalwarp(srcfile = sdnames[[x]], 
             dstfile = paste0("modis_type_", x, "_alb.tif"), 
             t_srs = prjstr, tr = res(r), te = bbox(sa.buf)[1:4])
  })  
  mod_1 <- raster("modis_type1_alb.tif")
} else {
  mod_1 <- raster(full_path(moddir, "modis_type_1_alb.tif"))
}

```

And then South Africa's landcover database (30 m)
```{r, eval = FALSE}
path <- "/u/lestes/spatial_data/sa_lc_2009/"
setwd("/u/lestes/spatial_data/")
lcnm <- c("sa_ag.tif", "sa_af.tif")
if(length(dir(p_data, pattern = "sa_.*._masked.tif")) != 2) {
  url <- "http://planet.uwc.ac.za/BGISdownloads/landcover_2009.zip"
  download.file(url, method="auto", destfile = "sa_landcover_2009.zip")
  unzip("landcover_2009.zip", exdir = "sa_lc_2009")

  setwd(path)
  gdal_translate(src_dataset = "landcover", dst_dataset = "sa_lc_2009.tif", 
                 of = "GTiff", ot = "BYTE")
  #system(paste("rm -r", full_path(path, "landcover")))
  #file.remove("../landcover_2009.zip")

  # process masks--extract extract landcover class 2, cultivation, and class 6,
  # plantations, resample, mask with gdal tools
  av <- c(2, 6)
  a <- "sa_lc_2009.tif"
  dang <- Sys.time() 
  salc_l <- lapply(1:2, function(j) {
    print(paste("extracting cover for", lcnm[j]))
    gdal_calc(cstr = paste("A==", av[j], sep = ""), x = list("A" = a), 
              type = "Byte", filename = lcnm[j])
    nm <- full_path(p_data, 
                    paste(gsub("\\.tif", "", lcnm[j]), "_masked.tif", sep = ""))
    ext <- sapply(c("xmin", "ymin", "xmax", "ymax"), function(x) {
      slot(extent(sa_r), x)
    })
    print(paste("warping and masking", gsub(p_data, "", nm)))
    gdalwarp(srcfile = lcnm[j], t_srs = projection(sa_r), dstfile = nm, 
             r = "average", ot = "Float32", te = ext, srcnodata = 255, 
             dstnodata = 255, tr = c(1000, 1000), of = "GTiff", verbose = TRUE,
             overwrite = TRUE)
    file.remove(lcnm[j])
    r <- raster(nm)
  })
  Sys.time() - dang

} else {
  lcmnm <- paste(gsub("\\.tif", "", lcnm), "_masked.tif", sep = "")
  salc_l <- lapply(lcmnm, function(x) raster(full_path(p_data, x)))
} 
setwd(p_data)
```

### Mask out sugarcance
```{r, eval = FALSE}
if(!file.exists(full_path(p_data, "kzn_cane_masked.tif"))) {
  path <- "/u/lestes/spatial_data/kzn_landcover/"
  setwd(kznpath)
  gdal_translate(src_dataset = "kznlc11v1w31", dst_dataset= "kznlandcover.tif",
                 of = "GTiff")
  system(paste("rm -r", full_path(path, "kznlc11v1w31")))
  kznlc <- raster(full_path(npath, "kznlandcover.tif"))

  # system call to gdal_calc.py to extract landcover class 2, cultivation
  kznlconm <- "kzn_cane.tif"  # file.remove(kznlconm)
  ss <- strsplit(kznlc@file@name, "/")
  kzninnm <- ss[[1]][length(ss[[1]])]
  # kzninnm <- strsplit(kznlc.ss@file@name, "/")[[1]][length(strsplit(kznlc.ss@file@name, "/")[[1]])]
  pcalc <- paste("gdal_calc.py -A ", kzninnm, " --outfile=", kznlconm, 
                 " --overwrite --calc='(A==9)|(A==10)'", sep = "")
  system.time(system(pcalc))  # 49 seconds
  kzn_cane <- raster(full_path(path, kznlconm))
  # cext <- extent(c(20000, 40000, -3220000, -3200000))  # crops to small extent for examination
  # kznlc.ss <- crop(kznlc, cext, filename = full_path(kznpath, "kznlandcover_crop.tif"), overwrite = TRUE)
  # plot(kznlc.ss)
  # scane1 <- (kznlc.ss == 9); scane2 <- (kznlc.ss == 10)
  # plot(scane1, add = TRUE, col = c("transparent", "red"))
  # plot(scane2, add = TRUE, col = c("transparent", "purple"))
  # chk <- crop(kzn_cane, cext)
  # plot(chk); plot(scane1, add = TRUE, col = c("transparent", "red"), legend = FALSE) 
  # plot(scane2, add = TRUE, col = c("transparent", "red"), legend = FALSE)
  # file.remove("kznlandcover_crop.tif")
  # rm(chk, kznlc.ss, scane1, scane2)
  file.remove("kznlandcover.tif")

  # Warp and resample to SA grid
  setwd(p_data)
  gdalwarp(srcfile = kzn_cane@file@name, t_srs = projection(sa_r), 
           dstfile = "kzn_cane_1km.tif", r = "average", ot = "Float32", 
           srcnodata = 255, tr = c(1000, 1000), of = "GTiff", 
           verbose = TRUE, overwrite = TRUE)
  kzn_1km <- raster(full_path(p_data, "kzn_cane_1km.tif"))
  kzn_1km_r <- resample(kzn_1km, sa_r, method = "bilinear")
  kzn_1km_m <- mask(kzn_1km_r, sa_r, 
                    filename = full_path(p_data, "kzn_cane_masked.tif"), 
                    overwrite = TRUE)
} else {
  kzn_1km_m <- raster(full_path(p_data, "kzn_cane_masked.tif"))
}

```

### MODIS cropland variants
Different variants of cropland fractions from MODIS classes 12 and 14 (from the IGBP-DIS scheme). Class 14 is a cropland mosaic class, with minimum of 10% and a max of 60%, so create variants of 10%, 35% (mean) and 60%. 
```{r, eval = FALSE}
if(any(!file.exists(full_path(p_data, 
                              dir(p_data, 
                                  pattern = "mod_1*.*1kmrect_sum.tif"))))) {
  # MODIS, using main classification scheme (IGBP-DIS)
  mod_1_crop <- stack(mod_1 == 12, mod_1 == 14)
  mod_1_cropb <- brick(mod_1_crop)
  names(mod_1_crop) <- c("cropland", "mosaic")

  # assign percentages
  minset <- c(100, 10)  # assume 10% is min crop cover for MODIS mosaic class
  meanset <- c(100, 35)  # assume 35% is mean crop cover
  maxset <- c(100, 60)  # assume 60% is mean crop cover  
  modsa_min <- assignLCPct(mod_1_cropb, minset, 
                           fname = full_path(p_data, "mod_1_min.tif"))
  modsa_mu <- assignLCPct(mod_1_cropb, meanset, 
                          fname = full_path(p_data, "mod_1_mu.tif"))
  modsa_max <- assignLCPct(mod_1_cropb, maxset, 
                           fname = full_path(p_data, "mod_1_max.tif"))

  # aggregate, resample, mask to SA grid
  modsa_list <- lapply(list(modsa_min, modsa_mu, modsa_max), function(x) {
    nm <- gsub("//.tif", "", x@file@name)
    print(paste("Processing", nm))
    print("...aggregating")
    agg <- aggregate(x, fact = 2, na.rm = TRUE)  
    print("...resampling")
    crop1km <- resample(agg, sa_r, method = "bilinear")
    print("...masking")
    crop1km <- mask(crop1km, mask = sa_r, filename = paste0(nm, "_1kmrect.tif"),
                    overwrite = TRUE)
    return(crop1km)
  })

  # And them create single raster with summed proportions
  modsa_list_sum <- lapply(modsa_list, function(x) {
    nm <- paste(gsub("\\.tif", "", x@file@name), "_sum.tif", sep = "")
    paste(nm)
    calc(x, sum, filename = nm, overwrite = TRUE)
  })
} else {
  modsa_list_sum <- lapply(full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
}  
setwd(p_data)
```

### GlobCover cropland variants
And the same for GlobCover. We want Globcover class 11 (irrigated cropland), class 14 (rainfed), 20 (50-70% cropland), and 30 (20-50% cropland). For the latter two classes we create variants assumung the minimum, mean, and max of earh class
```{r, eval = FALSE}
if(any(!file.exists(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif"))))) {
  l <- lapply(c(11, 14, 20, 30), function(x) r <- globSA_alb == x)  # pull out classes
  s <- stack(l)
  globSA_cropb <- brick(full_path(p_data, "globSA_crop.grd"))
  rm(s, l)

  # assign percentages
  minset <- c(100, 100, 50, 20)
  maxset <- c(100, 100, 70, 50)
  meanset <- c(100, 100, 60, 35)
  globsa_min <- assignLCPct(globSA_cropb, minset, fname = full_path(p_data, "globSA_300_min.tif"))
  globsa_mu <- assignLCPct(globSA_cropb, meanset, fname = full_path(p_data, "globSA_300_mu.tif"))
  globsa_max <- assignLCPct(globSA_cropb, maxset, fname = full_path(p_data, "globSA_300_max.tif"))
  glob300list <- lapply(dir(pattern = "globSA_300+.+tif")[c(2:3, 1)], raster)

  # Aggregate them and resample and mask to SA grid
  globsa_list <- lapply(list(globsa_min, globsa_mu, globsa_max), function(x) {
    nm <- gsub("_300|.tif", "", x@file@name)
    print(paste("Processing", nm))
    print("...aggregating")
    agg <- aggregate(x, fact = 3, na.rm = TRUE)  # equivalent to summing, dividing by 900
    print("...resampling")
    crop1km <- resample(agg, sa_r, method = "bilinear")
    print("...masking")
    crop1km <- mask(crop1km, mask = sa_r, filename = paste(nm, "_1kmrect.tif", sep = ""), 
                    overwrite = TRUE)
    return(crop1km)
  })

  # And them create single raster with summed proportions
  globsa_list_sum <- lapply(globsa_list, function(x) {
    nm <- paste(gsub("\\.tif", "", x@file@name), "_sum.tif", sep = "")
    paste(nm)
    calc(x, sum, filename = nm)
  })
} else {
  globsa_list_sum <- lapply(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
} 
```

### Further masking of reference maps

```{r, eval = FALSE}
# Bring in SA landcover sets and mask
cover2007 <- brick(full_path(p_data, "cover2007.tif"))
cover2011 <- brick(full_path(p_data, "cover2011.tif"))

# sum these also with and without horticulture
sumfun <- function(x) sum(x, na.rm = TRUE)
gti <- lapply(list(cover2007, cover2011), function(y) {
  r <- calc(y[[-6]], sumfun)
  r[r > 100] <- 100
  writeRaster(r, file = gsub("\\.tif", "sum.tif", y@file@name), overwrite = TRUE)
})

# Leave commented out in case want to bring back horticulture class
# gti_hort <- lapply(list(cover2007, cover2011), function(y) {
#   r <- calc(y, sumfun)
#   r[r > 100] <- 100
#   writeRaster(r, file = gsub("\\.tif", "sum_h.tif", y@file@name), overwrite = TRUE)
# })

# Remove communal farmlands
sust_mask <- !is.na(cover2007[[2]]) | !is.na(cover2011[[2]])  # remove communal areas from both dates
sust_mask[sust_mask > 0] <- NA

# Set up SA mask including sugarcane and forestry -- add in Mpumalanga sugar cane when it becomes available
m1 <- kzn_1km_m
m1[is.na(m1)] <- 0
m1[m1 > 0] <- NA
m2 <- salc_l[[2]]
m2[is.na(m2)] <- 0
m2[m2 > 0] <- NA
sa_masks <- sa_r > 0
sa_masks <- m1 + m2 + sa_masks
rm(m1, m2)
full_mask <- writeRaster(sa_masks + sust_mask, file = full_path(p_data, "mask.tif"), overwrite = TRUE)

# Apply masks
lapply(gti, function(x) {
  out_name <- paste(gsub("\\.tif", "", x@file@name), "_mask.tif", sep = "")
  mask(x, full_mask, file = out_name, overwrite = TRUE)
})

```

# Cropland errors, part 1
This section of the analysis calculates the differences between the reference cropland percentages and those from the various landcover products. It draws from results generated by the [cropland pre-processing](cropland-preprocess.html). Bias statistics calculated here are primarily for the supplementals.

## Analyses
```{r, eval = FALSE}
library(lmisc)
library(raster)
library(SAcropland)
p_root <- proj_root("SAcropland")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
```

### Load cropland test grids 
Derived from global landcover products (note: it turns out GLC-Share and geowiki are the same thing, so GLC-Share dropped)

```{r, eval = FALSE}
salc <- raster(full_path(p_data, "sa_ag_masked.tif")) * 100
ds <- full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)])
globsa_list_sum <- lapply(ds, raster)
fnames <- full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)])
modsa_list_sum <- lapply(fnames, raster)
# glcsa <- raster(full_path(p_data, "glcsa_masked.tif"))
geow <- raster(full_path(p_data, "geowikisa_masked.tif"))
lclist <- c(salc, globsa_list_sum, modsa_list_sum, geow)  # into list
names(lclist) <- c("sa30", "globmin", "globmu", "globmax", "modmin", "modmu",
                   "modmax", "geow")
```

#### Reference data
```{r, eval = FALSE}
load(full_path(p_data, "MZshapes.Rdata"))  
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
```

#### Apply NA masks
```{r, eval = FALSE}
sumna <- function(x) sum(x, na.rm = FALSE)
namask <- calc(stack(stack(gti), stack(lclist)), sumna)
namask[namask >= 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(namask2, sum)
namask[namask == 0] <- NA
# cellStats(namask, sum)
writeRaster(namask, filename = full_path(p_data, "namask.tif"), 
            overwrite = TRUE)

# mask out all joint NAs
gti <- lapply(gti, function(x) mask(x, namask))
lclist <- lapply(lclist, function(x) mask(x, namask))

```

### Calculate differences at different levels of aggregation
This step draws on functions I wrote to calculate the actual and absolute differences between a given "master" grid and lists of other grids. This comparison is done at multiple aggregations, ranging from 1-600 km resolution cropland averages. The lists have two levels of nesting, here the upper provides the resolution, the inner provides the different landcover datasets.

```{r, eval = FALSE}
# Aggregate rasters
fact <- c(5, 10, 25, 50, 100)
#fact <- 25

area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)  # cells/pixel
lc_agg <- aggregate_rast_list(fact, lclist)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters

# Calculate differences
# actual
dlist_act <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)),
                            gti_agg, lc_agg, "a - b", silent = FALSE)  
# absolute
dlist_abs <- rast_list_math(list(names(lclist), names(lc_agg), names(gti)), 
                            gti_agg, lc_agg, "abs(a - b)", silent = FALSE)  
dlist_gti <- rast_list_math(list("g2007", names(gti_agg), "g2011"), 
                            gti_agg, gti_agg, "a - b", silent = FALSE)

save(dlist_act, file = full_path(p_data, "d_grid_act.rda"))
save(dlist_abs, file = full_path(p_data, "d_grid_abs.rda"))

# Write out the gti cover raster and the actual bias rasters for running impact of bias examples in land
# surface model (25 km resolution)
r <- gti_agg$f25$g2011
projection(r) <- projection(salc)
writeRaster(r, filename = full_path(p_data, "gti_2011_25km.tif"))
b <- brick(stack(lapply(c("sa30", "globmu", "modmu", "geow"), function(x) {
  dlist_act[[x]]$f25$g2011
})))
projection(b) <- projection(salc)
writeRaster(b, filename = full_path(p_data, "lc_bias_25km.tif"))

# r <- raster(full_path(p_data, "gti_2011_25km.tif"))
# lcb <- brick(full_path(p_data, "lc_bias_25km.tif"))
# plot(r - lcb, axes = FALSE, box = FALSE)
# plot(lcb)
# plot(round(lc_agg$f25$geow - (r - lcb$lc_bias_25km.4), 7))

```

### Error versus cropland density
#### Weighting rasters 
Based on number of non-NA cells in each aggregated pixel, and then "bin" rasters, where each category represents a different cropland fractional cover range (0-5%, 5-10%, etc) 
```{r, eval = FALSE}

# cropland cover bins
binv <- seq(0, 100, 5)
gti_bins <- lapply(gti_agg, function(x) {
  lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE))
})
```

#### Calculate bias/MAE, pooling MODIS and GlobCover across all versions
Step 1. Reshape lists, so that all MODIS and GlobCover variants are pooled by scale
```{r, eval = FALSE}
# Set up indexing parameters
fact_all <- c(1, fact)# full aggregation levels
lev_vec <- paste("f", fact_all, sep = "") 
nm_rt <- c("sa30", "glob", "mod", "geow")

# Reshape the list of actual landcover differences
dlists <- lapply(list(dlist_act, dlist_abs), function(i) {
  resh <- lapply(nm_rt, function(x) {  # select sensor
    l2 <- lapply(lev_vec, function(y) {  # isolate by scale
      l3 <- lapply(i[grep(x, names(i))], function(z) z[[y]])  # sensor by scale
        named_out(l3, names(i)[grep(x, names(i))])
      })
    named_out(l2, lev_vec)
  })
  named_out(resh, nm_rt)
})
names(dlists) <- c("act", "abs")
rm(dlist_act, dlist_abs)  # remove original to make space

```

#### Then calculate bias/MAE across datasets

Note: this uses an older and slower version of the code. Subsequent analyses use the faster version based on data.tables.  
```{r, eval = FALSE}
dang <- Sys.time()
# i <- dlists[[1]]; x <- names(i)[1]; y <- lev_vec[5]; j <- 14; z <- 1; k <- 1
dstats_all <- lapply(dlists, function(i) {
  print("processing 1 of 2 main bias lists")
  dstats <- lapply(names(i), function(x) {
    print(paste("..", x))
    l1 <- lapply(lev_vec, function(y) {
      print(paste("....", y))
      d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(j) {
        print(paste("......bin", j))
        d2 <- do.call(rbind, lapply(1:length(i[[x]][[y]]), function(z) {
          print(paste("........variant", z))
          l3 <- i[[x]][[y]][[z]]
          d3 <- do.call(rbind, lapply(1:length(l3), function(k) {
            rs <- gti_bins[[y]][[k]] == j    
            rs[rs == 0] <- NA
            o <- rs * l3[[k]]
            w <- area_wgts[[y]][[1]] * rs
            oo <- cbind.data.frame("lc" = rep(names(i)[z], length(w)), 
                                   getValues(o), getValues(w))
          }))
          d3f <- d3[!is.na(d3[, 2]), ]
        }))
        if(nrow(d2) > (length(i[[x]][[y]]) * 10)) {
          qs <- c(nrow(d2), box_stats(d2[, 2], weighted = TRUE, 
                                      weight.opts = list("weights" = d2[, 3])))
        } else if((nrow(d2) > 0) & (nrow(d2) < length(i[[x]][[y]]) * 20)) {
          qs <- c(nrow(d2), rep(NA, 5), 
                  Hmisc::wtd.mean(d2[, 2], weights = d2[, 3], na.rm = FALSE))
        } else {
          qs <- rep(NA, 7)
        }
      }))
      cbind(binv[-1], d1)
    })
    named_out(l1, lev_vec)
  })
  named_out(dstats, names(i))
})
dut <- Sys.time() - dang  # 33 minutes
save(dstats_all, file = full_path(p_data, "bias_stats.rda"))
```

### Differences between 2007 and 2011 reference
```{r, eval = FALSE}
dstats_gti <- lapply(lev_vec, function(x) {
    #x <- lev_vec[[2]]
    i <-  dlist_gti$g2007[[x]][[1]]
    print(paste("..", x))
    d1 <- do.call(rbind, lapply(1:(length(binv) - 1), function(y) {
      #y <- 1
      print(paste("....", y))
      rs <- gti_bins[[x]][[2]] == y    
      rs[rs == 0] <- NA
      o <- rs * i
      w <- area_wgts[[x]][[1]] * rs
      oo <- cbind.data.frame(rep(y, length(w)), getValues(o), getValues(w))
      oof <- oo[!is.na(oo[, 2]), ]
      if(nrow(oof) > 10) {
        qs <- c(nrow(oof), box_stats(oof[, 2], weighted = TRUE, 
                                     weight.opts = list("weights" = oof[, 3])))
      } else if((nrow(oof) > 0) & (nrow(oof) < 10)) {
        qs <- c(nrow(oof), rep(NA, 5),  
                Hmisc::wtd.mean(oof[, 2], weights = oof[, 3], na.rm = FALSE))
      } else {
        qs <- rep(NA, 7)
      }
    }))
    cbind(binv[-1], d1)
})
save(dstats_gti, file = full_path(p_data, "bias_stats_gti.rda"))

```

Checks: compared number of non-NA cells coming out of each GTI version at each level of aggregation (fine); compared cell counts between previous committed version and this one (difference exists, but due to fact that gtimu is no longer considered; checked that previous versions of difference rasters and reshapes of them were referencing the correct GTI-landcover bias rasters (i.e. to make sure rasters weren't incorrectly indexed: all fine)); box_stats with weighted quantile option compared to previous commit where wtd.quantile was used separately (fine). Compared results from original roughed out version of approach for calculating bias statistics with results from new outputs list (using geowiki at 20 km aggregation as the comparison)

### 2011 reference-test bias/accuracy
Whole country and agricultural areas only
```{r, eval = FALSE}
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids

# Subset difference grids, for 2011
snms <- c("sa30", "globmu", "modmu", "geow")
lc_pct_diff <- lapply(dlist_act[snms], function(x) {
  sapply(x, function(y) list(y$g2011))
})

# construct agricultural area union mask
lev <- names(lc_pct_diff[[1]])
lcu <- lapply(lev, function(x) {
  print(paste(".", x))
  lcb <- lapply(snms, function(y) {
    print(paste("...", y))
    gti_gt0 <- Which(gti_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(lc_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev

bins <- lapply(gti_bins, function(x) x$g2011)
awgts <- lapply(area_wgts, function(x) x[[1]])

# calculate means for 2011
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)
a <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wma, "mua", "bias", TRUE)
mu_dt <- rbind(a, b)
mu <- extract_stat(lev, snms, "all", "mu", "bias", mu_dt)  # ag area means
mua <- extract_stat(lev, snms, "all", "mua", "bias", mu_dt)  # ag area abs mean
mu_nm <- extract_stat(lev, snms, "all", "mu.nm", "bias", mu_dt)  # country mean
mua_nm <- extract_stat(lev, snms, "all", "mua.nm", "bias", mu_dt)  # country abs
# across bin mean and abs mean
mu_bin <- mu_dt[!like(bin, "all") & bvals == "mu", mean(bias), by = .(ol, il)]
setnames(mu_bin, "V1", "bias")
mua_bin <- mu_dt[!like(bin, "all") & bvals == "mua", mean(bias), by = .(ol, il)]
setnames(mua_bin, "V1", "bias")
# mu_dt[il == "sa30" & bin != "all" & ol == "f1" & bvals == "mua", mean(bias)]

# check
# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
bv <- "mua"
for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
                c("globmu", "f1"))) {
  x <- chk[1] 
  y <- chk[2] 
  print(paste(".", x, "..", y))
  rs <- lcu[[y]][[x]] 
  rs[rs == 0] <- NA
  l3 <- abs(lc_pct_diff[[x]][[y]])
  o <- rs * l3
  w <- awgts[[y]][[1]] * rs
  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
  print(b[ol == y & il == x & bvals == bv & bin == "all", bias] ==
          round(wmu, 2))
}  # check

svec <- c("mu_nm", "mua_nm", "mu", "mua", "mu_bin", "mua_bin")
aa <- c(rep("Country", 2), rep("Agricultural", 2), rep("Density independent", 2))
bb <- c("Bias", "MAE", "Bias", "MAE", "Bias", "MAE")
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")

lcf_out11 <- do.call(rbind, lapply(1:length(svec), function(i) {
  ao <- do.call(rbind, lapply(snms, function(x) {
    ai <- sapply(lev, function(y) {
      get(svec[i])[ol == y & il == x, bias]
    })
    named_out(ai, paste(c(1, fact), "km"))
  }))
  #cbind.data.frame("Region" = aa[i], "Map" = mcap, a)
  cbind.data.frame(Region = aa[i], Metric = bb[i], "Map" = mcap, ao)
}))

caption <- paste("Biases and mean absolute errors (MAE) in cropland maps",
                 "relative to the 2011 reference map for each aggregation",
                 "scale calculated over the entire country, for the union of",
                 "agricultural regions (cropland $>$ 0), and as density", 
                 "independent means, wherein the mean bias/MAE values",
                 "for each of 20 cropland cover classes (representing",
                 "5\\% increments of cover 0\\% to 100\\% defined",
                 "by the reference map) were calculated and then averaged.")
lcfout_xtab <- xtable(lcf_out11, digits = 1, caption = caption)
print(lcfout_xtab, type = "latex", file = "paper/figures/cropland2011-bias.tex", 
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)

# do.call(rbind, lapply(snms, function(i) {
#   mu_dt[!like(bin, "all") & ol == "f1" & il == i & bvals == "mu.nm", N]  
# }))
mubias_2011 <- mu_dt
save(lcf_out11, mubias_2011, file = full_path(p_data, "lcf_out2011.rda"))

```

### 2007 reference-test bias/accuracy
Whole country and agricultural areas only
```{r, eval = FALSE}
# Subset difference grids, for 2011
snms <- c("sa30", "globmu", "modmu", "geow")
lc_pct_diff <- lapply(dlist_act[snms], function(x) {
  sapply(x, function(y) list(y$g2007))
})

# construct agricultural area union mask
lev <- names(lc_list[[1]])
lcu <- lapply(lev, function(x) {
  print(paste(".", x))
  lcb <- lapply(snms, function(y) {
    print(paste("...", y))
    gti_gt0 <- Which(gti_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(lc_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev

bins <- lapply(gti_bins, function(x) x$g2007)
awgts <- lapply(area_wgts, function(x) x[[1]])

# calculate means for 2007
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)
a <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, lc_pct_diff, wma, "mua", "bias", TRUE)
mu_dt <- rbind(a, b)
mu <- extract_stat(lev, snms, "all", "mu", "bias", mu_dt)
mua <- extract_stat(lev, snms, "all", "mua", "bias", mu_dt)
mu_nm <- extract_stat(lev, snms, "all", "mu.nm", "bias", mu_dt)
mua_nm <- extract_stat(lev, snms, "all", "mua.nm", "bias", mu_dt)
# across bin mean and abs mean
mu_bin <- mu_dt[!like(bin, "all") & bvals == "mu", mean(bias), by = .(ol, il)]
setnames(mu_bin, "V1", "bias")
mua_bin <- mu_dt[!like(bin, "all") & bvals == "mua", mean(bias), by = .(ol, il)]
setnames(mua_bin, "V1", "bias")

# check
# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
bv <- "mua"
for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
                c("globmu", "f1"))) {
  x <- chk[1] 
  y <- chk[2] 
  print(paste(".", x, "..", y))
  rs <- lcu[[y]][[x]] 
  rs[rs == 0] <- NA
  l3 <- abs(lc_pct_diff[[x]][[y]])
  o <- rs * l3
  w <- awgts[[y]][[1]] * rs
  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
  print(b[ol == y & il == x & bvals == bv & bin == "all", bias] ==
          round(wmu, 2))
}  # check

svec <- c("mu_nm", "mua_nm", "mu", "mua", "mu_bin", "mua_bin")
aa <- c(rep("Country", 2), rep("Agricultural", 2), rep("Density independent", 2))
bb <- c("Bias", "MAE", "Bias", "MAE", "Bias", "MAE")
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")

lcf_out07 <- do.call(rbind, lapply(1:length(svec), function(i) {
  ao <- do.call(rbind, lapply(snms, function(x) {
    ai <- sapply(lev, function(y) {
      get(svec[i])[ol == y & il == x, bias]
    })
    named_out(ai, paste(c(1, fact), "km"))
  }))
  #cbind.data.frame("Region" = aa[i], "Map" = mcap, a)
  cbind.data.frame(Region = aa[i], Metric = bb[i], "Map" = mcap, ao)
}))

caption <- paste("Biases and mean absolute errors (MAE) in cropland maps",
                 "relative to the 2007 reference map for each aggregation scale",
                 "calculated over the entire country, for the union of",
                 "agricultural regions, and as density", 
                 "independent means, wherein the mean bias/MAE values",
                 "for each of 20 cropland cover classes",
                 "(representing 5\\% increments of cover 0\\% to 100\\% defined",
                 "by the reference map) were calculated and then averaged.")
lcfout_xtab <- xtable(lcf_out07, digits = 1, caption = caption)
print(lcfout_xtab, type = "latex", file = "paper/figures/cropland2007-bias.tex", 
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)

```

# Cropland errors, part 2

The following provide further analyses and plots of bias and accuracy metrics calculated from spatial error maps created in [compare-landcover](compare-landcover.html) (which includes some supplementary metrics),

## Spatial biases
### Datasets

```{r, eval = FALSE}
library(raster)
library(lmisc)
library(R.utils)
library(RColorBrewer)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")

# Load in datasets
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
load(full_path(p_data, "d_grid_abs.rda"))  # actual diffence grids
load(full_path(p_data, "bias_stats.rda"))  # bias statistics
load(full_path(p_data, "bias_stats_gti.rda"))  # bias statistics
# cellStats(!is.na(dlist_act$modmu$f1$g2011), sum)
```

## Error maps
Plot maps of biases at different resolutions. To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and masked before plotting.   
```{r, eval = FALSE}
namask <- raster(full_path(p_data, "namask.tif"))
# cellStats(namask, sum)
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
lev <- c("f1", "f25", "f50", "f100")  # just show four levels

# disggregate selected rasters at selected levels
# x <- snms[3]; y <- lev[3]
disagg <- lapply(snms, function(x) {
  l1 <- lapply(lev, function(y) {
    if(y == "f1") {
      r <- dlist_act[[x]][[y]]$g2011
    } else {
      r <- disaggregate(dlist_act[[x]][[y]]$g2011, 
                        fact = as.numeric(gsub("f", "", y)))
      r <- raster::mask(crop(r, namask), namask)
    }
  })
  named_out(l1, lev)
})
names(disagg) <- snms

# writeRaster(disagg$globmu$f100, 
#             filename = "external/ext_data/test/disagg100km.tif")
```

### Plot of selected rasters and levels 
```{r, eval = FALSE}
# Plotting colors
# cols <- colorRampPalette(c("red", "grey80", "blue4"))
brks <- c(-100.1, -75, -50, -25, -10, -5, -1, 1, 5, 10, 25, 50, 75, 100.1)
cols <- c(rev(brewer.pal(length(brks) / 2, "Reds")[-1]), "grey80", 
          brewer.pal(length(brks) / 2, "Blues")[-1])
brklen <- length(brks) - 1
# cols <- cols(brklen)
# cols[c(1, length(cols))] <- c("darkred", "purple3")
legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "bias_map.pdf"), height = 7, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(3.7, 1.5, 1.5, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols, breaks = brks,
         legend = FALSE)
    if(j == 1) mtext(mcap[i], side = 2, line = -0.25, cex = cx)
    if(i == 1) mtext(lev2[j], side = 3, line = -0.5, cex = cx)
  }
}
flex_legend(ncuts = brklen, legend.text = legtext, 
            legend.vals = round(brks), longdims = c(0.2, 0.8), 
            shortdims = c(0.05, 0.01), colvec = cols, srt = c(270, 0), 
            horiz = TRUE, textside = "bottom", legend.pos = c(4, 7), 
            leg.adj = list(c(0.2, 0.35), c(0.5, -0.5)), cex.val = cx, 
            textcol = lcol, bordercol = lcol)
dev.off()

lev <- names(dlist_abs$sa30)
```
The map below was created by the code above, showing biases at 1 km, 25 km, 50 km, and 100 km from South Africa's landcover product, the mean estimates from 
GlobCover and MODIS, and Geowiki/GLC-SHARE.

<p align="center"><img src="../paper/figures/bias_map.pdf" width="400px"></p>

## Bias/MAE in relation to field density

### 1 km Bias/MAE with boxplots

For supplemental.
```{r, eval = FALSE}
#btype <- c("bias_act.pdf", "bias_abs.pdf")
# plot matrix
nc <- 4
m <- matrix(rep(c(rep(1, 8), rep(2, 8), rep(3, 8), rep(4, 8)), 2), nrow = 6, 
            ncol = 32, byrow = TRUE) 
cols <- c("red", "orange3", "green4", "blue", "yellow")
bcol <- c("grey50", "grey50", "grey50", "grey")
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
fcol <- "grey"
lev <- names(dstats_all$act$sa30)
lvec <- seq(5, 100, 5)
pp <- list("x" = c(-2, 21), "y" = c(0, 0), 
           "x21" = list(c(0, 9), c(9, 21)),
           "y2" = list(c(seq(-100, -10, 10), seq(10, 100, 10)), 
                       seq(10, 100, 10)), 
           "y1lwd" = list(2, 1),  
           "y2lwd" = 0.5, #c(1, 0.5, 1, 0.5, 0.5, 0.5, 0.5, 1, 0.5, 1),
           "y2lc" = list(c("black", rep("grey80", 4), "black", rep("grey80", 4), 
                           rep("grey80", 4), "black", rep("grey80", 4),"black"),
                         rep("grey80", 10)),
           "yl" = list(c(-100, 100), c(0, 100)), 
           "yl2" = c(-20, 60), "xl" = c(1, 20),
           "xaxl" = unlist(lapply(seq(5, 100, 10), function(x) c(x, " "))),
           "xaxl2" = ifelse(seq(5, 100, 5) %in% c(25, 50, 75), seq(5, 100, 5), 
                            " "),
           "yax1" = list(seq(-100, 100, 50), seq(0, 100, 20)), 
           "ycol" = c("grey70"))  # plot parameters

pdf(full_path(p_fig, "biases_1km.pdf"), width = 7, height = 5)
lmat <- rbind(m, m + max(m))
l <- layout(lmat)
par(mar = c(3, 0, 0, 0), oma = c(1, 4, 2, 2))
for(bt in 1:2) {
  snms <- names(dstats_all[[bt]])[names(dstats_all[[bt]]) != "glcsh"]
  dset <- dstats_all[[bt]][snms]
  for(i in 1:length(snms)) { i
    d <- dset[[snms[i]]]$f1
    n <- nrow(d)
    plot(pp$x, pp$y, ylim = pp$yl[[bt]], xlim = pp$xl, type = "l", 
         lwd = pp$y1lwd[[bt]], axes = FALSE, 
         xlab = "", ylab = "")
    if(i == 1) axis(2, at = pp$yax1[[bt]], labels = pp$yax1[[bt]], las = 2, 
                    mgp = c(1, 0.6, 0))
    axis(1, at = 1:20, labels = pp$xaxl, las = 2, mgp = c(1, 0.4, 0), 
         tcl = -0.3, pos = pp$yaxp[[bt]], cex = 0.8)
    for(z in 1:length(pp$y2[[bt]])) {
      lines(pp$x, rep(pp$y2[[bt]][z], 2), lwd =pp$y2lwd, col = pp$y2lc[[bt]][z])
    }
    for(k in 1:n) {
      a <- seq(grep("2.5", colnames(d)), by = 1, length.out = 6)
      if(!is.na(d[k, 2])) {
        boxplot_v(k, d[k, a], n, 150, 10, bcol, fcol, pcex = 0.5, 
                  lwd = c(1, 0.5, 2))
      }
    }
    lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = 3)
    if(i == 1) mtext(ifelse(bt == 1, "Bias (%)", "Mean absolute error (%)"), 
                     side = 2, line = 2.2, cex = 0.8)
    if(bt == 1) mtext(lcnms[i], side = 3, line = 0, col = cols[i])
  }
}
mtext("% Cropland", side = 1, outer = TRUE, line = -1, cex = 0.8)
dev.off()
```

### Bias/MAE with scale

Note: the mean values presented in the figure here is a straight mean across 5 percentile bins, which gives bias independent of cropland density. Supplementary tables report the bin-weighted biases. 
```{r, eval=FALSE}
#lev <- lev[!lev %in% c("f5", "f300")]
levnms <- paste(gsub("f", "", lev), "km")
m2 <- unlist(lapply(seq(1, by = 1, length = length(lev) * 4), function(x) {
  rep(x, nc)
}))
m3 <- do.call(rbind, lapply(seq(1, length(m2), length(lev) * nc), function(x) {
  do.call(rbind, rep(list(m2[x:(x + length(lev) * nc - 1)]), nc))
}))

yl <- seq(pp$yl2[1], pp$yl2[2], 10)
ylcol <- c("black", "grey", "black", rep("grey", 5), "black")
lt <- c(5, 1)
ylwd <- c(1.5, 2)

wm <- function(x, w) stats::weighted.mean(x, w, na.rm = TRUE)
wma <- function(x, w) stats::weighted.mean(abs(x), w)

pdf(full_path(p_fig, "biases_1-100km.pdf"), width = 7.5, height = 7)
l <- layout(m3)
# layout.show(l)
# i <- 1; j <- 1; bt <- 1
par(mar = c(0.5, 0, 1, 0.3), oma = c(5, 5, 2, 2))
for(i in 1:length(snms)) {
  highcounter <- rep(0, length(lev))
  for(j in 1:length(lev)) {
    plot(pp$x, pp$y, ylim = pp$yl2, xlim = pp$xl, type = "l", lwd = 1, 
         axes = FALSE, xlab = "", ylab = "", xaxs = "i")
    for(z in 1:length(yl)) {
      lines(pp$x, rep(yl[z], 2), lwd = pp$y2lwd, col = ylcol[z])
    }
    muv <- rep(NA, 2)
    muv2 <- rep(NA, 2)
    for(bt in 1:2) {
      d <- dstats_all[[bt]][[i]][[j]]
      n <- nrow(d)
      lines(1:n, d[, "mu"], pch = 20, col = cols[i], lwd = ylwd[bt], 
            lty = lt[bt])
      muv[bt] <- round(mean(d[, 8], na.rm = TRUE)) # unweighted means
      muv2[bt] <- round(wm(d[, 8], d[, 2]))  # weighted means
      toohigh <- which(d[, "mu"] > pp$yl2[2])
      if(length(toohigh) > 0) {
        highcounter[j] <- 1
        xi <- toohigh[which.max(d[toohigh, "mu"])]
        text(xi, y = pp$yl2[2] + 4, labels = round(d[xi, "mu"]), cex = 1.1,
             col = "grey50", xpd = NA)
      }
    }
    text(5, pp$yl2[2] - 5, labels = paste(muv, collapse = " / "), cex = 1.1, 
         col = "grey10", font = 4)
    text(5, pp$yl2[2] - 16, labels = paste(muv2, collapse = " / "), cex = 1.1, 
         col = "grey10")
    if(j == 1) axis(2, at = yl, labels = yl, las = 2, mgp = c(1, 0.6, 0))
    if(j == 1) mtext(lcnms[i], side = 2, line = 3, cex = 1, col = cols[i])
    if(i == 1) mtext(levnms[j], side = 3, line = -0.4, cex = 1.2)
    axis(1, at = c(1, 5, 10, 15, 20), labels = rep("", 5), mgp = c(1, 1, 0), 
         tcl = -0.7, lwd.ticks = 1.2, pos = pp$yl2[1])
    axis(1, at = 1:20, labels = rep("", 20), las = 2, mgp = c(1, 0.4, 0), 
         tcl = -0.3, pos = pp$yl2[1])
    if(i == length(snms)) {
      axis(1, at = 1:20, labels = pp$xaxl2, las = 2, mgp = c(1, 0.8, 0), 
           tcl = -0.3, pos = pp$yl2[1])
    }
  }
}
mtext("% Bias/MAE", side = 2, line = 2, cex = 0.8, outer = TRUE, adj = 0.5)
mtext("% Cropland", side = 1, outer = TRUE, line = 1.5, cex = 0.8)
par(xpd = NA)
x <- grconvertX(0.75, from = "ndc", to = "user")
y <- grconvertY(0.07, from = "ndc", to = "user")
legend(x = x, y = y, legend = c("Bias", "Mean absolute error"), lty = lt, 
       lwd = ylwd, bty = "n", cex = 1.2)
dev.off()

```

### Total error
Calculate how much cropland estimates differ between datasets
```{r, eval = FALSE}
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
gti <- lapply(gti, function(x) mask(x, namask))  # apply namask
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)

# recreate cropland percentages
lc2007 <- lapply(snms, function(x) gti$g2007 - dlist_act[[x]]$f1$g2007)
lc2011 <- lapply(snms, function(x) gti$g2011 - dlist_act[[x]]$f1$g2011)
# plot(lc2007[[2]] - lc2011[[2]])  # same thing, as it should be

# calculate km2 cropland in LC datasets
lckm2 <- sapply(lc2011, function(x) cellStats(x / 100, sum))
refkm2 <- sapply(gti, function(x) cellStats(x / 100, sum))
refkm2[2] / refkm2[1]  # 3.6 percent more in 2011 than 2007

# bias in total cropland relative to GTI
stats2007 <- round((refkm2[2] - lckm2) / refkm2[2] * 100, 1)
stats2011 <- round((refkm2[1] - lckm2) / refkm2[1] * 100, 1)

# N km^2 in country
#sapply(lc2007, function(x) cellStats(!is.na(x), sum))
# sapply(lc2011, function(x) cellStats(!is.na(x) , sum))
# sapply(gti, function(x) cellStats(!is.na(x) , sum))
totkm2 <- cellStats(!is.na(gti$g2011) , sum)
save(stats2007, stats2011, totkm2, refkm2, 
     file = full_path(p_data, "total-cropland.rda"))
```

Total square kilometers of South Africa being evaluated
```{r, echo = FALSE}
library(lmisc)
p_root <- proj_root("SAcropland")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
load(full_path(p_data, "total-cropland.rda"))
totkm2
```

Total square kilometers of cropland according to 2007 and 2011 GTI data
```{r, echo = FALSE}
refkm2
```

Percent bias in total cropland area estimates relative to 2007 GTI data
```{r, echo = FALSE}
stats2007
```

And relative to 2011 GTI data
```{r, echo = FALSE}
stats2011
```

## Cropland bias/MAE plots

### Primary method

Weighted by __actual__ cropland percentage, and then, with aggregation, by number of pixels being aggregated.

```{r, eval = FALSE}
# Create weights maks
fact <- c(5, 10, 25, 50, 100)
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
# cellStats(!is.na(gti$g2011), sum)
# cellStats(namask, sum)

namask2 <- !is.na(namask)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts, function(x) x[[1]])

# Subset difference grids, for 2011
snms <- c("sa30", "globmu", "modmu", "geow")
err2011 <- lapply(lev, function(x) {
  sapply(dlist_act[snms], function(y) list(y[[x]]$g2011))
})
names(err2011) <- lev

gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters
gti_agg11 <- lapply(gti_agg, function(x) x$g2011)

# calculate bias/MAE for data
a <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wm, "mu")
b <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wma, "mua")
bacc2011 <- rbind(a, b)

# check error stats - do they match alternate approaches?
for(i in c("f1", "f25", "f10")) {
  for(j in c("sa30", "modmu", "geow")) { 
    print(paste("cross-checking calculations in", i, j))
    a1 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
                      "mu", aweight = FALSE)[, j, with = FALSE]
    a2 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
                      "mu", rweight = FALSE, aweight = FALSE)[,j, with=FALSE]
    a3 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
                      "mu")[, j, with = FALSE]
    c1 <- getValues(err2011[[i]][[j]])
    w1 <- getValues(gti_agg11[[i]])
    w2 <- getValues(awgts[[i]])
    print("non-area weighted mean matches?")
    print(a1 == round(weighted.mean(c1, w1, na.rm = TRUE), 2))
    print("totally unweighted mean matches?")
    print(a2 == round(cellStats(err2011[[i]][[j]], mean), 2))
    print("double weighted mean matches?")
    print(a3 == round(weighted.mean(c1, w1 * w2, na.rm = TRUE), 2))
  }
}

# Subset difference grids, for 2007
snms <- c("sa30", "globmu", "modmu", "geow")
err2007 <- lapply(lev, function(x) {
  sapply(dlist_act[snms], function(y) list(y[[x]]$g2007))
})
names(err2007) <- lev
gti_agg07 <- lapply(gti_agg, function(x) x$g2007)

a <- bias_statsw_list(gti_agg07, awgts, err2007, snms, wm, "mu")
b <- bias_statsw_list(gti_agg07, awgts, err2007, snms, wma, "mua")
bacc2007 <- rbind(a, b)

fact <- c(1, 5, 10, 25, 50, 100)
alph <- c(225, 40)
x <- c(0, 3, 6, 9, 12, 15)
w <- 3 / 8
xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
o <- c(0, w)
xa <- sapply(x, function(x) x + xo)
cx <- c(1.25, 1, 1)
g1 <- "grey90"
reg <- c("Country", "Agricultural", "Density independent")
lcnms <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
mutype <- c("mua", "mu")
# svec <- paste(gsub("f", "", bacc2011[stnm == "mu", ol]), "km")
svec <- bacc2011[stnm == "mu", ol]
xl <- c(-0.5, 18)
yl <- c(-30, 50)
shd <- c(4.5, 10.5, 16.5)
cols <- c("red", "orange3", "green4", "blue")
colnms <- snms
pchs <- c("+", "o", "x")
yax <- seq(yl[1], yl[2], 5)
xax <- seq(1, 6, 5 / (length(yax) - 1))

pdf(full_path(p_fig, "cropland_bias_main.pdf"), height = 7, width = 7)
par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "")#,
for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
abline(h = yax, v = NULL, col = "grey80", lty = 1)
polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
# i <- 1; j <- 1; k <- 1
for(i in 1:length(svec)) {
  lv <- svec[i]
  for(j in 1:length(snms)) {
    # pchs1 <- pchs
    nm <- snms[j]
    for(k in 1:length(mutype)) {
      dat <- bacc2011[ol == lv & stnm == mutype[k], nm, with = FALSE][[1]]
      pcol <- makeTransparent(cols[j], alpha = alph[k])
      polyfunc2(xa[j, i] + o[k], c(0, dat), w = w, col = pcol, bcol = pcol)
    }
  }
}
axis(1, at = seq(1.5, 18.5, 3), labels = fact)
axis(2, at = yax, labels = yax, las = 2)
mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
legend(x = 12.4, y = -19.9, legend = lcnms, pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
legend(x = 11.9, y = -19.9, legend = rep("", 4), pch = 15, adj = 0, 
       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
       cex = 0.8)
text(x = 12.7, y = -20.5, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
text(x = 12.2, y = -20.5, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
dev.off()

# Difference between 2007 and 2011 bias and accuracy statistics
baccdiff <- bacc2011[, snms, with = FALSE] -  bacc2007[, snms, with = FALSE]
setnames(baccdiff, colnames(baccdiff), mcap)

# baccdiff <- round(rbind(baccdiff, baccdiff[, lapply(.SD, mean)]), 2)
baccdiff <- cbind("Resolution" = rep(paste(fact, "km"), 2), baccdiff)
xtable::xtable(baccdiff)
```

### Other calculations methods

Across country, agricultural region only, or bin-wise (cropland density classes) mean.
```{r, eval = FALSE}
load(full_path(p_data, "lcf_out2011.rda"))
lcf <- data.table(lcf_out11)

fact <- c(1, 5, 10, 25, 50, 100)
alph <- c(225, 40)
x <- c(0, 3, 6, 9, 12, 15)
w <- 3 / 8
xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
o <- c(0, w)
xa <- sapply(x, function(x) x + xo)
cx <- c(1.25, 1, 1)
g1 <- "grey90"
reg <- c("Country", "Agricultural", "Density independent")
lcnms <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
mutype <- c("MAE", "Bias")
svec <- colnames(lcf)[grep("km", colnames(lcf))]
xl <- c(-0.5, 18)
yl <- c(-40, 40)
shd <- c(4.5, 10.5, 16.5)
cols <- c("red", "orange3", "green4", "blue")
colnms <- names(lcf)[4:length(names(lcf))]
pchs <- c("+", "o", "x")
yax <- seq(yl[1], yl[2], 5)
xax <- seq(1, 6, 5 / (length(yax) - 1))

pdf(full_path(p_fig, "cropland_bias_region.pdf"), height = 7, width = 7)
par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "")#,
for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
abline(h = yax, v = NULL, col = "grey80", lty = 1)
polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
# i <- 1; j <- 1; k <- 1
for(i in 1:length(svec)) {
  lv <- colnms[i]
  for(j in 1:length(snms)) {
    pchs1 <- pchs
    nm <- lcnms[j]
    for(k in 1:length(mutype)) {
      #dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
      dat <- lcf[Metric == mutype[k] & Map == nm, lv, with = FALSE]
      pcol <- makeTransparent(cols[j], alpha = alph[k])
      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
      for(z in 1:length(reg)) {
        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
               pch = pchs[z], type = "pt")
      }
    }
  }
}

axis(1, at = seq(1.5, 18.5, 3), labels = fact)
axis(2, at = yax, labels = yax, las = 2)
mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
legend(x = 12.4, y = -19.9, legend = lcnms, pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
legend(x = 11.9, y = -19.9, legend = rep("", 4), pch = 15, adj = 0, 
       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
       cex = 0.8)
text(x = 12.7, y = -20.5, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
text(x = 12.2, y = -20.5, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
legend(x = 12.1, y = 40, 
       legend = c("Country", "Agricultural", "Density independent"), pch = pchs, 
       pt.cex = c(1.5, 1.5), bty = "n", cex = 0.8)
dev.off()

```

### Bin density plots

To see how many values contribute to bin-wise means.
```{r, eval = FALSE}
# Using the union of agricultural areas, rather than whole country
bin_size <- do.call(rbind, lapply(lev, function(i) {
  # i <- "f50"
  st <- mubias_2011[bin != "all" & ol == i & il == "sa30" & bvals == "mu", 
                    .(bin, N)]
  st[, bin := as.integer(bin)]  # convert bin to integers
  setkey(st, bin)  
  st[data.table(bin = 1:20), N]  # merge to bin key to allow NAs
}))

pdf(full_path(p_fig, "cropland_bins.pdf"), height = 7, width = 7)
par(mfrow = c(3, 2), mar = c(2, 2, 2, 2), oma = c(3, 3, 0, 0), 
    mgp = c(1, 0.25, 0), tcl = -0.1)
for(i in 1:nrow(bin_size)) {
  barplot(bin_size[i, ], las = 2, col = "grey70", names.arg = seq(5, 100, 5))
  # axis(1, at = 1:20, labels = seq(5, 100, 5), las = 2)
  mtext(side = 3, text = paste(fact[i], "km"))
}
mtext(side = 1, text = "% Cropland", outer = TRUE)
mtext(side = 2, text = "Bin count", outer = TRUE, line = 1)
dev.off()

# check bin frequencies starting a bit higher: 2.5% cover
binv <- seq(2.5, 97.5, 5)
gti_bins <- lapply(gti_agg, function(x) {
  lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE))
})
bins_25 <- lapply(gti_bins, function(x) freq(x$g2011))
barplot(bins_25$f1[1:19, 2])
```

```{r, eval=FALSE, echo=FALSE}
DF <- do.call(rbind.data.frame, lapply(snms[4], function(x) {
  dstats_all[[2]][[x]]$f1[, c(1, 8)]
}))
dlm <- lm(mu ~ poly(V1, degree = 2), data = DF)
plot(DF[, 1], DF[, 2])
lines(predict.lm(dlm, newdata = data.frame("V1" = 1:110)))
```

# Accuracy in relation to typical crop cover

## Data
```{r, message=FALSE, results='hold', warning=FALSE}
library(raster)
library(lmisc)
library(nlme)
library(mgcv)
library(ape)
library(fitdistrplus)
library(SAcropland)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")

# Load in datasets
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data2, "ZAF_adm2.Rdata"))  # magisterial districts
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
load(full_path(p_data, "d_grid_abs.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths)  # gti 2011
gti@crs <- sa.shp@proj4string
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

# Transform magisterial districts, remove Prince Edward Islands
md <- spTransform(gadm, CRSobj = sa.shp@proj4string)
md <- md[md$ID_2 != 313, ]
```

Mean area of magisterial district.
```{r}
round(mean(rgeos::gArea(md, byid = TRUE) / 1000000))
```

## Analyses

### Extract data for magisterial districts

Both cropland percentages and absolute errors. 
```{r, eval=FALSE}
# GTI data
# mean cropland cover
gti_md <- extract(x = gti, y = md, progress = "text")
save(gti_md, file = "external/ext_data/gti_md.rda")
# load("external/ext_data/gti_md.rda")
gti_mu <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
}))

# mean no zeros
gti_mu0 <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x[x > 0], na.rm = TRUE), 1))
}))

# how many contribute when non-cropland removed, and which are non-zero
nonzero <- sapply(gti_md, function(x) length(x[!is.na(x) & (x > 0)]))
whichnonzero <- sapply(gti_md, function(x) which(!is.na(x) & (x > 0)))

# median
gti_median <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(median(x, na.rm = TRUE), 1))
}))

gti_median0 <- t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), 
    round(median(x[x > 0], na.rm = TRUE), 1))
}))

par(mar = c(0, 0, 0, 0))
plot(md[323, ])
plot(gti, add = TRUE)
plot(md[323, ], add = TRUE)

# Absolute biases at 1 km
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_abs1km <- stack(lapply(dlist_abs[snms], function(x) x$f1$g2011))
# cellStats(!is.na(dlist_abs1km), sum)
dlist_abs1km@crs <- sa.shp@proj4string
dlist_abs1km_md <- extract(x = dlist_abs1km, y = md, progress = "text")
save(dlist_abs1km_md, file = "external/ext_data/dlist_abs1km_md.rda")
# load("external/ext_data/dlist_abs1km_md.rda")

# check counts, primarily of NA values for each dataset
dlist_counts <- t(sapply(dlist_abs1km_md, function(x) {
  apply(x, 2, function(y) {
    c(length(y), length(which(is.na(y))))#, round(mean(y, na.rm = TRUE), 1))
  })
}))

chkfun <- function(x, tol = 0.001) {
  return(apply(x, 1, function(x) abs(max(x) - min(x)) < tol))
}
a <- seq(1, 7, 2)
b <- seq(2, 8, 2)
dlist_counts[which(!chkfun(dlist_counts[, a])), a]
dlist_counts[which(!chkfun(dlist_counts[, b])), b]

# bind stats
# MD
# some have more NAs than others, so select which LC set has max NAs
nas <- apply(dlist_counts[, seq(2, 8, 2)], 1, max)
all(gti_mu[, 1] == dlist_counts[, 1])

# then calculate mean per LC set per district
# mean
dlist_mu <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
  round(colMeans(x, na.rm = TRUE), 1)
})))

# mean no zero
# dlist_mu0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
#   tab <- round(x)
#   round(apply(tab, 2, function(x) mean(x[x > 0], na.rm = TRUE)), 1)
# })))
dlist_mu0 <- cbind(nas, t(sapply(1:length(dlist_abs1km_md), function(x) {
  tab <- round(dlist_abs1km_md[[x]])
  sel <- whichnonzero[[x]]  # select out reference cropland pixels only
  round(apply(tab, 2, function(x) mean(x, na.rm = TRUE)), 1)
})))

# plot(dlist_mu0[, 2], dlist_mu02[, 2])
# i <- 2
# plot(gti_mu0[, 3], dlist_mu02[, i])
# plot(gti_mu0[, 3], dlist_mu0[, i])

# median
dlist_median <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
  round(matrixStats::colMedians(x, na.rm = TRUE), 1)
})))
colnames(dlist_median) <- c("nas", snms)

# dlist_median0 <- cbind(nas, t(sapply(dlist_abs1km_md, function(x) {
#   tab <- round(x)
#   round(apply(tab, 2, function(x) median(x[x > 0], na.rm = TRUE)), 1)
# })))
dlist_median0 <- cbind(nas, t(sapply(1:length(dlist_abs1km_md), function(x) {
  tab <- round(dlist_abs1km_md[[x]])
  sel <- whichnonzero[[x]]  # select out reference cropland pixels only
  round(apply(tab, 2, function(x) median(x, na.rm = TRUE)), 1)
})))

# bind function
bind_stats <- function(x, y, coords = xy, 
                       snms = c("sa30", "globmu", "modmu", "geow")) {
  dat <- cbind.data.frame(x, y)
  dat$nd <- apply(dat[, c(2, 4)], 1, max)  # max of GTI and LC NAs
  colnames(dat)[c(1, 3)] <- c("N", "gti")
  dat <- dat[, -2]
  dat <- cbind(coords, dat)
  dat$dmu4 <- round(rowMeans(dat[, snms]), 1)  # mean across all 4
  dat$dmu3 <- round(rowMeans(dat[, snms[snms != "globmu"]]), 1) # mu - globcov
  dat
}

# centroids for districts
xy <- as.data.frame(rgeos::gCentroid(md, byid = TRUE))  # get centroid coords

# bind to gti for comparisons 
mu <- bind_stats(x = gti_mu, y = dlist_mu)
mu$w <- mu$N - mu$nd  # weight - N contributing pixels
mu0 <- bind_stats(x = gti_mu0, y = dlist_mu0)
mu0$w <- nonzero  # weight - N contributing pixels (non-zeros)
med <- bind_stats(x = gti_median, y = dlist_median)
med$w <- med$N - med$nd  # weight - N contributing pixels
med0 <- bind_stats(x = gti_median0, y = dlist_median0)
med0$w <- nonzero  # weight - N contributing pixels (non-zeros)

```

### Run regressions and check for spatial autocorrelation
```{r, eval=FALSE}
# check the polygons and rows line up for coordinates
# i <- 300
# a <- extract(x = gti, y = md[i, ], progress = "text")
# length(a[[1]])
# length(which(is.na(a[[1]])))
# mean(a[[1]], na.rm = TRUE)
# mu[i, ]

proc_df <- function(x, nm) {
  nms <- which(colnames(x) %in% c("x", "y", "gti", nm, "w"))
  dat <- x[, nms]
  colnames(dat)[3:4] <- c("pv", "dv")
  return(dat)
}

nms <- colnames(mu)[match(c(snms, "dmu4", "dmu3"), colnames(mu))]
lm_mu <- lapply(nms, function(x) {
  dat <- proc_df(x = mu, nm = x)
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})

lm_mu0 <- lapply(nms, function(x) {
  dat <- proc_df(x = mu0, nm = x)
  dat <- dat[!is.na(dat$pv), ]
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})

lm_med <- lapply(nms, function(x) {
  dat <- proc_df(x = med, nm = x)
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})

lm_med0 <- lapply(nms, function(x) {
  dat <- proc_df(x = med0, nm = x)
  dat <- dat[!is.na(dat$pv), ]
  dlm <- lm(dv ~ poly(pv, degree = 2), weights = w, data = dat)
  return(list("model" = dlm, "dat" = dat))
})
```

### Plot the fitted LM models
```{r, eval=FALSE}
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
cols <- c("red", "orange3", "green4", "blue", "black", "grey50")
lw <- c(rep(1, 4), 3, 3)
lt <- c(rep(2, 4), 1, 1)
xl <- c(0, 100)
yl <- c(0, round(max(mu[, snms] / 10 )) * 10)

# mean cropland in districts
pdf(full_path(p_fig, "biases_magdist.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- mu[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_mu)) {
  dlm <- lm_mu[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()

# mean non-zero cropland
yl <- c(0, round(max(mu0[, snms] / 10, na.rm = TRUE)) * 10)
pdf(full_path(p_fig, "biases_magdist_no0.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- mu0[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_mu0)) {
  dlm <- lm_mu0[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()

# median cropland
yl <- c(0, round(max(med[, snms] / 10, na.rm = TRUE)) * 10)
pdf(full_path(p_fig, "biases_magdist_med.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- med[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_med)) {
  dlm <- lm_med[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()

# median of non-zero cropland
yl <- c(0, round(max(med0[, snms] / 10, na.rm = TRUE)) * 10)
pdf(full_path(p_fig, "biases_magdist_med0.pdf"), width = 7, height = 7)
par(oma = c(5, 0, 0, 0), mar = c(5, 3, 2, 3))
plot(xl, yl, pch = "", ylab = "% bias", xlab = "% cropland", xaxs = "i", 
    mgp = c(1.5, 0.5, 0), yaxs = "i")
for(j in 1:length(snms)) {
  d <- med0[, c("gti", snms[j])]
  points(d[, 1], d[, 2], pch = 20, col = cols[j], cex = 0.5)
}
for(j in 1:length(lm_med0)) {
  dlm <- lm_med0[[j]]$model
  lines(predict(dlm, newdata = data.frame("pv" = 1:110)), col = cols[j], 
        lwd = lw[j], lty = lt[j])
}
legend(x = 0, y = -5, mcap, pch = 20, col = cols, bty = "n", xpd = NA,
       title = "District means")
legend(x = 80, y = -5, c(mcap, "All 4", "No GlobCover"), lty = lt, lwd = lw,
       col = cols, bty = "n", xpd = NA, title = "Fit")
dev.off()
```

### Check for spatial autocorrelation in residuals
```{r, eval=FALSE}
dists <- as.matrix(dist(mu[, c("x", "y")]))
dists_inv <- 1 / dists
diag(dists_inv) <- 0
for(i in 1:length(snms)) {
  print(Moran.I(residuals(lm_mu[[i]]$model), dists_inv)$p.value)  
}

dists <- as.matrix(dist(lm_mu0[[1]]$dat[, c("x", "y")]))
dists_inv <- 1 / dists
diag(dists_inv) <- 0
for(i in 1:length(snms)) {
  print(Moran.I(residuals(lm_mu0[[i]]$model), dists_inv)$p.value)  
} 
```

Present and significant. To deal with it, first try to use:

### A generalized least squares (GLS) with spatial autocovariance terms.

```{r, eval=FALSE}
# mean of all data
i <- "geow"  # geowiki
dat <- proc_df(x = mu, nm = i)
dgls <- gls(dv ~ poly(pv, 2), weights = ~1 / w, data = dat)  # orthogonal polys
dgls_exp <- update(dgls, correlation = corExp(form = ~x + y, nugget = TRUE))
plot(Variogram(dgls, form = ~x+y, resType = "n"))
plot(Variogram(dgls_exp, form = ~x+y, resType = "n"))  # funny shape remains

# cropland only mean
dat <- proc_df(x = mu0, nm = i)
dat <- dat[!is.na(dat$pv), ]
dgls <- gls(dv ~ poly(pv, 2), weights = ~1 / w, data = dat)  # orthogonal polys
dgls_exp <- update(dgls, correlation = corExp(form = ~x + y, nugget = TRUE))
plot(Variogram(dgls, form = ~x+y, resType = "n"))
plot(Variogram(dgls_exp, form = ~x+y, resType = "n"))  # funny shape remains
```

That seemed inadequate for removing autocorrelation in residuals. The GLS did not seem to properly deal with residual autocorrelation, probably because the spatial structure is complex, given the pattern of rainfall and cropland in it.  A better approach is to include coordinates directly in the model, leveraging the 2-D smoothing capabilities of the GAM model in the mgcv library.

### Generalized additive model

First check fit of model family, and whether assumptions of normally distributed errors are met. 
```{r, eval=FALSE}
# stats.stackexchange.com/questions/99425/distribution-for-percentage-data
# quartz()
# dat <- mu0[, "sa30"]
dat <- mu[, "sa30"]
dat <- dat[!is.na(dat)]
dat[dat == 0] <- dat[dat == 0] + 0.1
descdist(dat, discrete = FALSE)  # check potential distributions

# Check what distributions work best on data (non-cropland removed)
dists <- c("norm", "lnorm", "exp", "cauchy", "gamma", "logis", "beta","weibull")
fit_distrs <- lapply(snms, function(x) {
  dat <- mu0[, x]
  dat <- dat[!is.na(dat)] / 100
  dat[dat == 0] <- dat[dat == 0] + 0.1
  fits <- lapply(dists, function(x) fitdist(dat, x))
  names(fits) <- dists
  fits
})
names(fit_distrs) <- snms
# lapply(fit_distrs, function(x) sort(sapply(x, function(y) y$aic)))
plot(fit_distrs[[2]]$lnorm)  # lognorm shows up as best fit

# Matters more on residuals from model fit, however. 
# dataset
dat <- proc_df(x = mu, nm = snms[4])  # interactively change response  
# dat <- proc_df(x = med0, nm = snms[1])  # and input data
dat <- dat[!is.na(dat$pv), ]
dat[, 4] <- dat[, 4] + 0.1  # add 1 to allow for log transform

# GAM assuming Gaussian
dgam <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
            method = "ML")
summary(dgam)
AIC(dgam)
gam.check(dgam)
summary(fitdist(residuals(dgam), "norm"))
plot(fitdist(residuals(dgam), "norm"))

# GAM Gaussian with log link
dgam2 <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
            family = gaussian(link = "log"), method = "ML")
summary(dgam2)
AIC(dgam2)
gam.check(dgam2)
summary(fitdist(residuals(dgam2), "norm"))
plot(fitdist(residuals(dgam2), "norm"))

# GAM w/Gamma
dgam3 <- gam(dv ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
            family = Gamma(link = "identity"), method = "ML")
summary(dgam3)
AIC(dgam3)  # bizarre large AIC
gam.check(dgam3)  # crazy deviance residuals
summary(fitdist(residuals(dgam3), "norm"))
plot(fitdist(residuals(dgam3), "norm"))

# GAM w/log
dgam4 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat, 
             method = "ML")
summary(dgam4)
AIC(dgam4)
gam.check(dgam4)
summary(fitdist(residuals(dgam4), "norm"))
plot(fitdist(residuals(dgam4), "norm"))
# termplot(dgam4, terms = "poly(pv, 2)", ask = FALSE)
# pred <- predict(dgam4, newdata = newdat, type = "response", se.fit = TRUE)

# above, with smooth fit instead of quadratic
dgam4b <- gam(log(dv + 0.0001) ~ s(pv, k = 3) + s(x, y), weights = w, 
              data = dat, method = "ML")
plot(dat$pv, dat$dv)
summary(dgam4b)
AIC(dgam4b)
gam.check(dgam4)
summary(fitdist(residuals(dgam4b), "norm"))
plot(fitdist(residuals(dgam4b), "norm"))

plot.gam(dgam4b, residuals = TRUE, select = 1, se = TRUE, all.terms = TRUE, 
         pages = 0)

```

Testing each of the main landcover sets for model shape, the best fit in terms of AIC and assumption of normality of residuals is from a log-normal model.

### Plot final results

Helper functions
```{r, eval = FALSE}
predfunc <- function(model, newdat) {
  pred <- predict(model, newdata = newdat, type = "response", se.fit = TRUE)
  pdat <- transform(newdat, fit = pred$fit)
  pdat <- transform(pdat, up = fit + (1.96 * pred$se.fit),
                    lo = fit - (1.96 * pred$se.fit))
  return(pdat)
}

termfunc <- function(model, newdat) {
  pred <- predict(model, newdata = newdat, type = "terms", se.fit = TRUE)
  pdat <- transform(newdat, fit = pred$fit[, 1] + coef(model)[1])
  pdat <- transform(pdat, up = fit + (1.96 * pred$se.fit[, 1]),
                    lo = fit - (1.96 * pred$se.fit[, 1]))
  return(pdat)
}

polyfunc <- function(pdat) {
  coord <- cbind.data.frame("x" = c(rev(pdat$pv), pdat$pv), 
                            "y" = c(rev(pdat$lo), pdat$up))
}

```

#### GAM fits
```{r, eval = FALSE}
gam_mods <- lapply(nms, function(i) {
  dat1 <- proc_df(x = mu0, nm = i)
  dat1 <- dat1[!is.na(dat1$pv), ]
  dat1$dv <- dat1$dv + 0.1  # adding 0.1 to allow log transform
  dat2 <- proc_df(x = mu, nm = i)
  dat2 <- dat2[!is.na(dat2$pv), ]
  dat2$dv <- dat2$dv + 0.1  # adding 0.1 to allow log transform
  dgam1 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat1, 
               method = "ML")
  dgam2 <- gam(log(dv) ~ poly(pv, 2) + s(x, y), weights = w, data = dat2, 
               method = "ML")
  newdat <- with(dat1, 
                 data.frame(pv = 0:100, "x" = mean(dat1$x), "y" = mean(dat1$y)))
  pdat1 <- predfunc(dgam1, newdat)
  pdat2 <- predfunc(dgam2, newdat)
  # pdat1 <- termfunc(dgam1, newdat)
  # pdat2 <- termfunc(dgam2, newdat)
  return(list("mumod" = dgam1, "medmod" = dgam2, "pdat1" = pdat1, 
              "pdat2" = pdat2,  "data1" = dat1, "data2" = dat2))  
})
```

#### GAM plots
```{r, eval = FALSE}
# plot(xl, yl2, pch = "", ylab = "% bias)", xlab = "% cropland", xaxs = "i",
#      yaxs = "i", yaxt = "n")
# round(log(seq(1, 90, 10)), 2)
# exp(seq(0, 4.5, 0.5))

ptcols <- sapply(cols, function(x) makeTransparent(x, 80))  # CI colors
lw <- c(rep(2, 4), 3, 3)
yl1 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat1$fit) * 1.1)), 1))
yl2 <- c(0, round(max(sapply(gam_mods, function(x) max(x$pdat2$fit) * 1.1)), 1))
pdf(full_path(p_fig, "biases_md_lnorm_gam_mu0.pdf"), width = 7, height = 4.5)
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
plot(xl, yl1, pch = "", ylab = "Mean Absolute Error (%)", 
     xlab = "% Cropland", xaxs = "i", mgp = c(1.5, 0.5, 0), yaxs = "i", 
     yaxt = "n")
axis(2, at = 0:4, round(exp(c(-10, 1:4))))
# for(i in 1:length(gam_mods)) {
for(i in 1:4) {
  m <- gam_mods[[i]]
  p <- polyfunc(m$pdat1)
  polygon(p$x, p$y, col = ptcols[i], border = NA)
  lines(fit ~ pv, data = m$pdat1, col = cols[i], lwd = 1.5, lty = 1)
}
# legend("topleft", c(mcap, "All 4", "No GlobCover"), lty = 1, lwd = 1.5,
#        col = cols, bty = "n", xpd = NA)#, title = "Fit")
legend("topleft", mcap, lty = 1, lwd = 1.5,
       col = cols, bty = "n", xpd = NA)#, title = "Fit")
dev.off()

pdf(full_path(p_fig, "biases_md_lnorm_gam_mu.pdf"), width = 7, height = 4.5)
par(oma = c(0, 0, 0, 0), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
plot(xl, yl2, pch = "", ylab = "Mean Absolute Error (%)", 
     xlab = "% Cropland", xaxs = "i", mgp = c(1.5, 0.5, 0), yaxs = "i", 
     yaxt = "n")
axis(2, at = 0:4, round(exp(c(-10, 1:4))))
#for(i in 1:length(gam_mods)) {
for(i in 1:4) {
  m <- gam_mods[[i]]
  p <- polyfunc(m$pdat2)
  polygon(p$x, p$y, col = ptcols[i], border = NA)
  lines(fit ~ pv, data = m$pdat2, col = cols[i], lwd = 1.5, lty = 1)
}
# legend("topleft", c(mcap, "All 4", "No GlobCover"), lty = 1, lwd = 1.5,
#        col = cols, bty = "n", xpd = NA)#, title = "Fit")
legend("topleft", mcap, lty = 1, lwd = 1.5,
       col = cols, bty = "n", xpd = NA)#, title = "Fit")
dev.off()

# Check p-values for all model terms
lapply(gam_mods, function(x) {
  m1 <- summary(x$mumod)
  m2 <- summary(x$medmod)
  rbind("mu" = round(c(m1$p.table[, 4], m1$s.pv), 3), 
        "med" = round(c(m2$p.table[, 4], m2$s.pv), 3))
})  # all terms significant at p < 0.001 except 2nd order term on GlobCov med
```

#### Final plot of magisterial districts

For supplemental.
```{r, eval=FALSE}
provs <- unique(md$ID_1)
greys <- paste0("grey", seq(15, 95, 10))
pdf(full_path(p_fig, "md_map.pdf"), width = 7, height = 7, bg = "transparent")
par(mar = rep(0, 4))
plot(sa.shp, lty = 0)
for(i in 1:length(provs)) {
  plot(md[md$ID_1 == provs[i], ], col = greys[i], #border = "transparent", 
       add = TRUE, lwd = 1)
}
dev.off()

```

# Error, bias, and accuracy in carbon stock estimates
Based on the Ruesch & Gibbs (2008) approach for estimating carbon density.

## Data
```{r, eval=FALSE}
library(raster)
library(rgdal)
library(lmisc)
library(data.table)
library(SAcropland)
library(RColorBrewer)
library(xtable)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")
p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")
```

### Carbon values

Starting with the carbon look-up tables used for Africa by Ruesch & Gibbs (2008), inputs downloaded from [here](http://cdiac.ornl.gov/epubs/ndp/global_carbon/carbon_documentation.html#methods).
```{r, eval=FALSE}
cfiles <- dir(p_carb, pattern = "m6", full.names = TRUE)
nms <- gsub("*.*m6|\\.txt", "", cfiles)
cval_list <- lapply(1:length(cfiles), function(x) {
  tab <- read.table(cfiles[x])
  tab <- tab[ c(1, 3)]
  colnames(tab) <- c("CL", nms[x])
  tab
})

# merge into single carbon table
mergefun <- function(x, y) merge(x, y, by = "CL", all.x = TRUE, all.y = TRUE)
ctab <- Reduce(mergefun, cval_list)

# Read in lookup table key
key <- readLines(dir(p_carb, pattern = "key", full.names = TRUE))[25:44]
key2 <- readLines(dir(p_carb, pattern = "key", full.names = TRUE))[47:57]
mtch <- sapply(gregexpr("[0-9]", key2), max)
lcs <- sapply(1:length(key2), function(x) substr(key2[x], 1, mtch[x]))
lcs <- gsub(" & ", ",", gsub("-", ":", lcs))

# Reshape and reduce according to carbon classes
ctab2 <- t(data.frame(sapply(lcs, function(x) {
  ind <- eval(parse(text = paste0("c(", x, ")")))
  as.numeric(as.vector(
    round(colMeans(ctab[ctab$CL %in% ind, -1], na.rm = TRUE))))
})))
rownames(ctab2) <- 1:nrow(ctab2)
colnames(ctab2) <- nms
ctab2 <- ctab2 * 0.01  # convert to tons/ha from 1000 kg/ha
ctab2 <- cbind(transform(lcs), ctab2)
colnames(ctab2)[1] <- "class"
```

### Further compress classes that have the same carbon value.

  + 1:3, 6:8 => 1 (broadleaf and mixed forests)
  + 4:5 => Drop
  + 9, 10; 17 => 2 (Secondary forests, forest/cropland mosaic)
  + 11, 12, 15 => 3 (shrublands)
  + 20:23 => drop (water, snow, artificial surfaces)
  + 19 => drop (bare areas)

```{r, eval=FALSE}
ctabf <- round(rbind(colMeans(ctab2[c(1, 3), 2:ncol(ctab2)], na.rm = TRUE), 
                     colMeans(ctab2[10:11, 2:ncol(ctab2)], na.rm = TRUE), 
                     ctab2[4:6, 2:ncol(ctab2)]))

# Read in ecofloristic regions and calculate their areas for Africa
ecoflora <- readOGR("external/ext_data/africa_ecofloristic_zones.sqlite", 
                    layer = "africa_ecofloristic_zones")
ecoflora@proj4string <- CRS("+proj=longlat +datum=WGS84 +no_defs")
afalb <- paste0("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 ", 
                "+y_0=0 +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")
ecoflora_alb <- spTransform(ecoflora, CRS = CRS(afalb))
ecoflora_num <- cbind.data.frame(unique(ecoflora_alb@data),  
                                 "ID" = c("06", "08", "09", "15", "NA", "18",
                                          "20", "16", "17", "WA", "19", "07"),
                                 stringsAsFactors = FALSE)
ecoflora_alb$ID <- rep("NA", nrow(ecoflora_alb))
head(ecoflora_alb)
ecoflora_alb$ID <- ecoflora_num[match(ecoflora_alb$gez_term, 
                                      ecoflora_num$gez_term), "ID"]
# ecoflora_alb@data <- sp::merge(ecoflora_alb@data, ecoflora_num, 
#                                by = "gez_term", sort = FALSE, keep = TRUE)
ecoflora_alb@data[sample(1:nrow(ecoflora_alb@data), 2), ]
ecoflora_alb$area <- round(rgeos::gArea(ecoflora_alb, byid = TRUE) / 1000000, 1)
ecoareas <- sapply(ecoflora_num$ID, function(x) {
  sum(ecoflora_alb@data[ecoflora_alb$ID == x, "area"])
})
ecoareas <- ecoareas[!names(ecoareas) %in% c("WA", "NA")]
ecowgts <- ecoareas / sum(ecoareas)
ecowgts <- ecowgts[sort(names(ecowgts))]

# par(mar = rep(0, 4))
# plot(ecoflora_alb)
# plot(ecoflora_alb[ecoflora_alb$ID == "17", ], col = "red", add = TRUE)

i <- 1:ncol(ctabf)
ctabf$mu <- round(sapply(1:nrow(ctabf), function(x) {
  sum(ctabf[x, ] * ecowgts, na.rm = TRUE)
}), 1)
ctabf <- cbind(ctabf, t(apply(ctabf[, i], 1, function(x) range(x, na.rm=TRUE))))
colnames(ctabf)[(ncol(ctabf) - 1):ncol(ctabf)] <- c("min", "max")
rownames(ctabf) <- 1:nrow(ctabf)
LC <- c("forest", "second", "shrub", "grass", "sparse")
ctabo <- data.frame(t(ctabf[, c("mu", "min", "max")]))  # reclass table
colnames(ctabo) <- LC

```

## Analysis
### Prepare raster data
```{r, eval=FALSE}
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100) 
plot(lc_list[[1]])
# cellStats(!is.na(lc_list[[1]]), sum)
plot(gti)
# tst <- raster("external/ext_data/geowikisa_masked.tif")  # check
# plot(round(tst / 100 - lc_list[[4]], 4)) # okay

# aggregate rasters
fact <- c(5, 10, 25, 50, 100)
lc_agg <- aggregate_rast_list(fact, lc_list)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, list("gti" = gti))  # GTI rasters

# need NA mask for weighting of aggregated value
# sumna <- function(x) sum(x, na.rm = FALSE)
# namask <- calc(stack(stack(gti), stack(lclist)), sumna)
# namask[namask > 0] <- 1
namask2 <- !is.na(namask)
# namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
# namask2 <- !is.na(namask)  # set NAs to zero
area_wgts <- aggregate_rast_list(fact, list(namask2), fun = sum)

# cropland cover bins, for looking at error as a function of cover
binv <- seq(0, 1, 0.05)
gti_bins <- lapply(gti_agg, function(x) {
  cut(x$gti, breaks = binv, include.lowest = TRUE)
})

```

### Create cropland carbon estimates
```{r, eval=FALSE}
# Apply carbon estimates
# Function for pixel-wise carbon density from crop & non-crop fractions
carbon <- function(fcrop, cropC, noncropC) {
  carb <- fcrop * cropC + noncropC * (1 - fcrop)
  return(round(carb, 2))
}

cc <- ctab2[ctab2$class == 16, 2]
gti_c <- lapply(gti_agg, function(x) {
  r <- x$gti
  s <- stack(lapply(1:ncol(ctabo), function(y) {
    carbon(r, cc, ctabo[1, y])
  }))
  names(s) <- colnames(ctabo)
  s
})  # gti

lc_c <- lapply(lc_agg, function(x) {
  lc <- lapply(x, function(j) {
    s <- stack(lapply(1:ncol(ctabo), function(y) {
      carbon(j, cc, ctabo[1, y])
    }))
    names(s) <- colnames(ctabo)
    s
  })
  names(lc) <- names(x)
  lc
})  # landcover datasets

# checks--right rasters being referenced? 
tst <- cbind(sample(1:5, 10, replace = TRUE), 
             sample(1:4, 10, replace = TRUE),
             sample(1:5, 10, replace = TRUE))
sapply(1:nrow(tst), function(i) {
  cellStats(lc_c[[tst[i, 1]]][[tst[i, 2]]][[tst[i, 3]]] - 
              carbon(lc_agg[[tst[i, 1]]][[tst[i, 2]]], cc, ctabo[1, tst[i, 3]]),
              sum)
}) # should be all zeroes

```

### Difference the carbon datasets
```{r, eval=FALSE}
# percent difference
pct_diff <- function(x, y) (x - y) / x * 100
c_pct_diff <- lapply(1:length(gti_c), function(x) {
  dif <- lapply(1:length(lc_c[[x]]), function(y) {
    s <- stack(lapply(1:nlayers(lc_c[[x]][[y]]), function(z) {
      p <- pct_diff(gti_c[[x]][[z]], lc_c[[x]][[y]][[z]])
    }))
    names(s) <- colnames(ctabo)
    s
  })
  names(dif) <- names(lc_c[[x]])
  dif
})
names(c_pct_diff) <- names(lc_c)

# checks - right rasters being referenced 
tst <- cbind(sample(1:5, 10, replace = TRUE), 
             sample(1:4, 10, replace = TRUE),
             sample(1:5, 10, replace = TRUE))
sapply(1:nrow(tst), function(i) {
  x <- gti_c[[tst[i, 1]]][[tst[i, 3]]]
  y <- lc_c[[tst[i, 1]]][[tst[i, 2]]][[tst[i, 3]]]
  z <- c_pct_diff[[tst[i, 1]]][[tst[i, 2]]][[tst[i, 3]]]
  cellStats(z - ((x - y) / x * 100), sum)
})  # zeroes

# Then for a map plot figure, calculate the mean pixel-wise percent difference
c_pct_diff_mu <- lapply(c_pct_diff, function(x) {
  lapply(x, function(y) calc(y, mean))
})

# # disggregate selected rasters at selected levels for plotting
# namask <- raster(full_path(p_data, "namask.tif")) # load in NA mask
lev <- names(c_pct_diff_mu) 
disagg <- lapply(snms, function(x) {
  l1 <- lapply(lev, function(y) {
    if(y == "f1") {
      r <- c_pct_diff_mu[[y]][[x]]
    } else {
      r <- raster::disaggregate(c_pct_diff_mu[[y]][[x]], 
                                fact = as.numeric(gsub("f", "", y)))
      r <- raster::mask(crop(r, namask), namask)
    }
  })
  named_out(l1, lev)
})
names(disagg) <- snms

stats <- lapply(disagg, function(x) {
  sapply(x, function(y) {
    c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
  })
})
```

## Outputs
### Carbon error maps
```{r, eval=FALSE}
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape

# Plotting colors
lims <- c(ceiling(min(sapply(stats, function(x) x[3, ]))), 
          floor(max(sapply(stats, function(x) x[21, ]))))
rng <- range(sapply(stats, range))
brks <- c(rng[1], lims[1], -45, -20, -10, -5, -1, 1, 5, 10, lims[2], rng[2])
#n_cols <- length(brks) - 1
colsall <- brewer.pal(n = 11, "Spectral")
cols <- c(colsall[1:6], "grey80", colsall[c(7, 9, 10, 11)])
#colorRampPalette(c("red", "tan", "grey80"))
#cols2 <- colorRampPalette(c("grey80", "green4", "blue"))
#cols <- c(cols1(8), c(cols2(5)))

# brks <- c(-100.1, -75, -50, -25, -10, 0, 10, 25, 50, 75, 100.1)
# n_cols <- length(brks) - 1
legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev <- names(disagg[[1]])[-c(2:3)]
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "carbon_bias_map.pdf"), height = 6, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]], add = TRUE, col = cols, 
         breaks = brks, legend = FALSE)
  if(j == 1) mtext(mcap[i], side = 2, line = 1, cex = cx)
  if(i == 1) mtext(lev2[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext, 
            legend.vals = round(brks), 
            longdims = c(0.2, 0.8), shortdims = c(0.06, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0, -0.5)), 
            cex.val = cx, textcol = lcol, bordercol = lcol)
dev.off()
```

### Carbon bias/accuracy
#### Calculate how much country-level carbon estimates differ between datasets
```{r, eval = FALSE}
lev_vec <- names(c_pct_diff)
# created mask for non-cropland areas, unioning GTI and each LC, filtering out
# areas of no-cropland (<1/2% total cover)
lc_union <- lapply(lev_vec, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(round(gti_agg[[x]]$gti * 100) > 0)
    lc_gt0 <- Which(round(lc_agg[[x]][[y]] * 100) > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lc_union) <- lev_vec
# plot(lc_union$f50$globmu)
# plot(gti_agg$f50$gti)

# calculate for whole country
tareas <- lapply(area_wgts, function(x) x[[1]] * 100)
gti_ctot <- sapply(1:length(gti_c), function(x) {
  cellStats(gti_c[[x]] * tareas[[x]], sum)
})  # gti total carbon stock
lc_c2 <- lapply(snms, function(x) {
  sapply(names(lc_c), function(y) lc_c[[y]][[x]])
})  # reshape lc_c2 -> landcover in outer list
names(lc_c2) <- snms
lc_ctot <- lapply(lc_c2, function(x) {
  sapply(1:length(x), function(y) {
    cellStats(x[[y]] * tareas[[y]], sum)
  })
})  # landcover total carbon stocks
totc_cntry <- lapply(lc_ctot, function(x) {
  stats <- (gti_ctot - x) / gti_ctot * 100
  colnames(stats) <- names(lc_c2$sa30)
  rownames(stats) <- names(lc_c2$sa30$f1)
  stats
})  # country-level percent differences: gti versus landcover
names(totc_cntry) <- snms

# for cropped areas only - note here we are masking on cropland fraction only,
# not on union of gti and each landcover map, which is needed below for mean 
# bias estimates
tot_area <- sum(freq(namask)[1, 2])  # sum of just the non-NA area
crop_areas <- lapply(lc_union$f1, freq)  
sapply(crop_areas, function(x) x[2, 2] / tot_area)  # 29, 53, 33, 31

lc_ag2 <- lapply(snms, function(x) {
  sapply(names(lc_agg), function(y) lc_agg[[y]][[x]])
})  # reshape lc_ag2 -> landcover fractions in outer list
gti_ctot2 <- sapply(1:length(gti_c), function(x) {
  msk <- gti_agg[[x]]$gti > 0.005  # farmland > 0.05% mask
  # msk <- gti_agg[[x]]$gti > 0.05  # farmland > 0.05% mask
  msked <- mask(gti_c[[x]] * tareas[[x]], msk, maskvalue = 0)
  cellStats(msked, sum)
})  # gti 
lc_ctot2 <- lapply(1:length(lc_c2), function(x) {
  xx <- lc_c2[[x]]  # recycle reshaped lc_c2 list
  jj <- lc_ag2[[x]]  # recycle reshaped lc_c2 list
  sapply(1:length(xx), function(y) {
    msk <- jj[[y]] > 0.005  # farmland > 0.05% mask
    # msk <- jj[[y]] > 0  # farmland > 0.05% mask
    msked <- mask(xx[[y]] * tareas[[y]], msk, maskvalue = 0)
    cellStats(msked, sum)
  })
})  # landcover carbon estimates

# but discrepancy will only be relevant at 1 km scale, because the carbon total 
# keeps increasing when cropland areas are the only ones being considered
totc_crop <- lapply(lc_ctot2, function(x) {
  stats <- (gti_ctot2 - x) / gti_ctot2 * 100
  colnames(stats) <- names(lc_c2$sa30)
  rownames(stats) <- names(lc_c2$sa30$f1)
  stats
})
names(totc_crop) <- snms

# Combine tables for output to supplementals, 1 km % differences for country
# and agricultural levels
totc_out <- rbind(t(sapply(totc_cntry, function(x) x[, 1])),
                  t(sapply(totc_crop, function(x) x[, 1])))
pnms <- c("Forest", "Secondary", "Shrubland", "Grassland", "Sparse")
knms <- rep(mcap, 2)
totc_out <- cbind.data.frame(knms, round(unname(totc_out), 2))
colnames(totc_out) <- c("Map", pnms)
totc_out <- cbind(Region = c(rep("Country", 4), rep("Agricultural", 4)), 
                  totc_out)
caption <- paste("Percent differences in total carbon stock estimates", 
                 "calculated from the reference maps and from each of the four",
                 "cropland maps. Differences are evaluated for total carbon",
                 "estimates either at the country scale or over just the",
                 "agricultural regions (cropland $>$0.05\\%), using",
                 "the carbon densities of 5 different cover types to provide", 
                 "the values for the non-agricultural portions of each pixel",  
                 "(cover types indicated by column names).")
totc_xtab <- xtable(totc_out, caption = caption, digits = 1)
print(totc_xtab, type = "latex", caption.placement = "top",
      file = "paper/figures/totC-bias.tex", include.rownames = FALSE)

```

### Calculate bias/MAE statistics

#### Primary method

Weighted by __actual__ cropland percentage, and then, with aggregation, by number of pixels being aggregated.

```{r, eval=FALSE}
# Helper functions to pass into data.table
# not used - switch on if quantiles needs
# bfn <- function(x, y) {
#   box_stats(x, weighted = TRUE, weight.opts = list("weights" = y))
# }
# bfna <- function(x, y) {
#   box_stats(abs(x), weighted = TRUE, weight.opts = list("weights" = y))
# }

wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)

# Bias
cb_statsw_mu <- rbindlist(lapply(lev_vec, function(x) {
  il <- rbindlist(lapply(snms, function(y) {
    ref <- gti_agg[[x]][[1]]
    awgts <- area_wgts[[x]][[1]]
    rerror <- c_pct_diff[[x]][[y]]
    bstats <- bias_statsw(ref, awgts, rerror, LC, wm, "Bias", rnd = 2, 
                          rweight = TRUE, aweight = TRUE, trim_wgt = TRUE)
    cbind("map" = y, bstats)
  }))
  # named_out(ol, snms)
  cbind("ol" = x, il)
}))

# calculate mean across cover types
cb_statsw_mu <- cbind(cb_statsw_mu, 
                      "All" = apply(cb_statsw_mu[, LC, with = FALSE], 1, mean))

# Accuracy
cb_statsw_mua <- rbindlist(lapply(lev_vec, function(x) { # x <- "f1"
  il <- rbindlist(lapply(snms, function(y) { # y <- "sa30"
    ref <- gti_agg[[x]][[1]]
    awgts <- area_wgts[[x]][[1]]
    rerror <- c_pct_diff[[x]][[y]]
    bstats <- bias_statsw(ref, awgts, rerror, LC, wma, "MAE", rnd = 2, 
                          rweight = TRUE, aweight = TRUE, trim_wgt = TRUE)
    cbind("map" = y, bstats)
  }))
  # named_out(ol, snms)
  cbind("ol" = x, il)
}))

# calculate mean across cover types
cb_statsw_mua <- cbind(cb_statsw_mua, 
                       "All" = apply(cb_statsw_mua[, LC, with=FALSE], 1, mean))
# rowMeans(as.data.frame(cb_statsw_mua[1, ])[, LC])

# check error stats - do they match alternate approaches?
for(i in c("f1", "f25", "f10")) { # i <- "f1"
  for(j in c("sa30", "modmu", "geow")) { # j <- "modmu"
    for(k in c("forest", "shrub", "sparse", "second")) {  # k <- "second"
      print(paste("cross-checking calculations in", i, j, k))
      ref <- gti_agg[[i]][[1]]
      awgts <- area_wgts[[i]][[1]]
      rerror <- c_pct_diff[[i]][[j]]
      a1 <- bias_statsw(ref, awgts, rerror, LC, wm, "mu", 
                        aweight = FALSE)[, get(k)]
      a2 <- bias_statsw(ref, awgts, rerror, LC, wm, "mu", 
                        rweight = FALSE, aweight = FALSE)[, get(k)]
      a3 <- bias_statsw(ref, awgts, rerror, LC, wm, "mu")[, get(k)]
      c1 <- getValues(c_pct_diff[[i]][[j]][[k]])
      w1 <- getValues(gti_agg[[i]][[1]])
      w2 <- getValues(area_wgts[[i]][[1]])
      print("...non-area weighted mean matches?")
      print(a1 == round(weighted.mean(c1, w1, na.rm = TRUE), 2))
      print("...totally unweighted mean matches?")
      print(a2 == round(cellStats(rerror[[k]], mean), 2))
      print("...double weighted mean matches?")
      print(a3 == round(weighted.mean(c1, w1 * w2, na.rm = TRUE), 2))
    }
  }
}

````

#### Secondary methods

```{r, eval=FALSE}
# selection variables
nms1 <- c(colnames(ctabo), "wgt")
nms2 <- colnames(ctabo)
# bvals <- names(box_stats(sample(1:100, 200, replace = TRUE)))
bvals <- "mu"

# check constancy of < 1/2% cropland being excluded
areas <- sapply(lc_agg, function(x) res(x[[1]])[1]^2 / 10000)
plot(areas, (areas * 0.005))  # scales
(areas * 0.005)  # 1/2 to 5000 ha

dang <- Sys.time()
  #print(paste("..", x))
cb_stats <- lapply(lev_vec, function(x) {  # level
  #x <- lev_vec[4]
  print(paste(".", x))
  #levr <- names(c_pct_diff[[x]])
  l1 <- lapply(names(c_pct_diff[[x]]), function(y) {
    #y <- levr[1]
    print(paste("...", y))
    lc <- c_pct_diff[[x]][[y]]
    lcmask <- lc_union[[x]][[y]]

    # stack raster bins, cropland bins, and landcover set
    s <- stack(list("bin" = gti_bins[[x]], "wgt" = area_wgts[[x]][[1]], 
                    "mask" = lc_union[[x]][[y]], c_pct_diff[[x]][[y]]))
    DT <- na.omit(as.data.table.raster(s))
    setkey(DT, "bin")

    # Potentially useful material deleted here: check repo prior to 17/10 if
    # needed

    fr1 <- data.table("bvals" = c("m", "m0", "ma", "ma0"))
    fr2 <- data.table("bin" = "all", "N" = nrow(DT))
    binl <- DT[, .N, by = bin]  # n obs per bin
    a <- round(rbind(DT[mask == 1, lapply(.SD, wm, wgt), .SDcols = nms1], 
                     DT[, lapply(.SD, wm, wgt), .SDcols = nms1], 
                     DT[mask == 1, lapply(.SD, wma, wgt), .SDcols = nms1],
                     DT[, lapply(.SD, wma, wgt), .SDcols = nms1]), 2)
    dtl <- list(DT[mask == 1, lapply(.SD, wm, wgt), by = bin, .SDcols = nms1],
                DT[, lapply(.SD, wm, wgt), by = bin, .SDcols = nms1], 
                DT[mask == 1, lapply(.SD, wma, wgt), by = bin, .SDcols = nms1], 
                DT[, lapply(.SD, wma, wgt), by = bin, .SDcols = nms1])
    b <- rbindlist(lapply(1:4, function(x) cbind(fr1[x], round(dtl[[x]], 2))))
    setkey(b, "bin")
    odt <- rbind(cbind(fr1, rbind(cbind(fr2, a))), binl[b])
  })
  named_out(l1, names(c_pct_diff[[x]]))
})      
dut <- Sys.time() - dang  # 3.65 minutes (vs 33 in earlier incarnation), 
# 24 sec for means only
names(cb_stats) <- lev_vec
#save(cb_stats, file = "external/ext_data/carbon_bias_tables.rda")
save(cb_stats, file = "external/ext_data/carbon_bias_tables_mus.rda")

namevec <- LC
stats <- cb_stats
i1 <- "all"
i2 <- "m"

cb_stats$f25$geow[bvals == "m" & bin != "all", lapply(.SD, mean), .SDcols = LC]
cb_stats$f25$geow[bvals == "m", ]
cb_stats$f25$geow[bvals == "m" & bin != "all", lapply(.SD, mean), 
                  .SDcols = "forest"]

extract_stat0 <- function(namevec, stats, i1, i2, type = "density") {
  estats <- do.call(rbind, lapply(namevec, function(i) {
      dat <- sapply(stats, function(x) {
      vals <- unlist(sapply(x, function(y) {
        if(type == "density") {
          v <- y[bin == i1 & bvals == i2, i, with = FALSE]
        } else if(type == "nodensity") {
          v <- y[bin != i1 & bvals == i2, lapply(.SD, mean), .SDcols = i]
        }
        if(nrow(v) == 0) v <- rbindlist(list(v, as.list(NA)))
        v
      }))
    })#)
  }))
  out <- cbind(do.call(rbind, strsplit(rownames(estats), "\\.")), 
               data.table(round(estats, 2)))
  #setnames(out, old = names(mu), new = "")
  return(out)
}
N <- function(namevec, stats, i1, i2) {
  estats <- data.table(do.call(rbind, lapply(namevec, function(i) {
    o <- sapply(stats, function(x) {
      sapply(x, function(y) y[bin == i1 & bvals == i2, N][1])
    })
  })))
  return(estats)
}
```

#### Extract statistics for secondary methods

For supplementals
```{r, eval=FALSE}
# cropland area only
a <- extract_stat0(LC, cb_stats, "all", "m")  # mean cropland only 
b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(a, c(2:1, 3:ncol(a)))
mu <- rbind(a, b)  # bias

a <- extract_stat0(LC, cb_stats, "all", "ma")  # mean abs cropland only
b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(b, c(2:1, 3:ncol(b)))
mua <- rbind(a, b)  # MAE

# whole country
a <- extract_stat0(LC, cb_stats, "all", "m0")  # mean whole country
b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(b, c(2:1, 3:ncol(b)))
mu0 <- rbind(a, b)  # bias

a <- extract_stat0(LC, cb_stats, "all", "ma0")  # mean abs whole country
b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(b, c(2:1, 3:ncol(b)))
mu0a <- rbind(a, b)  # MAE

# density independent
a <- extract_stat0(LC, cb_stats, "all", "m0", type = "nodensity")
b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(a, c(2:1, 3:ncol(a)))
mund <- rbind(a, b)  # bias

# check
all(round(sapply(lev_vec, function(x) {
  mean(cb_stats[[x]]$modmu[bvals == "m0" & bin != "all", forest])
}), 2) == mund[V2 == "forest" & V1 == "modmu", lev_vec, with = FALSE])

a <- extract_stat0(LC, cb_stats, "all", "ma0", type = "nodensity")
b <- cbind(V2 = "All", a[, lapply(.SD, mean), by = V1, .SDcols = lev_vec])
setcolorder(b, c(2:1, 3:ncol(b)))
munda <- rbind(a, b)  # MAE

# check
all(round(sapply(lev_vec, function(x) {
  mean(cb_stats[[x]]$modmu[bvals == "ma0" & bin != "all", forest])
}), 2) == munda[V2 == "forest" & V1 == "modmu", lev_vec, with = FALSE])

```

### Plot mean carbon bias/MAE against scale

#### Density-weighted bias and accuracy

```{r, eval = FALSE}
cols <- c("red", "orange3", "green4", "blue")
lcnms <- c("SA-LC", "GlobCover", "MODIS", "Geowiki")
pnms <- c("Forest", "Secondary", "Shrubland", "Grassland", "Sparse")
lw <- 1.5

# yl <- range(cb_statsw_mu[, LC, with = FALSE], 
#             cb_statsw_mua[, LC, with = FALSE])
# yl <- round(yl / 10) * 10
yl <- c(-150, 150)
yax <- seq(yl[1], yl[2], 10)
xax <- seq(1, 6, 5 / (length(yax) - 1))

alph <- c(225, 40)
LC2 <- c(LC[-grep("second|grass", LC)], "All")
x <- c(0, 3, 6, 9, 12, 15)
w <- 3 / 8
xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
xa <- sapply(x, function(x) x + xo)
mutype <- c("cb_statsw_mua", "cb_statsw_mu")

o <- c(0, w)
pchs <- c("*", "+", "o", "-")
cx <- c(2, 1.25, 1, 1)
g1 <- "grey90"

# ctabo
xl <- c(-0.5, 18)
shd <- c(4.5, 10.5, 16.5)

pdf(full_path(p_fig, "carbon_veg_scalew.pdf"), height = 7, width = 7)
par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "")#,
for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
abline(h = yax, v = NULL, col = "grey80", lty = 1)
polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey50")
for(i in 1:length(lev_vec)) { # i <- 1; j <- 1; k <- 1
  for(j in 1:length(snms)) {
    for(k in 1:length(mutype)) {
      dat <- get(mutype[k])[ol == lev_vec[i] & map == snms[j], LC2, with=FALSE]
      pcol <- makeTransparent(cols[j], alpha = alph[k])
      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
      for(z in 1:length(LC2)) {
        lpfunc(x[i] + xo[j] + o[k], dat[, z, with = FALSE], col = "black", 
               size = cx[z], pch = pchs[z], type = "pt")
      }
    }
  }
}
axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
axis(2, at = yax, labels = yax, las = 2)
mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
legend(x = 12.4, y = -110, legend = lcnms, pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
legend(x = 11.9, y = -110, legend = rep("", 4), pch = 15, adj = 0, 
       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
       cex = 0.8)
text(x = 12.7, y = -115, labels = "MAE", srt = 45, adj = c(0, 0), cex = 0.8)
text(x = 12.2, y = -115, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
legend(x = 12.1, y = 120, legend = c("Forest", "Shrubland", "Sparse", "Mean"), 
       pch = pchs, pt.cex = c(2, 1.5, 1.5, 2), bty = "n", cex = 0.8)
dev.off()
```

```{r, echo=FALSE, eval=FALSE}
# Agricultural area bias and accuracy, deleted here. Values in supplmentary table
# Check repo before 17/10/15 to retrieve
```

### Supplementary bias/accuracy tables
```{r, eval=FALSE}
# Statistics from cropland area-weighted measures, reshaped
cb_stats_s <- lapply(rev(mutype), function(i) {
  i0 <- do.call(rbind, lapply(snms, function(x) {
    i1 <- do.call(rbind, lapply(c(LC, "All"), function(y) {
      i2 <- do.call(cbind, lapply(lev_vec, function(z) {
        get(i)[ol == z & map == x, get(y)]
      }))
      cbind.data.frame(y, i2)
    }))
    i1 <- cbind.data.frame(x, i1)
    colnames(i1) <- c("Map", "Cover", lev_vec)
    data.table(i1)
  }))
})
names(cb_stats_s) <- c("Bias", "Accuracy")
muws <- copy(rbind(cbind(Metric = "Bias", cb_stats_s$Bias), 
                   cbind(Metric = "MAE", cb_stats_s$Accuracy)))
setnames(muws, lev_vec, paste(c(1, fact), "km"))
setkeyv(muws, c("Map", "Cover"))
for(i in 1:length(LC)) muws[Cover == LC[i], Cover := pnms[i]]
setkey(muws, "Map")
for(i in 1:length(snms)) muws[Map == snms[i], Map := mcap[i]]

# Statistics from non-croplands areas included in means, for supplementals
# mu0[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
# mu0a[, c(lev_vec) := lapply(.SD, round), .SDcols = lev_vec]
mu0s <- copy(rbind(cbind(Metric = "Bias", mu0), 
                   cbind(Metric = "MAE", mu0a)))
setnames(mu0s, c("V1", "V2", lev_vec), 
         c("Map", "Cover", paste(c(1, fact), "km")))
setkeyv(mu0s, c("Map", "Cover"))
for(i in 1:length(LC)) mu0s[Cover == LC[i], Cover := pnms[i]]
setkey(mu0s, "Map")
for(i in 1:length(snms)) mu0s[Map == snms[i], Map := mcap[i]]

# Statistics from agricultural areas
mus <- copy(rbind(cbind(Metric = "Bias", mu), 
                  cbind(Metric = "MAE", mua)))
setnames(mus, c("V1", "V2", lev_vec), 
         c("Map", "Cover", paste(c(1, fact), "km")))
setkeyv(mus, c("Map", "Cover"))
for(i in 1:length(LC)) mus[Cover == LC[i], Cover := pnms[i]]
setkey(mus, "Map")
for(i in 1:length(snms)) mus[Map == snms[i], Map := mcap[i]]

# Bind all three tables together
fulls <- muws 
# fulls <- rbind(cbind("Region" = "Density", muws), 
#                cbind("Region" = "Country", mu0s),
#                cbind("Region" = "Agricultural", mus))

caption <- paste("Biases and mean absolute errors, weighted by reference", 
                 "cropland density, for each of the test", 
                 "maps across aggregation scales and each possible landcover", 
                 "type sharing the pixel with cropland.")
# fulls_xtab <- xtable(fulls[order(Region, Metric, Cover)], digits = 1, 
fulls_xtab <- xtable(fulls[order(Metric, Cover)], digits = 1, 
                     caption = caption)
print(fulls_xtab, type = "latex", file = "paper/figures/C-bias-accuracy.tex", 
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE) 
#       add.to.row = list(pos = list(0), 
#                         command = "\\hline \\endhead"))

```

```{r, eval=FALSE, echo=FALSE}
# Plot means against cropland density, deleted because not used. Check repo
# prior to 17/10/15 if needed. 
```

# Error, bias, and accuracy in evapotranspiration estimates

Bias in PET estimates as calculated using different version of the cropland datasets to determine the vegetation properties in VIC.

## Analysis
### Prepare datasets

Bring in cropland datasets, ET grids, reproject the latter to SA Albers, mask, convert to monthly total ET, etc. 
```{r, eval = FALSE}
library(SAcropland)
library(RColorBrewer)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
p_et <- full_path(p_root, "external/ext_data/vic/")

# cropland data
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
anms <- c("gti", snms)
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100) 
# cellStats(!is.na(lc_list$sa30), sum)

# check
# plot(lc_list[[1]])
# plot(gti)
# tst <- raster("external/ext_data/geowikisa_masked.tif")  # check
# plot(round(tst / 100 - lc_list[[4]], 4)) # okay

# aggregate rasters
fact <- 25
lc_agg <- aggregate_rast_list(fact, lc_list)   # landcover rasters 
gti_agg <- aggregate_rast_list(fact, list("gti" = gti))  # GTI rasters

# Create namask to remove all NA areas across datasets
# sumna <- function(x) sum(x, na.rm = FALSE)
# namask <- calc(stack(gti, stack(lc_list)), sumna)
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(namask2, sum)
# # namaskt <- raster(full_path(p_data, "namask.tif")) # load in NA mask
# # namaskt2 <- !is.na(namaskt)  # set NAs to zero
# cellStats(is.na(gti), sum); cellStats(namask2, sum); cellStats(namask2t, sum)
# cellStats(namaskt, sum)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts$f25, function(x) {
  r <- x[[1]]
  projection(r) <- sa.shp@proj4string
  r
})  # reshape a bit into list for subsequent use
names(awgts) <- "f25"
mask25k <- awgts$f25 > 0  # mask for et datasets

# read in ET estimates, but drop first year because of VIC problem in January
etf <- dir(p_et, pattern = "nc", full.names = TRUE)
etb <- lapply(etf, function(x) brick(x))
etb <- lapply(etb, function(x) dropLayer(x, i = 1:12))

# reproject them to Albers
etba <- lapply(etb, function(x) {
  b <- projectRaster(x, awgts$f25)
})

# rename layers and convert to total month mm
dts <- seq.Date(as.Date("1981/1/1"), as.Date("2008/12/31"), "days")
allmos <- substr(dts, 6, 7)
allyrs <- substr(dts, 1, 4)
ndays <- unlist(lapply(unique(allyrs), function(x) {
  unname(sapply(unique(allmos), function(y) {
    length(which(allyrs == x & allmos == y))
  }))
}))  # n days in each month in time series

mos <- format(ISOdatetime(2000,1:12,1,0,0,0),"%b")
lnames <- sapply(1981:2008, function(x) paste0(mos, "_", substr(x, 3, 4)))
lnames <- lnames[1:length(lnames)]
# length(lnames) == nlayers(etba[[1]])
etba <- lapply(etba, function(x) {
  b <- mask(x, mask25k, maskvalue = 0)
  names(b) <- lnames
  b
})
names(etba) <- anms

# check that mm/month multiplies through correctly
tst <- etba[[1]] * ndays
plot(tst[[15]] - etba[[1]][[15]] * ndays[15])

# convert to monthly total mm
etba <- lapply(etba, function(x) x * ndays)
# plot(etba[[1]][[1]])

# date indices
yrs <- gsub("*.*_", "", lnames)
mos <- gsub("_.*", "", lnames)
yi <- t(sapply(unique(yrs), function(x) range(which(yrs == x))))  # year index
mi <- sapply(unique(mos), function(x) which(mos == x))  # month index
```

### Process ET time series

Figure out which months produce maximum ET values, calculate monthly means, overall country-wide time series, etc. 
```{r, eval=FALSE}
# Calculate mean, median, mode of dates in time series when PET max occurs
et_max <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    which.max(x[[ind]])
  }))
  s <- stack(calc(bi, max), calc(bi, modal), calc(bi, mean), calc(bi, median))
  names(s) <- c("maxmax", "mode", "mean", "median")
  s
})

# plot to see which statistic picks out PET peak date most effectively
plot(et_max$gti)  # all the same, not coherent
plot(et_max$gti %in% 6:9)  # median looks most sensible and coherent 
plot(et_max$gti - et_max$sa30)  # some differences in date of max
plot(et_max$gti - et_max$globmu) 
plot(et_max$gti - et_max$modmu)  
plot(et_max$gti - et_max$geow)

# Monthly mean values
et_momu <- lapply(etba, function(x) {
  bi <- stack(lapply(1:ncol(mi), function(y) {
    ind <- mi[, y]
    calc(x[[ind]], mean)
  }))
  names(bi) <- unique(mos)
  bi
})

# do a focal max to separate out noisy areas
et_medsm <- focal(et_max$gti$median, w = matrix(1, 3, 3), fun = modal, pad = 3)
buf <- (!is.na(et_max$gti$median) - !is.na(et_medsm)) * et_max$gti$median
et_medsm[is.na(et_medsm)] <- 0
et_medsm <- buf + et_medsm  # fill back in areas edged out by moving window
# plot((et_medsm > 0) - (et_momu$gti[[1]] > 0))

# 1 month before and after peak
et_meds_m1 <- et_medsm - 1
# plot(et_medsm - 1)  # a few zero areas
et_meds_m1[et_meds_m1 < 1] <- 12  # set to 12 (December) if median - 1 = 0
et_meds_p1 <- et_medsm + 1
# plot(et_medsm + 1)  # nothing greater than 12

# calculate 3 month peak ET
gs_peak <- lapply(et_momu, function(x) {
  et1 <- stackSelect(x, et_meds_m1)
  et2 <- stackSelect(x, et_medsm)
  et3 <- stackSelect(x, et_meds_p1)
  stack(et1, et2, et3)
})

# Annual mean values
et_amu <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    calc(x[[ind]], sum)
  }))
  et_ts <- cellStats(bi, mean)
  names(et_ts) <- 1981:2008
  list("gridmu" = calc(bi, mean), "cntry_ts" = et_ts)
})

# plot annual mean values
rbcol <- rainbow(4)
plot(1981:2008, et_amu$gti$cntry_ts, type = "l", las = 2)
for(i in 2:5) lines(1981:2008, et_amu[[i]]$cntry_ts, col = rbcol[i])

# figure out min_max years
mm_yr <- sapply(list(which.min, which.max), function(x) x(et_amu$gti$cntry_ts))

# annual totals
et_mutot <- lapply(etba, function(x) {
  bi <- stack(lapply(1:nrow(yi), function(y) {
    ind <- yi[y, 1]:yi[y, 2]
    calc(x[[ind]], sum)
  }))
})
et_mm <- lapply(et_mutot, function(x) x[[mm_yr]])
```

### Calculate ET differences
```{r, eval=FALSE}
# difference rasters and convert to percent
# between time series annual mean
ilist <- lapply(et_amu[2:5], function(x) x[[1]])
a <- et_amu$gti$gridmu
amu_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between time series mins
ilist <- lapply(et_mm[2:5], function(x) x[[1]])
a <- et_mm$gti[[1]]
mumin_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between time series maxs
ilist <- lapply(et_mm[2:5], function(x) x[[2]])
a <- et_mm$gti[[2]]
mumax_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between mean peak growing season ET (3 month total)
ilist <- lapply(gs_peak[2:5], function(x) calc(x, sum))
a <- calc(gs_peak$gti, sum)
gs3_diff <- lapply(ilist, function(b) {
  (a - b) / a * 100
})

# between median max growing season ET (3 month total)
ilist <- lapply(gs_peak[2:5], function(x) x[[2]])
a <- gs_peak$gti[[2]]
gsmx_diff <- lapply(ilist, function(b) {
  bl <- (a - b) / a * 100
})
```

### Plot difference maps
#### Reshape and disaggregate lists
```{r, eval = FALSE}
# function to reshape lists for use in disaggregate_rast_list
disaggl <- function(rl, lev = "f25") {
  bo <- lapply(rl, function(x) {
    b <- list(x)
    names(b) <- lev
    b
  })
  bo  
}

# reshape in list
# l2dag <- list(amu_diff, mumax_diff, mumin_diff, gs3_diff, gsmx_diff)
# names(l2dag) <- c("mu", "mumx", "mumn", "gs", "gspk")
l2dag <- list(amu_diff, mumax_diff, mumin_diff, gs3_diff)
names(l2dag) <- c("mu", "mumx", "mumn", "gs")
l2dag <- lapply(l2dag, disaggl)

# disaggregate
l2dagd <- lapply(l2dag, function(x) {
  disaggregate_rast_list(snms, "f25", x, namask)
})

# stats for plotting
dstats <- function(dlist) {
  statsd <- lapply(dlist, function(x) {
    sapply(x, function(y) {
      c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
    })
  })
}
```

#### Plot maps
```{r, eval=FALSE}
# calculate single set of breakranges for all sets
statsl <- lapply(l2dag, dstats)
rng <- do.call(cbind, lapply(statsl, function(x) {
  ilims <- c(ceiling(min(sapply(x, function(y) y[3, ]))), 
             floor(max(sapply(x, function(y) y[21, ]))))
  olims <- range(sapply(x, function(y) range(y)))
  c(olims[1], ilims, olims[2])
}))
rngdt <- data.table(t(rng))
rng <- unname(unlist(rngdt[, lapply(.SD, function(x) x[which.max(abs(x))])]))

fnm <- "et_bias_map4.pdf"
legtext <- rep("% Difference", 5)
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
# ecap <- c("Annual mean", "TS max", "TS min", "Peak 3", "Peak")
ecap <- c("Annual mean", "TS max", "TS min", "Peak")

rnder <- function(x) {
  dig <- nchar(max(round(x)))
  rmat <- cbind(1:8, c(1, 5, 10, 20, 50, 100, 100, 1000))
  rnd <- rmat[dig, 2]
  round(c(floor(x[1] / rnd), ceiling(x[2] / rnd))) * rnd 
}

# For Supplementals
# brks <- c(seq(floor(rng[1]), -5, by = 5), -1, 1, seq(5, ceiling(rng[4]), 5))
brks <- c(-25, -15, -10, -5, -3, -1, 1, 3, 5, 10, 15, 25)
# cvec <- c("red", "orange", "grey80", "green", "blue4")
# cols <- colorRampPalette(cvec)(length(brks) - 1)
# cols[c(1, length(cols))] <- c("darkred", "purple3")
cols <- brewer.pal(length(brks) - 1, name = "RdBu")
cols[6] <- "grey80"
pdf(full_path(p_fig, fnm), height = 6, width = 7)
par(mfcol = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(j in 1:length(l2dagd)) {
  print(names(l2dagd)[j])
  for(k in 1:length(snms)) {
    rl <- l2dagd[[j]][[k]]$f25
    print(paste("...", snms[k]))
    plot(sa.shp, lty = 0)
    image(rl, add = TRUE, col = cols, breaks = brks, legend = FALSE)
    if(j == 1) mtext(mcap[k], side = 2, line = 1, cex = cx)
    if(k == 1) mtext(ecap[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
            legend.vals = brks, longdims = c(0.2, 0.8), 
            shortdims = c(0.06, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 4), leg.adj = list(c(0.1, 0.3), c(-0.25, -0.5)), 
            cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
dev.off()

fnm <- "et_bias_map.pdf"
pdf(full_path(p_fig, fnm), height = 6, width = 7)
par(mfcol = c(2, 2), mar = c(0, 0, 0, 0), oma = c(4, 0, 1, 0))
for(j in 1) {
  print(names(l2dagd)[j])
  for(k in 1:length(snms)) {
    rl <- l2dagd[[j]][[k]]$f25
    print(paste("...", snms[k]))
    plot(sa.shp, lty = 0)
    image(rl, add = TRUE, col = cols, breaks = brks, legend = FALSE)
    mtext(mcap[k], side = 3, line = -1, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
            legend.vals = brks, longdims = c(0.2, 0.8), 
            shortdims = c(0.07, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 4), leg.adj = list(c(0.3, 0.3), c(-0.25, -0.75)), 
            cex.val = cx - 0.4, textcol = lcol, bordercol = lcol)
dev.off()

```

### Calculate bias/accuracy statistics

#### Primary method

Density dependent method

```{r, eval=FALSE}
# weighted mean functions
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)
snms <- c("sa30", "globmu", "modmu", "geow")

# Quick look at density dependent
# reshape for ET mu
err2011 <- lapply("f25", function(x) {
  sapply(l2dag$mu, function(y) list(y[[x]]))
})
names(err2011) <- "f25"

# calculate bias/MAE for data 
a <- bias_statsw(gti_agg$f25, awgts, err2011$f25, snms, wm, "mu")
b <- bias_statsw(gti_agg$f25, awgts, err2011$f25, snms, wma, "mua")
et_muw <- rbind(a, b)

# reshape for ET mu
err2011 <- lapply("f25", function(x) {
  sapply(l2dag$gs, function(y) list(y[[x]]))
})
names(err2011) <- "f25"

# calculate bias/MAE for data
a <- bias_statsw(gti_agg$f25, awgts, err2011$f25, snms, wm, "mu")
b <- bias_statsw(gti_agg$f25, awgts, err2011$f25, snms, wma, "mua")
et_gsw <- rbind(a, b)
```

Agricultural area method
```{r, eval=FALSE}
lev_vec <- "f25"
# create mask for non-cropland areas, unioning GTI and each LC, filtering out
# areas of no-cropland (<1/2% total cover)
lcu <- lapply(lev_vec, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(round(gti_agg[[x]]$gti * 100) > 0)
    lc_gt0 <- Which(round(lc_agg[[x]][[y]] * 100) > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev_vec

# cropland cover bins, for looking at error as a function of cover
binv <- seq(0, 1, 0.05)
bins <- lapply(gti_agg[[2]], function(x) {
  cut(x[[1]], breaks = binv, include.lowest = TRUE)
})
names(bins) <- lev_vec
# bins <- cut(gti_agg$f25$gti, breaks = binv, include.lowest = TRUE)

# calculate bias stats
vnms <- names(l2dag)
bstats <- lapply(vnms, function(x) {
  a <- bias_stats_list(bins, awgts, lcu, l2dag[[x]], wm, "mu", "bias", TRUE)
  b <- bias_stats_list(bins, awgts, lcu, l2dag[[x]], wma, "mua", "bias", TRUE)
  stat <- rbind(a, b)
})
names(bstats) <- vnms

# checks to see if stats correct
# bias_stats* functions correct
chk <- bias_stats(bins$f25, awgts$f25, lcu$f25$globmu, l2dag$mu$globmu$f25, wm,
                  "mu", "bias", TRUE)
all(chk[bvals == "mu", bias] == bstats$mu[il == "globmu" & bvals == "mu", bias])
# compared to raster
for(i in snms) {
  msk <- lcu$f25[[i]]
  msk[msk == 0] <- NA
  a <- l2dag$mu[[i]]$f25 * msk
  b <- awgts$f25 * msk
  d <- weighted.mean(values(a), values(b), na.rm = TRUE)
  print(round(d, 2) == bstats$mu[il == i & bvals == "mu" & bin == "all", bias])
}  # check

# extract stats
mu <- lapply(bstats, function(x) {
  extract_stat(lev_vec, snms, "all", "mu", "bias", x)
})
mua <- lapply(bstats, function(x) {
  extract_stat(lev_vec, snms, "all", "mua", "bias", x)
})

# reshape
mus <- rbindlist(lapply(names(mu), function(x) cbind(x, mu[[x]])))
mus[, ol := NULL]
setnames(mus, c("x", "il", "bias"), c("Variable", "Map", "Bias"))
setkeyv(mus, c("Map", "Variable"))
muas <- rbindlist(lapply(names(mua), function(x) cbind(x, mua[[x]])))
muas[, ol := NULL]
setnames(muas, c("x", "il", "bias"), c("Variable", "Map", "MAE"))
setkeyv(muas, c("Map", "Variable"))
mutab <- mus[muas]
mutab[, lapply(.SD, function(x) max(abs(x))), .SDcols = 3:4]
setkey(mutab, "Variable")
vnmr <- c("Annual Mean", "29-year Max", "29-year Min", "Peak")
for(i in 1:length(vnms)) mutab[Variable == vnms[i], Variable := vnmr[i]]
setkey(mutab, "Map")
for(i in 1:length(snms)) mutab[Map == snms[i], Map := mcap[i]]
mutab[order(-rank(Map))]

caption <- paste("Biases and mean absolute errors (as \\%) for",              
                 "evapotranspiration variables derived from a 29-year time",
                 "series calculated by the VIC model, including the average", 
                 "total ET for the", 
                 "3 months of the year when ET is highest, the annual mean", 
                 "and the minimum and maximum annual ETs in the time series.")
mutab_xtab <- xtable::xtable(mutab, digits = 1, caption = caption)
print(mutab_xtab, type = "latex", file = "paper/figures/et-bias.tex",
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)
```

The largest (in an absolute sense) mean actual bias is -0.6%, while the largest mean absolute bias is 1.25%, so not much here.

### Bias in relation to rainfall

We'll end by having a look at the relationship between bias and rainfall

```{r, eval = FALSE}
# rainfall layer prepared for SA crop subsidy analysis
load("external/ext_data/climate/mag_dist_clim.rda")
pre <- calc(pre2, mean)  # 31 year mean rainfall for Africa
p4s <- projection(raster(nrow = 10, ncol = 10)) # GCS
presa <- crop(pre, spTransform(sa.shp, p4s))  # crop to SA
projection(presa) <- p4s  
presa <- projectRaster(presa, mask25k)  # project to Albers
presa <- mask(presa, awgts$f25, maskvalue = 0)  # mask

plot(presa, breaks = seq(0, 1000, 100), col = topo.colors(9))
plot(abs(l2dag$mu$globmu$f25))
plot(presa, abs(l2dag$mu$globmu$f25))

```

# Error, bias, and accuracy of gridded crop yield & production data

Based on Ramankutty et al's (2008) procedure for calculating cropland area, followed by Monfreda et al's (2008) to disaggregate yields and harvested areas.

## Data

Use SA provinces with GTI to create the amount of agricultural land estimates per province, akin to provincial statistics used by Ramankutty et al (2008). Note: they mentioned SA has 11 administrative units they used, but the publications they cited [here](http://www.statssa.gov.za/publications/DiscussLandAccounts/DiscussLandAccounts.pdf) has just 9 provinces, so maybe they mistakenly counted Lesotho and Swaziland?

```{r, eval=FALSE}
library(SAcropland)
library(xtable)
library(RColorBrewer)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
#p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")
# load(full_path(p_data, "yield-bias.rda"))

# SA provinces
prov <- readOGR("external/ext_data/provinces.sqlite", layer = "provinces")
prov <- prov[prov$id_1 != 9, ]  # remove Princeton Edward Islands
prov[prov$id_1 == 10, "id_1"] <- 9 # reset WC to id 10
load("external/ext_data/MZshapes.Rdata")

# cropland data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
lc_list <- lapply(dlist_1km, function(x) gti - x / 100) 
# cellStats(!is.na(lc_list$sa30), sum)

# plot(lc_list[[1]])
# plot(gti)
# tst <- raster("external/ext_data/geowikisa_masked.tif")  # check
# plot(round(tst / 100 - lc_list[[4]], 4)) # okay

# Create namask to remove all NA areas across datasets
sumna <- function(x) sum(x, na.rm = FALSE)
# namask <- calc(stack(stack(gti), stack(lclist)), sumna)
# namask[namask >= 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(!is.na(gti), sum); cellStats(namask2, sum)
fact <- c(5, 10, 25, 50, 100)
```

## Analyses
### Cropland fractions
#### Get provincial $cf$ factors from 1 km GTI dataset
```{r, eval=FALSE}
# rasterize provinces and stack with gti raster, convert to data.table
# provr <- rasterize(prov, y = gti, field = "id_1", 
#                    filename = "external/ext_data/provinces.tif")
provr <- raster("external/ext_data/provinces.tif")
provr <- mask(provr, namask)
gtis <- stack(list("prov" = provr, "f" = gti))

# calculate fractions using raster::extract with provinces, to compare speed
# prov_est <- extract(gti, prov, progress = "text")
# prov_estNA <- lapply(prov_est, function(x) x[!is.na(x)])
# cftest <- cbind("carea" = sapply(prov_estNA, sum), 
#                 "parea" = sapply(prov_estNA, length), 
#                 "cf" = sapply(prov_estNA, sum) / sapply(prov_estNA, length))

# data.table calculations
gti_dt <- na.omit(as.data.table.raster(gtis, xy = TRUE))
setkey(gti_dt, prov)
cf <- gti_dt[, list("carea" = sum(f), "parea" = length(f), 
                    "cf" = sum(f) / length(f)), by = prov]
# cftest - cf[, 2:4, with = FALSE]  # close, but diff caused by few NAs in provr

# data.table approach is much more efficient

```

#### Calculate cropland fractions from 1 km landcover sets

Derived at 1 km resolution rather 10 km (as in Ramankutty et al, 2008), using data.tables to calculate provincial-level correction factors for each
```{r, eval=FALSE}
lcs <- stack(list("prov" = provr, stack(lc_list)))
lcs_dt <- na.omit(as.data.table.raster(lcs, xy = TRUE))
setkey(lcs_dt, prov)
fc <- function(x) sum(x) / length(x)  # fraction functions
lc_cf <- lapply(list(sum, length, fc), function(x) {
  lcs_dt[, lapply(.SD, x), by = prov, .SDcols = snms]
})
names(lc_cf) <- colnames(cf)[-1]
# lcs_dt[, .N, by = prov] == gti_dt[, .N, by = prov]

# correction factors
pcf <- cbind("prov" = prov$id_1, 
             sapply(snms, function(x) cf$cf / lc_cf$cf[, get(x)]))
pcf <- data.table(pcf)

# tst1 <- disaggregate(tst, fact = 10)
# namask <- raster("external/ext_data/namask.tif")
# tst1m <- mask(crop(tst1, namask), mask = namask)
# plot(tst1m - gti$cover2011sum_mask)

```

#### Re-scale landcover cropland fraction using $cf$ factors
```{r, eval=FALSE}
# multiply each lc fraction value by correction factor for the particular 
# province
lcs_dta <- copy(lcs_dt)
for(j in snms) {
  for(g in pcf$prov) {
    lcs_dta[prov == g, j := (get(j) * pcf[prov == g, get(j)]), with = FALSE] 
  }
}

# check to see if DT syntax correct
for(i in 1:9) {
  for(j in snms) {
    print(all(lcs_dta[prov == i, get(j)][1:10] == 
                (lcs_dt[prov == i, get(j)] * pcf[i, get(j)])[1:10]))
  }
}

# check to see if the factors resulted in same total area per province
fchk <- lcs_dta[, lapply(.SD, sum), by = prov, .SDcols = snms]
par(mfrow = c(2, 2), mar = rep(1, 4))
for(i in snms) plot(cf[, carea], fchk[, get(i)])  # yup

# how many pixels have fraction > 1 
for(i in 1:9) {
  for(j in snms) {
    print(paste(j, ":", i, ":", 
                lcs_dta[prov == i, length(which(get(j) > 1))] / 
                  lcs_dta[prov == i, .N]))
  }
}  # several in globcover, modis, and geow

# Adjust these to equal 1 
lcs_dta2 <- copy(lcs_dta)

# function to adjust fractions to fall within 1 while maintaining statistics
force1 <- function(x) {
  a <- x
  while(any(a > 1)) {
    reall <- sum(a[a > 1] - 1)  # total of parts of x > 1, to reallocate
    ind <- which(a < 1)  # parts of x less than 1
    a[a > 1] <- 1
    af <- (reall + sum(a[a < 1])) / sum(a[a < 1])
    a[a < 1] <- a[a < 1] * af
  }
  return(a)
}
# check if works on modis set
hist(lcs_dta[prov == 8, modmu])
hist(force1(lcs_dta[prov == 8, modmu]))
round(sum(force1(lcs_dta[prov == 8, modmu])), 2) == 
  round(sum(lcs_dta[prov == 8, modmu]), 2)

for(j in snms) {
  for(g in pcf$prov) {
    lcs_dta2[prov == g, j := force1(get(j)), with = FALSE] 
  }
}

# Check to see if all fractions <= 1 and that provincial sums are equal
for(i in 1:9) {
  for(j in snms) {
    print(paste(j, ":", i, ":", 
                lcs_dta2[prov == i, length(which(get(j) > 1))] / 
                  lcs_dta[prov == i, .N], ":", 
                round(lcs_dta2[prov == i, sum(get(j))], 4) == 
                  round(lcs_dta[prov == i, sum(get(j))], 4)))
  }
}  # check

# convert back to rasters
lc_adj <- dt_to_raster(lcs_dta2, CRSobj = sa.shp@proj4string)
lc_adj <- dropLayer(lc_adj, i = 1)
```

#### Aggregate adjusted rasters, compare cropland fractions
```{r, eval=FALSE}
tl <- lapply(1:4, function(x) lc_adj[[x]])
names(tl) <- snms
lc_agg <- aggregate_rast_list(fact, tl) # landcover rasters 
rm(tl)
gtic <- crop(gti, lc_adj$sa30)
gti_agg <- aggregate_rast_list(fact, list("gti" = gtic))  # GTI rasters

# set up masks and weights for assessing output bias and absolute errors
namask <- calc(stack(gtic, lc_agg$f1), sumna)
# cellStats(!is.na(gtic), sum); cellStats(!is.na(lc_agg$f1$sa30), sum)
namask[namask >= 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(!is.na(namask), sum); cellStats(namask2, sum)
fact <- c(5, 10, 25, 50, 100)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts, function(x) x[[1]])  # unlist on inner loop

# compare extents to make sure no offsets - look at in QGIS
# writeRaster(gti_agg$f1$gti, filename = "external/ext_data/test/gti_crop.tif")
# writeRaster(lc_agg$f1$sa30, filename = "external/ext_data/test/sa30_crop.tif")
# looks fine

# tst <- gti_agg$f5$gti - lc_agg$f5$sa30
# plot(tst * 100)
#pct_diff <- function(x, y) (x - y) / x * 100

# calculate differences between adjusted fraction and reference
cf_pct_diff <- lapply(snms, function(x) {
  dif <- lapply(names(lc_agg), function(y) {
    d <- gti_agg[[y]][[1]] - lc_agg[[y]][[x]]
  })
  named_out(dif, names(lc_agg))
})
names(cf_pct_diff) <- snms

# same, but as percent
cf_pct_diff2 <- lapply(snms, function(x) {
  dif <- lapply(names(lc_agg), function(y) {
    d <- (gti_agg[[y]][[1]] - lc_agg[[y]][[x]]) * 100
  })
  named_out(dif, names(lc_agg))
})
names(cf_pct_diff2) <- snms

# cropland cover bins, for looking at error as a function of cover, based on 
# original extent of gti data, for ease
binv <- seq(0, 1, 0.05)
bins <- lapply(gti_agg, function(x) {
  cut(x$gti, breaks = binv, include.lowest = TRUE)
})

lev <- names(cf_pct_diff[[1]])
lcu <- lapply(lev, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(gti_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(lc_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev

# calculate bias/MAE for data
# calculate means
wm <- function(x, w) stats::weighted.mean(x, w)
wma <- function(x, w) stats::weighted.mean(abs(x), w)

# reshape error raster for density-weighted bias/accuracy calculations
cf_pct_resh <- lapply(lev, function(x) {
  sapply(cf_pct_diff2, function(y) list(y[[x]]))
})
names(cf_pct_resh) <- lev

# bias/accuracy
a <- bias_statsw_list(gti_agg, awgts, cf_pct_resh, snms, wm, "mu")
b <- bias_statsw_list(gti_agg, awgts, cf_pct_resh, snms, wma, "mua")
lcf_err <- rbind(a, b)

# check
# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
# check error stats - do they match alternate approaches?
for(i in c("f1", "f25", "f10")) {
  for(j in c("sa30", "modmu", "geow")) { 
    print(paste("cross-checking calculations in", i, j))
    a1 <- bias_statsw(gti_agg[[i]]$gti, awgts[[i]], cf_pct_resh[[i]], snms, wm, 
                      "mu", aweight = FALSE)[, j, with = FALSE]
    a2 <- bias_statsw(gti_agg[[i]]$gti, awgts[[i]], cf_pct_resh[[i]], snms, wm, 
                      "mu", rweight = FALSE, aweight = FALSE)[,j, with=FALSE]
    a3 <- bias_statsw(gti_agg[[i]]$gti, awgts[[i]], cf_pct_resh[[i]], snms, wm, 
                      "mu")[, j, with = FALSE]
    c1 <- getValues(cf_pct_resh[[i]][[j]])
    w1 <- getValues(gti_agg[[i]]$gti)
    w2 <- getValues(awgts[[i]])
    print("non-area weighted mean matches?")
    print(a1 == round(weighted.mean(c1, w1, na.rm = TRUE), 2))
    print("totally unweighted mean matches?")
    print(a2 == round(cellStats(cf_pct_resh[[i]][[j]], mean), 2))
    print("double weighted mean matches?")
    print(a3 == round(weighted.mean(c1, w1 * w2, na.rm = TRUE), 2))
  }
}

# Removed whole-country and agricultural area accuracy measures. See repo prior
# to 18/10 if needed. 
svec <- c("mu", "mua")
aa <- c("Bias", "Accuracy")
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")

# Reshape for output table
lcf_out <- do.call(rbind, lapply(1:length(svec), function(i) {
  a <- do.call(rbind.data.frame, lapply(snms, function(x) {
    aa <- sapply(lev, function(y) {
      lcf_err[ol == y & stnm == svec[i], x, with = FALSE]
    })
    named_out(c(x, aa), c("Map", paste(c(1, fact), "km")))
  }))
  named_out(cbind.data.frame(aa[i], a), 
            c("Metric", "Map", paste(c(1, fact), "km")))
}))

# rename landcover sets for output
lcf_out <- as.data.table(lcf_out)
setkey(lcf_out, "Map")
for(i in 1:length(snms)) lcf_out[Map == snms[i], Map := mcap[i]]

# Output table
caption <- paste("Bias and mean absolute errors (MAE) in statistically", 
                 "constrained cropland maps across aggregation scales,",
                 "weighted by density of cropland", 
                 "cover in the reference map. ")    
lcfout_xtab <- xtable(lcf_out[order(Metric)], digits = 1, caption = caption)
print(lcfout_xtab, type = "latex", 
      file = "paper/figures/cropadj-bias-accuracy.tex", 
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)

# disaggregate for plotting
# namaskc <- crop(namask, gtic)
disagg <- disaggregate_rast_list(snms, lev, cf_pct_diff, namask)

stats <- lapply(disagg, function(x) {
  sapply(x, function(y) {
    c(cellStats(y * 100, mean), quantile(y * 100, seq(0, 1, 0.05)))
  })
})
```

#### Plot for supplementary data
```{r, eval=FALSE}
lims <- c(ceiling(min(sapply(stats, function(x) x[3, ]))), 
          floor(max(sapply(stats, function(x) x[21, ]))))
rng <- range(sapply(stats, range))
brks <- c(rng[1], -50, -30, -20, -10, -5, -1, 1, 5, 10, 20, 30, 50, rng[2])
cols <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks) - 1)

legtext <- "% Difference"
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev <- names(disagg[[1]])[-c(2:3)]
lev2 <- c("1 km", "25 km", "50 km", "100 km")
pdf(full_path(p_fig, "cropland_adj_bias_map2.pdf"), height = 6, width = 7)
par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
for(i in 1:length(snms)) {
  print(snms[i])
  for(j in 1:length(lev)) {
    print(lev[j])
    plot(sa.shp, lty = 0)
    plot(disagg[[snms[i]]][[lev[j]]] * 100, add = TRUE, col = cols, 
         breaks = brks, legend = FALSE)
  if(j == 1) mtext(mcap[i], side = 2, line = 1, cex = cx)
  if(i == 1) mtext(lev2[j], side = 3, line = 0, cex = cx)
  }
}
flex_legend(ncuts = length(brks) - 1, legend.text = legtext, 
            legend.vals = round(brks), 
            longdims = c(0.2, 0.8), shortdims = c(0.06, 0.01), 
            colvec = cols, #(length(brks) - 1), 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0, -0.5)), 
            cex.val = cx, textcol = lcol, bordercol = lcol)
dev.off()
```

### Yield disaggregation

Following Monfreda et al's (2008) methods. They appear to have used the 2002 district census, so we will use the more recent 2007 census.

#### Process magisterial districts and census data
```{r, eval=FALSE}
# Load in yield dataset and magisterial district polygons
load(full_path(p_data2, "ZAF_adm2.Rdata"))  # magisterial districts
yld <- fread(full_path(p_data, "maize_wheat_2007.csv"))

# Transform magisterial districts, remove Prince Edward Islands
md <- spTransform(gadm, CRSobj = sa.shp@proj4string)
md <- md[md$ID_2 != 313, ]
md@data <- md@data[, c("PID", "ID_2", "NAME_1", "NAME_2")]  # trim down columns
# fix a few names that occur more than once to make unique for merging properly
md@data[md$ID_2 == 43, "NAME_2"] <- "MiddelburgEC"
md@data[md$ID_2 == 143, "NAME_2"] <- "RichmondKZ"
md@data[md$ID_2 == 86, "NAME_2"] <- "HeidelbergG"

# mean MD area
round(mean(rgeos::gArea(md, byid = TRUE) / 1000000))  # 3445 km square

# check names in two datasets
# match(md@data$NAME_2, yld$district)
# match(yld$district, md@data$NAME_2)
# yld$district[which(!yld$district %in% md@data$NAME_2)]
# sort(md@data$NAME_2[which(!md@data$NAME_2 %in% yld$district)])

# some offline fixing of names ensues
# write.csv(md@data, file = "external/ext_data/mdrect.csv")
# writeOGR(md, dsn = "external/ext_data/mdcheck.sqlite", layer = "mdcheck", 
#          driver = "SQLite")

# reread fixed data, merge a few districts' data where there are synonyms or 1
# district subsumed by another--sum yields, because these districts were 
# probably broken into smaller pieces
# fix same names for merging properly
yld[district == "Middelburg" & province == "Eastern Cape", 
    district := "MiddelburgEC"]
yld[district == "Richmond" & province == "KwaZulu-Natal", 
    district := "RichmondKZ"]
yld[district == "Heidelberg" & province == "Gauteng", 
    district := "HeidelbergG"]

# sumna <- function(x) sum(x, na.rm = TRUE)
mrg <- rbindlist(lapply(unique(yld$merge)[-1], function(x) {
  yld[merge == x, lapply(.SD, sumna), .SDcols = grep("mz|wh", names(yld))]
}))
nms <- sapply(unique(yld$merge)[-1], function(x) {
  nm <- yld[merge == x, district]
})
mnms <- unname(sapply(nms, function(x) x[x %in% md$NAME_2][1]))

keep <- names(yld)[-c(2:5)]
yld2 <- rbind(yld[!district %in% unname(unlist(nms)), keep, with = FALSE], 
              cbind("district" = mnms, mrg))  # adjusted yield dataset

# merge with magisterial district data
mdyld <- sp::merge(md@data, yld2, by.x = "NAME_2", by.y = "district", 
                   all.x = TRUE)  # merge
mdyld <- mdyld[match(mdyld$NAME_2, md@data$NAME_2), ]  # reorder rows correctly
nrow(mdyld) == nrow(md@data)  # check

# calculate yields and total planted ha
mdyld$mzha <- rowSums(mdyld[, grep("mz_ha", colnames(mdyld))])
mdyld$mzyld <- round(rowSums(mdyld[, grep("mz_pr", colnames(mdyld))]) /
                       mdyld$mzha, 1)
mdyld$whha <- rowSums(mdyld[, grep("wh_ha", colnames(mdyld))])
mdyld$whyld <- round(rowSums(mdyld[, grep("wh_pr", colnames(mdyld))]) / 
                       mdyld$whha, 1)

# join to md data
m <- mdyld[match(md$ID_2, mdyld$ID_2), 
           c("ID_2", "NAME_2", "mzha", "mzyld", "whha", "whyld")]
colnames(m)[1:2] <- c("id", "name")
md@data <- cbind(md@data, m)  
all(md$NAME_2 == md$name); all(md$ID_2 == md$id)
md@data <- md@data[, -c(2:4)]

# plot to make sure districts didn't get reordered at all
par(mar = rep(0, 4))
plot(sa.shp)
plot(md, add = TRUE)
plot(md[md$id == 214, ], col = "red", add = TRUE)
plot(md[md$id == 327, ], col = "red", add = TRUE)
plot(md[md$id == 263, ], col = "red", add = TRUE)
plot(md[md$name == "Barberton", ], col = "red", add = TRUE)
```

#### Calculate individual crop fractions

Rasterize MDs and calculate individual crop fractions (for maize), for both GTI and adjusted LC cropland fractions
```{r, eval = FALSE}
# mdr <- rasterize(md, provr, field = "id",  
#                  filename = "external/ext_data/mag_dist.tif", overwrite = TRUE)
mdr <- raster("external/ext_data/mag_dist.tif")
mds <- stack(list("md" = crop(mdr, gti_agg$f1$gti), "gti" = gti_agg$f1$gti,
                  lc_adj))  # use gti and lc_adj  
# v <- values((lc_agg$f1$modmu > 1) * 1); sum(v[!is.na(v)])

# data.table calculations
md_dt <- as.data.table.raster(mds, xy = TRUE)
setkey(md_dt, md)
mdarea <- md_dt[, .N, by = md]  # how many km2 in each MD
setnames(mdarea, "N", "mdkm2") 
mdarea <- mdarea[!is.na(md)]  # remove NAs  
setkey(mdarea, md)  
md_dt <- na.omit(md_dt)  # remove NAs from fraction data.tables
varea <- md_dt[, .N, by = md]  # area of non-NA data in MDs
setnames(varea, "N", "vkm2")
setkey(varea, md)
# mdcf <- md_dt[, list("carea" = round(sum(f), 2), "parea" = length(f), 
#                         "cf" = round(sum(f) / length(f), 4)), by = md]
mdva <- mdarea[varea][, fmd:= round(vkm2 / mdkm2, 2)]  # md valid area
# mdcf[, fmd := mdarea[varea][, round(vkm2 / mdkm2, 2)]]  # fraction non-NA in md

# yield data for MDs
ydt <- data.table(md@data[, c("id", "mzha", "mzyld", "whha", "whyld")])
setnames(ydt, "id", "md")
setkey(ydt, "md")
mdvay <- mdva[ydt][is.na(mzha), c("mzha", "mzyld") := 0]
mdvay[is.na(whha), c("whha", "whyld") := 0]
mdvay[, c("mdkm2", "vkm2") := NULL]
# mdcfy[, mfrac := round(((mzha * fmd) / 100) / carea, 4)]  # maize fraction

# calculate crop fractions for each dataset
md_dta <- copy(md_dt)
# md_dt[, lapply(.SD, mean), by = md, .SDcols = c("gti", snms)]
md_lcl <- lapply(c("gti", snms), function(j) {
  DT <- md_dta[, list("carea" = round(sum(get(j)), 2),  # crop area
                      "parea" = length(get(j)),  # non-NA area in MD
                      "cf" = round(sum(get(j)) / length(get(j)), 4)),
               by = md]
  # calculate crop fraction, adjusting ha first by how much non-NA area there
  # is in MD
  DTy <- DT[mdvay][, mfrac := round(((mzha * fmd) / 100) / carea, 4)]
  DTy[is.na(mfrac), mfrac := 0]  # set mfrac to 0 in 0 cropland areas
  # print(length(which(DTy$mfrac > 1)))  # 5, 4, 15, 9, 4
  #DTy[mfrac > 1, mfrac := 1]  # set to 1 - mostly missing data causing > 1
  DTy
})
names(md_lcl) <- c("gti", snms)
# lapply(md_lcl, function(x) x[, sum(mzha)])
#length(which(is.na(md_lcl$modmu$mfrac)))
#length(which(md_lcl$sa30$mfrac > 1))

par(mfrow = c(2, 3))
for(i in 1:5) hist(md_lcl[[i]]$mfrac)
for(i in 1:5) hist(md_lcl[[i]]$carea)
plot(md_lcl[[1]]$carea, md_lcl[[3]]$carea)
for(i in 1:5) print(length(which(md_lcl[[i]]$mfrac > 1)))
```

Note: when assigning crop fractions, previously we set all fractions > 1 to 1, but now turned this off because Monfreda et al (2010) allow for double-cropping. This is likely not correct, but we are doing it here to be consistent.

#### Assign back crop fraction and yield to grid 
```{r, eval=FALSE}
# first in data.tables
asnms <- c("gti", snms)
crop_ay <- lapply(c("gti", snms), function(j) {
  #j <- snms[3]
  DTya <- copy(md_dt)[, list(x, y, md, "f" = get(j))]
  DTya <- DTya[md_lcl[[j]][, list(md, fmd, mfrac, mzyld)]]
  DTya[, fa := f * mfrac]  # this is Monfreda et al equation (pg. 10)
  #print(DTya[which(round(DTya$fa, 4) > 1)[1:4], ])  # check line
  DTya[, mzya := mzyld]  # adjusted yield variable
  DTya[fa == 0, mzya := 0] # set to 0 in pixels having no crop
  DTya
})
names(crop_ay) <- asnms
sapply(crop_ay, function(x) x[, round(sum(fa), 4), by = md][, V1])

# then back to rasters for aggregation
crop_ayr <- lapply(crop_ay, function(x) {
  dt_to_raster(x[, list(x, y, fa, mzya)], CRSobj = sa.shp@proj4string)
})

# redo NA mask because of some slight shifts
namask <- calc(stack(lapply(crop_ayr, function(x) x$fa)), sumna)
namask[namask >= 0] <- 1
namask2 <- !is.na(namask)  # set NAs to zero
# cellStats(!is.na(gti), sum); cellStats(namask2, sum)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts, function(x) x[[1]])  # unlist on inner loop
# sumna <- function(x, na.rm = na.rm) sum(x, na.rm = na.rm)
# amsk <- aggregate_rast_list(fact, list(namask2), fun = sumna)
# amsk <- lapply(amsk, function(x) x[[1]])  # unlist on inner loop
```

#### Aggregate yields and crop area to coarser resolutions
```{r, eval=FALSE}
# Yield aggregation - aggregating with weighting by crop fraction
# using a custom data.table function to allow this 
lev <- names(disagg[[1]])
wmfun <- parse(text = "sum((mzya * fa) / sum(fa, na.rm = TRUE), na.rm = TRUE)")
sr <- sa.shp@proj4string
d <- c(dim(gti_agg[[1]]$gti)[1:2], xres(gti_agg[[1]]$gti))  # dimensions
cyld <- lapply(lev[-1], function(i) {
  print(i)
  f <- as.integer(gsub("f", "", i))
  il <- lapply(c("gti", snms), function(j) {
    print(paste("...", j))
    rdt <- as.data.table(crop_ayr[[j]], xy = TRUE)
    o <- dt_aggregate(rdt, d[1], d[2], f, wmfun, d[3], "yw")
  })
  named_out(il, asnms)
})
names(cyld) <- lev[-1]

# convert back to rasters
cyldr <- lapply(cyld, function(i) lapply(i, function(j) dt_to_raster(j, sr)))
cyldf1 <- lapply("f1", function(i) {
  f1yld <- lapply(asnms, function(j) crop_ayr[[j]]$mzya)
  named_out(f1yld, asnms)
})
names(cyldf1) <- lev[1]
cyld_agg <- c(cyldf1, cyldr)
# plot(cyld_agg$f10$gti)

# check Monfreda's yield data to see how it deals with null areas
monf <- full_path("external/ext_data/monfreda_data/maize_HarvAreaYield_NetCDF",
                  "maize_AreaYieldProduction.nc")
monfmz <- brick(monf, level = 2)
plot(crop(monfmz, spTransform(sa.shp, monfmz@crs)), col = bpy.colors(20))
plot(crop(monfmz, spTransform(sa.shp, monfmz@crs)) > 0)  # almost all SA > 0

# Removed other versions of yield aggregation. Refer to repo prior to 18/10 to 
# to recover

## crop areas
carea <- lapply(1:5, function(x) crop_ayr[[x]]$fa)
names(carea) <- c("gti", snms)

# sumha <- function(x, na.rm) sum(x, na.rm)
carea_agg <- aggregate_rast_list(fact, carea) # crop area, mean aggregation
#carea_agg2 <- aggregate_rast_list(fact, carea, sum) # crop area, sum agg
# plot(carea_agg$f25$globmu)
# plot(carea_agg2$f25$globmu)

# area of each pixel at each aggregation scale
agg_res <- sapply(carea_agg, function(x) res(x$gti)[1]^2 / 10000)  
tareas <- lapply(awgts, function(x) x[[1]] * 100)

# plot(tareas$f50)
#plot((carea_agg$f50$gti * tareas$f50) - (carea_agg2$f50$gti * 100))  # equiv
# ta <- rep(1, 50 * 50)
# a <- runif(50 * 50, 0, 1)
# (sum(a) * 100) / (sum(ta) * 100)
# mean(a) * (sum(ta) * 100)

# Removed yat check. Refer to repo prior to 18/10 to recover

# check aggregated maize fractions
for(i in lev) {
  for(j in snms) {
    ck <- crop_ay[[j]][, sum(fa)] * 100
    d <- round((ck - cellStats(tareas[[i]] * carea_agg[[i]][[j]], sum)) / ck,2)
    print(d)
  }
}  # all equal
```

#### Crop production estimates at different aggregation scales

Including differences relative to reference versions.
```{r, eval=FALSE}
# Type A aggregation: multiplying aggregated yields by aggregated fractions
# with crop-fraction weighted aggregate yields
cpagg1 <- lapply(1:length(cyld_agg), function(x) {
  p <- lapply(1:length(cyld_agg[[x]]), function(y) {
    (tareas[[x]] * carea_agg[[x]][[y]]) * cyld_agg[[x]][[y]]
  })
  named_out(p, asnms)
})
names(cpagg1) <- names(agg_res)

# Removed production aggregation with 0-removed yieldsother. Refer to repo prior
# to 18/10 to recover

# Type B (formerly II) aggregation: crop production differences when aggregated
# from 1 km calculate production at 1 km
cprod <- lapply(1:length(cyldf1$f1), function(x) {
  cyldf1$f1[[x]] * (carea[[x]] * 100)
})
names(cprod) <- asnms
cpaggII <- aggregate_rast_list(fact, cprod, "sum") # aggregate production

# check production estimates for consistency
# pat <- list(cpagg1, cpagg3, cpaggII)
# ptype <- c("1", "3", "II")
pat <- list(cpagg1, cpaggII)
ptype <- c("1", "II")

# check production estimates
for(i in lev[-1]) {
  print(paste("factor", i))
  for(j in asnms) {
    print(paste("...", j))
    for(k in 1:length(ptype)) {
      print(paste("...... type", ptype[k], "agg"))
      ck <- cellStats(cprod[[j]], sum)  # weighted by fraction
      ck2 <- cellStats(cprod$gti, sum) # weighted by fraction
      v <- cellStats(pat[[k]][[i]][[j]], sum)
      d <- round((ck - v) / ck, 2)
      d2 <- round((ck2 - v) / ck2, 2)
      print(c(d, d2))
    }
  }  # all types consistent across scales
}

# sapply(cprod, function(x) cellStats(x, sum))
# sapply(carea, function(x) cellStats(x, sum))

DT <- na.omit(as.data.table.raster(s, xy = TRUE))
DT[, wgt := ref * awgts]  # calculate weights from ref and area weights

# Mean production in producing cells at different levels of aggregation. 
# No need to weight by cropland density, because production value already factors
# in area - just mask out zero production areas
mup <- sapply(lev, function(i) {
  msk <- cpagg1[[i]]$gti > 0
  msk[msk == 0] <- NA
  cpmsk <- msk * cpagg1[[i]]$gti
  cellStats(cpmsk, mean)
})

# mean yield in producing cells at different levels of aggregation, use density
# weighting here
muy <- sapply(lev, function(i) {
  # i <- lev[2]; print(i)
  s <- stack(cyld_agg[[i]]$gti, carea_agg[[i]]$gti, awgts[[i]])
  names(s) <- c("yld", "ref", "awgts")
  DT <- na.omit(as.data.table.raster(s, xy = TRUE))
  DT[, wgt := ref * awgts]  # calculate weights from ref and area weights
  DT[, weighted.mean(yld, wgt)]
})  # yield is 3.361 kg/ha at all scales

# Removed original mup and muy. Refer to repo prior to 18/10 to recover

# calculate differences between them
# check first that cpagg1 and II are equivalent, make cpaggII master
for(i in lev) {
  for(j in snms) {
    a <- (cpagg1[[i]]$gti - cpagg1[[i]][[j]]) - 
      (cpaggII[[i]]$gti - cpagg1[[i]][[j]])
    print(round(cellStats(a, sum), 5))
  }
}

# Production differences
# first standardize
# std_raster <- function(x) {
#   (x - cellStats(x, min)) / diff(cellStats(x, range))
# }
# cpagg_std <- lapply(cpagg1, function(x) lapply(x, function(y) std_raster(y)))
ilist <- list(names(cpagg1[[1]])[-1], names(cpagg1), "gti")
# list1 <- lapply(cpagg1, function(x) x[1])
# list2 <- lapply(cpagg1, function(x) x[-1])
list1 <- lapply(cpagg1, function(x) x[1])
list2 <- lapply(cpagg1, function(x) x[-1])
p_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1 agg

# Removed original p_diff2/3. Refer to repo prior to 18/10 to recover

# yield differences
list1 <- lapply(cyld_agg, function(x) x[1])
list2 <- lapply(cyld_agg, function(x) x[-1])
y_diff1 <- rast_list_math(ilist, list1, list2, expr = "a - b")  # type 1

# check grids (output in QGIS and examine for sensible results--they are)
# for(i in lev) {
#   writeRaster(y_diff1$globmu[[i]]$gti, 
#               file = full_path(p_data, paste0("test/globdiff", i, ".tif")))
# }
# for(i in lev) {
#   writeRaster(list1[[i]]$gti, 
#               file = full_path(p_data, paste0("test/gtiyldtst", i, ".tif")))
# }
# for(i in lev) {
#   writeRaster(list2[[i]]$globmu, 
#               file = full_path(p_data, paste0("test/globyld2tst", i, ".tif")))
# }

# Removed original y_diff2. Refer to repo prior to 18/10 to recover

# convert them to percent differences relative to mean pixel-wise production
# here only for map plotting.

pdiffs <- lapply(list(p_diff1), function(x) {
  l1 <- lapply(snms, function(y) {
    l2 <- lapply(lev, function(z) {
      x[[y]][[z]]$gti / mup[z] * 100
    })
    named_out(l2, lev)
  })  
  named_out(l1, snms)
})
names(pdiffs) <- "pd1"

ydiffs <- lapply(list(y_diff1), function(x) {
  l1 <- lapply(snms, function(y) {
    l2 <- lapply(lev, function(z) {
      x[[y]][[z]]$gti / muy[z] * 100
    })
    named_out(l2, lev)
  })  
  named_out(l1, snms)
})
names(ydiffs) <- "yd1"

# disaggregate for plotting
lev <- names(y_diff1[[1]])
p_diff1d <- disaggregate_rast_list(snms, lev, pdiffs$pd1, namask)
y_diff1d <- disaggregate_rast_list(snms, lev, ydiffs$yd1, namask)

# stats for plotting
dstats <- function(dlist) {
  statsd <- lapply(dlist, function(x) {
    sapply(x, function(y) {
      c(cellStats(y, mean), quantile(y, seq(0, 1, 0.05)))
    })
  })
}
stats_pd1 <- dstats(p_diff1d)
stats_yd1 <- dstats(y_diff1d)
```

### Plot difference maps
```{r, eval=FALSE}
dnms <- ls()[grep("[y_|p_]diff*.*d", ls())]
stnms <- ls()[grep("stats_", ls())]
disaggl <- lapply(dnms, function(x) get(x))
statsl <- lapply(stnms, function(x) get(x))
rng <- lapply(statsl, function(x) {
  ilims <- c(ceiling(min(sapply(x, function(y) y[3, ]))), 
             floor(max(sapply(x, function(y) y[21, ]))))
  olims <- range(sapply(x, function(y) range(y)))
  c(olims[1], ilims, olims[2])
})

fnms <- c("prod_bias_map.pdf", "yld_bias_map.pdf")  # output names
legtext <- rep("% Difference", 5)
cx <- 1.4
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
lev <- names(p_diff1d[[1]])[-c(2:3)]
lev2 <- c("1 km", "25 km", "50 km", "100 km")

rnder <- function(x) {
  dig <- nchar(max(round(x)))
  rmat <- cbind(1:8, c(1, 5, 10, 20, 50, 100, 100, 1000))
  rnd <- rmat[dig, 2]
  round(c(floor(x[1] / rnd), ceiling(x[2] / rnd))) * rnd 
}
div_breaks <- function(olim, ilim, ibrk) {
  il <- rnder(ilim) 
  ol <- c(floor(olim[1]), ceiling(olim[2]))
  brks <- c(ol[1], seq(il[1], -ibrk, abs(floor(il[1]) - -ibrk) / 4), 
            seq(ibrk, il[2], abs(ceiling(il[2]) - ibrk) / 4), ol[2])
  brks
}

rng2 <- lapply(rng, function(x) {
  x[2:3] <- c(-100, 100)
  x
})
brksl <- lapply(rng2, function(x) {
  c(floor(x[1:2]), c(-50, -25, -10, -5, -1, 1, 5, 10, 25, 50), ceiling(x[3:4]))
})
# brksl <- list(c(-1, -0.75, -0.5, -0.25, -0.1, -0.05, -0.01, 0.01, 0.05, 0.1, 
#                 0.25, 0.5, 0.75, 1), 
#               c(-10.7, -7, -5, -3, -2, -1, -0.1, 0.1, 1, 2, 3, 5, 7, 10.7))

for(i in 1:length(statsl)) {
  # i <- 2
  print(paste("figure", i, "of", length(statsl)))
  brks <- brksl[[i]]
  cols <- c(rev(brewer.pal(length(brks) / 2, "Reds")[-1]), "grey80", 
          brewer.pal(length(brks) / 2, "Blues")[-1])
  brklen <- length(brks) - 1
  pdf(full_path(p_fig, fnms[i]), height = 6, width = 7)
  par(mfrow = c(4, 4), mar = c(0, 0, 0, 0), oma = c(5, 5, 2, 0))
  for(j in 1:length(snms)) {
    print(snms[j])
    for(k in 1:length(lev)) {
      rl <- disaggl[[i]]
      print(lev[k])
      plot(sa.shp, lty = 0)
      plot(rl[[snms[j]]][[lev[k]]], add = TRUE, col = cols, 
                breaks = brks, legend = FALSE)
      if(k == 1) mtext(mcap[j], side = 2, line = 1, cex = cx)
      if(j == 1) mtext(lev2[k], side = 3, line = 0, cex = cx)
    }
  }
  flex_legend(ncuts = length(brks) - 1, legend.text = legtext[i], 
              legend.vals = brks, longdims = c(0.15, 0.8), 
              shortdims = c(0.07, 0.01), 
              colvec = cols, #(length(brks) - 1), 
              srt = c(270, 0), horiz = TRUE, textside = "bottom", 
              legend.pos = c(4, 5), leg.adj = list(c(0.25, 0), c(0, -0.5)),
              cex.val = cx, textcol = lcol, bordercol = lcol)
  dev.off()
}

lev <- names(p_diff1d[[1]])  # reset levs
```

Several versions of yield aggregation were tested in this analysis, each of which has a different impact on yield and production impacts. There are three types of yield aggregation that can be done (most commented out now after initial checking):

  1. __Type 1__: weighted mean averaging, with the weights provided by crop fractions. This produces correct country-level average yields (<1-2% different) across scales and between datasets, provided that the averaging is weighted by the cropland fractions that were aggregated to the level from which the country-level yield estimate is being calculated

  2. __Type 2__: Straight averaging, which results in diluted country-level yield estimates. Incorrect, and should be obvious that this wouldn't be done, so we haven't pursued it here.

  3. __Type 3__: or zero-removed averaging, where only areas having fraction of that crop contribute to the aggregated average. As with __Type 1__, this produces correct country level estimates when weighting by cropland fraction, but it does not produce correct production estimates, probably because it inflates the value of high-yielding irrigated areas out towards the dry west, where the crop fractions are small. (see below)

There are also two ways of aggregating production estimates:

  1.  __Type A__: Aggregating yield and crop area separately, then multiplying. This is correct with __Type 1__ yield aggregation, and producing identical production estimates to __Type II__, but not with the other two types of yield aggregation.

  2.  __Type B__: Calculating production at the base resolution, then aggregating by sum. This should be the correct way of doing it.

Why would someone do a yield aggregation, when you have the gridded estimates?  To calculate values and compare them to a coarser resolution product, for instance, or to get a country-level average.

Causes for bias concern to examine:

  + Bias in yield gap estimates, and how much closing that could affect contribution of closing gaps to boosting production.

  + Country-level statistics might cancel out bias, but if bias of one type of sign is spatially correlated with the driver of values in yield gaps, then that would bias larger-scale estimates of gap closure potential.

  + These concerns apply to all resolutions.

Type 2 and 3 yield aggregation are incorrect methods, and are not retained here (but see earlier commits of code, specifically prior to 18/10/2015. The other variants were used for comparison and for initial methods checks only.

## Bias/MAE statistics

### Primary method

Density weighted, following methods developed in [cropland-bias](croland-bias.html), but here metrics are based on residuals rather than percent errors, because of infinite values (harvested area/yield disaggregation methods mean that there are gridded yields/production estimates in some areas where no reference value exists, leading to infinite values if converting to percent)
```{r, eval=FALSE}

# yield bias/accuracy
# reshape
yerr <- lapply(lev, function(x) {
  sapply(y_diff1, function(y) list(y[[x]]$gti))
})
names(yerr) <- lev

# carea_gti <- lapply(carea_agg, function(x) x$gti)
a <- bias_statsw_list(gti_agg, awgts, yerr, snms, wm, "mu")
b <- bias_statsw_list(gti_agg, awgts, yerr, snms, wma, "mua")
yld_bacc <- rbind(a, b)

# production bias/accuracy
# reshape
perr <- lapply(lev, function(x) {
  sapply(p_diff1, function(y) list(y[[x]]$gti))
})
names(perr) <- lev

# carea_gti <- lapply(carea_agg, function(x) x$gti)
a <- bias_statsw_list(gti_agg, awgts, perr, snms, wm, "mu")
b <- bias_statsw_list(gti_agg, awgts, perr, snms, wma, "mua")
prod_bacc <- rbind(a, b)

# check error stats - do they match alternate approach?
ref <- lapply(gti_agg, function(x) x$gti)
test_err <- "perr" #"yerr"
evst <- "wma" # "wm"
for(i in c("f1", "f25", "f10")) {
  for(j in c("sa30", "modmu", "geow")) { 
    print(paste("cross-checking calculations in", i, j))
    a1 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms,  
                      get(evst), aweight = FALSE)[, j, with = FALSE]
    a2 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms, 
                      get(evst), rweight = FALSE, 
                      aweight = FALSE)[, j, with = FALSE]
    a3 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms, 
                      get(evst))[, j, with = FALSE]
    s <- stack(get(test_err)[[i]][[j]], ref[[i]], awgts[[i]])
    sv <- getValues(s)
    sv <- sv[which(!is.na(rowSums(sv))), ]  # remove NAs
    if(evst == "wm") {
      print(a1 == round(weighted.mean(sv[, 1], sv[, 2], na.rm = TRUE), 2))
      print(a2 == round(mean(sv[, 1]), 2))
      print(a3 == round(weighted.mean(sv[, 1], sv[, 2] * sv[, 3]), 2))
    } else if(evst == "wma"){
      print(a1 == round(weighted.mean(abs(sv[, 1]), sv[, 2], na.rm = TRUE), 2))
      print(a2 == round(mean(abs(sv[, 1])), 2))
      print(a3 == round(weighted.mean(abs(sv[, 1]), sv[, 2] * sv[, 3]), 2))
    }
  }
}

```

### Secondary method

Bias and accuracy within agricultural areas
```{r, eval=FALSE}
# create mask for non-cropland areas, unioning GTI and LC areas with maize
# estimates -- probably not needed if doing production estimates.  
lev <- names(y_diff1[[1]])

# first check aggregation to make yield aggregation variants have common area,
# so that just one lc_union is needed
# for(i in lev) {
#   for(j in snms) {
#     print(cellStats(cyld_agg[[i]][[j]] > 0 & cyld_agg3[[i]][[j]] == 0, sum))
#     #print(cellStats(carea_agg[[i]][[j]] > 0 & cyld_agg[[i]][[j]] == 0, sum))
#   }
# }  # all zeros

snms <- names(p_diff1)
lcu <- lapply(lev, function(x) {
  lcb <- lapply(snms, function(y) {
    gti_gt0 <- Which(carea_agg[[x]][[1]] > 0)
    lc_gt0 <- Which(carea_agg[[x]][[y]] > 0)
    all_gt0 <- gti_gt0 + lc_gt0
    all_gt0[all_gt0 > 0] <- 1
    all_gt0
  })
  named_out(lcb, snms)
})
names(lcu) <- lev

# cropland cover bins, for looking at error as a function of cover
binv <- seq(0, 1, 0.05)
bins <- lapply(gti_agg, function(x) {
  cut(x$gti, breaks = binv, include.lowest = TRUE)
})

# Many commented out checks removed here. Look at pre 18/10 commits to restore

# Spatial bias stats calculation on yield and production differences, not on
# percentages
# Many stats for alternate yield/production metrics removed here. 
# Look at pre 18/10 commits to restore
a <- bias_stats_list(bins, awgts, lcu, p_diff1, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, p_diff1, wma, "mua", "bias", TRUE)
d <- bias_stats_list(bins, awgts, lcu, p_diff1, sum, "sum", "bias", FALSE)
pd1_st <- rbind(a, b, d)

a <- bias_stats_list(bins, awgts, lcu, y_diff1, wm, "mu", "bias", TRUE)
b <- bias_stats_list(bins, awgts, lcu, y_diff1, wma, "mua", "bias", TRUE)
yd1_st <- rbind(a, b)

# Output statistics
p1mu <- extract_stat(lev, snms, "all", "mu", "bias", pd1_st)  # identical to p3
p1ma <- extract_stat(lev, snms, "all", "mua", "bias", pd1_st)  # ident to p3
y1mu <- extract_stat(lev, snms, "all", "mu", "bias", yd1_st)
y1ma <- extract_stat(lev, snms, "all", "mua", "bias", yd1_st)

# Older variant of code from compare-landcover.Rmd to evaluate whether newer DT
# version is finding correct results
i <- y_diff1 #pdiff_3 # p_diff1
jdt <- yd1_st # pd3_st # pd1_st
bv <- "mua" # bv <- "mu"
for(chk in list(c("modmu", "f25"), c("geow", "f10"), c("sa30", "f50"), 
                c("globmu", "f1"))) {
  x <- chk[1] 
  y <- chk[2] 
  print(paste(".", x, "..", y))
  rs <- lcu[[y]][[x]] 
  rs[rs == 0] <- NA
  l3 <- abs(i[[x]][[y]][[1]])
  # l3 <- i[[x]][[y]][[1]]
  o <- rs * l3
  w <- awgts[[y]][[1]] * rs
  wmu <- weighted.mean(getValues(o), getValues(w), na.rm = TRUE)
  print(jdt[ol == y & il == x & bvals == bv & bin == "all", bias] ==
          round(wmu, 2))
}  # check

```

### Bias/MAE plots

#### Density-weighted
```{r, eval = FALSE}
alph <- c(225, 60)
x <- c(0, 3, 6, 9, 12, 15)
w <- 3 / 8
xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
o <- c(0, w)
xa <- sapply(x, function(x) x + xo)
cx <- c(1.25, 1)
g1 <- "grey90"

# get stats data into shape
mulw <- list(prod_bacc[stnm == "mu"], yld_bacc[stnm == "mu"])
malw <- list(prod_bacc[stnm == "mua"], yld_bacc[stnm == "mua"])
malmulw <- list(malw, mulw)  # density-weighted stats, reordered
pyl <- c("mup", "muy")

xl <- c(-0.5, 18)
yl <- c(-30, 110)  # try it out, play with it 
shd <- c(4.5, 10.5, 16.5)
cols <- c("red", "orange3", "green4", "blue")
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
pchs <- c("+", "o")

yax <- seq(yl[1], yl[2], 10)
xax <- seq(1, 6, 5 / (length(yax) - 1))

pdf(full_path(p_fig, "yield_prod_bias.pdf"), height = 7, width = 7)
par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "")#,
for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
abline(h = yax, v = NULL, col = "grey80", lty = 1)
polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
for(ii in 1:length(lev)) { # ii <- 1; i <- 1; k <- 1
  lv <- lev[ii]
  for(i in 1:length(snms)) {
    pchs1 <- pchs
    nm <- snms[i]
    for(k in 1:length(malmulw)) {
      mm <- malmulw[[k]]
      # fetch bias stats, convert to percentage relative to mean
      # e.g. GTI mean yield
      v <- sapply(1:length(mm), function(j) {  # j <- 1
        round(mm[[j]][ol == lv, get(nm)] / get(pyl[j])[lv] * 100, 1)
      })
      pcol <- makeTransparent(cols[i], alpha = alph[k])
      polyfunc2(xa[i, ii] + o[k], range(v), w = w, col = pcol, bcol = pcol)
      xx <- rep(x[ii] + xo[i] + o[k], length(v))
      points(xx, v, pch = pchs, cex = cx)
    }
  }
}
axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
axis(2, at = yax, labels = yax, las = 2)
mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
legend(x = 12.4, y = -10, legend = lcnms, pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
legend(x = 11.9, y = -10, legend = rep("", 4), pch = 15, adj = 0, 
       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
       cex = 0.8)
text(x = 12.7, y = -12, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
text(x = 12.2, y = -12, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
legend(x = 12.1, y = 100, legend = c("Production", "Yield"), pch = pchs, 
       pt.cex = c(1.5, 1.5), bty = "n", cex = 0.8)
dev.off()
```

#### Agricultural area

Supplemental
```{r, eval=FALSE}
alph <- c(225, 60)

mula <- c("p1mu", "y1mu")
mala <- c("p1ma", "y1ma")
malmula <- list(mala, mula)
pyl <- c("mup", "muy")

yl <- c(-70, 80)
yax <- seq(yl[1], yl[2], 10)

pdf(full_path(p_fig, "yield_prod_bias_agric.pdf"), height = 7, width = 7)
par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "")
for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
abline(h = yax, v = NULL, col = "grey80", lty = 1)
polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
for(ii in 1:length(lev)) {
  lv <- lev[ii]
  for(i in 1:length(snms)) {
    pchs1 <- pchs
    nm <- snms[i]
    for(k in 1:length(malmula)) {
      mm <- malmula[[k]]
      # fetch bias stats, convert to percentage relative to mean
      # e.g. GTI mean yield
      v <- sapply(1:length(mm), function(j) {
        round(get(mm[j])[ol == lv & il == nm,  bias] / get(pyl[j])[lv] * 100,1)
      })
      pcol <- makeTransparent(cols[i], alpha = alph[k])
      polyfunc2(xa[i, ii] + o[k], range(v), w = w, col = pcol, bcol = pcol)
      xx <- rep(x[ii] + xo[i] + o[k], length(v))
      points(xx, v, pch = pchs, cex = cx)
    }
  }
}
axis(1, at = seq(1.5, 18.5, 3), labels = c(1, fact))
axis(2, at = yax, labels = yax, las = 2)
mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
legend(x = 12.4, y = -40, legend = lcnms, pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
legend(x = 11.9, y = -40, legend = rep("", 4), pch = 15, adj = 0, 
       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
       cex = 0.8)
text(x = 12.7, y = -42, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
text(x = 12.2, y = -42, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
legend(x = 12.1, y = 65, legend = c("Production", "Yield"), pch = pchs, 
       pt.cex = c(1.5, 1.5), bty = "n", cex = 0.8)

dev.off()

lapply(1:length(lev), function(ii) {
  lv <- lev[ii]
  lapply(1:length(snms), function(i) {
    nm <- snms[i]
    lapply(1:length(malmula), function(k) {
      mm <- malmula[[k]]
      v <- sapply(1:length(mm), function(j) {
        round(get(mm[j])[ol == lv & il == nm,  bias] / get(pyl[j])[lv] * 100,1)
      })
    })
  })
})

do.call(cbind, lapply(lev, function(x) {
  rbind(cbind.data.frame("Map" = y1mu[ol == x, il], 
                         "v" = round(y1mu[ol == x, bias] / muy[x] * 100, 1)),
        cbind.data.frame("Map" = p1mu[ol == x, il], 
                         "v" = round(p1mu[ol == x, bias] / mup[x] * 100, 1)))
}))
```

### Supplemental tables
```{r, eval = FALSE}
vnms <- c(rep("Yield", length(snms)), rep("Production", length(snms)))
aa <- c("Bias", "MAE")

# Density-weighted
out_tabw <- do.call(rbind, lapply(1:2, function(x) {  # x <- 1 
  out <- do.call(cbind, lapply(1:length(lev), function(ii) {  # ii <- 1
    lv <- lev[ii]  
    mm <- malmulw[2:1][[x]]  
    v <- do.call(rbind, lapply(length(mm):1, function(j) {  # j <- 1
      v2 <- do.call(rbind, lapply(1:length(snms), function(i) { # i <- 1
        nm <- snms[i]
        round(mm[[j]][ol == lv, get(nm)] / get(pyl[j])[lv] * 100, 1)
      }))
      rownames(v2) <- snms
      v2
    }))
  }))
  out <- cbind.data.frame(aa[x], mcap, "Variable" = vnms, unname(out))
  colnames(out) <- c("Metric", "Map", "Variable", paste(c(1, fact), "km"))
  out
}))

# Agricultural area
out_taba <- do.call(rbind, lapply(1:2, function(x) {
  out <- do.call(cbind, lapply(1:length(lev), function(ii) {
    lv <- lev[ii]  
    mm <- malmula[2:1][[x]] 
    v <- do.call(rbind, lapply(length(mm):1, function(j) {
      v2 <- do.call(rbind, lapply(1:length(snms), function(i) {
        nm <- snms[i]
        round(get(mm[j])[ol == lv & il == nm,  bias] / get(pyl[j])[lv] * 100,1)
      }))
      rownames(v2) <- snms
      v2
    }))
  }))
  out <- cbind.data.frame(aa[x], mcap, "Variable" = vnms, unname(out))
  colnames(out) <- c("Metric", "Map", "Variable", paste(c(1, fact), "km"))
  out
}))

# Join them
out_tab <- rbind(cbind("Region" = "Density", out_tabw), 
                 cbind("Region" = "Agricultural", out_taba))

caption <- paste("Biases and mean absolute errors (MAE) in disaggregated maize",
                 "yield and production (calculated from disaggregated yield",
                 "and harvested area estimates)",
                 "maps. Values for both density-weighted and agricultural", 
                 "areas bias and accuracy are presennted.",
                 "Bias and MAE were normalized to their", 
                 "respective mean values calculated from reference maps.")    
out_xtab <- xtable(out_tab, digits = 1, caption = caption)
print(out_xtab, type = "latex", file = "paper/figures/yldprod-bias.tex", 
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)

```

### The impacts of spatial patterns in bias

(Supplmentary analysis not presented). Using yield gap estimates as an indicator. First we'll bring in mean rainfall as a means of calculating a spatial correlation in yield gap estimates. Using data from the Princeton Global Forcing dataset, downloaded from the Africa Flood and Drought Monitor. Note: currently this is only explored here. 
```{r, eval=FALSE}
# rainfall layer prepared for SA crop subsidy analysis
load("external/ext_data/climate/mag_dist_clim.rda")
pre <- calc(pre2, mean)  # 31 year mean rainfall for Africa
p4s <- projection(raster(nrow = 10, ncol = 10)) # GCS
presa <- crop(pre, spTransform(sa.shp, p4s))  # crop to SA
projection(presa) <- p4s  
presa <- projectRaster(presa, lcu$f25$sa30)  # project to Albers
presa <- mask(presa, awgts$f25, maskvalue = 0)  # mask
presa1k <- disaggregate(presa, fact = 25)

# reclassify rainfalls to nearest 100 mm, tweak to catch lower and upper bound
rclmat <- cbind(seq(50, 750, 100), seq(150, 850, 100), seq(100, 800, 100))
rclmat[1] <- 15
rclmat[8, 2] <- 875
prebin <- crop(reclassify(presa1k, rclmat), namask)  # rainfall bins

# Assume a yield gap over-estimated by 100% in a particular location, say the 
# 300-400 mm isohyet
prez <- (prebin == 300) | (prebin == 400)
notprez <- (prebin < 300) | (prebin > 400)
g1 <- 0
g2 <- 1

ygap_bias <- sapply(lev, function(x) {
  if(x != "f1") { 
    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), 
                        fun = modal)
  } else {
    preagg <- prebin
  }
  prez <- (preagg == 300) | (preagg == 400)
  notprez <- (preagg < 300) | (preagg > 400)
  sapply(snms, function(y) {
    yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2*cyld_agg[[x]][[y]] * prez)
    pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
    pg2 <- tareas[[x]] * carea_agg[[x]][[y]] * yg2
    cellStats(pg - pg2, sum) / cellStats(pg, sum) * 100
  })
})

# Assume a yield gap over-estimated by 100% in 400 mm isohyet
ygap_bias2 <- sapply(lev, function(x) {
  if(x != "f1") { 
    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun=modal)
  } else {
    preagg <- prebin
  }
  prez <- preagg == 400
  notprez <- preagg != 400
  sapply(snms, function(y) {
    yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]]*prez)
    pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
    pg2 <- tareas[[x]] * carea_agg[[x]][[y]] * yg2
    cellStats(pg - pg2, sum) / cellStats(pg, sum) * 100
  })
})

# Assume a yield gap over-estimated by 100% in 500 mm isohyet
ygap_bias3 <- sapply(lev, function(x) {
  if(x != "f1") { 
    preagg <- aggregate(prebin, fact = as.numeric(gsub("f", "", x)), fun=modal)
  } else {
    preagg <- prebin
  }
  prez <- preagg == 500
  notprez <- preagg != 500
  sapply(snms, function(y) {
    yg <- (g1 * cyld_agg[[x]]$gti * notprez) + (g2 * cyld_agg[[x]]$gti * prez)
    yg2 <- (g1 * cyld_agg[[x]][[y]] * notprez) + (g2 * cyld_agg[[x]][[y]]*prez)
    pg <- tareas[[x]] * carea_agg[[x]]$gti * yg
    pg2 <- tareas[[x]] * carea_agg[[x]][[y]] * yg2
    cellStats(pg - pg2, sum) / cellStats(pg, sum) * 100
  })
})

# Print out combined results to latex table
scen <- c(rep("300-400 mm", nrow(ygap_bias)), rep("400 mm", nrow(ygap_bias)),
          rep("500 mm", nrow(ygap_bias)))
ygaps <- cbind.data.frame(scen, "sensor" = rep(snms, 3), 
                          rbind(ygap_bias, ygap_bias2, ygap_bias3))
ygap_xtab <- xtable::xtable(ygaps, digits = 1)
print(ygap_xtab, type = "latex", 
      file = "paper/figures/ygap-spat-bias.tex")

save(list = ls(), file = full_path(p_data, "yield-bias.rda"))
```

The actual impact of the spatial bias is fairly low, being largely cancelled out by biases in other direction in other areas. This illustrates the advantage of a statistically constrained approach to crop area estimates.

```{r, eval=FALSE, echo=FALSE}
# looking into pycnophylactic interpolation of correction factors
# install.packages("pycno")
require(pycno)

# pycno package example
nc_sids <- readOGR(system.file("shapes/sids.shp", package="maptools")[1], 
                   layer = "sids")
nc_sids@proj4string <- CRS("+proj=longlat +ellps=clrk66")
births74 <- pycno(nc_sids, nc_sids$BIR74, 0.05, converge = 1)
plot(raster(as(births74, "SpatialPixelsDataFrame")))
plot(nc_sids,add=TRUE)

# on out data
provr <- rasterize(prov, y = gti_agg$f10$gti, field = "id_1")
prov2 <- rasterToPolygons(provr, dissolve = TRUE)
colnames(prov2@data) <- "provid"
prov2@data <- cbind(prov2@data, do.call(cbind, pcf))

snms <- colnames(prov2@data)[2:5]
cfrs <- lapply(snms, function(x) {
  #x <- snms[1]
  cfr <- pycno(prov2, prov2@data[, x], 10000, converge = 1)
  cfr_spp <- as(cfr, "SpatialPixelsDataFrame")
  cfrr <- raster(cfr_spp)
  cfrrs <- resample(x = cfrr, provr, method = "ngb")
})
names(cfrs) <- snms
plot(cfrs$sa30)

reclmat <- cbind(1:9, pcf$sa30)  # reclass matrix
cfr <- reclassify(provr, reclmat)  # assign cf values to rasterized
tst <- (lc_agg$f10$sa30 * cfr)

# this interpolation does not make sense in terms of leading to a correction 
# that ends up with cropland fractions matching the provincial estimates of 
# cropland area, at least not given the equations provided by Ramankutty et al
# (2008). The provincial level correction factor should not be disaggregated so # that the summed correction factors add back up to the original correction
# factor.

cft <- 18 / sum(rep(2, 10))
sum(rep(2, 10) * cft)

```

# Agent-based model sensitivity to map error

Selection of magisterial districts and associated landcover sets for use in ABM example, and analysis of bias after initial allocation of agents performed based on available farmland estimated from 100 m downscaled versions of the landcover data.

## Data
```{r, eval=FALSE}
library(SAcropland)
library(raster)
library(rgdal)
library(lmisc)
library(readxl)
library(rgeos)

# Paths
p_root <- proj_root("SAcropland")
p_fig <- full_path(p_root, "SAcropland/paper/figures/")
p_data <- full_path(p_root, "SAcropland/external/ext_data/")
p_data2 <- full_path(p_root, "SAcropland/data/")

# landcover data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- raster(paths[2])  # gti 2011
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

# Reconstruct original landcover estimates
snms <- c("sa30", "globmu", "modmu", "geow")
dlist_1km <- lapply(dlist_act[snms], function(x) x$f1$g2011)
# dlist_1km <- stack(lapply(dlist_act[snms], function(x) x$f1$g2011))
lc_list <- lapply(dlist_1km, function(x) gti - x) 
# lc_list <- gti - dlist_1km 
# cellStats(!is.na(lc_list$sa30), sum)
lc_list <- stack(lc_list)

# magisterial district data cropland statistics (2011, but only to look at for 
# selection 
load("external/ext_data/gti_md.rda")
mu <- data.frame(t(sapply(gti_md, function(x) {
  c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
})))
colnames(mu) <- c("N", "na", "f")

# district shapes, filtered by N missing data, minimum fraction, and area
md <- readOGR("external/ext_data/mag_dists.sqlite", layer = "mag_dists")
mdfilt <- mu[mu$na < 20 & mu$f > 5 & mu$N < 1500, ]  
mdsel <- as.numeric(row.names(mdfilt))

# Examine, including in QGIS
# plot(md)
# plot(md[mdsel[mdsel < 314], ], col = "red", add = TRUE)  # filter out Cape ones
writeOGR(md[mdsel[mdsel < 314], ], dsn = "external/ext_data/md_abms.sqlite",
         layer = "md_abms", driver = "SQLite")

```

### Magisterial District selection

Looking at the data in QGIS, including some rainfall maps from earlier papers and the inter-tubes, it looks like the following districts are a good match:

   + Clocolan (269)
   + Ficksburg (274)
   + Fouriesburg (275)
   + Marquard (288)

```{r, eval=FALSE}
abm_md <- md[mdsel[mdsel %in% c(269, 274:275, 288)], ]
writeOGR(abm_md, dsn = "external/ext_data/md4.sqlite", layer = "md4", 
         driver = "SQLite")

# Deprecated: earlier used when running 2007 rather than 2011
# Let's compare the fractions between the GTI 2007 and GTI 2011 datasets for 
# interest's sake
# gti_abm <- extract(gti, abm_md)
# abm_mu <- data.frame(t(sapply(gti_abm, function(x) {
#   c(length(x), length(which(is.na(x))), round(mean(x, na.rm = TRUE), 1))
# })))
# colnames(abm_mu) <- c("N", "na", "f")
# abm_mu$f - mu[mdsel[mdsel %in% c(269, 274:275, 288)], "f"]
 # -1.0  0.0 -0.4 -0.2  # tinf differences!
rgeos::gArea(abm_md, byid = TRUE) / 10000 / 100
```

### Extract cropland percentage from selected districts
```{r, eval=FALSE}
gti4 <- crop(gti, extent(abm_md))
lc4 <- crop(lc_list, extent(abm_md))
cb <- brick(stack(gti4, lc4), filename = "external/ext_data/cropland-md4.tif", 
            overwrite = TRUE)
cpct <- raster::extract(gti4, abm_md)
dcrop_pct <- round(sapply(cpct, sum, na.rm = TRUE) / sapply(cpct, length), 1)
```
Cropland percentages:

  + Clocolan (269) - 45%
  + Ficksburg (274) - 39%
  + Fouriesburg (275) - 29%
  + Marquard (288) - 44%

### Map selected districts, cropland percentage, and error
```{r, eval = FALSE}
brks1 <- seq(0, 100, 10)
cols1 <- rev(terrain.colors(length(brks1) - 1))
brks2 <- c(-100, -80, -60, -40, -20, -10, -5, -1)
brks2 <- c(brks2, rev(abs(brks2)))
cols2 <- colorRampPalette(c("red", "grey80", "blue4"))(length(brks2) - 1)
cols2[c(1, length(cols2))] <- c("darkred", "purple3")
cx <- 1.1
lcol <- "black"
mcap <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")

# layout.show()
pdf(full_path(p_fig, "abm-selected-districts.pdf"), height = 7, width = 7)
# png(full_path(p_fig, "abm-selected-districts.png"), height = 700, width = 700)
par(mfrow = c(3, 2), mar = c(0, 0, 0, 0), oma = c(2, 0, 0, 0))
plot(md, col = "grey", border = "transparent")
plot(abm_md, col = "red", border = "grey30", add = TRUE)
par(mar = c(3, 0, 0, 0))
plot(abm_md)
plot(cb[[1]], axes = FALSE, box = FALSE, breaks = brks1, col = cols1, 
     add = TRUE, legend = FALSE)
plot(abm_md, add = TRUE)
polygonsLabel(abm_md, labels = paste0(abm_md$name_2, " (", 1:4, ")"), 
              method = "buffer", cex = 0.9)
for(i in 2:5) {
  plot(abm_md)
  plot(cb[[1]] - cb[[i]], axes = FALSE, box = FALSE, breaks = brks2, 
       col = cols2, add = TRUE, legend = FALSE)
  plot(abm_md, add = TRUE, lwd = 2)

  mtext(mcap[i - 1], side = 1, line = -4, adj = 0.7)
}
nct <- length(brks1) - 1
flex_legend(ncuts = nct, legend.text = "% cropland", legend.vals = rep("", nct),
            longdims = c(0.6, 0.9), shortdims = c(0.72, 0.012), colvec = cols1, 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 3), leg.adj = list(c(4.2, 0), c(-0.5, -0.5)), 
            cex.val = 1.1, textcol = lcol, bordercol = lcol)
tc <- rect_coords(0.6, 0.9, nct, constEWorNS = 0.72, resEWorNS = 0.02)
xs <- c(tc[[1]][1, 1], sapply(tc, function(x) x[2, 1]))
ys <- sapply(tc, function(x) x[1, 2])
text(xs, ys, labels = brks1, srt = 270, adj = c(-0.15, 0.5), cex = 1.1)
nct <- length(brks2) - 1
flex_legend(ncuts = nct, legend.text = "% bias", legend.vals = rep("", nct),
            longdims = c(0.2, 0.8), shortdims = c(0.05, 0.015), colvec = cols2, 
            srt = c(270, 0), horiz = TRUE, textside = "bottom", 
            legend.pos = c(4, 6), leg.adj = list(c(4.2, 0), c(-0.5, -0.5)), 
            cex.val = 1.1, textcol = lcol, bordercol = lcol)
tc <- rect_coords(0.2, 0.8, nct, constEWorNS = 0.047, resEWorNS = 0.012)
xs <- c(tc[[1]][1, 1], sapply(tc, function(x) x[2, 1]))
ys <- sapply(tc, function(x) x[1, 2])
text(xs, ys, labels = brks2, srt = 270, adj = c(-0.1, 0.5), cex = 1.1)
dev.off()
```

![figure](../paper/figures/abm-selected-districts.png)

## Select modeled yields for agent-based model

From earlier simulation, to give idea about variability
```{r, eval = FALSE}
dssat <- fread(full_path(p_data, "abm/dssat-yields.csv"))
dssat_red <- copy(dssat[])

dts <- c(2007288, 2007303, 2007318, 2007333, 2007348, 2007363, 2008013, 2008028)
dssat[PDAT %in% dts, ]

plot(1:31, 1:31, ylim = range(dssat$HWAH), pch = "", xlab = "YEAR", 
     ylab = "HWAH")
points(dssat[PDAT %in% dts & FNAM != "ZASP0001", HWAH])
boxplot(dssat[PDAT %in% dts & FNAM == "ZASP0001", HWAH])
boxplot(dssat[PDAT %in% dts & FNAM == "ZASP0002", HWAH])
boxplot(dssat[PDAT %in% dts & FNAM == "ZASP0003", HWAH])
dssat[PDAT %in% dts, mean(HWAH), by = FNAM]
dssat[PDAT %in% dts, mean(HWAH), by = .(CU, FNAM)]

plot(1:8, 1:8, ylim = range(dssat$HWAH), pch = "", xlab = "YEAR", 
     ylab = "HWAH")
for(i in unique(dssat$FNAM)) {
  dssat[PDAT %in% dts & CU == 1 & FNAM == i, lines(HWAH)]
}

# provincial ag census statistics from 2007
prodyld <- cbind.data.frame("md" = as.character(abm_md$name_2), 
                            "pha" = c(6219, 6203, 6391, 9404), 
                            "prod" = c(14781, 17434, 24366, 25578))
prodyld$yld <- prodyld$prod / prodyld$pha

outylds <- copy(dssat[PDAT %in% c(2007303, 2007348, 2008013) & 
                   FNAM != "ZASP0001", ])
outylds[PDAT %in% dts, mean(HWAH), by = .(FNAM)]
outylds[PDAT %in% dts, mean(HWAH), by = CU]
outylds[PDAT %in% dts, mean(HWAH), by = .(CU, FNAM)]
write.csv(outylds, file = full_path(p_data, "abm/abm-yields.csv"))
```

## Downscaling and agent allocation

Performed separately and described in the methods and supplemental.

## Analysis of agent allocation errors
```{r, eval = FALSE}
abias <- data.table(read_excel(full_path(p_data, "abm/stats.xlsx")))
# abias <- abias[Alg == 1]
setkey(abias, MD)
hhs <- data.table(read_excel(full_path(p_data, "abm/agent-bias2.xlsx"), 
                             sheet = 2))
setkey(hhs, MD)
abias <- abias[hhs]
abias

# calculate district level bias as percent
pctbias <- abias[var == "ag_area", 
                 lapply(list(sa30, globmu, modmu, geow), function(x) {
                   (gti - x) / gti * 100
                 })]

# HHs without land as percent total households
noland <- abias[var == "HHs_no_land", 
                lapply(list(sa30, globmu, modmu, geow), function(x) {
                  x / HHs * 100
                })]

# unallocated farmland
unalloc <- round(abias[var == "ag_area_unalloc", snms, with = FALSE] / 
  abias[var == "ag_area", snms, with = FALSE] * 100, 2)

# land deficit farmland
land_def <- round(abias[var == "land_def", snms, with = FALSE] / 
  abias[var == "ag_area", snms, with = FALSE] * 100, 2)

# households couldn't find enough land
not_enough <- round(abias[var == "HHs_cant_find", snms, with = FALSE] / 
                      (abias[var == "HHs_no_land", 
                             lapply(.SD, function(x) HH_gen - x), 
                             .SDcol = c(snms, "HH_gen")][, snms, with = FALSE]), 
                    2)

# production deficits
hhprod <- abias[var == "prod", gti / HHs]
mdprodlc <- abias[var == "prod", 
                  lapply(.(sa30, globmu, modmu, geow), function(x) x / HHs)]
prdef <- abias[var == "prod", lapply(.(sa30, globmu, modmu, geow), function(x) {
  (gti - x) / gti * 100
})]  # district total deficit
hhprodlc <- abias[var == "prod", 
                  lapply(.(sa30, globmu, modmu, geow), function(x) x / HHs)]
hhprod_def <- (hhprod - hhprodlc) / hhprod * 100  # household deficit

cexs <- seq(1, 2, 1 / 3)
# pdf(full_path(p_fig, "agent-bias.pdf"), height = 5, width = 3)
pdf(full_path(p_fig, "agent-bias.pdf"), height = 2.5, width = 5)
cx <- c(0.9, 1, 0.6)
# png(full_path(p_fig, "agent-bias.png"), height = 350, width = 700)
# par(mfrow = c(2, 1), mar = c(0.5, 2.5, 0.5, 1), mgp = c(1, 0.05, 0), 
#     tcl = -0.2, oma = c(3.5, 0, 0, 0))
par(mfrow = c(1, 3), mar = c(4, 2, 0.5, 0), mgp = c(0.9, 0.1, 0), 
    tcl = -0.2, oma = c(3.5, 0, 0, 0.1))
scols <- c("red", "orange3", "green4", "blue")
cols <- c(rep("red", 4), rep("orange3", 4), rep("green4", 4), rep("blue", 4))
plot(unlist(pctbias), unlist(unalloc), col = cols, cex = cexs, pch = 20, 
     ylab = "% land unallocated", xlab = "% error in cropland area", 
     cex.lab = cx[2], cex.axis = cx[1])
plot(unlist(pctbias), unlist(land_def), col = cols, cex = cexs, pch = 20, 
     ylab = "% land deficit", xlab = "% error in cropland area", 
     cex.lab = cx[2], cex.axis = cx[1])
plot(unlist(pctbias), unlist(hhprod_def), col = cols, cex = cexs, pch = 20, 
     ylab = "% food deficit", xlab = "% error in cropland area", 
     cex.lab = cx[2], cex.axis = cx[1])
par(xpd = NA)

yst <- seq(0.05, 0.23, 0.18 / 4)
xy <- cbind(grconvertX(rep(0.5, 5), from = "ndc"), 
            grconvertY(yst, from = "ndc"))
for(i in 2:length(yst)) {
  if(i == 5) {
    leg <- c("District", 1:4)
  } else {
    leg <- rep("", 4)
  }
  legend(x = xy[i, 1], y = xy[i, 2], legend = rep("",4), #legend = leg, 
         pch = 20, cex = cx[3], 
         horiz = TRUE, adj = c(4, -5.75),  
         col = scols[i - 1], pt.cex = cexs, bty = "n")#cex = cexs,)
}
yst <- seq(0.06, 0.24, 0.18 / 4)
xy <- cbind(grconvertX(rep(0.505, 5), from = "ndc"), 
            grconvertY(yst, from = "ndc"))
text(x = xy[, 1], y = xy[, 2], labels = c(mcap, "District"), 
     #srt = 90, #pos = 2, 
     #adj = c(-1, 1), 
     pos = 2, col = c(scols, "black"), cex = cx[1])
xst <- seq(0.53, 0.60, 0.07 / 3)
xy <- cbind(grconvertX(xst, from = "ndc"), 
            grconvertY(rep(yst[length(yst)], 4), from = "ndc"))
text(x = xy[, 1], y = xy[, 2], labels = 1:4, pos = 2, cex = cx[1])
dev.off()

```

![figure](../paper/figures/agent-bias.png)

# Miscellaneous additional calculations

```{r, eval=FALSE}
library(rgdal)
library(SAcropland)
library(rgeos)

p_dat <- full_path(full_path(proj_root("SAcropland"), "SAcropland"), 
                   "external/ext_data")

```

Area of sub-Saharan Africa and South Africa's share of that area
```{r, eval=FALSE}
eco <- readOGR(dsn = full_path(p_dat, "africa_ecofloristic_zones.sqlite"), 
               layer = "africa_ecofloristic_zones")
eco@proj4string <- CRS(projection(raster(nrow = 1, ncol = 1)))
af <- readOGR(dsn = full_path(p_dat, "africa_countries_alb.sqlite"), 
               layer = "africa_countries_alb")
af@data <- af@data[, c(1, 3:4)]
ecoalb <- spTransform(eco, af@proj4string)

# cut down countries
nms <- af$cntry_name[af$region == "Northern Africa"]
nms <- nms[nms != "Sudan"]  # keep Sudan
ssa <- af[!af$cntry_name %in% nms, ]
ssa <- ssa[-grep("Island|Sao|Verde|Mauritius|Helena|Comoros|Portugal|Seychell",
                 as.character(ssa@data$cntry_name)), ]
ssa <- gBuffer(ssa, width = 0)
# plot(ssa)

desert <- ecoalb[grep("desert", ecoalb$gez_term), ]
# plot(ecoalb[names(which.max(gArea(desert, byid = TRUE) / 10000 / 100)), ])
sahara <- as.numeric(names(which.max(gArea(desert, byid = TRUE))))

nosahara <- ecoalb[-sahara, ]
plot(nosahara, col = "red")

ovs <- which(!is.na(over(nosahara, ssa)))
ssaeco <- nosahara[ovs, ]
ssaeco <- ssaeco[-1, ]  # drop a couple Saharan mountains
ssaeco <- ssaeco[-1, ]
# plot(ssaeco[ssaeco@data$gez_term == "Tropical mountain system", ][1, ])
plot(ssa)
plot(ssaeco, col = "red", add = TRUE)

# calculate area of remainder
afarea <- gArea(ssaeco) / 10000
# plot(af[af$cntry_name == "South Africa", ])
saarea <- gArea(af[af$cntry_name == "South Africa", ]) / 10000
saarea / afarea

```

```{r, eval=FALSE}
a <- sample(1:100, size = 20)
b <- sample(1:80, size = 20)
mean(a - b)
sum(a) / sum(b)

```

### 2015-12-01 Rmd/compile-book.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

---
title: "Compile code Rmd into single PDF"
author: "Lyndon Estes"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Compile book

Solution from [here](http://stackoverflow.com/questions/25824795/how-to-combine-two-rmarkdown-rmd-files-into-a-single-output).
```{r, eval=FALSE}
d <- "Rmd"
a <- "---\ntitle: 'Code Appendix'\noutput:\n  pdf_document:\n    toc: yes\n"
b <- "    toc_depth: 4\n    highlight: tango\n    number_sections: yes\n---"
book_header <- readLines(textConnection(paste0(a, b)))
# book_header <- readLines(textConnection("---\ntitle: 'Title'\n---"))
old <- setwd(d)
if(length(grep("code.Rmd", list.files())) > 0){
  warning("code.Rmd already exists")
}
write(book_header, file = "code.Rmd", )
cfiles <- list.files(pattern = "[0-9].*.Rmd")
ttext <- NULL
for(i in 1:length(cfiles)){
  text <- readLines(cfiles[i])
  hspan <- grep("---", text)
  text <- text[-c(hspan[1]:hspan[2])]
  write(text, sep = "\n", file = "code.Rmd", append = TRUE)
}
```

Some editing of unified code file to remove back to top hyperlinks before knitting. 
```{r, eval = FALSE}
rmarkdown::render("code.Rmd", output_format = "pdf_document")  
```

### 2015-12-01 paper/SI.tex
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

\RequirePackage{lineno}  % http://www-d0.fnal.gov/Run2Physics/WWW/templates/lineno.html
%\documentclass[12pt]{iopart}
\documentclass[11pt, titlepage]{article}
\usepackage[top=20mm, bottom=20mm, left=23mm, right=23mm]{geometry}
%\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}  
% parameters from other files
\newcommand{\citetapos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}  % apostrophe's in cites
%\usepackage[square,numbers,sort&compress]{natbib}
\def\linenumberfont{\normalfont\small\sffamily}
\usepackage{graphicx}
\usepackage{epstopdf}
% From link below: next two lines allow caption to fill the full figure box, and caption formats to be changed (e.g. font)
% http://tex.stackexchange.com/questions/107350/caption-below-the-figure-and-aligned-with-left-side-of-figure
% http://ctan.mackichan.com/macros/latex/contrib/caption/caption-eng.pdf
\usepackage{caption}
\captionsetup[figure]{slc=off, font=footnotesize}
%\usepackage{amsmath}
\usepackage{longtable}
\usepackage{placeins}
\usepackage{multirow}
\usepackage[numbers,round,sort,compress]{natbib}
\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
% figures float at top of page
\makeatletter
\setlength{\@fptop}{0pt}
\setlength{\@fpbot}{0pt plus 1fil}
\makeatother

\begin{document}
%\bibliographystyle{unsrt}

\setpagewiselinenumbers
%\modulolinenumbers[5]
%\linenumbers
\begin{center}\section*{\Large Supplementary Methods and Results}\end{center}
%\begin{table}[ht]
%\centering
%\caption{Confidence classes for the forcing data based on bias-correction weights.}
%\begin{tabular}{cc}
%  \hline
%  Confidence class & Bias-correction weight \\ 
%  \hline
%High & 0.75-1.00 \\ 
%Medium-High & 0.50-0.75\\ 
%Medium-Low & 0.25-0.50\\ 
% Low & 0.05-0.25 \\ 
% None & $<$0.05 \\ 
%   \hline
%\end{tabular}
%\end{table}  
\section{\large Overview}
This document provides supplementary information on methods and results from the different components of the analysis described in the main text.  The code used to conduct the analysis is presented after sections 1-X.

\section{\large Reference data accuracy}

The accuracy of the reference vector dataset was assessed by Estes et al \cite{estes_platform_2015}, and described further here. The assessment was undertaken using a sub-sample of 609 1 km$^2$ grid cells, which were selected using a weighted randomized sampling scheme. Weights were derived from a logistic regression model of cropland occurrence probability in South Africa, with 1 corresponding to the lowest quartile of probability, and 4 the highest. The accuracy assessment included all cropland classes mapped within the reference dataset, which ranged from communal/smallholder fields to commercial row crops to orchards and other types of horticulture. A visual accuracy assessment was conducted within each 1 km$^2$ grid cell, wherein each cell was divided into 25 smaller cells of 200 X 200 m (4 ha), and then the proportion of each cell overlapping with cropfields visible in underlying, high resolution Google Maps imagery was then calculated to the nearest 5\% of coverage, using a finer 20-cell mesh overlaid on each sub-cell within which field presence/absence was recorded. The same procedure was then performed to assess coverage by reference map polygons, as well as the intersections and differences between cells determined to be occupied by i) visual assessment and ii) by the digitized polygons. These intersections and differences were used to calculate the area of true positives, false positives, true negatives, and false negatives within each 1 km$^2$ grid cell, which were then used to calculate accuracy measures.  In this case, accuracy was assessed using two landcover classes: cropland and non-cropland, thus a two class confusion matrix was constructed (Table 1) according to remote sensing classification accuracy guidelines set out by Oloffson et al. \cite{olofsson_good_2014}. In this case, the reference data were the visually interpreted crop field presences/absences, and the map data were the field boundary vectors being assessed (i.e. the data which provide the reference maps in this study), the class ``crop" refers to crop field presence and ``non-crop" to crop field absence.  The total area of vectorized crop fields was used to calculate the proportion \emph{W} of South Africa's total area mapped as crop fields, and of the areas ``mapped'' (by omission) as having other cover types.  These were then used to weight the proportions of each mapped class \emph{i} corresponding to each reference class \emph{j} by the total area \emph{A} mapped for class \emph{i}. The total accuracy (0.97, or 97\%) was calculated by summing the diagonal (bold in Table 1), as well as the producer's accuracy \emph{P} for each reference class \emph{j}, and the user's accuracy \emph{U} for each mapped class \emph{i}.

\begin{table}[!ht]
\centering
\caption{Confusion matrix for the assessment of the hand-digitized crop field boundaries used to generate reference cropland cover percentage values. Here \emph{Reference} denotes a visual assessment of crop field presence in high resolution imagery within a sample of 609 sites, while \emph{Map} refers to the vectorized crop field boundaries. The assessment had two classes: crop fields and non-crop fields, and proportions corresponding to these two categories are the proportions of areas determined by class agreements and disagreement in the 609 sites (see text above), which are weighted by the proportion \emph{W} of each class' total mapped area \emph{A$_i$} in South Africa. \emph{P} and \emph{U} provide the Producer's and User's accuracies, respectively. }
\begin{tabular}{rrrrrrr}
  \hline
  & & Reference (j) & & & & \\\cline{3-7}
  & & Crop & Non-crop & \emph{W}$_i$ & \emph{A}$_i$ (ha) & \emph{U}$_i$ \\ 
  \hline
  Map (i) & Crop & \textbf{0.108} & 0.007 & 0.114 & 14018567 & 0.943 \\ 
  & Non-crop & 0.021 & \textbf{0.865} & 0.886 & 108541733 & 0.977 \\ 
  & Total & 0.128 & 0.872 & 1.000 & 122560300 &  \\ 
  & \emph{P}$_j$ & 0.840 & 0.992 &  &  &  \\ 
   \hline
\end{tabular}
\end{table}

%\clearpage
\FloatBarrier
\section{\large Cropland map error analysis}
We examined the impact of several sources of uncertainty on our results.  The first is the potential temporal mismatch between the reference dataset and the landcover products we were testing.  We tested this by examining two versions of the reference dataset, the initial version created in 2007, and the updated version from 2011 used in the main analysis.  The 2011 version shows 3\% more cropland area than the 2007 version. To examine the effect of the over mapping bias and accuracy measures, we converted the 2007 vector maps to gridded cropland percentage and subtracted each test map, and then compared the differences in bias and mean absolute error (MAE) values calculated between both (Table 2). The largest difference was 1.63\%, between the 1 km SA-LC residuals, which means that the overestimation bias by SA-LC was actually greater relative to the 2007 version of the reference map.  Except for the corresponding MAE value, all other difference were $<$1\%.

% latex table generated in R 3.2.1 by xtable 1.7-4 package
% Wed Nov 25 15:57:07 2015
\begin{table}[!ht]
\caption{Differences in the bias and mean absolute errors values resulting from cropland percentage residuals calculated using 1) the 2011 reference map and 2) the 2007 reference map, with 2007 values subtracted from 2011 values. Differences from each test map and each scale of aggregation are shown. }
\centering
\begin{tabular}{rlrrrr}
  \hline
 & Resolution & SA-LC & GlobCover & MODIS & GeoWiki \\ 
  \hline
1 & 1 km & 1.63 & -0.80 & 0.12 & 0.64 \\ 
  2 & 5 km & 0.99 & -0.38 & 0.40 & 0.80 \\ 
  3 & 10 km & 0.89 & -0.27 & 0.44 & 0.77 \\ 
  4 & 25 km & 0.83 & -0.12 & 0.47 & 0.76 \\ 
  5 & 50 km & 0.79 & 0.04 & 0.47 & 0.73 \\ 
  6 & 100 km & 0.76 & 0.19 & 0.54 & 0.72 \\ 
  7 & 1 km & 1.30 & -0.77 & -0.10 & 0.27 \\ 
  8 & 5 km & -0.22 & -0.44 & -0.08 & 0.12 \\ 
  9 & 10 km & -0.48 & -0.34 & -0.08 & 0.12 \\ 
  10 & 25 km & -0.61 & -0.20 & -0.08 & 0.20 \\ 
  11 & 50 km & -0.67 & -0.09 & -0.36 & 0.22 \\ 
  12 & 100 km & -0.70 & 0.04 & -0.24 & 0.19 \\ 
   \hline
\end{tabular}
\end{table}

As a further exploration of uncertainty, we included the full range of variability that resulted from a) using either the 2007 or 2011 map to calculate bias and accuracy, and b) from different levels of cropland percentages used when converting MODIS and GlobCover mixed cropland classes into cropland percentage maps, following Fritz et al. \cite{fritz_mapping_2015}. We pooled the residuals from across each of these permutations, for each reference-test map combination, and examined how their overall mean values change with aggregation scale (Fig. 1), and also assessed the total distribution of errors within different class of cropland density, determined by dividing the 2011 1 km reference map into 20 different zones, or bins, of cropland density, ranging from 0-5\% cropland cover up to 95-100\% cover. We then calculated summary statistics from the pooled residuals and absolute values of residuals within each of these bins (Fig. 2).

\begin{figure}[!ht]
  \centering
     \includegraphics[width = 14cm]{figures/biases_1-100km.pdf} 
%     \vspace{-3cm}
      \caption{Biases and mean absolute errors (MAE) for each of the cropland maps as a function of cropland density (calculated using the 2011 reference maps) and aggregation scales. Rows present biases by map product, columns by aggregation scale.  Dash lines indicate bias at each level of cropland density, calculated in bins spanning 5\% of density (e.g. 0-5\% cropland cover, 5-10\%, etc.), while solid lines indicate the mean absolute error.  The black numbers in each plot area present the overall means of bias/MAE for each sensor-scale combination. The bin-wise and overall mean statistics were calculated from pooled map errors calculated from differences between the 2007 reference map and each cropland map (including all three variants--high, medium, and low--of the MODIS and GlobCover-derived cropland maps), and the 2011 reference map and each cropland map. }
      \label{fig:default}
\end{figure}

%\vspace{0.5 cm}
%\begin{equation}
%\hspace*{1.5 cm} 
%   \label{metric}
%     R_n  = \frac{(1 - \alpha)dswr + dlwr - \sigma * T^4}{\lambda}
%\end{equation}

\begin{figure}[!ht]
  \centering
     \includegraphics[width = 14cm]{figures/biases_1km.pdf} 
%     \vspace{-3cm}
      \caption{Biases and mean absolute errors (MAE) for each of the cropland maps at 1 km resolution, as a function of cropland density. Colored lines (color-coded to map product name) show the bias/MAE at each level of cropland density, calculated in bins spanning 5\% (e.g. 0-5\% cropland cover, 5-10\%, etc.). Box plots show the variability of bias in each bin (whiskers = 2.5 and 97.5 percentiles, box the inter-quartile, and grey bar in box the median). Biases are presented in the top row, and MAEs in the bottom row. Statistics were calculated from pooled map errors calculated from differences between the 2007 reference map and each cropland map (including all three variants--high, medium, and low--of the MODIS and GlobCover-derived cropland maps), and the 2011 reference map and each cropland map. }
      \label{fig:default}
\end{figure}

We also calculated how different methods of calculating the bias and accuracy statistics impacted out findings, using three different methods. First, we simply took the straight averages across the entire country, which substantially understates bias and accuracy because cropland covers only 10\% of the country, and all landcover products successfully discriminate the non-cropped regions.  We also compared the average error metrics extracted from within just the agricultural regions of the maps, defined here as the union of areas having $>$0.5\% cropland in the reference and each test map. This also tends to understate bias, because the area having $<$5\% cropland is much larger than areas having higher densities of cropland (see Fig. 4 frequencies of cells per five percentile bins of cropland density). Finally, we calculated a density-independent metrics, which is similar to the density weighted mean calculated for our main analysis, which was calculated by averaging the mean bias and MAE values within different levels of cropland cover (divided in increments of 5\%). The results of these three alternate metrics are shown in Figure 3.

\begin{figure}[!ht]
  \centering
     \includegraphics[width = 7cm]{figures/cropland_bias_region.pdf} 
%     \vspace{-3cm}
      \caption{A comparison of three alternate methods for calculating bias and accuracy (mean absolute error): as a straight average across the entire country, averaged over agricultural areas only (the union of areas defined as having $>$0.5\% cropland cover in the reference map and each test map), and independent of cropland density, wherein the mean bias/MAE values for each of 20 cropland cover classes (representing 5\% increments of cover 0\% to 100\% defined by the reference map) were calculated and then averaged.}
      \label{fig:default}
\end{figure}

%\input{figures/cropland2011-bias.tex}
%\input{figures/cropland2007-bias.tex}

\begin{figure}[!ht]
  \centering
     \includegraphics[width = 12cm]{figures/cropland_bins.pdf} 
%     \vspace{-3cm}
      \caption{Number of cells within each cropland density bin at each scale of aggregation, where bins represent 5\% increments of cropland cover (values on x-axis provide the upper limit of each bin). Bin values were based on the 2011 reference map, excluding areas with $<$0.5\% cropland.}
      \label{fig:default}
\end{figure}

We used magisterial district boundaries for South Africa (Fig. 5) to extract absolute residual values for agricultural areas ($>$0.5\% cropland cover) and 2011 reference cropland density values, the means of which respectively served as the response and predictor in the generalized additive model \cite{wood_mgcv:_2001} analysis (main text).

\begin{figure}[!ht]
  \centering
     \includegraphics[scale = 0.7]{figures/md_map.pdf} 
%     \vspace{-3cm}
      \caption{South Africa's magisterial districts.}
      \label{fig:default}
\end{figure}

\clearpage
\section{\large Carbon analysis}

To calculate carbon stocks using the method of Ruesch and Gibbs \cite{ruesch_new_2008}, we simplified their landcover-specific carbon density classes into six categories, based on the similarity of their class types and the fact that they had the same assigned carbon densities. Other classes were dropped because of their low level of, or lack of, occurrence. The class adjustments were as follows:

\begin{itemize}
\item Classes 1, 3, 6, 8 corresponding to broadleaf and mixed forests were reduced to a single forest class.
\item Classes 4 and 5 (needleaf forests) were dropped. 
\item Classes 9, 10, 17 (secondary forests, forest/cropland mosaic) where merged to a single secondary forest class.
\item Classes 20-23 (water, snow, ice, built-up areas) were dropped.
\item Class 19 (bare areas) was dropped.
\end{itemize}

All other classes (cropland, shrubland, spare vegetation) were retained. Each of these classes has a specific carbon density according to the ecofloristic zone it is found in, of which there are 10 for Africa (Zones 6-9 and 10-15; see \cite{ruesch_new_2008}). We calculated the area of each ecofloristic zone using their polygon boundary maps\footnote{available from cdiac.ornl.gov/ftp/global\_carbon/ecofloristic\_zones.zip}. For each of the simplified cover classes, we then calculated the mean ecofloristic zone carbon density value, weighted by ecofloristic zone area. We used these values to generate the different carbon maps, as described in the main text.

Maps of the mean residual differences between the reference map and each of the 5 derived carbon maps for each test map are shown in Figure 6. Table 3 provides the bias and mean absolute errors for the different carbon maps and how they vary with scale.

\begin{figure}[!ht]
  \centering
     \includegraphics[width = 12cm]{figures/carbon_bias_map.pdf} 
%     \vspace{-3cm}
      \caption{Spatial patterns of error (averaged across all five different possible cover types adjacent to cropland) in carbon stock estimates. }
      \label{fig:default}
\end{figure}

%\input{figures/totC-bias.tex}

\FloatBarrier
\input{figures/C-bias-accuracy.tex}

\clearpage
\section{\large Gridded maize yield and production}

For our crop yield and production analysis, we followed several steps to create the gridded yield and crop production maps. First, we calibrated cropland percentage maps so that they matched total cropland areas reported for coarser administrative units, following methods set out by Ramankutty et al \cite{ramankutty_farming_2008}:

\begin{enumerate}
 \item We extracted the 2011 reference percentages within each of South Africa's 9 provinces (the same units used by Ramankutty et al \cite{ramankutty_farming_2008}, converted these to proportions and summed them to calculate the ``reported'' cropland areas for each province. We did the same to calculate provincial cropland areas for each test map;  
  \item We divided the ``reported'' provincial cropland area estimates by provincial areas to calculate \emph{rpcf}, the reference provincial cropland fractions, and then did the same with test map provincial cropland areas to calculate \emph{tpcf}, the test map provincial cropland fractions; 
  \item A province-specific correction factor \emph{upcf} (\emph{rpcf/tpcf}) was then calculated and applied to each testmap cropland fraction (\emph{tcf}) in each pixel (x, y):

  \begin{equation}  
    \textrm{tcfa}_{x,y} = \textrm{upcf}_{x,y} * \textrm{tcf}_{x,y}
  \end{equation}

Yielding \emph{tcfa}, the calibrated test map cropland fraction.  
\end{enumerate}

The differences between the reference map cropland fraction and \emph{tfca} are shown in Figure 7 (where they were adjusted back to percentages), while Table 4 provides the corresponding bias and accuracy values (MAE).

We then used \emph{tcfa}, per Monfreda et al \cite{monfreda_farming_2008}, to disaggregate magisterial district-level reported maize yields and harvested areas \cite{statistics_south_africa_building_2007}, according to the following steps:

\begin{enumerate}
  \item We first disaggregated maize yields using the following formula:

  \begin{equation}
  \textrm{fcrop}_{x,y} = \textrm{tfca}_{x,y}\frac{\textrm{cmda}}{\textrm{tmda}}
  \end{equation}

  Where \emph{fcrop} is the fraction of each cell in each district harvested for maize, \emph{cmda} is the district-reported harvested area for maize, and \emph{tmda} is the total area of the magisterial district;
  \item The reported maize yield for each magisterial district was then assigned to each pixel in \emph{fcrop} in the district having non-zero maize harvested areas;
  \item Production estimates were calculated by multiplying yield and harvested areas.
\end{enumerate}

The maps of differences between reference-derived yield and production maps and those based on the test maps are shown in Figures 8 and 9, where the values were normalized to the country mean yield or production (calculated from the reference map).  Table 5 presents the bias and accuracy values for each reference-test comparison of gridded yield and production estimates, giving both the cropland density-weighted variants of bias and accuracy, and for comparison the same values calculated as straight averages within agricultural pixels (i.e. pixels in the reference or test map having $.$0.5\% cropland). In the former variant, yields show relatively much less bias than in the latter, where large yield biases were driven by the discrepancies in the low cropland density regions in the center of the country. In contrast, the density-weighted measures reveal large production biases, whereas the unweighted agricultural region measures show no bias in production estimates.  The density-weighted variant shows that error scales with cropland cover, as seen in Figure 2 in the main text.

\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/cropland_adj_bias_map2.pdf} 
      \caption{Errors in cropland maps adjusted using provincial cropland area statistics.}
      \label{fig:default}
\end{figure}

\FloatBarrier
\input{figures/cropadj-bias-accuracy.tex}

\clearpage
\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/yld_bias_map.pdf} 
      \caption{Errors (normalized to the reference-derived country mean) in disaggregated maize yield estimates.}
      \label{fig:default}
\end{figure}

\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/prod_bias_map.pdf} 
      \caption{Errors (normalized to reference-derived country mean) production estimates calculated from disaggregated maize yield and harvested area estimates.}
      \label{fig:default}
\end{figure}

\FloatBarrier
\input{figures/yldprod-bias.tex}

\section{\large Evapotranspiration analysis}
%\input{figures/et-bias.tex}
A number of the variables used by the Variable Infiltration Capacity \cite[VIC][]{liang_simple_1994} model are linked to a 0.25$^{\circ}$ resolution, AVHRR-derived landcover map. These include seasonal leaf area index (LAI) phenologies (Fig. 10), as well as other properties such as plant rooting depth and infiltration rates. Landcover properties therefore impact the model's simulation of water balance. To test how landcover map error impacts VIC simulations, we created five new versions of VIC's native landcover scheme, one based on the reference cropland map, and the other four on each of the test maps. For each version, we first reprojected the relevant 25 km cropland map to the geographic coordinate system used by VIC and resampled it to 0.25$^{\circ}$. We then adjusted VIC's landcover scheme by replacing its existing cropland percentages with those from our map, and then proportionally adjusted the remaining cover types in each cell to accommodate the changed cropland amounts. Since LAI is strongly linked to rainfall seasonality in South Africa, which varies between winter (May-August) rainfall in the west and southwest of the country and summer (October-March) rainfall in the east and northeast, we assigned LAI curves that peaked during the winter months to the cover types in the western half of the country, and those peaking in the summer months to the eastern half (Fig. 10).

\begin{figure}[!ht]
  \centering
     \includegraphics[width = 10cm]{figures/lai_seasonal_cycles.png} 
      \caption{Seasonal LAI curves for different cover types in VIC's landcover scheme, showing the phenologies used for the winter-rainfall portion of the country in the west and southwest (left column), and for the summer rainfall region in the country's easter to northeast (right column). }
      \label{fig:default}
\end{figure}

After making these adjustments, we conducted five different VIC simulations, one for each of the adjusted landcover schemes. The model was run at a daily time step for 28 years, from 1981-2008, from which monthly total evapotranspiration values were extracted.  We calculated from these results the average annual total ET, the maximum and minimum monthly ET observed throughout the entire time series, and the mean ET during the month which on average had the highest ET during the time series. We then found the differences between the reference and test map variants of each of these ET variables, and calculated their bias and MAE values (Fig. 11). The results are similar across the four variables, and we present the difference maps for the mean annual ET comparison in the main text.

\begin{figure}[ht]
  \centering
     \includegraphics[width = 12cm]{figures/et_bias_map4.pdf} 
      \caption{Maps of error between reference- and test map-derived ET estimates simulated by the VIC model. Four different variants of ET were assessed: annual mean (left column, also presented in Fig. 2 of the main text), the maximum (second column, TS max) and minimum (third column, TS min) ET values observed during the 30 year times, and mean ET during the month in which ET peaks (right column, Peak).}
      \label{fig:default}
\end{figure}

\section{\large Agent-based model assessment}

\subsection{\large Maize production}
The agent-based model simulates household food production using a look-up table that links maize yields to several different variables: planting date; cultivar (open-pollinated or hybrid); soil properties, and weather.  The look-up table itself was based on a series of yield simulations conducted by the DSSAT cropping system model run over 31 years (1979-2010) for a location in the southern Province of Zambia, corresponding to the region where household survey data were collected. The model was run for three dominant soil series\footnote{extracted from a gridded version of the ISRIC-WISE database, available from http://dssat.net/649}, using two open-pollinated and two hybrid cultivars (each pair representing a short- and medium-length growing season variety) whose coefficients were obtained from the inputs files for the PSIMs modeling platform \cite{elliott_parallel_2014}. Each cultivar was simulated under planting dates ranging from October 15th until January 30th, varying by 15 day increments, with row spacing fixed at 90 cm and planting density at 3.7 plants m$^2$, and 5 kg ha$^{-1}$ applied at planting. Models were run using weather data extracted from a bias-corrected version of the Princeton Global Meteorological forcing dataset \cite{sheffield_development_2006,chaney_spatial_2014,estes_changing_2014}.

\subsection{\large District selection}
For our analysis of how landcover map error impacts agent-based model results, we selected four districts in South Africa which had similar climatic characteristics ($\sim$800 mm annual rainfall) to the region in Zambia where the crop model simulations and household survey data were collected.  These were four contiguous districts along the western border of Lesotho, Clocolan, Ficksburg, Fouriesburg, and Marquard (Fig. 12, top left), which had between 29-45\% of their areas covered by cropland, according to the 2011 reference map (Fig. 12, top right).

\begin{figure}[!ht]
  \centering
     \includegraphics[width = 12cm]{figures/abm-selected-districts.pdf} 
      \caption{The location of the four selected magisterial districts (top left) used in evaluating agent allocation bias, the reference levels of cropland cover within those districts (top right), and the difference in cropland percentage between the reference and each of the four cropland maps (lower four panels). }
      \label{fig:default}
\end{figure}

\FloatBarrier
%\section*{References}
\bibliographystyle{pnas2011} 
{\footnotesize \bibliography{/Users/lestes/Dropbox/publications/full.bib}}
\end{document}

### 2015-12-01 paper/main.tex
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

@@
 
-\title{Landcover Data Limits Our Understanding of Global Change}
+\title{Landcover Data Limit Our Understanding of Global Change}
 %\title{Our Understanding of Global Change is Built on a Shaky Foundation}
@@
 %%% If significance statement need, then can use the below command otherwise just delete it.
-%\significancetext{LDE developed the concept of the study, conducted the analysis, data interpretation and drafted the manuscript. \color{red}{To be completed}}
+\significancetext{Remote sensing-derived landcover maps underpin much research into global environmental change, particularly in the most rapidly developing regions.
+These maps often contain substantial error, yet inadequate validation data makes it impossible to fully quantify map errors and how they affect our understanding of global change processes.
+Using a unique, high quality reference map of cropland distribution, this study comprehensively assesses landcover map bias and accuracy in a major agricultural region of sub-Saharan Africa, finding that map errors are strongly correlated with the landcover density, and can propagate substantial errors in downstream global change analyses.
+Users of landcover maps should select later generation products and aggregate to appropriate resolutions to avoid misleading insights and policies related to global change.
+}
 
@@
 First, in many regions the cover types of interest are smaller than the sensor resolution \cite[e.g.
-smallholder's farms][]{jain_mapping_2013,debats_generalized_2015,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \cite{sweeney_mapping_2015}.
+smallholder's farms][]{jain_mapping_2013,debats_generalized_2015,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \cite{sweeney_mapping_2015}, which are factors that increase mapping complexity \cite{yu_meta-discoveries_2014}.
 Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties (e.g.
@@
 
-These problems with landcover products are well known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \cite[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}.
+These problems with landcover products is known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \cite[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}.
 What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes.
-Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}.
-Our current understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data.
+Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are typically unavailable for Africa and other rapidly changing regions \cite{see_improved_2015}, which have also been fairly sparsely mapped \cite{yu_meta-discoveries_2014}.
+Our current understanding of map accuracy over such areas is therefore based on a small number of maps \cite{yu_meta-discoveries_2014}, which were assessed using either bottom-up tests made with a relatively small number of ground truth points, or from top-down "sanity checks" made in comparison to aggregated survey data.
 In terms of the former, there are well-developed, best-practice guidelines for selecting and using ground-truth samples to robustly quantify map error, but these are not commonly applied by the broader community of landcover map users (often due to lack of data), which makes it challenging to compare how error varies within and between maps \cite{stehman_global_2012, olofsson_making_2013,olofsson_good_2014,foody_status_2002}.
 The latter form of assessment are made at coarser administrative levels relative to reported statistics \cite[e.g.][]{fritz_comparison_2010}, which can also be suspect \cite{carletto_emperor_2013,fao_action_2013}.
-
-%Top-down accuracy assessments are also made from  "sanity checks" made in comparison to aggregated survey data. 
-%
-%
-%
-%This allows us to quantify between-map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.
-%
-%
-%
-%In practice, this means that we have an incomplete understanding of map error, and 
-%
-%
-%
-%
-%There are also well-defined guidelines for quantifying and accounting for map accuracy \cite{stehman_global_2012, olofsson_good_2014, olofsson_good_2014}.
-However, inadequate assessment of map error and their impacts on subsequent analyses remains quite common within the community of landcover map users \cite{olofsson_making_2013,olofsson_good_2014}.
-There is also an open question
-%
-%or from top-down "sanity checks" made in comparison to aggregated survey data.
-
-%This allows us to quantify between-map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.
 
@@
 Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g.
-3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \cite{fritz_geo-wiki:_2012,estes_projected_2013, debats_generalized_2015}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \cite{searchinger_high_2015}.
+3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \cite{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013, debats_generalized_2015}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \cite{searchinger_high_2015}.
 
@@
 
-We used this dataset as a reference layer for evaluating four landcover products representative of the type commonly used in Earth Systems research.
+We used this dataset as a reference layer for evaluating four landcover products representative of the type commonly used in global change studies.
 The first was South Africa's own 30 m resolution 2009 National Landcover map (SA-LC)\cite{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are typically available only for individual countries \cite[e.g.][]{fry_completion_2009}.
@@
 
-We then used these maps to conduct four further analyses typical of Earth System science research.
-The first two related to physical processes, namely the calculation of vegetative carbon stocks and the simulation of evapotranspiration by a hydrological model.
-The second two were socio-economic in nature: the estimation of agriculture yield and production, and an agent-based model assessment of household food security.
+We then used these maps to conduct four further analyses typical of global change research.
+The first two related to physical processes, namely the calculation of vegetative carbon stocks and the simulation of evapotranspiration.
+The second two were socio-economic in nature: the estimation of agriculture yield and production, and an agent-based model simulating household dynamics in dryland agroecosystems.
 The first and third analyses were relatively simple, in that the variable(s) of interest were mapped onto landcover using empirical relationships.
@@
 
-%\begin{figure*}[!b]
-%\centering
-%  \begin{minipage}[b]{.45\linewidth}
-%    \centerline{\includegraphics[scale=0.51]{figures/carbon_veg_scalew.pdf}}
-%    \caption{Caption}
-%    \label{afoto3}
-%  \end{minipage}\hfill%\qquad
-%  \begin{minipage}[b]{.45\linewidth}
-%    \centerline{\includegraphics[scale=0.51]{figures/et_bias_map.pdf}}
-%    \caption{Caption 2}
-%    \label{afoto4}
-%  \end{minipage}
-%\end{figure*}
-
 %\vspace{-0.75 cm}
-\begin{figure}[!h]
+\begin{figure}[h]
 \centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
@@
 
-\subsection{Agent-based simulation of food security}
+\subsection{Agent-based simulation of households in dryland agroecosystems}
 Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}.
@@
 In our example, we used an ABM of household food security that simulates food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}.
-The model is initialized such that each household is allocated its designated share of cropland, and then produces a crop over the course of one to many seasons.
+The model is initialized such that each household is allocated its designated share of cropland based on aggregated farm-level statistics, and then calculates household-level agricultural production based on maize yields simulated by the DSSAT cropping system model \cite{jones_modelling_2003}.
+Yield estimates vary as a function of soil properties, management inputs, and weather.
+The former two factors differ between households, as do the total areas of cropland, thus total production can vary substantially between households.
 Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g.
@@
 We examined three variables: the percent of unallocated cropland, land deficit, and food deficit.
-The percent of unallocated cropland measures how effective the model was in matching household agents to available cropland resources.
+The percent of unallocated cropland what share of the district's total cropland area was not assigned to any household, and is a measure of how effective the model was in matching household agents to available cropland resources.
 Where error was negative (indicating a cropland overestimate by the test maps), the amount of land left unallocated had a straight one-to-one relationship with the percentage of overestimation (Fig.
@@
 
-The second measure, land deficit, describes the amount of land that the model should have assigned but couldn't due to the mismatch between the cropland map and the statistical estimate of cropland holdings.
-Land deficit increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2 (Fig.
+The second measure, land deficit, is calculated from the total area of cropland that should have been allocated to households in each district, but wasn't, due to the mismatch between the cropland map and the statistical estimates of what total cropland holdings should have been (given the number of households and the size class distributions of cropland in the district).
+Land deficit, expressed as a percentage of the total area of cropland in the district (thus values exceeded 100\% when total deficit exceeded cropland area),  increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2 (Fig.
 6B)--and would become infinite in the case of a 100\% underestimate.
-This contrasted with the third measure, food deficit, the shortfall in the average amount of food that should have been produced by each household but wasn't because of the land deficit; this particular deficit increased linearly with the percentage of cropland underestimate (Fig.
+This contrasted with the third measure, food deficit, the percentage shortfall in the average amount of food production that should have been produced by each household but wasn't because of the land deficit; this particular deficit increased linearly with the percentage of cropland underestimate (Fig.
 6C).
@@
 \centerline{\includegraphics[width=.5\textwidth]{figures/figure6.pdf}}
-\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the percent of cropland that should have been allocated to households but wasn't, and C) the food deficit, or the average food production across households resulting from inadequate cropland allocation.
+\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and C) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
 Dot sizes correspond to district numbers, colors represent the landcover map.}
@@
 
-These last few points suggest two final guidelines.
-For high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \cite{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share.
-www.glcn.org} datasets for other cover types).
-But in either case some aggregation--1-25 km, depending on application and whether low bias or high accuracy is the objective--will still likely be necessary to achieve sufficiently low error.
-Additionally, these higher accuracy, lower bias products often represent just a single time period \cite[e.g.][]{fritz_mapping_2015,tuanmu_global_2014}, thus applications requiring higher frequency (e.g.
-annual), large area landcover maps will depend on MODIS or similar sensors, and therefore much coarser aggregations.
-
-%GeoWiki was the most accurate among the large scale landcover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
-
 \begin{table*}[!h]
@@
 \end{table*}%
+
+These last few points suggest two final guidelines.
+For high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \cite{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share.
+www.glcn.org} datasets for other cover types).
+But in either case some aggregation--1-25 km, depending on application and whether low bias or high accuracy is the objective--will still likely be necessary to achieve sufficiently low error.
+Additionally, these higher accuracy, lower bias products often represent just a single time period \cite[e.g.][]{fritz_mapping_2015,tuanmu_global_2014}, thus applications requiring higher frequency (e.g.
+annual), large area landcover maps will depend on MODIS or similar sensors, and therefore much coarser aggregations.
+
+%GeoWiki was the most accurate among the large scale landcover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
 
@@
 We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area (2.2 ha), and preserved the cropland area distributions by multiplying the total number of households by the frequencies.
-To create cropland surfaces for the, we disaggregated all five cropland maps to binary cropland/non-cropland cover types with 100 m resolution, in order to match the survey-recorded average field size of 1 ha.
+To create cropland surfaces for each district, we disaggregated all five cropland maps to binary cropland/non-cropland cover types with 100 m resolution, in order to match the survey-recorded average field size of 1 ha.
 In initializing, the model takes a weighted (by cropland area frequency) random draw of 100 households and places these within a district, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household.
 This process is iterated until all households are assigned cropland, or all available cropland is allocated.
-The model then runs separately for each district, with the household agents planting and managing a maize crop throughout the growing season, with possible final maize yields drawn from a look-up table that relates yield to particular climatic outcomes and management decisions.
-We ran the model for a single season, once per district and when initialized by each of the cropland maps (20 simulations total).
-\\
+The model then runs separately for each district, with the household agents planting and managing a maize crop throughout the growing season, with possible final maize yields drawn from a look-up table that relates yield to particular climatic outcomes and management decisions, as simulated by the DSSAT model.
+We ran each ABM simulation for a single season, once per district and when initialized by each of the cropland maps (20 simulations total).\\
 \indent In calculating map error statistics, and with the exception of the agent-based model results, we use the reference cropland percentages to calculate the weighted mean map error (bias) and mean absolute map error (accuracy).
@@
 
-%The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}.
-The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates.
-Maps based on these analyses underpin many assessments of crop productivity and production \cite[e.g.][]{foley_solutions_2011,licker_mind_2010}.
-
-%The reference dataset covered all of South Africa's field crop estate, from which we filtered out areas of communally owned farmlands and permanent tree crops to have a common basis of comparison across all landcover products, leaving us with map, after conversion to cropland percentage, that covered or 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.
-
-%The difference between total carbon stocks for the country made using any of the cropland maps were within +/-3\% of those based on the reference map, regardless of which cover type was adjacent to cropland (Table S1), as the large uncropped area of South Africa (Fig.
-1) dilutes the errors within the
-%(which is constrained by an absolute distance and the proximity of other agents)
-
-%\section{Digital RCD Analysis} 
 \end{materials}
@@
 
-% bib solution from here
-% http://tex.stackexchange.com/questions/167650/is-there-a-more-recent-bibliography-style-file-bst-for-pnas
-% https://github.com/jburon/pnas2011.bst
 \bibliographystyle{pnas2011} 
 {\footnotesize \bibliography{/Users/lestes/Dropbox/publications/full.bib}}
-
-%\begin{thebibliography}{10}
-%\bibitem{BN}
-%M.~Belkin and P.~Niyogi, {\em Using manifold structure for partially
-%  labelled classification}, Advances in NIPS, 15 (2003).
-%
-%\bibitem{BBG:EmbeddingRiemannianManifoldHeatKernel}
-%P.~B\'erard, G.~Besson, and S.~Gallot, {\em Embedding {R}iemannian
-%  manifolds by their heat kernel}, Geom. and Fun. Anal., 4 (1994),
-%  pp.~374--398.
-%\end{thebibliography}
 

### 2015-12-01 Rmd/build-reports.R

@@ -1,24 +1,0 @@
-# build-reports.R
-# Builds output Rmd files
-
-library(lmisc)
-require(knitr)
-require(markdown)
-
-p_root <- proj_root("SAcropland")
-p_vignettes <- full_path(p_root, "SAcropland/vignettes")
-
-# Create .md, .html, and .pdf files
-knit("vignettes/overview.Rmd", output = "vignettes/overview.md")
-markdownToHTML("vignettes/overview.md", "vignettes/overview.html", 
-               options=c("use_xhml"))
-
-knit("vignettes/compare-landcover.Rmd", 
-     output = "vignettes/compare-landcover.md")
-markdownToHTML("vignettes/compare-landcover.md", 
-               "vignettes/compare-landcover.html", options=c("use_xhml"))
-
-knit("vignettes/landcover-bias-plots.Rmd", 
-     output = "vignettes/landcover-bias-plots.md")
-markdownToHTML("vignettes/landcover-bias-plots.md", 
-               "vignettes/landcover-bias-plots.html", options=c("use_xhml"))

### 2015-12-01 external/ext_R/grid-fields-intersect-analyze.R

@@ -16,7 +16,7 @@ source("/u/sandbox/afmap/R/KMLAccuracyFunctions.R")
 
 # Connection
 drv <- dbDriver("PostgreSQL")
-con <- dbConnect(drv, dbname = "SouthAfricaSandbox", user = "postgis", password = "P0stG1S")
+con <- dbConnect(drv, dbname = "db", user = "user", password = "pw")
 
 # Projection information
 prjsrid <- 97490

### 2015-12-01 external/ext_R/grid-fields-intersect-postgis.R

@@ -34,7 +34,7 @@ if(test == "N") {
 
 # Connection
 drv <- dbDriver("PostgreSQL")
-con <- dbConnect(drv, dbname = "SouthAfricaSandbox", user = "postgis", password = "P0stG1S")
+con <- dbConnect(drv, dbname = "db", user = "user", password = "pw")
 
 prjsrid <- 97490
 prj.sql <- paste("select proj4text from spatial_ref_sys where srid=", prjsrid, sep = "")

### 2015-12-17 paper/SI.tex
Submitted version of manuscript and SI with Tom and Peng's updates

@@
 \captionsetup[figure]{slc=off, font=footnotesize}
-%\usepackage{amsmath}
+\usepackage{amsmath}
 \usepackage{longtable}
@@
 \usepackage[numbers,round,sort,compress]{natbib}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+%\usepackage[noend]{algpseudocode}
+\floatstyle{plain}
+\newfloat{myalgo}{tbhp}{mya}
+
+\newenvironment{Algorithm}[2][tbh]%
+{\begin{myalgo}[#1]
+\centering
+\begin{minipage}{#2}
+\begin{algorithm}[H]}%
+{\end{algorithm}
+\end{minipage}
+\end{myalgo}}
+
 \bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
@@
 
+\subsection{\large Model description}
+
+The following summary follows the Overview, Design and Details protocol \cite{grimm_standard_2006,polhill_using_2008} designed as a standardized method for describing individual-based models and agent-based models.
+
+\subsubsection{Purpose} 
+We developed an agent-based simulation to explore the intra-seasonal dynamics during the maize production of smallholder farmers under climate changes with an emphasis on household-level heterogeneity.
+
+\subsubsection{State variables and scales} 
+The primary analytical components in the model are actors (households) and cells (land).
+We designed a greedy algorithm that allocates land to households based on a heuristic rule that households tend to live near each other to form communities like villages.
+The simulation runs on four districts in South Africa.
+For each district, a set of synthetic household agents is created from the land cover data of that district and the farmer register in Monze District of Zambia.
+We extracted the frequencies of cultivated area from the farmer register using bins (in hectare) of 0-1, 1-2, 2-3, 3-4, 4-5, and 5-10.
+Because each of the land cells is 1 hectare (ha), we considered the households in the same bin having the same integer value of cultivated area.
+For example, households in the bin (0, 1] all have 1 ha of cropland, and households in the bin (5, 10] all have 8 ha of cropland.
+From this distribution of cultivated area (Figure~\ref{fig:histo}), we calculated its mean value as:
+
+\begin{displaymath} 
+\begin{split} & \textit{Mean value of cultivated area} \\ 
+& = \textit{area of 1}^{st}\textit{ bin} * 1 + \textit{area of 2}^{nd}\textit{bin} * 2 + \textit{area of 3}^{rd}\textit{ bin} * 3 + \textit{area of 4}^{th}\textit{ bin} * 4 \\ 
+& \quad  + \textit{area of 5}^{th}\textit{ bin} * 5 + + \textit{area of 6}^{th}\textit{ bin} * (6 + 10)/2 \\ & = 0.3308 *1 + 0.3591 * 2 + 0.1619 * 3 + 0.0835 * 4 + 0.0335 * 5 + 0.0062 * 5 * 8 \\ 
+& = 2.2854 \textit{ (ha)} 
+\end{split}
+\end{displaymath}
+
+\begin{figure}[!ht] 
+\centering 
+\includegraphics[width = 10cm]{figures/histogram.eps} 
+\caption{Histogram of cultivated area in
+the farmer register. } 
+\label{fig:histo} 
+\end{figure}
+
+We then used the mean value of cultivated area to divide the total area of cropland in each district to get the number of households (Table~\ref{tab:numOfHHs}).
+
+\begin{table}[!ht] 
+\centering 
+\caption{Number of households created and their total cultivated area.} 
+\centering
+\begin{tabular}{p{2.5cm}|p{2cm}|p{3cm}|p{3cm}|p{3.5cm}} 
+\hline District & Cropland area (ha) & Calculated number of households & Actual number of household agents & Total cultivated area of household agents \\
+\hline 
+clocolan & 54413.3 & 23810 & 23810 & 54410 \\ 
+ficksburg & 51363.93 & 22475 & 22475 & 51367 \\ 
+fouriesburg & 29402.99 & 12866 & 12865 & 29400 \\ 
+marquard & 55633.12 & 24343 & 24341 & 55615 \\ 
+\hline
+\end{tabular} 
+\label{tab:numOfHHs} 
+\end{table}
+
+Four key household initial characteristics define agents: (1) cultivated area (used primarily for maize production), (2) day of planting maize, (3) soil properties of land cells, and (4) seed type of maize.
+To determine (1), the household agents are assigned 1 ha, 2 ha, 3 ha, 4 ha, 5ha, and 8 ha, according to the frequencies we extracted from the farmer register previously.
+The results in Table 1 show a good matching between the total area of cropland from the land cover data and the total cultivated area of household agents.
+As for (2), (3), and (4), we create a distribution from our household survey that is shown in Figure~\ref{fig:distribution}.
+In addition to the initial characteristics, each household has two internal variables: household ID and harvest amount.
+Each land cell has the following variables: household ID (if assigned), land cover type (cropland or not), and revenue (if harvested).
+The harvest amount of a household is the sum of revenue of cropland cells managed by that household.
+
+\begin{figure}[!ht] 
+\centering 
+\includegraphics[width = 10cm]{figures/distribution.pdf} 
+\caption{Distribution of soil properties, maize type (cultivar), and planting day.
+The soil properties, cultivar, and planting day can be used to find the yield in the look-up table created from the DSSAT cropping system model.
+CU 1 indicates a local maize type and CU 3 indicates a hybrid one.
+Day 303 means the 303\textsuperscript{rd} day of the year.} %Peng: first mentioning of DSSAT
+\label{fig:distribution} \end{figure}
+
+The model runs biweekly for the growing season from October 2007 to April 2008, with all household harvested on April 1\textsuperscript{st} in 2008.
+The four districts in South Africa are represented at a 100m spatial resolution.
+We took the gridded cropland reference map at 1 km and disaggregated it into 100m land cells that are either cropland or not.
+%Peng: first mentioning of the cropland reference map %
+
+\subsubsection{Process overview and scheduling}
+
+We developed an algorithm shown in Algorithm~\ref{alg1}, to allocate the cropland cells to households.
+Our land allocation algorithm first chooses a number of seed households (HHs) in the procedure \mbox{ALLOCATE\_HH} and invokes \mbox{ALLOCATE\_MANY\_FARMLAND} to randomly assign unallocated cropland cells/patches to them.
+Then the algorithm randomly selects an unallocated cropland cell/patch that is adjacent to some already allocated cell/patch, and allocates it to the next household using \mbox{ALLOCATE\_MANY\_FARMLAND}.
+In this way, the households should be located close to each other to form communities.
+This is repeated until there is no unallocated land or there is no household that hasn't been assigned any land.
+\mbox{ALLOCATE\_MANY\_FARMLAND} repeatedly invokes \mbox{ALLOCATE\_FARMLAND} until required amount of cropland has been allocated to a household.
+Comments are denoted by right-pointing triangles.
+
+\begin{Algorithm}[t]{14cm} 
+\begin{algorithmic}[1] 
+\Procedure{allocate\_farmland}{$H$, $P$} \Comment{$H$: household, $P$: patch}
+\State $A \leftarrow $ the area of farmland needed by $H$ 
+\If {$A > $ the area of $P$}
+  \Comment{$P$ is fully occupied by $H$}  
+\State $occupiedRatio(P) \leftarrow 1 $ 
+\Else \State $occupiedRatio(P) \leftarrow (A - 1)$  
+  \Comment{$P$ is partially occupied by $H$}
+\EndIf 
+\State $N \leftarrow $ neighbor farmland (in radius $r$) of $P$  
+\Comment{$r$ is a global parameter of allocation radius} 
+\State $status(N) \leftarrow $ tentative seed patches 
+\EndProcedure 
+\Statex
+
+\Procedure {allocate\_many\_farmland}{$H$, $P$} \Comment{$H$: household, $P$: patch} 
+\State \textbf{Invoke} allocate\_farmland($H$, $P$) 
+\Repeat
+\State $searchRadius \leftarrow 700m $ \Comment{starting from a threshold value} 
+\State $UP \leftarrow $ a randomly selected unoccupied farmland within $searchRadius$ of $P$ 
+\State \textbf{Invoke} allocate\_farmland($H$, $UP$) 
+\If {$A > $ the area of $P$} 
+\State $searchRadius \leftarrow (searchRadius + 100m) $ 
+\EndIf 
+\Until{$H$ is assigned enough farmland $\vee$ $searchRadius == s$} 
+\Comment{$s$ is a global parameter of the maximum search radius} 
+\EndProcedure 
+\Statex
+
+\Procedure {allocate\_hh}{} 
+\State $i \leftarrow 1$ \Comment{the id of current household to be allocated} 
+\Repeat 
+\State $SH \leftarrow $ the $i$th household 
+\State $status(SH) \leftarrow $ seed household 
+\State $SP \leftarrow $ a randomly selected patch 
+\State \textbf{Invoke} allocate\_many\_farmland($SH$, $SP$) 
+\State $i \leftarrow (i+1) $ \Until{$i == numSeed\quad\vee\quad$ there is no unoccupied land}
+\Comment{$numSeed$ is a global parameter of the total number of seed households created during initialization} \Repeat 
+\State $SH \leftarrow $ the $i$th household 
+\State $TSP \leftarrow $ a randomly selected patch so that $status(TSP) == $ tentative seed patch $\wedge$
+$occupiedRatio(TSP) == 0$ 
+\State \textbf{Invoke} allocate\_many\_farmland($SH$, $TSP$) 
+\State $i \leftarrow (i+1) $
+\Until{$i == numHHs\quad\vee\quad$ there is no unoccupied land}
+\Comment{$numHHs$ is the total number of households} 
+\EndProcedure
+
+\end{algorithmic} 
+\caption{Algorithm to allocate cropland patches (cells) to households.} 
+\label{alg1} 
+\end{Algorithm}
+
+Once a cropland cell is allocated to a household agent, its soil property is determined by the characteristics of the household.
+The simulation runs biweekly and when the planting day of a household arrives, that household will plant maize on all of its cropland cells.
+When April 1st arrives, all households will harvest their cropland cells where the yield is calculated using the look-up table from the DSSAT cropping system.
+
+\subsection{\large Design concepts}
+
+\subsubsection{Emergence} 
+Cropland cells are allocated to households with seeding and geospatial proximity searching. The resulting household communities define the scope of labor sharing activities.
+
+\subsubsection{Adaptation}
+%Peng: Not true for the South Africa model:
+% The labor sharing module in the model (however disabled in South
+% Africa simulations) allows the households to adapt to climate changes
+% and food crisis through labor exchanges.
+% 
+\subsubsection{Fitness/objectives} 
+The objective of the model is to evaluate the impact of climate variability on food production, one component of food security.
+An important aspect of climate change impact research in agricultural systems is the ability of farmers to adapt to changing climate conditions.
+This means farmers must have some signal detection mechanism and then some process by which they modify their behavior based on their own experiences, or experiences they learn from others.
+However, in the version of the model used in this paper we reduced the overall complexity to focus primarily on the impact of crop bias in land cover data on model initialization (or rather sensitive the allocation of land to agents is to different land cover datasets).
+In particular, we do not employ agent-interactions or learning in the version of the model used here.
+
+\subsubsection{Prediction} 
+The look-up table assumes no deviation from optimal yields for the specified planting scenarios.
+In other words, households do not take into account the possibility of frost, pests, or other disturbances that affect yields in the real world.
+
+\subsubsection{Interaction} 
+Households interact with the cropland cells through planting and harvesting. Because labor sharing is turned off in the model there is no formal agent-to-agent interaction in the model.
+
+\subsubsection{Sensing} 
+Households do not sense the actions of other agents or environmental conditions in the version of the model used for this manuscript.
+
+\subsubsection{Model calibration} 
+The model was calibrated during two phases: land allocation and food production.
+After the cropland cells are allocated to households, the model is considered well calibrated when all households were allocated their appropriate area of cropland, and all cropland was allocated.
+The land allocation algorithm starts with an initial value of maximum searching scope.
+The model is then run iteratively, increasing the value of maximum searching scope if there are households not allocated any cropland and cropland unallocated.
+When harvesting is finished at April 1st, the total production is compared against the district level census data, and the average yield is compared against the post-harvest survey data.
+
 \FloatBarrier

### 2015-12-17 paper/main.tex
Submitted version of manuscript and SI with Tom and Peng's updates

@@
 
-\title{Landcover Data Limit Our Understanding of Global Change}
+\title{Errors in Landcover Data Limit Our Understanding of Global Change}
 %\title{Our Understanding of Global Change is Built on a Shaky Foundation}
@@
 Figure 3 in][]{foley_solutions_2011}.
-Such maps give an inflated sense of the precision with which such locations can be identified.
-Although the study in this example is careful to report aggregated statistics, and the map-makers stress that their products should not be used at the grid cell scale\footnote{www.earthstat.org\/wp-content\/uploads\/METADATA\_HarvestedAreaYield175Crops.pdf}, it is not inconceivable that policy-makers might make decisions based on their visual interpretations of maps.
+Such maps may give an inflated sense of the precision with which such locations can be identified.
+Although the study in this example is careful to report aggregated statistics and maps the per hectare values, and the map-makers stress that their products should not be used at the grid cell scale\footnote{www.earthstat.org\/wp-content\/uploads\/METADATA\_HarvestedAreaYield175Crops.pdf}, it is not inconceivable that users might make decisions based on their visual interpretations of maps.
 For this reason, it is advisable to aggregate presented maps until their accuracy reaches an acceptable standard (Table 1).

### 2015-12-27 paper/main.tex
Edited tex class to drop PNAS from footers

@@
 %% BASIC CLASS FILE
-\documentclass{pnastwo}
+\documentclass{pnastwo2}
 
@@
 %% For PNAS Only:
-\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
+%\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
 \copyrightyear{2008}
@@
 
-\contributor{Submitted to Proceedings of the National Academy of Sciences
-of the United States of America}
+%\contributor{Submitted to Proceedings of the National Academy of Sciences
+%of the United States of America}
 
@@
 %%% If significance statement need, then can use the below command otherwise just delete it.
-\significancetext{Remote sensing-derived landcover maps underpin much research into global environmental change, particularly in the most rapidly developing regions.
+%\significancetext{Remote sensing-derived landcover maps underpin much research into global environmental change, particularly in the most rapidly developing regions.
 These maps often contain substantial error, yet inadequate validation data makes it impossible to fully quantify map errors and how they affect our understanding of global change processes.

### 2016-03-13 Rmd/c5-accuracy-vs-density.Rmd
Manuscript revised for GCB format

@@
 namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
-gti <- mask(gti, namask)  # apply mask to GTI
+gti <- raster::mask(gti, namask)  # apply mask to GTI
 # cellStats(!is.na(gti), sum)

### 2016-03-13 paper/SI.tex
Manuscript revised for GCB format

@@
 \usepackage{multirow}
-\usepackage[numbers,round,sort,compress]{natbib}
+\usepackage[sort,compress]{natbib}
 \usepackage{algorithm}
@@
 
-\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
+%\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
 % figures float at top of page
@@
 
-The accuracy of the reference vector dataset was assessed by Estes et al \cite{estes_platform_2015}, and described further here.
+The accuracy of the reference vector dataset was assessed by \citet{estes_platform_2016}, and described further here.
 The assessment was undertaken using a sub-sample of 609 1 km$^2$ grid cells, which were selected using a weighted randomized sampling scheme.
@@
 These intersections and differences were used to calculate the area of true positives, false positives, true negatives, and false negatives within each 1 km$^2$ grid cell, which were then used to calculate accuracy measures.
-In this case, accuracy was assessed using two landcover classes: cropland and non-cropland, thus a two class confusion matrix was constructed (Table 1) according to remote sensing classification accuracy guidelines set out by Oloffson et al.
-\cite{olofsson_good_2014}.
+In this case, accuracy was assessed using two landcover classes: cropland and non-cropland, thus a two class confusion matrix was constructed (Table 1) according to remote sensing classification accuracy guidelines set out by \citet{olofsson_good_2014}.
 In this case, the reference data were the visually interpreted crop field presences/absences, and the map data were the field boundary vectors being assessed (i.e.
@@
 
-As a further exploration of uncertainty, we included the full range of variability that resulted from a) using either the 2007 or 2011 map to calculate bias and accuracy, and b) from different levels of cropland percentages used when converting MODIS and GlobCover mixed cropland classes into cropland percentage maps, following Fritz et al.
-\cite{fritz_mapping_2015}.
+As a further exploration of uncertainty, we included the full range of variability that resulted from a) using either the 2007 or 2011 map to calculate bias and accuracy, and b) from different levels of cropland percentages used when converting MODIS and GlobCover mixed cropland classes into cropland percentage maps, following \citet{fritz_mapping_2015}.
 We pooled the residuals from across each of these permutations, for each reference-test map combination, and examined how their overall mean values change with aggregation scale (Fig.
@@
 We used magisterial district boundaries for South Africa (Fig.
-5) to extract absolute residual values for agricultural areas ($>$0.5\% cropland cover) and 2011 reference cropland density values, the means of which respectively served as the response and predictor in the generalized additive model \cite{wood_mgcv:_2001} analysis (main text).
+5) to extract absolute residual values for agricultural areas ($>$0.5\% cropland cover) and 2011 reference cropland density values, the means of which respectively served as the response and predictor in the generalized additive model \citep{wood_mgcv:_2001} analysis (main text).
 
@@
 
-To calculate carbon stocks using the method of Ruesch and Gibbs \cite{ruesch_new_2008}, we simplified their landcover-specific carbon density classes into six categories, based on the similarity of their class types and the fact that they had the same assigned carbon densities.
+To calculate carbon stocks using the method of \citet{ruesch_new_2008}, we simplified their landcover-specific carbon density classes into six categories, based on the similarity of their class types and the fact that they had the same assigned carbon densities.
 Other classes were dropped because of their low level of, or lack of, occurrence.
@@
 All other classes (cropland, shrubland, spare vegetation) were retained.
-Each of these classes has a specific carbon density according to the ecofloristic zone it is found in, of which there are 10 for Africa (Zones 6-9 and 10-15; see \cite{ruesch_new_2008}).
+Each of these classes has a specific carbon density according to the ecofloristic zone it is found in, of which there are 10 for Africa \citep[Zones 6-9 and 10-15; see][]{ruesch_new_2008}.
 We calculated the area of each ecofloristic zone using their polygon boundary maps\footnote{available from cdiac.ornl.gov/ftp/global\_carbon/ecofloristic\_zones.zip}.
@@
 For our crop yield and production analysis, we followed several steps to create the gridded yield and crop production maps.
-First, we calibrated cropland percentage maps so that they matched total cropland areas reported for coarser administrative units, following methods set out by Ramankutty et al \cite{ramankutty_farming_2008}:
+First, we calibrated cropland percentage maps so that they matched total cropland areas reported for coarser administrative units, following methods set out by \citet{ramankutty_farming_2008}:
 
 \begin{enumerate}
- \item We extracted the 2011 reference percentages within each of South Africa's 9 provinces (the same units used by Ramankutty et al \cite{ramankutty_farming_2008}, converted these to proportions and summed them to calculate the ``reported'' cropland areas for each province.
+ \item We extracted the 2011 reference percentages within each of South Africa's 9 provinces \citep[the same units used by][]{ramankutty_farming_2008}, converted these to proportions and summed them to calculate the ``reported'' cropland areas for each province.
 We did the same to calculate provincial cropland areas for each test map;
@@
 
-We then used \emph{tcfa}, per Monfreda et al \cite{monfreda_farming_2008}, to disaggregate magisterial district-level reported maize yields and harvested areas \cite{statistics_south_africa_building_2007}, according to the following steps:
+We then used \emph{tcfa}, per \citet{monfreda_farming_2008}, to disaggregate magisterial district-level reported maize yields and harvested areas \citep{statistics_south_africa_commercial_2007}, according to the following steps:
 
@@
 %\input{figures/et-bias.tex}
-A number of the variables used by the Variable Infiltration Capacity \cite[VIC][]{liang_simple_1994} model are linked to a 0.25$^{\circ}$ resolution, AVHRR-derived landcover map.
+A number of the variables used by the Variable Infiltration Capacity \citep[VIC;][]{liang_simple_1994} model are linked to a 0.25$^{\circ}$ resolution, AVHRR-derived landcover map.
 These include seasonal leaf area index (LAI) phenologies (Fig.
@@
 The look-up table itself was based on a series of yield simulations conducted by the DSSAT cropping system model run over 31 years (1979-2010) for a location in the southern Province of Zambia, corresponding to the region where household survey data were collected.
-The model was run for three dominant soil series\footnote{extracted from a gridded version of the ISRIC-WISE database, available from http://dssat.net/649}, using two open-pollinated and two hybrid cultivars (each pair representing a short- and medium-length growing season variety) whose coefficients were obtained from the inputs files for the PSIMs modeling platform \cite{elliott_parallel_2014}.
+The model was run for three dominant soil series\footnote{extracted from a gridded version of the ISRIC-WISE database, available from http://dssat.net/649}, using two open-pollinated and two hybrid cultivars (each pair representing a short- and medium-length growing season variety) whose coefficients were obtained from the inputs files for the PSIMs modeling platform \citep{elliott_parallel_2014}.
 Each cultivar was simulated under planting dates ranging from October 15th until January 30th, varying by 15 day increments, with row spacing fixed at 90 cm and planting density at 3.7 plants m$^2$, and 5 kg ha$^{-1}$ applied at planting.
-Models were run using weather data extracted from a bias-corrected version of the Princeton Global Meteorological forcing dataset \cite{sheffield_development_2006,chaney_spatial_2014,estes_changing_2014}.
+Models were run using weather data extracted from a bias-corrected version of the Princeton Global Meteorological forcing dataset \citep{sheffield_development_2006,chaney_spatial_2014,estes_changing_2014}.
 
@@
 
-The following summary follows the Overview, Design and Details protocol \cite{grimm_standard_2006,polhill_using_2008} designed as a standardized method for describing individual-based models and agent-based models.
+The following summary follows the Overview, Design and Details protocol \citep{grimm_standard_2006,polhill_using_2008} designed as a standardized method for describing individual-based models and agent-based models.
 
@@
 %\section*{References}
-\bibliographystyle{pnas2011} 
+\bibliographystyle{gcbnourl} 
 {\footnotesize \bibliography{/Users/lestes/Dropbox/publications/full.bib}}

### 2016-03-13 paper/main.tex
Manuscript revised for GCB format

@@
 %% Version2: Oct 04, 2013
-
+%
 %% BASIC CLASS FILE
-\documentclass{pnastwo2}
-
+%\documentclass{pnas}
+%
 %% ADDITIONAL OPTIONAL STYLE FILES Font specification
-
+%
 %\usepackage{pnastwoF}
 %\usepackage{cite}
-\usepackage[numbers,round,sort,compress]{natbib}
-\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
-\usepackage{lineno}
-\setlength\linenumbersep{2pt}
+%\usepackage[numbers,round,sort,compress]{natbib}
+%\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
+%\usepackage{lineno}
+%\setlength\linenumbersep{2pt}
 %\usepackage{placeins}
@@
 %\setlength{\textfloatsep}{5pt plus 1.0pt minus 2.0pt}
+%\usepackage{multirow}
+%\setlength{\parskip}{1cm plus4mm minus3mm}
+%\usepackage{makecell}
+%\usepackage{boldline}
+%\setcellgapes{3pt}
+%\usepackage[flushleft]{threeparttable}
+%
+%% OPTIONAL MACRO DEFINITIONS
+%\def\s{\sigma}
+%%%%%%%%%%%%
+%% For PNAS Only:
+%\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
+%\copyrightyear{2008}
+%\issuedate{Issue Date}
+%\volume{Volume}
+%\issuenumber{Issue Number}
+%\setcounter{page}{2687} %Set page number here if desired
+%%%%%%%%%%%%
+%
+%
+%
+\documentclass[a4paper]{article}
+\usepackage[affil-it]{authblk}
+
+\usepackage{graphicx}
+\usepackage{tabularx}
+\usepackage{epstopdf}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsfonts}
+\usepackage{amsthm}
+%\usepackage{endfloat}
+\usepackage{float}
+\usepackage[sort,compress]{natbib}
+%\bibpunct{[}{]}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
+\usepackage{endnotes}
+\usepackage{setspace}
+\usepackage{verbatim}
+\usepackage[left=2.5cm, right=2.5cm, bottom=2cm, top=2cm]{geometry}
+\usepackage{times}
+\usepackage{helvet}
+\usepackage{courier}
+%\usepackage{mathtime}
+\usepackage{bm}
+\usepackage{url}
+%\usepackage{babel}
+\usepackage{dcolumn}
 \usepackage{multirow}
-\setlength{\parskip}{1cm plus4mm minus3mm}
 \usepackage{makecell}
@@
 \usepackage[flushleft]{threeparttable}
+\usepackage{wrapfig}
+\newcommand{\citetapos}[1]{\citeauthor{#1}'s \citeyearpar{#1}}
+%%%%%
 
-%% OPTIONAL MACRO DEFINITIONS
-\def\s{\sigma}
-%%%%%%%%%%%%
-%% For PNAS Only:
-%\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-\copyrightyear{2008}
-\issuedate{Issue Date}
-\volume{Volume}
-\issuenumber{Issue Number}
-%\setcounter{page}{2687} %Set page number here if desired
-%%%%%%%%%%%%
+\usepackage[
+  breaklinks=true,
+  colorlinks=true,
+  linkcolor=blue,anchorcolor=blue,
+  citecolor=blue,filecolor=blue,
+  menucolor=blue,pagecolor=blue,
+  urlcolor=blue]{hyperref}
 
-\begin{document}
+% Let's add todonotes and comments:
+%\usepackage[draft]{todonotes}
+%
+%% Select what to do with command \comment:  
+%% \newcommand{\comment}[1]{}  %comment not showed
+%\newcommand{\comment}[1]
+%{{[\bfseries \color{blue} #1]}} %comment showed
+
+\usepackage{lineno}
+\usepackage{float}
+\usepackage[anythingbreaks]{breakurl}
+
+%\begin{document}
 %\hspace*{\fill}
@@
 
-\author{Lyndon Estes{\affil{1}{Woodrow Wilson School, Princeton University, Princeton, NJ USA}\affil{2}{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}},
-Peng Chen\affil{3}{Indiana University, Bloomington, IN USA},
-Stephanie Debats\affil{2}{},
-Tom Evans\affil{3}{},
-Stefanus Ferreira\affil{4}{GeoTerraImage, Pretoria, RSA}{},
-Gabrielle Ragazzo\affil{5}{Ecology and Evolutionary Biology, Princeton University, Princeton, NJ USA},
-Justin Sheffield\affil{2}{},
-Adam Wolf\affil{5}{},
-Eric Wood\affil{2}{},
-\and
-Kelly Caylor\affil{2}{}}
+%\author{Lyndon Estes{\affil{1}{Woodrow Wilson School, Princeton University, Princeton, NJ USA}\affil{2}{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}},
+%Peng Chen\affil{3}{Indiana University, Bloomington, IN USA},
+%Stephanie Debats\affil{2}{},
+%Tom Evans\affil{3}{},
+%Stefanus Ferreira\affil{4}{GeoTerraImage, Pretoria, RSA}{},
+%Gabrielle Ragazzo\affil{5}{Ecology and Evolutionary Biology, Princeton University, Princeton, NJ USA},
+%Justin Sheffield\affil{2}{},
+%Adam Wolf\affil{5}{},
+%Eric Wood\affil{2}{},
+%\and
+%Kelly Caylor\affil{2}{}}
+
+\author[1,2]{Lyndon Estes}
+\author[3]{Peng Chen}
+\author[1]{Stephanie Debats}
+\author[3]{Tom Evans}
+\author[4]{Stefanus Ferreira}
+\author[5]{Gabrielle Ragazzo}
+\author[1]{Justin Sheffield}
+\author[5]{Adam Wolf}
+\author[1]{Eric Wood}
+\author[1]{Kelly Caylor}
+
+\affil[2]{Woodrow Wilson School, Princeton University, Princeton, NJ USA}
+\affil[1]{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}
+\affil[3]{Indiana University, Bloomington, IN USA}
+\affil[4]{GeoTerraImage, Pretoria, RSA}
+\affil[5]{Ecology and Evolutionary Biology, Princeton University, Princeton, NJ USA}
 
@@
 
+\begin{document}
 \maketitle
 
-\begin{article}
+%\begin{article}
 \begin{abstract}
@@
 
-\keywords{landcover | bias | remote sensing | agriculture | crop yield | harvested area | carbon | agent-based model | landscape}
+%\keywords{landcover | bias | remote sensing | agriculture | crop yield | harvested area | carbon | agent-based model | landscape}
 
@@
 
-\dropcap{T}he functioning of the Earth System is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface.
-Human endeavors are strongly governed by and shape landcover, whether it be felling forests for timber or burning savannas to rejuvenate forage, and landcover is a primary factor in many climate and biogeochemical processes \cite{lambin_dynamics_2003}.
-Our increasing modification of the Earth's surface \cite{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover.
-To fully understand these processes and the nature of global change, it is essential to know the nature and distribution of landcover.
+\section*{Introduction}
+The functioning of the Earth System is fundamentally connected to the characteristics of landcover \citep{lambin_modelling_1997}, the physical constituents of the terrestrial surface.
+Human endeavors are strongly governed by and shape landcover, and the characteristics of landcover govern many climate and biogeochemical processes \citep{lambin_dynamics_2003}.
+Our increasing modification of the Earth's surface \citep{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover.
+To fully understand these processes and the nature of global change, an accurate understanding of the nature and distribution of landcover is essential.
 
-This importance is understood by a growing number of social, economic, and natural scientists, who are using landcover data to advance understanding of food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, demographic shifts \cite{linard_assessing_2010}, and other important facets of global change.
+This importance is understood by a growing number of social, economic, and natural scientists, who are using landcover data to advance understanding of food security \citep{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \citep{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \citep{newbold_global_2015, luoto_predicting_2004}, demographic shifts \citep{linard_assessing_2010}, and other important facets of global change.
 
@@
 Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations.
-The reason for this is that landcover data can only practically be derived from satellite imaging, which has several important constraints--related to both physics and analytics--that propagate mapping errors.
-First, in many regions the cover types of interest are smaller than the sensor resolution \cite[e.g.
-smallholder's farms][]{jain_mapping_2013,debats_generalized_2015,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \cite{sweeney_mapping_2015}, which are factors that increase mapping complexity \cite{yu_meta-discoveries_2014}.
+The reason for this is that landcover data can only practically be derived from satellite imaging, which has several important constraints, related to both physics and analytics, that propagate mapping errors.
+First, in many regions the cover types of interest are smaller than the sensor resolution \citep[e.g.
+smallholder's farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \citep{sweeney_mapping_2015}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
 Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties (e.g.
-agriculture \cite{debats_generalized_2015,estes_platform_2015}) and thus be difficult for the classifier to distinguish, while discretizing a continuous cover type (e.g.
-dividing a forest into different canopy cover classes) can also promote classification error, particularly near class boundaries \cite{foody_status_2002}, as well as confusion about the actual extent of the cover type \cite{sexton_conservation_2015}.
-Furthermore, class definitions often vary between maps, complicating inter-comparison \cite{kuemmerle_challenges_2013}.
-Third, landcover map series are often used to detect changes \cite[e.g.][]{gross_monitoring_2013}, or may use within-class seasonality as a means for discriminating between classes (e.g.
-phenology of cropping versus surrounding vegetation \cite{sweeney_mapping_2015}), but seasonal variability and landcover changes can be easily confused.
-Given these multiple sources of error, landcover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}.
+agriculture \citep{debats_generalized_2016,estes_platform_2015}) and thus be difficult for the classifier to distinguish, while discretizing a continuous cover type (e.g.
+dividing a forest into different canopy cover classes) can also promote classification error, particularly near class boundaries \citep{foody_status_2002}, as well as confusion about the actual extent of the cover type \citep{sexton_conservation_2015}.
+Furthermore, class definitions often vary between maps, complicating inter-comparison \citep{kuemmerle_challenges_2013}.
+Third, landcover map series are often used to detect changes \citep[e.g.][]{gross_monitoring_2013}, or may use within-class seasonality as a means for discriminating between classes (e.g.
+phenology of cropping versus surrounding vegetation \citep{sweeney_mapping_2015}), but seasonal variability and landcover changes can be easily confused.
+Given these multiple sources of error, landcover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \citep{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}.
 These errors limit our ability to obtain granular, mechanistic understanding of global change processes.
 
-These problems with landcover products is known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \cite[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}.
+These problems with landcover products are known \citep{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \citep[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}.
 What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes.
-Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are typically unavailable for Africa and other rapidly changing regions \cite{see_improved_2015}, which have also been fairly sparsely mapped \cite{yu_meta-discoveries_2014}.
-Our current understanding of map accuracy over such areas is therefore based on a small number of maps \cite{yu_meta-discoveries_2014}, which were assessed using either bottom-up tests made with a relatively small number of ground truth points, or from top-down "sanity checks" made in comparison to aggregated survey data.
-In terms of the former, there are well-developed, best-practice guidelines for selecting and using ground-truth samples to robustly quantify map error, but these are not commonly applied by the broader community of landcover map users (often due to lack of data), which makes it challenging to compare how error varies within and between maps \cite{stehman_global_2012, olofsson_making_2013,olofsson_good_2014,foody_status_2002}.
-The latter form of assessment are made at coarser administrative levels relative to reported statistics \cite[e.g.][]{fritz_comparison_2010}, which can also be suspect \cite{carletto_emperor_2013,fao_action_2013}.
+Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are typically unavailable for Africa and other rapidly changing regions \citep{see_improved_2015}, which have also been fairly sparsely mapped \citep{yu_meta-discoveries_2014}.
+Our current understanding of map accuracy over such areas is therefore based on a handful of maps \citep{yu_meta-discoveries_2014}, which were assessed using either bottom-up tests made with a relatively small number of ground truth points, or from top-down "sanity checks" made in comparison to aggregated survey data.
+In terms of the former, there are well-developed, best-practice guidelines for selecting and using ground-truth samples to robustly quantify map error, but these are not commonly applied by the broader community of landcover map users (often due to lack of data), which makes it challenging to compare how error varies within and between maps \citep{stehman_global_2012, olofsson_making_2013,olofsson_good_2014,foody_status_2002}.
+The latter form of assessment is made at coarser administrative levels relative to reported statistics \citep[e.g.][]{fritz_comparison_2010}, which can also be suspect \citep{carletto_emperor_2013,fao_action_2013}.
 
-Since it is difficult to fully quantify landcover map errors, it is even more challenging to gauge their impact on downstream analyses, in which there is substantial risk of error amplification\cite{kuemmerle_challenges_2013}.
-There has been some work examining how map error influences climate simulations \cite{ge_impacts_2007}, agricultural land use patterns \cite{schmit_limitations_2006}, carbon flux measurements \cite{quaife_impact_2008}, human population estimates \cite{linard_assessing_2010}, and species distribution models \cite{tuanmu_global_2014}, but these either use simulated landcover errors \cite{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \cite{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}.
-One exception is a Belgian study \cite{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover errors bias measurements of agricultural land use patterns, but the study extent was fairly small and the validation data were discontiguous.
+Since it is difficult to fully quantify landcover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification\citep{kuemmerle_challenges_2013}.
+There has been some work examining how map error influences climate simulations \citep{ge_impacts_2007}, agricultural land use patterns \citep{schmit_limitations_2006}, carbon flux measurements \citep{quaife_impact_2008}, human population estimates \citep{linard_assessing_2010}, and species distribution models \citep{tuanmu_global_2014}, but these either use simulated landcover errors \citep{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \citep{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}.
+One exception is a Belgian study \citep{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover errors bias measurements of agricultural land use patterns, but the extent of this study was fairly small and the validation data were discontiguous.
 
 Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g.
-3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \cite{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013, debats_generalized_2015}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \cite{searchinger_high_2015}.
+3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \citep{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013,debats_generalized_2016}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \citep{searchinger_high_2015}.
 
-In this study, we take advantage of this recent growth in high resolution data to address the call to more thoroughly assess errors and in landcover maps \cite{kuemmerle_challenges_2013, olofsson_good_2014,olofsson_global_2012}, and further examine how these errors might impact our understanding of change in the world's most dynamic regions.
+In this study, we take advantage of this recent growth in high resolution data to address the call to more thoroughly assess errors and in landcover maps \citep{kuemmerle_challenges_2013, olofsson_good_2014,olofsson_global_2012}, and further examine how these errors might impact our understanding of change in the world's most dynamic regions.
 We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through downstream studies investigating both the physical and socioeconomic components of global change.
-Our objective is to provide scientists and policy-makers who use landcover maps an up-to-date assessment of their appropriate uses and limitations.
+Our objective is to provide scientists and policy-makers who use landcover data an up-to-date assessment of their appropriate uses and limitations.
 
 \vspace{-0.5 cm}
-\section{Overview of study area and analyses}
-In the late 2000s, the South African government commissioned a cropland map that was made by manually interpreting and digitizing fields visible within high resolution satellite imagery \cite{fourie_better_2009}.
+\section*{Materials and Methods}
+\subsection*{Overview of study area and reference data}
+In the late 2000s, the South African government commissioned a cropland map that was made by manually interpreting and digitizing fields visible within high resolution satellite imagery \citep{fourie_better_2009}.
 The resulting vectorized field boundaries provide highly accurate data on field sizes and distribution for the period 2009-2011.
-Measured in terms of its ability to distinguish crop fields from other cover types, this dataset is 97\% percent accurate at 1 km resolution, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\%  \cite[][SI]{estes_platform_2015}.
+Measured in terms of its ability to distinguish crop fields from other cover types, this dataset is 97\% percent accurate at 1 km resolution, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\% \citep[][SI]{estes_platform_2016}.
 
-This dataset is particularly valuable because South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \cite{searchinger_high_2015}, yet is notably lacking trustworthy maps of existing agriculture land \cite{fritz_comparison_2010}.
-Moreover, South Africa's large agricultural sector represents the diversity of farming systems found throughout SSA, ranging from large commercial operations to smallholder farms \cite{hardy_rainfed_2011,estes_using_2014}.
+This dataset is particularly valuable because South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \citep{searchinger_high_2015}, yet is notably lacking trustworthy maps of existing agriculture land \citep{fritz_comparison_2010}.
+Moreover, South Africa's large agricultural sector represents the diversity of farming systems found throughout SSA, ranging from large commercial operations to smallholder farms \citep{hardy_rainfed_2011,estes_using_2014}.
 
-We used this dataset as a reference layer for evaluating four landcover products representative of the type commonly used in global change studies.
-The first was South Africa's own 30 m resolution 2009 National Landcover map (SA-LC)\cite{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are typically available only for individual countries \cite[e.g.][]{fry_completion_2009}.
-The second and third were respectively the 300 m GlobCover 2009 \cite{arino_global_2012} and 500 m resolution 2011 MODIS Landcover products, which are widely used global-scale products \cite[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
-The fourth dataset was the new 1 km GeoWiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}, which incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
+\subsubsection*{Assessment of landcover map accuracy}
+We used the vectorize field boundaries as a reference layer for evaluating four landcover products representative of the type commonly used in global change studies.
+The first was South Africa's own 30 m resolution 2009 National Landcover map \citep[SA-LC][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are typically available only for individual countries \citep[e.g.][]{fry_completion_2009}.
+The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution 2011 MODIS Landcover products, which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
+The fourth dataset was the new 1 km GeoWiki hybrid-fusion cropland map for Africa \citep{fritz_mapping_2015}, which incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
 Since GeoWiki provides a continuous bounded value of landcover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.
 
+To develop the cropland reference map (see SI for accuracy assessment details), we intersected the original field boundary vector maps with a 1 km grid, and calculated the percent of each cell occupied by fields to create the gridded cropland reference map.
+In the resulting grid, we masked out areas classified as communal farmland, because only their outer perimeters were digitized \citep{fourie_better_2009}, which posed a risk of overestimating cropland extent.
+We also excluded the permanent tree crop class, to avoid confusion with landcover maps that exclude these in their cropland classes.
+For the same reason, we used other high resolution landcover maps for selected provinces of South Africa to mask out commercially afforested areas and sugarcane plantations, which were also not included in the reference map.
+The resulting masked reference grid covered 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ was cropland.
+
+We extracted the cropland classes from SA-LC, MODIS, and GlobCover, and converted these into percent cropland estimates at 1 km resolution.
+Both MODIS and GlobCover had mixed pixel classes, thus, following the approach used in creating the GeoWiki maps \citep{fritz_mapping_2015} we used the upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages.
+We used the mean map for the main analysis, but estimated error variability using all three versions (SI).
+
 We first evaluated the quality of the landcover product-derived percent cropland maps (hereafter referred to as the ``test maps'').
-Instead of the standard confusion matrix-based accuracy assessment metrics \cite{olofsson_good_2014,olofsson_making_2013}, which apply to categorical landcover maps, we assessed the bias and accuracy of the test maps based on the gridded residuals that resulted when each test map was subtracted from the reference map.
+Instead of the standard confusion matrix-based accuracy assessment metrics \citep{olofsson_good_2014,olofsson_making_2013}, which apply to categorical landcover maps, we assessed the bias and accuracy of the test maps based on the gridded residuals that resulted when each test map was subtracted from the reference map.
 Here bias is the mean residual value, weighted by the density of reference cropland, and accuracy is the mean of the absolute values of residuals (also weighted by cropland density), thus accuracy is inversely related to this value.
 We calculated these metrics for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km, in order to evaluate how bias and accuracy changes with scale.
+For these aggregated maps, we applied a further weight in calculating error metrics, the number of pixels contributing to each aggregated pixel, to prevent pixels close to national boundaries or where non-target cover types were masked out from having outsize influence on the statistics.
+
 We also assessed how landcover pattern impacts map performance by modeling the correlation between map accuracy and cropland density.
+To evaluate this relationship, we used magisterial district boundaries (n=354, mean area=3,445 km$^2$; SI) to provide a landscape-scaled unit for calculating characteristic cover density.
+We filtered out pixels with $<$0.05\% (0.5 ha) cropland, to prevent the much larger areas of non-agricultural districts from dominating the signal, extracted the absolute values of test map errors and corresponding reference cropland percentages, and calculated their district-wide means.
+The relationship between mean absolute error (response) and cropland density (predictor) was then modeled using a generalized additive model \citep{hastie_generalized_1990}, with each district weighted by its number of agricultural pixels \citep{wood_mgcv:_2001}.
+To account for potential spatial autocorrelation, we fit a two-dimensional smoothing spline to the coordinates of each district's centroid.
 
+\subsection*{Impact of map error on downstream analyses}
 We then used these maps to conduct four further analyses typical of global change research.
@@
 
+\subsubsection*{Biophysical analyses}
+\emph{Estimating vegetative carbon stocks: }
+To understand the carbon cycle and climate forcing due to land use change, it is important to have accurate, high resolution maps of vegetative carbon stocks \citep[][]{searchinger_high_2015}.
+One widely used vegetative carbon dataset is that of \citep{ruesch_new_2008}, who mapped estimated carbon density values for different vegetation types to the classes of a global landcover product.
+The resulting data were intended to provide a baseline for climate policy by the Intergovernmental Panel on Climate Change (IPCC), as well as input to other land use and biogeochemical analyses \citep{ruesch_new_2008}.
+
+%\vspace{-1 cm}
+We followed this method to create vegetative carbon maps for South Africa.
+Since our cropland percentage map provided no information on surrounding cover, we developed several variants representing potential surrounding cover types by assigning the average carbon densities of five biomes (forest, secondary forest, shrubland, grassland, and sparse vegetation \citep{ruesch_new_2008}) to the non-cropland fraction of our maps.
+These hypothetical maps represented the range in potential carbon densities, and allowed us to investigate how carbon estimates can vary as a function of i) test map errors and ii) the properties of neighboring cover types.
+We multiplied cropland densities by cropland fractions and added these to each of the five other densities multiplied by the residual non-cropland fractions to create five different carbon density maps.
+We aggregated each carbon map up to the five coarser resolutions for scaling comparisons (SI).
+
+To assess carbon estimation error, we subtracted test map-derived carbon maps from those based on the reference map, and calculated bias and accuracy scores using the method described for the cropland maps (see previous section).
+
+\vspace{4 pt}
+\noindent\emph{Evapotranspiration estimates: }
+Accurate estimation of hydrological fluxes is critical to understanding how land-atmosphere interactions impact the climate system and runoff \citep{liang_simple_1994}.
+Land surface hydrological models are used to simulate these processes, and depend on landcover maps to provide information on the characteristics of vegetation and other materials covering the surface, as these govern the rates of runoff, infiltration, and evapotranspiration.
+We used the Variable Infiltration Capacity \citep[VIC;][]{liang_simple_1994} land surface hydrology model run with the Africa Flood and Drought Monitor's meteorological data \citep{sheffield_drought_2013} to produce monthly gridded evapotranspiration estimates for South Africa for the years 1979-2010 at 25 km resolution.
+VIC's landcover scheme (derived from AVHRR) provides values for leaf area index (LAI), plant rooting depth, aerodynamic roughness, and several other variables that the modeluses to partition water vapor fluxes into their evaporative and transpirative components.
+We adjusted VIC's base map so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25$^{\circ}$ resolution), then ran one instance of VIC for each of the five landcover schemes.
+We then compared the mean annual ET produced by the reference map variant with those from the test maps to assess the degree to which map errors impact evapotranspiration values.
+
+\subsubsection*{Socio-economic analyses}
+\noindent\emph{Gridded crop yield and production data: }
+The spatial variability of crop productivity and production is critical for understanding a host of social, economic, and environmental issues, such as food security, trade, and the potential for agricultural expansion and intensification \citep{licker_mind_2010,monfreda_farming_2008}.
+The most reliable source of such data are national to sub-national agricultural statistics, which are provided for relatively coarse-scaled administrative boundaries.
+To obtain higher spatial resolutions, efforts have been made to disaggregate these statistics using gridded landcover data \citep{ramankutty_farming_2008,monfreda_farming_2008}.
+These disaggregated datasets, which are constrained to match the agricultural statistics within the boundaries of the areas for which they are reported, have seen increasing use because they are considered to be more accurate than single source methods, and provide consistent data on which to base studies of global change \citep{ramankutty_farming_2008, see_improved_2015}
+
+We used these methods to disaggregate the harvested area for maize (South Africa's largest crop \citep{estes_projected_2013}) on top of our reference and cropland maps, followed by yields, which were assigned to cells having harvested areas greater than zero.
+The first step in this process entails adjusting cropland percentages so that their totals match census-derived cropland area estimates \citep{ramankutty_farming_2008}.
+In place of census statistics, we used the reference map to calculate total cropland areas for South Africa's nine provinces, then adjusted the pixel-wise cropland percentages in the four test maps so that their province-wise sums matched these totals (SI).
+We then followed \citetapos{monfreda_farming_2008} procedure for disaggregating maize \citep[South Africa's largest crop;][]{estes_comparing_2013} planted area and yields onto the reference and adjusted test maps.
+The necessary statistics were obtained from magisterial district-level agricultural censuses conducted for the year 2007 \citep{statistics_south_africa_commercial_2007}.
+
+We then used these two layers to calculate maize production, and further aggregated the yield and production grids to 5, 10, 25, 50, and 100 km resolutions before quantifying the bias and accuracy of each test map's yield and production values.
+In this case, we could not convert cell-wise errors into percentages of the reference map values (because many cells had zero values for one map but not the other), so we calculated bias and accuracy from the map residuals and then normalized their values to the reference map means.
+
+%\indent To create the gridded maize yield and production maps, we followed Ramankutty et al.
+\citep{ramankutty_farming_2008} to first calibrate the test cropland maps against cropland areas reported by administrative-level agricultural censuses.
+
+\vspace{4 pt}
+\noindent\emph{Agent-based simulation of household food security: }
+Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \citep{berger_creating_2006}.
+To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistically represents the social and biophysical features of the study region \citep{berger_creating_2006}.
+
+In our example, we used an ABM of household food security that simulates food production by individual farming households (the agents) in response to weather and management capacity \citep{chen_dependency_2013}.
+The model is initialized such that each household is allocated a designated share of cropland derived from household survey statistics, and then calculates annual household-level agricultural production based on maize yields simulated by the DSSAT cropping system model \citep{jones_modelling_2003}.
+Yield estimates vary as a function of soil properties, management inputs, and weather.
+The former two factors differ between households, as do the total areas of cropland, thus total production can vary substantially between households.
+The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity.
+To achieve this, the model first takes a weighted (by cropland area frequency) random draw of 100 households and places these within the simulated landscape, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household.
+This process is iterated until all households are assigned cropland, or all available cropland is allocated.
+The model is thus considered to be well-calibrated when all households are allocated their appropriate area of cropland, and all cropland is allocated to households.
+
+Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g.
+districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of individual farmers.
+To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, SI) in the eastern part of the country with similar climate and 28-45\% of their areas devoted to cropland.
+The frequency distributions of household cropland holdings were derived from household survey data obtained from Zambia (no equivalent data were available for South Africa).
+We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area (2.2 ha), and preserved the cropland area distributions by multiplying the total number of households by the frequencies.
+To create cropland surfaces for each district, we disaggregated all five cropland maps to binary cropland/non-cropland cover types with 100 m resolution, in order to match the survey-recorded average field size of 1 ha.
+The model was then run separately for each district, with the household agents planting and managing a maize crop throughout the growing season, with possible final maize yields drawn from a look-up table that relates yield to particular climatic outcomes and management decisions, as simulated by the DSSAT model.
+
+We ran each ABM simulation for a single season, once per district and when initialized by each of the cropland maps (20 simulations total). 
+To examined how map errors impacted the land allocation process and household food production estimates, we calculated three variables: the percent of unallocated cropland, land deficit, and food deficit.
+The percent of unallocated cropland is the share of the district's total cropland area that was not assigned to any household, which measures how effective the model was in matching household agents to available cropland resources.
+Land deficit is the total area of cropland that should have been allocated to households in each district but wasn't, due to the mismatch between the cropland map and the survey-based estimates of what total cropland holdings should have been.
+Food deficit is the percentage shortfall in the average amount of food production that should have been produced by each household but wasn't because of the land deficit.
+
 \vspace{-0.5 cm}
-\section{Map quality}
-\subsection{Bias and accuracy}
-
-Our reference map indicated that crop fields covered nearly 10\% of the total study area in the 2009-2011 time period.
+\section*{Results}
+\subsection*{Map quality}
+\subsubsection*{Bias and accuracy}
+Our reference cropland map indicated that crop fields covered nearly 10\% of the total study area in the 2009-2011 time period.
 Subtracting each test map from the reference maps created pixel-wise residuals, where negative and positive values respectively represent overestimates and underestimates by the test map (Fig.
@@
 
-%\end{wrapfigure}
-
-%\begin{figure*}[htp]
-%\centering
-%    \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}\quad
-%     \subfigure[random caption 1]{\includegraphics[scale=0.49]{figures/bias_map.pdf}}
-%\end{figure*}
-%\linenumbers
 The most pronounced errors were in the MODIS and GlobCover maps, which showed large positive residuals in the center of the country where cropland is most concentrated (blue areas in Fig.
@@
 
-\subsection{Error and landcover density}
-Given that landcover characteristics can influence landcover map quality \cite{see_improved_2015, estes_platform_2015, debats_generalized_2015}, we set out to elucidate the relationship between cropland configuration and map error.
-We first calculated the average 1 km reference cropland percentage within each of South Africa's 354 magisterial districts (the finest administrative unit, averaging 3,445 km$^2$; SI), which yielded a landscape-scaled metric of characteristic cropland density, and similarly calculated the district-wise MAE for each test map.
-
-\begin{figure*}[hb]
+\begin{figure*}[!ht]
 %\begin{wrapfigure}{c}{1\textwidth}
-\centerline{\includegraphics[width=0.95\textwidth]{figures/figure1.pdf}}
+\centerline{\includegraphics[width=1\textwidth]{figures/figure1.pdf}}
 \vspace{-0.15 cm}
@@
 
-We then used a generalized additive model to evaluate the shape of the relationship between district MAE and cropland density.
-The model shows that map accuracy is typically lowest at intermediate levels of cropland density (50-60\% cover) for all but the GlobCover map (where accuracy continues to decline with cropland cover), and is highest when the landscape is dominated either by cropland or by another type (Fig.
+\subsubsection*{Error and landcover density}
+The generalized additive model revealed primarily non-linear relationships between district MAE and cropland density that were best approximated by a first order polynomial function of cropland density (for all four cropland maps: p$<$0.001 on both terms of quadratic and on smoothing function applied to district centroids; $>$ 85\% deviance explained).
+Map accuracy was typically lowest at intermediate levels of cropland density (50-60\% cover) for all but the GlobCover map (where accuracy continues to decline with cropland cover), and was highest where the landscape is dominated either by cropland or by another type (Fig.
 2).
@@
 
-\section{The impact of map error on physical analyses}
-\subsection{Estimating vegetative carbon stocks}
-To understand the carbon cycle and climate forcing due to land use change, it is important to have accurate, high resolution maps of vegetative carbon stocks \cite[][]{searchinger_high_2015}.
-One widely used vegetative carbon dataset is that of \cite{ruesch_new_2008}, who mapped estimated carbon density values for different vegetation types to the classes of a global landcover product.
-The resulting data were intended to provide a baseline for climate policy by the Intergovernmental Panel on Climate Change (IPCC), as well as input to other land use and biogeochemical analyses \cite{ruesch_new_2008}.
-
 %\vspace{-0.75 cm}
+%\begin{wrapfigure}{r}{0.5\textwidth}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
+%\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/biases_md_lnorm_gam_mu0.pdf}}
+\centerline{\includegraphics[width=.7\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
 \caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
@@
 \end{figure}
+%\end{wrapfigure}
+%\vspace{-0.75 cm}
 
-%\vspace{-1 cm}
-We followed this method to create vegetative carbon maps for South Africa.
-Since our maps represented cropland percentage, we developed several variants by assigning the carbon densities of different vegetation types (forest, secondary forest, shrubland, grassland, and sparse vegetation \cite{ruesch_new_2008}) to the non-cropland fraction of our maps.
-These hypothetical maps represented the range in potential carbon densities, and allowed us to investigate how carbon estimates can vary as a function of i) test map errors and ii) the properties of neighboring cover types.
-To assess carbon estimation error, we subtracted test map-derived carbon maps from those based on the reference map, and calculated bias and accuracy scores.
-
-The spatial patterns of test map errors transmitted into carbon estimation errors, but the sign varied as a function of the density of carbon adjacent to croplands (SI).
+\subsection*{The impact of map error on physical analyses}
+\subsubsection*{Carbon estimates}
+The spatial patterns of test map errors transmitted into substantial carbon estimation errors, with the sign varying as a function of the density of carbon adjacent to croplands (SI).
 Where cropland was underestimated and the surrounding cover type was more carbon dense than cropland, carbon density was overestimated, but when the cover type was less dense than croplands (e.g.
@@
 
-In terms of accuracy, MAE values were essentially the same as bias magnitudes, except for GeoWiki's, which were twice as large.
-The average MAE across vegetation classes was 47\% at 1 km, dropping to $<$10 only with 25 km of aggregation.
-In contrast, SA-LC's carbon estimates were twice as accurate at 1 km, and were slightly more accurate up to 25 km of aggregation were GeoWiki achieved parity.
-
 %\vspace{-0.75 cm}
 \begin{figure}[!h]
-\centerline{\includegraphics[width=.5\textwidth]{figures/figure3.pdf}}
+%\begin{wrapfigure}{r}{0.5\textwidth}
+%\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/figure3.pdf}}
+\centerline{\includegraphics[width=0.55\textwidth]{figures/figure3.pdf}}
 \caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map.
@@
 \end{figure}
+%\end{wrapfigure}
 
-\subsection{Evapotranspiration estimates}
-Accurate estimation of hydrological fluxes is critical to understanding how land-atmosphere interactions impact the climate system and runoff \cite{liang_simple_1994}.
-Land surface hydrological models, such as the Variable Infiltration Capacity \cite{liang_simple_1994}, are used to simulate these processes, and depend on landcover maps to provide information on the characteristics of vegetation and other materials covering the surface, as these govern the rates of runoff, infiltration, and evapo-transpiration.
-We used the VIC model to generate 25 km estimates of monthly evapotranspiration throughout South Africa, and examined how these were impacted by error in the test maps, which were used to determine landcover-specific values for leaf area index (LAI), plant rooting depth, aerodynamic roughness, and several other variables that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.
+In terms of accuracy, MAE values were essentially the same as bias magnitudes, except for GeoWiki's, which were twice as large.
+The average MAE across vegetation classes was 47\% at 1 km, dropping to $<$10 only with 25 km of aggregation.
+In contrast, SA-LC's carbon estimates were twice as accurate at 1 km, and were slightly more accurate up to 25 km of aggregation were GeoWiki achieved parity.
 
+\subsubsection*{Evapotranspiration estimates}
 Compared to the carbon analysis, the bias and accuracy in evapotranspiration (ET) calculated using the VIC model was negligible, averaging less than than +/-2\%.
@@
 These locations correspond primarily to the margins of major crop production regions--in the center is the westernmost boundary of the summer rainfall growing region, marked approximately by the 400 mm isohyet, where maize is the primary crop.
-The west coast hotspot falls at the western edge of the wheat-dominated winter rainfall region \cite{hardy_rainfed_2011}, where growing season rainfall is approximately 200 mm.
+The west coast hotspot falls at the western edge of the wheat-dominated winter rainfall region \citep{hardy_rainfed_2011}, where growing season rainfall is approximately 200 mm.
 
@@
 \begin{figure}[h]
-\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
+\centerline{\includegraphics[width=.6\textwidth]{figures/et_bias_map.pdf}}
 \caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
@@
 
-\section{Socio-economic analyses}
-\subsection{Gridded crop yield and production data}
-The spatial variability of crop productivity and production is critical for understanding a host of social, economic, and environmental issues, such as food security, trade, and the potential for agricultural expansion and intensification \cite{licker_mind_2010,monfreda_farming_2008}.
-The most reliable source of such data are national to sub-national agricultural statistics, which are provided for relatively coarse-scaled administrative boundaries.
-To obtain higher spatial resolutions, efforts have been made to disaggregate these statistics using gridded landcover data \cite{ramankutty_farming_2008,monfreda_farming_2008}.
-These disaggregated datasets, which are constrained to match the agricultural statistics within the boundaries of the areas for which they are reported, have seen increasing use because they are considered to be more accurate than single source methods, and provide consistent data on which to base studies of global change \cite{ramankutty_farming_2008, see_improved_2015}
-
-We used these same methods  \cite{ramankutty_farming_2008,monfreda_farming_2008} to disaggregate maize harvested area (South Africa's largest crop \cite{estes_projected_2013}) on top of our reference and cropland maps, followed by yields, which were assigned to cells have harvested areas greater than zero.
-We then used these two layers to calculate maize production, and further aggregated the yield and production grids to 5, 10, 25, 50, and 100 km resolutions before quantifying the bias and accuracy of each test map's yield and production values.
-In this case, we could not convert cell-wise errors into percentages of the reference map values (because many cells had zero values for one map but not the other), so we calculated bias and accuracy from the map residuals and then normalized their values to the reference map means.
-
+\subsection*{Socio-economic analyses}
+\subsubsection*{Gridded crop yield and production data}
 Maize yields disaggregated onto the test maps showed some marked differences relative to the reference map, but only at the margins of the major crop production areas where cropland is sparser (SI).
@@
 
-\subsection{Agent-based simulation of households in dryland agroecosystems}
-Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}.
-To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistic represents the social and biophysical features of the study region \cite{berger_creating_2006}.
-In our example, we used an ABM of household food security that simulates food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}.
-The model is initialized such that each household is allocated its designated share of cropland based on aggregated farm-level statistics, and then calculates household-level agricultural production based on maize yields simulated by the DSSAT cropping system model \cite{jones_modelling_2003}.
-Yield estimates vary as a function of soil properties, management inputs, and weather.
-The former two factors differ between households, as do the total areas of cropland, thus total production can vary substantially between households.
-Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g.
-districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers.
-To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, SI) in the eastern part of the country with similar climate and 28-45\% of their areas devoted to cropland.
-The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity.
-The number of households and the cropland area per household was derived from surveys of farming communities where all cropland is used.
-The model was thus considered to be well-calibrated when all households were allocated their appropriate area of cropland, and all cropland was allocated to households.
-
 %\vspace{-0.5 cm}
@@
 %\vspace{-0.75 cm}
-\centerline{\includegraphics[width=0.5\textwidth]{figures/yield_prod_bias.pdf}}
+\centerline{\includegraphics[width=0.55\textwidth]{figures/yield_prod_bias.pdf}}
 \caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates.
@@
 
-We used the reference map and each cropland map to separately initialize the model, and examined how map errors impacted the land allocation process and household food production estimates.
-We examined three variables: the percent of unallocated cropland, land deficit, and food deficit.
-The percent of unallocated cropland what share of the district's total cropland area was not assigned to any household, and is a measure of how effective the model was in matching household agents to available cropland resources.
-Where error was negative (indicating a cropland overestimate by the test maps), the amount of land left unallocated had a straight one-to-one relationship with the percentage of overestimation (Fig.
+\subsubsection*{Agent-based model of household food security}
+In terms of impact to agent-based model simulation, where cropland map errors were negative (indicating a cropland overestimate by the test maps), the percent of land left unallocated had a straight one-to-one relationship with the percentage of overestimation (Fig.
 6A).
-When cropland was underestimated, no land was left unallocated until the underestimation exceeded 50\%.
-The MODIS-based simulation for districts 1 and 2 was most pronounced for this tendency, where 5-10\% of cropland remaining unallocated despite the fact that cropland with such a high cropland underestimate (85\%) the majority of households were not allocated cropland.
-This non-linear relationship occurred because croplands tend to cluster, and when underestimated clusters tend to be small, resulting in islands of cropland that fall outside of the search radius within which cropland is sought when agents are seeded onto the landscape.
+When cropland was underestimated, all croplands were allocated up until the underestimation exceeded 50\%.
+The MODIS-based simulation for districts 1 and 2 was most pronounced for this tendency, with 5-10\% of cropland remaining unallocated despite the fact that the majority of households were not assigned cropland (because cropland was underestimated by 85\%).
+This non-linear relationship occurred because croplands tend to cluster, and when underestimated clusters tend to be small and isolated, they are more likely to fall outside of the search radius used by the model for allocating fields to households when they are initially seeded onto the landscape.
 
-The second measure, land deficit, is calculated from the total area of cropland that should have been allocated to households in each district, but wasn't, due to the mismatch between the cropland map and the statistical estimates of what total cropland holdings should have been (given the number of households and the size class distributions of cropland in the district).
-Land deficit, expressed as a percentage of the total area of cropland in the district (thus values exceeded 100\% when total deficit exceeded cropland area),  increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2 (Fig.
+Land deficit (the total area of cropland that should have been allocated to households in each district, but wasn't) increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2 (Fig.
 6B)--and would become infinite in the case of a 100\% underestimate.
-This contrasted with the third measure, food deficit, the percentage shortfall in the average amount of food production that should have been produced by each household but wasn't because of the land deficit; this particular deficit increased linearly with the percentage of cropland underestimate (Fig.
+This contrasted with food deficit (the percentage shortfall in the average amount of food production that should have been produced by each household but wasn't), which increased linearly with the percentage of cropland underestimate (Fig.
 6C).
@@
 \begin{figure}[!ht]
-\centerline{\includegraphics[width=.5\textwidth]{figures/figure6.pdf}}
+\centerline{\includegraphics[width=.55\textwidth]{figures/figure6.pdf}}
 \caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and C) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
@@
 
-\section{Discussion}
+\section*{Discussion}
 This spatially comprehensive, bottom-up landcover map assessment provides new insight into the extent and causes of map error, and its consequences for understanding global change.
@@
 Although this reference map is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission, SI), while temporal mismatches between the reference and test maps may account for some of the error we identified, our assessment (SI) suggests that such discrepancies do not appreciably impact our findings.
-Our results are also supported by previous, continental-scale assessments that revealed substantial errors and inconsistencies between landcover maps \cite{fritz_comparison_2010,gross_monitoring_2013}.
+Our results are also supported by previous, continental-scale assessments that revealed substantial errors and inconsistencies between landcover maps \citep{fritz_comparison_2010,gross_monitoring_2013}.
 These similarities, combined with the larger extents of these studies, also suggest that our findings are applicable beyond South Africa's borders.
 
-\subsection{General guidelines for selecting and using landcover data}
+\subsection*{General guidelines for selecting and using landcover data}
 Our findings regarding bias and accuracy suggest several guidelines for selecting and using landcover data in global change research (Table 1), and suggested aggregation scales.
@@
 The first general guideline is that landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions.
-Researchers might need to use these datasets for a variety of reasons, for instance, because they provide consistent, multiple landcover classes at a global scale, or allow changes to be tracked across large scales over time \cite{luoto_predicting_2004}.
+Researchers might need to use these datasets for a variety of reasons, for instance, because they provide consistent, multiple landcover classes at a global scale, or allow changes to be tracked across large scales over time \citep{luoto_predicting_2004}.
 The appropriate scale of aggregation then depends on which error property the user wants to minimize.
@@
 We found that accuracy is lowest where mixing of cover types is greatest (Fig.
-2), a result echoed in an Africa-wide change detection study made with MODIS \cite{gross_monitoring_2013}.
-Outside of South Africa, where smallholder-driven agriculture predominates \cite{lambin_estimating_????}, such mixed landscapes are more common, thus greater levels of aggregation may be needed over these areas.
+2), a result echoed in an Africa-wide change detection study made with MODIS \citep{gross_monitoring_2013}.
+Outside of South Africa, where smallholder-driven agriculture predominates \citep{lambin_estimating_2013}, such mixed landscapes are more common, thus greater levels of aggregation may be needed over these areas.
 Coarser aggregations are also necessary if the cover type of interest belongs primarily to a mixed class (as with GlobCover).
@@
 
-Unfortunately, Landsat-scaled maps are typically developed for specific countries, using varying methods, and can be hard to obtain (an exception to this is the newly released GLC30, which has a reported classification accuracy of 80\% \cite{chen_global_2015}).
-Our results also suggest that, despite their higher accuracy, they may not necessarily the least biased maps.
-Except for a few results (notably the 1 km carbon and maize production estimates), GeoWiki-based maps had the lowest bias across most scales, which is the largely the result of calibrating the cropland percentages to match reported agricultural statistics \cite{fritz_cropland_2011,fritz_mapping_2015}.
-This procedure is similar to the one we used in disaggregating maize harvested areas \cite{ramankutty_farming_2008, monfreda_farming_2008}, which led to fairly unbiased estimates of maize production and yield (Fig.
+Unfortunately, Landsat-scaled maps are typically developed for specific countries, using varying methods, and can be hard to obtain (an exception to this is the newly released GLC30, which has a reported classification accuracy of 80\% \citep{chen_global_2015}).
+Our results also suggest that, despite their higher accuracy, they may not necessarily be the least biased maps.
+Except for a few results (notably the 1 km carbon and maize production estimates), GeoWiki-based maps had the lowest bias across most scales, which is the largely the result of calibrating the cropland percentages to match reported agricultural statistics \citep{fritz_cropland_2011,fritz_mapping_2015}.
+This procedure is similar to the one we used in disaggregating maize harvested areas \citep{ramankutty_farming_2008, monfreda_farming_2008}, which led to fairly unbiased estimates of maize production and yield (Fig.
 5) above 1 km (except for GlobCover's production estimates).
@@
 %The consensus method mirrors the ensemble averaging used by other fields (e.g.
-crop \cite{asseng_uncertainty_2013}, climate \cite{giorgi_calculation_2002}, and ecological modeling \cite{araujo_ensemble_2007}) to increase prediction confidence.
+crop \citep{asseng_uncertainty_2013}, climate \citep{giorgi_calculation_2002}, and ecological modeling \citep{araujo_ensemble_2007}) to increase prediction confidence.
 
-Of course, for the constraint to be effective, the inventory data themselves must be faithful to reality, but in many places, particularly Africa, agricultural statistics are suspect \cite{carletto_emperor_2013, fao_action_2013}, which is a vulnerability of the fusion approach \cite{see_improved_2015}.
+Of course, for the constraint to be effective, the inventory data themselves must be faithful to reality, but in many places, particularly Africa, agricultural statistics are suspect \citep{carletto_emperor_2013, fao_action_2013}, which is a vulnerability of the fusion approach \citep{see_improved_2015}.
 Beyond this limitation, this method does not improve map accuracy, which is evident in GeoWiki's relatively high MAE for cropland percentage at 1 km (23\%, Fig.
@@
 The reason that accuracy is not improved is because the statistical constraint cannot correct the omission and commission errors within the landcover maps.
-To minimize these types of errors, different techniques are needed, such as creating a single consensus map from multiple landcover products \cite[e.g.][]{fritz_comparison_2010,tuanmu_global_2014}.
+To minimize these types of errors, different techniques are needed, such as creating a single consensus map from multiple landcover products \citep[e.g.][]{fritz_comparison_2010,tuanmu_global_2014}.
 This second kind of data fusion was used in the first stage of creating the GeoWiki map (before statistical adjustment), and it accounts for GeoWiki's higher accuracy scores relative to MODIS and GlobCover (Fig.
 1B, 3, \& 5).
-It may be particularly effective in minimizing underestimation errors that result when a mixed class type becomes dominant (as with GlobCover), or when dealing with substantial sub-pixel heterogeneity \cite{fritz_mapping_2015,tuanmu_global_2014}
+It may be particularly effective in minimizing underestimation errors that result when a mixed class type becomes dominant (as with GlobCover), or when dealing with substantial sub-pixel heterogeneity \citep{fritz_mapping_2015,tuanmu_global_2014}
 
@@
     \begin{center}
-      \begin{tabular}{p{4.5cm}lp{6cm}l}
+      \begin{tabular*}{\textwidth}{m{3.5cm}m{3cm}m{5.5cm}m{2.25cm}}
        \hline
@@
 \Xhline{1pt}
-High frequency change           & Maximize accuracy   & Coarse resolution sensor (e.g. MODIS) & $>$100 km\\
-detection                                                               &                                  & \textbf{High resolution sensor} (e.g. Landsat)   &  1-25 km\\
+High frequency            & Maximize accuracy   & Coarse resolution sensor (e.g. MODIS) & $>$100 km\\
+change detection                                                               &                                  & \textbf{High resolution sensor} (e.g. Landsat)   &  1-25 km\\
 \hline
@@
 \Xhline{1pt}                         
-       \end{tabular}
+       \end{tabular*}
        \begin{tablenotes}
@@
         \item *Aggregation ranges account for variation in the results from the different map analyses.
-        \item **This assumes that consensus-based fusion products \cite[e.g.][]{fritz_cropland_2011,fritz_mapping_2015,tuanmu_global_2014} incorporate maps from coarse resolution sensors, making them less accurate at finer scales than products created using high resolution sensors.
+        \item **This assumes that consensus-based fusion products \citep[e.g.][]{fritz_cropland_2011,fritz_mapping_2015,tuanmu_global_2014} incorporate maps from coarse resolution sensors, making them less accurate at finer scales than products created using high resolution sensors.
        \end{tablenotes}
@@
   \end{threeparttable}
-\end{table*}%
+\end{table*}
 
 These last few points suggest two final guidelines.
-For high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \cite{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share.
+For high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \citep{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share.
 www.glcn.org} datasets for other cover types).
 But in either case some aggregation--1-25 km, depending on application and whether low bias or high accuracy is the objective--will still likely be necessary to achieve sufficiently low error.
-Additionally, these higher accuracy, lower bias products often represent just a single time period \cite[e.g.][]{fritz_mapping_2015,tuanmu_global_2014}, thus applications requiring higher frequency (e.g.
+Additionally, these higher accuracy, lower bias products often represent just a single time period \citep[e.g.][]{fritz_mapping_2015,tuanmu_global_2014}, thus applications requiring higher frequency (e.g.
 annual), large area landcover maps will depend on MODIS or similar sensors, and therefore much coarser aggregations.
@@
 
-\subsection{Implications for global change analyses}
+\subsection*{Implications for global change analyses}
 Besides the preceding landcover data selection and usage guidelines, our findings also have broader implications for global change analyses and associated policy decisions.
 The simpler, most direct uses of landcover data have substantial potential to skew our understanding of global change processes.
-For example, the global carbon carbon density map of \cite{ruesch_new_2008}, of which we created facsimiles here, has been widely used, but our results suggest that any analyses of the spatial variability of carbon stocks, or of mean carbon densities, can be highly misleading if the map is not first sufficiently aggregated.
+For example, the global carbon carbon density map of \citep{ruesch_new_2008}, of which we created facsimiles here, has been widely used, but our results suggest that any analyses of the spatial variability of carbon stocks, or of mean carbon densities, can be highly misleading if the map is not first sufficiently aggregated.
 For example, a study of deforestation risks in the Democratic Republic of the Congo estimated forest rents as a function of carbon densities based on this map.
@@
 3).
-The same carbon map provided country-specific mean forest carbon densities for an evaluation of climate mitigation policies \cite{cattaneo_international_2010}, which our results suggest could be substantially underestimated (Fig.
+The same carbon map provided country-specific mean forest carbon densities for an evaluation of climate mitigation policies \citep{cattaneo_international_2010}, which our results suggest could be substantially underestimated (Fig.
 3).
-In contrast to these, an assessment of the congruence between carbon and biodiversity \cite{strassburg_global_2010} aggregated the carbon densities to $\sim$110 km, where accuracy is much higher.
+In contrast to these, an assessment of the congruence between carbon and biodiversity \citep{strassburg_global_2010} aggregated the carbon densities to $\sim$110 km, where accuracy is much higher.
 
-A related example can be seen in high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \cite[e.g.
+A related example can be seen in high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \citep[e.g.
 Figure 3 in][]{foley_solutions_2011}.
@@
 Here the results appear to be mixed.
-Map error appears to have had little impact on the overall bias and accuracy of VIC's evapotranspiration (ET) estimates, which contrasts with a related study that examined how landcover maps impact rainfall simulations \cite{ge_impacts_2007}.
+Map error appears to have had little impact on the overall bias and accuracy of VIC's evapotranspiration (ET) estimates, which contrasts with a related study that examined how landcover maps impact rainfall simulations \citep{ge_impacts_2007}.
 There are two reasons for this.
@@
 Despite the low values of the overall error metrics, the hotspots of ET errors (Fig.
-4) are important, because they occurred at the climatic margins of the major crop growing regions, where a substantial share of farms use irrigation.
-Since VIC does not simulate irrigation, the magnitude of these errors is likely to have been underestimated, because irrigation substantially increases latent heat flux and can substantially modify regional climate\cite{sacks_effects_2008,mueller_cooling_2015}.
-This also suggests the potential for map error to magnify in a coupled land-atmosphere model (which could explain the difference between this result and that of Ge et al.
-\cite{ge_impacts_2007}.), and thereby give misleading insight into the climatic effects of land use.
+4) may have outsize significance because they occurred at the climatic margins of the major crop growing regions, where a substantial share of farms are irrigated.
+Since VIC does not simulate irrigation, the magnitude of these errors is likely to have been underestimated, because irrigation substantially increases latent heat flux and can substantially modify regional climate \citep{sacks_effects_2008,mueller_cooling_2015}.
+This also suggests the potential for map error to magnify in a coupled land-atmosphere model \citep[which could explain the difference between this result and that of][]{ge_impacts_2007}, and thereby give misleading insight into the climatic effects of land use.
 
 Relative to the ET example, map errors had a much larger impact on the agent-based model's results.
-For the model's most important metric of food security--household-level crop production--the result was fairly straightforward for a class of models where the interactions of multiple agents often produce analytically intractable outcomes \cite{janssen_empirically_2006}.
+For the model's most important metric of food security--household-level crop production--the result was fairly straightforward for a class of models where the interactions of multiple agents often produce analytically intractable outcomes \citep{janssen_empirically_2006}.
 In this case, the model was only sensitive to cropland underestimates, which had the perfectly predictable effect of lowering average household food production, which would in turn overstate the degree of food \emph{in}security.
@@
 The errors in the cropland data we tested were large, and it is unlikely that current ABM-based studies, mostly of small regions, are based on such inaccurate landcover data.
-However, as computing power increases, so, too, does the feasible application extent of ABMs, which then become more reliant on larger-scaled, more erroneous, landcover data, with the attendant risk that insights into socio-ecological systems arising from these models could be highly misleading.
+However, as computing power increases, so, too, does the feasible application extent of ABMs, which will then become more reliant on larger-scaled, more erroneous, landcover data, with the attendant risk that insights into socio-ecological systems arising from these models could be highly misleading.
 
 Beyond these examples, there are other applications where landcover map errors could alter understanding and policy.
-One of these lies in assessments of land availability for agricultural expansion, which could be particularly erroneous if they use a ``residual approach'' \cite{lambin_estimating_????}, where existing cropland and other occupied cover types are used to filter a map of potential suitability.
+One of these lies in assessments of land availability for agricultural expansion, which could be particularly erroneous if they use a ``residual approach'' \citep{lambin_estimating_2013}, where existing cropland and other occupied cover types are used to filter a map of potential suitability.
 Our results suggest that the pronounced underestimation biases (Fig.
-1) we found could inflate land availability estimates, and therefore increase the risk of unjust land allocation policies \cite{rulli_global_2013}.
-Another example is the growing efforts to minimize the environmental and climate impacts on new agricultural expansion frontiers, particularly Africa's savannas, which requires identifying areas that provide the highest agricultural benefit for the lowest environmental cost \cite{searchinger_high_2015}.
-This sort of analysis entails finding ``win-win'' land use solutions, which are essentially areas that deviate from an expected development cost-benefit relationship; for example, regions of unusually low carbon density but high crop fertility \cite{searchinger_high_2015}.
-The validity of such analysis are entirely dependent In order to be effective, such analyses must be undertaken at the fine scales--hectares, rather than square kilometers--at which land use decisions are made, thus their value for informing land use policy is contingent on the accuracy of their underlying landcover/land use data \cite{searchinger_high_2015}.
+1) we found could inflate land availability estimates, and therefore increase the risk of unjust land allocation policies \citep{rulli_global_2013}.
+Another example is the growing efforts to minimize the environmental and climate impacts on new agricultural expansion frontiers, particularly Africa's savannas, which requires identifying areas that provide the highest agricultural benefit for the lowest environmental cost \citep{searchinger_high_2015}.
+This sort of analysis entails finding ``win-win'' land use solutions, which are essentially areas that deviate from an expected development cost-benefit relationship; for example, regions of unusually low carbon density but high crop fertility \citep{searchinger_high_2015}.
+The validity of such analysis are entirely dependent In order to be effective, such analyses must be undertaken at the fine scales--hectares, rather than square kilometers--at which land use decisions are made, thus their value for informing land use policy is contingent on the accuracy of their underlying landcover/land use data \citep{searchinger_high_2015}.
 The validity of such analyses, which must be undertaken at the fine-scales where land use decisions are made (hectares rather than square kilometers), depends on having high quality map data, in order to have confidence that the identified tradeoffs are based on real biophysical properties, not simply map error.
 
-\subsection{The way forward}
+\subsection*{The way forward}
 Our analysis demonstrates that the landcover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on how accuracy or low bias.
@@
 Beyond these guidelines, we recommend updating several important, global-scale datasets using the latest generation landcover maps.
-For example, rebuilding the widely used cropland distribution and yield maps \cite{monfreda_farming_2008, ramankutty_farming_2008} using the GeoWiki map could substantially improve our knowledge about current global agricultural distribution and productivity.
-Similarly, reconstructing the IPCC's Tier 1 vegetative carbon map \cite{ruesch_new_2008} using a combination of the GeoWiki and new global landcover consensus maps \cite[e.g.
+For example, rebuilding the widely used cropland distribution and yield maps \citep{monfreda_farming_2008, ramankutty_farming_2008} using the GeoWiki map could substantially improve our knowledge about current global agricultural distribution and productivity.
+Similarly, reconstructing the IPCC's Tier 1 vegetative carbon map \citep{ruesch_new_2008} using a combination of the GeoWiki and new global landcover consensus maps \citep[e.g.
 GLC-Share or][]{tuanmu_global_2014} the GLC-Share datasets would improve our confidence in regional to global assessments of vegetative carbon stocks.
-Alternatively, the new high-resolution, Landsat-based GLC30 \cite{chen_global_2015}, aggregated and converted to fractional cover classes at 1 km--and fused with high quality national landcover maps, where available \cite[similar to][]{fritz_mapping_2015,fritz_cropland_2011}--may provide a preferable base map for rebuilding these datasets.
+Alternatively, the new high-resolution, Landsat-based GLC30 \citep{chen_global_2015}, aggregated and converted to fractional cover classes at 1 km--and fused with high quality national landcover maps, where available \citep[similar to][]{fritz_mapping_2015,fritz_cropland_2011}--may provide a preferable base map for rebuilding these datasets.
 
 Fusing human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power appears to offer the greatest promise for improving the quality of landcover maps.
-The GeoWiki project and new high resolution global forest change maps \cite{hansen_high-resolution_2013,sexton_conservation_2015} are prominent examples of these newer methods.
-Another approach is that of the Mapping Africa project\footnote{http://mappingafrica.princeton.edu}, which, inspired by quality of the reference map used in this study, enlists internet-based workers to manually digitize cropland boundaries \cite{estes_platform_2015}.
-Merging this crowdsourced approach for creating vector maps of discrete cover types with newer computer vision-based mapping algorithms \cite{debats_generalized_2015} is a near-term goal of this project, as the former can interactively provide high quality training and test data (by using humans' superior capacity to delineate discrete cover types in noisy images \cite{estes_platform_2015}) for the latter, which has the advantage of speed and the ability to handle high-dimensional datasets.
+The GeoWiki project and new high resolution global forest change maps \citep{hansen_high-resolution_2013,sexton_conservation_2015} are prominent examples of these newer methods.
+Another approach is that of the Mapping Africa project\footnote{http://mappingafrica.princeton.edu}, which, inspired by quality of the reference map used in this study, enlists internet-based workers to manually digitize cropland boundaries \citep{estes_platform_2016}.
+Merging this crowdsourced approach for creating vector maps of discrete cover types with newer computer vision-based mapping algorithms \citep{debats_generalized_2016} is a near-term goal of this project, as the former can interactively provide high quality training and test data (by using humans' superior capacity to delineate discrete cover types in noisy images \citep{estes_platform_2016}) for the latter, which has the advantage of speed and the ability to handle high-dimensional datasets.
 Approaches such as these will be needed to create the next generation of landcover maps, so that we can gain a clearer, unbiased understanding of global change.
@@
 
-\begin{materials}
-\section{Methods} 
-To develop the cropland reference map (see SI for accuracy assessment details), we intersected the original field boundary vector maps with a 1 km grid, and calculated the percent of each cell occupied by fields to create the gridded cropland reference map.
-In the resulting grid, we masked out areas classified as communal farmland, because only their outer perimeters were digitized \cite{fourie_better_2009}, which posed a risk of overestimating cropland extent.
-We also excluded the permanent tree crop class, to avoid confusion with landcover maps that exclude these in their cropland classes.
-For the same reason, we used other high resolution landcover maps for selected provinces of South Africa to mask out commercially afforested areas and sugarcane plantations, which were also not included in the reference map.
-The resulting masked reference grid covered 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ was cropland.\\
-\indent We extracted the cropland classes from SA-LC, MODIS, and GlobCover, and converted these into percent cropland estimates at 1 km resolution (GeoWiki is already a cropland percentage map).
-Both MODIS and GlobCover had mixed pixel classes, thus, following the approach used in creating the GeoWiki maps \cite{fritz_mapping_2015} we used the upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages.
-We used the mean map for the main analysis, but estimated error variability using all three versions (SI).\\
-\indent To uncover the relationship between map error and cover density, we used magisterial district boundaries (n=354, mean area=3,445 km$^2$; SI) to provide a landscape-scaled unit for calculating characteristic cover density.
-We filtered out pixels with $<$0.05\% (0.5 ha) cropland, to prevent the much larger areas of non-agricultural districts from drowning out any relationship.
-We extracted the absolute values of test map errors and corresponding reference cropland percentages, calculated their district-wide means, and then modeled the relationship between mean absolute error (response) and cropland density (predictor) as a polynomial function within a generalized additive model.
-We accounted for potential spatial autocorrelation by fitting a two-dimensional smoothing spline to the coordinates of each district's centroid, and weighted the model by the number of agricultural pixels per district \cite{wood_mgcv:_2001}.\\
-\indent To develop the carbon maps, we followed the methods used by Ruesch and Gibbs \cite{ruesch_new_2008} to create their vegetative carbon map.
-We took the average of the biome specific carbon density values they provided for African forests, secondary forests, shrublands, grasslands, sparse habitats, and croplands.
-We multiplied cropland densities by cropland fractions and added these to each of the five other densities multiplied by the residual non-cropland fractions to create five different carbon density maps.
-We aggregated each carbon map up to the five coarser resolutions for scaling comparisons (SI).
-\\
-\indent We used the Variable Infiltration Capacity \cite{liang_simple_1994} land surface hydrology model run with the Africa Flood and Drought Monitor's meteorological data \cite{sheffield_drought_2013} to produce monthly gridded evapotranspiration estimates for South Africa for the years 1979-2010.
-We adjusted VIC's native landcover scheme (derived from AVHRR) so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25$^{\circ}$ resolution).
-We ran one instance of VIC for each of the five landcover permutations, and compared the mean annual ET produced by the reference map variant with those from the test maps.
-\\
-\indent To create the gridded maize yield and production maps, we followed Ramankutty et al.
-\cite{ramankutty_farming_2008} to first calibrate the test cropland maps against cropland areas reported by administrative-level agricultural censuses.
-In place of census reports, we used the reference map to calculate total cropland areas for South Africa's nine provinces, then adjusted the pixel-wise cropland percentages in the four test maps so that their province-wise sums matched these totals (SI).
-We then followed Monfreda et al's procedure \cite{monfreda_farming_2008} for disaggregating maize \cite[South Africa's largest crop][]{estes_comparing_2013} planted area and yields onto the reference and adjusted test maps.
-The necessary statistics were obtained from magisterial district-level agricultural censuses conducted for the year 2007 \cite{statistics_south_africa_commercial_2007}.
-We then aggregated gridded harvested areas and yields (weighted by the corresponding harvested area) to each coarser resolution, and multiplied yield by harvested area to calculate production.\\
-\indent The agent-based model of food security used here calculates average household-level food production over multiple seasons based on the biophysical environment (soil properties, weather), management inputs (planting decisions, labor availability), and farm assets (total field area) \cite{chen_dependency_2013}.
-The frequency distributions of household cropland holdings were derived from household survey data obtained from Zambia (no equivalent data were available for South Africa).
-We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area (2.2 ha), and preserved the cropland area distributions by multiplying the total number of households by the frequencies.
-To create cropland surfaces for each district, we disaggregated all five cropland maps to binary cropland/non-cropland cover types with 100 m resolution, in order to match the survey-recorded average field size of 1 ha.
-In initializing, the model takes a weighted (by cropland area frequency) random draw of 100 households and places these within a district, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household.
-This process is iterated until all households are assigned cropland, or all available cropland is allocated.
-The model then runs separately for each district, with the household agents planting and managing a maize crop throughout the growing season, with possible final maize yields drawn from a look-up table that relates yield to particular climatic outcomes and management decisions, as simulated by the DSSAT model.
-We ran each ABM simulation for a single season, once per district and when initialized by each of the cropland maps (20 simulations total).\\
-\indent In calculating map error statistics, and with the exception of the agent-based model results, we use the reference cropland percentages to calculate the weighted mean map error (bias) and mean absolute map error (accuracy).
-For the comparisons between aggregated map values, we applied a further weight--the number of pixels contributing to each aggregated pixel, to prevent pixels close to national boundaries or where non-target cover types were masked out from having outsize influence on the statistics.
+%\begin{materials}
 
-\end{materials}
+%\end{materials}
 
@@
 
-\begin{acknowledgments}
-This work was supported by funds from the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program
-(NNX15AC64G), and the National Science Foundation (SES-1360463 and BCS-1026776).
-\end{acknowledgments}
+\section*{Acknowledgments}
+This work was supported by funds from the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463 and BCS-1026776).
+We thank Tobias Kuemmerle for his helpful review of our study and suggestions for improvement.
+%\end{acknowledgments}
 
-\bibliographystyle{pnas2011} 
+\bibliographystyle{gcbnourl} 
 {\footnotesize \bibliography{/Users/lestes/Dropbox/publications/full.bib}}
 
-\end{article}
+%\end{article}
 

### 2016-12-06 Rmd/c10-location-impact.Rmd
Revisions through 6/12/2016: 1. Manuscript into methods 2. Recompiled library, dealing with bug in raster::mask (masked out by different library) 3. Revised c6-carbon, c8-yield-production to save yield and carbon grids. 4. Created c10-location-impact.Rmd, new analysis to estimate difference in location due to landcover error in yield and carbon stocks

---
title: "Yield and production bias/accuracy"
author: "Lyndon Estes"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc: yes 
    number_sections: true
---

# Impact of cropland errors on location

Evaluated in terms of how landcover map error impacts the ability to locate areas having a particular quantity of interest.

## Data

Derived carbon (section 6) and yield estimates (section 8).

```{r, eval=FALSE}
library(SAcropland)
library(xtable)
library(RColorBrewer)
library(spatstat)
p_root <- full_path(proj_root("SAcropland"), "SAcropland")
p_fig <- full_path(p_root, "paper/figures/")
p_data <- full_path(p_root, "external/ext_data/")
p_data2 <- full_path(p_root, "data/")
#p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")
# load(full_path(p_data, "yield-bias.rda"))

# SA provinces
prov <- readOGR("external/ext_data/provinces.sqlite", layer = "provinces")
prov <- prov[prov$id_1 != 9, ]  # remove Princeton Edward Islands
prov[prov$id_1 == 10, "id_1"] <- 9 # reset WC to id 10
load("external/ext_data/MZshapes.Rdata")

# yield and carbon data
load("external/ext_data/yieldprod-1km.rda")
load("external/ext_data/carbon-tier1.rda")

```

## Differences in location

If you are interested in selecting areas of highest yield/production/carbon potential, how far off are you relative to "truth". Use top quintile/decline to assess this, and then calculate mean nearest neighbor distance between two pixels of each map.

### Identify high value areas
```{r, eval = FALSE}
# top decile of...
# ...yield
yclass <- lapply(cyld_agg$f1, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

# ...production
pclass <- lapply(cpagg1$f1, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

# ...carbon
cdat <- c("gti" = gti_c$f1, lc_c$f1)  # combine C 1 km sets
cdatf <- lapply(cdat, function(x) {  # mask out non-cropland areas and take mu C
  msk <- x$forest < 99.6
  cmu <- calc(x, mean)
  raster::mask(cmu, msk, maskvalue = 0)
})
cclass <- lapply(cdatf, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

# carbon efficiencies (C / yield)
effclass <- lapply(1:5, function(x) { # x <- 1
  cyld <- cyld_agg$f1[[x]]
  cyld[cyld == 0] <- NA
  cdatfc <- crop(cdatf[[x]], cyld)
  eff <- cdatfc / cyld  # carbon efficiency
  # eff[is.infinite(eff)] <- NA
  rq <- quantile(eff, c(0, 0.9, 1))
  types <- cut(eff, breaks = rq)
  types > 1
})
names(effclass) <- names(cclass)

```

### Calculate nearest-neighbor distances

Between GTI high value areas and each of the four LC-derived datasets. 
```{r, eval = FALSE}

# calculate distance between GTI and each other dataset's classes. 
# pms2a <- expand.grid(rep(list(0:1), 5))[-1, ] 
# pms2b <- pms2a[which(rowSums(pms2a) == 2 & pms2a$Var1 == 1), ]
# pms2b$cl <- c("YC1", "YC2", "YC3", "YC4")
# rownames(pms2b) <- 1:4

# Define neighborhood window
sa_owin <- owin(xrange = bbox(sa.shp)[1, ], yrange = bbox(sa.shp)[2, ],
                unitname = "meter")
# which1 <- function(x) x == 1  # selector function for r to xy

mudists <- lapply(list(yclass, pclass, cclass, effclass), function(x) {
  # x <- yclass
  mudist <- sapply(2:5, function(y) {  # y <- 2
    print(paste0("...", y))
    a <- rasterToPoints(x$gti, fun = function(x) x == 1)
    b <- rasterToPoints(x[[y]], fun = function(x) x == 1)
    pps <- lapply(list(a, b), function(z) {
      ppp(x = z[, 1], y = z[, 2], window = sa_owin)
    })
    names(pps) <- c("gti", names(x)[y])
    reord <- sort(sapply(pps, function(i) i$n), decreasing = TRUE)
    ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
    mean(ppnn$dist) / 1000
  })
  names(mudist) <- names(x)[2:5]
  mudist
})  
names(mudists) <- c("yield", "prod", "carb", "eff")

# plot(stack(yclass))
# plot(is.na(yclass[[1]]) - is.na(yclass[[2]]))
# plot(is.na(cyld_agg$f1$gti) - is.na(cyld_agg$f1$geow))

# output plot
cols <- c("red", "orange3", "green4", "blue")
pcol <- makeTransparent(cols, alpha = c(60))
plot(1:4, pch = 20, col = pcol)
bs <- do.call(rbind, mudists[c(1, 3)])
lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")

pdf(full_path(p_fig, "location_offset.pdf"), height = 4, width = 4, 
    bg = "transparent")
par(mar = c(4, 4, 1, 0), bg = "transparent")
bp <- barplot(bs, beside = TRUE, col = ggplot2:::interleave(cols, pcol), las = 2,
              border = "transparent", xaxt = "n", ylab = "km", ylim = c(0, 16), 
              yaxt = "n")
axis(1, at = colMeans(bp), labels = lcnms, lty = 0)
axis(2, at = seq(0, 16, 2), labels = seq(0, 16, 2), las = 2)
legend(x = 9.5, y = 16, legend = rep("", 4), pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0, horiz = TRUE)
legend(x = 9.5, y = 15.25, legend = rep("", 4), pch = 15, adj = 0, col = pcol, 
       pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
text(x = 10, y = 15.1, labels = "Yield", pos = 2, cex= 0.8)
text(x = 10, y = 14.35, labels = "Carbon", pos = 2, cex = 0.8)
dev.off()

```

### 2016-12-06 Rmd/c2-landcover-preprocess-2.Rmd
Revisions through 6/12/2016: 1. Manuscript into methods 2. Recompiled library, dealing with bug in raster::mask (masked out by different library) 3. Revised c6-carbon, c8-yield-production to save yield and carbon grids. 4. Created c10-location-impact.Rmd, new analysis to estimate difference in location due to landcover error in yield and carbon stocks

@@
 sumfun <- function(x) sum(x, na.rm = TRUE)
+# gti <- lapply(list(cover2007, cover2011), function(y) {
+#   r <- calc(y[[-6]], sumfun)
+#   r[r > 100] <- 100
+#   writeRaster(r, file = gsub("\\.tif", "sum.tif", y@file@name), overwrite = TRUE)
+# }) 
 gti <- lapply(list(cover2007, cover2011), function(y) {
-  r <- calc(y[[-6]], sumfun)
-  r[r > 100] <- 100
-  writeRaster(r, file = gsub("\\.tif", "sum.tif", y@file@name), overwrite = TRUE)
+  raster(gsub("\\.tif", "sum.tif", y@file@name))
 })
@@
 # Leave commented out in case want to bring back horticulture class
-# gti_hort <- lapply(list(cover2007, cover2011), function(y) {
-#   r <- calc(y, sumfun)
-#   r[r > 100] <- 100
+gti_hort <- lapply(list(cover2007, cover2011), function(y) {
+  r <- calc(y, sumfun)
+  r[r > 100] <- 100
   # writeRaster(r, file = gsub("\\.tif", "sum_h.tif", y@file@name), overwrite = TRUE)
-# })
+  r
+})
+
+# calculate areas
+cellStats(cover2011[[2]] / 100, sum) / 
+  cellStats(calc(cover2011 / 100, sumfun), sum) * 100  # communal areas 18.7\% cropland
+cellStats(cover2011[[6]] / 100, sum) / 
+  cellStats(calc(cover2011 / 100, sumfun), sum) * 100  # horticulture 3.1\% of cropland
 

### 2016-12-06 Rmd/c6-carbon.Rmd
Revisions through 6/12/2016: 1. Manuscript into methods 2. Recompiled library, dealing with bug in raster::mask (masked out by different library) 3. Revised c6-carbon, c8-yield-production to save yield and carbon grids. 4. Created c10-location-impact.Rmd, new analysis to estimate difference in location due to landcover error in yield and carbon stocks

@@
 
+# save for further analysis
+save(gti_c, lc_c, file = full_path(p_data, "carbon-tier1.rda"))
 ```

### 2016-12-06 Rmd/c8-yield-production.Rmd
Revisions through 6/12/2016: 1. Manuscript into methods 2. Recompiled library, dealing with bug in raster::mask (masked out by different library) 3. Revised c6-carbon, c8-yield-production to save yield and carbon grids. 4. Created c10-location-impact.Rmd, new analysis to estimate difference in location due to landcover error in yield and carbon stocks

@@
 library(RColorBrewer)
+library(spatstat)
 p_root <- full_path(proj_root("SAcropland"), "SAcropland")
@@
                   "maize_AreaYieldProduction.nc")
-monfmz <- brick(monf, level = 2)
+# monfmz <- brick(monf, level = 2, varname = "maizeData")
+monfmz <- raster(monf, level = 2)
+
 plot(crop(monfmz, spTransform(sa.shp, monfmz@crs)), col = bpy.colors(20))
@@
 
-DT <- na.omit(as.data.table.raster(s, xy = TRUE))
-DT[, wgt := ref * awgts]  # calculate weights from ref and area weights
+# DT <- na.omit(as.data.table.raster(s, xy = TRUE))
+# DT[, wgt := ref * awgts]  # calculate weights from ref and area weights
 
@@
 presa <- projectRaster(presa, lcu$f25$sa30)  # project to Albers
-presa <- mask(presa, awgts$f25, maskvalue = 0)  # mask
+presa <- raster::mask(presa, awgts$f25, maskvalue = 0)  # mask
 presa1k <- disaggregate(presa, fact = 25)
@@
 save(list = ls(), file = full_path(p_data, "yield-bias.rda"))
+save(cyld_agg, cpagg1, file = full_path(p_data, "yieldprod-1km.rda"))
+
 ```

### 2016-12-06 paper/main.tex
Version submitted to GCB
Revisions through 6/12/2016: 1. Manuscript into methods 2. Recompiled library, dealing with bug in raster::mask (masked out by different library) 3. Revised c6-carbon, c8-yield-production to save yield and carbon grids. 4. Created c10-location-impact.Rmd, new analysis to estimate difference in location due to landcover error in yield and carbon stocks

@@
-%% PNAStwoS.tex
-%% Sample file to use for PNAS articles prepared in LaTeX
-%% For two column PNAS articles
-%% Version1: Apr 15, 2008
-%% Version2: Oct 04, 2013
-%
-%% BASIC CLASS FILE
-%\documentclass{pnas}
-%
-%% ADDITIONAL OPTIONAL STYLE FILES Font specification
-%
-%\usepackage{pnastwoF}
-%\usepackage{cite}
-%\usepackage[numbers,round,sort,compress]{natbib}
-%\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
-%\usepackage{lineno}
-%\setlength\linenumbersep{2pt}
-%\usepackage{placeins}
-%\usepackage{float}
-%\restylefloat*{figure}
-%\usepackage{dblfloatfix}  % http://tex.stackexchange.com/questions/235623/placing-a-figure-in-the-bottom-of-a-page-spanning-the-two-columns-of-an-ieee-doc 
-%\usepackage{subfigure}
-%\usepackage{wrapfig}
-%\usepackage{caption}
-%\setlength\belowcaptionskip{-2cm}
-%\setlength{\textfloatsep}{5pt plus 1.0pt minus 2.0pt}
-%\usepackage{multirow}
-%\setlength{\parskip}{1cm plus4mm minus3mm}
-%\usepackage{makecell}
-%\usepackage{boldline}
-%\setcellgapes{3pt}
-%\usepackage[flushleft]{threeparttable}
-%
-%% OPTIONAL MACRO DEFINITIONS
-%\def\s{\sigma}
-%%%%%%%%%%%%
-%% For PNAS Only:
-%\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-%\copyrightyear{2008}
-%\issuedate{Issue Date}
-%\volume{Volume}
-%\issuenumber{Issue Number}
-%\setcounter{page}{2687} %Set page number here if desired
-%%%%%%%%%%%%
-%
-%
-%
 \documentclass[a4paper]{article}
 \usepackage[affil-it]{authblk}    
-
 \usepackage{graphicx}
@@
 
-% Let's add todonotes and comments:
-%\usepackage[draft]{todonotes}
-%
-%% Select what to do with command \comment:  
-%% \newcommand{\comment}[1]{}  %comment not showed
-%\newcommand{\comment}[1]
-%{{[\bfseries \color{blue} #1]}} %comment showed
-
 \usepackage{lineno}
@@
 
-\title{Errors in Landcover Data Limit Our Understanding of Global Change}
-%\title{Our Understanding of Global Change is Built on a Shaky Foundation}
-
-%\author{Lyndon Estes{\affil{1}{Woodrow Wilson School, Princeton University, Princeton, NJ USA}\affil{2}{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}},
-%Peng Chen\affil{3}{Indiana University, Bloomington, IN USA},
-%Stephanie Debats\affil{2}{},
-%Tom Evans\affil{3}{},
-%Stefanus Ferreira\affil{4}{GeoTerraImage, Pretoria, RSA}{},
-%Gabrielle Ragazzo\affil{5}{Ecology and Evolutionary Biology, Princeton University, Princeton, NJ USA},
-%Justin Sheffield\affil{2}{},
-%Adam Wolf\affil{5}{},
-%Eric Wood\affil{2}{},
-%\and
-%Kelly Caylor\affil{2}{}}
-
+\title{How Cropland Data Errors Impact Our Understanding of Socioeconomic and Environment Change}
 \author[1,2]{Lyndon Estes}
@@
 \author[4]{Stefanus Ferreira}
-\author[5]{Gabrielle Ragazzo}
+\author[5]{Tobias Kuemmerle}
+\author[6]{Gabrielle Ragazzo}
 \author[1]{Justin Sheffield}
-\author[5]{Adam Wolf}
+\author[6]{Adam Wolf}
 \author[1]{Eric Wood}
-\author[1]{Kelly Caylor}
+\author[7]{Kelly Caylor}
 
@@
 \affil[4]{GeoTerraImage, Pretoria, RSA}
-\affil[5]{Ecology and Evolutionary Biology, Princeton University, Princeton, NJ USA}
+\affil[5]{Geography Department, Humboldt University, 10099 Berlin, Germany}
+\affil[6]{Arable Labs, Princeton, NJ USA}
+\affil[7]{Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA USA}
 
+\date{}
 %\contributor{Submitted to Proceedings of the National Academy of Sciences
@@
 %%% If significance statement need, then can use the below command otherwise just delete it.
-%\significancetext{Remote sensing-derived landcover maps underpin much research into global environmental change, particularly in the most rapidly developing regions.
+%\significancetext{Remote sensing-derived land cover maps underpin much research into global environmental change, particularly in the most rapidly developing regions.
 These maps often contain substantial error, yet inadequate validation data makes it impossible to fully quantify map errors and how they affect our understanding of global change processes.
-Using a unique, high quality reference map of cropland distribution, this study comprehensively assesses landcover map bias and accuracy in a major agricultural region of sub-Saharan Africa, finding that map errors are strongly correlated with the landcover density, and can propagate substantial errors in downstream global change analyses.
-Users of landcover maps should select later generation products and aggregate to appropriate resolutions to avoid misleading insights and policies related to global change.
+Using a unique, high quality reference map of cropland distribution, this study comprehensively assesses land cover map bias and accuracy in a major agricultural region of sub-Saharan Africa, finding that map errors are strongly correlated with the land cover density, and can propagate substantial errors in downstream global change analyses.
+Users of land cover maps should select later generation products and aggregate to appropriate resolutions to avoid misleading insights and policies related to global change.
 }
@@
 \begin{abstract}
-{Research into global environmental change is often built upon landcover maps, which can have substantial errors.
-How these errors impact our understanding of global change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions.
-We used a high quality, national-scale cropland map to assess bias and accuracy in current generation landcover maps and four examples of ``downstream'' (landcover-dependent) analyses, two related to physical processes (estimates of vegetative carbon stocks and evapotranspiration) and two to socio-economic processes (gridded crop production estimates and agent-based simulation of household food security), at resolutions ranging from 1-100 km.
-We found that the landcover maps have substantial errors below 25 km resolution, with substantial biases and inaccuracies, particularly from maps derived from coarse resolution sensors.
+{Research into global environmental change is often built upon land cover data, particularly croplands maps, which can have substantial errors.
+How these errors impact our understanding of environmental change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions.
+We used a high quality, national-scale cropland map to assess bias and accuracy within the cropland classes of current generation land cover datasets and four examples of ``downstream'' (land cover-dependent) analyses, two related to physical processes (estimates of vegetative carbon stocks and evapotranspiration) and two to socio-economic processes (gridded crop production estimates and agent-based simulation of household food security), at resolutions ranging from 1-100 km.
+We found that cropland maps have substantial errors below 25 km resolution, with substantial biases and inaccuracies, particularly from maps derived from coarse resolution sensors.
 In some cases these error metrics remain high even when maps are aggregated to 100 km.
@@
 evapotranspiration estimates) the results can be relatively insensitive.
-To avoid these errors, and thus minimize the risk of misunderstanding global change processes or misinforming policy, substantial map aggregation of landcover maps from their base resolution is often needed.
-We provide several guidelines that can help landcover map users select appropriate datasets and  aggregation scales under different use cases.
-
-%To achieve low \emph{bias} in a map-wide statistic (e.g.
-average carbon density), then finer resolutions can be maintained, but if high \emph{accuracy} is desired (e.g.
-spatial comparisons of map values, or change detection), then coarser pixels are needed.
-%Map accuracy is lowest at intermediate levels of cover, thus if the mapped region has highly heterogenous cover (e.g.
-regions dominated by smallholder-farming), then greater aggregation is also necessary.
-From a dataset perspective, maps derived from coarse resolution sensors (e.g.
-MODIS) often need to be aggregated to at \textcolor{red}{50-100} km resolution for bias to be reduced to within +\-10\%, and \textcolor{red}{50-100+} km for mean absolute error (MAE), a measure of accuracy, to be <10\%.
-For comparable rates of bias and error, higher resolution imaging systems (e.g.
-Landsat) produce maps that require less aggregation (1-25 km).
-For regional to global analyses, new fusion products, such as the GeoWiki cropland map, which constrain maps so that the sum of their pixels match a total reported for a larger scale (e.g.
-government cropped area statistics for administrative districts), need just 5 km of aggregation to produce unbiased maps for most applications.
-However, this technique does less to improve accuracy, which may still require $\geq$25 km or more aggregation to achieve <10\% MAE.
-Widely used global datasets for carbon, crop area, and crop yield should be rebuilt on GeoWiki and similar maps, while new methods that fuse human judgement, the latest computer vision-based algorithms, and growing high resolution image archives can help produce the next generation of landcover maps.
+To avoid these errors, and thus minimize the risk of misunderstanding global change processes or misinforming policy, substantial map aggregation of land cover maps from their base resolution is often needed.
+NEW TEXT HERE.
 }
@@
 
-%\keywords{landcover | bias | remote sensing | agriculture | crop yield | harvested area | carbon | agent-based model | landscape}
+%\keywords{land cover | bias | remote sensing | agriculture | crop yield | harvested area | carbon | agent-based model | landscape}
 
@@
 \section*{Introduction}
-The functioning of the Earth System is fundamentally connected to the characteristics of landcover \citep{lambin_modelling_1997}, the physical constituents of the terrestrial surface.
-Human endeavors are strongly governed by and shape landcover, and the characteristics of landcover govern many climate and biogeochemical processes \citep{lambin_dynamics_2003}.
-Our increasing modification of the Earth's surface \citep{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover.
-To fully understand these processes and the nature of global change, an accurate understanding of the nature and distribution of landcover is essential.
+The functioning of the Earth System is fundamentally connected to the characteristics of land cover \citep{lambin_modelling_1997}.
+Our increasing modification of the Earth's surface \citep{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through land cover.
+To fully understand these processes, an accurate understanding of the nature and distribution of land cover is essential.
 
-This importance is understood by a growing number of social, economic, and natural scientists, who are using landcover data to advance understanding of food security \citep{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \citep{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \citep{newbold_global_2015, luoto_predicting_2004}, demographic shifts \citep{linard_assessing_2010}, and other important facets of global change.
+This importance is understood by a growing number of social, economic, and natural scientists, who are using land cover data to advance understanding of food security \citep{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \citep{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \citep{newbold_global_2015, luoto_predicting_2004}, demographic shifts \citep{linard_assessing_2010}, and other important facets of Earth System processes.
 
-The value of the insights resulting from such studies depends upon the veracity of their underlying landcover data, much as a house requires a solid foundation in order to remain standing.
-Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations.
-The reason for this is that landcover data can only practically be derived from satellite imaging, which has several important constraints, related to both physics and analytics, that propagate mapping errors.
-First, in many regions the cover types of interest are smaller than the sensor resolution \citep[e.g.
-smallholder's farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \citep{sweeney_mapping_2015}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
-Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties (e.g.
-agriculture \citep{debats_generalized_2016,estes_platform_2015}) and thus be difficult for the classifier to distinguish, while discretizing a continuous cover type (e.g.
-dividing a forest into different canopy cover classes) can also promote classification error, particularly near class boundaries \citep{foody_status_2002}, as well as confusion about the actual extent of the cover type \citep{sexton_conservation_2015}.
+The value of the insights resulting from such studies depends upon the veracity of their underlying land cover data, much as a house requires a solid foundation in order to remain standing.
+Unfortunately, the evidence to date indicates that this house has shaky foundations.
+The reason for this is that land cover data can only practically be derived from satellite imaging, which has several important constraints that propagate mapping errors.
+First, in many regions the spatial arrangement of cover types is smaller than the sensor resolution \citep[e.g.
+smallholder's farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
+Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties \citep[e.g.
+croplands or savannas;][]{debats_generalized_2016,estes_platform_2016}) and thus be difficult for the classifier to distinguish.
+Discretizing a continuous cover type (e.g.
+dividing a forest into different canopy cover classes) can promote classification error, particularly near class boundaries \citep{foody_status_2002}, as well as confusion about the actual extent of the cover type \citep{sexton_conservation_2015}.
 Furthermore, class definitions often vary between maps, complicating inter-comparison \citep{kuemmerle_challenges_2013}.
-Third, landcover map series are often used to detect changes \citep[e.g.][]{gross_monitoring_2013}, or may use within-class seasonality as a means for discriminating between classes (e.g.
-phenology of cropping versus surrounding vegetation \citep{sweeney_mapping_2015}), but seasonal variability and landcover changes can be easily confused.
-Given these multiple sources of error, landcover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \citep{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}.
-These errors limit our ability to obtain granular, mechanistic understanding of global change processes.
+Third, land cover maps are often used to detect changes \citep[e.g.][]{gross_monitoring_2013}, but seasonal variability and land cover changes can be easily confused.
+Given these multiple sources of error, land cover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \citep{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011,fritz_need_2013}.
+These errors limit our ability to obtain granular, mechanistic understanding of processes related to global change.
 
-These problems with landcover products are known \citep{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \citep[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}.
-What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes.
-Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are typically unavailable for Africa and other rapidly changing regions \citep{see_improved_2015}, which have also been fairly sparsely mapped \citep{yu_meta-discoveries_2014}.
-Our current understanding of map accuracy over such areas is therefore based on a handful of maps \citep{yu_meta-discoveries_2014}, which were assessed using either bottom-up tests made with a relatively small number of ground truth points, or from top-down "sanity checks" made in comparison to aggregated survey data.
-In terms of the former, there are well-developed, best-practice guidelines for selecting and using ground-truth samples to robustly quantify map error, but these are not commonly applied by the broader community of landcover map users (often due to lack of data), which makes it challenging to compare how error varies within and between maps \citep{stehman_global_2012, olofsson_making_2013,olofsson_good_2014,foody_status_2002}.
-The latter form of assessment is made at coarser administrative levels relative to reported statistics \citep[e.g.][]{fritz_comparison_2010}, which can also be suspect \citep{carletto_emperor_2013,fao_action_2013}.
+These problems with land cover products are known \citep{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \citep[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2016}.
+Likewise, the importance of assessing the accuracy of land cover maps is increasingly recognized, and there are well-developed, best-practice guidelines for gathering and using ground-truth samples to robustly quantify map error \citep{foody_status_2002,olofsson_making_2013,olofsson_good_2014,stehman_global_2012}.
+Because comprehensive, spatially representative ground truth data are typically unavailable for rapidly changing regions \citep{see_improved_2015}, what remains an open question is exactly how much the maps researchers typically use deviate from actual land cover, and how this in turn impacts our understanding of environmental change processes.
+Our current understanding of map accuracy over such areas is often based on scarce information or top-down "sanity checks" made in comparison to aggregated survey data \citep{yu_meta-discoveries_2014,larsen_taken_2015}.
 
-Since it is difficult to fully quantify landcover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification\citep{kuemmerle_challenges_2013}.
-There has been some work examining how map error influences climate simulations \citep{ge_impacts_2007}, agricultural land use patterns \citep{schmit_limitations_2006}, carbon flux measurements \citep{quaife_impact_2008}, human population estimates \citep{linard_assessing_2010}, and species distribution models \citep{tuanmu_global_2014}, but these either use simulated landcover errors \citep{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived landcover maps \citep{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}.
-One exception is a Belgian study \citep{schmit_limitations_2006} that used ground-collected farm parcel data to assess how landcover errors bias measurements of agricultural land use patterns, but the extent of this study was fairly small and the validation data were discontiguous.
+%Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are typically unavailable for Africa and other rapidly changing regions \citep{see_improved_2015}, which have also been fairly sparsely mapped \citep{yu_meta-discoveries_2014}.
+
+Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
+Some work has examined how map error influences climate simulations \citep{ge_impacts_2007}, agricultural land use patterns \citep{schmit_limitations_2006}, carbon flux measurements \citep{quaife_impact_2008}, human population estimates \citep{linard_assessing_2010}, and species distribution models \citep{tuanmu_global_2014}, but these either use simulated land cover errors \citep{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived land cover maps \citep{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}, or, in cases where spatially explicit ground validation maps were used for assessment, cover small and discontiguous extents \citep{schmit_limitations_2006}.
 
 Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g.
-3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \citep{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013,debats_generalized_2016}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \citep{searchinger_high_2015}.
+3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \citep{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013,debats_generalized_2016}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the land cover products that have become commonplace in global change research--and which are often used to make policy decisions \citep{searchinger_high_2015}.
 
-In this study, we take advantage of this recent growth in high resolution data to address the call to more thoroughly assess errors and in landcover maps \citep{kuemmerle_challenges_2013, olofsson_good_2014,olofsson_global_2012}, and further examine how these errors might impact our understanding of change in the world's most dynamic regions.
-We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through downstream studies investigating both the physical and socioeconomic components of global change.
-Our objective is to provide scientists and policy-makers who use landcover data an up-to-date assessment of their appropriate uses and limitations.
+In this study, we take advantage of this recent growth in data to address the call to more thoroughly assess errors in land cover maps \citep{kuemmerle_challenges_2013, olofsson_good_2014,olofsson_global_2012}, and further examine how these errors might impact our understanding of socioeconomic and environmental conditions.
+Using a unique, high-quality, spatially extensive map of South African croplands, which was created by expert mappers delineating individual fields visible within high resolution imagery, we conduct spatially comprehensive, bottom-up analyses to answer the following two questions: 1) What is the extent of error in several widely used land cover products?; 2) How do these errors propagate through downstream biophysical and socioeconomic studies?
+
+The answers to these questions provide important insights into how cropland datasets can influence our understanding socioeconomic and environmental processes in South Africa, as well as more broadly throughout the rest of sub-Saharan Africa (SSA), where our current knowledge of the extent and distribution of cropland relies heavily on land cover maps \citep[][]{fritz_comparison_2010, see_improved_2015}.
+
+%Given this dependency and the fact that SSA is poised to undergo a dramatic agricultural expansion during the next several decades \citep{searchinger_high_2015, estes_reconciling_2016}, these findings should also have relevance for contain implications for our broader ability to grasp global change processes.
+
+%The answers to these questions hold significance that extends beyond the borders of South Africa, for several reasons.
+First, South Africa is one of the major agricultural centers in sub-Saharan Africa (SSA), where a rapid agricultural expansion has begun that will play a major role in shaping the future trajectory of global change \citep{searchinger_high_2015, estes_reconciling_2016}.
+Second, South African agriculture spans the diversity of farming systems found in SSA, but has a more advanced commercial sector; it is likely that SSA's broader cropland estate will increasingly resemble South Africa's.
+Third, baseline understanding of the distribution and extent of agricultural land use in SSA is poor because existing land cover maps and agricultural census data are both inaccurate \citep{fritz_comparison_2010, carletto_emperor_2013}.
+By quantifying how cropland datasets can influence our understanding of socioeconomic and environmental processes in South Africa, our study therefore provides insights
+
+%First, South Africa's agricultural sector spans the diversity of farming systems found in sub-Saharan Africa (SSA), but at a more advanced stage of development \citep{hardy_rainfed_2011,estes_using_2014} one of the major agricultural centers in sub-Saharan Africa (SSA), where a rapid agricultural expansion has begun that will play a major role in shaping the future trajectory of global change \citep{searchinger_high_2015, estes_reconciling_2016}.
+Second, our baseline understanding of the distribution and extent of agricultural land use in this region is poor because existing land cover maps and agricultural census data are both inaccurate \citep{fritz_comparison_2010, carletto_emperor_2013}.
+Because South Africa's agricultural sector spans the diversity of farming systems found throughout SSA \citep{hardy_rainfed_2011,estes_using_2014}, our study provides insight into the important role and impact that land cover datasets have on understanding global change dynamics in this critical region.
+
+%The answers to these questions can provide important insights into how cropland datasets can influence broader understanding of global change processes, even though our study is limited to a single country.
+This The reason for this is two-fold.
+First, South Africa is one of the major agricultural centers in sub-Saharan Africa (SSA), where a rapid agricultural expansion has begun that will play a major role in shaping the future trajectory of global change \citep{searchinger_high_2015, estes_reconciling_2016}.
+Our baseline understanding of the distribution and extent of agricultural land use in this region is poor because existing land cover maps and agricultural census data are both inaccurate \citep{fritz_comparison_2010, carletto_emperor_2013}.
+Because South Africa's agricultural sector spans the diversity of farming systems found throughout SSA \citep{hardy_rainfed_2011,estes_using_2014}, our study provides insight into the important role and impact that land cover datasets have on understanding global change dynamics in this critical region.
+
+%, our findings can provide important insight into how errors and biases in cropland datasets may influence our broader understanding of change in key components of the Earth System.
+
+%We focus on sub-Saharan Africa, one of the world's most rapidly changing regions, where a rapid agricultural expansion is underway \citep{searchinger_high_2015,estes_reconciling_2016} yet the accuracy of maps of existing agriculture land is particularly poor \citep{fritz_comparison_2010}.
+
+%We specifically quantify the error and bias are in cropland cover estimates associated with different land cover products, how these vary with scale, and , and how do these vary with scale and by land cover product; 2.
+How do these errors and biases
+
+%We use a unique, high-quality map of South African croplands to conduct a spatially comprehensive, bottom-up quantification of error in several widely used land cover maps, and how these errors can propagate through downstream biophysical and socioeconomic studies.
+South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \citep{searchinger_high_2015,estes_reconciling_2016}, yet is notably lacking trustworthy maps of existing agriculture land \citep{fritz_comparison_2010}.
+
+%Moreover, South Africa's large agricultural sector represents the diversity of farming systems found throughout SSA, ranging from large commercial operations to smallholder farms \citep{hardy_rainfed_2011,estes_using_2014}.
+As such, our findings can provide important insight into how cropland datasets may influence our broader understanding of environmental change.
 
@@
 \section*{Materials and Methods}
-\subsection*{Overview of study area and reference data}
+\subsection*{Datasets}
 In the late 2000s, the South African government commissioned a cropland map that was made by manually interpreting and digitizing fields visible within high resolution satellite imagery \citep{fourie_better_2009}.
-The resulting vectorized field boundaries provide highly accurate data on field sizes and distribution for the period 2009-2011.
-Measured in terms of its ability to distinguish crop fields from other cover types, this dataset is 97\% percent accurate at 1 km resolution, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\% \citep[][SI]{estes_platform_2016}.
+The resulting vectorized field boundaries provide unique, highly accurate data on field sizes and distribution for 2009-2011.
+The accuracy of this dataset was evaluated in a previous study \citep{estes_platform_2016}, in which visual assessment of cropland presence/absence was made within 15,225 4 ha plots (25 sub-plots within 609 1 km$^2$ grids) placed over high resolution satellite imagery, and compared to coverage by the vector boundaries.
+Measured in terms of its ability to distinguish crop fields from other cover types, the results showed these data to be 97\% percent accurate, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\% \citep[see SI for more description, and][]{estes_platform_2016}.
 
-This dataset is particularly valuable because South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \citep{searchinger_high_2015}, yet is notably lacking trustworthy maps of existing agriculture land \citep{fritz_comparison_2010}.
+%This dataset is particularly valuable because South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \citep{searchinger_high_2015}, yet is notably lacking trustworthy maps of existing agriculture land \citep{fritz_comparison_2010}.
 Moreover, South Africa's large agricultural sector represents the diversity of farming systems found throughout SSA, ranging from large commercial operations to smallholder farms \citep{hardy_rainfed_2011,estes_using_2014}.
 
-\subsubsection*{Assessment of landcover map accuracy}
-We used the vectorize field boundaries as a reference layer for evaluating four landcover products representative of the type commonly used in global change studies.
-The first was South Africa's own 30 m resolution 2009 National Landcover map \citep[SA-LC][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are typically available only for individual countries \citep[e.g.][]{fry_completion_2009}.
-The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution 2011 MODIS Landcover products, which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
-The fourth dataset was the new 1 km GeoWiki hybrid-fusion cropland map for Africa \citep{fritz_mapping_2015}, which incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.
-Since GeoWiki provides a continuous bounded value of landcover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.
+We used these vector data as a reference for evaluating four land cover products representative of the type commonly used in global change studies and related areas of research.
+The first was South Africa's own 30 m resolution 2009 National land cover map \citep[SA-LC][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are created for individual countries \citep[e.g.][]{fry_completion_2009}.
+Although global-scale, Landsat-derived maps have recently become available \citep{chen_global_2015}, their reported accuracy for cultivated areas is lower (80-85\% user's accuracy) than that achieved by more intensive, national to sub-national products \citep[e.g.
+90\% user's accuracy][]{sweeney_mapping_2015}.
+The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution MODIS land cover data (for 2011, the final year of the reference interval), which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
+The fourth dataset was the 1 km GeoWiki cropland map for Africa \citep{fritz_mapping_2015}, which fuses the cropland classes of multiple landcover products (including the three assessed here), with the resulting cropland percentages adjusted until their totals sum to match cropland areas reported in national agricultural statistics.
+This ``hybrid-fusion'' approach represents the state of the art in mapping agricultural land cover.
+Since GeoWiki provides a continuous bounded value of land cover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.
 
-To develop the cropland reference map (see SI for accuracy assessment details), we intersected the original field boundary vector maps with a 1 km grid, and calculated the percent of each cell occupied by fields to create the gridded cropland reference map.
-In the resulting grid, we masked out areas classified as communal farmland, because only their outer perimeters were digitized \citep{fourie_better_2009}, which posed a risk of overestimating cropland extent.
-We also excluded the permanent tree crop class, to avoid confusion with landcover maps that exclude these in their cropland classes.
-For the same reason, we used other high resolution landcover maps for selected provinces of South Africa to mask out commercially afforested areas and sugarcane plantations, which were also not included in the reference map.
+%is a ``fusion'' product, in that it merges the results of the first three datasets (and several others), using i) the extent of the agreement between each landcover product and ii) each product's reliability, as determined by crowdsourced validation data, to assign a likelihood of cropland presence, and then calibrating that likelihood into a cropland percentage through calibration against national agricultural statistics.
+and represents the current state-of-the-art in land cover mapping.
+
+To develop percent coverage from the cropland reference map, we intersected the original field boundary vector maps with a 1 km grid, and calculated the percent of each cell occupied by fields.
+In the resulting grid, we masked out areas classified as communal farmland (18.7\% of mapped croplands in the reference data), because only their outer perimeters were digitized \citep{fourie_better_2009}, which risked overestimating cropland extent because the vectors also enclose uncropped areas between adjacent smallholder fields.
+We also excluded from the analysis areas covered by permanent tree crops, sugarcane plantations, and commercially afforested areas, because these three classes were not common to all five cropland datasets.
+To remove permanent tree crops (3.1\% of cropland area), we masked out reference vectors labelled as such, and to exclude the other two cover types, which are not included in the reference data, we relied on information from two other datasets.
+We used a 20 m resolution landcover map of KwaZulu-Natal to mask out the primarily coastal sugarcane farms \citep[93-100\% user's and 76-98\% producer's accuracy for sugarcane classes;][]{geoterraimage_2011_2013}, and used the SA-LC dataset to filter out areas of commercial afforestation, which are mainly located in South Africa's montane areas and do not overlap with arable croplands.
+In both cases, we aggregated these classes, and then masked out any 1 km pixels that had $>$0\% cover of each class.
 The resulting masked reference grid covered 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ was cropland.
@@
 We extracted the cropland classes from SA-LC, MODIS, and GlobCover, and converted these into percent cropland estimates at 1 km resolution.
-Both MODIS and GlobCover had mixed pixel classes, thus, following the approach used in creating the GeoWiki maps \citep{fritz_mapping_2015} we used the upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages.
+Both MODIS and GlobCover had mixed/mosaic classes of cropland and other covers, thus we followed \citet{fritz_mapping_2015} by creating upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages.
 We used the mean map for the main analysis, but estimated error variability using all three versions (SI).
 
-We first evaluated the quality of the landcover product-derived percent cropland maps (hereafter referred to as the ``test maps'').
-Instead of the standard confusion matrix-based accuracy assessment metrics \citep{olofsson_good_2014,olofsson_making_2013}, which apply to categorical landcover maps, we assessed the bias and accuracy of the test maps based on the gridded residuals that resulted when each test map was subtracted from the reference map.
+\vspace{-0.3 cm}
+\subsubsection*{Assessment of land cover map accuracy}
+\vspace{-0.2 cm}
+We first evaluated the quality of the land cover product-derived percent cropland maps (hereafter referred to as the ``test maps'').
+Instead of the standard confusion matrix-based accuracy metrics \citep{olofsson_good_2014,olofsson_making_2013}, which apply to categorical land cover maps, we assessed the bias and accuracy of the test maps based on the gridded residuals that resulted when each test map was subtracted from the reference map.
 Here bias is the mean residual value, weighted by the density of reference cropland, and accuracy is the mean of the absolute values of residuals (also weighted by cropland density), thus accuracy is inversely related to this value.
-We calculated these metrics for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km, in order to evaluate how bias and accuracy changes with scale.
+We calculated these metrics for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km resolutions, in order to evaluate how bias and accuracy changes with observational grain.
 For these aggregated maps, we applied a further weight in calculating error metrics, the number of pixels contributing to each aggregated pixel, to prevent pixels close to national boundaries or where non-target cover types were masked out from having outsize influence on the statistics.
 
-We also assessed how landcover pattern impacts map performance by modeling the correlation between map accuracy and cropland density.
+We also assessed how land cover pattern impacts map performance by modeling the correlation between map accuracy and cropland density.
 To evaluate this relationship, we used magisterial district boundaries (n=354, mean area=3,445 km$^2$; SI) to provide a landscape-scaled unit for calculating characteristic cover density.
@@
 \subsection*{Impact of map error on downstream analyses}
-We then used these maps to conduct four further analyses typical of global change research.
-The first two related to physical processes, namely the calculation of vegetative carbon stocks and the simulation of evapotranspiration.
-The second two were socio-economic in nature: the estimation of agriculture yield and production, and an agent-based model simulating household dynamics in dryland agroecosystems.
-The first and third analyses were relatively simple, in that the variable(s) of interest were mapped onto landcover using empirical relationships.
-The second and fourth relied on more complex numerical methods, where landcover was one of several variables needed to run each model.
+We then used these maps to conduct four analyses typical of global change research: 1) estimation of carbon stocks, 2) simulation of evapotranspiration, 3) disaggregation of crop yield and production, and 4) simulating household dynamics using an agent-based model.
+The first and third analyses were relatively simple, in that the variable(s) of interest were mapped onto land cover using empirical relationships.
+The second and fourth relied on more complex numerical methods, where land cover was one of several variables needed to run each model.
 For the simpler analyses, we examined how results were influenced by map aggregation, while for the more complex cases, our assessments were confined to each numerical model's standard output resolution.
 
-\subsubsection*{Biophysical analyses}
-\emph{Estimating vegetative carbon stocks: }
+\vspace{-0.3 cm}
+\subsubsection*{Estimating vegetative carbon stocks}
+\vspace{-0.2 cm}
 To understand the carbon cycle and climate forcing due to land use change, it is important to have accurate, high resolution maps of vegetative carbon stocks \citep[][]{searchinger_high_2015}.
-One widely used vegetative carbon dataset is that of \citep{ruesch_new_2008}, who mapped estimated carbon density values for different vegetation types to the classes of a global landcover product.
+One widely used vegetative carbon dataset is that of \citep{ruesch_new_2008}, who mapped estimated carbon density values for different vegetation types to the classes of a global land cover product.
 The resulting data were intended to provide a baseline for climate policy by the Intergovernmental Panel on Climate Change (IPCC), as well as input to other land use and biogeochemical analyses \citep{ruesch_new_2008}.
@@
 
-\vspace{4 pt}
-\noindent\emph{Evapotranspiration estimates: }
+\vspace{-0.3 cm}
+\subsubsection*{Estimating evapotranspiration}
+\vspace{-0.2 cm}
 Accurate estimation of hydrological fluxes is critical to understanding how land-atmosphere interactions impact the climate system and runoff \citep{liang_simple_1994}.
-Land surface hydrological models are used to simulate these processes, and depend on landcover maps to provide information on the characteristics of vegetation and other materials covering the surface, as these govern the rates of runoff, infiltration, and evapotranspiration.
+Land surface hydrological models are used to simulate these processes, and depend on land cover maps to provide information on the characteristics of vegetation and other materials covering the surface, as these govern the rates of runoff, infiltration, and evapotranspiration.
 We used the Variable Infiltration Capacity \citep[VIC;][]{liang_simple_1994} land surface hydrology model run with the Africa Flood and Drought Monitor's meteorological data \citep{sheffield_drought_2013} to produce monthly gridded evapotranspiration estimates for South Africa for the years 1979-2010 at 25 km resolution.
-VIC's landcover scheme (derived from AVHRR) provides values for leaf area index (LAI), plant rooting depth, aerodynamic roughness, and several other variables that the modeluses to partition water vapor fluxes into their evaporative and transpirative components.
-We adjusted VIC's base map so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25$^{\circ}$ resolution), then ran one instance of VIC for each of the five landcover schemes.
+VIC's land cover scheme (derived from AVHRR) provides values for leaf area index (LAI), plant rooting depth, aerodynamic roughness, and several other variables that the modeluses to partition water vapor fluxes into their evaporative and transpirative components.
+We adjusted VIC's base map so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25$^{\circ}$ resolution), then ran one instance of VIC for each of the five land cover schemes.
 We then compared the mean annual ET produced by the reference map variant with those from the test maps to assess the degree to which map errors impact evapotranspiration values.
 
-\subsubsection*{Socio-economic analyses}
-\noindent\emph{Gridded crop yield and production data: }
-The spatial variability of crop productivity and production is critical for understanding a host of social, economic, and environmental issues, such as food security, trade, and the potential for agricultural expansion and intensification \citep{licker_mind_2010,monfreda_farming_2008}.
-The most reliable source of such data are national to sub-national agricultural statistics, which are provided for relatively coarse-scaled administrative boundaries.
-To obtain higher spatial resolutions, efforts have been made to disaggregate these statistics using gridded landcover data \citep{ramankutty_farming_2008,monfreda_farming_2008}.
+\vspace{-0.3 cm}
+\subsubsection*{Disaggregating crop yield and production statistics}
+\vspace{-0.2 cm}
+The spatial variability of yield and production is critical for understanding food security, trade, and the potential for agricultural expansion and intensification \citep{licker_mind_2010,monfreda_farming_2008}.
+The most reliable source of such data are national to sub-national agricultural statistics, often available only at relatively coarse-scaled administrative boundaries.
+*****************HERE**************To obtain higher spatial resolutions, efforts have been made to disaggregate these statistics using gridded land cover data \citep{ramankutty_farming_2008,monfreda_farming_2008}.
 These disaggregated datasets, which are constrained to match the agricultural statistics within the boundaries of the areas for which they are reported, have seen increasing use because they are considered to be more accurate than single source methods, and provide consistent data on which to base studies of global change \citep{ramankutty_farming_2008, see_improved_2015}
@@
 
-\subsubsection*{Error and landcover density}
+\subsubsection*{Error and land cover density}
 The generalized additive model revealed primarily non-linear relationships between district MAE and cropland density that were best approximated by a first order polynomial function of cropland density (for all four cropland maps: p$<$0.001 on both terms of quadratic and on smoothing function applied to district centroids; $>$ 85\% deviance explained).
@@
 \caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and C) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
-Dot sizes correspond to district numbers, colors represent the landcover map.}
+Dot sizes correspond to district numbers, colors represent the land cover map.}
 \label{afoto}
@@
 \section*{Discussion}
-This spatially comprehensive, bottom-up landcover map assessment provides new insight into the extent and causes of map error, and its consequences for understanding global change.
+This spatially comprehensive, bottom-up land cover map assessment provides new insight into the extent and causes of map error, and its consequences for understanding global change.
 The analysis was made possible by a spatially comprehensive, high accuracy dataset that likely provides the truest measure of cropland area and distribution currently available for this region.
 Although this reference map is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission, SI), while temporal mismatches between the reference and test maps may account for some of the error we identified, our assessment (SI) suggests that such discrepancies do not appreciably impact our findings.
-Our results are also supported by previous, continental-scale assessments that revealed substantial errors and inconsistencies between landcover maps \citep{fritz_comparison_2010,gross_monitoring_2013}.
+Our results are also supported by previous, continental-scale assessments that revealed substantial errors and inconsistencies between land cover maps \citep{fritz_comparison_2010,gross_monitoring_2013}.
 These similarities, combined with the larger extents of these studies, also suggest that our findings are applicable beyond South Africa's borders.
 
-\subsection*{General guidelines for selecting and using landcover data}
-Our findings regarding bias and accuracy suggest several guidelines for selecting and using landcover data in global change research (Table 1), and suggested aggregation scales.
+\subsection*{General guidelines for selecting and using land cover data}
+Our findings regarding bias and accuracy suggest several guidelines for selecting and using land cover data in global change research (Table 1), and suggested aggregation scales.
 These guidelines vary according to the intended application, what error property the users wants to minimize or maximize, and what tradeoff the user is willing to make between error and resolution.
@@
 
-The first general guideline is that landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions.
-Researchers might need to use these datasets for a variety of reasons, for instance, because they provide consistent, multiple landcover classes at a global scale, or allow changes to be tracked across large scales over time \citep{luoto_predicting_2004}.
+The first general guideline is that land cover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions.
+Researchers might need to use these datasets for a variety of reasons, for instance, because they provide consistent, multiple land cover classes at a global scale, or allow changes to be tracked across large scales over time \citep{luoto_predicting_2004}.
 The appropriate scale of aggregation then depends on which error property the user wants to minimize.
@@
 5).
-The reason that accuracy is not improved is because the statistical constraint cannot correct the omission and commission errors within the landcover maps.
-To minimize these types of errors, different techniques are needed, such as creating a single consensus map from multiple landcover products \citep[e.g.][]{fritz_comparison_2010,tuanmu_global_2014}.
+The reason that accuracy is not improved is because the statistical constraint cannot correct the omission and commission errors within the land cover maps.
+To minimize these types of errors, different techniques are needed, such as creating a single consensus map from multiple land cover products \citep[e.g.][]{fritz_comparison_2010,tuanmu_global_2014}.
 This second kind of data fusion was used in the first stage of creating the GeoWiki map (before statistical adjustment), and it accounts for GeoWiki's higher accuracy scores relative to MODIS and GlobCover (Fig.
@@
   \begin{threeparttable}
-    \caption{Guidelines for selecting and using different classes of landcover product (based on the characteristics of the satellite sensor(s) and map-making methods), according to the general purpose of the application, the error minimization objective, and the range of aggregation needed to achieve the error objective (here we use a maximum acceptable error magnitude of 10\%).
-The preferred landcover product class for each application is indicated in boldface.
+    \caption{Guidelines for selecting and using different classes of land cover product (based on the characteristics of the satellite sensor(s) and map-making methods), according to the general purpose of the application, the error minimization objective, and the range of aggregation needed to achieve the error objective (here we use a maximum acceptable error magnitude of 10\%).
+The preferred land cover product class for each application is indicated in boldface.
 }
@@
        \hline
-Application                     & Objective & Source of Landcover Map & Minimum aggregation* \\
+Application                     & Objective & Source of land cover Map & Minimum aggregation* \\
 \Xhline{1pt}
@@
 Additionally, these higher accuracy, lower bias products often represent just a single time period \citep[e.g.][]{fritz_mapping_2015,tuanmu_global_2014}, thus applications requiring higher frequency (e.g.
-annual), large area landcover maps will depend on MODIS or similar sensors, and therefore much coarser aggregations.
+annual), large area land cover maps will depend on MODIS or similar sensors, and therefore much coarser aggregations.
 
-%GeoWiki was the most accurate among the large scale landcover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
+%GeoWiki was the most accurate among the large scale land cover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
 
 \subsection*{Implications for global change analyses}
-Besides the preceding landcover data selection and usage guidelines, our findings also have broader implications for global change analyses and associated policy decisions.
-The simpler, most direct uses of landcover data have substantial potential to skew our understanding of global change processes.
+Besides the preceding land cover data selection and usage guidelines, our findings also have broader implications for global change analyses and associated policy decisions.
+The simpler, most direct uses of land cover data have substantial potential to skew our understanding of global change processes.
 For example, the global carbon carbon density map of \citep{ruesch_new_2008}, of which we created facsimiles here, has been widely used, but our results suggest that any analyses of the spatial variability of carbon stocks, or of mean carbon densities, can be highly misleading if the map is not first sufficiently aggregated.
@@
 
-What then, of the impacts of landcover map error on more complex analysis?
+What then, of the impacts of land cover map error on more complex analysis?
 Here the results appear to be mixed.
-Map error appears to have had little impact on the overall bias and accuracy of VIC's evapotranspiration (ET) estimates, which contrasts with a related study that examined how landcover maps impact rainfall simulations \citep{ge_impacts_2007}.
+Map error appears to have had little impact on the overall bias and accuracy of VIC's evapotranspiration (ET) estimates, which contrasts with a related study that examined how land cover maps impact rainfall simulations \citep{ge_impacts_2007}.
 There are two reasons for this.
 First, map accuracy increased nearly two-fold by aggregating to 25 km, the resolution of VIC.
-Second, VIC's landcover-derived variables (e.g.
-LAI curves, effective rooting depth) could dampen the effect of map errors on ET calculations when the values of these variables in the adjacent landcovers are similar to that of cropland, which is the case in much of VIC's landcover scheme for South Africa.
-In other regions, where the matrix landcover's LAI values depart substantially from cropland's (e.g.
+Second, VIC's land cover-derived variables (e.g.
+LAI curves, effective rooting depth) could dampen the effect of map errors on ET calculations when the values of these variables in the adjacent land covers are similar to that of cropland, which is the case in much of VIC's land cover scheme for South Africa.
+In other regions, where the matrix land cover's LAI values depart substantially from cropland's (e.g.
 in forest), the ET errors would likely be much larger.
@@
 Overestimates did not matter in this case, because the number of households remained fixed, and their cropland holdings were constrained to match the survey statistics (but this would matter for models that allow these parameters to vary).
-Less predictable was the finding that the spatial distribution of cover can impact model parameterization, with land being left unallocated even when needed, thus such models also require spatial, not just statistical, accuracy in their base landcover maps.
+Less predictable was the finding that the spatial distribution of cover can impact model parameterization, with land being left unallocated even when needed, thus such models also require spatial, not just statistical, accuracy in their base land cover maps.
 
-The errors in the cropland data we tested were large, and it is unlikely that current ABM-based studies, mostly of small regions, are based on such inaccurate landcover data.
-However, as computing power increases, so, too, does the feasible application extent of ABMs, which will then become more reliant on larger-scaled, more erroneous, landcover data, with the attendant risk that insights into socio-ecological systems arising from these models could be highly misleading.
+The errors in the cropland data we tested were large, and it is unlikely that current ABM-based studies, mostly of small regions, are based on such inaccurate land cover data.
+However, as computing power increases, so, too, does the feasible application extent of ABMs, which will then become more reliant on larger-scaled, more erroneous, land cover data, with the attendant risk that insights into socio-ecological systems arising from these models could be highly misleading.
 
-Beyond these examples, there are other applications where landcover map errors could alter understanding and policy.
+Beyond these examples, there are other applications where land cover map errors could alter understanding and policy.
 One of these lies in assessments of land availability for agricultural expansion, which could be particularly erroneous if they use a ``residual approach'' \citep{lambin_estimating_2013}, where existing cropland and other occupied cover types are used to filter a map of potential suitability.
@@
 This sort of analysis entails finding ``win-win'' land use solutions, which are essentially areas that deviate from an expected development cost-benefit relationship; for example, regions of unusually low carbon density but high crop fertility \citep{searchinger_high_2015}.
-The validity of such analysis are entirely dependent In order to be effective, such analyses must be undertaken at the fine scales--hectares, rather than square kilometers--at which land use decisions are made, thus their value for informing land use policy is contingent on the accuracy of their underlying landcover/land use data \citep{searchinger_high_2015}.
+The validity of such analysis are entirely dependent In order to be effective, such analyses must be undertaken at the fine scales--hectares, rather than square kilometers--at which land use decisions are made, thus their value for informing land use policy is contingent on the accuracy of their underlying land cover/land use data \citep{searchinger_high_2015}.
 The validity of such analyses, which must be undertaken at the fine-scales where land use decisions are made (hectares rather than square kilometers), depends on having high quality map data, in order to have confidence that the identified tradeoffs are based on real biophysical properties, not simply map error.
@@
 \subsection*{The way forward}
-Our analysis demonstrates that the landcover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on how accuracy or low bias.
-We provide some basic guidelines (Table 1) for selecting and using landcover data that can help increase confidence in the conclusions drawn from landcover-based studies.
+Our analysis demonstrates that the land cover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on how accuracy or low bias.
+We provide some basic guidelines (Table 1) for selecting and using land cover data that can help increase confidence in the conclusions drawn from land cover-based studies.
 
-Beyond these guidelines, we recommend updating several important, global-scale datasets using the latest generation landcover maps.
+Beyond these guidelines, we recommend updating several important, global-scale datasets using the latest generation land cover maps.
 For example, rebuilding the widely used cropland distribution and yield maps \citep{monfreda_farming_2008, ramankutty_farming_2008} using the GeoWiki map could substantially improve our knowledge about current global agricultural distribution and productivity.
-Similarly, reconstructing the IPCC's Tier 1 vegetative carbon map \citep{ruesch_new_2008} using a combination of the GeoWiki and new global landcover consensus maps \citep[e.g.
+Similarly, reconstructing the IPCC's Tier 1 vegetative carbon map \citep{ruesch_new_2008} using a combination of the GeoWiki and new global land cover consensus maps \citep[e.g.
 GLC-Share or][]{tuanmu_global_2014} the GLC-Share datasets would improve our confidence in regional to global assessments of vegetative carbon stocks.
-Alternatively, the new high-resolution, Landsat-based GLC30 \citep{chen_global_2015}, aggregated and converted to fractional cover classes at 1 km--and fused with high quality national landcover maps, where available \citep[similar to][]{fritz_mapping_2015,fritz_cropland_2011}--may provide a preferable base map for rebuilding these datasets.
+Alternatively, the new high-resolution, Landsat-based GLC30 \citep{chen_global_2015}, aggregated and converted to fractional cover classes at 1 km--and fused with high quality national land cover maps, where available \citep[similar to][]{fritz_mapping_2015,fritz_cropland_2011}--may provide a preferable base map for rebuilding these datasets.
 
-Fusing human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power appears to offer the greatest promise for improving the quality of landcover maps.
+Fusing human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power appears to offer the greatest promise for improving the quality of land cover maps.
 The GeoWiki project and new high resolution global forest change maps \citep{hansen_high-resolution_2013,sexton_conservation_2015} are prominent examples of these newer methods.
@@
 Merging this crowdsourced approach for creating vector maps of discrete cover types with newer computer vision-based mapping algorithms \citep{debats_generalized_2016} is a near-term goal of this project, as the former can interactively provide high quality training and test data (by using humans' superior capacity to delineate discrete cover types in noisy images \citep{estes_platform_2016}) for the latter, which has the advantage of speed and the ability to handle high-dimensional datasets.
-Approaches such as these will be needed to create the next generation of landcover maps, so that we can gain a clearer, unbiased understanding of global change.
+Approaches such as these will be needed to create the next generation of land cover maps, so that we can gain a clearer, unbiased understanding of global change.
 
@@
 \bibliographystyle{gcbnourl} 
-{\footnotesize \bibliography{/Users/lestes/Dropbox/publications/full.bib}}
+{\footnotesize \bibliography{/Users/lestes/Dropbox/publications/fullbib.bib}}
 

### 2016-12-06 R/landcover-functions.R

@@ -63,7 +63,7 @@ disaggregate_rast_list <- function(names1, names2, rlist, rmask) {
         r <- rlist[[x]][[y]][[1]]
       } else {
         r <- disaggregate(rlist[[x]][[y]][[1]], fact = fact)
-        r <- mask(crop(r, rmask), rmask)
+        r <- raster::mask(crop(r, rmask), rmask)
       }
     })
     named_out(l1, names2)

### 2017-01-27 Rmd/c10-location-impact.Rmd
Edits to manuscript to 27 Jan, 2017. Added/edited code in c3-cropland-1.Rmd (saved aggregated cropland and area weights raster lists to output rda files), c4-cropland-2.Rmd (calculated overall cropland areas by dataset and with different scales of aggregation).

@@
 load("external/ext_data/carbon-tier1.rda")
-
+load("external/ext_data/cropland-1km.rda")
 ```
@@
 # top decile of...
+# ...cropland
+crclass <- lapply(cropl, function(x) {  # x <- cyld_agg$f1[[1]]
+  r <- x * (x > 0)
+  r[r == 0] <- NA
+  rq <- quantile(r, c(0, 0.9, 1))
+  types <- cut(r, breaks = rq)
+  types > 1
+})
+
 # ...yield
@@
 
-mudists <- lapply(list(yclass, pclass, cclass, effclass), function(x) {
-  # x <- yclass
+mudists <- lapply(list(crclass, yclass, pclass, cclass, effclass), function(x) {
+  # x <- cclass
   mudist <- sapply(2:5, function(y) {  # y <- 2
@@
 })  
-names(mudists) <- c("yield", "prod", "carb", "eff")
+names(mudists) <- c("cropland", "yield", "prod", "carb", "eff")
 
@@
 cols <- c("red", "orange3", "green4", "blue")
-pcol <- makeTransparent(cols, alpha = c(60))
+pcol <- makeTransparent(cols, alpha = c(120))
+pcol2 <- makeTransparent(cols, alpha = c(30))
+
 plot(1:4, pch = 20, col = pcol)
-bs <- do.call(rbind, mudists[c(1, 3)])
+bs <- do.call(rbind, mudists[c(1, 2, 4)])
 lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
 
-pdf(full_path(p_fig, "location_offset.pdf"), height = 4, width = 4, 
+pdf(full_path(p_fig, "location_offset.pdf"), height = 3.5, width = 4, 
     bg = "transparent")
-par(mar = c(4, 4, 1, 0), bg = "transparent")
-bp <- barplot(bs, beside = TRUE, col = ggplot2:::interleave(cols, pcol), las = 2,
-              border = "transparent", xaxt = "n", ylab = "km", ylim = c(0, 16), 
-              yaxt = "n")
-axis(1, at = colMeans(bp), labels = lcnms, lty = 0)
-axis(2, at = seq(0, 16, 2), labels = seq(0, 16, 2), las = 2)
-legend(x = 9.5, y = 16, legend = rep("", 4), pch = 15, col = cols, adj = 0,
+par(mar = c(2, 3.5, 1, 0), bg = "transparent")
+bp <- barplot(bs, beside = TRUE, col = ggplot2:::interleave(cols, pcol, pcol2), 
+              las = 2, border = "transparent", xaxt = "n", ylab = "", 
+              ylim = c(0, 18), yaxt = "n")
+axis(1, at = colMeans(bp), labels = lcnms, lty = 0, cex.axis = 0.9, 
+     mgp = c(2, 0.25, 0))
+axis(2, at = seq(0, 18, 2), labels = seq(0, 18, 2), las = 2)
+mtext("km", side = 2, line = 2.25)
+legend(x = 12.5, y = 18, legend = rep("", 4), pch = 15, col = cols, adj = 0,
        pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0, horiz = TRUE)
-legend(x = 9.5, y = 15.25, legend = rep("", 4), pch = 15, adj = 0, col = pcol, 
+legend(x = 12.5, y = 17.25, legend = rep("", 4), pch = 15, adj = 0, col = pcol, 
        pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
-text(x = 10, y = 15.1, labels = "Yield", pos = 2, cex= 0.8)
-text(x = 10, y = 14.35, labels = "Carbon", pos = 2, cex = 0.8)
+legend(x = 12.5, y = 16.5, legend = rep("", 4), pch = 15, adj = 0, col = pcol2, 
+       pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
+text(x = 13, y = 17, labels = "Cropland", pos = 2, cex= 0.8)
+text(x = 13, y = 16.25, labels = "Yield", pos = 2, cex = 0.8)
+text(x = 13, y = 15.5, labels = "Carbon", pos = 2, cex= 0.8)
 dev.off()

### 2017-01-27 Rmd/c4-cropland-2.Rmd
Edits to manuscript to 27 Jan, 2017. Added/edited code in c3-cropland-1.Rmd (saved aggregated cropland and area weights raster lists to output rda files), c4-cropland-2.Rmd (calculated overall cropland areas by dataset and with different scales of aggregation).

@@
 load(full_path(p_data, "bias_stats_gti.rda"))  # bias statistics
+load(full_path(p_data, "agg_cropland.rda"))  # aggregated cropland
 # cellStats(!is.na(dlist_act$modmu$f1$g2011), sum)
@@
 ```{r, eval = FALSE}
-paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
-gti <- lapply(paths, raster)
-names(gti) <- c("g2007", "g2011")
-gti <- lapply(gti, function(x) mask(x, namask))  # apply namask
-snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
+# paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
+# gti <- lapply(paths, raster)
+# names(gti) <- c("g2007", "g2011")
+# gti <- lapply(gti, function(x) raster::mask(x, namask))  # apply namask
+# snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
 
 # recreate cropland percentages
-lc2007 <- lapply(snms, function(x) gti$g2007 - dlist_act[[x]]$f1$g2007)
-lc2011 <- lapply(snms, function(x) gti$g2011 - dlist_act[[x]]$f1$g2011)
+# lc2007 <- lapply(snms, function(x) gti$g2007 - dlist_act[[x]]$f1$g2007)
+# lc2011 <- lapply(snms, function(x) gti$g2011 - dlist_act[[x]]$f1$g2011)
 # plot(lc2007[[2]] - lc2011[[2]])  # same thing, as it should be
@@
 # calculate km2 cropland in LC datasets
-lckm2 <- sapply(lc2011, function(x) cellStats(x / 100, sum))
-refkm2 <- sapply(gti, function(x) cellStats(x / 100, sum))
-refkm2[2] / refkm2[1]  # 3.6 percent more in 2011 than 2007
+awgts <- lapply(area_wgts, function(x) x[[1]])
+lckm2 <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
+  gtia <- cellStats(gti_agg[[x]]$g2011 / 100 * awgts[[x]], sum)
+  lca <- sapply(snms, function(y) {
+    cellStats(lc_agg[[x]][[y]] / 100 * awgts[[x]], sum)
+  })
+  o <- c(gtia, lca)
+}))
 
-# bias in total cropland relative to GTI
-stats2007 <- round((refkm2[2] - lckm2) / refkm2[2] * 100, 1)
-stats2011 <- round((refkm2[1] - lckm2) / refkm2[1] * 100, 1)
+# lckm2 <- sapply(lc_agg$f1[snms], function(x) cellStats(x / 100, sum))
+# lckm2 <- sapply(lc2011, function(x) cellStats(x / 100, sum))
+# refkm2 <- sapply(gti, function(x) cellStats(x / 100, sum))
+
+# area of each GTI dataset
+refkm2 <- sapply(gti_agg$f1, function(x) cellStats(x / 100, sum))
+refkm2[2] / refkm2[1]  # 3.7 percent more in 2011 than 2007
+
+# bias in total cropland relative to GTI, at 1 km scale
+stats2007 <- round((refkm2[2] - lckm2[1, ]) / refkm2[2] * 100, 1)
+stats2011 <- round((refkm2[1] - lckm2[2, ]) / refkm2[1] * 100, 1)
 
@@
 # sapply(gti, function(x) cellStats(!is.na(x) , sum))
-totkm2 <- cellStats(!is.na(gti$g2011) , sum)
+
+totkm2 <- cellStats(!is.na(gti_agg$f1$g2011) , sum)
 save(stats2007, stats2011, totkm2, refkm2, 
@@
 ```{r, eval = FALSE}
-# Create weights maks
+# Create weights mask
 fact <- c(5, 10, 25, 50, 100)

### 2017-01-27 paper/main.tex
Edits to manuscript to 27 Jan, 2017. Added/edited code in c3-cropland-1.Rmd (saved aggregated cropland and area weights raster lists to output rda files), c4-cropland-2.Rmd (calculated overall cropland areas by dataset and with different scales of aggregation).

@@
 
-This importance is understood by a growing number of social, economic, and natural scientists, who are using land cover data to advance understanding of food security \citep{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \citep{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \citep{newbold_global_2015, luoto_predicting_2004}, demographic shifts \citep{linard_assessing_2010}, and other important facets of Earth System processes.
+work in \citep{verburg_challenges_2011}, who says how fundamental landcover is to understanding global change.
 
-The value of the insights resulting from such studies depends upon the veracity of their underlying land cover data, much as a house requires a solid foundation in order to remain standing.
+This importance is understood by a growing number of social, economic, and natural scientists, who are using land cover data to advance understanding of food security \citep{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \citep{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \citep{newbold_global_2015, luoto_predicting_2004}, demographic shifts \citep{linard_assessing_2010}, and other important facets of Earth System processes (we hereafter collectively refer to this sort of work as Earth System Science (ESS), of which global change research is a subset).
+
+The value of the insights resulting from ESS studies depends upon the veracity of their underlying land cover data, much as a house requires a solid foundation in order to remain standing.
 Unfortunately, the evidence to date indicates that this house has shaky foundations.
@@
 First, in many regions the spatial arrangement of cover types is smaller than the sensor resolution \citep[e.g.
-smallholder's farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
+smallholder's farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015,fritz_identifying_2008}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
 Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties \citep[e.g.
-croplands or savannas;][]{debats_generalized_2016,estes_platform_2016}) and thus be difficult for the classifier to distinguish.
+croplands or savannas;][]{fritz_identifying_2008,verburg_challenges_2011,debats_generalized_2016}) and thus be difficult for the classifier to distinguish.
 Discretizing a continuous cover type (e.g.
 dividing a forest into different canopy cover classes) can promote classification error, particularly near class boundaries \citep{foody_status_2002}, as well as confusion about the actual extent of the cover type \citep{sexton_conservation_2015}.
-Furthermore, class definitions often vary between maps, complicating inter-comparison \citep{kuemmerle_challenges_2013}.
+Furthermore, class definitions often vary between maps, complicating inter-comparison \citep{kuemmerle_challenges_2013,fritz_identifying_2008}.
 Third, land cover maps are often used to detect changes \citep[e.g.][]{gross_monitoring_2013}, but seasonal variability and land cover changes can be easily confused.
@@
 Likewise, the importance of assessing the accuracy of land cover maps is increasingly recognized, and there are well-developed, best-practice guidelines for gathering and using ground-truth samples to robustly quantify map error \citep{foody_status_2002,olofsson_making_2013,olofsson_good_2014,stehman_global_2012}.
-Because comprehensive, spatially representative ground truth data are typically unavailable for rapidly changing regions \citep{see_improved_2015}, what remains an open question is exactly how much the maps researchers typically use deviate from actual land cover, and how this in turn impacts our understanding of environmental change processes.
+Because comprehensive, spatially representative ground truth data are typically unavailable for rapidly changing regions \citep{see_improved_2015,kuemmerle_challenges_2013}, what remains an open question is exactly how much the maps researchers typically use deviate from actual land cover, and how this in turn impacts our understanding of environmental change processes.
 Our current understanding of map accuracy over such areas is often based on scarce information or top-down "sanity checks" made in comparison to aggregated survey data \citep{yu_meta-discoveries_2014,larsen_taken_2015}.
@@
 Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
+Although a number of previous studies have examined how land cover maps errors can propagate, these are primarily assessed using either simulated errors, relative differences in existing land cover maps, or ground validation data covering relatively small areas.
+\citep[e.g.][]{ge_impacts_2007, linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014,schmit_limitations_2006}.
+
+%Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
 Some work has examined how map error influences climate simulations \citep{ge_impacts_2007}, agricultural land use patterns \citep{schmit_limitations_2006}, carbon flux measurements \citep{quaife_impact_2008}, human population estimates \citep{linard_assessing_2010}, and species distribution models \citep{tuanmu_global_2014}, but these either use simulated land cover errors \citep{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived land cover maps \citep{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}, or, in cases where spatially explicit ground validation maps were used for assessment, cover small and discontiguous extents \citep{schmit_limitations_2006}.
@@
 \section*{Materials and Methods}
+\vspace{-0.3 cm}
 \subsection*{Datasets}
+\vspace{-0.2 cm}
 In the late 2000s, the South African government commissioned a cropland map that was made by manually interpreting and digitizing fields visible within high resolution satellite imagery \citep{fourie_better_2009}.
@@
 \vspace{-0.3 cm}
-\subsubsection*{Assessment of land cover map accuracy}
+\subsection*{Assessment of cropland map errors}
 \vspace{-0.2 cm}
@@
 Instead of the standard confusion matrix-based accuracy metrics \citep{olofsson_good_2014,olofsson_making_2013}, which apply to categorical land cover maps, we assessed the bias and accuracy of the test maps based on the gridded residuals that resulted when each test map was subtracted from the reference map.
-Here bias is the mean residual value, weighted by the density of reference cropland, and accuracy is the mean of the absolute values of residuals (also weighted by cropland density), thus accuracy is inversely related to this value.
+Here bias is the mean residual value, weighted by the density of reference cropland, and accuracy is the mean of the absolute values of residuals (also weighted by cropland density), thus lower values signify higher accuracy.
 We calculated these metrics for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km resolutions, in order to evaluate how bias and accuracy changes with observational grain.
@@
 
+\vspace{-0.3 cm}
 \subsection*{Impact of map error on downstream analyses}
+\vspace{-0.2 cm}
 We then used these maps to conduct four analyses typical of global change research: 1) estimation of carbon stocks, 2) simulation of evapotranspiration, 3) disaggregation of crop yield and production, and 4) simulating household dynamics using an agent-based model.
@@
 The most reliable source of such data are national to sub-national agricultural statistics, often available only at relatively coarse-scaled administrative boundaries.
-*****************HERE**************To obtain higher spatial resolutions, efforts have been made to disaggregate these statistics using gridded land cover data \citep{ramankutty_farming_2008,monfreda_farming_2008}.
-These disaggregated datasets, which are constrained to match the agricultural statistics within the boundaries of the areas for which they are reported, have seen increasing use because they are considered to be more accurate than single source methods, and provide consistent data on which to base studies of global change \citep{ramankutty_farming_2008, see_improved_2015}
+To obtain higher spatial resolutions, disaggregation of these statistics using gridded land cover data is common \citep{ramankutty_farming_2008, monfreda_farming_2008,schierhorn_post-soviet_2013}.
 
-We used these methods to disaggregate the harvested area for maize (South Africa's largest crop \citep{estes_projected_2013}) on top of our reference and cropland maps, followed by yields, which were assigned to cells having harvested areas greater than zero.
+We used these methods to first disaggregate the harvested area for maize (South Africa's largest crop \citep{estes_projected_2013}) onto our cropland maps, followed by yields, which were assigned to cells having harvested areas greater than zero.
 The first step in this process entails adjusting cropland percentages so that their totals match census-derived cropland area estimates \citep{ramankutty_farming_2008}.
@@
 
-\vspace{4 pt}
-\noindent\emph{Agent-based simulation of household food security: }
-Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \citep{berger_creating_2006}.
-To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistically represents the social and biophysical features of the study region \citep{berger_creating_2006}.
+\vspace{-0.3 cm}
+\subsubsection*{Agent-based simulation of household food security }
+\vspace{-0.2 cm}
+Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making, to analyze socio-ecological system dynamics, and to facilitate improved policy-making \citep{berger_creating_2006}.
+To obtain robust insights, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistically represents the social and biophysical features of the study region \citep{berger_creating_2006}.
 
-In our example, we used an ABM of household food security that simulates food production by individual farming households (the agents) in response to weather and management capacity \citep{chen_dependency_2013}.
-The model is initialized such that each household is allocated a designated share of cropland derived from household survey statistics, and then calculates annual household-level agricultural production based on maize yields simulated by the DSSAT cropping system model \citep{jones_modelling_2003}.
-Yield estimates vary as a function of soil properties, management inputs, and weather.
-The former two factors differ between households, as do the total areas of cropland, thus total production can vary substantially between households.
-The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity.
+In our example, we used an ABM of household food security that simulates food production by individual farming households \citep[the agents; ][]{chen_dependency_2013}.
+The model (described briefly here and in more detail in the SI) is initialized so that each household is allocated a designated share of cropland (the number of households and their cropland areas are derived from survey statistics), on which annual household's crop production (maize) is simulated as a function of its field area, local weather, soil properties, and management actions.
+As these factors differ between households, household-level crop production can vary substantially.
+The initialization process iteratively assigns households to the landscape as a function of neighbor and cropland proximity, ensuring that households are grouped into communities and that their fields are within a realistic proximity.
 To achieve this, the model first takes a weighted (by cropland area frequency) random draw of 100 households and places these within the simulated landscape, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household.
 This process is iterated until all households are assigned cropland, or all available cropland is allocated.
-The model is thus considered to be well-calibrated when all households are allocated their appropriate area of cropland, and all cropland is allocated to households.
+The model is considered to be well-calibrated when all households are allocated cropland, and all cropland is allocated to households.
 
 Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g.
-districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of individual farmers.
-To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, SI) in the eastern part of the country with similar climate and 28-45\% of their areas devoted to cropland.
-The frequency distributions of household cropland holdings were derived from household survey data obtained from Zambia (no equivalent data were available for South Africa).
-We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area (2.2 ha), and preserved the cropland area distributions by multiplying the total number of households by the frequencies.
-To create cropland surfaces for each district, we disaggregated all five cropland maps to binary cropland/non-cropland cover types with 100 m resolution, in order to match the survey-recorded average field size of 1 ha.
-The model was then run separately for each district, with the household agents planting and managing a maize crop throughout the growing season, with possible final maize yields drawn from a look-up table that relates yield to particular climatic outcomes and management decisions, as simulated by the DSSAT model.
+district, rather than country, scales) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of individual farmers.
+To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$) in eastern South Africa with similar climate and 28-45\% of cropland coverage.
+The frequency distributions of household cropland holdings were derived from survey data obtained from Zambia (no equivalent data were available for South Africa).
+We used these statistics to calculate the ``true'' number of households per district, as well as household cropland area distributions.
+To create cropland surfaces for each district, we disaggregated all five cropland maps to 100 m, binary cropland/non-cropland cover maps (to match the survey-recorded average field size of 1 ha).
+We ran the model separately for each district, with household agents planting and managing a maize crop throughout the growing season.
 
 We ran each ABM simulation for a single season, once per district and when initialized by each of the cropland maps (20 simulations total). 
-To examined how map errors impacted the land allocation process and household food production estimates, we calculated three variables: the percent of unallocated cropland, land deficit, and food deficit.
+To examine how map errors impacted the land allocation process and household food production estimates, we calculated three variables: the percent of unallocated cropland, land deficit, and food deficit.
 The percent of unallocated cropland is the share of the district's total cropland area that was not assigned to any household, which measures how effective the model was in matching household agents to available cropland resources.
@@
 
+\vspace{-0.3 cm}
+\subsection*{The impact of map error on identifying specific locations}
+\vspace{-0.2 cm}
+The bias and accuracy metrics reflect the degree to quantitative estimates are influenced by land cover map errors.
+However, land cover data may also be used to identify specific locations \citep[e.g.
+areas of high agricultural potential and low ecological cost;][]{estes_reconciling_2016}, as opposed to general quantities.
+It is therefore important to also assess how map errors can impact the ability to accurately locate specific features of interest.
+To evaluate this, we calculated the mean euclidean distance (in km) between pixels representing a specific feature within the test maps relative to their nearest neighboring pixels representing the same feature in the reference map.
+The features in this analysis were simply those locations falling within the upper deciles of a) cropland cover, b) carbon density (based on the average carbon density of non-crop vegetation types), and c) crop yield.
+We confined this analysis to the 1 km resolution maps, as higher spatial resolutions are typically required when maps are used to identify locations rather than estimating quantities \citep[e.g.][]{estes_reconciling_2016}.
+
 \vspace{-0.5 cm}
 \section*{Results}
-\subsection*{Map quality}
-\subsubsection*{Bias and accuracy}
+\vspace{-0.3 cm}
+\subsection*{Cropland map errors}
+\vspace{-0.2 cm}
 Our reference cropland map indicated that crop fields covered nearly 10\% of the total study area in the 2009-2011 time period.
@@
 
-\subsubsection*{Error and land cover density}
+%\subsection*{Relationship between map error and cropland density}
 The generalized additive model revealed primarily non-linear relationships between district MAE and cropland density that were best approximated by a first order polynomial function of cropland density (for all four cropland maps: p$<$0.001 on both terms of quadratic and on smoothing function applied to district centroids; $>$ 85\% deviance explained).
@@
 
-\subsection*{The impact of map error on physical analyses}
+\vspace{-0.3 cm}
+\subsection*{The impact of map error on downstream analyses}
+\vspace{-0.2 cm}
 \subsubsection*{Carbon estimates}
+\vspace{-0.2 cm}
 The spatial patterns of test map errors transmitted into substantial carbon estimation errors, with the sign varying as a function of the density of carbon adjacent to croplands (SI).
@@
 
+\vspace{-0.3 cm}
 \subsubsection*{Evapotranspiration estimates}
+\vspace{-0.2 cm}
 Compared to the carbon analysis, the bias and accuracy in evapotranspiration (ET) calculated using the VIC model was negligible, averaging less than than +/-2\%.
@@
 
-\subsection*{Socio-economic analyses}
-\subsubsection*{Gridded crop yield and production data}
+\vspace{-0.3 cm}
+\subsubsection*{Downscaling crop yield and production data}
+\vspace{-0.2 cm}
 Maize yields disaggregated onto the test maps showed some marked differences relative to the reference map, but only at the margins of the major crop production areas where cropland is sparser (SI).
@@
 %\vspace{-0.5 cm}
-\begin{figure}[!hb]
+\begin{figure}[!ht]
 %\vspace{-0.75 cm}
@@
 
+\vspace{-0.3 cm}
 \subsubsection*{Agent-based model of household food security}
+\vspace{-0.2 cm}
 In terms of impact to agent-based model simulation, where cropland map errors were negative (indicating a cropland overestimate by the test maps), the percent of land left unallocated had a straight one-to-one relationship with the percentage of overestimation (Fig.
@@
 
-\begin{figure}[!ht]
+\begin{figure}[ht]
 \centerline{\includegraphics[width=.55\textwidth]{figures/figure6.pdf}}
@@
 
+\vspace{-0.3 cm}
+\subsection*{Location errors}
+\vspace{-0.2 cm}
+The average distance between areas containing the highest cropland densities (upper decile) in the reference map and those delineated by the test maps ranged from 1.1 km for SA-LC to 18.2 km for GlobCover, with MODIS (10.1 km) and GeoWiki (2.8 km) had intermediate offsets (Fig.
+7).
+Locational errors in maps indicating the highest yielding areas showed a similar pattern, with a range of 0.8-14.2 km (SA-LC and GlobCover) and intermediate errors of 5.8-7.5 km (GeoWiki and MODIS).
+For areas of highest carbon density, locations identified by the MODIS-derived map were most distant from those shown by the reference map (11.3 km), followed by GeoWiki (7.4 km), GlobCover (6.8 km), and SA-LC (3.7 km).
+
+%The analyses above primarily dealt with impact of landcover map errors on estimating quantitative values from maps.
+However, there are many applications in which land cover data are needed to identify specific locations, such as areas to be targeted for development, where quantitative information is used to target specific locations \citep{estes_reconciling_2016}.
+
+\begin{figure}[!ht]
+\centerline{\includegraphics[width=.55\textwidth]{figures/location_offset.pdf}}
+\caption{Average nearest neighbor distances (in km) between pixels representing features identified by the reference map versus those identified from the test maps.
+Bar colors indicate the different features (and thus contributing maps), which were delineated by selecting pixels with values greater than the 90th percentile: densest cropland (solid bars); highest maize yield (medium transparent bars); highest carbon density (most transparent bars).}
+\label{afoto}
+\end{figure}
+
+\vspace{-0.5 cm}
 \section*{Discussion}
-This spatially comprehensive, bottom-up land cover map assessment provides new insight into the extent and causes of map error, and its consequences for understanding global change.
-The analysis was made possible by a spatially comprehensive, high accuracy dataset that likely provides the truest measure of cropland area and distribution currently available for this region.
+\vspace{-0.2 cm}
+
+\subsection*{Error and its propagation}
+\subsubsection*{Quantities}
+Aggregation can minimize error, but issue of MAUP, and might not be the application needed.
+
+\subsubsection*{Identifying locations}
+A few to 10s of km. Doesn't sounds like much, but if need to precisely identify a location, can make a big difference. 
+Sensitivity is generally within the same range for each application. 
+ABM results indicate locational errors also.
+
+\subsection*{Sources of error}
+Mixed classes - MODIS versus GlobCover. 
+Pixel size
+Mosaic classes - cropland density/error relationship
+Within-class variability (cropland particularly pronounced)
+
+caveats: mention among MAUP concerns that relying on percentages helps avoid some of the issues of MAUP that relate to rasterization error and trying to aggregate classes (cites for this).
+Of course doing this loses one of the main sources of geographic information (configuration), but this is one of the many aspects that leads to the MAUP problem in any case.
+
+Reasons for not evaluating GLC30 (lower accuracy)
+
+\citep{verburg_challenges_2011} Relationship between class accuracy and heterogeneity of the landscape.
+
+cite frey and smith (2007) as downstream analyses.
+
+\subsection*{Broader implications}
+
+Broader conclusions are necessarily limited, because of single country involved in this analysis. 
+Parochial study. But suggests important findings for rest of Africa.
+
+However, support from previous studies.
+Properties revealed here in this study are relevant.  Mosaic classes.  Background vegetation similar to cropland (savanna, grassland). Should expect even high error.
+
+How to deal with it: Limited
+
+Synthetic assessments can help you determine how much impact error can make.
+
+Implications for different types of analyses???
+Previous studies as backup. Fritz et al, etc.
+
+\subsection*{Conclusion/way forward}
+
+Carefully evaluate a product for its sensitivity for a particular application \citep{fritz_highlighting_2011}.
+Here we show sensitivity of several important products for a variety of applications.
+Restricted geographically, so limited inference, but some general inferences may be drawn.
+Carefully crafted Landsat products for particular scale (higher density of validation points, local knowledge).
+
+Solutions: For fine features below sensor resolution, convert to quantitative measures such as SMA (tough for cropland though, but maybe with MESMA for different crop categories).
+For homogenous cover types with little within-class variability.
+
+Between-date classification error metrics (cite to Pontius, Verburg brings it up also)
+
+Care in aggregation methods worth emphasizing \citep{verburg_challenges_2011,dendoncker_exploring_2008,raj_analysing_2013,moody_influence_1995}, can lead to bias, which is influenced by spatial arrangement of cover types.
+
+points to make:
+
+Fusion of census data with RS: improve quantity estimates. But careful for accuracy of census data. Large
+
+Match dataset to application; merge complementary sources where available \citep{verburg_challenges_2011}
+
+Fritz and See (2008) about comparing maps with different themes\\
+
+\citep{nijland_optimizing_2009}.
+Spatial scale of pixel matters with respect to mapping accuracy.
+Average local variance is related to size of features being mapped.
+Averaging (smoothing) pixel value is also a better way of increasing accuracy (this is in estimating vegetation parameters).
+``Pixels are sufficiently small to separate the studied vegetation from other land cover units.
+This is especially important in complex scenes.''  Good point to make for ignoring larger pixel sensors.
+On the other hand, for certain applications, like estimating vegetation that might be better to have larger pixel sizes.
+
+They recommend experimental determination of optimal pixel size for all applications.
+Experimental determination of impact of error might also be appropriate.
+Careful consideration of pixel size emphasized, and shape of segments.
+
+This is for both discussion of MAUP issues and recommendations for way forward.
+
+MAUP in our case is minimized--it is at the level imposed by sensor resolution, but it is minimized by converting to percentage error estimates through aggregation [think I need to show this by showing area estimation with scale].
+Admittedly we do not deal as much with spatial aspects of MAUP, cite numerous studies that look at this.
+But our weighted assessment of accuracy does give some indication of spatial error--it is akin to the likely error experienced at any given point on the map.
+Weighting corrects for unequal spatial distribution of cover type.
+
+\citep{nelson_analysing_2001} Talks about ecological fallacy, scale, etc.  With spatial data, results should presented both in map form as well as tabulated results.
+
+Our contribution: 
+Motivation for our paper found in here. Our contribution is to more explicitly quantify, over a larger scale, the types of uncertainties in land cover data.
+
+Classification is scale-dependent.
+
+\citep{hay_comparison_2003}
+Up-scaling should be guided by the intrinsic scale of image objects that compose a scene.
+
+We show how error changes with scale, but that does not mean that coarser scales lead to more accurate findings.
+Coarser scale may be less appropriate, or provide misleading results \citep{nelson_analysing_2001}.
+
+Dimensions of the pixel should also be examined, although we didn't do that here.
+To a certain extent, this applies more to issues of category than percent coverage (hope for support on this).
+MAUP problem is also enforced by sensor resolution \citep{hay_comparison_2003,marceau_remote_1994}, so above that we are minimizing issue of aggregation by taking percent coverage.
+
+But zonation/pixel-size chosen should be user and purpose-designed.
+
+\citep{marceau_remote_1999}
+Areal units are modified by changing the sensor as well as by aggregating pixels.
+
+Prior studies
+%Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
+Some work has examined how map error influences climate simulations \citep{ge_impacts_2007}, agricultural land use patterns \citep{schmit_limitations_2006}, carbon flux measurements \citep{quaife_impact_2008}, human population estimates \citep{linard_assessing_2010}, and species distribution models \citep{tuanmu_global_2014}, but these either use simulated land cover errors \citep{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived land cover maps \citep{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}, or, in cases where spatially explicit ground validation maps were used for assessment, cover small and discontiguous extents \citep{schmit_limitations_2006}.
+Add in \cite{dendoncker_exploring_2008} as one related to Schmitt using comprehensive ground-truth data to assess scaling errors.  Part of mini-review.
+
+%This spatially comprehensive, bottom-up land cover map assessment provides new insight into the extent and causes of map error, and its consequences for understanding global change.
+The analysis was made possible by a high accuracy cropland map that likely provides the truest measure of cropped area and distribution currently available for this region.
 Although this reference map is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission, SI), while temporal mismatches between the reference and test maps may account for some of the error we identified, our assessment (SI) suggests that such discrepancies do not appreciably impact our findings.
 Our results are also supported by previous, continental-scale assessments that revealed substantial errors and inconsistencies between land cover maps \citep{fritz_comparison_2010,gross_monitoring_2013}.
-These similarities, combined with the larger extents of these studies, also suggest that our findings are applicable beyond South Africa's borders.
-
-\subsection*{General guidelines for selecting and using land cover data}
-Our findings regarding bias and accuracy suggest several guidelines for selecting and using land cover data in global change research (Table 1), and suggested aggregation scales.
+These similarities, combined with the larger extents of these studies, also suggest that our findings may be applicable beyond South Africa's borders.
+%
+%\subsection*{General guidelines for selecting and using land cover data}
+%Our findings regarding bias and accuracy suggest several guidelines for selecting and using land cover data in global change research (Table 1), and suggested aggregation scales.
 These guidelines vary according to the intended application, what error property the users wants to minimize or maximize, and what tradeoff the user is willing to make between error and resolution.
@@
 1, 3, 5) allow different thresholds to be  evaluated.
-
-The first general guideline is that land cover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions.
+%
+%The first general guideline is that land cover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions.
 Researchers might need to use these datasets for a variety of reasons, for instance, because they provide consistent, multiple land cover classes at a global scale, or allow changes to be tracked across large scales over time \citep{luoto_predicting_2004}.
@@
 1, 3, 5).
-
-The characteristics of cover at the landscape scale may also require greater levels of aggregation to achieve sufficient accuracy.
+%
+%The characteristics of cover at the landscape scale may also require greater levels of aggregation to achieve sufficient accuracy.
 We found that accuracy is lowest where mixing of cover types is greatest (Fig.
@@
 In South Africa's farming regions, this can exceed 1000 km$^2$.
-
-Maps derived from higher resolution sensors, such as the SA-LC dataset, do not have this mixture problem because the pixels are fine enough to be assigned to a single cover type.
+%
+%Maps derived from higher resolution sensors, such as the SA-LC dataset, do not have this mixture problem because the pixels are fine enough to be assigned to a single cover type.
 This characteristic leads to higher accuracy, which is what we observed here: the SA-LC dataset needed the least aggregation--just 5-10 km--to achieve $<$10\% MAE in most of the example applications we tested (note, however, another guideline in this result: even with high resolution maps some aggregation is needed to achieve higher accuracy).
-
-%needed This observation leads to a second rule of thumb: users should select datasets from higher resolution sensors where possible (assuming they were carefully developed), as these needed much less aggregation to return accurate estimates for most applications we tested here.
+%
+%%needed This observation leads to a second rule of thumb: users should select datasets from higher resolution sensors where possible (assuming they were carefully developed), as these needed much less aggregation to return accurate estimates for most applications we tested here.
 For $<$10\% MAE, just 5-10 km of aggregation were needed (however, this is still much coarser than the 30 m resolution of the sensor itself).
-
-Unfortunately, Landsat-scaled maps are typically developed for specific countries, using varying methods, and can be hard to obtain (an exception to this is the newly released GLC30, which has a reported classification accuracy of 80\% \citep{chen_global_2015}).
+%
+%Unfortunately, Landsat-scaled maps are typically developed for specific countries, using varying methods, and can be hard to obtain (an exception to this is the newly released GLC30, which has a reported classification accuracy of 80\% \citep{chen_global_2015}).
 Our results also suggest that, despite their higher accuracy, they may not necessarily be the least biased maps.
@@
 This result, together with GeoWiki's unbiased cropland estimates (Fig 1.B), indicates the value of fusing inventory data with remote sensing to reduce map bias.
-
-%The consensus method mirrors the ensemble averaging used by other fields (e.g.
+%
+%%The consensus method mirrors the ensemble averaging used by other fields (e.g.
 crop \citep{asseng_uncertainty_2013}, climate \citep{giorgi_calculation_2002}, and ecological modeling \citep{araujo_ensemble_2007}) to increase prediction confidence.
-
-Of course, for the constraint to be effective, the inventory data themselves must be faithful to reality, but in many places, particularly Africa, agricultural statistics are suspect \citep{carletto_emperor_2013, fao_action_2013}, which is a vulnerability of the fusion approach \citep{see_improved_2015}.
+%
+%Of course, for the constraint to be effective, the inventory data themselves must be faithful to reality, but in many places, particularly Africa, agricultural statistics are suspect \citep{carletto_emperor_2013, fao_action_2013}, which is a vulnerability of the fusion approach \citep{see_improved_2015}.
 Beyond this limitation, this method does not improve map accuracy, which is evident in GeoWiki's relatively high MAE for cropland percentage at 1 km (23\%, Fig.
@@
 
-\begin{table*}[!h]
-  \begin{threeparttable}
-    \caption{Guidelines for selecting and using different classes of land cover product (based on the characteristics of the satellite sensor(s) and map-making methods), according to the general purpose of the application, the error minimization objective, and the range of aggregation needed to achieve the error objective (here we use a maximum acceptable error magnitude of 10\%).
+%\begin{table*}[!h]
+%  \begin{threeparttable}
+%    \caption{Guidelines for selecting and using different classes of land cover product (based on the characteristics of the satellite sensor(s) and map-making methods), according to the general purpose of the application, the error minimization objective, and the range of aggregation needed to achieve the error objective (here we use a maximum acceptable error magnitude of 10\%).
 The preferred land cover product class for each application is indicated in boldface.
 }
-    \begin{center}
-      \begin{tabular*}{\textwidth}{m{3.5cm}m{3cm}m{5.5cm}m{2.25cm}}
-       \hline
-Application                     & Objective & Source of land cover Map & Minimum aggregation* \\
-\Xhline{1pt}
-High frequency            & Maximize accuracy   & Coarse resolution sensor (e.g. MODIS) & $>$100 km\\
-change detection                                                               &                                  & \textbf{High resolution sensor} (e.g. Landsat)   &  1-25 km\\
-\hline
-Mean values         & Minimize bias            & Coarse resolution sensor  & 25-50 km \\
-(e.g.
+%    \begin{center}
+%      \begin{tabular*}{\textwidth}{m{3.5cm}m{3cm}m{5.5cm}m{2.25cm}}
+%       \hline
+%Application                     & Objective & Source of land cover Map & Minimum aggregation* \\
+%\Xhline{1pt}
+%High frequency            & Maximize accuracy   & Coarse resolution sensor (e.g. MODIS) & $>$100 km\\
+%change detection                                                               &                                  & \textbf{High resolution sensor} (e.g. Landsat)   &  1-25 km\\
+%\hline
+%Mean values         & Minimize bias            & Coarse resolution sensor  & 25-50 km \\
+%(e.g.
 carbon density)                                                              &                                   &  \multicolumn{1}{r}{\emph{if mixed class dominates}}                                    & $>$100 km \\
-                                                              &                                   & \textbf{Fusion product: statistical constraint} & 1-5 km\\
-                                                              &                                   & High resolution sensor & 1-10 km\\
-\hline
-Spatial comparisons & Maximize accuracy  & Coarse resolution sensor & 25-100 km\\
- (e.g.
+%                                                              &                                   & \textbf{Fusion product: statistical constraint} & 1-5 km\\
+%                                                              &                                   & High resolution sensor & 1-10 km\\
+%\hline
+%Spatial comparisons & Maximize accuracy  & Coarse resolution sensor & 25-100 km\\
+% (e.g.
 crop production)                                                               &                                   &  \multicolumn{1}{r}{\emph{if mixed class dominates}}                                    & $>$100 km \\
-                                                             &                                   & Fusion product: map consensus** & 10-50 km\\
-                                                              &                                   & \textbf{High resolution sensor} & 1-25 km\\
-\Xhline{1pt}                         
-       \end{tabular*}
-       \begin{tablenotes}
-        \small
-        \item *Aggregation ranges account for variation in the results from the different map analyses.
-        \item **This assumes that consensus-based fusion products \citep[e.g.][]{fritz_cropland_2011,fritz_mapping_2015,tuanmu_global_2014} incorporate maps from coarse resolution sensors, making them less accurate at finer scales than products created using high resolution sensors.
-       \end{tablenotes}
-     \end{center}
-     \label{default}
-  \end{threeparttable}
-\end{table*}
+%                                                             &                                   & Fusion product: map consensus** & 10-50 km\\
+%                                                              &                                   & \textbf{High resolution sensor} & 1-25 km\\
+%\Xhline{1pt}                         
+%       \end{tabular*}
+%       \begin{tablenotes}
+%        \small
+%        \item *Aggregation ranges account for variation in the results from the different map analyses.
+%        \item **This assumes that consensus-based fusion products \citep[e.g.][]{fritz_cropland_2011,fritz_mapping_2015,tuanmu_global_2014} incorporate maps from coarse resolution sensors, making them less accurate at finer scales than products created using high resolution sensors.
+%       \end{tablenotes}
+%     \end{center}
+%     \label{default}
+%  \end{threeparttable}
+%\end{table*}
+%
 
-These last few points suggest two final guidelines.
+%These last few points suggest two final guidelines.
 For high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \citep{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share.
@@
 annual), large area land cover maps will depend on MODIS or similar sensors, and therefore much coarser aggregations.
-
-%GeoWiki was the most accurate among the large scale land cover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
-
-\subsection*{Implications for global change analyses}
-Besides the preceding land cover data selection and usage guidelines, our findings also have broader implications for global change analyses and associated policy decisions.
+%
+%%GeoWiki was the most accurate among the large scale land cover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
+%
+%\subsection*{Implications for global change analyses}
+%Besides the preceding land cover data selection and usage guidelines, our findings also have broader implications for global change analyses and associated policy decisions.
 The simpler, most direct uses of land cover data have substantial potential to skew our understanding of global change processes.
@@
 In contrast to these, an assessment of the congruence between carbon and biodiversity \citep{strassburg_global_2010} aggregated the carbon densities to $\sim$110 km, where accuracy is much higher.
-
-A related example can be seen in high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \citep[e.g.
+%
+%A related example can be seen in high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \citep[e.g.
 Figure 3 in][]{foley_solutions_2011}.
@@
 For this reason, it is advisable to aggregate presented maps until their accuracy reaches an acceptable standard (Table 1).
-
-What then, of the impacts of land cover map error on more complex analysis?
+%
+%What then, of the impacts of land cover map error on more complex analysis?
 Here the results appear to be mixed.
@@
 The carbon example illustrates how cover properties can mute or amplify map error.
-
-Despite the low values of the overall error metrics, the hotspots of ET errors (Fig.
+%
+%Despite the low values of the overall error metrics, the hotspots of ET errors (Fig.
 4) may have outsize significance because they occurred at the climatic margins of the major crop growing regions, where a substantial share of farms are irrigated.
@@
 This also suggests the potential for map error to magnify in a coupled land-atmosphere model \citep[which could explain the difference between this result and that of][]{ge_impacts_2007}, and thereby give misleading insight into the climatic effects of land use.
-
-Relative to the ET example, map errors had a much larger impact on the agent-based model's results.
+%
+%Relative to the ET example, map errors had a much larger impact on the agent-based model's results.
 For the model's most important metric of food security--household-level crop production--the result was fairly straightforward for a class of models where the interactions of multiple agents often produce analytically intractable outcomes \citep{janssen_empirically_2006}.
@@
 Less predictable was the finding that the spatial distribution of cover can impact model parameterization, with land being left unallocated even when needed, thus such models also require spatial, not just statistical, accuracy in their base land cover maps.
-
-The errors in the cropland data we tested were large, and it is unlikely that current ABM-based studies, mostly of small regions, are based on such inaccurate land cover data.
+%
+%The errors in the cropland data we tested were large, and it is unlikely that current ABM-based studies, mostly of small regions, are based on such inaccurate land cover data.
 However, as computing power increases, so, too, does the feasible application extent of ABMs, which will then become more reliant on larger-scaled, more erroneous, land cover data, with the attendant risk that insights into socio-ecological systems arising from these models could be highly misleading.
-
-Beyond these examples, there are other applications where land cover map errors could alter understanding and policy.
+%
+%Beyond these examples, there are other applications where land cover map errors could alter understanding and policy.
 One of these lies in assessments of land availability for agricultural expansion, which could be particularly erroneous if they use a ``residual approach'' \citep{lambin_estimating_2013}, where existing cropland and other occupied cover types are used to filter a map of potential suitability.
@@
 The validity of such analyses, which must be undertaken at the fine-scales where land use decisions are made (hectares rather than square kilometers), depends on having high quality map data, in order to have confidence that the identified tradeoffs are based on real biophysical properties, not simply map error.
-
-\subsection*{The way forward}
-Our analysis demonstrates that the land cover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on how accuracy or low bias.
+%
+%\subsection*{The way forward}
+%Our analysis demonstrates that the land cover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on how accuracy or low bias.
 We provide some basic guidelines (Table 1) for selecting and using land cover data that can help increase confidence in the conclusions drawn from land cover-based studies.
-
-Beyond these guidelines, we recommend updating several important, global-scale datasets using the latest generation land cover maps.
+%
+%Beyond these guidelines, we recommend updating several important, global-scale datasets using the latest generation land cover maps.
 For example, rebuilding the widely used cropland distribution and yield maps \citep{monfreda_farming_2008, ramankutty_farming_2008} using the GeoWiki map could substantially improve our knowledge about current global agricultural distribution and productivity.
@@
 Alternatively, the new high-resolution, Landsat-based GLC30 \citep{chen_global_2015}, aggregated and converted to fractional cover classes at 1 km--and fused with high quality national land cover maps, where available \citep[similar to][]{fritz_mapping_2015,fritz_cropland_2011}--may provide a preferable base map for rebuilding these datasets.
-
-Fusing human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power appears to offer the greatest promise for improving the quality of land cover maps.
+%
+%Fusing human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power appears to offer the greatest promise for improving the quality of land cover maps.
 The GeoWiki project and new high resolution global forest change maps \citep{hansen_high-resolution_2013,sexton_conservation_2015} are prominent examples of these newer methods.
@@
 This work was supported by funds from the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463 and BCS-1026776).
-We thank Tobias Kuemmerle for his helpful review of our study and suggestions for improvement.
 %\end{acknowledgments}

### 2017-01-27 Rmd/c4-cropland-2.Rmd
Added output to c4-cropland-2.Rmd for area estimates with each level of aggregation for each cropland dataset.

@@
 }))
+colnames(lckm2)[1] <- "gti"
+rownames(lckm2) <- lev2
 
@@
 stats2011
+```
+
+Cropland area by dataset and scale
+```{r, echo = FALSE}
+round(lckm2)
 ```

### 2017-03-20 Rmd/c10-location-impact.Rmd
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

@@
 # ...cropland
-crclass <- lapply(cropl, function(x) {  # x <- cyld_agg$f1[[1]]
+crclass <- lapply(cropl, function(x) {  # x <- cropl$sa30
   r <- x * (x > 0)
@@
 # ...carbon
-cdat <- c("gti" = gti_c$f1, lc_c$f1)  # combine C 1 km sets
+cdat <- c(gti_c$f1, lc_c$f1)
 cdatf <- lapply(cdat, function(x) {  # mask out non-cropland areas and take mu C
@@
 })
+names(cclass) <- names(pclass)
 
@@
 
+# average km off
+mudvec <- unlist(mudists)
+median(mudvec[!like(names(mudvec), "sa30")])
+# mean(mudvec)
+
 # plot(stack(yclass))

### 2017-03-20 Rmd/c4-cropland-2.Rmd
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

@@
 ```{r, eval = FALSE}
-# paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
-# gti <- lapply(paths, raster)
-# names(gti) <- c("g2007", "g2011")
-# gti <- lapply(gti, function(x) raster::mask(x, namask))  # apply namask
-# snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
+fact <- c(5, 10, 25, 50, 100)
+paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
+gti <- lapply(paths, raster)
+names(gti) <- c("g2007", "g2011")
+gti <- lapply(gti, function(x) raster::mask(x, namask))  # apply namask
+snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
+
+gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters
+gti_agg11 <- lapply(gti_agg, function(x) x$g2011)
 
@@
 lckm2 <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
-  gtia <- cellStats(gti_agg[[x]]$g2011 / 100 * awgts[[x]], sum)
+  gtia <- cellStats(gti_agg11[[x]] / 100 * awgts[[x]], sum)
   lca <- sapply(snms, function(y) {
@@
 colnames(lckm2)[1] <- "gti"
-rownames(lckm2) <- lev2
+rownames(lckm2) <- lev
+
+bias_statsw
+
+# mean (stays about the same)
+lckmmu <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
+  gtia <- cellStats(gti_agg11[[x]] / 100, mean)
+  lca <- sapply(snms, function(y) {
+    cellStats(lc_agg[[x]][[y]] / 100, mean)
+  })
+  o <- c(gtia, lca)
+}))
+colnames(lckmmu)[1] <- "gti"
+rownames(lckmmu) <- lev
+
+# variance declines with scale, as expected
+lckmvar <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
+  gtia <- cellStats(gti_agg11[[x]] / 100, var)
+  lca <- sapply(snms, function(y) {
+    cellStats(lc_agg[[x]][[y]] / 100, var)
+  })
+  o <- c(gtia, lca)
+}))
+colnames(lckmvar)[1] <- "gti"
+rownames(lckmvar) <- lev
 
@@
 
-totkm2 <- cellStats(!is.na(gti_agg$f1$g2011) , sum)
+totkm2 <- cellStats(!is.na(gti_agg$f1$g2011), sum)
 save(stats2007, stats2011, totkm2, refkm2, 
@@
 
-gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters
-gti_agg11 <- lapply(gti_agg, function(x) x$g2011)
+# gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters
+# gti_agg11 <- lapply(gti_agg, function(x) x$g2011)
 
@@
 bacc2011 <- rbind(a, b)
+save(bacc2011, file = fp(p_data, "cropland-werr-2011.rda"))
 

### 2017-03-20 Rmd/c7-et.Rmd
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

@@
 etba <- lapply(etba, function(x) {
-  b <- mask(x, mask25k, maskvalue = 0)
+  b <- raster::mask(x, mask25k, maskvalue = 0)
   names(b) <- lnames
@@
       caption.placement = "top", include.rownames = FALSE)
+
+et_err <- mutab
+save(et_err, file = fp(p_data, "et-err-bias.rda"))
 ```

### 2017-03-20 Rmd/c8-yield-production.Rmd
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

@@
 provr <- raster("external/ext_data/provinces.tif")
-provr <- mask(provr, namask)
+provr <- raster::mask(provr, namask)
 gtis <- stack(list("prov" = provr, "f" = gti))
@@
 # cftest - cf[, 2:4, with = FALSE]  # close, but diff caused by few NAs in provr 
-
 # data.table approach is much more efficient
-
 ```
@@
 dnms <- ls()[grep("[y_|p_]diff*.*d", ls())]
-stnms <- ls()[grep("stats_", ls())]
+stnms <- ls()[grep("^stats_", ls())]
 disaggl <- lapply(dnms, function(x) get(x))
 statsl <- lapply(stnms, function(x) get(x))
-rng <- lapply(statsl, function(x) {
+rng <- lapply(statsl, function(x) { # x <- statsl[[1]]
   ilims <- c(ceiling(min(sapply(x, function(y) y[3, ]))), 
@@
 evst <- "wma" # "wm"
-for(i in c("f1", "f25", "f10")) {
+for(i in c("f1", "f25", "f100")) {
   for(j in c("sa30", "modmu", "geow")) { 
@@
 save(list = ls(), file = full_path(p_data, "yield-bias.rda"))
+save(out_tab, file = fp(p_data, "yield-prod-accbias.rda"))
 save(cyld_agg, cpagg1, file = full_path(p_data, "yieldprod-1km.rda"))

### 2017-03-20 paper/SI.tex
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

@@
 This document provides supplementary information on methods and results from the different components of the analysis described in the main text.
-The code used to conduct the analysis is presented after sections 1-X.
+The code used to conduct the analysis is presented after section 7.
 
@@
 4 frequencies of cells per five percentile bins of cropland density).
-Finally, we calculated a density-independent metrics, which is similar to the density weighted mean calculated for our main analysis, which was calculated by averaging the mean bias and MAE values within different levels of cropland cover (divided in increments of 5\%).
+Finally, we calculated a density-independent metric, which is similar to the density weighted mean calculated for our main analysis, which was calculated by averaging the mean bias and MAE values within different levels of cropland cover (divided in increments of 5\%).
 The results of these three alternate metrics are shown in Figure 3.

### 2017-03-20 paper/main.tex
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

@@
 
-\title{How Cropland Data Errors Impact Our Understanding of Socioeconomic and Environment Change}
-\author[1,2]{Lyndon Estes}
-\author[3]{Peng Chen}
-\author[1]{Stephanie Debats}
-\author[3]{Tom Evans}
-\author[4]{Stefanus Ferreira}
-\author[5]{Tobias Kuemmerle}
-\author[6]{Gabrielle Ragazzo}
-\author[1]{Justin Sheffield}
-\author[6]{Adam Wolf}
-\author[1]{Eric Wood}
-\author[7]{Kelly Caylor}
+\title{How Landcover Errors Impact Our Understanding of Socioeconomic and Environment Change}
+\author[1,2,3]{Lyndon Estes}
+\author[4]{Peng Chen}
+\author[2]{Stephanie Debats}
+\author[4]{Tom Evans}
+\author[5]{Stefanus Ferreira}
+\author[6]{Tobias Kuemmerle}
+\author[2]{Gabrielle Ragazzo}
+\author[2,7]{Justin Sheffield}
+\author[8]{Adam Wolf}
+\author[2]{Eric Wood}
+\author[2,9]{Kelly Caylor}
 
-\affil[2]{Woodrow Wilson School, Princeton University, Princeton, NJ USA}
-\affil[1]{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}
-\affil[3]{Indiana University, Bloomington, IN USA}
-\affil[4]{GeoTerraImage, Pretoria, RSA}
-\affil[5]{Geography Department, Humboldt University, 10099 Berlin, Germany}
-\affil[6]{Arable Labs, Princeton, NJ USA}
-\affil[7]{Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA USA}
+\affil[1]{Woodrow Wilson School, Princeton University, Princeton, NJ USA}
+\affil[2]{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}
+\affil[3]{Graduate School of Geography, Clark University, Worcester, MA USA}
+\affil[4]{Indiana University, Bloomington, IN USA}
+\affil[5]{GeoTerraImage, Pretoria, RSA}
+\affil[6]{Geography Department, Humboldt University, 10099 Berlin, Germany}
+\affil[7]{Geography and Environment, University of Southampton, Southampton, United Kingdom}
+\affil[8]{Arable Labs, Princeton, NJ USA}
+\affil[9]{Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA USA}
 
@@
 \begin{abstract}
-{Research into global environmental change is often built upon land cover data, particularly croplands maps, which can have substantial errors.
-How these errors impact our understanding of environmental change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions.
+{Land cover maps increasingly underlie research into socioeconomic and environmental processes, including global change.
+It is known that map errors impact our understanding of these processes, but quantifying these impacts is difficult because many areas lack adequate reference data.
+We used a high accuracy map of South African cropland to assess 1) the magnitude of error in several current generation land cover maps, and 2) how these errors propagate in downstream studies.
+We first quantified pixel-wise errors in the cropland classes of four widely used land cover maps at resolutions ranging from 1-100 km, then calculated errors in several representative ``downstream'' (map-based) analyses, including estimates of vegetative carbon stocks and crop production, simulated evapotranspiration, and agent-based assessment of household food security.
+We also assessed maps' spatial accuracy based on how precisely they could be used to locate specific landscape features.
+We found that cropland maps can have substantial biases and poor accuracy at all resolutions (e.g.
+at 1 km resolution, up to $\sim$45\% underestimates of cropland (bias) and nearly 50\% mean absolute error (accuracy); at 100 km, up to 15\% underestimates and nearly 20\% MAE).
+National-scale maps derived from higher resolution imagery were most accurate, followed by multi-map fusion products.
+Constraining mapped values to match survey statistics proved effective at minimizing bias (provided the statistics are accurate).
+Downstream errors could be substantially amplified or muted, depending on the values ascribed to cropland-adjacent covers (e.g.
+with forest as adjacent cover, carbon map error was 200-500\% greater than in the foundational cropland maps, but $\sim$40\% less for sparse cover types).
+The average locational error was 6 km (600\%).
+These findings provide deeper insight into the causes and consequences of land cover map error, including their potential broader regional and policy consequences, and suggest several recommendations for land cover map users.
+%
+%
+%These maps can have substantial error that impact , which may impact our understanding of environmental change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions.
 We used a high quality, national-scale cropland map to assess bias and accuracy within the cropland classes of current generation land cover datasets and four examples of ``downstream'' (land cover-dependent) analyses, two related to physical processes (estimates of vegetative carbon stocks and evapotranspiration) and two to socio-economic processes (gridded crop production estimates and agent-based simulation of household food security), at resolutions ranging from 1-100 km.
@@
 \section*{Introduction}
+\vspace{-0.3 cm}
 The functioning of the Earth System is fundamentally connected to the characteristics of land cover \citep{lambin_modelling_1997}.
 Our increasing modification of the Earth's surface \citep{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through land cover.
-To fully understand these processes, an accurate understanding of the nature and distribution of land cover is essential.
+To fully understand these processes, including the forces driving climate change \citep{verburg_challenges_2011}, it is essential to have an accurate understanding of the nature and distribution of land cover.
 
-work in \citep{verburg_challenges_2011}, who says how fundamental landcover is to understanding global change.
+This importance is understood by a growing number of social, economic, and natural scientists, who are using land cover data to advance understanding of food security \citep{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \citep{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \citep{newbold_global_2015, luoto_predicting_2004}, demographic shifts \citep{linard_assessing_2010}, and other important facets of Earth System processes.
 
-This importance is understood by a growing number of social, economic, and natural scientists, who are using land cover data to advance understanding of food security \citep{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \citep{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \citep{newbold_global_2015, luoto_predicting_2004}, demographic shifts \citep{linard_assessing_2010}, and other important facets of Earth System processes (we hereafter collectively refer to this sort of work as Earth System Science (ESS), of which global change research is a subset).
+% (we hereafter collectively refer to this sort of work as Earth System Science (ESS), of which global change research is a subset).
 
-The value of the insights resulting from ESS studies depends upon the veracity of their underlying land cover data, much as a house requires a solid foundation in order to remain standing.
+The value of the insights resulting from such studies depends upon the veracity of their underlying land cover data \citep{verburg_challenges_2011}, much as a house requires a solid foundation in order to remain standing.
 Unfortunately, the evidence to date indicates that this house has shaky foundations.
@@
 First, in many regions the spatial arrangement of cover types is smaller than the sensor resolution \citep[e.g.
-smallholder's farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015,fritz_identifying_2008}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
+smallholders' farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015,fritz_identifying_2008}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
 Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties \citep[e.g.
@@
 Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g.
-3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \citep{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013,debats_generalized_2016}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the land cover products that have become commonplace in global change research--and which are often used to make policy decisions \citep{searchinger_high_2015}.
+3DRobotics, DJIA} to new high resolution satellite arrays \citep{drusch_sentinel-2:_2012,hand_startup_2015} and better mapping methods \citep{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013,debats_generalized_2016}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the land cover products that have become commonplace in global change research--and which are often used to make policy decisions \citep{searchinger_high_2015}.
 
 In this study, we take advantage of this recent growth in data to address the call to more thoroughly assess errors in land cover maps \citep{kuemmerle_challenges_2013, olofsson_good_2014,olofsson_global_2012}, and further examine how these errors might impact our understanding of socioeconomic and environmental conditions.
-Using a unique, high-quality, spatially extensive map of South African croplands, which was created by expert mappers delineating individual fields visible within high resolution imagery, we conduct spatially comprehensive, bottom-up analyses to answer the following two questions: 1) What is the extent of error in several widely used land cover products?; 2) How do these errors propagate through downstream biophysical and socioeconomic studies?
+Using a unique, high-quality, map of South African croplands, which was created by expert mappers delineating individual fields visible within high resolution imagery, we conduct spatially comprehensive, bottom-up analyses to answer the following two questions: 1) What is the extent of error in several widely used land cover products?; 2) How do these errors propagate through downstream biophysical and socioeconomic studies?
 
@@
 90\% user's accuracy][]{sweeney_mapping_2015}.
-The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution MODIS land cover data (for 2011, the final year of the reference interval), which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
+The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution MODIS land cover \citep{friedl_modis_2010} data (for 2011, the final year of the reference interval), which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
 The fourth dataset was the 1 km GeoWiki cropland map for Africa \citep{fritz_mapping_2015}, which fuses the cropland classes of multiple landcover products (including the three assessed here), with the resulting cropland percentages adjusted until their totals sum to match cropland areas reported in national agricultural statistics.
@@
 We then followed \citetapos{monfreda_farming_2008} procedure for disaggregating maize \citep[South Africa's largest crop;][]{estes_comparing_2013} planted area and yields onto the reference and adjusted test maps.
-The necessary statistics were obtained from magisterial district-level agricultural censuses conducted for the year 2007 \citep{statistics_south_africa_commercial_2007}.
+The necessary statistics were obtained from magisterial district-level agricultural censuses conducted for the year 2007 \citep{statistics_south_africa_building_2007}.
 
@@
 areas of high agricultural potential and low ecological cost;][]{estes_reconciling_2016}, as opposed to general quantities.
-It is therefore important to also assess how map errors can impact the ability to accurately locate specific features of interest.
+It is therefore important to also assess how map errors can impact the spatial accuracy of maps, or the ability to accurately locate specific features of interest.
 To evaluate this, we calculated the mean euclidean distance (in km) between pixels representing a specific feature within the test maps relative to their nearest neighboring pixels representing the same feature in the reference map.
@@
 \vspace{-0.2 cm}
-Our reference cropland map indicated that crop fields covered nearly 10\% of the total study area in the 2009-2011 time period.
+Our 1 km reference cropland map indicated that crop fields covered 104,304 km$^2$ (nearly 10\%) of the total study area in the 2009-2011 time period, with corresponding cropland area estimates of 131,390, 82,358, 77,090, and 110,272 (GeoWiki) km$^2$ resulting from the SA-LC, GlobCover, MODIS, and GeoWiki maps, respectively.
+Cropland area estimates from both the reference and test maps were constant for all levels of aggregation.
+
 Subtracting each test map from the reference maps created pixel-wise residuals, where negative and positive values respectively represent overestimates and underestimates by the test map (Fig.
 1A).
-
 The most pronounced errors were in the MODIS and GlobCover maps, which showed large positive residuals in the center of the country where cropland is most concentrated (blue areas in Fig.
@@
 The magnitude of mean absolute error (MAE) was somewhat higher in all cases.
-The GeoWiki map, in contrast, was the least biased overall, showing just a ~7\% bias at 1 km and near 0 for all other scales of aggregation, although its accuracy (23\% MAE) was only half as good as SA-LC's at 1 km (11\% MAE), which despite its uniform overestimation bias (Fig.
+The GeoWiki map, in contrast, was the least biased overall, showing just a $\sim$7\% bias at 1 km and near 0 for all other scales of aggregation, although its accuracy (23\% MAE) was only half as good as SA-LC's at 1 km (11\% MAE), which despite its uniform overestimation bias (Fig.
 1A) was the most accurate map at aggregation scales $<10$km.
@@
 \vspace{-0.2 cm}
-The average distance between areas containing the highest cropland densities (upper decile) in the reference map and those delineated by the test maps ranged from 1.1 km for SA-LC to 18.2 km for GlobCover, with MODIS (10.1 km) and GeoWiki (2.8 km) had intermediate offsets (Fig.
+The average distance between areas containing the highest cropland densities (upper decile) in the reference map and those delineated by the test maps ranged from 1.1 km for SA-LC to 18.2 km for GlobCover, with MODIS (10.1 km) and GeoWiki (2.8 km) having intermediate displacements (Fig.
 7).
@@
 \vspace{-0.2 cm}
+%What We Did and how it is different
 
-\subsection*{Error and its propagation}
-\subsubsection*{Quantities}
-Aggregation can minimize error, but issue of MAUP, and might not be the application needed.
+The preceding analyses contributes to a substantial body of existing work devoted to understanding the nature and consequences of error within landcover maps \citep[e.g.][]{fritz_highlighting_2011,verburg_challenges_2011,olofsson_making_2013}.
+A number of prior studies have assessed land cover map errors, either through point-based accuracy assessments \citep[e.g.][]{olofsson_making_2013,frey_how_2007-1,foody_status_2002}, by evaluating between-map discrepancies \citep[e.g.][]{fritz_highlighting_2011,fritz_identifying_2008,fritz_comparison_2010}, or by comparing map-derived estimates to aggregated statistics \citep[e.g.][]{larsen_taken_2015,fritz_comparison_2010,yu_meta-discoveries_2014}.
+Several studies have used contiguous ground truth maps to assess land cover errors, but these covered relatively small regions \cite[$<$3000 km$^2$, or $<$0.03\% of the area covered here;][]{dendoncker_exploring_2008,schmit_limitations_2006}.
 
-\subsubsection*{Identifying locations}
-A few to 10s of km. Doesn't sounds like much, but if need to precisely identify a location, can make a big difference. 
-Sensitivity is generally within the same range for each application. 
-ABM results indicate locational errors also.
+Previous work has also examined how map errors influence estimates of environmental and socioeconomic variables, including simulated rainfall \citep{ge_impacts_2007}, carbon stocks and emissions \citep{goetz_mapping_2009, quaife_impact_2008, olofsson_making_2013,jain_co2_2013}, nitrogen fluxes \citep{jain_co2_2013,nol_effect_2008}, human population density \citep{linard_assessing_2010}, species distributions \citep{tuanmu_global_2014}, and landscape patterns \citep{langford_map_2006}.
+The majority of these used either point validation, map inter-comparison, or a combination of both to directly assess errors in the measurement of interest \citep{goetz_mapping_2009, olofsson_making_2013, jain_co2_2013, linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}.
+A smaller number specifically examined how either simulated map errors \citep{ge_impacts_2007,langford_map_2006} or differences relative to small area ground truth maps \citep[$<$1000 m;][]{nol_effect_2008} propagate into the downstream estimates.
 
-\subsection*{Sources of error}
-Mixed classes - MODIS versus GlobCover. 
-Pixel size
-Mosaic classes - cropland density/error relationship
-Within-class variability (cropland particularly pronounced)
+Our study builds on these previous efforts by providing a large-area, spatially continuous quantification of cropland classification errors, and examines how these actual errors influence several common downstream applications.
+The spatially comprehensive nature of these analyses provides deeper insight into the causes and consequences of error than would otherwise be obtained from either a point-based accuracy assessment or through the imposition of simulated error.
+By assessing errors within a continuous estimate of cropland, we were also able to examine how pixel-wise errors change with scale, while minimizing the confounding effects of aggregating a categorical variable \citep[][and see discussion in subsequent Recommendations section]{moody_influence_1995, marceau_remote_1999}.
+These analyses were enabled by a high accuracy reference map that likely provides the truest measure of cropland area and distribution for this region.
+Although this reference map is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission, SI), while temporal mismatches between the reference and test maps may account for some of the error we identified, our assessment (SI) suggests that such discrepancies do not appreciably impact our findings.
 
-caveats: mention among MAUP concerns that relying on percentages helps avoid some of the issues of MAUP that relate to rasterization error and trying to aggregate classes (cites for this).
-Of course doing this loses one of the main sources of geographic information (configuration), but this is one of the many aspects that leads to the MAUP problem in any case.
+\vspace{-0.3 cm} 
+\subsection*{Sources of error in cropland maps}
+\vspace{-0.2 cm}
+Our findings showed that the most accurate cropland estimates, across all spatial scales, were produced by SA-LC followed by GeoWiki.
+In the reverse order, these two datasets were also the least biased, with GeoWiki having effectively zero bias at 5 km or higher of aggregation.
+GlobCover was highly inaccurate and biased at all scales, with MODIS estimates being nearly as inaccurate but substantially less biased.
+These error patterns are attributable to two factors: sensor resolution and methodology.
+With respect to the former, SA-LC's higher accuracy is largely because the 30 m (0.09 ha) resolution of Landsat imagery is smaller than the average area of South African crop fields.
+Previous work in agricultural remote sensing has shown that the sensor resolution should be finer than average field size to accurately estimate both the area and location of croplands\citep{ozdogan_resolution_2006,pax-lenney_effect_1997}.
+When pixel sizes are small relative to the objects being mapped, the number of ``mixed pixels'' (those where the spectral signature is defined by more than one cover) is relatively small, and their number naturally increases as sensor resolution decreases \citep{ozdogan_resolution_2006}.
 
-Reasons for not evaluating GLC30 (lower accuracy)
+Mixed pixels introduce another potential source of error related to methodology, which stems from the need to define thresholds for allocating pixels to different cover types \citep{ozdogan_resolution_2006}.
+This error is evident in the MODIS and GlobCover results, both of which cope with the mixing problem by assigning sub-pixel proportions of cropland to mosaic classes \citep{friedl_modis_2010,arino_global_2012}.
+These classes place upper thresholds on cropland, causing underestimation error where actual cropland proportions are higher  (Fig.
+2B).
+Over South African croplands, the GlobCover map was dominated by its mosaic pixel classes, leading to substantial underestimation bias that persisted even with aggregation.
+MODIS, on the other hand, classified more areas as pure cropland, and thus had lower underestimation bias.
 
-\citep{verburg_challenges_2011} Relationship between class accuracy and heterogeneity of the landscape.
+The modeled relationships between map accuracy and cropland density (Fig.
+2) further demonstrate how pixel mixing and class definition influence error.
+For MODIS, SA-LC, and GeoWiki, error was highest where pixels were evenly divided between cropland and other cover types, reflecting earlier work showing that classification accuracy is lowest when cover types are most mixed \citep{verburg_challenges_2011,gross_monitoring_2013}.
+On the other hand, GlobCover was least accurate over pixels with 100\% cropland, an error pattern imposed by the proportions defining cropland within GlobCover's dominant mosaic classes; these set an upper bound that necessarily led to underestimation error over dense croplands.
+Previous work by \citet{ozdogan_resolution_2006} shows that the threshold used for separating agricultural from non-agricultural classes is a significant source of error.
 
-cite frey and smith (2007) as downstream analyses.
+GeoWiki's cropland errors demonstrate three other ways in which methodology affects error pattern.
+The first is that constraining remotely sensed cropland proportions to match census-statistics \citep{fritz_mapping_2015,fritz_highlighting_2011} can be an effective way to reduce bias.
+The second is that this constraint cannot correct the errors of commission and omission within the individual landcover datasets that were merged to create the GeoWiki map \citep{fritz_mapping_2015}, which is why its accuracy is relatively low at 1 km resolution (Fig.
+1B).
+However, GeoWiki's accuracy was substantially higher than MODIS and GlobCover's (Fig.
+1B, 3, \& 5), which reveals the third point, namely that the landcover fusion process used to create GeoWiki does help minimize such error.
+Fusion may be particularly effective for minimizing underestimation errors caused by mixed classes (as with GlobCover) in areas of substantial sub-pixel heterogeneity \citep{fritz_mapping_2015,tuanmu_global_2014}.
+This fusion approach mirrors the ensemble methods used by various modeling sub-disciplines (e.g.
+crop \citep{asseng_uncertainty_2013}, climate \citep{giorgi_calculation_2002}, and ecological modeling \citep{araujo_ensemble_2007}) to increase prediction confidence.
 
-\subsection*{Broader implications}
+\vspace{-0.3 cm}
+\subsection*{Error propagation in downstream products}
+\vspace{-0.2 cm}
+Cropland map errors were either amplified or muted within the various downstream applications we assessed.
+Both tendencies were evident in the Tier-1 carbon maps, the simplest of the downstream methods.
+Using the ratio between each carbon map's accuracy score (MAE) and that of its foundational cropland map to calculate error propagation ($>$1 means error was exacerbated, $<$1 means error muted), we see that errors in the 1 km carbon map errors were 200 (SA-LC) to 500\% (GlobCover) larger than cropland map errors when forest was the adjacent cover type, but $\sim$40\% lower for the sparse cover type.
+Aggregation helped to reduce error magnitude, but carbon maps nevertheless had 30-50\% more error than cropland maps at 100 km resolution when forest or shrubland were sharing the pixel (the error ratio associated with the sparse vegetation carbon class remained relatively constant with aggregation).
 
-Broader conclusions are necessarily limited, because of single country involved in this analysis. 
-Parochial study. But suggests important findings for rest of Africa.
+The error propagation patterns in the carbon map were therefore determined by the differences between the carbon density of cropland and that of the adjacent cover types; forest and shrubland have higher carbon densities than cropland, whereas the sparse cover type is lower.
+A similarity in the values assigned to the cover types adjacent to cropland may also explain the low error rates in the evapotranspiration estimates produced by the VIC model, which were all 90-95\% lower than those in the input cropland maps.
+This dampening of error contrasts with results from elsewhere showing that map errors can substantially alter rainfall simulations \citep{ge_impacts_2007}.
+In the case of the ET simulations, VIC's map-related variables (e.g.
+LAI curves, effective rooting depth) were relatively similar between cropland and the adjacent landcover types, which we did not alter beyond adjusting their percentages to accommodate altered cropland proportions.
 
-However, support from previous studies.
-Properties revealed here in this study are relevant.  Mosaic classes.  Background vegetation similar to cropland (savanna, grassland). Should expect even high error.
+The disaggregated yield and crop production maps we created also showed both error amplification and muting, which in this case depended on the particular analysis.
+Errors within the yields maps were uniformly lower (50-90\% less at 1 km) than those in the input cropland maps, whereas errors in the production maps were 70-100\% higher at 1 km, and actually were exacerbated at intermediate levels of aggregation (10-25 km) for GlobCover, MODIS, and GeoWiki (170-290\%); this latter tendency mostly caused by accuracy in the production maps improving more slowly with aggregation (Fig.
+5) than in the original cropland maps (Fig.
+1B).
+These contrasting results reflect differences within disaggregation methods \citep{monfreda_farming_2008}.
+First, the yield methodology is a simple form of disaggregation that paints district-level yields onto pixels with $>$0\% cropland within each district without attempting to map within-district yield variability.
+This simplicity means that it is only sensitive to errors in classifying cropland presence/absence, but not to errors in cropland proportions.
+Production, on the other hand, is calculated from disaggregated harvested area maps, which are created by proportionally allocating district-level crop harvested area onto cropland fractions, making them highly sensitive errors in cropland proportion.
+Of particular note is that harvested area maps are subject not only to this statistical constraint (that pixel-wise harvested area fractions sum to district totals), but cropland fractions are also adjusted to match district-level cropland area estimates \citep[see Methods; ][]{ramankutty_farming_2008}.
+This suggests that statistical constraints are therefore not necessarily helpful in preventing pixel-level error propagation.
 
-How to deal with it: Limited
+Sensitivity to cropland area also was evident in the agent-based model's results.
+The ABM's most important metric of food security--household-level crop production--was only impacted by underestimates of cropland area, which lowered the models' estimates of average household production (in turn overstating the degree of food \emph{in}security) because individual households were allocated insufficient cropland.
+Cropland overestimates did not cause the opposite effect, because total households were constant and the allocation routine prevented cropland holdings from exceeding their assigned, census-derived hectarage.
+The less predictable result was that the model could leave cropland unallocated when there were substantial cropland underestimates (e.g.
+MODIS in Fig.
+6A), which was an error caused by the spatial arrangement of croplands in the district.
+MODIS cropland in those districts was confined to relatively small islands, meaning that they fell outside the allocation algorithm's search window, thus leaving those croplands unassigned to specific households.
 
-Synthetic assessments can help you determine how much impact error can make.
+This latter finding highlights the types of errors that caused by spatial inaccuracies in land cover maps, which was explicitly evaluated by the analysis of spatial distance between pixels containing upper decile of values in reference and test maps (Fig.
+7).
+The relative size of offsets tends to follow the patterns of accuracy seen in other assessments, with SA-LC producing the most spatially accurate results, and GlobCover the least, with the exception that GlobCover's 90\textsuperscript{th} percentile carbon density locations are closer to those in the reference map than those of MODIS or GeoWiki.
+Numerically, none of these nearest neighbor differences seem large, but they are akin to the root mean square error term used when measuring geometric distortion in a satellite image.
+Under this conception, the error in all but SA-LC's cropland and yield examples (Fig.
+7) is $>$3 pixels (or 300\%), with an overall average of nearly 6 pixels.
 
-Implications for different types of analyses???
-Previous studies as backup. Fritz et al, etc.
+%the result was fairly straightforward for a class of models where the interactions of multiple agents often produce analytically intractable outcomes \citep{janssen_empirically_2006}.
+In this case, the model
 
-\subsection*{Conclusion/way forward}
+%Yield disaggregation and production methods reinforce to a certain extent how statistical constrain can minimize bias at the pixel level.
+[check this against accuracy].
+But can still be off as a function of cropland errors.
+Highlight by production, and locational accuracy measures.
 
-Carefully evaluate a product for its sensitivity for a particular application \citep{fritz_highlighting_2011}.
-Here we show sensitivity of several important products for a variety of applications.
-Restricted geographically, so limited inference, but some general inferences may be drawn.
-Carefully crafted Landsat products for particular scale (higher density of validation points, local knowledge).
-
-Solutions: For fine features below sensor resolution, convert to quantitative measures such as SMA (tough for cropland though, but maybe with MESMA for different crop categories).
-For homogenous cover types with little within-class variability.
-
-Between-date classification error metrics (cite to Pontius, Verburg brings it up also)
-
-Care in aggregation methods worth emphasizing \citep{verburg_challenges_2011,dendoncker_exploring_2008,raj_analysing_2013,moody_influence_1995}, can lead to bias, which is influenced by spatial arrangement of cover types.
-
-points to make:
-
-Fusion of census data with RS: improve quantity estimates. But careful for accuracy of census data. Large
-
-Match dataset to application; merge complementary sources where available \citep{verburg_challenges_2011}
-
-Fritz and See (2008) about comparing maps with different themes\\
-
-\citep{nijland_optimizing_2009}.
-Spatial scale of pixel matters with respect to mapping accuracy.
-Average local variance is related to size of features being mapped.
-Averaging (smoothing) pixel value is also a better way of increasing accuracy (this is in estimating vegetation parameters).
-``Pixels are sufficiently small to separate the studied vegetation from other land cover units.
-This is especially important in complex scenes.''  Good point to make for ignoring larger pixel sensors.
-On the other hand, for certain applications, like estimating vegetation that might be better to have larger pixel sizes.
-
-They recommend experimental determination of optimal pixel size for all applications.
-Experimental determination of impact of error might also be appropriate.
-Careful consideration of pixel size emphasized, and shape of segments.
-
-This is for both discussion of MAUP issues and recommendations for way forward.
-
-MAUP in our case is minimized--it is at the level imposed by sensor resolution, but it is minimized by converting to percentage error estimates through aggregation [think I need to show this by showing area estimation with scale].
-Admittedly we do not deal as much with spatial aspects of MAUP, cite numerous studies that look at this.
-But our weighted assessment of accuracy does give some indication of spatial error--it is akin to the likely error experienced at any given point on the map.
-Weighting corrects for unequal spatial distribution of cover type.
-
-\citep{nelson_analysing_2001} Talks about ecological fallacy, scale, etc.  With spatial data, results should presented both in map form as well as tabulated results.
-
-Our contribution: 
-Motivation for our paper found in here. Our contribution is to more explicitly quantify, over a larger scale, the types of uncertainties in land cover data.
-
-Classification is scale-dependent.
-
-\citep{hay_comparison_2003}
-Up-scaling should be guided by the intrinsic scale of image objects that compose a scene.
-
-We show how error changes with scale, but that does not mean that coarser scales lead to more accurate findings.
-Coarser scale may be less appropriate, or provide misleading results \citep{nelson_analysing_2001}.
-
-Dimensions of the pixel should also be examined, although we didn't do that here.
-To a certain extent, this applies more to issues of category than percent coverage (hope for support on this).
-MAUP problem is also enforced by sensor resolution \citep{hay_comparison_2003,marceau_remote_1994}, so above that we are minimizing issue of aggregation by taking percent coverage.
-
-But zonation/pixel-size chosen should be user and purpose-designed.
-
-\citep{marceau_remote_1999}
-Areal units are modified by changing the sensor as well as by aggregating pixels.
-
-Prior studies
-%Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
-Some work has examined how map error influences climate simulations \citep{ge_impacts_2007}, agricultural land use patterns \citep{schmit_limitations_2006}, carbon flux measurements \citep{quaife_impact_2008}, human population estimates \citep{linard_assessing_2010}, and species distribution models \citep{tuanmu_global_2014}, but these either use simulated land cover errors \citep{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived land cover maps \citep{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}, or, in cases where spatially explicit ground validation maps were used for assessment, cover small and discontiguous extents \citep{schmit_limitations_2006}.
-Add in \cite{dendoncker_exploring_2008} as one related to Schmitt using comprehensive ground-truth data to assess scaling errors.  Part of mini-review.
-
-%This spatially comprehensive, bottom-up land cover map assessment provides new insight into the extent and causes of map error, and its consequences for understanding global change.
-The analysis was made possible by a high accuracy cropland map that likely provides the truest measure of cropped area and distribution currently available for this region.
-Although this reference map is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission, SI), while temporal mismatches between the reference and test maps may account for some of the error we identified, our assessment (SI) suggests that such discrepancies do not appreciably impact our findings.
-Our results are also supported by previous, continental-scale assessments that revealed substantial errors and inconsistencies between land cover maps \citep{fritz_comparison_2010,gross_monitoring_2013}.
-These similarities, combined with the larger extents of these studies, also suggest that our findings may be applicable beyond South Africa's borders.
-%
-%\subsection*{General guidelines for selecting and using land cover data}
-%Our findings regarding bias and accuracy suggest several guidelines for selecting and using land cover data in global change research (Table 1), and suggested aggregation scales.
-These guidelines vary according to the intended application, what error property the users wants to minimize or maximize, and what tradeoff the user is willing to make between error and resolution.
-For the purpose of clarity in these guidelines, we assume that a maximum error magnitude of 10\% is acceptable, but the results of our analyses (Fig.
-1, 3, 5) allow different thresholds to be  evaluated.
-%
-%The first general guideline is that land cover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions.
-Researchers might need to use these datasets for a variety of reasons, for instance, because they provide consistent, multiple land cover classes at a global scale, or allow changes to be tracked across large scales over time \citep{luoto_predicting_2004}.
-The appropriate scale of aggregation then depends on which error property the user wants to minimize.
-If the objective is to minimize systematic errors in order to calculate an unbiased mean (e.g.
-average carbon density), then at least 25-50 km of aggregation is needed to reduce map bias to within +/-10\% (Table 1, Fig.
-1B \& 3).
-If the user wants to maximize accuracy, which is important for comparing differences between locations on a single map, or between dates at fixed locations, then even coarser aggregations are needed to reduce mean absolute error below 10\% (Fig.
-1, 3, 5).
-%
-%The characteristics of cover at the landscape scale may also require greater levels of aggregation to achieve sufficient accuracy.
-We found that accuracy is lowest where mixing of cover types is greatest (Fig.
-2), a result echoed in an Africa-wide change detection study made with MODIS \citep{gross_monitoring_2013}.
-Outside of South Africa, where smallholder-driven agriculture predominates \citep{lambin_estimating_2013}, such mixed landscapes are more common, thus greater levels of aggregation may be needed over these areas.
-Coarser aggregations are also necessary if the cover type of interest belongs primarily to a mixed class (as with GlobCover).
-This leads to underestimation bias (Fig.
-2) and high inaccuracy, which will persist until the aggregated pixel exceeds the characteristic size of the geographical area within which the cover type dominates.
-In South Africa's farming regions, this can exceed 1000 km$^2$.
-%
-%Maps derived from higher resolution sensors, such as the SA-LC dataset, do not have this mixture problem because the pixels are fine enough to be assigned to a single cover type.
-This characteristic leads to higher accuracy, which is what we observed here: the SA-LC dataset needed the least aggregation--just 5-10 km--to achieve $<$10\% MAE in most of the example applications we tested (note, however, another guideline in this result: even with high resolution maps some aggregation is needed to achieve higher accuracy).
-%
-%%needed This observation leads to a second rule of thumb: users should select datasets from higher resolution sensors where possible (assuming they were carefully developed), as these needed much less aggregation to return accurate estimates for most applications we tested here.
-For $<$10\% MAE, just 5-10 km of aggregation were needed (however, this is still much coarser than the 30 m resolution of the sensor itself).
-%
-%Unfortunately, Landsat-scaled maps are typically developed for specific countries, using varying methods, and can be hard to obtain (an exception to this is the newly released GLC30, which has a reported classification accuracy of 80\% \citep{chen_global_2015}).
-Our results also suggest that, despite their higher accuracy, they may not necessarily be the least biased maps.
+%Our results also suggest that, despite their higher accuracy, they may not necessarily be the least biased maps.
 Except for a few results (notably the 1 km carbon and maize production estimates), GeoWiki-based maps had the lowest bias across most scales, which is the largely the result of calibrating the cropland percentages to match reported agricultural statistics \citep{fritz_cropland_2011,fritz_mapping_2015}.
@@
 This result, together with GeoWiki's unbiased cropland estimates (Fig 1.B), indicates the value of fusing inventory data with remote sensing to reduce map bias.
-%
-%%The consensus method mirrors the ensemble averaging used by other fields (e.g.
-crop \citep{asseng_uncertainty_2013}, climate \citep{giorgi_calculation_2002}, and ecological modeling \citep{araujo_ensemble_2007}) to increase prediction confidence.
-%
-%Of course, for the constraint to be effective, the inventory data themselves must be faithful to reality, but in many places, particularly Africa, agricultural statistics are suspect \citep{carletto_emperor_2013, fao_action_2013}, which is a vulnerability of the fusion approach \citep{see_improved_2015}.
-Beyond this limitation, this method does not improve map accuracy, which is evident in GeoWiki's relatively high MAE for cropland percentage at 1 km (23\%, Fig.
-1B), and by all maps' (except SA-LC's) highly inaccurate maize production estimates (Fig.
-5).
-The reason that accuracy is not improved is because the statistical constraint cannot correct the omission and commission errors within the land cover maps.
-To minimize these types of errors, different techniques are needed, such as creating a single consensus map from multiple land cover products \citep[e.g.][]{fritz_comparison_2010,tuanmu_global_2014}.
-This second kind of data fusion was used in the first stage of creating the GeoWiki map (before statistical adjustment), and it accounts for GeoWiki's higher accuracy scores relative to MODIS and GlobCover (Fig.
-1B, 3, \& 5).
-It may be particularly effective in minimizing underestimation errors that result when a mixed class type becomes dominant (as with GlobCover), or when dealing with substantial sub-pixel heterogeneity \citep{fritz_mapping_2015,tuanmu_global_2014}
 
-%\begin{table*}[!h]
-%  \begin{threeparttable}
-%    \caption{Guidelines for selecting and using different classes of land cover product (based on the characteristics of the satellite sensor(s) and map-making methods), according to the general purpose of the application, the error minimization objective, and the range of aggregation needed to achieve the error objective (here we use a maximum acceptable error magnitude of 10\%).
-The preferred land cover product class for each application is indicated in boldface.
-}
-%    \begin{center}
-%      \begin{tabular*}{\textwidth}{m{3.5cm}m{3cm}m{5.5cm}m{2.25cm}}
-%       \hline
-%Application                     & Objective & Source of land cover Map & Minimum aggregation* \\
-%\Xhline{1pt}
-%High frequency            & Maximize accuracy   & Coarse resolution sensor (e.g. MODIS) & $>$100 km\\
-%change detection                                                               &                                  & \textbf{High resolution sensor} (e.g. Landsat)   &  1-25 km\\
-%\hline
-%Mean values         & Minimize bias            & Coarse resolution sensor  & 25-50 km \\
-%(e.g.
-carbon density)                                                              &                                   &  \multicolumn{1}{r}{\emph{if mixed class dominates}}                                    & $>$100 km \\
-%                                                              &                                   & \textbf{Fusion product: statistical constraint} & 1-5 km\\
-%                                                              &                                   & High resolution sensor & 1-10 km\\
-%\hline
-%Spatial comparisons & Maximize accuracy  & Coarse resolution sensor & 25-100 km\\
-% (e.g.
-crop production)                                                               &                                   &  \multicolumn{1}{r}{\emph{if mixed class dominates}}                                    & $>$100 km \\
-%                                                             &                                   & Fusion product: map consensus** & 10-50 km\\
-%                                                              &                                   & \textbf{High resolution sensor} & 1-25 km\\
-%\Xhline{1pt}                         
-%       \end{tabular*}
-%       \begin{tablenotes}
-%        \small
-%        \item *Aggregation ranges account for variation in the results from the different map analyses.
-%        \item **This assumes that consensus-based fusion products \citep[e.g.][]{fritz_cropland_2011,fritz_mapping_2015,tuanmu_global_2014} incorporate maps from coarse resolution sensors, making them less accurate at finer scales than products created using high resolution sensors.
-%       \end{tablenotes}
-%     \end{center}
-%     \label{default}
-%  \end{threeparttable}
-%\end{table*}
-%
+%There are two reasons for this. First, map accuracy increased nearly two-fold by aggregating to 25 km, the resolution of VIC.
 
-%These last few points suggest two final guidelines.
-For high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \citep{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share.
-www.glcn.org} datasets for other cover types).
-But in either case some aggregation--1-25 km, depending on application and whether low bias or high accuracy is the objective--will still likely be necessary to achieve sufficiently low error.
-Additionally, these higher accuracy, lower bias products often represent just a single time period \citep[e.g.][]{fritz_mapping_2015,tuanmu_global_2014}, thus applications requiring higher frequency (e.g.
-annual), large area land cover maps will depend on MODIS or similar sensors, and therefore much coarser aggregations.
-%
-%%GeoWiki was the most accurate among the large scale land cover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.
-%
-%\subsection*{Implications for global change analyses}
-%Besides the preceding land cover data selection and usage guidelines, our findings also have broader implications for global change analyses and associated policy decisions.
-The simpler, most direct uses of land cover data have substantial potential to skew our understanding of global change processes.
-For example, the global carbon carbon density map of \citep{ruesch_new_2008}, of which we created facsimiles here, has been widely used, but our results suggest that any analyses of the spatial variability of carbon stocks, or of mean carbon densities, can be highly misleading if the map is not first sufficiently aggregated.
-For example, a study of deforestation risks in the Democratic Republic of the Congo estimated forest rents as a function of carbon densities based on this map.
-Although the map was aggregated to $\sim$50 km, we found that mean absolute errors can still reach 40\% at this scale (Fig.
-3).
-The same carbon map provided country-specific mean forest carbon densities for an evaluation of climate mitigation policies \citep{cattaneo_international_2010}, which our results suggest could be substantially underestimated (Fig.
-3).
-In contrast to these, an assessment of the congruence between carbon and biodiversity \citep{strassburg_global_2010} aggregated the carbon densities to $\sim$110 km, where accuracy is much higher.
-%
+\vspace{-0.3 cm}
+\subsection*{Broader implications}
+\vspace{-0.2 cm}
+Although our study focused on a single country and a subset of possible land cover-derived analyses, its findings should have broader geographic and practical relevance.
+In geographical terms, the key question is whether the error patterns revealed here will be reproduced outside of South Africa's borders?
+If one considers the rest of Sub-Saharan Africa (SSA), the answer is almost certainly yes.
+Previous work showing substantial disagreement between different cropland maps in other SSA countries \citep{fritz_comparison_2010} supports this contention.
+Another reason is that farming elsewhere in SSA is dominated by smallholders whose fields are substantially smaller than those in South Africa \citep{samberg_subnational_2016}, which increases mixed pixel classification error.
+Additionally, smallholders' fields often contain residual trees, which create a park-like appearance that classifiers struggle to distinguish from the savannas that dominate the region \citep{estes_reconciling_2016,debats_generalized_2016,sweeney_mapping_2015}.
+
+These factors suggest that the cropland map errors are likely to be even larger than we found, as well as errors in downstream products.
+For example, carbon map errors should be on the higher end of those found here (Fig.
+3), given the higher potential for error in the base cropland maps, combined with the fact that SSA's croplands lie mostly within savanna or forest biomes where the differences in carbon density between croplands and native vegetation would be higher\citep{searchinger_high_2015}.
+A presumably greater difference between the LAI and rooting depths of crops and these dominant vegetation types may also increase ET estimation errors.
+
+In terms of broader practical implications, these findings also suggest how maps errors could impact understanding of social and environmental processes and related policy.
+For example, assessments of land availability for new agricultural development could be misleading if they use a ``residual approach'' \citep{lambin_estimating_2013}, in which potential lands are identified by masking out existing croplands and un-cultivatable lands \citep[e.g.][]{estes_reconciling_2016}.
+In such cases, cropland underestimates (such as those of MODIS and GlobCover; Fig.
+1), could inflate land availability estimates, and thereby encourage erroneous land allocation policies \citep{rulli_global_2013}.
+Similarly, spatial errors in the cropland maps (Fig.
+6) could cause the wrong land to be developed, by mislocating areas with preferred development characteristics \citep[e.g.
+a combination of high agricultural potential and low environmental cost;][]{estes_reconciling_2016}.
+In addition to these possibilities, maps errors may misinform land use-focused emissions policies informed by analyses that rely on Tier-1 carbon maps \citep[e.g.][]{phelps_agricultural_2013,cattaneo_international_2010}.
+Disaggregated yield and harvested area maps could also mislead efforts to close crop production gaps, if specific interventions are targeted using finely resolved maps \citep[e.g.
+the 10 km map shown in Figure 3 in ][]{foley_solutions_2011}.
+Improper understanding of complex, coupled human-natural systems could also result from agent-based models that have calibration errors caused by land cover maps.
+
+\vspace{-0.3 cm}
+\subsection*{Recommendations}
+\vspace{-0.2 cm}
+These findings suggest several recommendations for using land cover maps, complementing those suggested in earlier work \citep{verburg_challenges_2011}.
+First, to minimize error, users should typically prefer maps derived from imagery with resolutions finer than the scale of individual objects in the cover(s) of interest, assuming that they were created with rigorous classification methods, report appropriate error metrics \citep{olofsson_good_2014}, and are thematically appropriate for the intended use \citep{verburg_challenges_2011}.
+Finer resolution not only helps to improve classification accuracy, but can minimize the \emph{aggregation problem}, one of two components of the modifiable area unit problem \citep[MAUP; ][]{openshaw_million_1979}, in which the shape and placement of the non-overlapping units used to extract map values influence the results of analyses of those values \citep{dark_modifiable_2007, marceau_scale_1999}.
+In remote sensing, an image's pixels define the fundamental mapping unit (by summarizing reflectance or radar/lidar sensors), and the pixels' dimensions influence analyses as a function of how closely they match the characteristic shapes and scales of natural features \citep{dark_modifiable_2007}.
+However, if the sensor resolution is fine enough, then mismatches can be minimized--a natural feature's shape can be approximated by aggregating several square pixels--giving the analyst greater ability to minimize errors associated with this aspect of MAUP \citep{dark_modifiable_2007, hay_comparison_2003}.
+
+Of course, high quality, fine-scaled maps such as SA-LC, which was carefully developed for South Africa, do not exist for many countries.
+In the absence of these, the new 30 m GLOBELAND30 \citep{chen_global_2015} provides global coverage for the periods 2000 and 2010, and may prove useful in smaller regions lacking their own focused maps.
+Users should first conduct their own accuracy assessments within their regions of interest to develop an appropriate understanding of their error.
+It is also becoming increasingly feasible for researchers to make their own high resolution maps, due to the rapidly expanding spatial and temporal coverage of high resolution imagery \citep[e.g.][]{hand_startup_2015,drusch_sentinel-2:_2012}, growing accessibility of high performance computing \citep[e.g.][]{pekel_high-resolution_2016}, and improvements to classification algorithms, particularly those designed to deal with the high intra-class variability inherent to certain cover types (e.g.
+cropland) and may be exacerbated by greater image resolution \citep[e.g.][]{debats_generalized_2016,sweeney_mapping_2015}.
+
+If existing high-resolution maps are not available or cannot be readily made, users interested in agricultural land cover may choose to use a new fusion map, such as Geowiki and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}.
+Although these maps are relatively coarse-scaled, the fusion methodology helps to greatly minimize error.
+By calibrating against inventory data, these maps can also reduce pixel level biases (e.g.
+Fig.
+1B).
+However, this adjustment technique surely introduces confounding errors where census-reported statistics are unreliable, such as in some SSA countries.
+\citep{carletto_emperor_2013,carletto_guesstimates_2015}.
+
+Lastly, if a particularly erroneous map is all that is available, and the user is chiefly interested in minimizing cell-wise error and less concerned about the spatial configuration of cover types, then this may be achieved by aggregating maps expressing continuous values (e.g.
+fractional cover) to coarser scales.
+This approach must be undertaken with care, as its poses a number of complications related to the \emph{scale problem}, the other half of the MAUP \citep{openshaw_million_1979}.
+These include progressive decline in variance with increasing scale \citep[even if means remain constant;][]{dark_modifiable_2007}, and reduced efficiency in estimating regression parameters from coarser map values \citep[][]{avelino_goldilocks_2016}.
+Nevertheless, aggregating continuous variables can reduce pixel errors without biasing some statistical properties (e.g.
+total cropland area remains constant using the aggregation approach demonstrated here; see Results: Cropland map errors), and has been shown to reduce other MAUP-related analytical problems \citep{avelino_goldilocks_2016}.
+However, the user must ensure that the scale of aggregation is appropriate for the analysis at hand.
+
+\vspace{-0.3 cm}
+\section*{Acknowledgments}
+\vspace{-0.2 cm}
+This work was supported by funds from the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (EAR-1 534544, SES-1360463, and BCS-1026776).
+%\end{acknowledgments}
+
+%Care in aggregation methods worth emphasizing \citep{verburg_challenges_2011,dendoncker_exploring_2008,raj_analysing_2013,moody_influence_1995}, can lead to bias, which is influenced by spatial arrangement of cover types.
+
+%[maybe] For fine features below sensor resolution, convert to quantitative measures such as SMA (tough for cropland though, but maybe with MESMA for different crop categories).
+For homogenous cover types with little within-class variability.
+
+%Marceau et al citing previous work that finer sensors will be needed to map crop field locations. Waldner et al (2016), gives field size requirements mentioned by GeoGlam
+
+%Match dataset to application; merge complementary sources where available \citep{verburg_challenges_2011}.
+
+%Our analysis demonstrates that the land cover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on high accuracy or low bias.
+
+%Synthetic assessments can help you determine how much impact error can make.
+
+%Carefully evaluate a product for its sensitivity for a particular application (Fritz et al, 2011, highlighting_2011).
+Restricted geographically, so limited inference, but some general inferences may be drawn.
+Carefully crafted Landsat products for particular scale (higher density of validation points, local knowledge).
+
+%Of course doing this loses one of the main sources of geographic information (configuration), but this is one of the many aspects that leads to the MAUP problem in any case.
+
+%Reasons for not evaluating GLC30 (lower accuracy). Other landcover products may be better, and we didnÕt evaluate these.
+
+%The resolution of the sensor relative to the typical scale of features in a landscape affects classification accuracy, and this relationship varies between features \citep{moody_influence_1995,marceau_scale_1999, marceau_scale_1999}.
+For agricultural fields, coarser resolution sensors can produce reasonably accurate estimates of total agricultural error, but are inaccurate at mapping the locations of fields \citep{pax-lenney_effect_1997,ozdogan_resolution_2006}
+
+%The 250 m resolution update of the GeoWiki map  \citep{waldner_unified_2016}, which we did not evaluate here, may help minimize errors associated with the former, but will still be exposed to the latter.
+
+%Caveats, including one related to ABM errors here?
+
+%And where background veg is similar, cite Sweeney et al, 2015, and show that similar finding in Frey and Smith (2006) for high latitude savannas and croplands. 
+%Accuracy is lowest where cropland cover is intermediate.
+
+%To a certain extent reflects configuration of landscape elements.
+
+%Our analysis allows us to assessment allows us to thoroughly examine and quantify the spatial patterning of errors
+
 %A related example can be seen in high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \citep[e.g.
@@
 For this reason, it is advisable to aggregate presented maps until their accuracy reaches an acceptable standard (Table 1).
-%
-%What then, of the impacts of land cover map error on more complex analysis?
-Here the results appear to be mixed.
-Map error appears to have had little impact on the overall bias and accuracy of VIC's evapotranspiration (ET) estimates, which contrasts with a related study that examined how land cover maps impact rainfall simulations \citep{ge_impacts_2007}.
-There are two reasons for this.
-First, map accuracy increased nearly two-fold by aggregating to 25 km, the resolution of VIC.
-Second, VIC's land cover-derived variables (e.g.
-LAI curves, effective rooting depth) could dampen the effect of map errors on ET calculations when the values of these variables in the adjacent land covers are similar to that of cropland, which is the case in much of VIC's land cover scheme for South Africa.
-In other regions, where the matrix land cover's LAI values depart substantially from cropland's (e.g.
-in forest), the ET errors would likely be much larger.
-The carbon example illustrates how cover properties can mute or amplify map error.
-%
-%Despite the low values of the overall error metrics, the hotspots of ET errors (Fig.
-4) may have outsize significance because they occurred at the climatic margins of the major crop growing regions, where a substantial share of farms are irrigated.
-Since VIC does not simulate irrigation, the magnitude of these errors is likely to have been underestimated, because irrigation substantially increases latent heat flux and can substantially modify regional climate \citep{sacks_effects_2008,mueller_cooling_2015}.
-This also suggests the potential for map error to magnify in a coupled land-atmosphere model \citep[which could explain the difference between this result and that of][]{ge_impacts_2007}, and thereby give misleading insight into the climatic effects of land use.
-%
-%Relative to the ET example, map errors had a much larger impact on the agent-based model's results.
-For the model's most important metric of food security--household-level crop production--the result was fairly straightforward for a class of models where the interactions of multiple agents often produce analytically intractable outcomes \citep{janssen_empirically_2006}.
-In this case, the model was only sensitive to cropland underestimates, which had the perfectly predictable effect of lowering average household food production, which would in turn overstate the degree of food \emph{in}security.
-Overestimates did not matter in this case, because the number of households remained fixed, and their cropland holdings were constrained to match the survey statistics (but this would matter for models that allow these parameters to vary).
-Less predictable was the finding that the spatial distribution of cover can impact model parameterization, with land being left unallocated even when needed, thus such models also require spatial, not just statistical, accuracy in their base land cover maps.
-%
+
+%\subsection*{Caveats}
+
+%Properties revealed here in this study are relevant.  Mosaic classes.  Background vegetation similar to cropland (savanna, grassland).
+
+%With the exception of the ABM-related example, these findings apply to error at the pixel level. We do not explicitly evaluate impacts on aggregate measures (such as total production).
+
 %The errors in the cropland data we tested were large, and it is unlikely that current ABM-based studies, mostly of small regions, are based on such inaccurate land cover data.
 However, as computing power increases, so, too, does the feasible application extent of ABMs, which will then become more reliant on larger-scaled, more erroneous, land cover data, with the attendant risk that insights into socio-ecological systems arising from these models could be highly misleading.
-%
-%Beyond these examples, there are other applications where land cover map errors could alter understanding and policy.
-One of these lies in assessments of land availability for agricultural expansion, which could be particularly erroneous if they use a ``residual approach'' \citep{lambin_estimating_2013}, where existing cropland and other occupied cover types are used to filter a map of potential suitability.
-Our results suggest that the pronounced underestimation biases (Fig.
-1) we found could inflate land availability estimates, and therefore increase the risk of unjust land allocation policies \citep{rulli_global_2013}.
-Another example is the growing efforts to minimize the environmental and climate impacts on new agricultural expansion frontiers, particularly Africa's savannas, which requires identifying areas that provide the highest agricultural benefit for the lowest environmental cost \citep{searchinger_high_2015}.
-This sort of analysis entails finding ``win-win'' land use solutions, which are essentially areas that deviate from an expected development cost-benefit relationship; for example, regions of unusually low carbon density but high crop fertility \citep{searchinger_high_2015}.
-The validity of such analysis are entirely dependent In order to be effective, such analyses must be undertaken at the fine scales--hectares, rather than square kilometers--at which land use decisions are made, thus their value for informing land use policy is contingent on the accuracy of their underlying land cover/land use data \citep{searchinger_high_2015}.
-The validity of such analyses, which must be undertaken at the fine-scales where land use decisions are made (hectares rather than square kilometers), depends on having high quality map data, in order to have confidence that the identified tradeoffs are based on real biophysical properties, not simply map error.
-%
-%\subsection*{The way forward}
-%Our analysis demonstrates that the land cover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on how accuracy or low bias.
-We provide some basic guidelines (Table 1) for selecting and using land cover data that can help increase confidence in the conclusions drawn from land cover-based studies.
-%
-%Beyond these guidelines, we recommend updating several important, global-scale datasets using the latest generation land cover maps.
-For example, rebuilding the widely used cropland distribution and yield maps \citep{monfreda_farming_2008, ramankutty_farming_2008} using the GeoWiki map could substantially improve our knowledge about current global agricultural distribution and productivity.
-Similarly, reconstructing the IPCC's Tier 1 vegetative carbon map \citep{ruesch_new_2008} using a combination of the GeoWiki and new global land cover consensus maps \citep[e.g.
-GLC-Share or][]{tuanmu_global_2014} the GLC-Share datasets would improve our confidence in regional to global assessments of vegetative carbon stocks.
-Alternatively, the new high-resolution, Landsat-based GLC30 \citep{chen_global_2015}, aggregated and converted to fractional cover classes at 1 km--and fused with high quality national land cover maps, where available \citep[similar to][]{fritz_mapping_2015,fritz_cropland_2011}--may provide a preferable base map for rebuilding these datasets.
-%
-%Fusing human judgement with expanding high resolution satellite data archives, improved classification algorithms, and expanding computing power appears to offer the greatest promise for improving the quality of land cover maps.
-The GeoWiki project and new high resolution global forest change maps \citep{hansen_high-resolution_2013,sexton_conservation_2015} are prominent examples of these newer methods.
-Another approach is that of the Mapping Africa project\footnote{http://mappingafrica.princeton.edu}, which, inspired by quality of the reference map used in this study, enlists internet-based workers to manually digitize cropland boundaries \citep{estes_platform_2016}.
-Merging this crowdsourced approach for creating vector maps of discrete cover types with newer computer vision-based mapping algorithms \citep{debats_generalized_2016} is a near-term goal of this project, as the former can interactively provide high quality training and test data (by using humans' superior capacity to delineate discrete cover types in noisy images \citep{estes_platform_2016}) for the latter, which has the advantage of speed and the ability to handle high-dimensional datasets.
-Approaches such as these will be needed to create the next generation of land cover maps, so that we can gain a clearer, unbiased understanding of global change.
 
-% $\theta$ changes from 0 to 1. (see Figure \ref{afoto}).That means we are considering $\theta$  of the form
-%For these solutions we have the following
+%This spatially comprehensive, bottom-up land cover map assessment provides new insight into the extent and causes of map error, and its consequences for understanding global change.
 
-%\begin{remark}
-%Note that equation \eqref{theta} specifies  the function $\varphi$
-%up to an error of order $\delta$. Theorem 1 provides an evolution
-%equation for the function $\varphi$ up to an error of order
-%$\delta |log \delta|$.
-%\end{remark}
-
-%(see \eqref{weaksol}).
-
-%\begin{materials}
-
-%\end{materials}
-
-%\appendix[App 1]
-
-%\appendix
-%This is an example of an appendix without a title.
-
-\section*{Acknowledgments}
-This work was supported by funds from the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (SES-1360463 and BCS-1026776).
-%\end{acknowledgments}
+%The analysis was made possible by a high accuracy cropland map that likely provides the truest measure of cropped area and distribution currently available for this region.
+Although this reference map is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission, SI), while temporal mismatches between the reference and test maps may account for some of the error we identified, our assessment (SI) suggests that such discrepancies do not appreciably impact our findings.
+Our results are also supported by previous, continental-scale assessments that revealed substantial errors and inconsistencies between land cover maps \citep{fritz_comparison_2010,gross_monitoring_2013}.
+These similarities, combined with the larger extents of these studies, also suggest that our findings may be applicable beyond South Africa's borders.
 

### 2017-03-22 paper/main.tex
Some more edits to Discussion and new title

@@
 
-\title{How Landcover Errors Impact Our Understanding of Socioeconomic and Environment Change}
+%\title{A Large-Area, Spatially Continuous Quantification of Error in Land Cover Maps and Associated Socioeconomic and Environmental Analyses Relevant to Global Change}
+\title{A Large-Area, Spatially Continuous Assessment of Land Cover Map Error and Its Impact on Downstream Analyses Relevant to Global Change}
+%\title{How Land Cover Errors Can Impact Understanding of Socioeconomic and Environment Phenomena}
 \author[1,2,3]{Lyndon Estes}
@@
 \begin{abstract}
-{Land cover maps increasingly underlie research into socioeconomic and environmental processes, including global change.
-It is known that map errors impact our understanding of these processes, but quantifying these impacts is difficult because many areas lack adequate reference data.
+{Land cover maps increasingly underlie research into socioeconomic and environmental patterns and processes, including global change.
+It is known that map errors impact our understanding of these phenomena, but quantifying these impacts is difficult because many areas lack adequate reference data.
 We used a high accuracy map of South African cropland to assess 1) the magnitude of error in several current generation land cover maps, and 2) how these errors propagate in downstream studies.
@@
 The value of the insights resulting from such studies depends upon the veracity of their underlying land cover data \citep{verburg_challenges_2011}, much as a house requires a solid foundation in order to remain standing.
-Unfortunately, the evidence to date indicates that this house has shaky foundations.
+Unfortunately, the evidence indicates that this house has shaky foundations \citep[e.g.][]{fritz_highlighting_2011}.
 The reason for this is that land cover data can only practically be derived from satellite imaging, which has several important constraints that propagate mapping errors.
@@
 Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties \citep[e.g.
-croplands or savannas;][]{fritz_identifying_2008,verburg_challenges_2011,debats_generalized_2016}) and thus be difficult for the classifier to distinguish.
+croplands or savannas;][]{fritz_identifying_2008,verburg_challenges_2011,debats_generalized_2016} and can thus be difficult for the classifier to distinguish.
 Discretizing a continuous cover type (e.g.
@@
 Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
-Although a number of previous studies have examined how land cover maps errors can propagate, these are primarily assessed using either simulated errors, relative differences in existing land cover maps, or ground validation data covering relatively small areas.
-\citep[e.g.][]{ge_impacts_2007, linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014,schmit_limitations_2006}.
+Although a number of previous studies have examined how land cover maps errors can propagate, these are primarily assessed using either simulated errors, relative differences in existing land cover maps, or ground validation data covering relatively small areas \citep[e.g.][]{ge_impacts_2007, linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014,schmit_limitations_2006}.
 
@@
 The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution MODIS land cover \citep{friedl_modis_2010} data (for 2011, the final year of the reference interval), which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
-The fourth dataset was the 1 km GeoWiki cropland map for Africa \citep{fritz_mapping_2015}, which fuses the cropland classes of multiple landcover products (including the three assessed here), with the resulting cropland percentages adjusted until their totals sum to match cropland areas reported in national agricultural statistics.
+The fourth dataset was the 1 km GeoWiki cropland map for Africa \citep{fritz_mapping_2015}, which fuses the cropland classes of multiple landcover products (including the three assessed here), with the resulting cropland percentages adjusted until their totals sum to match cropland areas reported in national agricultural statistics (this method also incorporated additional categorical training data collected via crowdsourcing).
 This ``hybrid-fusion'' approach represents the state of the art in mapping agricultural land cover.
-Since GeoWiki provides a continuous bounded value of land cover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.
+Since GeoWiki provides a continuous bounded value of land cover (cropland percentage), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.
 
@@
 \vspace{-0.2 cm}
-These findings suggest several recommendations for using land cover maps, complementing those suggested in earlier work \citep{verburg_challenges_2011}.
-First, to minimize error, users should typically prefer maps derived from imagery with resolutions finer than the scale of individual objects in the cover(s) of interest, assuming that they were created with rigorous classification methods, report appropriate error metrics \citep{olofsson_good_2014}, and are thematically appropriate for the intended use \citep{verburg_challenges_2011}.
-Finer resolution not only helps to improve classification accuracy, but can minimize the \emph{aggregation problem}, one of two components of the modifiable area unit problem \citep[MAUP; ][]{openshaw_million_1979}, in which the shape and placement of the non-overlapping units used to extract map values influence the results of analyses of those values \citep{dark_modifiable_2007, marceau_scale_1999}.
+These findings suggest a number of recommendations for using land cover maps, complementing those suggested in earlier work \citep{verburg_challenges_2011}, and beginning with those related to the selection of base land cover maps.
+First, to minimize error, users should typically prefer maps derived from imagery with resolutions finer than the scale of individual objects in the cover(s) of interest, assuming that they were created with rigorous classification methods accompanied by appropriate error metrics \citep{olofsson_good_2014}, and are thematically appropriate for the intended use \citep{verburg_challenges_2011}.
+Finer resolution not only helps to improve classification accuracy (see preceding section on \emph{Sources of error in cropland maps)}, but can minimize the \emph{aggregation problem}, one of two components of the modifiable area unit problem \citep[MAUP; ][]{openshaw_million_1979}, in which the shape and placement of the non-overlapping units used to extract map values influence the results of analyses of those values \citep{dark_modifiable_2007, marceau_scale_1999}.
 In remote sensing, an image's pixels define the fundamental mapping unit (by summarizing reflectance or radar/lidar sensors), and the pixels' dimensions influence analyses as a function of how closely they match the characteristic shapes and scales of natural features \citep{dark_modifiable_2007}.
@@
 In the absence of these, the new 30 m GLOBELAND30 \citep{chen_global_2015} provides global coverage for the periods 2000 and 2010, and may prove useful in smaller regions lacking their own focused maps.
-Users should first conduct their own accuracy assessments within their regions of interest to develop an appropriate understanding of their error.
-It is also becoming increasingly feasible for researchers to make their own high resolution maps, due to the rapidly expanding spatial and temporal coverage of high resolution imagery \citep[e.g.][]{hand_startup_2015,drusch_sentinel-2:_2012}, growing accessibility of high performance computing \citep[e.g.][]{pekel_high-resolution_2016}, and improvements to classification algorithms, particularly those designed to deal with the high intra-class variability inherent to certain cover types (e.g.
+Users should first conduct their own, more-focussed accuracy assessments to better understand error rates within their region of interest.
+Another option is to make one's own high resolution maps; although harder to do, this is becoming increasingly feasible due to the rapidly expanding spatial and temporal coverage of high resolution imagery \citep[e.g.][]{hand_startup_2015,drusch_sentinel-2:_2012}, growing accessibility of high performance computing \citep[e.g.][]{pekel_high-resolution_2016}, and improvements to classification algorithms, particularly those designed to deal with the high intra-class variability inherent to certain cover types (e.g.
 cropland) and may be exacerbated by greater image resolution \citep[e.g.][]{debats_generalized_2016,sweeney_mapping_2015}.
 
-If existing high-resolution maps are not available or cannot be readily made, users interested in agricultural land cover may choose to use a new fusion map, such as Geowiki and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}.
-Although these maps are relatively coarse-scaled, the fusion methodology helps to greatly minimize error.
+If existing high-resolution maps are not available or cannot be readily made, users interested in agricultural land cover may choose to use a new fusion map, such as GeoWiki and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}.
+Although these maps are relatively coarse-scaled, the fusion methodology helps to greatly minimize error, as does the use of fractional cover values.
 By calibrating against inventory data, these maps can also reduce pixel level biases (e.g.
@@
 1B).
-However, this adjustment technique surely introduces confounding errors where census-reported statistics are unreliable, such as in some SSA countries.
-\citep{carletto_emperor_2013,carletto_guesstimates_2015}.
+However, users should be aware that this adjustment technique introduces confounding errors where census-reported statistics are unreliable, such as in some SSA countries \citep{carletto_emperor_2013,carletto_guesstimates_2015}.
 
-Lastly, if a particularly erroneous map is all that is available, and the user is chiefly interested in minimizing cell-wise error and less concerned about the spatial configuration of cover types, then this may be achieved by aggregating maps expressing continuous values (e.g.
+If a particularly erroneous map is all that is available, and the user is chiefly interested in minimizing cell-wise error and less concerned about the spatial configuration of cover types, then this may be achieved by aggregating maps expressing continuous values (e.g.
 fractional cover) to coarser scales.
@@
 However, the user must ensure that the scale of aggregation is appropriate for the analysis at hand.
+
+Finally, users (or makers) of downstream products should rigorously ascertain how error propagates from the base land cover map into their derived maps \citep{verburg_challenges_2011}.
+In some cases, the providers of ``off-the-shelf'' products report pixel-level uncertainty values \citep[e.g.][]{ramankutty_farming_2008}, which may be sufficient.
+In other cases, downstream maps may lack quantified confidence intervals \citep[e.g.][]{monfreda_farming_2008}.
+Although such maps may provide guidelines for appropriate usage\footnote{www.earthstat.org\/wp-content\/uploads\/METADATA\_HarvestedAreaYield175Crops.pdf}, which our analyses help to further illustrate, users should undertake their own two-step error assessments.
+The most straightforward way may be to use a Monte Carlo approach that generates artificial datasets \citep[e.g.][]{avelino_goldilocks_2016} with introduced error, drawn from either reported or user-determined accuracy statistics, for both the base land cover and the downstream maps.
+Quantifying error propagation in this (or another appropriate) manner is important to understanding how map error may influence subsequent understanding of the phenomenon of interest.
 

### 2017-03-24 paper/SI.tex
Final edits from Tobias incorporated into main text and spillover into SI.

@@
 
+\subsection{\large Survey statistics}
+The frequency distributions of household cropland holdings were derived from survey data obtained from Zambia (no equivalent data were available for South Africa).
+We used these statistics to calculate the ``true'' number of households per district, as well as household cropland area distributions.
+The average field size in these studies was 1 ha, which determined the resolution of the base cropland maps (100 m).
+
 \subsection{\large Maize production}

### 2017-03-24 paper/main.tex
Final edits from Tobias incorporated into main text and spillover into SI.

@@
 %\title{A Large-Area, Spatially Continuous Quantification of Error in Land Cover Maps and Associated Socioeconomic and Environmental Analyses Relevant to Global Change}
-\title{A Large-Area, Spatially Continuous Assessment of Land Cover Map Error and Its Impact on Downstream Analyses Relevant to Global Change}
+\title{A Large-Area, Spatially Continuous Assessment of Land Cover Map Error and Its Impact on Downstream Analyses}
 %\title{How Land Cover Errors Can Impact Understanding of Socioeconomic and Environment Phenomena}
@@
 \author[5]{Stefanus Ferreira}
-\author[6]{Tobias Kuemmerle}
+\author[6,7]{Tobias Kuemmerle}
 \author[2]{Gabrielle Ragazzo}
-\author[2,7]{Justin Sheffield}
-\author[8]{Adam Wolf}
+\author[2,8]{Justin Sheffield}
+\author[9]{Adam Wolf}
 \author[2]{Eric Wood}
-\author[2,9]{Kelly Caylor}
+\author[2,10]{Kelly Caylor}
 
@@
 \affil[6]{Geography Department, Humboldt University, 10099 Berlin, Germany}
-\affil[7]{Geography and Environment, University of Southampton, Southampton, United Kingdom}
-\affil[8]{Arable Labs, Princeton, NJ USA}
-\affil[9]{Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA USA}
+\affil[7]{Integrative Research Institute for Transformations in Human-Environment Systems, Humboldt University, 10099 Berlin, Germany}
+\affil[8]{Geography and Environment, University of Southampton, Southampton, United Kingdom}
+\affil[9]{Arable Labs, Princeton, NJ USA}
+\affil[10]{Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA USA}
 
@@
 It is known that map errors impact our understanding of these phenomena, but quantifying these impacts is difficult because many areas lack adequate reference data.
-We used a high accuracy map of South African cropland to assess 1) the magnitude of error in several current generation land cover maps, and 2) how these errors propagate in downstream studies.
-We first quantified pixel-wise errors in the cropland classes of four widely used land cover maps at resolutions ranging from 1-100 km, then calculated errors in several representative ``downstream'' (map-based) analyses, including estimates of vegetative carbon stocks and crop production, simulated evapotranspiration, and agent-based assessment of household food security.
-We also assessed maps' spatial accuracy based on how precisely they could be used to locate specific landscape features.
+We used a highly accurate, high-resolution map of South African cropland to assess 1) the magnitude of error in several current generation land cover maps, and 2) how these errors propagate in downstream studies.
+We first quantified pixel-wise errors in the cropland classes of four widely used land cover maps at resolutions ranging from 1 to 100 km, then calculated errors in several representative ``downstream'' (map-based) analyses, including assessments of vegetative carbon stocks, evapotranspiration, crop production, and household food security.
+We also evaluated maps' spatial accuracy based on how precisely they could be used to locate specific landscape features.
 We found that cropland maps can have substantial biases and poor accuracy at all resolutions (e.g.
@@
 Constraining mapped values to match survey statistics proved effective at minimizing bias (provided the statistics are accurate).
-Downstream errors could be substantially amplified or muted, depending on the values ascribed to cropland-adjacent covers (e.g.
+Errors in downstream analyses could be substantially amplified or muted, depending on the values ascribed to cropland-adjacent covers (e.g.
 with forest as adjacent cover, carbon map error was 200-500\% greater than in the foundational cropland maps, but $\sim$40\% less for sparse cover types).
@@
 Our increasing modification of the Earth's surface \citep{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through land cover.
-To fully understand these processes, including the forces driving climate change \citep{verburg_challenges_2011}, it is essential to have an accurate understanding of the nature and distribution of land cover.
-
+To fully understand these processes, it is essential to have an accurate understanding of the nature and distribution of land cover \citep{verburg_challenges_2011}.
 This importance is understood by a growing number of social, economic, and natural scientists, who are using land cover data to advance understanding of food security \citep{lark_cropland_2015,wright_recent_2013, licker_mind_2010}, carbon cycling \citep{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \citep{newbold_global_2015, luoto_predicting_2004}, demographic shifts \citep{linard_assessing_2010}, and other important facets of Earth System processes.
-
-% (we hereafter collectively refer to this sort of work as Earth System Science (ESS), of which global change research is a subset).
 
@@
 First, in many regions the spatial arrangement of cover types is smaller than the sensor resolution \citep[e.g.
-smallholders' farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015,fritz_identifying_2008}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
+smallholders' fields][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015,fritz_identifying_2008}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
 Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties \citep[e.g.
@@
 Likewise, the importance of assessing the accuracy of land cover maps is increasingly recognized, and there are well-developed, best-practice guidelines for gathering and using ground-truth samples to robustly quantify map error \citep{foody_status_2002,olofsson_making_2013,olofsson_good_2014,stehman_global_2012}.
-Because comprehensive, spatially representative ground truth data are typically unavailable for rapidly changing regions \citep{see_improved_2015,kuemmerle_challenges_2013}, what remains an open question is exactly how much the maps researchers typically use deviate from actual land cover, and how this in turn impacts our understanding of environmental change processes.
+Because comprehensive, spatially representative ground truth data are typically unavailable for rapidly changing regions \citep{see_improved_2015,kuemmerle_challenges_2013}, what remains an open question is exactly how much the maps researchers typically use deviate from actual land cover, how this affects our analyses based on these maps, and how this in turn impacts our understanding of environmental change processes.
 Our current understanding of map accuracy over such areas is often based on scarce information or top-down "sanity checks" made in comparison to aggregated survey data \citep{yu_meta-discoveries_2014,larsen_taken_2015}.
 
-%Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are typically unavailable for Africa and other rapidly changing regions \citep{see_improved_2015}, which have also been fairly sparsely mapped \citep{yu_meta-discoveries_2014}.
-
 Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
-Although a number of previous studies have examined how land cover maps errors can propagate, these are primarily assessed using either simulated errors, relative differences in existing land cover maps, or ground validation data covering relatively small areas \citep[e.g.][]{ge_impacts_2007, linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014,schmit_limitations_2006}.
-
-%Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
-Some work has examined how map error influences climate simulations \citep{ge_impacts_2007}, agricultural land use patterns \citep{schmit_limitations_2006}, carbon flux measurements \citep{quaife_impact_2008}, human population estimates \citep{linard_assessing_2010}, and species distribution models \citep{tuanmu_global_2014}, but these either use simulated land cover errors \citep{ge_impacts_2007} or compare relevant differences in estimates between different satellite-derived land cover maps \citep{linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}, or, in cases where spatially explicit ground validation maps were used for assessment, cover small and discontiguous extents \citep{schmit_limitations_2006}.
+Although previous studies have examined how map errors propagate, these are primarily assessed using either simulated errors, relative differences in existing land cover maps, or ground validation data covering relatively small areas \citep[e.g.][]{ge_impacts_2007, linard_assessing_2010, quaife_impact_2008, tuanmu_global_2014, schmit_limitations_2006}.
 
 Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g.
-3DRobotics, DJIA} to new high resolution satellite arrays \citep{drusch_sentinel-2:_2012,hand_startup_2015} and better mapping methods \citep{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013,debats_generalized_2016}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the land cover products that have become commonplace in global change research--and which are often used to make policy decisions \citep{searchinger_high_2015}.
+3DRobotics, DJIA} to new high resolution satellite arrays \citep{drusch_sentinel-2:_2012,hand_startup_2015} and better mapping methods \citep{fritz_geo-wiki:_2012,estes_projected_2013,debats_generalized_2016}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the land cover products that have become commonplace in global change research--and which are often used to make policy decisions \citep{searchinger_high_2015}.
 
 In this study, we take advantage of this recent growth in data to address the call to more thoroughly assess errors in land cover maps \citep{kuemmerle_challenges_2013, olofsson_good_2014,olofsson_global_2012}, and further examine how these errors might impact our understanding of socioeconomic and environmental conditions.
-Using a unique, high-quality, map of South African croplands, which was created by expert mappers delineating individual fields visible within high resolution imagery, we conduct spatially comprehensive, bottom-up analyses to answer the following two questions: 1) What is the extent of error in several widely used land cover products?; 2) How do these errors propagate through downstream biophysical and socioeconomic studies?
+Using a unique, high-resolution, high-quality map of South African croplands, which was created by expert mappers delineating individual fields visible within high resolution imagery, we conduct spatially comprehensive, bottom-up analyses to answer the following two questions: 1) What is the extent of error in several widely used land cover products?; 2) How do these errors propagate through downstream biophysical and socioeconomic analyses?
 
-The answers to these questions provide important insights into how cropland datasets can influence our understanding socioeconomic and environmental processes in South Africa, as well as more broadly throughout the rest of sub-Saharan Africa (SSA), where our current knowledge of the extent and distribution of cropland relies heavily on land cover maps \citep[][]{fritz_comparison_2010, see_improved_2015}.
-
-%Given this dependency and the fact that SSA is poised to undergo a dramatic agricultural expansion during the next several decades \citep{searchinger_high_2015, estes_reconciling_2016}, these findings should also have relevance for contain implications for our broader ability to grasp global change processes.
-
-%The answers to these questions hold significance that extends beyond the borders of South Africa, for several reasons.
-First, South Africa is one of the major agricultural centers in sub-Saharan Africa (SSA), where a rapid agricultural expansion has begun that will play a major role in shaping the future trajectory of global change \citep{searchinger_high_2015, estes_reconciling_2016}.
-Second, South African agriculture spans the diversity of farming systems found in SSA, but has a more advanced commercial sector; it is likely that SSA's broader cropland estate will increasingly resemble South Africa's.
-Third, baseline understanding of the distribution and extent of agricultural land use in SSA is poor because existing land cover maps and agricultural census data are both inaccurate \citep{fritz_comparison_2010, carletto_emperor_2013}.
-By quantifying how cropland datasets can influence our understanding of socioeconomic and environmental processes in South Africa, our study therefore provides insights
-
-%First, South Africa's agricultural sector spans the diversity of farming systems found in sub-Saharan Africa (SSA), but at a more advanced stage of development \citep{hardy_rainfed_2011,estes_using_2014} one of the major agricultural centers in sub-Saharan Africa (SSA), where a rapid agricultural expansion has begun that will play a major role in shaping the future trajectory of global change \citep{searchinger_high_2015, estes_reconciling_2016}.
-Second, our baseline understanding of the distribution and extent of agricultural land use in this region is poor because existing land cover maps and agricultural census data are both inaccurate \citep{fritz_comparison_2010, carletto_emperor_2013}.
-Because South Africa's agricultural sector spans the diversity of farming systems found throughout SSA \citep{hardy_rainfed_2011,estes_using_2014}, our study provides insight into the important role and impact that land cover datasets have on understanding global change dynamics in this critical region.
-
-%The answers to these questions can provide important insights into how cropland datasets can influence broader understanding of global change processes, even though our study is limited to a single country.
-This The reason for this is two-fold.
-First, South Africa is one of the major agricultural centers in sub-Saharan Africa (SSA), where a rapid agricultural expansion has begun that will play a major role in shaping the future trajectory of global change \citep{searchinger_high_2015, estes_reconciling_2016}.
-Our baseline understanding of the distribution and extent of agricultural land use in this region is poor because existing land cover maps and agricultural census data are both inaccurate \citep{fritz_comparison_2010, carletto_emperor_2013}.
-Because South Africa's agricultural sector spans the diversity of farming systems found throughout SSA \citep{hardy_rainfed_2011,estes_using_2014}, our study provides insight into the important role and impact that land cover datasets have on understanding global change dynamics in this critical region.
-
-%, our findings can provide important insight into how errors and biases in cropland datasets may influence our broader understanding of change in key components of the Earth System.
-
-%We focus on sub-Saharan Africa, one of the world's most rapidly changing regions, where a rapid agricultural expansion is underway \citep{searchinger_high_2015,estes_reconciling_2016} yet the accuracy of maps of existing agriculture land is particularly poor \citep{fritz_comparison_2010}.
-
-%We specifically quantify the error and bias are in cropland cover estimates associated with different land cover products, how these vary with scale, and , and how do these vary with scale and by land cover product; 2.
-How do these errors and biases
-
-%We use a unique, high-quality map of South African croplands to conduct a spatially comprehensive, bottom-up quantification of error in several widely used land cover maps, and how these errors can propagate through downstream biophysical and socioeconomic studies.
-South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \citep{searchinger_high_2015,estes_reconciling_2016}, yet is notably lacking trustworthy maps of existing agriculture land \citep{fritz_comparison_2010}.
-
-%Moreover, South Africa's large agricultural sector represents the diversity of farming systems found throughout SSA, ranging from large commercial operations to smallholder farms \citep{hardy_rainfed_2011,estes_using_2014}.
-As such, our findings can provide important insight into how cropland datasets may influence our broader understanding of environmental change.
+The answers to these questions provide important insights into how cropland datasets can influence our understanding socioeconomic and environmental processes in South Africa, as well as more broadly throughout sub-Saharan Africa (SSA), where our current knowledge of the extent and distribution of cropland relies heavily on land cover maps \citep[][]{fritz_comparison_2010, see_improved_2015}.
 
@@
 The resulting vectorized field boundaries provide unique, highly accurate data on field sizes and distribution for 2009-2011.
-The accuracy of this dataset was evaluated in a previous study \citep{estes_platform_2016}, in which visual assessment of cropland presence/absence was made within 15,225 4 ha plots (25 sub-plots within 609 1 km$^2$ grids) placed over high resolution satellite imagery, and compared to coverage by the vector boundaries.
+The accuracy of this dataset was evaluated in a previous study \citep{estes_platform_2016}, in which visual assessment of cropland presence/absence was made within 15,225 individual 4 ha plots (25 sub-plots within 609 1 km$^2$ grids) placed over high resolution satellite imagery, and compared to coverage by the vector boundaries.
 Measured in terms of its ability to distinguish crop fields from other cover types, the results showed these data to be 97\% percent accurate, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\% \citep[see SI for more description, and][]{estes_platform_2016}.
-
-%This dataset is particularly valuable because South Africa represents nearly 6\% of sub-Saharan Africa's (SSA) area, which is a region that is poised to undergo rapid agricultural expansion \citep{searchinger_high_2015}, yet is notably lacking trustworthy maps of existing agriculture land \citep{fritz_comparison_2010}.
-Moreover, South Africa's large agricultural sector represents the diversity of farming systems found throughout SSA, ranging from large commercial operations to smallholder farms \citep{hardy_rainfed_2011,estes_using_2014}.
 
@@
 \vspace{-0.2 cm}
-We then used these maps to conduct four analyses typical of global change research: 1) estimation of carbon stocks, 2) simulation of evapotranspiration, 3) disaggregation of crop yield and production, and 4) simulating household dynamics using an agent-based model.
+We then used the reference and test maps to conduct four analyses typical of global change research: 1) estimation of carbon stocks, 2) simulation of evapotranspiration, 3) disaggregation of crop yield and production, and 4) simulating household dynamics using an agent-based model.
 The first and third analyses were relatively simple, in that the variable(s) of interest were mapped onto land cover using empirical relationships.
@@
 \vspace{-0.2 cm}
-To understand the carbon cycle and climate forcing due to land use change, it is important to have accurate, high resolution maps of vegetative carbon stocks \citep[][]{searchinger_high_2015}.
+To understand the carbon cycle and climate forcing due to land cover change, it is important to have accurate, high resolution maps of vegetative carbon stocks \citep[][]{searchinger_high_2015}.
 One widely used vegetative carbon dataset is that of \citep{ruesch_new_2008}, who mapped estimated carbon density values for different vegetation types to the classes of a global land cover product.
@@
 We multiplied cropland densities by cropland fractions and added these to each of the five other densities multiplied by the residual non-cropland fractions to create five different carbon density maps.
-We aggregated each carbon map up to the five coarser resolutions for scaling comparisons (SI).
+We aggregated each carbon map up to 5, 10, 25, 50, and 100 km resolutions for scaling comparisons (SI).
 
@@
 \vspace{-0.2 cm}
-The spatial variability of yield and production is critical for understanding food security, trade, and the potential for agricultural expansion and intensification \citep{licker_mind_2010,monfreda_farming_2008}.
+The spatial variability of crop yield and production is critical for understanding food security, trade, and the potential for agricultural expansion and intensification \citep{licker_mind_2010,monfreda_farming_2008}.
 The most reliable source of such data are national to sub-national agricultural statistics, often available only at relatively coarse-scaled administrative boundaries.
@@
 We used these methods to first disaggregate the harvested area for maize (South Africa's largest crop \citep{estes_projected_2013}) onto our cropland maps, followed by yields, which were assigned to cells having harvested areas greater than zero.
-The first step in this process entails adjusting cropland percentages so that their totals match census-derived cropland area estimates \citep{ramankutty_farming_2008}.
+The first step in this process entails adjusting cropland percentages so that their totals match census-derived cropland area estimates \citep{ramankutty_farming_2008,schierhorn_post-soviet_2013}.
 In place of census statistics, we used the reference map to calculate total cropland areas for South Africa's nine provinces, then adjusted the pixel-wise cropland percentages in the four test maps so that their province-wise sums matched these totals (SI).
@@
 \vspace{-0.2 cm}
-Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making, to analyze socio-ecological system dynamics, and to facilitate improved policy-making \citep{berger_creating_2006}.
+Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making, to analyze socio-ecological system dynamics, and to test policy impact \citep{berger_creating_2006}.
 To obtain robust insights, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistically represents the social and biophysical features of the study region \citep{berger_creating_2006}.
@@
 In our example, we used an ABM of household food security that simulates food production by individual farming households \citep[the agents; ][]{chen_dependency_2013}.
-The model (described briefly here and in more detail in the SI) is initialized so that each household is allocated a designated share of cropland (the number of households and their cropland areas are derived from survey statistics), on which annual household's crop production (maize) is simulated as a function of its field area, local weather, soil properties, and management actions.
-As these factors differ between households, household-level crop production can vary substantially.
+The model (described in more detail in the SI) is initialized so that each household is allocated a share of cropland, based on household number and cropland area estimated derived from survey statistics.
+Annual household crop production (maize) is simulated as a function of its field area, local weather, soil properties, and management actions, all of which can vary between households.
 The initialization process iteratively assigns households to the landscape as a function of neighbor and cropland proximity, ensuring that households are grouped into communities and that their fields are within a realistic proximity.
-To achieve this, the model first takes a weighted (by cropland area frequency) random draw of 100 households and places these within the simulated landscape, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household.
+To achieve this, the model first randomly places 100 households onto the simulated landscape, allocating each household its required cropland pixels, which must be within 1.5 km of the household.
 This process is iterated until all households are assigned cropland, or all available cropland is allocated.
@@
 
-Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g.
-district, rather than country, scales) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of individual farmers.
-To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$) in eastern South Africa with similar climate and 28-45\% of cropland coverage.
-The frequency distributions of household cropland holdings were derived from survey data obtained from Zambia (no equivalent data were available for South Africa).
-We used these statistics to calculate the ``true'' number of households per district, as well as household cropland area distributions.
-To create cropland surfaces for each district, we disaggregated all five cropland maps to 100 m, binary cropland/non-cropland cover maps (to match the survey-recorded average field size of 1 ha).
-We ran the model separately for each district, with household agents planting and managing a maize crop throughout the growing season.
-
-We ran each ABM simulation for a single season, once per district and when initialized by each of the cropland maps (20 simulations total).
-To examine how map errors impacted the land allocation process and household food production estimates, we calculated three variables: the percent of unallocated cropland, land deficit, and food deficit.
-The percent of unallocated cropland is the share of the district's total cropland area that was not assigned to any household, which measures how effective the model was in matching household agents to available cropland resources.
-Land deficit is the total area of cropland that should have been allocated to households in each district but wasn't, due to the mismatch between the cropland map and the survey-based estimates of what total cropland holdings should have been.
-Food deficit is the percentage shortfall in the average amount of food production that should have been produced by each household but wasn't because of the land deficit.
+Like many spatial ABMs, the model is computationally intensive and requires high spatial resolution to match the scale of individual fields, and therefore applied to smaller regions (e.g.
+districts).
+To meet these computational needs, we selected four contiguous magisterial districts (1,040-1,343 km$^2$) in eastern South Africa with similar climate and 28-45\% of cropland coverage.
+To create cropland surfaces for each district, we disaggregated all five cropland maps into 100 m binary cropland/non-cropland cover maps, and then ran the model separately for each district and with each cropland map (20 simulations total).
+To examine how map errors impacted the land allocation process and household food production estimates, we calculated three variables: the percent of unallocated cropland, land deficit, and food deficit, which respectively represent: 1) the share of district total cropland that was not assigned to any household (a measure of the model's effectiveness in matching households to available cropland); 2) the total area of cropland that should have been allocated to households in each district but wasn't (due to mismatches between the cropland map and the survey-based estimates of total cropland holdings); 3) the percentage shortfall in the average amount of food production that should have been produced by each household but wasn't because of the land deficit.
 
@@
 It is therefore important to also assess how map errors can impact the spatial accuracy of maps, or the ability to accurately locate specific features of interest.
-To evaluate this, we calculated the mean euclidean distance (in km) between pixels representing a specific feature within the test maps relative to their nearest neighboring pixels representing the same feature in the reference map.
+To evaluate this, we calculated the mean Euclidean distance (in km) between pixels representing a specific feature within the test maps relative to their nearest neighboring pixels representing the same feature in the reference map.
 The features in this analysis were simply those locations falling within the upper deciles of a) cropland cover, b) carbon density (based on the average carbon density of non-crop vegetation types), and c) crop yield.
@@
 2).
-In other words, accuracy is generally lowest when cropland cover is mixed evenly with other cover types.
+In other words, accuracy was lowest when cropland cover was mixed evenly with other cover types.
 GlobCover's accuracy continued to decrease with cropland density because the dominant agricultural cover class contributing to the test map was defined as 50-70\% crops mingled with other vegetation, thus the maximum percentage was constrained by this mixture range.
@@
 
-Production biases were generally slightly higher, but still low, for most datasets, with the exception of GlobCover, which had a large underestimation bias of $>$60\% (relative to mean production) at 1 km, which remained above 10\% even at 100 km of aggregation.
+Production biases were generally higher, but still low, for most datasets, with the exception of GlobCover, which had a large underestimation bias of $>$60\% (relative to mean production) at 1 km, which remained above 10\% even at 100 km of aggregation.
 MODIS production bias was above 20\% at 1 km, but declined to below 10\% at higher levels of aggregation.
@@
 
-The preceding analyses contributes to a substantial body of existing work devoted to understanding the nature and consequences of error within landcover maps \citep[e.g.][]{fritz_highlighting_2011,verburg_challenges_2011,olofsson_making_2013}.
-A number of prior studies have assessed land cover map errors, either through point-based accuracy assessments \citep[e.g.][]{olofsson_making_2013,frey_how_2007-1,foody_status_2002}, by evaluating between-map discrepancies \citep[e.g.][]{fritz_highlighting_2011,fritz_identifying_2008,fritz_comparison_2010}, or by comparing map-derived estimates to aggregated statistics \citep[e.g.][]{larsen_taken_2015,fritz_comparison_2010,yu_meta-discoveries_2014}.
-Several studies have used contiguous ground truth maps to assess land cover errors, but these covered relatively small regions \cite[$<$3000 km$^2$, or $<$0.03\% of the area covered here;][]{dendoncker_exploring_2008,schmit_limitations_2006}.
+The preceding analyses contributes to existing work investigating land cover map error and its consequences \citep[e.g.][]{fritz_highlighting_2011,verburg_challenges_2011,olofsson_making_2013}.
+Previous studies have assessed map errors either by using point-based accuracy assessments \citep[e.g.][]{olofsson_making_2013,frey_how_2007-1,foody_status_2002}, by evaluating between-map discrepancies \citep[e.g.][]{fritz_highlighting_2011,fritz_identifying_2008,fritz_comparison_2010}, or by comparing map-derived estimates to aggregated statistics \citep[e.g.][]{larsen_taken_2015,fritz_comparison_2010,yu_meta-discoveries_2014}.
+A smaller number based their assessments on contiguous ground truth maps, but these covered relatively small regions \cite[$<$3000 km$^2$, or $<$0.03\% of the area covered here;][]{dendoncker_exploring_2008,schmit_limitations_2006}.
 
-Previous work has also examined how map errors influence estimates of environmental and socioeconomic variables, including simulated rainfall \citep{ge_impacts_2007}, carbon stocks and emissions \citep{goetz_mapping_2009, quaife_impact_2008, olofsson_making_2013,jain_co2_2013}, nitrogen fluxes \citep{jain_co2_2013,nol_effect_2008}, human population density \citep{linard_assessing_2010}, species distributions \citep{tuanmu_global_2014}, and landscape patterns \citep{langford_map_2006}.
-The majority of these used either point validation, map inter-comparison, or a combination of both to directly assess errors in the measurement of interest \citep{goetz_mapping_2009, olofsson_making_2013, jain_co2_2013, linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}.
-A smaller number specifically examined how either simulated map errors \citep{ge_impacts_2007,langford_map_2006} or differences relative to small area ground truth maps \citep[$<$1000 m;][]{nol_effect_2008} propagate into the downstream estimates.
+Other studies have also examined how map errors impact downstream analyses, including simulated rainfall \citep{ge_impacts_2007}, carbon stocks and emissions \citep{goetz_mapping_2009, quaife_impact_2008, olofsson_making_2013,jain_co2_2013}, nitrogen fluxes \citep{jain_co2_2013,nol_effect_2008}, human population density \citep{linard_assessing_2010}, species distributions \citep{tuanmu_global_2014}, and landscape patterns \citep{langford_map_2006}.
+The majority of these used either point validation, map inter-comparison, or a combination of both to assess errors \citep{goetz_mapping_2009, olofsson_making_2013, jain_co2_2013, linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014}.
+Others have used simulated map errors \citep{ge_impacts_2007,langford_map_2006} or differences relative to small area ground truth maps \citep[$<$1000 m;][]{nol_effect_2008} to examine error propagation.
 
-Our study builds on these previous efforts by providing a large-area, spatially continuous quantification of cropland classification errors, and examines how these actual errors influence several common downstream applications.
+Our study builds on and goes beyond these previous efforts by providing a large-area, spatially continuous quantification of cropland classification errors, and by examining how these actual errors influence several common downstream applications.
 The spatially comprehensive nature of these analyses provides deeper insight into the causes and consequences of error than would otherwise be obtained from either a point-based accuracy assessment or through the imposition of simulated error.
@@
 Our findings showed that the most accurate cropland estimates, across all spatial scales, were produced by SA-LC followed by GeoWiki.
-In the reverse order, these two datasets were also the least biased, with GeoWiki having effectively zero bias at 5 km or higher of aggregation.
+In the reverse order, these two datasets were also the least biased, with GeoWiki having effectively zero bias at aggregation levels of 5 km or coarser.
 GlobCover was highly inaccurate and biased at all scales, with MODIS estimates being nearly as inaccurate but substantially less biased.
@@
 Both tendencies were evident in the Tier-1 carbon maps, the simplest of the downstream methods.
-Using the ratio between each carbon map's accuracy score (MAE) and that of its foundational cropland map to calculate error propagation ($>$1 means error was exacerbated, $<$1 means error muted), we see that errors in the 1 km carbon map errors were 200 (SA-LC) to 500\% (GlobCover) larger than cropland map errors when forest was the adjacent cover type, but $\sim$40\% lower for the sparse cover type.
+Using the ratio between each carbon map's accuracy score (MAE) and that of its foundational cropland map to calculate error propagation (values $>$1 means error was exacerbated, $<$1 means it was muted), we see that errors in the 1 km carbon map errors were 200 (SA-LC) to 500\% (GlobCover) larger than cropland map errors when forest was the adjacent cover type, but $\sim$40\% lower for the sparse cover type.
 Aggregation helped to reduce error magnitude, but carbon maps nevertheless had 30-50\% more error than cropland maps at 100 km resolution when forest or shrubland were sharing the pixel (the error ratio associated with the sparse vegetation carbon class remained relatively constant with aggregation).
@@
 The disaggregated yield and crop production maps we created also showed both error amplification and muting, which in this case depended on the particular analysis.
-Errors within the yields maps were uniformly lower (50-90\% less at 1 km) than those in the input cropland maps, whereas errors in the production maps were 70-100\% higher at 1 km, and actually were exacerbated at intermediate levels of aggregation (10-25 km) for GlobCover, MODIS, and GeoWiki (170-290\%); this latter tendency mostly caused by accuracy in the production maps improving more slowly with aggregation (Fig.
+Errors within the yields maps were uniformly lower (50-90\% less at 1 km) than those in the input cropland maps, whereas errors in the production maps were 70-100\% higher at 1 km, and actually were exacerbated at intermediate levels of aggregation (10-25 km) for GlobCover, MODIS, and GeoWiki (170-290\%).
+This latter tendency was mostly caused by accuracy in the production maps improving more slowly with aggregation (Fig.
 5) than in the original cropland maps (Fig.
@@
 This simplicity means that it is only sensitive to errors in classifying cropland presence/absence, but not to errors in cropland proportions.
-Production, on the other hand, is calculated from disaggregated harvested area maps, which are created by proportionally allocating district-level crop harvested area onto cropland fractions, making them highly sensitive errors in cropland proportion.
+Production, on the other hand, is calculated from disaggregated harvested area maps, which are created by proportionally allocating district-level crop harvested area onto cropland fractions, making them highly sensitive to errors in cropland proportion.
 Of particular note is that harvested area maps are subject not only to this statistical constraint (that pixel-wise harvested area fractions sum to district totals), but cropland fractions are also adjusted to match district-level cropland area estimates \citep[see Methods; ][]{ramankutty_farming_2008}.
@@
 
-Sensitivity to cropland area also was evident in the agent-based model's results.
+Sensitivity to cropland area also was evident in the food security model.
 The ABM's most important metric of food security--household-level crop production--was only impacted by underestimates of cropland area, which lowered the models' estimates of average household production (in turn overstating the degree of food \emph{in}security) because individual households were allocated insufficient cropland.
 Cropland overestimates did not cause the opposite effect, because total households were constant and the allocation routine prevented cropland holdings from exceeding their assigned, census-derived hectarage.
-The less predictable result was that the model could leave cropland unallocated when there were substantial cropland underestimates (e.g.
+The less predictable result was that the model sometimes left cropland unallocated when maps substantially underestimated cropland area (e.g.
 MODIS in Fig.
-6A), which was an error caused by the spatial arrangement of croplands in the district.
+6A).
+This initialization error was caused by the spatial arrangement of croplands in the district.
 MODIS cropland in those districts was confined to relatively small islands, meaning that they fell outside the allocation algorithm's search window, thus leaving those croplands unassigned to specific households.
@@
 Under this conception, the error in all but SA-LC's cropland and yield examples (Fig.
-7) is $>$3 pixels (or 300\%), with an overall average of nearly 6 pixels.
-
-%the result was fairly straightforward for a class of models where the interactions of multiple agents often produce analytically intractable outcomes \citep{janssen_empirically_2006}.
-In this case, the model
-
-%Yield disaggregation and production methods reinforce to a certain extent how statistical constrain can minimize bias at the pixel level.
-[check this against accuracy].
-But can still be off as a function of cropland errors.
-Highlight by production, and locational accuracy measures.
-
-%Our results also suggest that, despite their higher accuracy, they may not necessarily be the least biased maps.
-Except for a few results (notably the 1 km carbon and maize production estimates), GeoWiki-based maps had the lowest bias across most scales, which is the largely the result of calibrating the cropland percentages to match reported agricultural statistics \citep{fritz_cropland_2011,fritz_mapping_2015}.
-This procedure is similar to the one we used in disaggregating maize harvested areas \citep{ramankutty_farming_2008, monfreda_farming_2008}, which led to fairly unbiased estimates of maize production and yield (Fig.
-5) above 1 km (except for GlobCover's production estimates).
-This result, together with GeoWiki's unbiased cropland estimates (Fig 1.B), indicates the value of fusing inventory data with remote sensing to reduce map bias.
-
-%There are two reasons for this. First, map accuracy increased nearly two-fold by aggregating to 25 km, the resolution of VIC.
+7) is $>$3 pixels (or 300\%), with an overall average of nearly six pixels.
 
@@
 \vspace{-0.2 cm}
-Although our study focused on a single country and a subset of possible land cover-derived analyses, its findings should have broader geographic and practical relevance.
-In geographical terms, the key question is whether the error patterns revealed here will be reproduced outside of South Africa's borders?
-If one considers the rest of Sub-Saharan Africa (SSA), the answer is almost certainly yes.
+Although our study focused on a single country and a subset of possible land cover-derived analyses, its findings highlight issues with broader geographic and practical relevance.
+In geographical terms, the key question is whether the error patterns revealed here will be similar outside of South Africa?
+Land systems in many regions differ substantially from those in southern Africa, and further research on the bias and accuracy of cropland maps is thus desirable.
+Yet, if one considers the rest of Sub-Saharan Africa (SSA), a region notorious for its lack of high-quality cropland maps, the answer is almost certainly yes.
 Previous work showing substantial disagreement between different cropland maps in other SSA countries \citep{fritz_comparison_2010} supports this contention.
@@
 For example, carbon map errors should be on the higher end of those found here (Fig.
-3), given the higher potential for error in the base cropland maps, combined with the fact that SSA's croplands lie mostly within savanna or forest biomes where the differences in carbon density between croplands and native vegetation would be higher\citep{searchinger_high_2015}.
+3), given the higher potential for error in the base cropland maps, combined with the fact that SSA's croplands lie mostly within savanna or forest biomes where the differences in carbon density between croplands and native vegetation would be higher \citep{searchinger_high_2015}.
 A presumably greater difference between the LAI and rooting depths of crops and these dominant vegetation types may also increase ET estimation errors.
@@
 In such cases, cropland underestimates (such as those of MODIS and GlobCover; Fig.
-1), could inflate land availability estimates, and thereby encourage erroneous land allocation policies \citep{rulli_global_2013}.
+1), could inflate estimates of available land, and thereby encourage erroneous land policy \citep{rulli_global_2013}.
 Similarly, spatial errors in the cropland maps (Fig.
-6) could cause the wrong land to be developed, by mislocating areas with preferred development characteristics \citep[e.g.
-a combination of high agricultural potential and low environmental cost;][]{estes_reconciling_2016}.
+6) could cause the wrong land to be developed, by mis-locating areas with preferred development characteristics \citep[e.g.
+high agricultural potential and low environmental cost;][]{estes_reconciling_2016, gasparri_emerging_2015}.
 In addition to these possibilities, maps errors may misinform land use-focused emissions policies informed by analyses that rely on Tier-1 carbon maps \citep[e.g.][]{phelps_agricultural_2013,cattaneo_international_2010}.
@@
 the 10 km map shown in Figure 3 in ][]{foley_solutions_2011}.
-Improper understanding of complex, coupled human-natural systems could also result from agent-based models that have calibration errors caused by land cover maps.
+Improper understanding of complex, coupled human-natural systems could also result from models that have calibration errors caused by land cover maps.
 
@@
 \vspace{-0.2 cm}
-These findings suggest a number of recommendations for using land cover maps, complementing those suggested in earlier work \citep{verburg_challenges_2011}, and beginning with those related to the selection of base land cover maps.
-First, to minimize error, users should typically prefer maps derived from imagery with resolutions finer than the scale of individual objects in the cover(s) of interest, assuming that they were created with rigorous classification methods accompanied by appropriate error metrics \citep{olofsson_good_2014}, and are thematically appropriate for the intended use \citep{verburg_challenges_2011}.
-Finer resolution not only helps to improve classification accuracy (see preceding section on \emph{Sources of error in cropland maps)}, but can minimize the \emph{aggregation problem}, one of two components of the modifiable area unit problem \citep[MAUP; ][]{openshaw_million_1979}, in which the shape and placement of the non-overlapping units used to extract map values influence the results of analyses of those values \citep{dark_modifiable_2007, marceau_scale_1999}.
-In remote sensing, an image's pixels define the fundamental mapping unit (by summarizing reflectance or radar/lidar sensors), and the pixels' dimensions influence analyses as a function of how closely they match the characteristic shapes and scales of natural features \citep{dark_modifiable_2007}.
-However, if the sensor resolution is fine enough, then mismatches can be minimized--a natural feature's shape can be approximated by aggregating several square pixels--giving the analyst greater ability to minimize errors associated with this aspect of MAUP \citep{dark_modifiable_2007, hay_comparison_2003}.
+Our findings suggest a number of recommendations for using land cover maps, complementing those suggested in earlier work \citep{verburg_challenges_2011}.
+First, to minimize error, users should typically prefer maps derived from imagery with resolutions substantially finer than the scale of individual objects of interest (e.g.
+agricultural fields), assuming that available maps were created with rigorous classification methods accompanied by appropriate error metrics \citep{olofsson_good_2014}, and are thematically appropriate for the intended use \citep{verburg_challenges_2011}.
+Finer resolution not only helps to improve classification accuracy (see \emph{Sources of error in cropland maps)}, but can minimize the \emph{aggregation problem}, one of two fundamental components of the modifiable area unit problem \citep[MAUP; ][]{openshaw_million_1979}, in which the shape and placement of the non-overlapping units used to extract map values influence analyses of those values \citep{dark_modifiable_2007, marceau_scale_1999}.
+In remote sensing, the image's pixels define the fundamental mapping unit, and mismatches between the pixels' dimensions and the characteristic shapes and scales of natural features impact subsequent analysis \citep{dark_modifiable_2007}.
+However, if the sensor resolution is fine enough, such mismatches can be minimized--a natural feature's shape can be approximated by aggregating several square pixels--giving the analyst greater ability to minimize errors associated with this aspect of MAUP \citep{dark_modifiable_2007, hay_comparison_2003}.
 
 Of course, high quality, fine-scaled maps such as SA-LC, which was carefully developed for South Africa, do not exist for many countries.
-In the absence of these, the new 30 m GLOBELAND30 \citep{chen_global_2015} provides global coverage for the periods 2000 and 2010, and may prove useful in smaller regions lacking their own focused maps.
-Users should first conduct their own, more-focussed accuracy assessments to better understand error rates within their region of interest.
-Another option is to make one's own high resolution maps; although harder to do, this is becoming increasingly feasible due to the rapidly expanding spatial and temporal coverage of high resolution imagery \citep[e.g.][]{hand_startup_2015,drusch_sentinel-2:_2012}, growing accessibility of high performance computing \citep[e.g.][]{pekel_high-resolution_2016}, and improvements to classification algorithms, particularly those designed to deal with the high intra-class variability inherent to certain cover types (e.g.
+Development of a new generation of Landsat/Sentinel-based (i.e., 30m resolution or finer) land-cover maps is underway, as exemplified by the new 30 m GLOBELAND30 map.
+These maps will likely prove very useful for countries and regions lacking their own focused maps, as well as for cross-border analyses.
+To assess whether such maps are fit for the specific purpose, users should first conduct their own, thematically-focussed accuracy assessments to better understand error rates within their region of interest.
+%Another option is to make one's own high resolution maps; although harder to do, this is becoming increasingly feasible due to the rapidly expanding spatial and temporal coverage of high resolution imagery \citep[e.g.][]{hand_startup_2015,drusch_sentinel-2:_2012}, growing accessibility of high performance computing \citep[e.g.][]{pekel_high-resolution_2016}, and improvements to classification algorithms, particularly those designed to deal with the high intra-class variability inherent to certain cover types (e.g.
 cropland) and may be exacerbated by greater image resolution \citep[e.g.][]{debats_generalized_2016,sweeney_mapping_2015}.
 
-If existing high-resolution maps are not available or cannot be readily made, users interested in agricultural land cover may choose to use a new fusion map, such as GeoWiki and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}.
+If high quality, high-resolution maps are not available, users interested in agricultural cover could select a new fusion map, such as GeoWiki and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}.
 Although these maps are relatively coarse-scaled, the fusion methodology helps to greatly minimize error, as does the use of fractional cover values.
@@
 However, users should be aware that this adjustment technique introduces confounding errors where census-reported statistics are unreliable, such as in some SSA countries \citep{carletto_emperor_2013,carletto_guesstimates_2015}.
+Alternatively, users could directly apply the fusion methodology \citep{fritz_cropland_2011} to create their own improved maps.
 
-If a particularly erroneous map is all that is available, and the user is chiefly interested in minimizing cell-wise error and less concerned about the spatial configuration of cover types, then this may be achieved by aggregating maps expressing continuous values (e.g.
+If a particularly erroneous map is all that is available, and the user is chiefly interested in minimizing cell-wise error and less concerned about the spatial configuration of cover, then this may be achieved by aggregating maps expressing continuous values (e.g.
 fractional cover) to coarser scales.
-This approach must be undertaken with care, as its poses a number of complications related to the \emph{scale problem}, the other half of the MAUP \citep{openshaw_million_1979}.
-These include progressive decline in variance with increasing scale \citep[even if means remain constant;][]{dark_modifiable_2007}, and reduced efficiency in estimating regression parameters from coarser map values \citep[][]{avelino_goldilocks_2016}.
+This approach must be undertaken with care, as its poses a number of complications related to the \emph{scale problem}, the other half of MAUP \citep{openshaw_million_1979}, which include progressive declines in variance with increasing scale \citep[even if means remain constant;][]{dark_modifiable_2007}, and reduced efficiency in estimating regression parameters from coarser map values \citep[][]{avelino_goldilocks_2016}.
 Nevertheless, aggregating continuous variables can reduce pixel errors without biasing some statistical properties (e.g.
 total cropland area remains constant using the aggregation approach demonstrated here; see Results: Cropland map errors), and has been shown to reduce other MAUP-related analytical problems \citep{avelino_goldilocks_2016}.
-However, the user must ensure that the scale of aggregation is appropriate for the analysis at hand.
+However, the user must ensure that the scale of aggregation is appropriate for the particular analysis.
 
@@
 In other cases, downstream maps may lack quantified confidence intervals \citep[e.g.][]{monfreda_farming_2008}.
-Although such maps may provide guidelines for appropriate usage\footnote{www.earthstat.org\/wp-content\/uploads\/METADATA\_HarvestedAreaYield175Crops.pdf}, which our analyses help to further illustrate, users should undertake their own two-step error assessments.
-The most straightforward way may be to use a Monte Carlo approach that generates artificial datasets \citep[e.g.][]{avelino_goldilocks_2016} with introduced error, drawn from either reported or user-determined accuracy statistics, for both the base land cover and the downstream maps.
-Quantifying error propagation in this (or another appropriate) manner is important to understanding how map error may influence subsequent understanding of the phenomenon of interest.
+Although such maps may provide guidelines for appropriate usage\footnote{www.earthstat.org\/wp-content\/uploads\/METADATA\_HarvestedAreaYield175Crops.pdf}, which our analyses help to further illustrate, users should undertake their own error propagation assessments.
+The most straightforward way may be to use a Monte Carlo approach that generates artificial datasets \citep[e.g.][]{avelino_goldilocks_2016}, using introduced errors drawn from reported or user-determined accuracy statistics, for both the base land cover and the downstream maps.
+Quantifying error propagation is important to understanding how map error may influence subsequent understanding of the phenomenon of interest.
 
@@
 This work was supported by funds from the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (EAR-1 534544, SES-1360463, and BCS-1026776).
+We  thank two anonymous reviewers for their constructive and helpful comments on earlier manuscript versions.
 %\end{acknowledgments}

### 2017-03-26 paper/main.tex
Final edits (pending any last minute co-author comments)

@@
 The average locational error was 6 km (600\%).
-These findings provide deeper insight into the causes and consequences of land cover map error, including their potential broader regional and policy consequences, and suggest several recommendations for land cover map users.
-%
-%
-%These maps can have substantial error that impact , which may impact our understanding of environmental change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions.
-We used a high quality, national-scale cropland map to assess bias and accuracy within the cropland classes of current generation land cover datasets and four examples of ``downstream'' (land cover-dependent) analyses, two related to physical processes (estimates of vegetative carbon stocks and evapotranspiration) and two to socio-economic processes (gridded crop production estimates and agent-based simulation of household food security), at resolutions ranging from 1-100 km.
-We found that cropland maps have substantial errors below 25 km resolution, with substantial biases and inaccuracies, particularly from maps derived from coarse resolution sensors.
-In some cases these error metrics remain high even when maps are aggregated to 100 km.
-These errors can be substantially amplified in downstream analyses (e.g.
-the carbon and crop production analyses), but in cases where the values of different cover types that underpin the analysis are similar (e.g.
-evapotranspiration estimates) the results can be relatively insensitive.
-To avoid these errors, and thus minimize the risk of misunderstanding global change processes or misinforming policy, substantial map aggregation of land cover maps from their base resolution is often needed.
-NEW TEXT HERE.
-}
+These findings provide deeper insight into the causes and consequences of land cover map error, including their potential broader regional and policy consequences, and suggest several recommendations for land cover map users.}
 \end{abstract}
@@
 Unfortunately, the evidence indicates that this house has shaky foundations \citep[e.g.][]{fritz_highlighting_2011}.
-The reason for this is that land cover data can only practically be derived from satellite imaging, which has several important constraints that propagate mapping errors.
-First, in many regions the spatial arrangement of cover types is smaller than the sensor resolution \citep[e.g.
-smallholders' fields][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015,fritz_identifying_2008}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
+The reason is that land cover data can only practically be derived from satellite imaging, which has several constraints that propagate mapping errors.
+First, in many regions the characteristic scales of cover features are smaller than the sensor resolution \citep[e.g.
+smallholders' fields][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the covers of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015,fritz_identifying_2008}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}.
 Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties \citep[e.g.
@@
 Furthermore, class definitions often vary between maps, complicating inter-comparison \citep{kuemmerle_challenges_2013,fritz_identifying_2008}.
-Third, land cover maps are often used to detect changes \citep[e.g.][]{gross_monitoring_2013}, but seasonal variability and land cover changes can be easily confused.
-Given these multiple sources of error, land cover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \citep{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011,fritz_need_2013}.
+Third, land cover maps are often used to detect changes \citep[e.g.][]{gross_monitoring_2013}, but seasonal variability and cover changes can be easily confused.
+Given these multiple sources of error, land cover maps are often inaccurate and disagree widely between products, particularly in the world's most rapidly developing regions \citep{fritz_comparison_2010,fritz_cropland_2011,fritz_need_2013}.
 These errors limit our ability to obtain granular, mechanistic understanding of processes related to global change.
@@
 Likewise, the importance of assessing the accuracy of land cover maps is increasingly recognized, and there are well-developed, best-practice guidelines for gathering and using ground-truth samples to robustly quantify map error \citep{foody_status_2002,olofsson_making_2013,olofsson_good_2014,stehman_global_2012}.
-Because comprehensive, spatially representative ground truth data are typically unavailable for rapidly changing regions \citep{see_improved_2015,kuemmerle_challenges_2013}, what remains an open question is exactly how much the maps researchers typically use deviate from actual land cover, how this affects our analyses based on these maps, and how this in turn impacts our understanding of environmental change processes.
+Because comprehensive, spatially representative ground truth data are typically unavailable for rapidly changing regions \citep{see_improved_2015,kuemmerle_challenges_2013}, what remains an open question is exactly how much the maps researchers typically use deviate from actual land cover, how this affects analyses based on these maps, and how this in turn impacts our understanding of processes being studied.
 Our current understanding of map accuracy over such areas is often based on scarce information or top-down "sanity checks" made in comparison to aggregated survey data \citep{yu_meta-discoveries_2014,larsen_taken_2015}.
 
-Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
+Since it is difficult to fully quantify map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}.
 Although previous studies have examined how map errors propagate, these are primarily assessed using either simulated errors, relative differences in existing land cover maps, or ground validation data covering relatively small areas \citep[e.g.][]{ge_impacts_2007, linard_assessing_2010, quaife_impact_2008, tuanmu_global_2014, schmit_limitations_2006}.
@@
 Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g.
-3DRobotics, DJIA} to new high resolution satellite arrays \citep{drusch_sentinel-2:_2012,hand_startup_2015} and better mapping methods \citep{fritz_geo-wiki:_2012,estes_projected_2013,debats_generalized_2016}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the land cover products that have become commonplace in global change research--and which are often used to make policy decisions \citep{searchinger_high_2015}.
+3DRobotics, DJIA} to new high resolution satellite arrays \citep{drusch_sentinel-2:_2012,hand_startup_2015} and better mapping methods \citep{fritz_geo-wiki:_2012,estes_projected_2013,debats_generalized_2016}, are providing means to more comprehensively interrogate the accuracy and biases in land cover products that have become commonplace in global change research, and which often underpin policy decisions \citep{searchinger_high_2015}.
 
@@
 \vspace{-0.2 cm}
-In the late 2000s, the South African government commissioned a cropland map that was made by manually interpreting and digitizing fields visible within high resolution satellite imagery \citep{fourie_better_2009}.
-The resulting vectorized field boundaries provide unique, highly accurate data on field sizes and distribution for 2009-2011.
-The accuracy of this dataset was evaluated in a previous study \citep{estes_platform_2016}, in which visual assessment of cropland presence/absence was made within 15,225 individual 4 ha plots (25 sub-plots within 609 1 km$^2$ grids) placed over high resolution satellite imagery, and compared to coverage by the vector boundaries.
-Measured in terms of its ability to distinguish crop fields from other cover types, the results showed these data to be 97\% percent accurate, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\% \citep[see SI for more description, and][]{estes_platform_2016}.
+In the late 2000s, the South African government commissioned a cropland map that was made by manually interpreting and digitizing fields in high resolution satellite imagery \citep{fourie_better_2009}.
+The resulting vectorized field boundaries provide unique, highly accurate data on field sizes and distribution for the 2009-2011 period.
+A previous study evaluated the accuracy of this dataset \citep{estes_platform_2016}, using a visual assessment of cropland presence/absence in high resolution satellite images within 15,225 individual 4 ha plots (25 sub-plots within 609 1 km$^2$ grids) to evaluate the ability of the vector boundaries to distinguish crop fields from other cover types.
+The results showed these data to be 97\% percent accurate, with user's accuracies of 94\% and 98\% for the cropland and non-cropland classes, and producer's accuracies of 84 and 99\% \citep[see SI for more description, and][]{estes_platform_2016}.
 
-We used these vector data as a reference for evaluating four land cover products representative of the type commonly used in global change studies and related areas of research.
-The first was South Africa's own 30 m resolution 2009 National land cover map \citep[SA-LC][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are created for individual countries \citep[e.g.][]{fry_completion_2009}.
-Although global-scale, Landsat-derived maps have recently become available \citep{chen_global_2015}, their reported accuracy for cultivated areas is lower (80-85\% user's accuracy) than that achieved by more intensive, national to sub-national products \citep[e.g.
+We used these vector data as a reference for evaluating four land cover products representative of those commonly used in global change studies and related areas of research.
+The first was South Africa's 30 m resolution 2009 National land cover map \citep[SA-LC][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are created for individual countries \citep[e.g.][]{fry_completion_2009}.
+Although global-scale, Landsat-derived maps have recently become available \citep{chen_global_2015}, their reported accuracy for cultivated areas is lower (80-85\% user's accuracy) than those of more intensive, national to sub-national products \citep[e.g.
 90\% user's accuracy][]{sweeney_mapping_2015}.
@@
 This ``hybrid-fusion'' approach represents the state of the art in mapping agricultural land cover.
-Since GeoWiki provides a continuous bounded value of land cover (cropland percentage), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.
+Since GeoWiki provides a continuous value of land cover (cropland percentage), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded cropland percentages.
 
@@
 
-To develop percent coverage from the cropland reference map, we intersected the original field boundary vector maps with a 1 km grid, and calculated the percent of each cell occupied by fields.
-In the resulting grid, we masked out areas classified as communal farmland (18.7\% of mapped croplands in the reference data), because only their outer perimeters were digitized \citep{fourie_better_2009}, which risked overestimating cropland extent because the vectors also enclose uncropped areas between adjacent smallholder fields.
-We also excluded from the analysis areas covered by permanent tree crops, sugarcane plantations, and commercially afforested areas, because these three classes were not common to all five cropland datasets.
-To remove permanent tree crops (3.1\% of cropland area), we masked out reference vectors labelled as such, and to exclude the other two cover types, which are not included in the reference data, we relied on information from two other datasets.
+To convert the reference data to percentages, we intersected the field boundary vectors with a 1 km grid and calculated the percent of each cell covered by fields.
+In the resulting grid, we masked out areas classified as communal farmland (18.7\% of cropland area in the reference data), because only their outer perimeters were digitized \citep{fourie_better_2009}, which risked overestimating cropland extent because those vectors enclosed uncropped inter-field areas.
+We also excluded areas under permanent tree crops, sugarcane plantations, and commercially afforestation, because these classes were not common to all five cropland datasets.
+To remove permanent tree crops (3.1\% of cropland area), we masked out reference vectors labeled as such, and to exclude the other two cover types, which are not included in the reference data, we relied on information from two other datasets.
 We used a 20 m resolution landcover map of KwaZulu-Natal to mask out the primarily coastal sugarcane farms \citep[93-100\% user's and 76-98\% producer's accuracy for sugarcane classes;][]{geoterraimage_2011_2013}, and used the SA-LC dataset to filter out areas of commercial afforestation, which are mainly located in South Africa's montane areas and do not overlap with arable croplands.
@@
 
-%Care in aggregation methods worth emphasizing \citep{verburg_challenges_2011,dendoncker_exploring_2008,raj_analysing_2013,moody_influence_1995}, can lead to bias, which is influenced by spatial arrangement of cover types.
-
-%[maybe] For fine features below sensor resolution, convert to quantitative measures such as SMA (tough for cropland though, but maybe with MESMA for different crop categories).
-For homogenous cover types with little within-class variability.
-
-%Marceau et al citing previous work that finer sensors will be needed to map crop field locations. Waldner et al (2016), gives field size requirements mentioned by GeoGlam
-
-%Match dataset to application; merge complementary sources where available \citep{verburg_challenges_2011}.
-
-%Our analysis demonstrates that the land cover data that increasingly inform much global change research and policy can be substantially misleading, depending on the dataset selected, its application case, the scale to which it is aggregated, and whether the insight being drawn from the map depends on high accuracy or low bias.
-
-%Synthetic assessments can help you determine how much impact error can make.
-
-%Carefully evaluate a product for its sensitivity for a particular application (Fritz et al, 2011, highlighting_2011).
-Restricted geographically, so limited inference, but some general inferences may be drawn.
-Carefully crafted Landsat products for particular scale (higher density of validation points, local knowledge).
-
-%Of course doing this loses one of the main sources of geographic information (configuration), but this is one of the many aspects that leads to the MAUP problem in any case.
-
-%Reasons for not evaluating GLC30 (lower accuracy). Other landcover products may be better, and we didnÕt evaluate these.
-
-%The resolution of the sensor relative to the typical scale of features in a landscape affects classification accuracy, and this relationship varies between features \citep{moody_influence_1995,marceau_scale_1999, marceau_scale_1999}.
-For agricultural fields, coarser resolution sensors can produce reasonably accurate estimates of total agricultural error, but are inaccurate at mapping the locations of fields \citep{pax-lenney_effect_1997,ozdogan_resolution_2006}
-
-%The 250 m resolution update of the GeoWiki map  \citep{waldner_unified_2016}, which we did not evaluate here, may help minimize errors associated with the former, but will still be exposed to the latter.
-
-%Caveats, including one related to ABM errors here?
-
-%And where background veg is similar, cite Sweeney et al, 2015, and show that similar finding in Frey and Smith (2006) for high latitude savannas and croplands. 
-%Accuracy is lowest where cropland cover is intermediate.
-
-%To a certain extent reflects configuration of landscape elements.
-
-%Our analysis allows us to assessment allows us to thoroughly examine and quantify the spatial patterning of errors
-
-%A related example can be seen in high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \citep[e.g.
-Figure 3 in][]{foley_solutions_2011}.
-Such maps may give an inflated sense of the precision with which such locations can be identified.
-Although the study in this example is careful to report aggregated statistics and maps the per hectare values, and the map-makers stress that their products should not be used at the grid cell scale\footnote{www.earthstat.org\/wp-content\/uploads\/METADATA\_HarvestedAreaYield175Crops.pdf}, it is not inconceivable that users might make decisions based on their visual interpretations of maps.
-For this reason, it is advisable to aggregate presented maps until their accuracy reaches an acceptable standard (Table 1).
-
-%\subsection*{Caveats}
-
-%Properties revealed here in this study are relevant.  Mosaic classes.  Background vegetation similar to cropland (savanna, grassland).
-
-%With the exception of the ABM-related example, these findings apply to error at the pixel level. We do not explicitly evaluate impacts on aggregate measures (such as total production).
-
-%The errors in the cropland data we tested were large, and it is unlikely that current ABM-based studies, mostly of small regions, are based on such inaccurate land cover data.
-However, as computing power increases, so, too, does the feasible application extent of ABMs, which will then become more reliant on larger-scaled, more erroneous, land cover data, with the attendant risk that insights into socio-ecological systems arising from these models could be highly misleading.
-
-%This spatially comprehensive, bottom-up land cover map assessment provides new insight into the extent and causes of map error, and its consequences for understanding global change.
-
-%The analysis was made possible by a high accuracy cropland map that likely provides the truest measure of cropped area and distribution currently available for this region.
-Although this reference map is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission, SI), while temporal mismatches between the reference and test maps may account for some of the error we identified, our assessment (SI) suggests that such discrepancies do not appreciably impact our findings.
-Our results are also supported by previous, continental-scale assessments that revealed substantial errors and inconsistencies between land cover maps \citep{fritz_comparison_2010,gross_monitoring_2013}.
-These similarities, combined with the larger extents of these studies, also suggest that our findings may be applicable beyond South Africa's borders.
-
 \bibliographystyle{gcbnourl} 
@@
 
-%\begin{figure}
-%\centerline{\includegraphics[width=.4\textwidth]{figsamp.eps}}
-%\caption{LKB1 phosphorylates Thr-172 of AMPK$\alpha$ \textit{in vitro}
-%and activates its kinase activity.}\label{afoto}
-%\end{figure}
-%
-%\begin{figure*}[ht]
-%\begin{center}
-%\centerline{\includegraphics[width=.7\textwidth]{figsamp.eps}}
-%\caption{LKB1 phosphorylates Thr-172 of AMPK$\alpha$ \textit{in vitro}
-%and activates its kinase activity.}\label{afoto2}
-%\end{center}
-%\end{figure*}
-%
-%\begin{table}[h]
-%\caption{Repeat length of longer allele by age of onset class.
-%This is what happens when the text continues.}
-%\begin{tabular}{@{\vrule height 10.5pt depth4pt  width0pt}lrcccc}
-%&\multicolumn5c{Repeat length}\\
-%\noalign{\vskip-11pt}
-%Age of onset,\\
-%\cline{2-6}
-%\vrule depth 6pt width 0pt years&\multicolumn1c{\it n}&Mean&SD&Range&Median\\
-%\hline
-%Juvenile, 2$-$20&40&60.15& 9.32&43$-$86&60\\
-%Typical, 21$-$50&377&45.72&2.97&40$-$58&45\\
-%Late, $>$50&26&41.85&1.56&40$-$45&42\tablenote{The no. of wells for all samples was 384. Genotypes were
-%determined by mass spectrometric assay. The $m_t$ value indicates the
-%average number of wells positive for the over represented allele.}
-%\\
-%\hline
-%\end{tabular}
-%\end{table}
-%
-%
-%\begin{table*}[ht]
-%\caption{Summary of the experimental results}
-%\begin{tabular*}{\hsize}
-%{@{\extracolsep{\fill}}rrrrrrrrrrrrr}
-%\multicolumn{3}{l}{Parameters}&
-%\multicolumn{5}{c}{Averaged Results}&
-%\multicolumn{5}{c}{Comparisons}\cr
-%\hline
-%\multicolumn1c{$n$}&\multicolumn1c{$S^*_{MAX}$}&
-%\multicolumn1c{$t_1$}&\multicolumn1c{\ $r_1$}&
-%\multicolumn1c{\ $m_1$}&\multicolumn1c{$t_2$}&
-%\multicolumn1c{$r_2$}&\multicolumn1c{$m_2$}
-%&\multicolumn1c{$t_{lb}$}&\multicolumn1c{\ \ $t_1/t_2$}&
-%$r_1/r_2$&$m_1/m_2$&
-%$t_1/t_{lb}$\cr
-%\hline
-%10\tablenote{Stanford Synchrotron Radiation Laboratory (Stanford University,
-%Stanford, CA)}&1\quad &4&.0007&4&4&.0020&4&4&1.000&.333&1.000&1.000\cr
-%10\tablenote{$R_{\rm FREE}=R$ factor for the $\sim 5$\% of the randomly
-%chosen unique ref\/lections not used in the ref\/inement.}&5\quad &50&.0008&8&50&.0020&12&49&.999&.417&.698&1.020\cr
-%100\tablenote{Calculated for all observed data}&20\quad &2840975&.0423&95&2871117&.1083&521&---&
-%.990&.390&.182&---\ \ \cr
-%\hline
-%\end{tabular*}
-%\end{table*}
-
 \end{document}

### 2017-03-26 Rmd/code.Rmd
Update and re-compile code book for supplementals

@@
 sumfun <- function(x) sum(x, na.rm = TRUE)
+# gti <- lapply(list(cover2007, cover2011), function(y) {
+#   r <- calc(y[[-6]], sumfun)
+#   r[r > 100] <- 100
+#   writeRaster(r, file = gsub("\\.tif", "sum.tif", y@file@name), overwrite = TRUE)
+# }) 
 gti <- lapply(list(cover2007, cover2011), function(y) {
-  r <- calc(y[[-6]], sumfun)
+  raster(gsub("\\.tif", "sum.tif", y@file@name))
+})
+
+# Leave commented out in case want to bring back horticulture class
+gti_hort <- lapply(list(cover2007, cover2011), function(y) {
+  r <- calc(y, sumfun)
   r[r > 100] <- 100
-  writeRaster(r, file = gsub("\\.tif", "sum.tif", y@file@name), overwrite = TRUE)
-})
-
-# Leave commented out in case want to bring back horticulture class
-# gti_hort <- lapply(list(cover2007, cover2011), function(y) {
-#   r <- calc(y, sumfun)
-#   r[r > 100] <- 100
   # writeRaster(r, file = gsub("\\.tif", "sum_h.tif", y@file@name), overwrite = TRUE)
-# })
+  r
+})
+
+# calculate areas
+cellStats(cover2011[[2]] / 100, sum) / 
+  cellStats(calc(cover2011 / 100, sumfun), sum) * 100  # communal areas 18.7\% cropland
+cellStats(cover2011[[6]] / 100, sum) / 
+  cellStats(calc(cover2011 / 100, sumfun), sum) * 100  # horticulture 3.1\% of cropland
 
@@
 # mask out all joint NAs
-gti <- lapply(gti, function(x) mask(x, namask))
-lclist <- lapply(lclist, function(x) mask(x, namask))
-
+gti <- lapply(gti, function(x) raster::mask(x, namask))
+lclist <- lapply(lclist, function(x) raster::mask(x, namask))
+cropl <- c(gti$g2011, lclist[c("sa30", "globmu", "modmu", "geow")])
+names(cropl)[1] <- "gti"
+save(cropl, file = "external/ext_data/cropland-1km.rda")
 ```
@@
 
+save(lc_agg, gti_agg, area_wgts, file = full_path(p_data, "agg_cropland.rda"))
 save(dlist_act, file = full_path(p_data, "d_grid_act.rda"))
@@
 ```{r, eval = FALSE}
+fact <- c(5, 10, 25, 50, 100)
 paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
@@
 names(gti) <- c("g2007", "g2011")
-gti <- lapply(gti, function(x) mask(x, namask))  # apply namask
+gti <- lapply(gti, function(x) raster::mask(x, namask))  # apply namask
 snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
 
+gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters
+gti_agg11 <- lapply(gti_agg, function(x) x$g2011)
+
 # recreate cropland percentages
-lc2007 <- lapply(snms, function(x) gti$g2007 - dlist_act[[x]]$f1$g2007)
-lc2011 <- lapply(snms, function(x) gti$g2011 - dlist_act[[x]]$f1$g2011)
+# lc2007 <- lapply(snms, function(x) gti$g2007 - dlist_act[[x]]$f1$g2007)
+# lc2011 <- lapply(snms, function(x) gti$g2011 - dlist_act[[x]]$f1$g2011)
 # plot(lc2007[[2]] - lc2011[[2]])  # same thing, as it should be
@@
 # calculate km2 cropland in LC datasets
-lckm2 <- sapply(lc2011, function(x) cellStats(x / 100, sum))
-refkm2 <- sapply(gti, function(x) cellStats(x / 100, sum))
-refkm2[2] / refkm2[1]  # 3.6 percent more in 2011 than 2007
-
-# bias in total cropland relative to GTI
-stats2007 <- round((refkm2[2] - lckm2) / refkm2[2] * 100, 1)
-stats2011 <- round((refkm2[1] - lckm2) / refkm2[1] * 100, 1)
+awgts <- lapply(area_wgts, function(x) x[[1]])
+lckm2 <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
+  gtia <- cellStats(gti_agg11[[x]] / 100 * awgts[[x]], sum)
+  lca <- sapply(snms, function(y) {
+    cellStats(lc_agg[[x]][[y]] / 100 * awgts[[x]], sum)
+  })
+  o <- c(gtia, lca)
+}))
+colnames(lckm2)[1] <- "gti"
+rownames(lckm2) <- lev
+
+bias_statsw
+
+# mean (stays about the same)
+lckmmu <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
+  gtia <- cellStats(gti_agg11[[x]] / 100, mean)
+  lca <- sapply(snms, function(y) {
+    cellStats(lc_agg[[x]][[y]] / 100, mean)
+  })
+  o <- c(gtia, lca)
+}))
+colnames(lckmmu)[1] <- "gti"
+rownames(lckmmu) <- lev
+
+# variance declines with scale, as expected
+lckmvar <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
+  gtia <- cellStats(gti_agg11[[x]] / 100, var)
+  lca <- sapply(snms, function(y) {
+    cellStats(lc_agg[[x]][[y]] / 100, var)
+  })
+  o <- c(gtia, lca)
+}))
+colnames(lckmvar)[1] <- "gti"
+rownames(lckmvar) <- lev
+
+# lckm2 <- sapply(lc_agg$f1[snms], function(x) cellStats(x / 100, sum))
+# lckm2 <- sapply(lc2011, function(x) cellStats(x / 100, sum))
+# refkm2 <- sapply(gti, function(x) cellStats(x / 100, sum))
+
+# area of each GTI dataset
+refkm2 <- sapply(gti_agg$f1, function(x) cellStats(x / 100, sum))
+refkm2[2] / refkm2[1]  # 3.7 percent more in 2011 than 2007
+
+# bias in total cropland relative to GTI, at 1 km scale
+stats2007 <- round((refkm2[2] - lckm2[1, ]) / refkm2[2] * 100, 1)
+stats2011 <- round((refkm2[1] - lckm2[2, ]) / refkm2[1] * 100, 1)
 
@@
 # sapply(gti, function(x) cellStats(!is.na(x) , sum))
-totkm2 <- cellStats(!is.na(gti$g2011) , sum)
+
+totkm2 <- cellStats(!is.na(gti_agg$f1$g2011), sum)
 save(stats2007, stats2011, totkm2, refkm2, 
@@
 
+Cropland area by dataset and scale
+```{r, echo = FALSE}
+round(lckm2)
+```
+
 ## Cropland bias/MAE plots
@@
   # msk <- gti_agg[[x]]$gti > 0.05  # farmland > 0.05% mask
-  msked <- mask(gti_c[[x]] * tareas[[x]], msk, maskvalue = 0)
+  msked <- raster::mask(gti_c[[x]] * tareas[[x]], msk, maskvalue = 0)
   cellStats(msked, sum)
@@
     # msk <- jj[[y]] > 0  # farmland > 0.05% mask
-    msked <- mask(xx[[y]] * tareas[[y]], msk, maskvalue = 0)
+    msked <- raster::mask(xx[[y]] * tareas[[y]], msk, maskvalue = 0)
     cellStats(msked, sum)
@@
                   totc_out)
-caption <- paste("Percent differences in total carbon stock estimates", 
+capt <- paste("Percent differences in total carbon stock estimates", 
               "calculated from the reference maps and from each of the four",
@@
               "(cover types indicated by column names).")
-totc_xtab <- xtable(totc_out, caption = caption, digits = 1)
+totc_xtab <- xtable(totc_out, caption = capt, digits = 1)
 print(totc_xtab, type = "latex", caption.placement = "top",
@@
 }
-
+save(cb_statsw_mua, cb_statsw_mu, file = fp(p_data, "carbon-werr.rda"))
 ````
@@
 save(list = ls(), file = full_path(p_data, "yield-bias.rda"))
+save(out_tab, file = fp(p_data, "yield-prod-accbias.rda"))
+save(cyld_agg, cpagg1, file = full_path(p_data, "yieldprod-1km.rda"))
+
 ```
@@
 
+# Impact of cropland errors on location
+
+Evaluated in terms of how landcover map error impacts the ability to locate areas having a particular quantity of interest.
+
+## Data
+
+Derived carbon (section 6) and yield estimates (section 8).
+
+```{r, eval=FALSE}
+library(SAcropland)
+library(xtable)
+library(RColorBrewer)
+library(spatstat)
+p_root <- full_path(proj_root("SAcropland"), "SAcropland")
+p_fig <- full_path(p_root, "paper/figures/")
+p_data <- full_path(p_root, "external/ext_data/")
+p_data2 <- full_path(p_root, "data/")
+#p_carb <- full_path(p_root, "SAcropland/external/ext_data/carbon/")
+# load(full_path(p_data, "yield-bias.rda"))
+
+# SA provinces
+prov <- readOGR("external/ext_data/provinces.sqlite", layer = "provinces")
+prov <- prov[prov$id_1 != 9, ]  # remove Princeton Edward Islands
+prov[prov$id_1 == 10, "id_1"] <- 9 # reset WC to id 10
+load("external/ext_data/MZshapes.Rdata")
+
+# yield and carbon data
+load("external/ext_data/yieldprod-1km.rda")
+load("external/ext_data/carbon-tier1.rda")
+load("external/ext_data/cropland-1km.rda")
+```
+
+## Differences in location
+
+If you are interested in selecting areas of highest yield/production/carbon potential, how far off are you relative to "truth".
+Use top quintile/decline to assess this, and then calculate mean nearest neighbor distance between two pixels of each map.
+
+### Identify high value areas
+```{r, eval = FALSE}
+# top decile of...
+# ...cropland
+crclass <- lapply(cropl, function(x) {  # x <- cropl$sa30
+  r <- x * (x > 0)
+  r[r == 0] <- NA
+  rq <- quantile(r, c(0, 0.9, 1))
+  types <- cut(r, breaks = rq)
+  types > 1
+})
+
+# ...yield
+yclass <- lapply(cyld_agg$f1, function(x) {  # x <- cyld_agg$f1[[1]]
+  r <- x * (x > 0)
+  r[r == 0] <- NA
+  rq <- quantile(r, c(0, 0.9, 1))
+  types <- cut(r, breaks = rq)
+  types > 1
+})
+
+# ...production
+pclass <- lapply(cpagg1$f1, function(x) {  # x <- cyld_agg$f1[[1]]
+  r <- x * (x > 0)
+  r[r == 0] <- NA
+  rq <- quantile(r, c(0, 0.9, 1))
+  types <- cut(r, breaks = rq)
+  types > 1
+})
+
+# ...carbon
+cdat <- c(gti_c$f1, lc_c$f1)
+cdatf <- lapply(cdat, function(x) {  # mask out non-cropland areas and take mu C
+  msk <- x$forest < 99.6
+  cmu <- calc(x, mean)
+  raster::mask(cmu, msk, maskvalue = 0)
+})
+cclass <- lapply(cdatf, function(x) {  # x <- cyld_agg$f1[[1]]
+  r <- x * (x > 0)
+  r[r == 0] <- NA
+  rq <- quantile(r, c(0, 0.9, 1))
+  types <- cut(r, breaks = rq)
+  types > 1
+})
+names(cclass) <- names(pclass)
+
+# carbon efficiencies (C / yield)
+effclass <- lapply(1:5, function(x) { # x <- 1
+  cyld <- cyld_agg$f1[[x]]
+  cyld[cyld == 0] <- NA
+  cdatfc <- crop(cdatf[[x]], cyld)
+  eff <- cdatfc / cyld  # carbon efficiency
+  # eff[is.infinite(eff)] <- NA
+  rq <- quantile(eff, c(0, 0.9, 1))
+  types <- cut(eff, breaks = rq)
+  types > 1
+})
+names(effclass) <- names(cclass)
+
+```
+
+### Calculate nearest-neighbor distances
+
+Between GTI high value areas and each of the four LC-derived datasets. 
+```{r, eval = FALSE}
+
+# calculate distance between GTI and each other dataset's classes. 
+# pms2a <- expand.grid(rep(list(0:1), 5))[-1, ] 
+# pms2b <- pms2a[which(rowSums(pms2a) == 2 & pms2a$Var1 == 1), ]
+# pms2b$cl <- c("YC1", "YC2", "YC3", "YC4")
+# rownames(pms2b) <- 1:4
+
+# Define neighborhood window
+sa_owin <- owin(xrange = bbox(sa.shp)[1, ], yrange = bbox(sa.shp)[2, ],
+                unitname = "meter")
+# which1 <- function(x) x == 1  # selector function for r to xy
+
+mudists <- lapply(list(crclass, yclass, pclass, cclass, effclass), function(x) {
+  # x <- cclass
+  mudist <- sapply(2:5, function(y) {  # y <- 2
+    print(paste0("...", y))
+    a <- rasterToPoints(x$gti, fun = function(x) x == 1)
+    b <- rasterToPoints(x[[y]], fun = function(x) x == 1)
+    pps <- lapply(list(a, b), function(z) {
+      ppp(x = z[, 1], y = z[, 2], window = sa_owin)
+    })
+    names(pps) <- c("gti", names(x)[y])
+    reord <- sort(sapply(pps, function(i) i$n), decreasing = TRUE)
+    ppnn <- nncross(pps[[names(reord)[1]]], pps[[names(reord)[2]]])
+    mean(ppnn$dist) / 1000
+  })
+  names(mudist) <- names(x)[2:5]
+  mudist
+})  
+names(mudists) <- c("cropland", "yield", "prod", "carb", "eff")
+
+# average km off
+mudvec <- unlist(mudists)
+median(mudvec[!like(names(mudvec), "sa30")])
+# mean(mudvec)
+
+# plot(stack(yclass))
+# plot(is.na(yclass[[1]]) - is.na(yclass[[2]]))
+# plot(is.na(cyld_agg$f1$gti) - is.na(cyld_agg$f1$geow))
+
+# output plot
+cols <- c("red", "orange3", "green4", "blue")
+pcol <- makeTransparent(cols, alpha = c(120))
+pcol2 <- makeTransparent(cols, alpha = c(30))
+
+plot(1:4, pch = 20, col = pcol)
+bs <- do.call(rbind, mudists[c(1, 2, 4)])
+lcnms <- c("SA LC", "GlobCover", "MODIS", "Geowiki")
+
+pdf(full_path(p_fig, "location_offset.pdf"), height = 3.5, width = 4, 
+    bg = "transparent")
+par(mar = c(2, 3.5, 1, 0), bg = "transparent")
+bp <- barplot(bs, beside = TRUE, col = ggplot2:::interleave(cols, pcol, pcol2), 
+              las = 2, border = "transparent", xaxt = "n", ylab = "", 
+              ylim = c(0, 18), yaxt = "n")
+axis(1, at = colMeans(bp), labels = lcnms, lty = 0, cex.axis = 0.9, 
+     mgp = c(2, 0.25, 0))
+axis(2, at = seq(0, 18, 2), labels = seq(0, 18, 2), las = 2)
+mtext("km", side = 2, line = 2.25)
+legend(x = 12.5, y = 18, legend = rep("", 4), pch = 15, col = cols, adj = 0,
+       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0, horiz = TRUE)
+legend(x = 12.5, y = 17.25, legend = rep("", 4), pch = 15, adj = 0, col = pcol, 
+       pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
+legend(x = 12.5, y = 16.5, legend = rep("", 4), pch = 15, adj = 0, col = pcol2, 
+       pt.cex = 1.5, bty = "n", cex = 0.8, horiz = TRUE, x.intersp = 0)
+text(x = 13, y = 17, labels = "Cropland", pos = 2, cex= 0.8)
+text(x = 13, y = 16.25, labels = "Yield", pos = 2, cex = 0.8)
+text(x = 13, y = 15.5, labels = "Carbon", pos = 2, cex= 0.8)
+dev.off()
+
+```
+
 # Miscellaneous additional calculations
 
-```{r, eval=FALSE}
+```{r, message=FALSE}
 library(rgdal)
@@
 library(rgeos)
+library(lmisc)
 
@@
 
-Area of sub-Saharan Africa and South Africa's share of that area
-```{r, eval=FALSE}
-eco <- readOGR(dsn = full_path(p_dat, "africa_ecofloristic_zones.sqlite"), 
+## Area of sub-Saharan Africa and South Africa's share of that area
+```{r}
+eco <- readOGR(dsn = fp(p_dat, "africa_ecofloristic_zones.sqlite"), 
                layer = "africa_ecofloristic_zones")
 eco@proj4string <- CRS(projection(raster(nrow = 1, ncol = 1)))
-af <- readOGR(dsn = full_path(p_dat, "africa_countries_alb.sqlite"), 
+af <- readOGR(dsn = fp(p_dat, "africa_countries_alb.sqlite"), 
                layer = "africa_countries_alb")
-af@data <- af@data[, c(1, 3:4)]
+af@data <- af@data[, c(1, 3:4, 6:7)]
 ecoalb <- spTransform(eco, af@proj4string)
@@
 nosahara <- ecoalb[-sahara, ]
-plot(nosahara, col = "red")
+# plot(nosahara, col = "red")
 
@@
 # plot(ssaeco[ssaeco@data$gez_term == "Tropical mountain system", ][1, ])
-plot(ssa)
-plot(ssaeco, col = "red", add = TRUE)
+# plot(ssa)
+# plot(ssaeco, col = "red", add = TRUE)
 
@@
 
-```{r, eval=FALSE}
+```{r, eval=FALSE, echo = FALSE}
 a <- sample(1:100, size = 20)
@@
 sum(a) / sum(b)
-
-```
+```
+
+## Calculation of error propagation
+```{r}
+load(fp(p_dat, "cropland-werr-2011.rda"))
+load(fp(p_dat, "yield-prod-accbias.rda"))
+load(fp(p_dat, "et-err-bias.rda"))
+load(fp(p_dat, "carbon-werr.rda"))
+
+cland <- bacc2011
+yprod <- as.data.table(out_tab)[Region == "Density"]
+carb <- rbind(cb_statsw_mu, cb_statsw_mua)
+
+# Reshape error stats
+# yield/production 
+yp1 <- melt(yprod, id.vars = c("Map", "Metric", "Variable"), 
+            measure.vars = names(yprod)[5:10])
+yp2 <- dcast(yp1, formula = Metric + Variable + variable ~ Map)
+setcolorder(yp2, neworder = c(1:3, ncol(yp2), 5:6, 4))
+setnames(yp2, names(yp2), c("stnm", "var", "ol", names(cland)[-c(1:2)]))
+
+snms <- names(cland)[-c(1:2)]
+lev <- cland[stnm == "mu", ol]
+
+# Production and yield propagation
+yp_prop <- rbindlist(lapply(c("Yield", "Production"), function(x) {  
+  o <- data.frame(t(sapply(snms, function(y) {  # y <- "sa30"
+    emag <- yp2[var == x & stnm == "MAE", get(y)] / 
+      cland[stnm == "mua", get(y)] 
+    c(emag, mean(emag))
+  })))
+  o <- cbind(snms, round(o, 2))
+  colnames(o) <- c("map", lev, "mu")
+  data.table(cbind.data.frame("var" = x, o))
+}))
+
+# Carbon differences
+c_prop <- rbindlist(lapply(snms, function(x) {  # x <- "sa30"
+  dat <- carb[stnm == "MAE" & map == x, names(carb)[-c(1:3)], with = FALSE]
+  o <- dat[, lapply(.SD, function(y) y / cland[stnm == "mua", get(x)]), 
+           .SDcols = names(dat)]
+  o <- round(rbind(o, data.table(t(colMeans(o)))), 2)
+  o <- cbind("ol" = c(lev, "mu"), o)
+  cbind("map" = x, o)
+  # dat[, second] / cland[stnm == "mua", get(x)]
+}))
+
+# et propagation
+et_prop <- round(et_err[Variable == "Annual Mean", MAE] / 
+  cland[ol == "f25" & stnm == "mua", snms, with = FALSE], 2)
+
+yp_prop
+c_prop
+et_prop
+```

### 2017-04-25 paper/main.tex
Revert back to auth-date citation for GCB

@@
 \usepackage[sort,compress]{natbib}
+%\usepackage[numbers,sort,compress]{natbib}
 %\bibpunct{[}{]}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
@@
 
-\begin{figure*}[!ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\centerline{\includegraphics[width=1\textwidth]{figures/figure1.pdf}}
-\vspace{-0.15 cm}
-\caption{(A) Errors in the percent cropland estimates resulting from each of the four test maps relative to the reference map at different scale of pixel aggregation.
+%\begin{figure*}[!ht]
+%%\begin{wrapfigure}{c}{1\textwidth}
+%\centerline{\includegraphics[width=1\textwidth]{figures/figure1.pdf}}
+%\vspace{-0.15 cm}
+%\caption{(A) Errors in the percent cropland estimates resulting from each of the four test maps relative to the reference map at different scale of pixel aggregation.
 Rows indicate the test map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation.
@@
 Bias estimates are indicated by the semi-transparent bars, accuracy (lower is more accurate) by the solid bars, with bar colors coded to specific cropland maps.}
-\label{afoto1}
-\end{figure*}
+%\label{afoto1}
+%\end{figure*}
 
@@
 
-%\vspace{-0.75 cm}
-%\begin{wrapfigure}{r}{0.5\textwidth}
-\begin{figure}[!h]
-%\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/biases_md_lnorm_gam_mu0.pdf}}
-\centerline{\includegraphics[width=.7\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
-\caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
+%%\vspace{-0.75 cm}
+%%\begin{wrapfigure}{r}{0.5\textwidth}
+%\begin{figure}[!h]
+%%\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/biases_md_lnorm_gam_mu0.pdf}}
+%\centerline{\includegraphics[width=.7\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
+%\caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
 Prediction curves are color-coded to the different test maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}\label{afoto2}
-\end{figure}
-%\end{wrapfigure}
-%\vspace{-0.75 cm}
+%\end{figure}
+%%\end{wrapfigure}
+%%\vspace{-0.75 cm}
 
@@
 
-%\vspace{-0.75 cm}
-\begin{figure}[!h]
-%\begin{wrapfigure}{r}{0.5\textwidth}
-%\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/figure3.pdf}}
-\centerline{\includegraphics[width=0.55\textwidth]{figures/figure3.pdf}}
-\caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map.
+%%\vspace{-0.75 cm}
+%\begin{figure}[!h]
+%%\begin{wrapfigure}{r}{0.5\textwidth}
+%%\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/figure3.pdf}}
+%\centerline{\includegraphics[width=0.55\textwidth]{figures/figure3.pdf}}
+%\caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map.
 Bias estimates (represented by symbols) fall within the semi-transparent floating bars, while accuracies are contained in the solid bars.
@@
 MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.}
-\label{afoto}
-\end{figure}
-%\end{wrapfigure}
+%\label{afoto}
+%\end{figure}
+%%\end{wrapfigure}
 
@@
 All but MODIS underestimated ET by 5-15\% in the northern tip of the country.
-
-%\vspace{-0.75 cm}
-\begin{figure}[h]
-\centerline{\includegraphics[width=.6\textwidth]{figures/et_bias_map.pdf}}
-\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
-\end{figure}
+%
+%%\vspace{-0.75 cm}
+%\begin{figure}[h]
+%\centerline{\includegraphics[width=.6\textwidth]{figures/et_bias_map.pdf}}
+%\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
+%\end{figure}
 
@@
 
-%\vspace{-0.5 cm}
-\begin{figure}[!ht]
-%\vspace{-0.75 cm}
-\centerline{\includegraphics[width=0.55\textwidth]{figures/yield_prod_bias.pdf}}
-\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates.
+%%\vspace{-0.5 cm}
+%\begin{figure}[!ht]
+%%\vspace{-0.75 cm}
+%\centerline{\includegraphics[width=0.55\textwidth]{figures/yield_prod_bias.pdf}}
+%\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates.
 Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.
 Symbols code the different variables (production and yield), normalized to their respective means.}
-\label{afoto}
-\end{figure}
+%\label{afoto}
+%\end{figure}
 
@@
 
-\begin{figure}[ht]
-\centerline{\includegraphics[width=.55\textwidth]{figures/figure6.pdf}}
-\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and C) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
+%\begin{figure}[ht]
+%\centerline{\includegraphics[width=.55\textwidth]{figures/figure6.pdf}}
+%\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and C) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
 Dot sizes correspond to district numbers, colors represent the land cover map.}
-\label{afoto}
-\end{figure}
+%\label{afoto}
+%\end{figure}
 
@@
 
-\begin{figure}[!ht]
-\centerline{\includegraphics[width=.55\textwidth]{figures/location_offset.pdf}}
-\caption{Average nearest neighbor distances (in km) between pixels representing features identified by the reference map versus those identified from the test maps.
+%\begin{figure}[!ht]
+%\centerline{\includegraphics[width=.55\textwidth]{figures/location_offset.pdf}}
+%\caption{Average nearest neighbor distances (in km) between pixels representing features identified by the reference map versus those identified from the test maps.
 Bar colors indicate the different features (and thus contributing maps), which were delineated by selecting pixels with values greater than the 90th percentile: densest cropland (solid bars); highest maize yield (medium transparent bars); highest carbon density (most transparent bars).}
-\label{afoto}
-\end{figure}
+%\label{afoto}
+%\end{figure}
 
@@
 \bibliographystyle{gcbnourl} 
+%\bibliographystyle{prsb} 
 {\footnotesize \bibliography{/Users/lestes/Dropbox/publications/fullbib.bib}}

### 2017-08-22 Rmd/c2-landcover-preprocess-2.Rmd
Various revisions on way to revising (minor) for GCB

@@
 p_root <- proj_root("SAcropland")
-p_fig <- full_path(p_root, "SAcropland/external/figures/")
-p_data <- full_path(p_root, "SAcropland/external/ext_data/")
+p_fig <- fp(p_root, "SAcropland/external/figures/")
+p_data <- fp(p_root, "SAcropland/external/ext_data/")
 ```
@@
 # Get SA shape
-if(!file.exists(full_path(p_data, "MZshapes.Rdata"))) {
+if(!file.exists(fp(p_data, "MZshapes.Rdata"))) {
   system("wget http://dl.dropboxusercontent.com/u/31890709/MZshapes.Rdata")
 }
-load(full_path(p_data, "MZshapes.Rdata"))
+load(fp(p_data, "MZshapes.Rdata"))
 sa_buf <- gBuffer(sa.shp, width = 3000)
@@
 # Create 1 km grid from sa1kilo to rectify grids against that 
-if(!file.exists(full_path(p_data, "sagrid.tif"))) {  # already made? 
+if(!file.exists(fp(p_data, "sagrid.tif"))) {  # already made? 
   sql <- paste("SELECT gid, id, ST_AsText(ST_Centroid(geom)) as center", 
@@
   sa_r <- rasterFromXYZ(pointsXYZ)
-  projection(sa_r) <- sa.buf@proj4string
-  sa_r <- writeRaster(sa_r, filename = full_path(p_data, "sagrid.tif"), 
+  projection(sa_r) <- sa_buf@proj4string
+  sa_r <- writeRaster(sa_r, filename = fp(p_data, "sagrid.tif"), 
                       overwrite = TRUE)
 } else {
-  sa_r <- raster(full_path(p_data, "sagrid.tif"))
+  sa_r <- raster(fp(p_data, "sagrid.tif"))
 }
@@
 ```{r, eval = FALSE}
-if(!file.exists(full_path(p_data, "glcsa_masked.tif"))) {  # see if file exists,  
+if(!file.exists(fp(p_data, "glcsa_masked.tif"))) {  # see if file exists,  
   # to avoid redoing step
@@
   unzip("glc_share.zip", exdir = "glc_share")
-  glc <- raster(full_path(path, "glc_shv10_02.Tif"))
+  glc <- raster(fp(path, "glc_shv10_02.Tif"))
   projection(glc) <- gcsstr
-  glcSA <- crop(glc, y=sa_buf_gcs, file = full_path(path, "glcSA.tif"), 
+  glcSA <- crop(glc, y=sa_buf_gcs, file = fp(path, "glcSA.tif"), 
                 overwrite = TRUE)  # crop to SA
   gdalwarp(srcfile = glcSA@file@name, 
-           dstfile = full_path(path, "glcSA_alb.tif"), t_srs = prjstr, 
+           dstfile = fp(path, "glcSA_alb.tif"), t_srs = prjstr, 
            tr = c(1000, 1000), r = "bilinear")
-  glcSA_alb <- raster(full_path(path, "glcSA_alb.tif"))
+  glcSA_alb <- raster(fp(path, "glcSA_alb.tif"))
 
@@
   glcsa <- resample(glcSA_alb, sa_r, method = "bilinear", 
-                    filename = full_path(path, "glcSA_alb_rect.tif"))
-  glcsa_m <- mask(glcsa, sa_r, file = full_path(p_data, "glcsa_masked.tif"),
+                    filename = fp(path, "glcSA_alb_rect.tif"))
+  glcsa_m <- mask(glcsa, sa_r, file = fp(p_data, "glcsa_masked.tif"),
                   overwrite = TRUE)
 } else {
-  glcsa_m <- raster(full_path(p_data, "glcsa_masked.tif"))
+  glcsa_m <- raster(fp(p_data, "glcsa_masked.tif"))
 }
 
-if(!file.exists(full_path(p_data, "geowikisa_masked.tif"))) {  # see if file exists, to avoid redoing step 
+if(!file.exists(fp(p_data, "geowikisa_masked.tif"))) {  # see if file exists, to avoid redoing step 
    # first downloaded manually from password protected geo-wiki site
@@
   unzip("cropland_hybrid_14052014v8.zip", exdir= "geowiki") 
-  geowiki <- raster(full_path(path, "Hybrid_14052014V8.img"))
+  geowiki <- raster(fp(path, "Hybrid_14052014V8.img"))
   projection(geowiki) <- gcsstr
   geowikiSA <- crop(geowiki, y=sa_buf_gcs, 
-                    file = full_path(path, "geowikiSA.tif"))  # crop to SA
+                    file = fp(path, "geowikiSA.tif"))  # crop to SA
   gdalwarp(srcfile = geowikiSA@file@name, 
-           dstfile = full_path(path, "geowikiSA_alb.tif"), t_srs = prjstr, 
+           dstfile = fp(path, "geowikiSA_alb.tif"), t_srs = prjstr, 
            tr = c(1000, 1000), r = "bilinear")
-  geowikisa_alb <- raster(full_path(path, "geowikiSA_alb.tif"))
+  geowikisa_alb <- raster(fp(path, "geowikiSA_alb.tif"))
   geowikisa <- resample(geowikisa_alb, sa_r, method = "bilinear", # Warp to SA
-                        filename = full_path(path, "geowikisa_alb_rect.tif"))    
+                        filename = fp(path, "geowikisa_alb_rect.tif"))    
   geowikisa_m <- mask(glcsa, sa_r, 
-                      file = full_path(p_data, "geowikisa_masked.tif"), 
+                      file = fp(p_data, "geowikisa_masked.tif"), 
                       overwrite = TRUE)
 } else {
-  geowikisa_m <- raster(full_path(p_data, "geowikisa_masked.tif"))
+  geowikisa_m <- raster(fp(p_data, "geowikisa_masked.tif"))
 }
@@
 path <- "/u/lestes/spatial_data/globcover_2009/"
-if(!file.exists(full_path(path, "globSA_alb.tif"))) {
+if(!file.exists(fp(path, "globSA_alb.tif"))) {
   url <- "http://due.esrin.esa.int/globcover/LandCover2009/Globcover2009_V2.3_Global_.zip"
@@
   unzip("globcover_2009.zip", exdir= "globcover_2009")
-  glob <- raster(full_path(path, "GLOBCOVER_L4_200901_200912_V2.3.tif"))
+  glob <- raster(fp(path, "GLOBCOVER_L4_200901_200912_V2.3.tif"))
   globSA <- crop(glob, y=sa_buf_gcs)  # crop to extent of SA
   gdalwarp(srcfile = globSA@file@name, 
-           dstfile = full_path(path, "globSA_alb.tif"), t_srs = prjstr, 
+           dstfile = fp(path, "globSA_alb.tif"), t_srs = prjstr, 
            tr = c(300, 300))
-  globSA_alb <- raster(full_path(path, "globSA_alb.tif"))
+  globSA_alb <- raster(fp(path, "globSA_alb.tif"))
 } else {
-  globSA_alb <- raster(full_path(path, "globSA_alb.tif"))
+  globSA_alb <- raster(fp(path, "globSA_alb.tif"))
 }
@@
 moddir <- "/u/lestes/spatial_data/MCD12Q1/"
-if(!file.exists(full_path(moddir, "modis_type_1_alb.tif"))) {
+if(!file.exists(fp(moddir, "modis_type_1_alb.tif"))) {
   tiles <- expand.grid("h" = c(19, 20), "v" = c(11, 12))
@@
 
+  # Quick convert of one MODIS set to get funny resolution
+  setwd(moddir)
+  gdal_translate(src_dataset = modis[1], dst_dataset= "modis1.tif", sd_index= 1)
+  r <- raster("modis1.tif")
+  r <- res(r)
+
   # Translate to geotiff
@@
              dstfile = paste0("modis_type_", x, "_alb.tif"), 
-             t_srs = prjstr, tr = res(r), te = bbox(sa.buf)[1:4])
+             t_srs = prjstr, tr = res(r), te = bbox(sa_buf)[1:4])
   })  
-  mod_1 <- raster("modis_type1_alb.tif")
+  mod_1 <- raster("modis_type_1_alb.tif")
 } else {
-  mod_1 <- raster(full_path(moddir, "modis_type_1_alb.tif"))
+  mod_1 <- raster(fp(moddir, "modis_type_1_alb.tif"))
 }
@@
   url <- "http://planet.uwc.ac.za/BGISdownloads/landcover_2009.zip"
-  download.file(url, method="auto", destfile = "sa_landcover_2009.zip")
+  download.file(url, method="auto", destfile = "landcover_2009.zip")
   unzip("landcover_2009.zip", exdir = "sa_lc_2009")
@@
                  of = "GTiff", ot = "BYTE")
-  #system(paste("rm -r", full_path(path, "landcover")))
+  #system(paste("rm -r", fp(path, "landcover")))
   #file.remove("../landcover_2009.zip")
@@
               type = "Byte", filename = lcnm[j])
-    nm <- full_path(p_data, 
+    nm <- fp(p_data, 
                     paste(gsub("\\.tif", "", lcnm[j]), "_masked.tif", sep = ""))
@@
   lcmnm <- paste(gsub("\\.tif", "", lcnm), "_masked.tif", sep = "")
-  salc_l <- lapply(lcmnm, function(x) raster(full_path(p_data, x)))
+  salc_l <- lapply(lcmnm, function(x) raster(fp(p_data, x)))
 } 
@@
 ```{r, eval = FALSE}
-if(!file.exists(full_path(p_data, "kzn_cane_masked.tif"))) {
+if(!file.exists(fp(p_data, "kzn_cane_masked.tif"))) {
   path <- "/u/lestes/spatial_data/kzn_landcover/"
@@
                  of = "GTiff")
-  system(paste("rm -r", full_path(path, "kznlc11v1w31")))
-  kznlc <- raster(full_path(npath, "kznlandcover.tif"))
+  system(paste("rm -r", fp(path, "kznlc11v1w31")))
+  kznlc <- raster(fp(npath, "kznlandcover.tif"))
 
@@
   system.time(system(pcalc))  # 49 seconds
-  kzn_cane <- raster(full_path(path, kznlconm))
+  kzn_cane <- raster(fp(path, kznlconm))
   # cext <- extent(c(20000, 40000, -3220000, -3200000))  # crops to small extent for examination
-  # kznlc.ss <- crop(kznlc, cext, filename = full_path(kznpath, "kznlandcover_crop.tif"), overwrite = TRUE)
+  # kznlc.ss <- crop(kznlc, cext, filename = fp(kznpath, "kznlandcover_crop.tif"), overwrite = TRUE)
   # plot(kznlc.ss)
@@
            verbose = TRUE, overwrite = TRUE)
-  kzn_1km <- raster(full_path(p_data, "kzn_cane_1km.tif"))
+  kzn_1km <- raster(fp(p_data, "kzn_cane_1km.tif"))
   kzn_1km_r <- resample(kzn_1km, sa_r, method = "bilinear")
   kzn_1km_m <- mask(kzn_1km_r, sa_r, 
-                    filename = full_path(p_data, "kzn_cane_masked.tif"), 
+                    filename = fp(p_data, "kzn_cane_masked.tif"), 
                     overwrite = TRUE)
 } else {
-  kzn_1km_m <- raster(full_path(p_data, "kzn_cane_masked.tif"))
+  kzn_1km_m <- raster(fp(p_data, "kzn_cane_masked.tif"))
 }
@@
 ```{r, eval = FALSE}
-if(any(!file.exists(full_path(p_data, 
-                              dir(p_data, 
-                                  pattern = "mod_1*.*1kmrect_sum.tif"))))) {
+if(any(!file.exists(fp(p_data, 
+                       dir(p_data, pattern = "mod_1*.*1kmrect_sum.tif"))))) {
   # MODIS, using main classification scheme (IGBP-DIS)
@@
   modsa_min <- assignLCPct(mod_1_cropb, minset, 
-                           fname = full_path(p_data, "mod_1_min.tif"))
+                           fname = fp(p_data, "mod_1_min.tif"))
   modsa_mu <- assignLCPct(mod_1_cropb, meanset, 
-                          fname = full_path(p_data, "mod_1_mu.tif"))
+                          fname = fp(p_data, "mod_1_mu.tif"))
   modsa_max <- assignLCPct(mod_1_cropb, maxset, 
-                           fname = full_path(p_data, "mod_1_max.tif"))
+                           fname = fp(p_data, "mod_1_max.tif"))
 
@@
 } else {
-  modsa_list_sum <- lapply(full_path(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
+  fnms <- fp(p_data, dir(p_data, "mod_1*.*1kmrect_sum.tif")[c(2:3, 1)])
+  modsa_list_sum <- lapply(fnms, raster)
 }  
@@
 We want Globcover class 11 (irrigated cropland), class 14 (rainfed), 20 (50-70% cropland), and 30 (20-50% cropland).
-For the latter two classes we create variants assumung the minimum, mean, and max of earh class
+For the latter two classes we create variants assumung the minimum, mean, and max of each class
 ```{r, eval = FALSE}
-if(any(!file.exists(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif"))))) {
+if(any(!file.exists(fp(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif"))))) {
   l <- lapply(c(11, 14, 20, 30), function(x) r <- globSA_alb == x)  # pull out classes
   s <- stack(l)
-  globSA_cropb <- brick(full_path(p_data, "globSA_crop.grd"))
+  globSA_cropb <- brick(s, filename = fp(p_data, "globSA_crop.grd"))
   rm(s, l)
@@
   meanset <- c(100, 100, 60, 35)
-  globsa_min <- assignLCPct(globSA_cropb, minset, fname = full_path(p_data, "globSA_300_min.tif"))
-  globsa_mu <- assignLCPct(globSA_cropb, meanset, fname = full_path(p_data, "globSA_300_mu.tif"))
-  globsa_max <- assignLCPct(globSA_cropb, maxset, fname = full_path(p_data, "globSA_300_max.tif"))
+  globsa_min <- assignLCPct(globSA_cropb, minset, 
+                            fname = fp(p_data, "globSA_300_min.tif"))
+  globsa_mu <- assignLCPct(globSA_cropb, meanset, 
+                           fname = fp(p_data, "globSA_300_mu.tif"))
+  globsa_max <- assignLCPct(globSA_cropb, maxset, 
+                            fname = fp(p_data, "globSA_300_max.tif"))
   glob300list <- lapply(dir(pattern = "globSA_300+.+tif")[c(2:3, 1)], raster)
@@
 } else {
-  globsa_list_sum <- lapply(full_path(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
+  globsa_list_sum <- lapply(fp(p_data, dir(p_data, "globSA*.*1kmrect_sum.tif")[c(2:3, 1)]), raster)
 } 
@@
 # Bring in SA landcover sets and mask
-cover2007 <- brick(full_path(p_data, "cover2007.tif"))
-cover2011 <- brick(full_path(p_data, "cover2011.tif"))
+cover2007 <- brick(fp(p_data, "cover2007.tif"))
+cover2011 <- brick(fp(p_data, "cover2011.tif"))
 
@@
 rm(m1, m2)
-full_mask <- writeRaster(sa_masks + sust_mask, file = full_path(p_data, "mask.tif"), overwrite = TRUE)
+full_mask <- writeRaster(sa_masks + sust_mask, file = fp(p_data, "mask.tif"), overwrite = TRUE)
 

### 2017-08-31 Rmd/c2-landcover-preprocess-2.Rmd
Gigantic first commit of croplandbias repo after being cloned and renamed from SAcropland. Significant changes: 1. Rerrangement of data structure structure--key intermediate inputs/outputs will be kept in inst/extdata/subfolder to install as R library; some other smaller ones are in data/ 2. Paper and figures are in inst/paper 3. Code has been re-run from c1-c9 (yield-production). Numbering of code chapters has been changed because landcover dataset download steps were placed first in standalone chapter (c1-landcover-fetch) 4. Analysis chapters are now stored as vignettes. Rmd folder is staged

@@
 
+Additonal note: This part of the analysis was undertaken on the Linux VM mapper.princeton.edu, using `gdal 1.11`.
+For full reproducibility of results, this version should be used (`gdalwarp` appears to have different behaviour in older versions when no `-te` (extent) values are passed, which makes the warp slightly different between old and new versions due to different extents after warping).
+
 ## Analyses
@@
 library(RCurl)
-library(RPostgreSQL)
+# library(RPostgreSQL) # used for initial extraction of sagrid
 library(rgdal)
@@
 library(raster)
-library(lmisc)
 ```
@@
 ```{r, eval = FALSE}
+rm(list = ls())
 #setwd("/u/lestes/analyses/SAcropland/")
-p_root <- proj_root("SAcropland")
-p_fig <- fp(p_root, "SAcropland/external/figures/")
-p_data <- fp(p_root, "SAcropland/external/ext_data/")
+p_root <- proj_root("croplandbias")
+p_data <- fp(p_root, "croplandbias/inst/extdata/landcover")
+p_dl <- fp(p_root, "croplandbias/external/ext_data/landcover/")
+
+# Note: installed from homebrew gdal 1.11.5, which is older than version with
+# QGIS on Sierra (2.1)
+getOption("gdalUtils_gdalPath")[[1]]$path
+
 ```
@@
 # Get SA shape
-if(!file.exists(fp(p_data, "MZshapes.Rdata"))) {
-  system("wget http://dl.dropboxusercontent.com/u/31890709/MZshapes.Rdata")
-}
-load(fp(p_data, "MZshapes.Rdata"))
-sa_buf <- gBuffer(sa.shp, width = 3000)
-
-# Connection
-drv <- dbDriver("PostgreSQL")
-con <- dbConnect(drv, dbname = "SouthAfricaSandbox", user = "postgis", 
-                 password = "P0stG1S")
+data(sashp)
+sa_buf <- gBuffer(sashp, width = 3000)
 
 # projections, etc
-prjsrid <- 97490  # Albers
-prj_sql <- paste0("select proj4text from spatial_ref_sys where srid=", prjsrid)
-prjstr <- dbGetQuery(con, prj_sql)$proj4text
-gcsid <- 4326  # GCS
-gcs_sql <- paste0("select proj4text from spatial_ref_sys where srid=", gcsid)
-gcsstr <- dbGetQuery(con, gcs_sql)$proj4text
+prjstr <- projection(sashp)
+epsg <- data.table(make_EPSG())
+gcsstr <- epsg[like(code, "4326"), prj4]
 
@@
 
+# Connection (deprecated, from old version of mapper.princeton.edu, since 
+# updated to have whole Africa grid)
+# drv <- dbDriver("PostgreSQL")
+# con <- dbConnect(drv, dbname = "SouthAfricaSandbox", user = "postgis", 
+#                  password = "P0stG1S")
+
 # Create 1 km grid from sa1kilo to rectify grids against that 
-if(!file.exists(fp(p_data, "sagrid.tif"))) {  # already made? 
-  sql <- paste("SELECT gid, id, ST_AsText(ST_Centroid(geom)) as center", 
-               "FROM sa1kilo")
-  geom.tab <- dbGetQuery(con, sql)
-  coord.mat <- do.call(rbind, lapply(1:nrow(geom.tab), function(y) {
-    strip <- gsub("POINT|\\(|\\)", "", geom.tab[y, 3])
-    coord.mat <- do.call(rbind, strsplit(strsplit(strip, ",")[[1]], " "))
-  }))
-  class(coord.mat) <- "numeric"
-  point.tab <- cbind(geom.tab[, 1:2], coord.mat)
-  colnames(point.tab)[3:4] <- c("x", "y") 
-  pointsXYZ <- point.tab[, c(3:4, 2)]
-  sa_r <- rasterFromXYZ(pointsXYZ)
-  projection(sa_r) <- sa_buf@proj4string
-  sa_r <- writeRaster(sa_r, filename = fp(p_data, "sagrid.tif"), 
-                      overwrite = TRUE)
-} else {
+# sql <- paste("SELECT gid, id, ST_AsText(ST_Centroid(geom)) as center", 
+#              "FROM sa1kilo")
+# geom.tab <- dbGetQuery(con, sql)
+# coord.mat <- do.call(rbind, lapply(1:nrow(geom.tab), function(y) {
+#   strip <- gsub("POINT|\\(|\\)", "", geom.tab[y, 3])
+#   coord.mat <- do.call(rbind, strsplit(strsplit(strip, ",")[[1]], " "))
+# }))
+# class(coord.mat) <- "numeric"
+# point.tab <- cbind(geom.tab[, 1:2], coord.mat)
+# colnames(point.tab)[3:4] <- c("x", "y") 
+# pointsXYZ <- point.tab[, c(3:4, 2)]
+# r <- rasterFromXYZ(pointsXYZ)
+# projection(r) <- sa_buf@proj4string
+# sa_r <- writeRaster(r, filename = fp(p_data, "sagrid.tif"), 
+#                     overwrite = TRUE)
 sa_r <- raster(fp(p_data, "sagrid.tif"))
-}
+
 ```
@@
 ### Bring in landcover datasets
-Let's start with the improved/fusion datasets, GLC share and geo-wiki's
+Let's start with the improved/fusion dataset, geo-wiki (note: GLC-Share was also initially tested, but it turned out to be identical to geo-wiki)
+
 ```{r, eval = FALSE}
-if(!file.exists(fp(p_data, "glcsa_masked.tif"))) {  # see if file exists,  
-  # to avoid redoing step
-  path <- "/u/lestes/spatial_data/glc_share/"
-  url <- "http://www.fao.org/geonetwork/srv/en/resources.get?id=47948&fname=GlcShare_v10_02.zip&acces=private"
+setwd(p_dl)
+path <- fp(p_dl, "glc_share")
+# url <- "http://www.fao.org/geonetwork/srv/en/resources.get?id=47948&fname=GlcShare_v10_02.zip&acces=private"
+url <- paste0("http://www.fao.org/geonetwork/srv/en/resources.get?id=47948&",
+              "fname=GlcShare_v10_02.zip&access=private")
 download.file(url, method = "auto", destfile = "glc_share.zip")
@@
                 overwrite = TRUE)
-} else {
-  glcsa_m <- raster(fp(p_data, "glcsa_masked.tif"))
-}
 
+```
+
+```{r, eval = FALSE}
 if(!file.exists(fp(p_data, "geowikisa_masked.tif"))) {  # see if file exists, to avoid redoing step 
    # first downloaded manually from password protected geo-wiki site
-  path <- "/u/lestes/spatial_data/geowiki/"
+  # path <- "/u/lestes/spatial_data/geowiki/"
+  path <- fp(p_dl, "geowiki/")
   unzip("cropland_hybrid_14052014v8.zip", exdir= "geowiki") 
+  # geowiki <- raster(fp(path, "Hybrid_14052014V8.img"))
   geowiki <- raster(fp(path, "Hybrid_14052014V8.img"))
-  projection(geowiki) <- gcsstr
-  geowikiSA <- crop(geowiki, y=sa_buf_gcs, 
-                    file = fp(path, "geowikiSA.tif"))  # crop to SA
-  gdalwarp(srcfile = geowikiSA@file@name, 
+  # projection(geowiki) <- gcsstr
+  # crs(geowiki) <- CRS(gcsstr)
+  # geowikiSA <- crop(geowiki, y=sa_buf_gcs,
+  #                   file = fp(path, "geowikiSA.tif"))  # crop to SA
+  geowikiSA <- crop(geowiki, y=sa_buf_gcs, overwrite = TRUE, 
+                    file = fp(p_dl, "geowiki/geowikiSA.tif"))  # crop to SA
+  gdalwarp(srcfile = filename(geowikiSA), 
            dstfile = fp(path, "geowikiSA_alb.tif"), t_srs = prjstr, 
@@
                         filename = fp(path, "geowikisa_alb_rect.tif"))    
-  geowikisa_m <- mask(glcsa, sa_r, 
+  # geowikisa_m <- mask(glcsa, sa_r, 
+  geowikisa_m <- mask(geowikisa, sa_r, 
                       file = fp(p_data, "geowikisa_masked.tif"), 
@@
 }
+
+# test
+tpath <- "croplandbias/external/ext_data/geowikisa_masked.tif"
+geowikisa_mo <- raster(fp(p_root, tpath))
+plot(geowikisa_mo - geowikisa_m)
+
 ```
 
-And the standard products, Globcover 2009 and MODIS
+And the standard products, Globcover 2009 and MODIS. Note the URL used originally has been superseded by the one commented out
 ```{r, eval = FALSE}
@@
 if(!file.exists(fp(path, "globSA_alb.tif"))) {
-  url <- "http://due.esrin.esa.int/globcover/LandCover2009/Globcover2009_V2.3_Global_.zip"
+  url <- paste0("http://due.esrin.esa.int/globcover/LandCover2009/", 
+                "Globcover2009_V2.3_Global_.zip")
+  # url <- "http://due.esrin.esa.int/files/Globcover2009_V2.3_Global_.zip"
   download.file(url, method="auto", destfile = "globcover_2009.zip")
   unzip("globcover_2009.zip", exdir= "globcover_2009")
-  glob <- raster(fp(path, "GLOBCOVER_L4_200901_200912_V2.3.tif"))
-  globSA <- crop(glob, y=sa_buf_gcs)  # crop to extent of SA
-  gdalwarp(srcfile = globSA@file@name, 
-           dstfile = fp(path, "globSA_alb.tif"), t_srs = prjstr, 
-           tr = c(300, 300))
-  globSA_alb <- raster(fp(path, "globSA_alb.tif"))
+  # glob <- raster(fp(path, "GLOBCOVER_L4_200901_200912_V2.3.tif"))
+  glob <- raster(fp(p_dl,
+                    "globcover_2009/GLOBCOVER_L4_200901_200912_V2.3.tif"))
+  # globSA <- crop(glob, y=sa_buf_gcs)  # crop to extent of SA
+  # e <- as(extent(c(-776722.9, 920077.1, -3831848, -2385248)),
+  #                "SpatialPolygons")
+  # crs(e) <- CRS(prjstr)
+  # e <- spTransform(e, CRS(gcsstr))
+  # globSA <- crop(glob, y = e, overwrite = TRUE,
+  globSA <- crop(glob, y = sa_buf_gcs, overwrite = TRUE,
+                 filename = fp(p_dl, "globcover_2009/globSA.tif"))  # crop
+  # gdalwarp(srcfile = filename(globSA), overwrite = TRUE, 
+  gdalwarp(srcfile = filename(globSA), overwrite = TRUE, 
+           dstfile = fp(p_dl, "globSA_alb.tif"), t_srs = prjstr,
+           # tap = TRUE,
+           # dstfile = fp(path, "globSA_alb.tif"), t_srs = prjstr, 
+           tr = c(300, 300))#, te = extent(e)[1:4])
+  # globSA_alb <- raster(fp(path, "globSA_alb.tif"))
+  globSA_alb <- raster(fp(p_dl, "globSA_alb.tif"))
 } else {
@@
 # MODIS landcover 
-moddir <- "/u/lestes/spatial_data/MCD12Q1/"
+# moddir <- "/u/lestes/spatial_data/MCD12Q1/"
+moddir <- fp(p_dl, "MCD12Q1")
 if(!file.exists(fp(moddir, "modis_type_1_alb.tif"))) {
@@
   setwd(moddir)
-  gdal_translate(src_dataset = modis[1], dst_dataset= "modis1.tif", sd_index= 1)
+  gdal_translate(src_dataset = modis[1], dst_dataset= "modis1.tif", 
+                 sd_index= 1)
   r <- raster("modis1.tif")
-  r <- res(r)
 
@@
   mod_1 <- raster("modis_type_1_alb.tif")
+  file.remove(dir(moddir, pattern = "MCD12Q1"))
 } else {
@@
 ```{r, eval = FALSE}
-path <- "/u/lestes/spatial_data/sa_lc_2009/"
-setwd("/u/lestes/spatial_data/")
+# path <- "/u/lestes/spatial_data/sa_lc_2009/"
+# setwd("/u/lestes/spatial_data/")
+path <- fp(p_dl, "sa_lc_2009/")
 lcnm <- c("sa_ag.tif", "sa_af.tif")
@@
               type = "Byte", filename = lcnm[j])
-    nm <- fp(p_data, 
-                    paste(gsub("\\.tif", "", lcnm[j]), "_masked.tif", sep = ""))
+    # nm <- fp(p_data, paste0(gsub("\\.tif", "", lcnm[j]), "_masked.tif"))
+    nm <- fp(path, paste0(gsub("\\.tif", "", lcnm[j]), "_masked.tif"))
     ext <- sapply(c("xmin", "ymin", "xmax", "ymax"), function(x) {
@@
     })
-    print(paste("warping and masking", gsub(p_data, "", nm)))
+    print(paste("warping and masking", gsub(path, "", nm)))
     gdalwarp(srcfile = lcnm[j], t_srs = projection(sa_r), dstfile = nm, 

### 2017-08-31 Rmd/c9-agent-based-model.Rmd
Gigantic first commit of croplandbias repo after being cloned and renamed from SAcropland. Significant changes: 1. Rerrangement of data structure structure--key intermediate inputs/outputs will be kept in inst/extdata/subfolder to install as R library; some other smaller ones are in data/ 2. Paper and figures are in inst/paper 3. Code has been re-run from c1-c9 (yield-production). Numbering of code chapters has been changed because landcover dataset download steps were placed first in standalone chapter (c1-landcover-fetch) 4. Analysis chapters are now stored as vignettes. Rmd folder is staged

--- 2015-11-02 Rmd/agent-allocation-bias.Rmd
@@
 ---
-title: "Bias in Allocation of ABM Agents"
+title: "Agent-based model sensitivity"
 author: "Lyndon Estes"
@@
 
-# Overview
+# Agent-based model sensitivity to map error
 
@@
 
-# Data
+## Data
 ```{r, eval=FALSE}
@@
 
-## Magisterial District selection
+### Magisterial District selection
 
@@
 
-### Map selected districts, cropland percentage, and bias
+### Map selected districts, cropland percentage, and error
 ```{r, eval = FALSE}
@@
 
-# Select modeled yields for agent-based model
+## Select modeled yields for agent-based model
 
@@
 
-# Downscaling and agent allocation
+## Downscaling and agent allocation
 
-Performed separately and described in the methods.
+Performed separately and described in the methods and supplemental.
 
-# Analysis of agent allocation bias
+## Analysis of agent allocation errors
 ```{r, eval = FALSE}

### 2017-08-31 Rmd/c99-misc.Rmd
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.
Gigantic first commit of croplandbias repo after being cloned and renamed from SAcropland. Significant changes: 1. Rerrangement of data structure structure--key intermediate inputs/outputs will be kept in inst/extdata/subfolder to install as R library; some other smaller ones are in data/ 2. Paper and figures are in inst/paper 3. Code has been re-run from c1-c9 (yield-pr

--- 2015-08-27 Rmd/misc.Rmd
@@
 ---
-title: "Miscallaneous calculations"
+title: "Miscellaneous calculations"
 author: "Lyndon Estes"
@@
 
-Various small calculations made for the analysis.
+# Miscellaneous additional calculations
 
-```{r, eval=FALSE}
+```{r, message=FALSE}
 library(rgdal)
@@
 sum(a) / sum(b)
+```
 
+## Calculation of error propagation
+```{r}
+load(fp(p_dat, "cropland-werr-2011.rda"))
+load(fp(p_dat, "yield-prod-accbias.rda"))
+load(fp(p_dat, "et-err-bias.rda"))
+load(fp(p_dat, "carbon-werr.rda"))
+
+cland <- bacc2011
+yprod <- as.data.table(out_tab)[Region == "Density"]
+carb <- rbind(cb_statsw_mu, cb_statsw_mua)
+
+# Reshape error stats
+# yield/production 
+yp1 <- melt(yprod, id.vars = c("Map", "Metric", "Variable"), 
+            measure.vars = names(yprod)[5:10])
+yp2 <- dcast(yp1, formula = Metric + Variable + variable ~ Map)
+setcolorder(yp2, neworder = c(1:3, ncol(yp2), 5:6, 4))
+setnames(yp2, names(yp2), c("stnm", "var", "ol", names(cland)[-c(1:2)]))
+
+snms <- names(cland)[-c(1:2)]
+lev <- cland[stnm == "mu", ol]
+
+# Production and yield propagation
+yp_prop <- rbindlist(lapply(c("Yield", "Production"), function(x) {  
+  o <- data.frame(t(sapply(snms, function(y) {  # y <- "sa30"
+    emag <- yp2[var == x & stnm == "MAE", get(y)] / 
+      cland[stnm == "mua", get(y)] 
+    c(emag, mean(emag))
+  })))
+  o <- cbind(snms, round(o, 2))
+  colnames(o) <- c("map", lev, "mu")
+  data.table(cbind.data.frame("var" = x, o))
+}))
+
+# Carbon differences
+c_prop <- rbindlist(lapply(snms, function(x) {  # x <- "sa30"
+  dat <- carb[stnm == "MAE" & map == x, names(carb)[-c(1:3)], with = FALSE]
+  o <- dat[, lapply(.SD, function(y) y / cland[stnm == "mua", get(x)]), 
+           .SDcols = names(dat)]
+  o <- round(rbind(o, data.table(t(colMeans(o)))), 2)
+  o <- cbind("ol" = c(lev, "mu"), o)
+  cbind("map" = x, o)
+  # dat[, second] / cland[stnm == "mua", get(x)]
+}))
+
+# et propagation
+et_prop <- round(et_err[Variable == "Annual Mean", MAE] / 
+  cland[ol == "f25" & stnm == "mua", snms, with = FALSE], 2)
+
+yp_prop
+c_prop
+et_prop
 ```

### 2017-08-31 Rmd/compile-book.Rmd
Gigantic first commit of croplandbias repo after being cloned and renamed from SAcropland. Significant changes: 1. Rerrangement of data structure structure--key intermediate inputs/outputs will be kept in inst/extdata/subfolder to install as R library; some other smaller ones are in data/ 2. Paper and figures are in inst/paper 3. Code has been re-run from c1-c9 (yield-production). Numbering of code chapters has been changed because landcover dataset download steps were placed first in standalone chapter (c1-landcover-fetch) 4. Analysis chapters are now stored as vignettes. Rmd folder is staged

@@
 cfiles <- list.files(pattern = "[0-9].*.Rmd")
+reord <- order(as.numeric(gsub("c|-.*", "", cfiles)))
+cfiles <- cfiles[(1:length(cfiles))[reord]]
 ttext <- NULL

### 2017-08-31 inst/paper/figures/cropadj-bias-accuracy.tex
Gigantic first commit of croplandbias repo after being cloned and renamed from SAcropland. Significant changes: 1. Rerrangement of data structure structure--key intermediate inputs/outputs will be kept in inst/extdata/subfolder to install as R library; some other smaller ones are in data/ 2. Paper and figures are in inst/paper 3. Code has been re-run from c1-c9 (yield-production). Numbering of code chapters has been changed because landcover dataset download steps were placed first in standalone chapter (c1-landcover-fetch) 4. Analysis chapters are now stored as vignettes. Rmd folder is staged

% latex table generated in R 3.4.1 by xtable 1.8-2 package
% Wed Aug 30 17:02:57 2017
\begin{longtable}{llrrrrrr}
\caption{Bias and mean absolute errors (MAE) in statistically constrained cropland maps across aggregation scales, weighted by density of cropland cover in the reference map. } \\ 
  \hline
Metric & Map & 1 km & 5 km & 10 km & 25 km & 50 km & 100 km \\ 
  \hline
Bias & GLC-Share & 9.7 & 1.1 & 0.6 & 0.4 & 0.5 & 0.1 \\ 
  Bias & GlobCover & 34.5 & 18.3 & 14.5 & 10.6 & 7.6 & 4.6 \\ 
  Bias & MODIS & 17.8 & 5.5 & 3.2 & 1.3 & 0.1 & -1.3 \\ 
  Bias & SA-LC & 6.6 & 2.7 & 2.1 & 1.6 & 1.1 & 0.6 \\ 
  Accuracy & GLC-Share & 23.8 & 12.6 & 9.4 & 6.8 & 4.8 & 3.0 \\ 
  Accuracy & GlobCover & 42.3 & 27.3 & 23.3 & 18.8 & 15.6 & 11.2 \\ 
  Accuracy & MODIS & 33.8 & 21.5 & 18.4 & 15.3 & 12.7 & 10.6 \\ 
  Accuracy & SA-LC & 11.4 & 6.0 & 4.7 & 3.7 & 2.8 & 1.9 \\ 
   \hline
\hline
\end{longtable}

### 2017-08-31 inst/paper/figures/et-bias.tex
Gigantic first commit of croplandbias repo after being cloned and renamed from SAcropland. Significant changes: 1. Rerrangement of data structure structure--key intermediate inputs/outputs will be kept in inst/extdata/subfolder to install as R library; some other smaller ones are in data/ 2. Paper and figures are in inst/paper 3. Code has been re-run from c1-c9 (yield-production). Numbering of code chapters has been changed because landcover dataset download steps were placed first in standalone chapter (c1-landcover-fetch) 4. Analysis chapters are now stored as vignettes. Rmd folder is staged

% latex table generated in R 3.4.1 by xtable 1.8-2 package
% Sun Aug 27 21:56:10 2017
\begin{longtable}{llrr}
\caption{Biases and mean absolute errors (as \%) for evapotranspiration variables derived from a 29-year time series calculated by the VIC model, including the average total ET for the 3 months of the year when ET is highest, the annual mean and the minimum and maximum annual ETs in the time series.} \\ 
  \hline
Variable & Map & Bias & MAE \\ 
  \hline
Peak & GLC-Share & 0.2 & 0.8 \\ 
  Annual Mean & GLC-Share & 0.3 & 0.7 \\ 
  29-year Min & GLC-Share & 0.3 & 0.6 \\ 
  29-year Max & GLC-Share & 0.3 & 0.8 \\ 
  Peak & GlobCover & 0.1 & 1.2 \\ 
  Annual Mean & GlobCover & -0.1 & 1.0 \\ 
  29-year Min & GlobCover & -0.2 & 0.9 \\ 
  29-year Max & GlobCover & 0.2 & 1.2 \\ 
  Peak & MODIS & -0.5 & 0.9 \\ 
  Annual Mean & MODIS & -0.6 & 0.8 \\ 
  29-year Min & MODIS & -0.6 & 0.7 \\ 
  29-year Max & MODIS & -0.5 & 0.8 \\ 
  Peak & SA-LC & 0.3 & 0.7 \\ 
  Annual Mean & SA-LC & 0.5 & 0.6 \\ 
  29-year Min & SA-LC & 0.4 & 0.5 \\ 
  29-year Max & SA-LC & 0.4 & 0.6 \\ 
   \hline
\hline
\end{longtable}

### 2017-08-31 inst/paper/figures/totC-bias.tex
Gigantic first commit of croplandbias repo after being cloned and renamed from SAcropland. Significant changes: 1. Rerrangement of data structure structure--key intermediate inputs/outputs will be kept in inst/extdata/subfolder to install as R library; some other smaller ones are in data/ 2. Paper and figures are in inst/paper 3. Code has been re-run from c1-c9 (yield-production). Numbering of code chapters has been changed because landcover dataset download steps were placed first in standalone chapter (c1-landcover-fetch) 4. Analysis chapters are now stored as vignettes. Rmd folder is staged

--- 2015-09-24 paper/figures/totC-bias.tex
@@
-% latex table generated in R 3.2.1 by xtable 1.7-4 package
-% Wed Sep 23 22:19:49 2015
+% latex table generated in R 3.4.1 by xtable 1.8-2 package
+% Mon Aug 28 06:26:20 2017
 \begin{table}[ht]
@@
   Country & MODIS & -2.6 & -2.5 & -2.5 & -0.1 & 2.1 \\ 
-  Country & GeoWiki & 0.6 & 0.5 & 0.5 & 0.0 & -0.5 \\ 
+  Country & GLC-Share & 0.6 & 0.5 & 0.5 & 0.0 & -0.5 \\ 
   Agricultural & SA-LC & -2.0 & -2.7 & -2.8 & -10.6 & -14.9 \\ 
@@
   Agricultural & MODIS & -1.6 & -0.8 & -0.7 & 8.4 & 13.3 \\ 
-  Agricultural & GeoWiki & 7.7 & 7.3 & 7.2 & 2.9 & 0.5 \\ 
+  Agricultural & GLC-Share & 7.7 & 7.3 & 7.2 & 2.9 & 0.5 \\ 
    \hline

### 2017-08-31 paper/SI.tex
Gigantic first commit of croplandbias repo after being cloned and renamed from SAcropland. Significant changes: 1. Rerrangement of data structure structure--key intermediate inputs/outputs will be kept in inst/extdata/subfolder to install as R library; some other smaller ones are in data/ 2. Paper and figures are in inst/paper 3. Code has been re-run from c1-c9 (yield-production). Numbering of code chapters has been changed because landcover dataset download steps were placed first in standalone chapter (c1-landcover-fetch) 4. Analysis chapters are now stored as vignettes. Rmd folder is staged

@@
 %\documentclass[12pt]{iopart}
-\documentclass[11pt, titlepage]{article}
+\documentclass[12pt, titlepage]{article}
 \usepackage[top=20mm, bottom=20mm, left=23mm, right=23mm]{geometry}
@@
 
+\usepackage{setspace}
+\doublespacing
+\usepackage{fontspec}
+\setmainfont{Times New Roman}
+
 %\bibpunct{(}{)}{,}{n}{,}{,}  % https://xianblog.wordpress.com/tag/natbib/ (allows natbib with PNAS)
@@
 
-\setpagewiselinenumbers
+%\setpagewiselinenumbers
 %\modulolinenumbers[5]
-%\linenumbers
 \begin{center}\section*{\Large Supplementary Methods and Results}\end{center}
@@
 %\end{table}  
+\linenumbers
 \section{\large Overview}
@@
 % Wed Nov 25 15:57:07 2015
+
+\begin{spacing}{1.0}
 \begin{table}[!ht]
@@
 \end{table}
+\end{spacing}
 
@@
 0-5\% cropland cover, 5-10\%, etc.), while solid lines indicate the mean absolute error.
-The black numbers in each plot area present the overall means of bias/MAE for each sensor-scale combination.
+The black numbers in each plot area present the overall means of bias/MAE for each sensor-scale combination; bold numbers provide the mean of bin-wise means of bias/MAE, while non-bold numbers represent the mean of bin-wise means weighted by the number of cells contributing to each bin.
 The bin-wise and overall mean statistics were calculated from pooled map errors calculated from differences between the 2007 reference map and each cropland map (including all three variants--high, medium, and low--of the MODIS and GlobCover-derived cropland maps), and the 2011 reference map and each cropland map.
@@
 4 frequencies of cells per five percentile bins of cropland density).
-Finally, we calculated a density-independent metric, which is similar to the density weighted mean calculated for our main analysis, which was calculated by averaging the mean bias and MAE values within different levels of cropland cover (divided in increments of 5\%).
+Finally, we calculated a density-independent metric, which produces values similar to the density-weighted means presented in our main analysis, which was calculated by averaging the mean bias and MAE values within different levels of cropland cover (binned into increments of 5\%, excluding areas with $<$0.5\% cropland).
 The results of these three alternate metrics are shown in Figure 3.

### 2017-08-31 paper/main.tex
Gigantic first commit of croplandbias repo after being cloned and renamed from SAcropland. Significant changes: 1. Rerrangement of data structure structure--key intermediate inputs/outputs will be kept in inst/extdata/subfolder to install as R library; some other smaller ones are in data/ 2. Paper and figures are in inst/paper 3. Code has been re-run from c1-c9 (yield-production). Numbering of code chapters has been changed because landcover dataset download steps were placed first in standalone chapter (c1-landcover-fetch) 4. Analysis chapters are now stored as vignettes. Rmd folder is staged

--- 2017-03-26 paper/main.tex
@@
-\documentclass[a4paper]{article}
+\documentclass[12 pt, titlepage, a4paper]{article}
 \usepackage[affil-it]{authblk}    
@@
 
+\usepackage{setspace}
+\usepackage{fontspec}
+\setmainfont{Times New Roman}
+
 %\begin{document}
@@
 %\title{A Large-Area, Spatially Continuous Quantification of Error in Land Cover Maps and Associated Socioeconomic and Environmental Analyses Relevant to Global Change}
+\vspace{-5cm}
 \title{A Large-Area, Spatially Continuous Assessment of Land Cover Map Error and Its Impact on Downstream Analyses}
 %\title{How Land Cover Errors Can Impact Understanding of Socioeconomic and Environment Phenomena}
-\author[1,2,3]{Lyndon Estes}
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+\author[1,2,3\footnote{+1-202-431-0496 (p) | +1-508-793-8881 (f) | LEstes@clarku.edu}]{Lyndon Estes}
 \author[4]{Peng Chen}
-\author[2]{Stephanie Debats}
+\author[3]{Stephanie Debats}
 \author[4]{Tom Evans}
@@
 \author[6,7]{Tobias Kuemmerle}
-\author[2]{Gabrielle Ragazzo}
-\author[2,8]{Justin Sheffield}
+\author[3]{Gabrielle Ragazzo}
+\author[3,8]{Justin Sheffield}
 \author[9]{Adam Wolf}
 \author[2]{Eric Wood}
-\author[2,10]{Kelly Caylor}
+\author[3,10]{Kelly Caylor}
 
-\affil[1]{Woodrow Wilson School, Princeton University, Princeton, NJ USA}
-\affil[2]{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}
-\affil[3]{Graduate School of Geography, Clark University, Worcester, MA USA}
+\affil[1]{Graduate School of Geography, Clark University, Worcester, MA USA}
+\affil[2]{Woodrow Wilson School, Princeton University, Princeton, NJ USA}
+\affil[3]{Civil and Environmental Engineering, Princeton University, Princeton, NJ USA}
 \affil[4]{Indiana University, Bloomington, IN USA}
@@
 \maketitle 
+\doublespacing
+\renewcommand*{\thefootnote}{\arabic{footnote}}
 
@@
 Instead of the standard confusion matrix-based accuracy metrics \citep{olofsson_good_2014,olofsson_making_2013}, which apply to categorical land cover maps, we assessed the bias and accuracy of the test maps based on the gridded residuals that resulted when each test map was subtracted from the reference map.
-Here bias is the mean residual value, weighted by the density of reference cropland, and accuracy is the mean of the absolute values of residuals (also weighted by cropland density), thus lower values signify higher accuracy.
+Here bias is the mean residual value, weighted by the density of reference cropland (to condition the analysis on cropland prevalence; see the SI for an assessment of error using other, non-weighted, variants of these measures), and accuracy is the mean of the absolute values of residuals (also weighted by cropland density), thus lower values signify higher accuracy.
 We calculated these metrics for the original 1 km resolution, and for maps that were further aggregated to 5, 10, 25, 50, and 100 km resolutions, in order to evaluate how bias and accuracy changes with observational grain.
@@
 Since our cropland percentage map provided no information on surrounding cover, we developed several variants representing potential surrounding cover types by assigning the average carbon densities of five biomes (forest, secondary forest, shrubland, grassland, and sparse vegetation \citep{ruesch_new_2008}) to the non-cropland fraction of our maps.
-These hypothetical maps represented the range in potential carbon densities, and allowed us to investigate how carbon estimates can vary as a function of i) test map errors and ii) the properties of neighboring cover types.
 We multiplied cropland densities by cropland fractions and added these to each of the five other densities multiplied by the residual non-cropland fractions to create five different carbon density maps.
 We aggregated each carbon map up to 5, 10, 25, 50, and 100 km resolutions for scaling comparisons (SI).
+To assess carbon estimation error, we subtracted test map-derived carbon maps from those based on the reference map, and calculated bias and accuracy scores using the method described for the cropland maps (see previous section).
 
-To assess carbon estimation error, we subtracted test map-derived carbon maps from those based on the reference map, and calculated bias and accuracy scores using the method described for the cropland maps (see previous section).
+Although this analysis ascribed hypothetical carbon densities to non-cropland areas, the selected values represent the range in potential carbon stocks in landscapes containing arable agriculture (error metrics from this analysis defined as pixels containing cropland, pixels without cropland do not influence the results), which allowed us to counterfactually investigate how carbon estimates are affected by the interaction of i) test map errors and ii) the properties of neighboring cover types.
 
@@
 We used the Variable Infiltration Capacity \citep[VIC;][]{liang_simple_1994} land surface hydrology model run with the Africa Flood and Drought Monitor's meteorological data \citep{sheffield_drought_2013} to produce monthly gridded evapotranspiration estimates for South Africa for the years 1979-2010 at 25 km resolution.
-VIC's land cover scheme (derived from AVHRR) provides values for leaf area index (LAI), plant rooting depth, aerodynamic roughness, and several other variables that the modeluses to partition water vapor fluxes into their evaporative and transpirative components.
-We adjusted VIC's base map so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25$^{\circ}$ resolution), then ran one instance of VIC for each of the five land cover schemes.
-We then compared the mean annual ET produced by the reference map variant with those from the test maps to assess the degree to which map errors impact evapotranspiration values.
+VIC's land cover scheme (derived from AVHRR) provides values for leaf area index (LAI), plant rooting depth, aerodynamic roughness, and several other variables that the model uses to partition water vapor fluxes into their evaporative and transpirative components.
+We adjusted VIC's base map so that its cropland fractions matched those of the 25 km reference and test maps (each reprojected and resampled to VIC's 0.25$^{\circ}$ resolution), and correspondingly altered the fractions of non-cropland cover types to accommodate the adjusted cropland fractions.
+We then ran one instance of VIC for each of the five land cover schemes, and compared the mean annual ET produced by the reference map variant with those from the test maps to assess the degree to which map errors impact evapotranspiration values.
 
@@
 
-We used these methods to first disaggregate the harvested area for maize (South Africa's largest crop \citep{estes_projected_2013}) onto our cropland maps, followed by yields, which were assigned to cells having harvested areas greater than zero.
+We used these methods to first disaggregate the harvested area for maize \citep[South Africa's largest crop;][]{estes_projected_2013} onto our cropland maps, followed by yields, which were assigned to cells having harvested areas greater than zero.
 The first step in this process entails adjusting cropland percentages so that their totals match census-derived cropland area estimates \citep{ramankutty_farming_2008,schierhorn_post-soviet_2013}.
 In place of census statistics, we used the reference map to calculate total cropland areas for South Africa's nine provinces, then adjusted the pixel-wise cropland percentages in the four test maps so that their province-wise sums matched these totals (SI).
-We then followed \citetapos{monfreda_farming_2008} procedure for disaggregating maize \citep[South Africa's largest crop;][]{estes_comparing_2013} planted area and yields onto the reference and adjusted test maps.
+We then followed \citetapos{monfreda_farming_2008} procedure for disaggregating planted area and yields onto the reference and adjusted test maps.
 The necessary statistics were obtained from magisterial district-level agricultural censuses conducted for the year 2007 \citep{statistics_south_africa_building_2007}.
@@
 All but MODIS underestimated ET by 5-15\% in the northern tip of the country.
-
+%
 %\vspace{-0.75 cm}
@@
 6A).
-This initialization error was caused by the spatial arrangement of croplands in the district.
-MODIS cropland in those districts was confined to relatively small islands, meaning that they fell outside the allocation algorithm's search window, thus leaving those croplands unassigned to specific households.
+This initialization error was caused by the spatial arrangement of croplands in the district, which in the MODIS maps was clumped in relatively small islands within the four selected districts.
+This error interacted with the ABM's household placement routine, in that some cropland islands fell beyond the 1.5 km search radius of their nearest randomly placed households, resulting in those croplands being left unallocated.
+This result demonstrates how map errors can propagate through more complex models as a function of both model assumptions (here the choice of search radius) and model structures (the randomized household placement routine).
 
-This latter finding highlights the types of errors that caused by spatial inaccuracies in land cover maps, which was explicitly evaluated by the analysis of spatial distance between pixels containing upper decile of values in reference and test maps (Fig.
+This latter finding also highlights the types of errors that can be caused by spatial inaccuracies in land cover maps, which was explicitly evaluated by the analysis of spatial distance between pixels containing upper decile of values in reference and test maps (Fig.
 7).
@@
 The most straightforward way may be to use a Monte Carlo approach that generates artificial datasets \citep[e.g.][]{avelino_goldilocks_2016}, using introduced errors drawn from reported or user-determined accuracy statistics, for both the base land cover and the downstream maps.
+For downstream products based on more complex models, users should examine how particular model assumptions or structures interact with land cover map errors.
 Quantifying error propagation is important to understanding how map error may influence subsequent understanding of the phenomenon of interest.
@@
 \vspace{-0.2 cm}
-This work was supported by funds from the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (EAR-1 534544, SES-1360463, and BCS-1026776).
-We  thank two anonymous reviewers for their constructive and helpful comments on earlier manuscript versions.
+This work was supported by the Princeton Environmental Institute Grand Challenges program, the NASA New Investigator Program (NNX15AC64G), and the National Science Foundation (EAR-1534544, SES-1360463, and BCS-1026776).
+We thank six anonymous reviewers for their constructive comments on earlier versions of the manuscript.
+Disclosure: S.F.
+is a director of GeoTerraImage, the owner of the reference map used in this study.
 %\end{acknowledgments}

### 2017-08-31 R/utilities.R

@@ -79,3 +79,19 @@ lpfunc <- function(x, y, w = NULL, col, size = 1, type, pch = NULL) {
     stop("Need to specify either 'pt' or 'line'")
   }
 }
+
+#' Helps get paths to system files installed with package
+#' @param x Name of system file to find
+#' @param y Sub-directory under inst/extdata where file is stored (details)
+#' @return Full path to system file
+#' @details For argument y, use "lc" for landcover, "yl" for "yield", and "cl" 
+#' for "climate".
+#' @keywords internal
+#' @export 
+spathfunc <- function(x, y = "lc") {
+  mat <- cbind(c("lc", "yl", "cl"), c("landcover", "yieldprod", "climate"))
+  base <- mat[which(mat[, 1] == y), 2]
+  system.file(paste0("extdata/", base), x, package = "croplandbias")
+}
+
+

### 2017-08-31 external/ext_R/grid-fields-intersect-analyze.R
Committing deletes from old external/ext_data and ext_r folders, and a few modifies to commit before removing from repo (paper/notes/...; external/test)

@@ -1,289 +1,0 @@
-##############################################################################################################
-# Title      : SAGridFieldsIntersect_analyze.R
-# Purpose    : Post-processing and analyzing results of intersections between fields and SA grid
-# Author     : Lyndon Estes
-# Draws from : SAGridFieldsIntersect_postgis.R; SAGridFieldsIntersect.R
-# Used by    : 
-# Notes      : 
-##############################################################################################################
-
-# Libraries
-suppressMessages(library(RPostgreSQL))
-suppressMessages(library(rgdal))
-suppressMessages(library(rgeos))
-suppressMessages(library(raster))
-source("/u/sandbox/afmap/R/KMLAccuracyFunctions.R")
-
-# Connection
-drv <- dbDriver("PostgreSQL")
-con <- dbConnect(drv, dbname = "db", user = "user", password = "pw")
-
-# Projection information
-prjsrid <- 97490
-prj.sql <- paste("select proj4text from spatial_ref_sys where srid=", prjsrid, sep = "")
-prjstr <- dbGetQuery(con, prj.sql)$proj4text
-
-# table names
-tab2007 <- "sa_fg_isect_2007"
-tab2011 <- "sa_fg_isect_2011"
-
-# Calculate area of each intersection
-lapply(list(tab2007, tab2011), function(x) {  # add column for area
-  d <- dbSendQuery(con, paste("ALTER TABLE",  x, "ADD COLUMN area numeric"))
-})
-lapply(list(tab2007, tab2011), function(x) {  # calculate area
-  print(paste("processing", x))
-  dbSendQuery(con, paste("UPDATE ", x, "SET area=ST_Area(geom)/10000")) 
-})
-lapply(list(tab2007, tab2011), function(x) {   # clean up tables
-  print(paste("processing", x))
-  dbSendQuery(con, paste("VACUUM ANALYZE",  x))
-})
-
-# vals <- lapply(list(tab2007, tab2011), function(x) {
-#   d <- dbGetQuery(con, paste("SELECT gid FROM", x, "WHERE area > 0"))$gid
-# })
-
-# A small check (all the same)
[…]

### 2017-08-31 external/ext_R/grid-fields-intersect-postgis.R

@@ -1,140 +1,0 @@
-##############################################################################################################
-# Title      : SAGridFieldsIntersect_postgis.R
-# Purpose    : Workflow between R, and postgis to clean up polygons and calculate intersections
-# Author     : Lyndon Estes
-# Draws from : SAGridFieldsIntersect_postgis.R; SAGridFieldsIntersect.R
-# Used by    : 
-# Notes      : 
-##############################################################################################################
-
-# Libraries
-suppressMessages(library(RPostgreSQL))
-suppressMessages(library(rgdal))
-suppressMessages(library(rgeos))
-# suppressMessages(library(raster))
-source("/u/sandbox/afmap/R/KMLAccuracyFunctions.R")
-
-test <- "N"
-
-if(test == "N") {
-  args <- commandArgs(TRUE)
-  if(length(args) != 5) stop("Wrong number of arguments")
-  fldname_in <- args[1]  # name of data table to create to hold cleaned fields in
-  fldname_out <- args[2]  # name of data table to create to hold cleaned fields in
-  grname_in <- args[3]  # name of input grid for comparison
-  grname_out <- args[4]  # name of input grid for comparison
-  isectname <- args[5]  # name of output table with intersections
-} else if(test == "Y") {
-  fldname_in <- "sa_flds_2011_alb"
-  fldname_out <- "sa_flds_2011_alb_cl"
-  isectname <- "sa_fg_isect_2011" # "sa_fg_isect_2007"
-  grname_in <- "sa1kilo"  # name of input grid for comparison
-  grname_out <- "sa_isect_grid"  # name of input grid for comparison
-}
-
-# Connection
-drv <- dbDriver("PostgreSQL")
-con <- dbConnect(drv, dbname = "db", user = "user", password = "pw")
-
-prjsrid <- 97490
-prj.sql <- paste("select proj4text from spatial_ref_sys where srid=", prjsrid, sep = "")
-prjstr <- dbGetQuery(con, prj.sql)$proj4text
-
-# First clean up SA fields datatables, starting with duplicate of it
-if(!dbExistsTable(con, name = fldname_out)) {
[…]

### 2017-08-31 external/test/function-devel.R
Removed external/test and paper folders from repo, as these are replaced by new structure

@@ -1,66 +1,0 @@
-# function-devel.R
-
-#' Generic function to create adjacent rectangular polygons from vectors of coordinates passed to it
-#' 
-#' @param x Vector of x coordinates, arranged in ascending order
-#' @param y Vector of y coordinates, arrange in acending order
-#' @return A polygon "chain": segmented polygons with interior vertical x or horizontal y axis
-#' @details This is modeled on the rect_coords function used by flex_legend, but the resulting polygon 
-#' can be of arbitrary shape. It can be used to make either a horizontal or vertical chain of polygons.  If 
-#' horizontal, then x should be an ascending vector providing one x value per polygon 
-poly_chain <- function(x, y) {
-  rects <- lapply(1:(length(x) - 1), function(z) {
-    rbind("ll" = c("x" = x[z], y2[z]),
-          "ul" = c("x" = x[z], y1[z]),
-          "ur" = c("x" = x[z + 1], y1[z + 1]),    
-          "lr" = c("x" = x[z + 1], y2[z + 1]), 
-          "ll" = c("x" = x[z], y2[z]))
-  })
-  return(rects)
-}
-#' Generic function to create adjacent rectangular polygons from vectors of coordinates passed to it
-#' 
-#' @param x Vector of x coordinates, arranged in ascending order
-#' @param y Vector of y coordinates, arrange in acending order
-#' @param fillcol A 1 to n - 1 length vector of colors for filling. n = length of x or y (the longest)
-#' @param linecol A 1 to n - 1 length vector of polygon border colors. n = length of x or y (the longest)
-#' @return A polygon "chain": segmented polygons with interior vertical x or horizontal y axis
-#' @details This is modeled on the rect_coords function used by flex_legend, but the resulting polygon 
-#' can be of arbitrary shape. It can be used to make either a horizontal (i.e. a sequence of polygons
-#' separated on the X-axis, with the long dimension filling the Y-axis) or vertical chain of polygons (the
-#' opposite). To make the former, provide a vector giving the X-coordinates separating each polygon, and a 2 
[…]

### 2017-08-31 R/utilities.R
Halfway updating c10, after having copied it into vignettes from Rmd. Updated spathfunc to read into abm sub-folder of inst/extdata

@@ -84,12 +84,13 @@ lpfunc <- function(x, y, w = NULL, col, size = 1, type, pch = NULL) {
 #' @param x Name of system file to find
 #' @param y Sub-directory under inst/extdata where file is stored (details)
 #' @return Full path to system file
-#' @details For argument y, use "lc" for landcover, "yl" for "yield", and "cl" 
-#' for "climate".
+#' @details For argument y, use "lc" for landcover, "yl" for "yield", "cl" 
+#' for "climate", "ab" for "abm". 
 #' @keywords internal
 #' @export 
 spathfunc <- function(x, y = "lc") {
-  mat <- cbind(c("lc", "yl", "cl"), c("landcover", "yieldprod", "climate"))
+  mat <- cbind(c("lc", "yl", "cl", "ab"), 
+               c("landcover", "yieldprod", "climate", "abm"))
   base <- mat[which(mat[, 1] == y), 2]
   system.file(paste0("extdata/", base), x, package = "croplandbias")
 }

### 2017-08-31 R/utilities.R
Updated spathfunc to include carbon sub-directory. Updated package version accordingly. Prompted by starting to work through c11 (formerly c10), which I moved out of Rmd to vignettes.

@@ -85,12 +85,12 @@ lpfunc <- function(x, y, w = NULL, col, size = 1, type, pch = NULL) {
 #' @param y Sub-directory under inst/extdata where file is stored (details)
 #' @return Full path to system file
 #' @details For argument y, use "lc" for landcover, "yl" for "yield", "cl" 
-#' for "climate", "ab" for "abm". 
+#' for "climate", "ab" for "abm", "cb" for "carbon"
 #' @keywords internal
 #' @export 
 spathfunc <- function(x, y = "lc") {
-  mat <- cbind(c("lc", "yl", "cl", "ab"), 
-               c("landcover", "yieldprod", "climate", "abm"))
+  mat <- cbind(c("lc", "yl", "cl", "ab", "cb"), 
+               c("landcover", "yieldprod", "climate", "abm", "carbon"))
   base <- mat[which(mat[, 1] == y), 2]
   system.file(paste0("extdata/", base), x, package = "croplandbias")
 }

### 2017-08-31 R/utilities.R
Updated spathfunc to point to et sub-directory. Version number updated. Incremental fixes to paths and read-in approaches in c12

@@ -85,12 +85,12 @@ lpfunc <- function(x, y, w = NULL, col, size = 1, type, pch = NULL) {
 #' @param y Sub-directory under inst/extdata where file is stored (details)
 #' @return Full path to system file
 #' @details For argument y, use "lc" for landcover, "yl" for "yield", "cl" 
-#' for "climate", "ab" for "abm", "cb" for "carbon"
+#' for "climate", "ab" for "abm", "cb" for "carbon", "et" for "et"
 #' @keywords internal
 #' @export 
 spathfunc <- function(x, y = "lc") {
-  mat <- cbind(c("lc", "yl", "cl", "ab", "cb"), 
-               c("landcover", "yieldprod", "climate", "abm", "carbon"))
+  mat <- cbind(c("lc", "yl", "cl", "ab", "cb", "et"), 
+               c("landcover", "yieldprod", "climate", "abm", "carbon", "et"))
   base <- mat[which(mat[, 1] == y), 2]
   system.file(paste0("extdata/", base), x, package = "croplandbias")
 }

### 2017-09-01 inst/paper/main.tex
Re-ran c-9; primarily edited main fig for tick conformity and renaming to fig5

--- 2017-08-31 paper/main.tex
@@
 We found that cropland maps can have substantial biases and poor accuracy at all resolutions (e.g.
-at 1 km resolution, up to $\sim$45\% underestimates of cropland (bias) and nearly 50\% mean absolute error (accuracy); at 100 km, up to 15\% underestimates and nearly 20\% MAE).
+at 1 km resolution, up to $\sim$45\% underestimates of cropland (bias) and nearly 50\% mean absolute error  (MAE, describing accuracy); at 100 km, up to 15\% underestimates and nearly 20\% MAE).
 National-scale maps derived from higher resolution imagery were most accurate, followed by multi-map fusion products.
@@
 The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution MODIS land cover \citep{friedl_modis_2010} data (for 2011, the final year of the reference interval), which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
-The fourth dataset was the 1 km GeoWiki cropland map for Africa \citep{fritz_mapping_2015}, which fuses the cropland classes of multiple landcover products (including the three assessed here), with the resulting cropland percentages adjusted until their totals sum to match cropland areas reported in national agricultural statistics (this method also incorporated additional categorical training data collected via crowdsourcing).
+The fourth dataset was the 1 km Global Land Cover Share (GLC-Share)\citep{fritz_mapping_2015}, which fuses the cropland classes of multiple landcover products (including the three assessed here), with the resulting cropland percentages adjusted until their totals sum to match cropland areas reported in national agricultural statistics (this method also incorporated additional categorical training data collected via crowdsourcing).
 This ``hybrid-fusion'' approach represents the state of the art in mapping agricultural land cover.
@@
 The most straightforward way may be to use a Monte Carlo approach that generates artificial datasets \citep[e.g.][]{avelino_goldilocks_2016}, using introduced errors drawn from reported or user-determined accuracy statistics, for both the base land cover and the downstream maps.
-For downstream products based on more complex models, users should examine how particular model assumptions or structures interact with land cover map errors.
+For downstream products based on more complex models, users should examine how land cover map errors interact with particular model assumptions or structures, and alter these where necessary and possible to minimize confounding effects.
 Quantifying error propagation is important to understanding how map error may influence subsequent understanding of the phenomenon of interest.

### 2017-09-04 inst/paper/SI.tex
Big commit of figures, and after revising lc figure again in c12.

--- 2017-08-31 paper/SI.tex
@@
 These intersections and differences were used to calculate the area of true positives, false positives, true negatives, and false negatives within each 1 km$^2$ grid cell, which were then used to calculate accuracy measures.
-In this case, accuracy was assessed using two landcover classes: cropland and non-cropland, thus a two class confusion matrix was constructed (Table 1) according to remote sensing classification accuracy guidelines set out by \citet{olofsson_good_2014}.
+In this case, accuracy was assessed using two land cover classes: cropland and non-cropland, thus a two class confusion matrix was constructed (Table 1) according to remote sensing classification accuracy guidelines set out by \citet{olofsson_good_2014}.
 In this case, the reference data were the visually interpreted crop field presences/absences, and the map data were the field boundary vectors being assessed (i.e.
@@
 \FloatBarrier
+\section{\large Input land cover data}
+The land cover data used in this analysis are illustrated in Figure~\ref{fig:lcfig}, which shows their original form prior to being converted to 1 km cropland percentages over a small region of South Africa.
+
+\begin{figure}[!ht]
+  \centering
+     \includegraphics[width = 17cm]{figures/lc-demo-fig.png} 
+%     \vspace{-3cm}
+      \caption{The original form of the reference dataset (field boundary polygons) and land cover datasets evaluated in this study, illustrated in relation to an underlying high resolution satellite image (from Google Maps, second map from left on top row) selected from a sub-region (red area in top left map) of South Africa.
+The original cover types (excluding a few classes having <10 pixels in the selected location) comprising the SA-LC, GlobCover, and MODIS land cover datasets are shown (GLC-Share is provided as cropland percentages), with the legend at the right providing the corresponding acronym for each class.
+For SA-LC, these classes are natural vegetation (NAT), cultivation (CUL), built (BLT), and water (H20).
+For GlobCover, the classes are mosaic forest or shrubland (50-70\%) / grassland (20-50\%) (FSG), closed to open herbaceous vegetation (COH), sparse vegetation (SVG), mosaic vegetation (50-70\%) / cropland (20-50\%) (MVC), and water (H20).
+For MODIS, the classes are open shrubland (OSH), grassland (GRL), cropland (CLD), and cropland/natural vegetation mosaic (CNV)}
+      \label{fig:lcfig}
+\end{figure}
+
 \section{\large Cropland map error analysis}
 We examined the impact of several sources of uncertainty on our results.
-The first is the potential temporal mismatch between the reference dataset and the landcover products we were testing.
+The first is the potential temporal mismatch between the reference dataset and the land cover products we were testing.
 We tested this by examining two versions of the reference dataset, the initial version created in 2007, and the updated version from 2011 used in the main analysis.
@@
 
+\vspace{10 pt}
 As a further exploration of uncertainty, we included the full range of variability that resulted from a) using either the 2007 or 2011 map to calculate bias and accuracy, and b) from different levels of cropland percentages used when converting MODIS and GlobCover mixed cropland classes into cropland percentage maps, following \citet{fritz_mapping_2015}.
-We pooled the residuals from across each of these permutations, for each reference-test map combination, and examined how their overall mean values change with aggregation scale (Fig.
-1), and also assessed the total distribution of errors within different class of cropland density, determined by dividing the 2011 1 km reference map into 20 different zones, or bins, of cropland density, ranging from 0-5\% cropland cover up to 95-100\% cover.
-We then calculated summary statistics from the pooled residuals and absolute values of residuals within each of these bins (Fig.
-2).
+We pooled the residuals from across each of these permutations, for each reference-test map combination, and examined how their overall mean values change with aggregation scale (Fig.~\ref{fig:bias1-100}), and also assessed the total distribution of errors within different class of cropland density, determined by dividing the 2011 1 km reference map into 20 different zones, or bins, of cropland density, ranging from 0-5\% cropland cover up to 95-100\% cover.
+We then calculated summary statistics from the pooled residuals and absolute values of residuals within each of these bins (Fig.~\ref{fig:biases1km}).
 
@@
 }
-      \label{fig:default}
+      \label{fig:bias1-100}
 \end{figure}
@@
 }
-      \label{fig:default}
+      \label{fig:biases1km}
 \end{figure}
@@
 We also calculated how different methods of calculating the bias and accuracy statistics impacted out findings, using three different methods.
-First, we simply took the straight averages across the entire country, which substantially understates bias and accuracy because cropland covers only 10\% of the country, and all landcover products successfully discriminate the non-cropped regions.
+First, we simply took the straight averages across the entire country, which substantially understates bias and accuracy because cropland covers only 10\% of the country, and all land cover products successfully discriminate the non-cropped regions.
 We also compared the average error metrics extracted from within just the agricultural regions of the maps, defined here as the union of areas having $>$0.5\% cropland in the reference and each test map.
-This also tends to understate bias, because the area having $<$5\% cropland is much larger than areas having higher densities of cropland (see Fig.
-4 frequencies of cells per five percentile bins of cropland density).
+This also tends to understate bias, because the area having $<$5\% cropland is much larger than areas having higher densities of cropland (see Fig.~\ref{fig:cropbins} for frequencies of cells per five percentile bins of cropland density).
 Finally, we calculated a density-independent metric, which produces values similar to the density-weighted means presented in our main analysis, which was calculated by averaging the mean bias and MAE values within different levels of cropland cover (binned into increments of 5\%, excluding areas with $<$0.5\% cropland).
-The results of these three alternate metrics are shown in Figure 3.
+The results of these three alternate metrics are shown in Figure ~\ref{fig:biasregion}.
 
@@
       \caption{A comparison of three alternate methods for calculating bias and accuracy (mean absolute error): as a straight average across the entire country, averaged over agricultural areas only (the union of areas defined as having $>$0.5\% cropland cover in the reference map and each test map), and independent of cropland density, wherein the mean bias/MAE values for each of 20 cropland cover classes (representing 5\% increments of cover 0\% to 100\% defined by the reference map) were calculated and then averaged.}
-      \label{fig:default}
+      \label{fig:biasregion}
 \end{figure}
@@
 Bin values were based on the 2011 reference map, excluding areas with $<$0.5\% cropland.}
-      \label{fig:default}
+      \label{fig:cropbins}
 \end{figure}
 
-We used magisterial district boundaries for South Africa (Fig.
-5) to extract absolute residual values for agricultural areas ($>$0.5\% cropland cover) and 2011 reference cropland density values, the means of which respectively served as the response and predictor in the generalized additive model \citep{wood_mgcv:_2001} analysis (main text).
+We used magisterial district boundaries for South Africa (Fig.~\ref{fig:mdists}) to extract absolute residual values for agricultural areas ($>$0.5\% cropland cover) and 2011 reference cropland density values, the means of which respectively served as the response and predictor in the generalized additive model \citep{wood_mgcv:_2001} analysis (main text).
 
@@
       \caption{South Africa's magisterial districts.}
-      \label{fig:default}
+      \label{fig:mdists}
 \end{figure}
@@
 
-To calculate carbon stocks using the method of \citet{ruesch_new_2008}, we simplified their landcover-specific carbon density classes into six categories, based on the similarity of their class types and the fact that they had the same assigned carbon densities.
+To calculate carbon stocks using the method of \citet{ruesch_new_2008}, we simplified their land cover-specific carbon density classes into six categories, based on the similarity of their class types and the fact that they had the same assigned carbon densities.
 Other classes were dropped because of their low level of, or lack of, occurrence.
@@
 
-Maps of the mean residual differences between the reference map and each of the 5 derived carbon maps for each test map are shown in Figure 6.
+Maps of the mean residual differences between the reference map and each of the 5 derived carbon maps for each test map are shown in Figure~\ref{fig:carbon}.
 Table 3 provides the bias and mean absolute errors for the different carbon maps and how they vary with scale.
@@
       \caption{Spatial patterns of error (averaged across all five different possible cover types adjacent to cropland) in carbon stock estimates. }
-      \label{fig:default}
+      \label{fig:carbon}
 \end{figure}
@@
 
-The differences between the reference map cropland fraction and \emph{tfca} are shown in Figure 7 (where they were adjusted back to percentages), while Table 4 provides the corresponding bias and accuracy values (MAE).
+The differences between the reference map cropland fraction and \emph{tfca} are shown in Figure~\ref{fig:cropadjbias} (where they were adjusted back to percentages), while Table 4 provides the corresponding bias and accuracy values (MAE).
 
@@
 
-The maps of differences between reference-derived yield and production maps and those based on the test maps are shown in Figures 8 and 9, where the values were normalized to the country mean yield or production (calculated from the reference map).
+The maps of differences between reference-derived yield and production maps and those based on the test maps are shown in Figures~\ref{fig:yld} and~\ref{fig:prod}, where the values were normalized to the country mean yield or production (calculated from the reference map).
 Table 5 presents the bias and accuracy values for each reference-test comparison of gridded yield and production estimates, giving both the cropland density-weighted variants of bias and accuracy, and for comparison the same values calculated as straight averages within agricultural pixels (i.e.
@@
       \caption{Errors in cropland maps adjusted using provincial cropland area statistics.}
-      \label{fig:default}
+      \label{fig:cropadjbias}
 \end{figure}
@@
       \caption{Errors (normalized to the reference-derived country mean) in disaggregated maize yield estimates.}
-      \label{fig:default}
+      \label{fig:yld}
 \end{figure}
@@
       \caption{Errors (normalized to reference-derived country mean) production estimates calculated from disaggregated maize yield and harvested area estimates.}
-      \label{fig:default}
+      \label{fig:prod}
 \end{figure}
@@
 %\input{figures/et-bias.tex}
-A number of the variables used by the Variable Infiltration Capacity \citep[VIC;][]{liang_simple_1994} model are linked to a 0.25$^{\circ}$ resolution, AVHRR-derived landcover map.
-These include seasonal leaf area index (LAI) phenologies (Fig.
-10), as well as other properties such as plant rooting depth and infiltration rates.
-Landcover properties therefore impact the model's simulation of water balance.
-To test how landcover map error impacts VIC simulations, we created five new versions of VIC's native landcover scheme, one based on the reference cropland map, and the other four on each of the test maps.
+A number of the variables used by the Variable Infiltration Capacity \citep[VIC;][]{liang_simple_1994} model are linked to a 0.25$^{\circ}$ resolution, AVHRR-derived land cover map.
+These include seasonal leaf area index (LAI) phenologies (Fig.~\ref{fig:lai}), as well as other properties such as plant rooting depth and infiltration rates.
+Land cover and cover properties therefore impact the model's simulation of water balance.
+To test how land cover map error impacts VIC simulations, we created five new versions of VIC's native land cover scheme, one based on the reference cropland map, and the other four on each of the test maps.
 For each version, we first reprojected the relevant 25 km cropland map to the geographic coordinate system used by VIC and resampled it to 0.25$^{\circ}$.
-We then adjusted VIC's landcover scheme by replacing its existing cropland percentages with those from our map, and then proportionally adjusted the remaining cover types in each cell to accommodate the changed cropland amounts.
-Since LAI is strongly linked to rainfall seasonality in South Africa, which varies between winter (May-August) rainfall in the west and southwest of the country and summer (October-March) rainfall in the east and northeast, we assigned LAI curves that peaked during the winter months to the cover types in the western half of the country, and those peaking in the summer months to the eastern half (Fig.
-10).
+We then adjusted VIC's land cover scheme by replacing its existing cropland percentages with those from our map, and then proportionally adjusted the remaining cover types in each cell to accommodate the changed cropland amounts.
+Since LAI is strongly linked to rainfall seasonality in South Africa, which varies between winter (May-August) rainfall in the west and southwest of the country and summer (October-March) rainfall in the east and northeast, we assigned LAI curves that peaked during the winter months to the cover types in the western half of the country, and those peaking in the summer months to the eastern half (Fig.~\ref{fig:lai}).
 
@@
      \includegraphics[width = 10cm]{figures/lai_seasonal_cycles.png} 
-      \caption{Seasonal LAI curves for different cover types in VIC's landcover scheme, showing the phenologies used for the winter-rainfall portion of the country in the west and southwest (left column), and for the summer rainfall region in the country's easter to northeast (right column).
+      \caption{Seasonal LAI curves for different cover types in VIC's land cover scheme, showing the phenologies used for the winter-rainfall portion of the country in the west and southwest (left column), and for the summer rainfall region in the country's easter to northeast (right column).
 }
-      \label{fig:default}
+      \label{fig:lai}
 \end{figure}
 
-After making these adjustments, we conducted five different VIC simulations, one for each of the adjusted landcover schemes.
+After making these adjustments, we conducted five different VIC simulations, one for each of the adjusted land cover schemes.
 The model was run at a daily time step for 28 years, from 1981-2008, from which monthly total evapotranspiration values were extracted.
 We calculated from these results the average annual total ET, the maximum and minimum monthly ET observed throughout the entire time series, and the mean ET during the month which on average had the highest ET during the time series.
-We then found the differences between the reference and test map variants of each of these ET variables, and calculated their bias and MAE values (Fig.
-11).
+We then found the differences between the reference and test map variants of each of these ET variables (Fig.~\ref{fig:et}), and calculated their bias and MAE values.
 The results are similar across the four variables, and we present the difference maps for the mean annual ET comparison in the main text.
@@
 2 of the main text), the maximum (second column, TS max) and minimum (third column, TS min) ET values observed during the 30 year times, and mean ET during the month in which ET peaks (right column, Peak).}
-      \label{fig:default}
+      \label{fig:et}
 \end{figure}
@@
 \subsection{\large District selection}
-For our analysis of how landcover map error impacts agent-based model results, we selected four districts in South Africa which had similar climatic characteristics ($\sim$800 mm annual rainfall) to the region in Zambia where the crop model simulations and household survey data were collected.
-These were four contiguous districts along the western border of Lesotho, Clocolan, Ficksburg, Fouriesburg, and Marquard (Fig.
-12, top left), which had between 29-45\% of their areas covered by cropland, according to the 2011 reference map (Fig.
-12, top right).
+For our analysis of how land cover map error impacts agent-based model results, we selected four districts in South Africa which had similar climatic characteristics ($\sim$800 mm annual rainfall) to the region in Zambia where the crop model simulations and household survey data were collected.
+These were four contiguous districts along the western border of Lesotho, Clocolan, Ficksburg, Fouriesburg, and Marquard (Fig.~\ref{fig:abmdists}, top left), which had between 29-45\% of their areas covered by cropland, according to the 2011 reference map (Fig.~\ref{fig:abmdists}, top right).
 
@@
 }
-      \label{fig:default}
+      \label{fig:abmdists}
 \end{figure}
@@
 \includegraphics[width = 10cm]{figures/histogram.eps} 
-\caption{Histogram of cultivated area in
-the farmer register. } 
+\caption{Histogram of cultivated area in the farmer register. } 
 \label{fig:histo} 

### 2017-09-04 inst/paper/main.tex
Big commit of figures, and after revising lc figure again in c12.

@@
 National-scale maps derived from higher resolution imagery were most accurate, followed by multi-map fusion products.
-Constraining mapped values to match survey statistics proved effective at minimizing bias (provided the statistics are accurate).
+Constraining mapped values to match survey statistics may be effective at minimizing bias (provided the statistics are accurate).
 Errors in downstream analyses could be substantially amplified or muted, depending on the values ascribed to cropland-adjacent covers (e.g.
-with forest as adjacent cover, carbon map error was 200-500\% greater than in the foundational cropland maps, but $\sim$40\% less for sparse cover types).
+with forest as adjacent cover, carbon map error was 200-500\% greater than in input cropland maps, but $\sim$40\% less for sparse cover types).
 The average locational error was 6 km (600\%).
-These findings provide deeper insight into the causes and consequences of land cover map error, including their potential broader regional and policy consequences, and suggest several recommendations for land cover map users.}
+These findings provide deeper insight into the causes and potential consequences of land cover map error, and suggest several recommendations for land cover map users.}
 \end{abstract}
@@
 
-These problems with land cover products are known \citep{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \citep[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2016}.
+These problems with land cover products are known \citep{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \citep[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2016,fritz_mapping_2015}.
 Likewise, the importance of assessing the accuracy of land cover maps is increasingly recognized, and there are well-developed, best-practice guidelines for gathering and using ground-truth samples to robustly quantify map error \citep{foody_status_2002,olofsson_making_2013,olofsson_good_2014,stehman_global_2012}.
@@
 
-We used these vector data as a reference for evaluating four land cover products representative of those commonly used in global change studies and related areas of research.
-The first was South Africa's 30 m resolution 2009 National land cover map \citep[SA-LC][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are created for individual countries \citep[e.g.][]{fry_completion_2009}.
+We used these vector data as a reference for evaluating four land cover products representative of those commonly used in global change studies and related areas of research (see the SI for an illustration of all five datasets in their original form).
+The first was South Africa's 30 m resolution 2009 National land cover map \citep[SA-LC;][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are created for individual countries \citep[e.g.][]{fry_completion_2009}.
 Although global-scale, Landsat-derived maps have recently become available \citep{chen_global_2015}, their reported accuracy for cultivated areas is lower (80-85\% user's accuracy) than those of more intensive, national to sub-national products \citep[e.g.
@@
 The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution MODIS land cover \citep{friedl_modis_2010} data (for 2011, the final year of the reference interval), which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}.
-The fourth dataset was the 1 km Global Land Cover Share (GLC-Share)\citep{fritz_mapping_2015}, which fuses the cropland classes of multiple landcover products (including the three assessed here), with the resulting cropland percentages adjusted until their totals sum to match cropland areas reported in national agricultural statistics (this method also incorporated additional categorical training data collected via crowdsourcing).
-This ``hybrid-fusion'' approach represents the state of the art in mapping agricultural land cover.
-Since GeoWiki provides a continuous value of land cover (cropland percentage), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded cropland percentages.
+The fourth dataset was the 1 km Global Land Cover Share \citep[GLC-Share;][]{latham_global_2014}, which harmonizes and merges the classes of the best available high and medium resolution products for each country/region, providing separate maps for cropland and 10 other thematic classes.
+Among the datasets included in the GLC-Share's cropland map is the GeoWiki cropland percentage map \citep{waldner_unified_2016, fritz_mapping_2015}, which provided an important input to GLC-Share's cropland layer \citep{latham_global_2014}.
+The GeoWiki map was also constructed using fusion techniques to merge multiple cropland datasets (including SA-LC, MODIS, and GlobCover), but further calibrated to match cropland areas reported in national agricultural statistics, and validated by crowdsourced volunteers \citep{fritz_mapping_2015}.
+GLC-Share and the multi-map fusion methodology underlying it (and the GeoWiki map it incorporates) represents a state of the art approach for mapping agricultural land cover.
+Since GLC-Share provides a continuous value of land cover (cropland percentage), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded cropland percentages.
 
@@
 We extracted the cropland classes from SA-LC, MODIS, and GlobCover, and converted these into percent cropland estimates at 1 km resolution.
-Both MODIS and GlobCover had mixed/mosaic classes of cropland and other covers, thus we followed \citet{fritz_mapping_2015} by creating upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages.
+Both MODIS and GlobCover had mixed/mosaic classes of cropland and other covers, thus we followed \citet{fritz_mapping_2015} in creating upper, mean, and lower cropland estimates from these classes to produce three versions of the gridded percentages.
 We used the mean map for the main analysis, but estimated error variability using all three versions (SI).
@@
 We followed this method to create vegetative carbon maps for South Africa.
-Since our cropland percentage map provided no information on surrounding cover, we developed several variants representing potential surrounding cover types by assigning the average carbon densities of five biomes (forest, secondary forest, shrubland, grassland, and sparse vegetation \citep{ruesch_new_2008}) to the non-cropland fraction of our maps.
+Since our cropland percentage map provided no information on surrounding cover, we developed several variants representing potential surrounding cover types by assigning the average carbon densities of five biomes \citep[forest, secondary forest, shrubland, grassland, and sparse vegetation;][]{ruesch_new_2008}) to the non-cropland fraction of our maps.
 We multiplied cropland densities by cropland fractions and added these to each of the five other densities multiplied by the residual non-cropland fractions to create five different carbon density maps.
@@
 
-Although this analysis ascribed hypothetical carbon densities to non-cropland areas, the selected values represent the range in potential carbon stocks in landscapes containing arable agriculture (error metrics from this analysis defined as pixels containing cropland, pixels without cropland do not influence the results), which allowed us to counterfactually investigate how carbon estimates are affected by the interaction of i) test map errors and ii) the properties of neighboring cover types.
+Although this analysis ascribed hypothetical carbon densities to non-cropland areas, the selected values represent the range in potential carbon stocks in landscapes containing arable agriculture (the method for calculating error metrics only affects pixels containing cropland; pixels without cropland do not influence the results), which allowed us to counterfactually investigate how carbon estimates are affected by the interaction of i) test map errors and ii) the properties of neighboring cover types.
 
@@
 \vspace{-0.2 cm}
-Our 1 km reference cropland map indicated that crop fields covered 104,304 km$^2$ (nearly 10\%) of the total study area in the 2009-2011 time period, with corresponding cropland area estimates of 131,390, 82,358, 77,090, and 110,272 (GeoWiki) km$^2$ resulting from the SA-LC, GlobCover, MODIS, and GeoWiki maps, respectively.
+Our 1 km reference cropland map indicated that crop fields covered 104,304 km$^2$ (nearly 10\%) of the total study area in the 2009-2011 time period, with corresponding cropland area estimates of 131,390, 82,358, 77,090, and 110,272 (GeoWiki) km$^2$ resulting from the SA-LC, GlobCover, MODIS, and GLC-Share maps, respectively.
 Cropland area estimates from both the reference and test maps were constant for all levels of aggregation.
@@
 The magnitude of mean absolute error (MAE) was somewhat higher in all cases.
-The GeoWiki map, in contrast, was the least biased overall, showing just a $\sim$7\% bias at 1 km and near 0 for all other scales of aggregation, although its accuracy (23\% MAE) was only half as good as SA-LC's at 1 km (11\% MAE), which despite its uniform overestimation bias (Fig.
+The GLC-Share map, in contrast, was the least biased overall, showing just a $\sim$7\% bias at 1 km and near 0 for all other scales of aggregation, although its accuracy (23\% MAE) was only half as good as SA-LC's at 1 km (11\% MAE), which despite its uniform overestimation bias (Fig.
 1A) was the most accurate map at aggregation scales $<10$km.
-Above this, GeoWiki became slightly more accurate, having $<$5\% MAE at 100 km resolution.
-The reason GeoWiki had relatively poor accuracy at 1 km resolution was due to the high heterogeneous error pattern, which traded between positive and negative residuals over short distances, thereby inflating MAE at this scale.
+Above this, GLC-Share became slightly more accurate, having $<$5\% MAE at 100 km resolution.
+The reason GLC-Share had relatively poor accuracy at 1 km resolution was due to the highly heterogeneous error pattern, which traded between positive and negative residuals over short distances, thereby inflating MAE at this scale.
 
@@
 %\begin{wrapfigure}{c}{1\textwidth}
-\centerline{\includegraphics[width=1\textwidth]{figures/figure1.pdf}}
+\centerline{\includegraphics[width=1\textwidth]{figures/fig1.pdf}}
 \vspace{-0.15 cm}
@@
 %\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/biases_md_lnorm_gam_mu0.pdf}}
-\centerline{\includegraphics[width=.7\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
+\centerline{\includegraphics[width=.7\textwidth]{figures/fig2.pdf}}
 \caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
@@
 However, when forest was adjacent then bias was a three- to five-fold multiple of cropland map bias (Fig.
-1B).
+1b).
 At the most extreme, GlobCover's bias was -276\% at 1 km, but even SA-LC and GeoWiki had biases of 22\% and -50\%, respectively.
@@
 %\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/figure3.pdf}}
-\centerline{\includegraphics[width=0.55\textwidth]{figures/figure3.pdf}}
+\centerline{\includegraphics[width=0.55\textwidth]{figures/fig3.pdf}}
 \caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map.
@@
 \begin{figure}[h]
-\centerline{\includegraphics[width=.6\textwidth]{figures/et_bias_map.pdf}}
+\centerline{\includegraphics[width=.6\textwidth]{figures/fig4.pdf}}
 \caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
@@
 %\vspace{-0.75 cm}
-\centerline{\includegraphics[width=0.55\textwidth]{figures/yield_prod_bias.pdf}}
+\centerline{\includegraphics[width=0.55\textwidth]{figures/fig5.pdf}}
 \caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates.
@@
 \begin{figure}[ht]
-\centerline{\includegraphics[width=.55\textwidth]{figures/figure6.pdf}}
+\centerline{\includegraphics[width=.55\textwidth]{figures/fig6.pdf}}
 \caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and C) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
@@
 \begin{figure}[!ht]
-\centerline{\includegraphics[width=.55\textwidth]{figures/location_offset.pdf}}
+\centerline{\includegraphics[width=.55\textwidth]{figures/fig7.pdf}}
 \caption{Average nearest neighbor distances (in km) between pixels representing features identified by the reference map versus those identified from the test maps.
@@
 The preceding analyses contributes to existing work investigating land cover map error and its consequences \citep[e.g.][]{fritz_highlighting_2011,verburg_challenges_2011,olofsson_making_2013}.
-Previous studies have assessed map errors either by using point-based accuracy assessments \citep[e.g.][]{olofsson_making_2013,frey_how_2007-1,foody_status_2002}, by evaluating between-map discrepancies \citep[e.g.][]{fritz_highlighting_2011,fritz_identifying_2008,fritz_comparison_2010}, or by comparing map-derived estimates to aggregated statistics \citep[e.g.][]{larsen_taken_2015,fritz_comparison_2010,yu_meta-discoveries_2014}.
+Previous studies have assessed map errors either by using point-based accuracy assessments \citep[e.g.][]{olofsson_making_2013,frey_how_2007,foody_status_2002}, by evaluating between-map discrepancies \citep[e.g.][]{fritz_highlighting_2011,fritz_identifying_2008,fritz_comparison_2010}, or by comparing map-derived estimates to aggregated statistics \citep[e.g.][]{larsen_taken_2015,fritz_comparison_2010,yu_meta-discoveries_2014}.
 A smaller number based their assessments on contiguous ground truth maps, but these covered relatively small regions \cite[$<$3000 km$^2$, or $<$0.03\% of the area covered here;][]{dendoncker_exploring_2008,schmit_limitations_2006}.

### 2017-09-04 inst/extR/grid-fields-intersect-analyze.R
Added some data under inst into repo, committing some ignores.

@@ -1,0 +1,289 @@
+##############################################################################################################
+# Title      : SAGridFieldsIntersect_analyze.R
+# Purpose    : Post-processing and analyzing results of intersections between fields and SA grid
+# Author     : Lyndon Estes
+# Draws from : SAGridFieldsIntersect_postgis.R; SAGridFieldsIntersect.R
+# Used by    : 
+# Notes      : 
+##############################################################################################################
+
+# Libraries
+suppressMessages(library(RPostgreSQL))
+suppressMessages(library(rgdal))
+suppressMessages(library(rgeos))
+suppressMessages(library(raster))
+source("/u/sandbox/afmap/R/KMLAccuracyFunctions.R")
+
+# Connection
+drv <- dbDriver("PostgreSQL")
+con <- dbConnect(drv, dbname = "db", user = "user", password = "pw")
+
+# Projection information
+prjsrid <- 97490
+prj.sql <- paste("select proj4text from spatial_ref_sys where srid=", prjsrid, sep = "")
+prjstr <- dbGetQuery(con, prj.sql)$proj4text
+
+# table names
+tab2007 <- "sa_fg_isect_2007"
+tab2011 <- "sa_fg_isect_2011"
+
+# Calculate area of each intersection
+lapply(list(tab2007, tab2011), function(x) {  # add column for area
+  d <- dbSendQuery(con, paste("ALTER TABLE",  x, "ADD COLUMN area numeric"))
+})
+lapply(list(tab2007, tab2011), function(x) {  # calculate area
+  print(paste("processing", x))
+  dbSendQuery(con, paste("UPDATE ", x, "SET area=ST_Area(geom)/10000")) 
+})
+lapply(list(tab2007, tab2011), function(x) {   # clean up tables
+  print(paste("processing", x))
+  dbSendQuery(con, paste("VACUUM ANALYZE",  x))
+})
+
+# vals <- lapply(list(tab2007, tab2011), function(x) {
+#   d <- dbGetQuery(con, paste("SELECT gid FROM", x, "WHERE area > 0"))$gid
+# })
+
+# A small check (all the same)
[…]

### 2017-09-04 inst/extR/grid-fields-intersect-postgis.R

@@ -1,0 +1,140 @@
+##############################################################################################################
+# Title      : SAGridFieldsIntersect_postgis.R
+# Purpose    : Workflow between R, and postgis to clean up polygons and calculate intersections
+# Author     : Lyndon Estes
+# Draws from : SAGridFieldsIntersect_postgis.R; SAGridFieldsIntersect.R
+# Used by    : 
+# Notes      : 
+##############################################################################################################
+
+# Libraries
+suppressMessages(library(RPostgreSQL))
+suppressMessages(library(rgdal))
+suppressMessages(library(rgeos))
+# suppressMessages(library(raster))
+source("/u/sandbox/afmap/R/KMLAccuracyFunctions.R")
+
+test <- "N"
+
+if(test == "N") {
+  args <- commandArgs(TRUE)
+  if(length(args) != 5) stop("Wrong number of arguments")
+  fldname_in <- args[1]  # name of data table to create to hold cleaned fields in
+  fldname_out <- args[2]  # name of data table to create to hold cleaned fields in
+  grname_in <- args[3]  # name of input grid for comparison
+  grname_out <- args[4]  # name of input grid for comparison
+  isectname <- args[5]  # name of output table with intersections
+} else if(test == "Y") {
+  fldname_in <- "sa_flds_2011_alb"
+  fldname_out <- "sa_flds_2011_alb_cl"
+  isectname <- "sa_fg_isect_2011" # "sa_fg_isect_2007"
+  grname_in <- "sa1kilo"  # name of input grid for comparison
+  grname_out <- "sa_isect_grid"  # name of input grid for comparison
+}
+
+# Connection
+drv <- dbDriver("PostgreSQL")
+con <- dbConnect(drv, dbname = "db", user = "user", password = "pw")
+
+prjsrid <- 97490
+prj.sql <- paste("select proj4text from spatial_ref_sys where srid=", prjsrid, sep = "")
+prjstr <- dbGetQuery(con, prj.sql)$proj4text
+
+# First clean up SA fields datatables, starting with duplicate of it
+if(!dbExistsTable(con, name = fldname_out)) {
[…]

### 2017-09-04 inst/paper/main.tex
Updated all vignettes so the rm(list = ls()) appeared above setting variable e (for running chunks all at once), which was causing some vignettes to fail at knit time. Rebuilt package.

@@
 \vspace{-0.2 cm}
-Our 1 km reference cropland map indicated that crop fields covered 104,304 km$^2$ (nearly 10\%) of the total study area in the 2009-2011 time period, with corresponding cropland area estimates of 131,390, 82,358, 77,090, and 110,272 (GeoWiki) km$^2$ resulting from the SA-LC, GlobCover, MODIS, and GLC-Share maps, respectively.
+Our 1 km reference cropland map indicated that crop fields covered 104,304 km$^2$ (nearly 10\%) of the total study area in the 2009-2011 time period, with corresponding cropland area estimates of 131,390, 82,358, 77,090, and 110,272 km$^2$ resulting from the SA-LC, GlobCover, MODIS, and GLC-Share maps, respectively.
 Cropland area estimates from both the reference and test maps were constant for all levels of aggregation.

### 2017-09-04 inst/paper/main.tex
Redid fig. 3 in c7, so that values for carbon bias/mae that extend beyond plot range are added directly to plot, not with post-hoc editing. Rebuilt package. Removed totC_bias.tex from repo (not included in paper)

@@
 districts).
-To meet these computational needs, we selected four contiguous magisterial districts (1,040-1,343 km$^2$) in eastern South Africa with similar climate and 28-45\% of cropland coverage.
+To meet these computational needs, we selected four contiguous magisterial districts (1,037-1,329 km$^2$) in eastern South Africa with similar climate and 28-45\% of cropland coverage.
 To create cropland surfaces for each district, we disaggregated all five cropland maps into 100 m binary cropland/non-cropland cover maps, and then ran the model separately for each district and with each cropland map (20 simulations total).
@@
 Subtracting each test map from the reference maps created pixel-wise residuals, where negative and positive values respectively represent overestimates and underestimates by the test map (Fig.
-1A).
+1a).
 The most pronounced errors were in the MODIS and GlobCover maps, which showed large positive residuals in the center of the country where cropland is most concentrated (blue areas in Fig.
-1A), and negative residuals (red areas) along the eastern and northern margins.
+1a), and negative residuals (red areas) along the eastern and northern margins.
 These patterns translated into substantial map bias (Fig.
-1B), with GlobCover and MODIS mean error exceeding 45\% and 25\% respectively at 1 km resolution, meaning that each map tends to underestimate cropland by that amount at that resolution.
+1b), with GlobCover and MODIS mean error exceeding 45\% and 25\% respectively at 1 km resolution, meaning that each map tends to underestimate cropland by that amount at that resolution.
 This bias declined with each level of map aggregation, being reduced to nearly 15\% for GlobCover and 5\% for MODIS at 100 km.
@@
 The GLC-Share map, in contrast, was the least biased overall, showing just a $\sim$7\% bias at 1 km and near 0 for all other scales of aggregation, although its accuracy (23\% MAE) was only half as good as SA-LC's at 1 km (11\% MAE), which despite its uniform overestimation bias (Fig.
-1A) was the most accurate map at aggregation scales $<10$km.
+1a) was the most accurate map at aggregation scales $<10$km.
 Above this, GLC-Share became slightly more accurate, having $<$5\% MAE at 100 km resolution.
@@
 \vspace{-0.15 cm}
-\caption{(A) Errors in the percent cropland estimates resulting from each of the four test maps relative to the reference map at different scale of pixel aggregation.
+\caption{(a) Errors in the percent cropland estimates resulting from each of the four test maps relative to the reference map at different scale of pixel aggregation.
 Rows indicate the test map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation.
 White indicates areas where areas under communal farmlands or permanent tree crops were removed from analysis.
-(B) The bias (mean error) and accuracy (mean absolute error [MAE]) of each test map at each scale of aggregation, weighted by the percentage cropland in each cell of the reference map.
+(b) The bias (mean error) and accuracy (mean absolute error [MAE]) of each test map at each scale of aggregation, weighted by the percentage cropland in each cell of the reference map.
 Bias estimates are indicated by the semi-transparent bars, accuracy (lower is more accurate) by the solid bars, with bar colors coded to specific cropland maps.}
@@
 %\subsection*{Relationship between map error and cropland density}
-The generalized additive model revealed primarily non-linear relationships between district MAE and cropland density that were best approximated by a first order polynomial function of cropland density (for all four cropland maps: p$<$0.001 on both terms of quadratic and on smoothing function applied to district centroids; $>$ 85\% deviance explained).
+The generalized additive model revealed primarily non-linear relationships between district MAE and cropland density that were best approximated by a first order polynomial function of cropland density (for all four cropland maps: p$<$0.001 on both terms of quadratic and on smoothing function applied to district centroids; $>$85\% deviance explained).
 Map accuracy was typically lowest at intermediate levels of cropland density (50-60\% cover) for all but the GlobCover map (where accuracy continues to decline with cropland cover), and was highest where the landscape is dominated either by cropland or by another type (Fig.
@@
 1b).
-At the most extreme, GlobCover's bias was -276\% at 1 km, but even SA-LC and GeoWiki had biases of 22\% and -50\%, respectively.
+At the most extreme, GlobCover's bias was -276\% at 1 km, but even SA-LC and GLC-Share had biases of 22\% and -50\%, respectively.
 Bias could be substantial even for the least carbon dense vegetation type (sparse), as evidenced by the 15-25\% mean error at 1 km for MODIS and GlobCover under this class.
-The mean bias across the different potential adjacent vegetation classes ranged between -20 for GeoWiki and -123\% for GlobCover at 1 km (with MODIS in between these), while SA-LC's average bias was 11\%.
+The mean bias across the different potential adjacent vegetation classes ranged between -20 for GLC-Share and -123\% for GlobCover at 1 km (with MODIS in between these), while SA-LC's average bias was 11\%.
 Biases declined fairly rapidly with aggregation, with all datasets having an average (across cover types) bias magnitude of $<$10\% at $\geq$25 km of aggregation, except for GlobCover, which was -12\% at 100 km (SI).
-As with cropland percentages, GeoWiki produced the least biased carbon density estimates above 1 km resolution.
+As with cropland percentages, GLC-Share produced the least biased carbon density estimates above 1 km resolution.
 
@@
 
-In terms of accuracy, MAE values were essentially the same as bias magnitudes, except for GeoWiki's, which were twice as large.
+In terms of accuracy, MAE values were essentially the same as bias magnitudes, except for GLC-Share's, which were twice as large.
 The average MAE across vegetation classes was 47\% at 1 km, dropping to $<$10 only with 25 km of aggregation.
-In contrast, SA-LC's carbon estimates were twice as accurate at 1 km, and were slightly more accurate up to 25 km of aggregation were GeoWiki achieved parity.
+In contrast, SA-LC's carbon estimates were twice as accurate at 1 km, and were slightly more accurate up to 25 km of aggregation were GLC-Share achieved parity.
 
@@
 
-SA-LC and GeoWiki also resulted in ET errors estimates along the southern and western coasts, but here the tendency was to underestimate ET, while biases in the center of the country were either negligible to absent.
+SA-LC and GLC-Share also resulted in ET errors estimates along the southern and western coasts, but here the tendency was to underestimate ET, while biases in the center of the country were either negligible to absent.
 All but MODIS underestimated ET by 5-15\% in the northern tip of the country.
@@
 Here all datasets but SA-LC had MAE values of $\geq$30\% below 25 km of aggregation (Fig.
-5), reaching as high as 100\% for GlobCover at 1 km, followed by 65\% for MODIS and 45\% for GeoWiki.
+5), reaching as high as 100\% for GlobCover at 1 km, followed by 65\% for MODIS and 45\% for GLC-Share.
 SA-LC estimated production was most accurate, having between 10-20\% MAE between 1 and 10 km, and $<$10\% at 25 km and higher.
@@
 \vspace{-0.2 cm}
-The average distance between areas containing the highest cropland densities (upper decile) in the reference map and those delineated by the test maps ranged from 1.1 km for SA-LC to 18.2 km for GlobCover, with MODIS (10.1 km) and GeoWiki (2.8 km) having intermediate displacements (Fig.
+The average distance between areas containing the highest cropland densities (upper decile) in the reference map and those delineated by the test maps ranged from 1.1 km for SA-LC to 18.2 km for GlobCover, with MODIS (10.1 km) and GLC-Share (2.8 km) having intermediate displacements (Fig.
 7).
-Locational errors in maps indicating the highest yielding areas showed a similar pattern, with a range of 0.8-14.2 km (SA-LC and GlobCover) and intermediate errors of 5.8-7.5 km (GeoWiki and MODIS).
-For areas of highest carbon density, locations identified by the MODIS-derived map were most distant from those shown by the reference map (11.3 km), followed by GeoWiki (7.4 km), GlobCover (6.8 km), and SA-LC (3.7 km).
+Locational errors in maps indicating the highest yielding areas showed a similar pattern, with a range of 0.8-14.2 km (SA-LC and GlobCover) and intermediate errors of 5.8-7.5 km (GLC-Share and MODIS).
+For areas of highest carbon density, locations identified by the MODIS-derived map were most distant from those shown by the reference map (11.3 km), followed by GLC-Share (7.4 km), GlobCover (6.8 km), and SA-LC (3.7 km).
 
@@
 \vspace{-0.2 cm}
-Our findings showed that the most accurate cropland estimates, across all spatial scales, were produced by SA-LC followed by GeoWiki.
-In the reverse order, these two datasets were also the least biased, with GeoWiki having effectively zero bias at aggregation levels of 5 km or coarser.
+Our findings showed that the most accurate cropland estimates, across all spatial scales, were produced by SA-LC followed by GLC-Share.
+In the reverse order, these two datasets were also the least biased, with GLC-Share having effectively zero bias at aggregation levels of 5 km or coarser.
 GlobCover was highly inaccurate and biased at all scales, with MODIS estimates being nearly as inaccurate but substantially less biased.
@@
 2) further demonstrate how pixel mixing and class definition influence error.
-For MODIS, SA-LC, and GeoWiki, error was highest where pixels were evenly divided between cropland and other cover types, reflecting earlier work showing that classification accuracy is lowest when cover types are most mixed \citep{verburg_challenges_2011,gross_monitoring_2013}.
+For MODIS, SA-LC, and GLC-Share, error was highest where pixels were evenly divided between cropland and other cover types, reflecting earlier work showing that classification accuracy is lowest when cover types are most mixed \citep{verburg_challenges_2011,gross_monitoring_2013}.
 On the other hand, GlobCover was least accurate over pixels with 100\% cropland, an error pattern imposed by the proportions defining cropland within GlobCover's dominant mosaic classes; these set an upper bound that necessarily led to underestimation error over dense croplands.
@@
 
-GeoWiki's cropland errors demonstrate three other ways in which methodology affects error pattern.
+GLC-Share's cropland errors demonstrate three other ways in which methodology affects error pattern.
 The first is that constraining remotely sensed cropland proportions to match census-statistics \citep{fritz_mapping_2015,fritz_highlighting_2011} can be an effective way to reduce bias.
-The second is that this constraint cannot correct the errors of commission and omission within the individual landcover datasets that were merged to create the GeoWiki map \citep{fritz_mapping_2015}, which is why its accuracy is relatively low at 1 km resolution (Fig.
+The second is that this constraint cannot correct the errors of commission and omission within the individual landcover datasets that were merged to create the GLC-Share map \citep{fritz_mapping_2015}, which is why its accuracy is relatively low at 1 km resolution (Fig.
 1B).
-However, GeoWiki's accuracy was substantially higher than MODIS and GlobCover's (Fig.
-1B, 3, \& 5), which reveals the third point, namely that the landcover fusion process used to create GeoWiki does help minimize such error.
+However, GLC-Share's accuracy was substantially higher than MODIS and GlobCover's (Fig.
+1B, 3, \& 5), which reveals the third point, namely that the landcover fusion process used to create GLC-Share does help minimize such error.
 Fusion may be particularly effective for minimizing underestimation errors caused by mixed classes (as with GlobCover) in areas of substantial sub-pixel heterogeneity \citep{fritz_mapping_2015,tuanmu_global_2014}.
@@
 The disaggregated yield and crop production maps we created also showed both error amplification and muting, which in this case depended on the particular analysis.
-Errors within the yields maps were uniformly lower (50-90\% less at 1 km) than those in the input cropland maps, whereas errors in the production maps were 70-100\% higher at 1 km, and actually were exacerbated at intermediate levels of aggregation (10-25 km) for GlobCover, MODIS, and GeoWiki (170-290\%).
+Errors within the yields maps were uniformly lower (50-90\% less at 1 km) than those in the input cropland maps, whereas errors in the production maps were 70-100\% higher at 1 km, and actually were exacerbated at intermediate levels of aggregation (10-25 km) for GlobCover, MODIS, and GLC-Share (170-290\%).
 This latter tendency was mostly caused by accuracy in the production maps improving more slowly with aggregation (Fig.
@@
 7).
-The relative size of offsets tends to follow the patterns of accuracy seen in other assessments, with SA-LC producing the most spatially accurate results, and GlobCover the least, with the exception that GlobCover's 90\textsuperscript{th} percentile carbon density locations are closer to those in the reference map than those of MODIS or GeoWiki.
+The relative size of offsets tends to follow the patterns of accuracy seen in other assessments, with SA-LC producing the most spatially accurate results, and GlobCover the least, with the exception that GlobCover's 90\textsuperscript{th} percentile carbon density locations are closer to those in the reference map than those of MODIS or GLC-Share.
 Numerically, none of these nearest neighbor differences seem large, but they are akin to the root mean square error term used when measuring geometric distortion in a satellite image.
@@
 
-If high quality, high-resolution maps are not available, users interested in agricultural cover could select a new fusion map, such as GeoWiki and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}.
+If high quality, high-resolution maps are not available, users interested in agricultural cover could select a new fusion map, such as GLC-Share and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}.
 Although these maps are relatively coarse-scaled, the fusion methodology helps to greatly minimize error, as does the use of fractional cover values.

### 2017-09-04 inst/paper/main.tex
Checking manuscript through fig2 results. Redid c6 to double-check gam fit.

@@
 \vspace{-0.2 cm}
-The bias and accuracy metrics reflect the degree to quantitative estimates are influenced by land cover map errors.
+The bias and accuracy metrics reflect the degree to which quantitative estimates are influenced by land cover map errors.
 However, land cover data may also be used to identify specific locations \citep[e.g.
@@
 These patterns translated into substantial map bias (Fig.
-1b), with GlobCover and MODIS mean error exceeding 45\% and 25\% respectively at 1 km resolution, meaning that each map tends to underestimate cropland by that amount at that resolution.
+1b), with GlobCover and MODIS mean bias exceeding 45\% and 25\% respectively at 1 km resolution, meaning that each map tends to underestimate cropland by that amount at that resolution.
 This bias declined with each level of map aggregation, being reduced to nearly 15\% for GlobCover and 5\% for MODIS at 100 km.
@@
 The generalized additive model revealed primarily non-linear relationships between district MAE and cropland density that were best approximated by a first order polynomial function of cropland density (for all four cropland maps: p$<$0.001 on both terms of quadratic and on smoothing function applied to district centroids; $>$85\% deviance explained).
-Map accuracy was typically lowest at intermediate levels of cropland density (50-60\% cover) for all but the GlobCover map (where accuracy continues to decline with cropland cover), and was highest where the landscape is dominated either by cropland or by another type (Fig.
+Map accuracy was typically lowest at intermediate levels of cropland density (50-60\% cover) for all but the GlobCover map (where accuracy continues to decline with cropland cover), and was highest where the landscape was dominated either by cropland or by another type (Fig.
 2).

### 2017-09-04 inst/paper/main.tex
Re-ran c7 as check on data.table results (c9 also). Rebuilt package

@@
 1b).
-At the most extreme, GlobCover's bias was -276\% at 1 km, but even SA-LC and GLC-Share had biases of 22\% and -50\%, respectively.
+At the most extreme, GlobCover's bias was -276\% at 1 km, but even SA-LC and GLC-Share had biases of 22\% and -46\%, respectively.
 Bias could be substantial even for the least carbon dense vegetation type (sparse), as evidenced by the 15-25\% mean error at 1 km for MODIS and GlobCover under this class.
@@
 
-In terms of accuracy, MAE values were essentially the same as bias magnitudes, except for GLC-Share's, which were twice as large.
-The average MAE across vegetation classes was 47\% at 1 km, dropping to $<$10 only with 25 km of aggregation.
-In contrast, SA-LC's carbon estimates were twice as accurate at 1 km, and were slightly more accurate up to 25 km of aggregation were GLC-Share achieved parity.
+In terms of accuracy, MAE values were essentially the same as bias magnitudes, except for GLC-Share's, which were twice as large; GLC-Share's average MAE across vegetation classes was 47\% at 1 km, dropping to $<$10 only with 25 km of aggregation.
+In contrast, SA-LC's carbon estimates were twice as accurate at 1 km, and were slightly more accurate up to 25 km of aggregation, where GLC-Share achieved parity.
 
@@
 This low accuracy relative to the gridded yield measures relates to the disaggregation process for harvested area, which allocates a fractional value to each pixel, which is itself a fraction.
-The process of adjusting the gridded values so that their totals match reported statistics does relatively little to correct the map's underlying commission or omission errors, and this constraint in fact appears to shorten the distance between negative and positive residuals (SI), thereby increasing absolute errors.
+The process of adjusting the gridded values so that their totals match reported statistics does relatively little to correct the map's underlying commission or omission errors, and this constraint in fact appears to shorten the spatial distance between negative and positive residuals (SI), thereby increasing absolute errors.
 
@@
 In terms of impact to agent-based model simulation, where cropland map errors were negative (indicating a cropland overestimate by the test maps), the percent of land left unallocated had a straight one-to-one relationship with the percentage of overestimation (Fig.
-6A).
+6a).
 When cropland was underestimated, all croplands were allocated up until the underestimation exceeded 50\%.
@@
 Land deficit (the total area of cropland that should have been allocated to households in each district, but wasn't) increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2 (Fig.
-6B)--and would become infinite in the case of a 100\% underestimate.
+6b)--and would become infinite in the case of a 100\% underestimate.
 This contrasted with food deficit (the percentage shortfall in the average amount of food production that should have been produced by each household but wasn't), which increased linearly with the percentage of cropland underestimate (Fig.
-6C).
-
-%randomly siting 100 household agents within each district, and allocating the nearest two cropland pixels to each household.
-The remaining agents are then iteratively assigned unallocated cropland pixels within a 1.5 km radius of existing agents' fields, and this process continues until all agents are assigned cropland, or all available cropland is allocated.
-This initialization process
-%\FloatBarrier
+6c).
 
@@
 \centerline{\includegraphics[width=.55\textwidth]{figures/fig6.pdf}}
-\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and C) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
+\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of a) the percent of cropland in each district that was not allocated to any household, b) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and c) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
 Dot sizes correspond to district numbers, colors represent the land cover map.}
@@
 For areas of highest carbon density, locations identified by the MODIS-derived map were most distant from those shown by the reference map (11.3 km), followed by GLC-Share (7.4 km), GlobCover (6.8 km), and SA-LC (3.7 km).
-
-%The analyses above primarily dealt with impact of landcover map errors on estimating quantitative values from maps.
-However, there are many applications in which land cover data are needed to identify specific locations, such as areas to be targeted for development, where quantitative information is used to target specific locations \citep{estes_reconciling_2016}.
 
@@
 
-GLC-Share's cropland errors demonstrate three other ways in which methodology affects error pattern.
-The first is that constraining remotely sensed cropland proportions to match census-statistics \citep{fritz_mapping_2015,fritz_highlighting_2011} can be an effective way to reduce bias.
-The second is that this constraint cannot correct the errors of commission and omission within the individual landcover datasets that were merged to create the GLC-Share map \citep{fritz_mapping_2015}, which is why its accuracy is relatively low at 1 km resolution (Fig.
-1B).
+GLC-Share's cropland errors demonstrate three other possible ways in which methodology affects error pattern.
+The first is that constraining remotely sensed cropland proportions to match census-statistics may be effective in reducing bias (this point is attributable to the GeoWiki product \citep{fritz_mapping_2015,fritz_highlighting_2011}, a statistically constrained map that is a major component of GLC-Share, although it is unclear how much GeoWiki's results dominate those of other maps in the GLC-Share fusion process).
+The second is that such a constraint cannot correct the errors of commission and omission within the individual landcover datasets that were merged to create the GLC-Share map \citep{fritz_mapping_2015}, which is why its accuracy is relatively low at 1 km resolution (Fig.
+1c).
 However, GLC-Share's accuracy was substantially higher than MODIS and GlobCover's (Fig.
-1B, 3, \& 5), which reveals the third point, namely that the landcover fusion process used to create GLC-Share does help minimize such error.
+1b, 3, \& 5), which reveals the third point, namely that the landcover fusion process used to create GLC-Share does help minimize such error.
 Fusion may be particularly effective for minimizing underestimation errors caused by mixed classes (as with GlobCover) in areas of substantial sub-pixel heterogeneity \citep{fritz_mapping_2015,tuanmu_global_2014}.
@@
 Of course, high quality, fine-scaled maps such as SA-LC, which was carefully developed for South Africa, do not exist for many countries.
-Development of a new generation of Landsat/Sentinel-based (i.e., 30m resolution or finer) land-cover maps is underway, as exemplified by the new 30 m GLOBELAND30 map.
+Development of a new generation of Landsat/Sentinel-based (i.e., 30m resolution or finer) land-cover maps is underway, as exemplified by the new 30 m GLOBELAND30 map \citep{chen_global_2015}.
 These maps will likely prove very useful for countries and regions lacking their own focused maps, as well as for cross-border analyses.
@@
 
-If high quality, high-resolution maps are not available, users interested in agricultural cover could select a new fusion map, such as GLC-Share and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}.
+If high quality, high-resolution maps are not available, users interested in agricultural cover could select a new fusion map, such as GLC-Share, GeoWiki, or its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}.
 Although these maps are relatively coarse-scaled, the fusion methodology helps to greatly minimize error, as does the use of fractional cover values.
-By calibrating against inventory data, these maps can also reduce pixel level biases (e.g.
+Calibrating against inventory data may also reduce pixel level biases (e.g.
 Fig.
-1B).
+1b).
 However, users should be aware that this adjustment technique introduces confounding errors where census-reported statistics are unreliable, such as in some SSA countries \citep{carletto_emperor_2013,carletto_guesstimates_2015}.

### 2017-09-05 inst/paper/main.tex
Redid c9 to double-check some SI values (0 production bias in agricultural area version). It was fine--confirmed with checks 1051-1071. Updated package.

@@
 
+\section*{Supporting information}
+Additional information on supplemental methods and results can be found in the Supplementary Methods and Results document.
+Additionally, the paper, code, and all non-proprietary data are available as an installable R package at https://github.com/agroimpacts/croplandbias.
+
 \bibliographystyle{gcbnourl} 

### 2017-09-05 inst/paper/main_submission.tex
Final updates before resubmission. New version of latex source, with figures stripped out (figures converted manually in adobe to EPS for submission, not versioned here).  Did various small retests of files and figures, including last minute check on correctness of fig6 in on c10 plots, using code below to make sure the right datasets were associated the right results (they were)
tnm <- "modmu"
plot(abias[var == "ag_area", (gti - get(tnm)) / gti * 100],
     round(abias[var == "land_def", get(tnm)] /
             abias[var == "ag_area", get(tnm)] * 100, 2),
     xlim = c(-30, 100), ylim = c(0,

--- 2017-09-05 inst/paper/main.tex
@@
 \date{}
-%\contributor{Submitted to Proceedings of the National Academy of Sciences
-%of the United States of America}
-
-%%%Newly updated.
-%%% If significance statement need, then can use the below command otherwise just delete it.
-%\significancetext{Remote sensing-derived land cover maps underpin much research into global environmental change, particularly in the most rapidly developing regions.
-These maps often contain substantial error, yet inadequate validation data makes it impossible to fully quantify map errors and how they affect our understanding of global change processes.
-Using a unique, high quality reference map of cropland distribution, this study comprehensively assesses land cover map bias and accuracy in a major agricultural region of sub-Saharan Africa, finding that map errors are strongly correlated with the land cover density, and can propagate substantial errors in downstream global change analyses.
-Users of land cover maps should select later generation products and aggregate to appropriate resolutions to avoid misleading insights and policies related to global change.
-}
 
@@
 
-%\keywords{land cover | bias | remote sensing | agriculture | crop yield | harvested area | carbon | agent-based model | landscape}
-
-%\abbreviations{GTI, GeoTerraImage; SSA, sub-Saharan Africa}
-\linenumbers
+%\linenumbers
 
@@
 Since GLC-Share provides a continuous value of land cover (cropland percentage), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded cropland percentages.
-
-%is a ``fusion'' product, in that it merges the results of the first three datasets (and several others), using i) the extent of the agreement between each landcover product and ii) each product's reliability, as determined by crowdsourced validation data, to assign a likelihood of cropland presence, and then calibrating that likelihood into a cropland percentage through calibration against national agricultural statistics.
-and represents the current state-of-the-art in land cover mapping.
 
@@
 
-%\indent To create the gridded maize yield and production maps, we followed Ramankutty et al.
-\citep{ramankutty_farming_2008} to first calibrate the test cropland maps against cropland areas reported by administrative-level agricultural censuses.
-
 \vspace{-0.3 cm}
@@
 
-\begin{figure*}[!ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\centerline{\includegraphics[width=1\textwidth]{figures/fig1.pdf}}
-\vspace{-0.15 cm}
-\caption{(a) Errors in the percent cropland estimates resulting from each of the four test maps relative to the reference map at different scale of pixel aggregation.
+%\begin{figure*}[!ht]
+%\centerline{\includegraphics[width=1\textwidth]{figures/fig1.pdf}}
+%\vspace{-0.15 cm}
+%\caption{(a) Errors in the percent cropland estimates resulting from each of the four test maps relative to the reference map at different scale of pixel aggregation.
 Rows indicate the test map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation.
@@
 %\begin{wrapfigure}{r}{0.5\textwidth}
-\begin{figure}[!h]
-%\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/biases_md_lnorm_gam_mu0.pdf}}
-\centerline{\includegraphics[width=.7\textwidth]{figures/fig2.pdf}}
-\caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=.7\textwidth]{figures/fig2.pdf}}
+%\caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
 Prediction curves are color-coded to the different test maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}\label{afoto2}
-\end{figure}
+%\end{figure}
 %\end{wrapfigure}
@@
 
-%\vspace{-0.75 cm}
-\begin{figure}[!h]
-%\begin{wrapfigure}{r}{0.5\textwidth}
-%\centerline{\includegraphics[keepaspectratio,scale = 0.4]{figures/figure3.pdf}}
-\centerline{\includegraphics[width=0.55\textwidth]{figures/fig3.pdf}}
-\caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map.
+%\begin{figure}[!h]
+%\centerline{\includegraphics[width=0.55\textwidth]{figures/fig3.pdf}}
+%\caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map.
 Bias estimates (represented by symbols) fall within the semi-transparent floating bars, while accuracies are contained in the solid bars.
@@
 MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.}
-\label{afoto}
-\end{figure}
-%\end{wrapfigure}
+%\label{afoto}
+%\end{figure}
 
@@
 All but MODIS underestimated ET by 5-15\% in the northern tip of the country.
-%
-%\vspace{-0.75 cm}
-\begin{figure}[h]
-\centerline{\includegraphics[width=.6\textwidth]{figures/fig4.pdf}}
-\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
-\end{figure}
+
+%begin{figure}[h]
+%\centerline{\includegraphics[width=.6\textwidth]{figures/fig4.pdf}}
+%\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
+%\end{figure}
 
@@
 
-%\vspace{-0.5 cm}
-\begin{figure}[!ht]
-%\vspace{-0.75 cm}
-\centerline{\includegraphics[width=0.55\textwidth]{figures/fig5.pdf}}
-\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates.
+%\begin{figure}[!ht]
+%\centerline{\includegraphics[width=0.55\textwidth]{figures/fig5.pdf}}
+%\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates.
 Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.
 Symbols code the different variables (production and yield), normalized to their respective means.}
-\label{afoto}
-\end{figure}
+%\label{afoto}
+%\end{figure}
 
@@
 
-\begin{figure}[ht]
-\centerline{\includegraphics[width=.55\textwidth]{figures/fig6.pdf}}
-\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of a) the percent of cropland in each district that was not allocated to any household, b) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and c) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
+%\begin{figure}[ht]
+%\centerline{\includegraphics[width=.55\textwidth]{figures/fig6.pdf}}
+%\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of a) the percent of cropland in each district that was not allocated to any household, b) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and c) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
 Dot sizes correspond to district numbers, colors represent the land cover map.}
-\label{afoto}
-\end{figure}
+%\label{afoto}
+%\end{figure}
 
@@
 
-\begin{figure}[!ht]
-\centerline{\includegraphics[width=.55\textwidth]{figures/fig7.pdf}}
-\caption{Average nearest neighbor distances (in km) between pixels representing features identified by the reference map versus those identified from the test maps.
+%\begin{figure}[!ht]
+%\centerline{\includegraphics[width=.55\textwidth]{figures/fig7.pdf}}
+%\caption{Average nearest neighbor distances (in km) between pixels representing features identified by the reference map versus those identified from the test maps.
 Bar colors indicate the different features (and thus contributing maps), which were delineated by selecting pixels with values greater than the 90th percentile: densest cropland (solid bars); highest maize yield (medium transparent bars); highest carbon density (most transparent bars).}
-\label{afoto}
-\end{figure}
+%\label{afoto}
+%\end{figure}
 
@@
 Additional information on supplemental methods and results can be found in the Supplementary Methods and Results document.
-Additionally, the paper, code, and all non-proprietary data are available as an installable R package at https://github.com/agroimpacts/croplandbias.
+The paper manuscript source, code, and all non-proprietary data are available as part of an R package at https://github.com/agroimpacts/croplandbias.
 
@@
 
+\section*{Captions}
+\noindent
+Fig.
+1.
+Errors in the percent cropland estimates resulting from each of the four test maps relative to the reference map at different scale of pixel aggregation.
+Rows indicate the test map being assessed (by subtraction from the reference map), while columns refer to resolution of aggregation.
+White indicates areas where areas under communal farmlands or permanent tree crops were removed from analysis.
+(b) The bias (mean error) and accuracy (mean absolute error [MAE]) of each test map at each scale of aggregation, weighted by the percentage cropland in each cell of the reference map.
+Bias estimates are indicated by the semi-transparent bars, accuracy (lower is more accurate) by the solid bars, with bar colors coded to specific cropland maps.
+
+\vspace{10pt}
+\noindent Fig.
+2.
+The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model.
+Prediction curves are color-coded to the different test maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.
+
+\vspace{10pt}
+\noindent Fig.
+3.
+Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map.
+Bias estimates (represented by symbols) fall within the semi-transparent floating bars, while accuracies are contained in the solid bars.
+Bar colors are coded to specific cropland map, symbols indicate which cover type was used to calculate cropland-adjacent carbon density.
+The bar represents the mean biases calculated across each of the 5 cover types.
+Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2).
+MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.
+
+\vspace{10pt}
+\noindent Fig.
+4.
+Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.
+
+\vspace{10pt}
+\noindent Fig.
+5.
+Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates.
+Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.
+Symbols code the different variables (production and yield), normalized to their respective means.
+
+\vspace{10pt}
+\noindent Fig.
+6.
+Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of a) the percent of cropland in each district that was not allocated to any household, b) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and c) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation.
+Dot sizes correspond to district numbers, colors represent the land cover map.
+
+\vspace{10pt}
+\noindent Fig.
+7.
+Average nearest neighbor distances (in km) between pixels representing features identified by the reference map versus those identified from the test maps.
+Bar colors indicate the different features (and thus contributing maps), which were delineated by selecting pixels with values greater than the 90th percentile: densest cropland (solid bars); highest maize yield (medium transparent bars); highest carbon density (most transparent bars).
+
 \end{document}

### 2017-09-13 inst/paper/main.tex
Fixed minor wording for type-setting in Fig. 3 caption

@@
 The bar represents the mean biases calculated across each of the 5 cover types.
-Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2).
+Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (see SI for all values).
 MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.}
@@
 Additional information on supplemental methods and results can be found in the Supplementary Methods and Results document.
-Additionally, the paper, code, and all non-proprietary data are available as an installable R package at https://github.com/agroimpacts/croplandbias.
+The paper manuscript, code, and all non-proprietary data are available as part of an R package at https://github.com/agroimpacts/croplandbias.
 

### 2017-09-13 inst/paper/main_submission.tex
Fixed minor wording for type-setting in Fig. 3 caption

@@
 The bar represents the mean biases calculated across each of the 5 cover types.
-Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2).
+Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (see SI for all values).
 MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.
